---
title: "Calculate Iron and Zinc Bioavailability"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script calculates micronutrient bioavailability for children 6-59 months in SIVESNU 2018, applying:

1.  Load SIVESNU data with integrated Chispitas supplementation
2.  Calculate total nutrient intake by aggregating maize and non-maize sources
3.  Apply Miller equation to estimate Total Absorbed Zinc (TAZ) from combined diet + Chispitas
4.  Apply uniform 5% iron absorption factor based on predominantly plant-based dietary patterns
5.  Generate diagnostic tables and visualizations
6.  Save processed dataset maintaining three-tier variable structure

**Key methodological decisions**:

-   **Miller equation version**: Estimate total absorbed zinc among children 6-59 months using equation developed by Miller et al. (2015).

-   **Miller equation applicability**: The equation was derived from studies including micronutrient powder trials, making it appropriate for estimating absorption from diet + Chispitas combined.

# Setup

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  survey,             # Complex survey methodology
  srvyr,              # Tidy wrapper for survey package
  gt,                 # Table formatting and presentation
  tidyverse           # General purpose data manipulation
)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter
drop   <- base::drop

# Set display options
options(scipen = 999, digits = 4)

source(here("02_code", "00_functions", "00_functions.R"))

# transfer/
ensure_output_dir(here("01_data", "02_processed", "transfer"))

# web_resources/
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

# Phase A: Data Loading and Parameters

## A1. Load SIVESNU Data with Chispitas

Load SIVESNU 2018 children's data with integrated Chispitas supplementation from Script 04_02.

```{r}
#| label: a1-load-data

# Load SIVESNU data with Chispitas integration
df_input <- read_parquet(

here("01_data", "02_processed", 
     "transfer", "04_02_sivesnu_with_chispitas.parquet")
)
```

## A2. Define Miller Equation Parameters

The Miller equation for children estimates Total Absorbed Zinc (TAZ) as a function of age and total dietary zinc (TDZ). 

**Equation**:

$$TAZ = 0.5 \times \left( K_T + A \cdot Age^E + TDZ - \sqrt{(K_T + A \cdot Age^E + TDZ)^2 - 4 \cdot (A \cdot Age^E) \cdot TDZ} \right)$$

Where:

-   $TAZ$ = Total Absorbed Zinc (mg/day)
-   $TDZ$ = Total Daily Zinc from all sources (mg/day)
-   $Age$ = Age in months
-   $K_T$ = 2.42 (saturation constant)
-   $A$ = 0.633 (age coefficient)
-   $E$ = 0.311 (age exponent)

```{r}
#| label: a2-define-miller-parameters

# Miller equation parameters for children (6-59 months)
# Source: Engle-Stone R, UC Davis (2025, personal communication)
MILLER_K_T <- 2.42
MILLER_A   <- 0.633
MILLER_E   <- 0.311
```

## A3. Analyze Dietary Composition for Iron Bioavailability Decision

Based on the distribution of the proportion of total food consumption from plant-based sources in this population, summarised in the figure below, we apply a uniform 5% iron absorption rate.

```{r}
#| label: fig-a3-plant-percentage-distribution
#| fig-cap: "Distribution of plant-based food percentage in children's diets (SIVESNU 2018)"
#| fig-width: 9
#| fig-height: 6

# =============================================================================
# A3. Histogram: Plant-Based Food Percentage Distribution
# =============================================================================

# Define Azahar color palette
color_azul_primary <- "#1f78b4"

# Filter valid data for plotting (removes NA and non-finite values)
df_plot <- df_input |>
  filter(!is.na(per_vegetales) & is.finite(per_vegetales))

# Calculate survey-weighted statistics for reference lines
svy_temp <- df_plot |>
  as_survey_design(ids = conglomerado, weights = pesonino)

weighted_median <- svyquantile(~per_vegetales, design = svy_temp, 
                                quantiles = 0.50, na.rm = TRUE) |>
  coef() * 100

weighted_terciles <- svyquantile(~per_vegetales, design = svy_temp, 
                                  quantiles = c(1/3, 2/3), na.rm = TRUE) |>
  coef() * 100

export(df_plot, here("01_data", "02_processed", "web_resources", "04_03_fig_a3_plant_percentage_distribution.rds"))

# Create histogram
ggplot(df_plot, aes(x = per_vegetales * 100)) +
  geom_histogram(
    binwidth = 2.5,
    fill = color_azul_primary,
    color = "white",
    alpha = 0.8
  ) +
  geom_vline(
    xintercept = weighted_terciles[1], 
    linetype = "dashed", 
    color = "#3AAF85", 
    linewidth = 0.8
  ) +
  geom_vline(
    xintercept = weighted_terciles[2], 
    linetype = "dashed", 
    color = "#3AAF85", 
    linewidth = 0.8
  ) +
  geom_vline(
    xintercept = weighted_median, 
    linetype = "solid", 
    color = "#1a3a5c", 
    linewidth = 1
  ) +
  annotate(
    "text", 
    x = weighted_terciles[1] - 2, 
    y = Inf, 
    label = paste0("T1: ", round(weighted_terciles[1], 1), "%"),
    hjust = 1, 
    vjust = 1.5, 
    size = 3, 
    color = "#236C51"
  ) +
  annotate(
    "text", 
    x = weighted_terciles[2] + 2, 
    y = Inf, 
    label = paste0("T2: ", round(weighted_terciles[2], 1), "%"),
    hjust = 0, 
    vjust = 1.5, 
    size = 3, 
    color = "#236C51"
  ) +
  labs(
    title = "Distribution of Plant-Based Food Consumption",
    subtitle = "Children 6-59 months, SIVESNU 2018",
    x = "Percentage of Plant-Based Food Consumption",
    y = "Frequency",
    caption = "Vertical lines: median (solid), tercile cutoffs (dashed). Source: SIVESNU 2018."
  ) +
  scale_x_continuous(
    breaks = seq(0, 100, 10),
    labels = function(x) paste0(x, "%")
  ) +
  coord_cartesian(xlim = c(0, 100)) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 13, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    plot.caption = element_text(size = 9, color = "grey50", hjust = 0),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4")
  )

# Cleanup temporary objects
rm(svy_temp, weighted_median, weighted_terciles, df_plot, color_azul_primary)
```

```{r}
#| label: tbl-a3-plant-percentage-summary
#| tbl-cap: "Survey-weighted distribution of plant-based food percentage (SIVESNU 2018)"

# =============================================================================
# A3. Summary Statistics Table
# Survey-weighted estimates: Mean, SD, Median, IQR, Terciles
# =============================================================================

# Create survey design object
svy_plant <- df_input |>
  filter(!is.na(per_vegetales) & is.finite(per_vegetales)) |>
  as_survey_design(ids = conglomerado, weights = pesonino)

# Calculate survey-weighted statistics
weighted_mean <- svymean(~per_vegetales, design = svy_plant, na.rm = TRUE)
mean_val <- coef(weighted_mean) * 100

# Unweighted SD
unwt_sd <- sd(df_input$per_vegetales, na.rm = TRUE) * 100

# Quantiles: Q25, Median, Q75
weighted_quantiles <- svyquantile(
  ~per_vegetales, 
  design = svy_plant, 
  quantiles = c(0.25, 0.50, 0.75), 
  na.rm = TRUE
)
q25_val <- coef(weighted_quantiles)[1] * 100
median_val <- coef(weighted_quantiles)[2] * 100
q75_val <- coef(weighted_quantiles)[3] * 100

# Terciles
weighted_terciles <- svyquantile(
  ~per_vegetales, 
  design = svy_plant, 
  quantiles = c(1/3, 2/3), 
  na.rm = TRUE
)
t1_val <- coef(weighted_terciles)[1] * 100
t2_val <- coef(weighted_terciles)[2] * 100

# Sample size
n_sample <- nrow(df_input |> filter(!is.na(per_vegetales)))

# Build summary table
summary_plant_pct <- tibble(
  statistic = c(
    "Mean (± SD)",
    "Median [Q25, Q75]",
    "IQR",
    "Tercile 1 cutoff (33.3%)",
    "Tercile 2 cutoff (66.7%)"
  ),
  value = c(
    sprintf("%.1f%% (± %.1f%%)", mean_val, unwt_sd),
    sprintf("%.1f%% [%.1f%%, %.1f%%]", median_val, q25_val, q75_val),
    sprintf("%.1f%%", q75_val - q25_val),
    sprintf("%.1f%%", t1_val),
    sprintf("%.1f%%", t2_val)
  ),
  interpretation = c(
    "Population average plant-food consumption",
    "Central tendency and spread",
    "Interquartile range",
    "Lower third of population below this threshold",
    "Upper third of population above this threshold"
  )
)

export(summary_plant_pct, here("01_data", "02_processed", "web_resources", "04_03_tbl_a3_plant_percentage_summary.rds"))

# Create GT table
summary_plant_pct |>
  gt() |>
  tab_header(
    title = "Plant-Based Food Percentage Distribution",
    subtitle = sprintf("Children 6-59 months (N = %d), SIVESNU 2018", n_sample)
  ) |>
  cols_label(
    statistic = "Statistic",
    value = "Value",
    interpretation = "Interpretation"
  ) |>
  tab_spanner(
    label = "Survey-Weighted Estimates",
    columns = c(value, interpretation)
  ) |>
  tab_footnote(
    footnote = "Survey weights applied (pesonino). Tercile cutoffs per Allen et al. (2020) H-AR methodology.",
    locations = cells_title(groups = "subtitle")
  ) |>
  tab_source_note(
    source_note = md("**Source**: SIVESNU 2018. SD estimated from unweighted data.")
  ) |>
  cols_width(
    statistic ~ px(180),
    value ~ px(180),
    interpretation ~ px(280)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )

# Cleanup
rm(svy_plant, weighted_mean, mean_val, unwt_sd,
   weighted_quantiles, q25_val, median_val, q75_val,
   weighted_terciles, t1_val, t2_val, summary_plant_pct, n_sample)
```

::: callout-note
### Methodological Decision

**Finding**: The distribution shows that the great majority of children have diets above 70% plant-based food consumption, and the lower tercile cutoff also falls within the "very high phytate" category.

**Decision**: Based on predominantly plant-based dietary patterns in the population, we apply a uniform **low iron absorption rate of 5%** to the entire population, following Allen et al. (2020) H-AR recommendations for unrefined/high-phytate diets.

**Iron requirement benchmark**: With 5% absorption and a recommended intake of 10 mg/day (Allen et al., 2020), the absorbed iron requirement is **0.5 mg/day**.
:::

------------------------------------------------------------------------

# Phase B: Nutrient Calculations

## B1. Calculate Total Nutrient Intake

Aggregate maize and non-maize sources to obtain total daily intake for all nutrients. 

```{r}
#| label: b1-calculate-total-intake

df_totals <- df_input |>
  mutate(
    # === TOTAL NUTRIENT INTAKE (maize + non-maize) ===
    # Iron (mg/day) - non-maize includes Chispitas
    fe_mg_total = fe_mg_maiz + fe_mg_nomaiz,
    
    # Zinc (mg/day) - non-maize includes Chispitas
    zn_mg_total = zn_mg_maiz + zn_mg_nomaiz,
    
    # Digestible protein (g/day)
    prot_g_total = prot_g_maiz + prot_g_nomaiz,
    
    # Lysine (mg/day)
    lys_mg_total = lys_mg_maiz + lys_mg_nomaiz,
    
    # Tryptophan (mg/day)
    trp_mg_total = trp_mg_maiz + trp_mg_nomaiz,
    
    # Energy (kcal/day)
    ene_kcal_total = ene_kcal_maiz + ene_kcal_nomaiz
  )
```

## B2. Apply Miller Equation for Zinc

Calculate Total Absorbed Zinc (TAZ) using the Miller equation for children. The equation is applied to total zinc intake (diet + Chispitas).

```{r}
#| label: b2-apply-miller-equation

df_miller <- df_totals |>
  mutate(
    # === MILLER EQUATION COMPONENTS ===
    # Age-dependent absorption capacity
    miller_age_factor = MILLER_A * (edad_mes ^ MILLER_E),
    
    # Sum of constants and zinc intake
    miller_sum = MILLER_K_T + miller_age_factor + zn_mg_total,
    
    # Discriminant for quadratic formula
    miller_discriminant = (miller_sum ^ 2) - (4 * miller_age_factor * zn_mg_total),
    
    # === TOTAL ABSORBED ZINC (TAZ) ===
    taz_mg = 0.5 * (miller_sum - sqrt(miller_discriminant)),
    
    # === FRACTIONAL ABSORPTION (for diagnostics) ===
    zn_fractional_absorption = taz_mg / zn_mg_total
  ) |>
  # Remove intermediate calculation columns
  select(-miller_age_factor, -miller_sum, -miller_discriminant)
```

## B3. Apply Iron Bioavailability Factor

Apply uniform 5% iron absorption rate to total iron intake. This conservative factor reflects the predominantly plant-based dietary patterns observed in this population.

```{r}
#| label: b3-apply-iron-bioavailability

df_bioavailable <- df_miller |>
  mutate(
    # === IRON ABSORPTION FACTOR (uniform) ===
    fe_absorption_factor = 0.05,
    
    # === ABSORBED IRON ===
    fe_absorbed_mg = fe_mg_total * fe_absorption_factor
  )
```

------------------------------------------------------------------------

# Phase C: Diagnostic Tables

## C1. Zinc Absorption by Age Group (Miller)

```{r}
#| label: tbl-c1-miller-by-age
#| tbl-cap: "Zinc Absorption by Age Group (Miller Equation)"

# Create survey design for weighted estimates
svy_bioavailable <- df_bioavailable |>
 as_survey_design(

    ids = conglomerado,
    weights = pesonino
  )

# TAZ by age group
tbl_c1_miller_by_age <- svy_bioavailable |>
  mutate(
    age_group = cut(
      edad_mes,
      breaks = c(6, 12, 24, 36, 48, 60),
      labels = c("6-11 mo", "12-23 mo", "24-35 mo", "36-47 mo", "48-59 mo"),
      include.lowest = TRUE
    )
  ) |>
  group_by(age_group) |>
  summarise(
    n = n(),
    mean_zn_intake = survey_mean(zn_mg_total, vartype = "ci"),
    mean_taz = survey_mean(taz_mg, vartype = "ci"),
    mean_absorption_pct = survey_mean(zn_fractional_absorption * 100, vartype = "ci")
  )

export(tbl_c1_miller_by_age, here("01_data", "02_processed", "web_resources", "04_03_tbl_c1_miller_by_age.rds"))

tbl_c1_miller_by_age |>
  gt() |>
  tab_header(
    title = "Zinc Absorption by Age Group",
    subtitle = "Miller equation results for children 6-59 months"
  ) |>
  cols_label(
    age_group = "Age Group",
    n = "N",
    mean_zn_intake = "Zn Intake",
    mean_zn_intake_low = "95% CI",
    mean_zn_intake_upp = "",
    mean_taz = "TAZ",
    mean_taz_low = "95% CI",
    mean_taz_upp = "",
    mean_absorption_pct = "Absorption",
    mean_absorption_pct_low = "95% CI",
    mean_absorption_pct_upp = ""
  ) |>
  cols_merge(columns = c(mean_zn_intake_low, mean_zn_intake_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(mean_taz_low, mean_taz_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(mean_absorption_pct_low, mean_absorption_pct_upp), pattern = "[{1}, {2}]") |>
  fmt_number(columns = c(mean_zn_intake, mean_taz), decimals = 2) |>
  fmt_number(columns = c(mean_zn_intake_low, mean_taz_low), decimals = 2) |>
  fmt_number(columns = c(mean_absorption_pct, mean_absorption_pct_low), decimals = 1) |>
  tab_spanner(label = "Zinc Intake (mg/d)", columns = c(mean_zn_intake, mean_zn_intake_low)) |>
  tab_spanner(label = "Absorbed Zinc (mg/d)", columns = c(mean_taz, mean_taz_low)) |>
  tab_spanner(label = "Fractional (%)", columns = c(mean_absorption_pct, mean_absorption_pct_low)) |>
  tab_source_note("TAZ = Total Absorbed Zinc. Absorption increases with age per Miller equation.") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(tbl_c1_miller_by_age)
```

## C2. Iron Absorption Summary

```{r}
#| label: tbl-c2-iron-summary
#| tbl-cap: "Iron Intake and Absorption Summary"

tbl_c2_iron_summary <- svy_bioavailable |>
  summarise(
    n = n(),
    mean_fe_intake = survey_mean(fe_mg_total, vartype = "ci"),
    mean_fe_absorbed = survey_mean(fe_absorbed_mg, vartype = "ci")
  ) |>
  mutate(
    absorption_factor = "5% (low absorption)",
    absorbed_requirement = "0.5 mg/day"
  )

export(tbl_c2_iron_summary, here("01_data", "02_processed", "web_resources", "04_03_tbl_c2_iron_summary.rds"))

tbl_c2_iron_summary |>
  gt() |>
  tab_header(
    title = "Iron Intake and Absorption",
    subtitle = "Children 6-59 months, uniform 5% absorption factor"
  ) |>
  cols_label(
    n = "N",
    mean_fe_intake = "Mean Intake",
    mean_fe_intake_low = "95% CI",
    mean_fe_intake_upp = "",
    mean_fe_absorbed = "Mean Absorbed",
    mean_fe_absorbed_low = "95% CI",
    mean_fe_absorbed_upp = "",
    absorption_factor = "Absorption Factor",
    absorbed_requirement = "Requirement (absorbed)"
  ) |>
  cols_merge(columns = c(mean_fe_intake_low, mean_fe_intake_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(mean_fe_absorbed_low, mean_fe_absorbed_upp), pattern = "[{1}, {2}]") |>
  fmt_number(columns = c(mean_fe_intake, mean_fe_intake_low, mean_fe_absorbed, mean_fe_absorbed_low), decimals = 2) |>
  tab_spanner(label = "Iron Intake (mg/d)", columns = c(mean_fe_intake, mean_fe_intake_low)) |>
  tab_spanner(label = "Absorbed Iron (mg/d)", columns = c(mean_fe_absorbed, mean_fe_absorbed_low)) |>
  tab_spanner(label = "Methodology", columns = c(absorption_factor, absorbed_requirement)) |>
  tab_source_note("Absorption factor based on Allen et al. (2020) H-AR for unrefined diets.") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(tbl_c2_iron_summary)
```

## C3. Overall Bioavailability Summary

```{r}
#| label: tbl-c3-bioavailability-summary
#| tbl-cap: "Micronutrient Bioavailability Summary"

tbl_c3_bioavailability_summary <- svy_bioavailable |>
  summarise(
    # Zinc
    zn_intake_mean = survey_mean(zn_mg_total, vartype = "ci"),
    taz_mean = survey_mean(taz_mg, vartype = "ci"),
    zn_abs_pct = survey_mean(zn_fractional_absorption * 100, vartype = "ci"),
    # Iron
    fe_intake_mean = survey_mean(fe_mg_total, vartype = "ci"),
    fe_absorbed_mean = survey_mean(fe_absorbed_mg, vartype = "ci"),
    fe_abs_pct = survey_mean(fe_absorption_factor * 100, vartype = "ci")
  ) |>
  pivot_longer(everything(), names_to = "metric", values_to = "value") |>
  mutate(
    nutrient = case_when(
      str_detect(metric, "^zn|^taz") ~ "Zinc",
      str_detect(metric, "^fe") ~ "Iron"
    ),
    measure = case_when(
      str_detect(metric, "intake_mean$") ~ "Total Intake (mg/d)",
      str_detect(metric, "^taz_mean$|absorbed_mean$") ~ "Absorbed (mg/d)",
      str_detect(metric, "abs_pct$") ~ "Absorption Rate (%)",
      str_detect(metric, "_low$") ~ "ci_low",
      str_detect(metric, "_upp$") ~ "ci_upp"
    ),
    ci_type = case_when(
      str_detect(metric, "_low$") ~ "low",
      str_detect(metric, "_upp$") ~ "upp",
      TRUE ~ "est"
    ),
    measure_clean = str_remove(metric, "(_low|_upp)$")
  ) |>
  filter(ci_type == "est") |>
  select(nutrient, measure, value) |>
  filter(!is.na(measure)) |>
  pivot_wider(names_from = nutrient, values_from = value)

export(tbl_c3_bioavailability_summary, here("01_data", "02_processed", "web_resources", "04_03_tbl_c3_bioavailability_summary.rds"))

tbl_c3_bioavailability_summary |>
  gt() |>
  tab_header(
    title = "Micronutrient Bioavailability Summary",
    subtitle = "Survey-weighted estimates for children 6-59 months"
  ) |>
  cols_label(
    measure = "Metric",
    Zinc = "Zinc",
    Iron = "Iron"
  ) |>
  fmt_number(columns = c(Zinc, Iron), decimals = 2) |>
  tab_source_note("Zinc: Miller equation (age-dependent). Iron: Assumed uniform 5% absorption based on dietary patterns") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(tbl_c3_bioavailability_summary)
```

------------------------------------------------------------------------

# Phase D: Export

```{r}
#| label: d-save-output

# Prepare final dataset
df_output <- df_bioavailable |>
  select(
    # === Identifiers ===
    nboletahogar, conglomerado, pesonino,
    
    # === Demographics ===
    parentesco, sexo, edad_an, edad_mes, edad_grupos,
    
    # === Geographic ===
    departamento, area,
    
    # === Household characteristics ===
    miembros_hogar_original, n_cuartos_original,
    grado_estudios, grado_estudios_hogar,
    propiedad, tipo_vivienda,
    material_paredes, material_techo, material_piso,
    
    # === Basic services ===
    cocina, tipo_sanitario, fuente_agua, recoleccion_basura,
    
    # === Household assets ===
    electricidad, televisor, telefonia_fija, telefonia_celular, computadora,
    
    # === Food security (FIES) ===
    fies1, fies2, fies3, fies4, fies5, fies6, fies7, fies8,
    fies_raw_score,
    
    # === Nutritional intake by source (mg/day or kcal/day) ===
    fe_mg_maiz, fe_mg_nomaiz,
    zn_mg_maiz, zn_mg_nomaiz,
    prot_g_maiz, prot_g_nomaiz,
    lys_mg_maiz, lys_mg_nomaiz,
    trp_mg_maiz, trp_mg_nomaiz,
    ene_kcal_maiz, ene_kcal_nomaiz,

    # === Raw transfer-model predictions (preserved from 04_01) ===
    # Pre-FIES, pre-quantile-mapping, pre-alignment. Diagnostics only.
    fe_mg_maiz_raw, fe_mg_nomaiz_raw,
    zn_mg_maiz_raw, zn_mg_nomaiz_raw,
    prot_g_maiz_raw, prot_g_nomaiz_raw,
    lys_mg_maiz_raw, lys_mg_nomaiz_raw,
    trp_mg_maiz_raw, trp_mg_nomaiz_raw,
    ene_kcal_maiz_raw, ene_kcal_nomaiz_raw,

    # === Dietary composition ===
    per_vegetales,
    
    # === Bioavailable micronutrients ===
    # Zinc (Miller equation)
    zn_mg_total,
    taz_mg,
    zn_fractional_absorption,
    # Iron
    fe_mg_total,
    fe_absorption_factor,
    fe_absorbed_mg,
    
    # === Total intake (other nutrients) ===
    prot_g_total, lys_mg_total, trp_mg_total, ene_kcal_total,
    
    # === Chispitas supplementation ===
    receives_chispitas,
    adherence_factor_final,
    fe_chispitas_effective_mg,
    zn_chispitas_effective_mg
  )

# Save final dataset
write_parquet(
  df_output,
  here("01_data", "02_processed", 
       "transfer", "04_03_sivesnu_bioavailable.parquet")
)
```

# Notes

## Methodology

### Miller Equation for Zinc

The Miller equation estimates Total Absorbed Zinc (TAZ) among children based on age and total zinc intake:

$$TAZ = 0.5 \times \left( K_T + A \cdot Age^E + TDZ - \sqrt{(K_T + A \cdot Age^E + TDZ)^2 - 4 \cdot (A \cdot Age^E) \cdot TDZ} \right)$$

Parameters used:

-   $K_T = 2.42$ (saturation constant)
-   $A = 0.633$ (age coefficient)
-   $E = 0.311$ (age exponent)

Key characteristics:

-   Absorption capacity increases with age
-   Fractional absorption decreases at higher zinc intakes (saturation effect)
-   Validated for combined dietary + supplemental zinc (including MNP studies)

### Iron Bioavailability

Iron absorption is estimated using a uniform 5% factor based on the predominantly plant-based dietary patterns observed in this population.

**Methodology justification**:

1. **Dietary analysis**: Survey-weighted analysis places the plant-food share of the diet in the very-high-phytate range for the great majority of children, as reported in the Phase A distribution.

2. **Reference framework**: Allen et al. (2020) H-AR values recommend 5% absorption for "low absorption / unrefined diet" scenarios with high phytate content.

3. **Absorbed iron requirement**: With 5% absorption and 10 mg/day recommended intake, the absorbed iron requirement is 0.5 mg/day.

This absorbed-iron requirement is consistent with the bioavailability methodology adopted for predominantly plant-based child populations.

## Limitations

1.  **Iron bioavailability factors**: A uniform 5% absorption rate is applied to the entire population; individual variation in absorption is not modelled. 

2.  **Chispitas mineral form**: Modeled iron and zinc absorption does not account for potential differences in absorption of zinc sulfate and ferrous fumarate in Chispitas relative to iron and zinc from dietary sources.

## References

-   Allen LH, Carriquiry AL, Murphy SP (2020). Perspective: Proposed Harmonized Nutrient Reference Values for Populations. *Advances in Nutrition* 11:469-483. doi: 10.1093/advances/nmz096.

-   Miller LV, Krebs NF, Hambidge KM (2007). A mathematical model of zinc absorption in humans as a function of dietary zinc and phytate. *Journal of Nutrition* 137(1):135-141.

-   International Zinc Nutrition Consultative Group (IZiNCG) (2004). Technical Brief No. 3: Assessment of the risk of zinc deficiency in populations.

-   WHO/FAO (2004). Vitamin and mineral requirements in human nutrition, 2nd edition.

-   Engle-Stone R (2025). Personal communication regarding the application of the Miller equation for children receiving micronutrient powders.

-   Soto Méndez MJ (2025). Personal communication regarding iron bioavailability methodology for the Guatemalan child population.
