---
title: "Prepare SIVESNU Nutrient Data - Transfer Models Application"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script applies the nutrient transfer models trained in Module 3 (ENCOVI 2023) to SIVESNU 2018 data, generating predicted nutrient intake values for all individuals in the population.

Main operations:

1. Load SIVESNU 2018 household and individual data
2. Load Module 3 transfer models, Yeo-Johnson transformations, and ENCOVI reference data
3. Harmonize and recode SIVESNU variables to match ENCOVI structure
4. Impute missing values
5. Apply Yeo-Johnson transformations to continuous predictors
6. Predict nutrient values using Module 3 models (iron, zinc, protein, lysine, tryptophan, energy, plant percentage)
7. Filter to the target population (children 6-59 months)
8. Apply quantile mapping against an age-matched ENCOVI reference
9. Validate distributions and save processed dataset

**Key methodological decision**: Quantile mapping is age-matched. The target population is filtered to children 6-59 months before calibration, and the ENCOVI reference is restricted to the same age range, so that source and reference describe comparable populations.

# Setup

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  janitor,            # Data cleaning utilities
  haven,              # Manage Stata files
  bestNormalize,      # Yeo-Johnson predict() S3 method dispatch
  car,                # Yeo-Johnson diagnostic profiles (boxCox)
  cowplot,            # Fig dispositions
  gt,                 # Table formatting
  mirai,              # Persistent daemon pool for parallel quantile mapping
  RhpcBLASctl,        # BLAS/OpenMP thread control inside daemons
  tidyverse           # General purpose data manipulation
)

# Packages used via :: (not loaded to minimize memory footprint)
# missRanger — missRanger() via impute_rf_parallel() for RF-based imputation
# ggpubr     — get_legend() for plot composition

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file
source(here("02_code", "00_functions", "00_functions.R"))

# transfer/
ensure_output_dir(here("01_data", "02_processed", "transfer"))

# web_resources/
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

# Phase A: Data Loading

## A1. Load SIVESNU Raw Data and ENCOVI Reference Data

Load SIVESNU 2018 household and individual databases and ENCOVI 2023 reference nutrient distributions for quantile mapping calibration.

```{r}
#| label: a1-load-data

# Load SIVESNU 2018 household data
sivesnu_hogar <- read_parquet(
  here("01_data", "01_raw", "SIVESNU_2018", "sivesnu_gt18_hogar_publico.parquet")
) |>
  clean_names() |>
  zap_label() |>
  zap_labels() |>
  zap_formats()

# Load SIVESNU 2018 individual data
sivesnu_individuos <- read_parquet(
  here("01_data", "01_raw", "SIVESNU_2018", "sivesnu_gt18_miembros_publico.parquet")
) |>
  clean_names() |>
  zap_label() |>
  zap_labels() |>
  zap_formats()

# Load SIVESNU children data to get age in months for filtering
sivesnu_nino_edad <- read_parquet(
 here("01_data", "01_raw", "SIVESNU_2018", "sivesnu_gt18_nino_publico.parquet")
) |>
  clean_names() |>
  zap_label() |>
  zap_labels() |>
  zap_formats() |>
  transmute(
    nboletahogar,
    conglomerado,
    edad_mes,
    sexo_nino = case_when(
      sexonino == 1 ~ "Hombre",
      sexonino == 2 ~ "Mujer",
      TRUE ~ NA_character_
    )
  )

# Load ENCOVI 2023 reference data for quantile mapping
# This dataset contains the observed nutritional distributions from ENCOVI.
#
# The reference is restricted to the target age range. Nutrient intake scales
# with body size, and ENCOVI children aged 6-59 months consume roughly half of
# what the all-age population does (median ratios 0.57 for maize-source and
# 0.47-0.49 for non-maize components). Calibrating child predictions against an
# all-age distribution therefore mis-scales them, and does so by a different
# amount per nutrient, which distorts the composition of any index built from
# them rather than merely its level.
df_encovi_reference <- read_parquet(
  here("01_data", "02_processed",
       "transfer", "01_04_individual_nutritional_intake_complete.parquet")
) |>
  filter(edad_meses >= 6, edad_meses < 60) |>
  # Select only nutritional variables needed for quantile mapping
  transmute(
    departamento,
    fe_maiz = fe_individuo_dia_maiz, 
    fe_no_maiz = fe_individuo_dia_no_maiz,
    zn_maiz = zn_individuo_dia_maiz,
    zn_no_maiz = zn_individuo_dia_no_maiz,
    prot_apro_maiz = prot_apro_individuo_dia_maiz,
    prot_apro_no_maiz = prot_apro_individuo_dia_no_maiz,
    lys_maiz = lys_individuo_dia_maiz,
    lys_no_maiz = lys_individuo_dia_no_maiz,
    trp_maiz = trp_individuo_dia_maiz,
    trp_no_maiz = trp_individuo_dia_no_maiz, 
    kcal_maiz = kcal_individuo_dia_maiz,
    kcal_no_maiz = kcal_individuo_dia_no_maiz
  ) |>
  drop_na()
```

## A2. Load Transfer Models

Load Module 3 nutrient transfer models (iron, zinc, protein, lysine, tryptophan, energy, plant percentage) and Yeo-Johnson transformation objects for continuous predictors.

```{r}
#| label: a2-load-models

# Load Yeo-Johnson transformation objects
yj_transforms <- import(
  here("01_data", "02_processed", 
       "models", "03_01_yj_transforms.rds"),
  trust = TRUE
)

# Extract individual transformation objects
yj_miembros_hogar <- yj_transforms$yj_miembros_hogar
yj_n_cuartos <- yj_transforms$yj_n_cuartos

# Load iron models (maize and non-maize)
iron_models <- import(
  here("01_data", "02_processed", 
       "models", "03_02_iron_models.rds"),
  trust = TRUE
)
modelo_fe_maiz <- iron_models$model_fe_maize
modelo_fe_no_maiz <- iron_models$model_fe_non_maize

# Load zinc models (maize and non-maize)
zinc_models <- import(
  here("01_data", "02_processed", 
       "models", "03_03_zinc_models.rds"),
  trust = TRUE
)
modelo_zn_maiz <- zinc_models$model_zn_maize
modelo_zn_no_maiz <- zinc_models$model_zn_non_maize

# Load digestible protein models (maize and non-maize)
protein_models <- import(
  here("01_data", "02_processed", 
       "models", "03_04_digestible_protein_models.rds"),
  trust = TRUE
)
modelo_prot_apro_maiz <- protein_models$model_prot_apro_maize
modelo_prot_apro_no_maiz <- protein_models$model_prot_apro_non_maize

# Load lysine models (maize and non-maize)
lysine_models <- import(
  here("01_data", "02_processed", 
       "models", "03_05_lysine_models.rds"),
  trust = TRUE
)
modelo_lys_maiz <- lysine_models$model_lys_maize
modelo_lys_no_maiz <- lysine_models$model_lys_non_maize

# Load tryptophan models (maize and non-maize)
tryptophan_models <- import(
  here("01_data", "02_processed", 
       "models", "03_06_tryptophan_models.rds"),
  trust = TRUE
)
modelo_trp_maiz <- tryptophan_models$model_trp_maize
modelo_trp_no_maiz <- tryptophan_models$model_trp_non_maize

# Load energy models (maize and non-maize)
energy_models <- import(
  here("01_data", "02_processed", 
       "models", "03_07_energy_models.rds"),
  trust = TRUE
)
modelo_kcal_maiz <- energy_models$model_kcal_maize
modelo_kcal_no_maiz <- energy_models$model_kcal_non_maize

# Load plant percentage model
plant_model <- import(
  here("01_data", "02_processed", 
       "models", "03_08_plant_percentage_model.rds"),
  trust = TRUE
)
modelo_vegetales <- plant_model$model_plant_percentage

# Clean auxiliary list objects (keep only extracted models)
rm(yj_transforms, iron_models, zinc_models, protein_models, 
   lysine_models, tryptophan_models, energy_models, plant_model)
```

---

# Phase B: Variable Harmonization

## B1. Merge and Recode SIVESNU Variables

Recode SIVESNU 2018 variables to match ENCOVI 2023 structure and coding. This includes factor level alignment, variable renaming, and creation of derived variables needed for transfer model prediction.

```{r}
#| label: b1-merge-and-recode

# Merge household and individual tables, recode variables
df_merge <- sivesnu_individuos |>
  left_join(sivesnu_hogar, by = join_by(nboletahogar, conglomerado)) |>
  transmute(
    # Identifiers
    nboletahogar, 
    conglomerado, 
    # Survey weight (SIVESNU 2018 has no sampling weights, assign equal weight)
    pesohogar = 1,
    # Department
    departamento = factor(
      departamento,
      levels = c("Guatemala", "", "Sacatepéquez", "Chimaltenango", 
                 "Escuintla", "Santa Rosa", "Sololá", "Totonicapán", "Quezaltenango", 
                 "Suchitepéquez", "Retalhuleu", "San Marcos", "Huehuetenango", "Quiché", 
                 "Baja Verapaz", "Alta Verapaz", "Petén", "Izabal", "Zacapa", "Chiquimula", 
                 "Jalapa", "Jutiapa"),
      labels = c("Guatemala", "El Progreso", "Sacatepéquez", "Chimaltenango", 
                 "Escuintla", "Santa Rosa", "Sololá", "Totonicapán", "Quetzaltenango", 
                 "Suchitepéquez", "Retalhuleu", "San Marcos", "Huehuetenango", "Quiché", 
                 "Baja Verapaz", "Alta Verapaz", "Petén", "Izabal", "Zacapa", "Chiquimula", 
                 "Jalapa", "Jutiapa"), 
      ordered = TRUE
    ),
    # Urban/rural area
    area = factor(
      ruralurbano, 
      levels = c(1, 2), 
      labels = c("Urbana", "Rural")
    ),
    # Household size
    miembros_hogar = num_miembros_hogar,
    # Housing ownership
    propiedad = factor(
      ifelse(vivienda_propriedad == "", 96, vivienda_propriedad),
      levels = c(1, 2, 3, 4, 96), 
      labels = c("Propia y totalmente pagada", "Propia y pagandola a plazos", 
                 "Alquilada", "Cedida o prestada", "Otro")
    ),
    # Wall material
    material_paredes = factor(
      ifelse(vivienda_material_pared == "" | vivienda_material_pared == 8 | is.na(vivienda_material_pared), 
             96, vivienda_material_pared),
      levels = c(6, 3, 2, 5, 4, 1, 7, 96), 
      labels = c("Ladrillo", "Block", "Adobe", "Madera", 
                 "Lamina metalica", "Bajareque", "Lepa, palo o caña", "Otro")
    ),
    # Roof material
    material_techo = factor(
      ifelse(vivienda_material_techo == "" | is.na(vivienda_material_techo), 96, vivienda_material_techo),
      levels = c(5, 3, 4, 2, 1, 96), 
      labels = c("Concreto", "Lamina metalica", "Asbesto cemento", 
                 "Teja", "Paja, palma o similar", "Otro")
    ),
    # Floor material
    material_piso = factor(
      ifelse(vivienda_material_piso == "" | is.na(vivienda_material_piso), 96, vivienda_material_piso),
      levels = c(6, 5, 2, 7, 4, 3, 1, 96), 
      labels = c("Ladrillo ceramico", "Ladrillo de cemento", "Ladrillo de barro", 
                 "Torta de cemento", "Tierra", "Madera", "Tierra", "Tierra")
    ),
    # Housing type
    tipo_vivienda = factor(
      ifelse(vivienda_tipo == "" | is.na(vivienda_tipo), NA, vivienda_tipo),
      levels = c(1, 2, 3, 4, 5), 
      labels = c("Casa formal", "Apartamento", "Cuarto en casa de vecindad", 
                 "Rancho", "Casa improvisada")
    ),
    # Number of rooms
    n_cuartos = vivienda_cuartos,
    # Kitchen type
    cocina = factor(
      cocina_tiene, 
      levels = c(1, 2), 
      labels = c("De uso exclusivo", "De uso compartido")
    ),
    # Water source
    fuente_agua = factor(
      ifelse(agua_fuente %in% c("", 9), 96, agua_fuente),
      levels = c(1, 3, 2, 5, 6, 8, 7, 96), 
      labels = c("Tubería dentro de la vivienda", "Tubería fuera de la vivienda", 
                 "Chorro publico", "Pozo perforado", "Río, lago, manantial", 
                 "Camion cisterna", "Agua de lluvia", "Otro")
    ),
    # Sanitation type
    tipo_sanitario = factor(
      case_when(
        serv_san_tipo == 1 ~ "Uso exclusivo",
        serv_san_tipo == 2 ~ "Uso compartido",
        serv_san_tipo == 3 ~ "Uso compartido",
        serv_san_tipo == 4 ~ "Uso exclusivo",
        serv_san_tipo == 5 ~ "Uso exclusivo",
        serv_san_tipo == 6 ~ "Uso compartido",
        serv_san_tipo == 96 ~ "Uso compartido",
        TRUE ~ "Uso compartido"
      ),
      levels = c("Uso exclusivo", "Uso compartido")
    ),
    # Waste disposal
    recoleccion_basura = factor(
      ifelse(basura_deshacer == "", 96, basura_deshacer),
      levels = c(3, 5, 4, 2, 1, 96), 
      labels = c("Servicio municipal", "Servicio privado", 
                 "Servicio municipal", "La queman o la entierran", 
                 "La tiran en cualquier lugar", "La tiran en cualquier lugar")
    ),
    # Electricity access
    electricidad = factor(
      bienes_luz,
      levels = c("Yes", "No"), 
      labels = c("Si", "No")
    ),
    # TV ownership
    televisor = factor(
      bienes_television,
      levels = c("Yes", "No"), 
      labels = c("Si", "No")
    ),
    # Landline phone
    telefonia_fija = factor(
      bienes_telefonofija,
      levels = c("Yes", "No"), 
      labels = c("Si", "No")
    ),
    # Cell phone
    telefonia_celular = factor(
      bienes_celular,
      levels = c("Yes", "No"), 
      labels = c("Si", "No")
    ),
    # Computer
    computadora = factor(
      bienes_computadora,
      levels = c("Yes", "No"), 
      labels = c("Si", "No")
    ),
    # Relationship to household head
    parentesco = factor(
      case_when(
        parentesco == 1 ~ "Jefe del hogar",
        parentesco == 2 ~ "Esposo(a) o compañero(a)",
        parentesco %in% c(3, 10) ~ "Hijo(a)",
        parentesco == 4 ~ "Yerno o nuera",
        parentesco == 5 ~ "Nieto(a)",
        parentesco == 6 ~ "Padre o madre",
        parentesco == 7 ~ "Suegro(a)",
        parentesco == 8 ~ "Hermano(a)",
        parentesco == 9 ~ "Otro pariente",
        parentesco == 11 ~ "Empleado de casa particular",
        parentesco == 12 ~ "Otro no pariente",
        TRUE ~ "Otro pariente"
      )
    ),
    # Sex
    sexo = factor(
      sexo,
      levels = c(1, 2), 
      labels = c("Hombre", "Mujer")
    ),
    # Age
    edad = edad,
    # Age groups
    edad_grupos = factor(
      cut(
        edad, 
        breaks = c(-Inf, 0.5, 1, 2, 4, 7, 10, 14, 18, 30, 60, Inf), 
        labels = c("de 0 a 6 meses",
                   "de 6 meses a 1 año",
                   "de 1 a 2 años",
                   "de 2 a 4 años",
                   "de 4 a 7 años",
                   "de 7 a 10 años",
                   "de 10 a 14 años",
                   "de 14 a 18 años",
                   "de 18 a 30 años",
                   "de 30 a 60 años",
                   "más de 60 años")
      )
    ),
    # Education level
    grado_estudios = fct_relevel(
      factor(
        case_when(
          grado_estudio %in% c("", "9-No sabe") ~ "Desconocido",
          grado_estudio == "0-Ninguno" ~ "Ninguno",
          grado_estudio == "4-Alfabetización" ~ "Solo sabe leer",
          grado_estudio == "1-Primaria" ~ "Primaria",
          grado_estudio == "2-Secundaria" ~ "Secundaria",
          grado_estudio == "3-Superior/universitaria" ~ "Superior/universitaria"
        ),
        ordered = TRUE
      ), 
      c("Desconocido", "Ninguno", "Solo sabe leer", "Primaria", "Secundaria", "Superior/universitaria")
    ),
    # Education numeric for household average
    grado_estudios_num = case_when(
      grado_estudios == "Desconocido" ~ 1,
      grado_estudios == "Ninguno" ~ 1,
      grado_estudios == "Solo sabe leer" ~ 2,
      grado_estudios == "Primaria" ~ 4,
      grado_estudios == "Secundaria" ~ 7,
      grado_estudios == "Superior/universitaria" ~ 10
    ),
    # Keep education numeric only for household heads and spouses
    grado_estudios_num = if_else(
      parentesco %in% c("Jefe del hogar", "Esposo(a) o compañero(a)"),  
      grado_estudios_num, 
      NA
    ),
    # Recode Food Insecurity Experience Scale (FIES) variables:
    # 1=Yes→1, 2=No→0, 98/99→NA
    across(
      starts_with("fies"), 
      ~ factor(
        case_when(
          . == 1 ~ "Yes",
          . == 2 ~ "No",
          . %in% c(98, 99) ~ NA_character_,
          TRUE ~ NA_character_
        ),
        levels = c("No", "Yes")
      )
    )
  ) |>
  # Remove rows with missing identifiers or weights
  drop_na(conglomerado, pesohogar) |>
  # Calculate household-level education (mean of head and spouse)
  mutate(
    grado_estudios_hogar = mean(grado_estudios_num, na.rm = TRUE),
    .by = "nboletahogar", 
    .keep = "unused"
  ) |>
  # Remove Stata labels and formats for clean processing
  zap_label() |>
  zap_labels() |>
  zap_formats()
```

## B2. Impute Missing Values

Apply Random Forest-based multivariate imputation to handle missing values in predictor and FIES variables. This method preserves variable correlations, handles mixed data types (continuous and categorical), and provides out-of-bag error estimates for imputation quality.

Implementation uses the `impute_rf_parallel()` utility function, which wraps `missRanger` with a multi-threaded `ranger` engine and returns imputed values alongside out-of-bag error estimates (OOBE for continuous variables, PFC for categorical).

```{r}
#| label: tbl-b2-impute-missing
#| tbl-cap: "Random Forest imputation quality metrics"

# Separate predictors and FIES variables for imputation
predictors_to_impute <- df_merge |>
  select(
    parentesco, sexo, edad, edad_grupos, departamento, area,
    miembros_hogar, grado_estudios, grado_estudios_hogar, propiedad,
    tipo_vivienda, material_paredes, material_techo, material_piso,
    n_cuartos, cocina, tipo_sanitario,
    fuente_agua, recoleccion_basura, electricidad,
    televisor, telefonia_fija, telefonia_celular, computadora,
    starts_with("fies")
  ) |>
  # Convert ordered factors to regular factors for ranger compatibility
  mutate(
    departamento = factor(departamento, ordered = FALSE),
    grado_estudios = factor(grado_estudios, ordered = FALSE)
  ) |>
  # Convert to data frame
  as.data.frame()

# =============================================================================
# B2. Run Random Forest Imputation on Predictor and FIES Variables
# =============================================================================
# Uses impute_rf_parallel() (00_functions.R, section 6), which wraps
# missRanger with a multi-threaded ranger engine and returns a list
# containing ximp (imputed tibble) and OOBerror (OOBE, PFC).

# --- Run Random Forest imputation ---
set.seed(789) 

imputed_result <- impute_rf_parallel(
  data = predictors_to_impute,
  maxiter = 10,
  ntree = 100,
  verbose = FALSE
)

# Extract imputed data
predictors_imputed <- imputed_result$ximp

# Store imputation quality metrics for table
imputation_metrics <- imputed_result$OOBerror

# Report each metric only when a variable of that type required imputation
n_continuous <- imputed_result$n_continuous
n_categorical <- imputed_result$n_categorical

# Create quality metrics summary table
tbl_b2_impute_missing <- tibble(
  metric = c(
    "OOBE (Continuous Variables)",
    "PFC (Categorical Variables)",
    "Number of Iterations",
    "Trees per Iteration"
  ),
  value = c(
    if (n_continuous > 0) imputation_metrics["OOBE"] else NA_real_,
    if (n_categorical > 0) imputation_metrics["PFC"] else NA_real_,
    10,
    100
  ),
  interpretation = c(
    if (n_continuous > 0) {
      if_else(imputation_metrics["OOBE"] < 0.5, "Acceptable", "Review required")
    } else {
      "No continuous variables required imputation"
    },
    if (n_categorical > 0) {
      if_else(imputation_metrics["PFC"] < 0.2, "Acceptable", "Review required")
    } else {
      "No categorical variables required imputation"
    },
    "Fixed parameter",
    "Fixed parameter"
  )
)

export(tbl_b2_impute_missing, here("01_data", "02_processed", "web_resources", "04_01_tbl_b2_impute_missing.rds"))

tbl_b2_impute_missing |>
  gt() |>
  tab_header(
    title = "Missing Value Imputation Quality Assessment",
    subtitle = "Out-of-bag error estimates from Random Forest imputation (includes FIES variables)"
  ) |>
  cols_label(
    metric = "Metric",
    value = "Value",
    interpretation = "Quality Assessment"
  ) |>
  fmt_number(
    columns = value,
    decimals = 4
  ) |>
  sub_missing(
    columns = value,
    missing_text = "N/A"
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "Acceptable"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "Review required"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#D1ECF1"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "No continuous variables required imputation"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#D1ECF1"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "No categorical variables required imputation"
    )
  ) |>
  tab_footnote(
    footnote = "OOBE: out-of-bag error across continuous variables, the square root of the variance-weighted mean of per-variable unexplained variance. Values < 0.5 indicate acceptable imputation quality. N/A when no continuous variables had missing values.",
    locations = cells_body(columns = metric, rows = 1)
  ) |>
  tab_footnote(
    footnote = "PFC: Proportion of Falsely Classified entries. Values < 0.2 indicate acceptable imputation quality for categorical variables. N/A when no categorical variables had missing values.",
    locations = cells_body(columns = metric, rows = 2)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Clean auxiliary objects from quality summary (keep predictors_imputed
# and imputed_result alive; they are consumed by B2b below).
rm(
  predictors_to_impute,
  imputation_metrics,
  n_continuous,
  n_categorical,
  tbl_b2_impute_missing
)
```

## B2b. Completion of Residual Missing Values

The Random Forest imputation engine can leave a residual `NaN` for a small subset of rows when the out-of-bag prediction collapses inside a terminal node with zero within-leaf variance. This affects continuous discrete-scale predictors.

Downstream analyses (Yeo-Johnson transformation, model prediction, quantile mapping) operate on a complete predictor matrix, so residual missing values are completed with a variable-specific rule:

- **Numeric / integer variables**: median of the observed values, which preserves the support of discrete-scale ordinal variables (e.g. years of schooling take values such as 1, 1.5, 2, 2.5, ...).
- **Factor variables**: mode of the observed values.

Completing rather than discarding keeps every record in the predicted distributions. The count of values completed by each rule is reported in the table below.

```{r}
#| label: tbl-b2b-completion-summary
#| tbl-cap: "Residual missing value completion after Random Forest imputation"

# --- Identify residual missing values ---
residual_summary <- predictors_imputed |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n_completed") |>
  filter(n_completed > 0) |>
  mutate(
    rule = vapply(
      variable,
      function(v) {
        x <- predictors_imputed[[v]]
        if (is.numeric(x)) "Median" else "Mode"
      },
      character(1)
    )
  ) |>
  arrange(desc(n_completed))

# --- Apply completion rules ---
mode_value <- function(x) {
  observed <- x[!is.na(x)]
  tab <- table(observed)
  names(tab)[which.max(tab)]
}

predictors_complete <- predictors_imputed |>
  mutate(across(
    where(is.numeric),
    ~ if_else(is.na(.), median(., na.rm = TRUE), .)
  )) |>
  mutate(across(
    where(is.factor),
    ~ {
      mv <- mode_value(.)
      replace(., is.na(.), mv)
    }
  ))

# --- Build summary with post-completion integrity row -----------------------
n_residual_after <- sum(is.na(predictors_complete))

completion_report <- residual_summary |>
  bind_rows(
    tibble(
      variable    = "\u2014 Residual NAs after completion \u2014",
      n_completed = n_residual_after,
      rule        = if_else(n_residual_after == 0L,
                            "Integrity check: passed",
                            "Integrity check: review required")
    )
  )

# --- Export summary for web resources ---
export(
  completion_report,
  here("01_data", "02_processed", "web_resources", "04_01_tbl_b2b_completion_summary.rds")
)

# --- Present summary as GT table ---
completion_report |>
  gt() |>
  tab_header(
    title = "Residual Missing Value Completion",
    subtitle = "Values completed by variable-specific rule after RF imputation"
  ) |>
  cols_label(
    variable = "Variable",
    n_completed = "N Completed",
    rule = "Completion Rule"
  ) |>
  fmt_number(
    columns = n_completed,
    decimals = 0,
    use_seps = TRUE
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      rows = rule == "Integrity check: passed"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(
      rows = rule == "Integrity check: review required"
    )
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# --- Combine complete predictors with identifiers ---
df_imputed <- df_merge |>
  select(nboletahogar, conglomerado, pesohogar) |>
  bind_cols(predictors_complete)

# Clean auxiliary objects
rm(
  imputed_result,
  predictors_imputed,
  predictors_complete,
  residual_summary,
  completion_report,
  n_residual_after
)
```

---

# Phase C: Apply Transformations

Apply Yeo-Johnson transformations to continuous predictors (`miembros_hogar` [number of household members] and `n_cuartos` [number of rooms]) using transformation objects fitted on ENCOVI 2023 data. These transformations normalize skewed distributions before prediction.

```{r}
#| label: fig-c-yj-miembros-hogar-pre
#| fig-cap: "Yeo-Johnson transformation profile for household members (pre-transformation)"

export(df_imputed, here("01_data", "02_processed", "web_resources", "04_01_fig_c_yj_miembros_hogar_pre.rds"))

# Evaluate distribution before transformation
boxCox(
  lm(miembros_hogar ~ 1, data = df_imputed),
  main = "Household Members - Pre-Transformation",
  family = "yjPower"
)
```

```{r}
#| label: fig-c-yj-miembros-hogar-post
#| fig-cap: "Yeo-Johnson transformation profile for household members (post-transformation)"

# Apply transformation to evaluate post-transformation profile
miembros_hogar_transformed <- predict(yj_miembros_hogar, newdata = df_imputed$miembros_hogar)

export(tibble(miembros_hogar_transformed = miembros_hogar_transformed), here("01_data", "02_processed", "web_resources", "04_01_fig_c_yj_miembros_hogar_post.rds"))

# Verify transformation improved normality
boxCox(
  lm(miembros_hogar_transformed ~ 1),
  main = "Household Members - Post-Transformation",
  family = "yjPower"
)

# Clean temporary object
rm(miembros_hogar_transformed)
```

```{r}
#| label: fig-c-yj-n-cuartos-pre
#| fig-cap: "Yeo-Johnson transformation profile for number of rooms (pre-transformation)"

export(df_imputed, here("01_data", "02_processed", "web_resources", "04_01_fig_c_yj_n_cuartos_pre.rds"))

# Evaluate distribution before transformation
boxCox(
  lm(n_cuartos ~ 1, data = df_imputed),
  main = "Number of Rooms - Pre-Transformation",
  family = "yjPower"
)
```

```{r}
#| label: fig-c-yj-n-cuartos-post
#| fig-cap: "Yeo-Johnson transformation profile for number of rooms (post-transformation)"

# Apply transformation to evaluate post-transformation profile
n_cuartos_transformed <- predict(yj_n_cuartos, newdata = df_imputed$n_cuartos)

export(tibble(n_cuartos_transformed = n_cuartos_transformed), here("01_data", "02_processed", "web_resources", "04_01_fig_c_yj_n_cuartos_post.rds"))

# Verify transformation improved normality
boxCox(
  lm(n_cuartos_transformed ~ 1),
  main = "Number of Rooms - Post-Transformation",
  family = "yjPower"
)

# Clean temporary object
rm(n_cuartos_transformed)
```

```{r}
#| label: apply-c-yj-transformations

# Apply Yeo-Johnson transformations to create modeling-ready dataset
df_yj <- df_imputed |>
  mutate(
    # Keep original values for reference
    miembros_hogar_original = miembros_hogar,
    n_cuartos_original = n_cuartos,
    # Apply transformations
    miembros_hogar = predict(yj_miembros_hogar, newdata = miembros_hogar),
    n_cuartos = predict(yj_n_cuartos, newdata = n_cuartos)
  )
```

---

# Phase D: Nutrient Prediction

Apply Module 3 transfer models to predict nutrient intake values for all SIVESNU individuals. Models predict maize and non-maize components separately for each nutrient.

**Note**: GLM models with log link return predictions in log scale, requiring exp() back-transformation. Exception: plant percentage model uses identity link (no transformation needed).

## D1. Predict Nutrient Values

```{r}
#| label: d1-predict-nutritional-values
 
# Align nominal predictor levels with those used during model training in
# Module 03. "Otro" is retired from tipo_vivienda because the level carries
# no observations in either survey and is absent from the trained models.
df_yj_predict <- df_yj |>
  mutate(
    tipo_vivienda = fct_drop(tipo_vivienda, only = "Otro")
  )
 
# Create list of all loaded models for automatic prediction
lista_modelos <- mget(ls(pattern = "^modelo_"))
 
# Predict nutritional values using all models
df_predicted <- df_yj |>
  bind_cols(
    imap_dfc(lista_modelos, ~ {
      # Remove 'modelo_' prefix from column name
      nuevo_nombre <- str_replace(.y, "modelo_", "")
      # Change 'vegetales' to 'per_vegetales'
      if (nuevo_nombre == "vegetales") nuevo_nombre <- "per_vegetales"
      # Predict and strip svystat attributes
      # Note: per_vegetales model uses identity link (no back-transform needed)
      #       All other models use log link (require exp() back-transform)
      predictions <- if (nuevo_nombre == "per_vegetales") {
        as.numeric(predict(.x, newdata = df_yj_predict))
      } else {
        exp(as.numeric(predict(.x, newdata = df_yj_predict)))
      }
      # Return clean numeric column
      tibble(!!nuevo_nombre := predictions)
    })
  )
 
# Clean auxiliary objects
rm(df_yj_predict, lista_modelos)
```

## D2. Filter Target Population

Filter to children 6-59 months, the target population for the stunting analysis.

The filter precedes the quantile mapping. Calibration maps the quantiles of a source distribution onto those of a reference, so the two describe comparable populations: the reference loaded in A1 is restricted to ENCOVI children of the same age range, and filtering first keeps source and reference age-consistent.

```{r}
#| label: d2-filter-children

# Filter to children 6-59 months using fuzzy join on household + demographics
df_children <- df_predicted |>
  left_join(
    sivesnu_nino_edad,
    by = c("nboletahogar", "conglomerado"),
    relationship = "many-to-many"
  ) |>
  filter(
    sexo == sexo_nino | is.na(sexo_nino),
    edad >= 0 & edad <= 5
  ) |>
  filter(edad_mes >= 6 & edad_mes <= 59) |>
  distinct(nboletahogar, conglomerado, sexo, edad, .keep_all = TRUE)

# Clean auxiliary object
rm(sivesnu_nino_edad)
```

## D3. Apply Quantile Mapping with FIES Adjustment

Apply pre-quantile mapping adjustment based on FIES food security status, then calibrate predicted distributions to match the ENCOVI reference distributions using quantile mapping.

**FIES-based adjustment rationale:**
- Food Secure (score=0): Higher food access → predictions adjusted +5%
- Mild Food Insecurity (score 1-3): Slight decrease in quantity and/or quality of food consumed → predictions adjusted +2%
- Moderate Food Insecurity (score 4-6): Moderate reduction in quantity and/or quality of food consumed → predictions adjusted -2%
- Severe Food Insecurity (score 7-8): Have run out of food for a day or more → predictions adjusted -5%

```{r}
#| label: d3-apply-quantile-mapping

# Define nutritional variables for adjustment
nutritional_vars <- c(
  "fe_maiz", "fe_no_maiz", 
  "zn_maiz", "zn_no_maiz",
  "prot_apro_maiz", "prot_apro_no_maiz",
  "lys_maiz", "lys_no_maiz",
  "trp_maiz", "trp_no_maiz",
  "kcal_maiz", "kcal_no_maiz"
)

# Calculate FIES raw score and apply food security adjustment
df_qm_adjusted <- df_children |>
  mutate(
    # Calculate FIES raw score (sum of affirmative responses)
    fies_raw_score = (fies1 == "Yes") + (fies2 == "Yes") + (fies3 == "Yes") + 
                     (fies4 == "Yes") + (fies5 == "Yes") + (fies6 == "Yes") + 
                     (fies7 == "Yes") + (fies8 == "Yes"),
    # Apply FIES-based adjustment to nutritional variables
    across(
      all_of(nutritional_vars),
      ~ case_when(
        fies_raw_score == 0 ~ .x * 1.05,                              # Food Secure: +5%
        fies_raw_score >= 1 & fies_raw_score <= 3 ~ .x * 1.02,        # Mild FI: +2%
        fies_raw_score >= 4 & fies_raw_score <= 6 ~ .x * 0.98,        # Moderate FI: -2%
        fies_raw_score >= 7 ~ .x * 0.95,                              # Severe FI: -5%
        TRUE ~ .x
      ),
      .names = "{.col}_fies_adj"
    )
  )

# Create filtered reference distributions (remove zeros and Tukey outliers).
# This is a fast, vectorized operation; kept sequential.
reference_filtered <- map(
  df_encovi_reference[nutritional_vars],
  filter_reference_distribution
)

# Define variable names for mapping
fies_adjusted_vars <- paste0(nutritional_vars, "_fies_adj")
qm_adjusted_vars <- paste0(nutritional_vars, "_qm_adj")

# =============================================================================
# Apply quantile mapping with gamma distribution (parallel via mirai)
# =============================================================================
# The 12 calls to quantile_mapping_gamma() are independent: each operates on
# a distinct (source, reference) pair with no shared state. The bottleneck
# inside each call is two iterative MLE fits via fitdistrplus::fitdist.
# Since the age filter now precedes this step the source is the child
# subpopulation rather than the full survey, so the fits are considerably
# cheaper; the pool is retained because the 12 mappings still parallelise
# cleanly and the setup cost is negligible.
#
# Execution is delegated to a local mirai daemon pool sized to the number
# of nutrient pairs, capped by available physical cores. The pool is torn
# down on exit of the local() block, leaving no lingering processes in the
# user's R session.
df_qm_adjusted[qm_adjusted_vars] <- local({

  physical_cores <- parallel::detectCores(logical = FALSE)
  n_daemons      <- min(length(qm_adjusted_vars), max(1, physical_cores - 4))
  mirai_profile  <- "qm_gamma_04_01"

  mirai::daemons(n_daemons, .compute = mirai_profile)
  on.exit(mirai::daemons(0, .compute = mirai_profile), add = TRUE)

  # Constrain BLAS/OpenMP threads inside each daemon and import the
  # quantile mapping function so it is available without re-sourcing
  # 00_functions.R from inside the worker.
  mirai::everywhere(
    {
      if (requireNamespace("RhpcBLASctl", quietly = TRUE)) {
        RhpcBLASctl::blas_set_num_threads(1)
        RhpcBLASctl::omp_set_num_threads(1)
      }
    },
    .compute = mirai_profile
  )

  # Build paired argument list and dispatch
  pairs <- map2(
    df_qm_adjusted[fies_adjusted_vars],
    reference_filtered,
    ~ list(source_distribution = .x, reference_distribution = .y)
  )

  results <- mirai::mirai_map(
    .x       = pairs,
    .f       = function(p, qm_fun) {
      qm_fun(
        source_distribution    = p$source_distribution,
        reference_distribution = p$reference_distribution
      )
    },
    .args    = list(qm_fun = quantile_mapping_gamma),
    .compute = mirai_profile
  )[]

  # Restore original names (mirai_map returns positional names)
  names(results) <- qm_adjusted_vars
  results
})

# Clean auxiliary objects
rm(reference_filtered, fies_adjusted_vars)
```

## D4. Apply Nutrient Alignment

Apply rank-based alignment to the maize-source variables.

The alignment realigns each individual's maize-source values against their overall non-maize nutrient profile, under the assumption of intra-individual nutrient coherence. Maize dependency is itself a marker of household resources, so the maize-source block and the outcome share that gradient before the alignment.

```{r}
#| label: d4-apply-nutritional-alignment

# Define variable groups for alignment
maize_vars_qm <- c(
  "fe_maiz_qm_adj", "zn_maiz_qm_adj", "prot_apro_maiz_qm_adj",
  "lys_maiz_qm_adj", "trp_maiz_qm_adj", "kcal_maiz_qm_adj"
)

nomaize_vars_qm <- c(
  "fe_no_maiz_qm_adj", "zn_no_maiz_qm_adj", "prot_apro_no_maiz_qm_adj",
  "lys_no_maiz_qm_adj", "trp_no_maiz_qm_adj", "kcal_no_maiz_qm_adj"
)

# Apply rank-based nutritional coherence alignment
# This preserves marginal distributions while correcting correlation direction
df_children_aligned <- align_maize_nutrition(
  data = df_qm_adjusted,
  maize_vars = maize_vars_qm,
  nomaize_vars = nomaize_vars_qm,
  keep_diagnostics = FALSE
)

# Create final adjusted variables with consistent naming
df_final_adjusted <- df_children_aligned |>
  mutate(
    # Final maize variables (from alignment)
    fe_maiz_adjusted = fe_maiz_qm_adj_aligned,
    zn_maiz_adjusted = zn_maiz_qm_adj_aligned,
    prot_apro_maiz_adjusted = prot_apro_maiz_qm_adj_aligned,
    lys_maiz_adjusted = lys_maiz_qm_adj_aligned,
    trp_maiz_adjusted = trp_maiz_qm_adj_aligned,
    kcal_maiz_adjusted = kcal_maiz_qm_adj_aligned,
    
    # Final non-maize variables (direct from QM, no alignment needed)
    fe_no_maiz_adjusted = fe_no_maiz_qm_adj,
    zn_no_maiz_adjusted = zn_no_maiz_qm_adj,
    prot_apro_no_maiz_adjusted = prot_apro_no_maiz_qm_adj,
    lys_no_maiz_adjusted = lys_no_maiz_qm_adj,
    trp_no_maiz_adjusted = trp_no_maiz_qm_adj,
    kcal_no_maiz_adjusted = kcal_no_maiz_qm_adj
  ) |>
  # Remove intermediate variables
  select(-ends_with("_fies_adj"), -ends_with("_qm_adj"), -ends_with("_qm_adj_aligned"))

# Update final_adjusted_vars for downstream use
final_adjusted_vars <- paste0(nutritional_vars, "_adjusted")

# Clean auxiliary objects
rm(df_children, df_children_aligned, maize_vars_qm, nomaize_vars_qm, qm_adjusted_vars)
```

---

# Phase E: Final Dataset Preparation

## E1. Prepare Final Output

Select and organize variables for the final SIVESNU nutrient dataset. This dataset will be used in subsequent scripts for synthetic population generation and stunting model application.

```{r}
#| label: e1-prepare-final-output

# Raw transfer-model predictions, exported under the naming convention used
# from 04_02 onwards plus an explicit `_raw` suffix. These columns hold the
# model output before the FIES adjustment, the quantile mapping and the
# maize-source alignment, and are preserved for downstream diagnostics.
raw_prediction_vars <- c(
  fe_mg_maiz_raw      = "fe_maiz",
  fe_mg_nomaiz_raw    = "fe_no_maiz",
  zn_mg_maiz_raw      = "zn_maiz",
  zn_mg_nomaiz_raw    = "zn_no_maiz",
  prot_g_maiz_raw     = "prot_apro_maiz",
  prot_g_nomaiz_raw   = "prot_apro_no_maiz",
  lys_mg_maiz_raw     = "lys_maiz",
  lys_mg_nomaiz_raw   = "lys_no_maiz",
  trp_mg_maiz_raw     = "trp_maiz",
  trp_mg_nomaiz_raw   = "trp_no_maiz",
  ene_kcal_maiz_raw   = "kcal_maiz",
  ene_kcal_nomaiz_raw = "kcal_no_maiz"
)

# Select and organize final variables
df_sivesnu_nutritional <- df_final_adjusted |>
  select(
    # === Identifiers ===
    nboletahogar, conglomerado, pesohogar,
    
    # === Demographics ===
    parentesco, sexo, edad, edad_grupos,
    
    # === Geographic ===
    departamento, area,
    
    # === Household characteristics ===
    miembros_hogar_original, n_cuartos_original, 
    grado_estudios, grado_estudios_hogar,
    propiedad, tipo_vivienda,
    material_paredes, material_techo, material_piso,
    
    # === Basic services ===
    cocina, tipo_sanitario, fuente_agua, recoleccion_basura,
    
    # === Household assets ===
    electricidad, televisor, telefonia_fija, telefonia_celular, computadora,
    
    # === Food security (FIES) ===
    fies1, fies2, fies3, fies4, fies5, fies6, fies7, fies8,
    fies_raw_score,
    
    # === Nutritional values (raw transfer-model predictions) ===
    all_of(raw_prediction_vars),
    
    # === Nutritional values (FIES + quantile mapping adjusted) ===
    all_of(final_adjusted_vars),
    
    # === Plant-based food percentage ===
    per_vegetales
  )
```

## E2. Dataset Structure Summary

```{r}
#| label: tbl-e2-final-validation
#| tbl-cap: "Final dataset structure summary"

# Generate summary of final dataset
tbl_e2_final_validation <- tibble(
  category = c(
    "Identifiers",
    "Demographics",
    "Geographic",
    "Household characteristics",
    "Basic services",
    "Household assets",
    "Food security (FIES)",
    "Nutritional (raw predictions)",
    "Nutritional (adjusted)",
    "Other",
    "TOTAL"
  ),
  n_variables = c(
    3,   # nboletahogar, conglomerado, pesohogar
    4,   # parentesco, sexo, edad, edad_grupos
    2,   # departamento, area
    8,   # miembros_hogar_original, n_cuartos_original, grado_estudios, etc.
    4,   # cocina, tipo_sanitario, fuente_agua, recoleccion_basura
    5,   # electricidad, televisor, telefonia_fija, telefonia_celular, computadora
    9,   # fies1-8 + fies_raw_score
    12,  # raw_prediction_vars
    12,  # final_adjusted_vars
    1,   # per_vegetales
    ncol(df_sivesnu_nutritional)
  ),
  observations = c(
    rep(nrow(df_sivesnu_nutritional), 11)
  )
)

export(tbl_e2_final_validation, here("01_data", "02_processed", "web_resources", "04_01_tbl_e2_final_validation.rds"))

tbl_e2_final_validation |>
  gt() |>
  tab_header(
    title = "Final Dataset Structure",
    subtitle = "SIVESNU 2018 with predicted nutritional values"
  ) |>
  cols_label(
    category = "Variable Category",
    n_variables = "N Variables",
    observations = "N Observations"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = category == "TOTAL")
  ) |>
  tab_options(
    table.font.size = 11,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(tbl_e2_final_validation)
```

## E3. Validation and Quality Assessment

Assess the quality of the nutrient transfer by comparing predicted SIVESNU distributions with ENCOVI reference distributions. Successful transfer should show similar distribution shapes between populations.

```{r}
#| label: fig-e3-density-comparison
#| fig-cap: "Distribution comparison: SIVESNU predicted vs ENCOVI reference"
#| fig-width: 10
#| fig-height: 8

# Prepare ENCOVI reference data
df_encovi_plot <- df_encovi_reference |>
  transmute(
    source = "ENCOVI 2023 (Reference)",
    fe_total = fe_maiz + fe_no_maiz,
    zn_total = zn_maiz + zn_no_maiz,
    prot_total = prot_apro_maiz + prot_apro_no_maiz,
    lys_total = lys_maiz + lys_no_maiz,
    trp_total = trp_maiz + trp_no_maiz,
    kcal_total = kcal_maiz + kcal_no_maiz
  )

# Prepare SIVESNU predicted data
df_sivesnu_plot <- df_qm_adjusted |>
  transmute(
    source = "SIVESNU 2018 (Predicted)",
    fe_total = fe_maiz_qm_adj + fe_no_maiz_qm_adj,
    zn_total = zn_maiz_qm_adj + zn_no_maiz_qm_adj,
    prot_total = prot_apro_maiz_qm_adj + prot_apro_no_maiz_qm_adj,
    lys_total = lys_maiz_qm_adj + lys_no_maiz_qm_adj,
    trp_total = trp_maiz_qm_adj + trp_no_maiz_qm_adj,
    kcal_total = kcal_maiz_qm_adj + kcal_no_maiz_qm_adj
  )

# Combine datasets
df_comparison <- bind_rows(df_encovi_plot, df_sivesnu_plot) |>
  mutate(source = factor(source, levels = c("ENCOVI 2023 (Reference)", "SIVESNU 2018 (Predicted)")))

export(df_comparison, here("01_data", "02_processed", "web_resources", "04_01_fig_e3_density_comparison.rds"))

# Define color palette
color_palette <- c("ENCOVI 2023 (Reference)" = "#5C946E", "SIVESNU 2018 (Predicted)" = "#E07A5F")

# Define nutrients and labels
nutrients <- c("fe_total", "zn_total", "prot_total", "lys_total", "trp_total", "kcal_total")
nutrient_labels <- c(
  "fe_total" = "Iron (mg/day)",
  "zn_total" = "Zinc (mg/day)", 
  "prot_total" = "Digestible Protein (g/day)",
  "lys_total" = "Lysine (mg/day)",
  "trp_total" = "Tryptophan (mg/day)",
  "kcal_total" = "Energy (kcal/day)"
)

# Create individual density plots (without legend)
plot_list <- map(nutrients, ~ {
  upper_limit <- quantile(df_comparison[[.x]], 0.95, na.rm = TRUE)
  
  df_comparison |>
    filter(.data[[.x]] > 0, .data[[.x]] <= upper_limit) |>
    ggplot(aes(x = .data[[.x]], fill = source)) +
    geom_density(alpha = 0.6, color = NA) +
    scale_fill_manual(values = color_palette, drop = FALSE) +
    labs(title = nutrient_labels[.x], x = NULL, y = "Density") +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 11, color = "azure4", face = "bold"),
      axis.title = element_text(size = 9, color = "azure4"),
      axis.text = element_text(size = 8, color = "azure4"),
      legend.position = "none",
      panel.grid.minor = element_blank()
    )
})

# Create a plot just to extract the legend
legend_plot <- df_comparison |>
  ggplot(aes(x = fe_total, fill = source)) +
  geom_density(alpha = 0.6) +
  scale_fill_manual(values = color_palette, drop = FALSE) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 11, color = "azure4")
  ) +
  guides(fill = guide_legend(override.aes = list(alpha = 0.6)))

# Extract legend using ggpubr (more reliable)
legend_grob <- ggpubr::get_legend(legend_plot)

# Arrange plots in grid
plot_grid <- plot_grid(plotlist = plot_list, ncol = 3, nrow = 2)

# Combine with legend
plot_grid(plot_grid, legend_grob, ncol = 1, rel_heights = c(1, 0.08))

# Cleanup auxiliary objects
rm(df_encovi_plot, df_sivesnu_plot, df_comparison, color_palette,
   nutrient_labels, plot_list, legend_plot, legend_grob, plot_grid)
```

```{r}
#| label: tbl-e3-distribution-comparison
#| tbl-cap: "Distribution comparison: SIVESNU predicted vs ENCOVI reference"
#| warning: FALSE

# Function to calculate KS test and statistics (without filtering zeros)
compare_distributions <- function(encovi_vec, sivesnu_vec, nutrient_name, component) {
  
  # Remove only NAs, keep zeros for median calculation
  encovi_clean <- encovi_vec[!is.na(encovi_vec)]
  sivesnu_clean <- sivesnu_vec[!is.na(sivesnu_vec)]
  
  # For KS test, filter positive values only (KS requires comparable distributions)
  encovi_ks <- encovi_clean[encovi_clean > 0]
  sivesnu_ks <- sivesnu_clean[sivesnu_clean > 0]
  
  # KS test
 ks_result <- ks.test(encovi_ks, sivesnu_ks)
  
  tibble(
    nutrient = nutrient_name,
    component = component,
    median_encovi = median(encovi_clean),
    median_sivesnu = median(sivesnu_clean),
    ratio = median(sivesnu_clean) / median(encovi_clean),
    ks_statistic = ks_result$statistic,
    p_value = ks_result$p.value
  )
}

# Calculate for all nutrients. Both sides are children 6-59 months: the
# reference was restricted in A1 and the source in D2, so this compares the
# calibrated SIVESNU children against the ENCOVI children they were mapped onto.
comparison_results <- bind_rows(
  # Maize components
  compare_distributions(df_encovi_reference$fe_maiz, df_qm_adjusted$fe_maiz_qm_adj, "Iron (mg/day)", "Maize"),
  compare_distributions(df_encovi_reference$zn_maiz, df_qm_adjusted$zn_maiz_qm_adj, "Zinc (mg/day)", "Maize"),
  compare_distributions(df_encovi_reference$prot_apro_maiz, df_qm_adjusted$prot_apro_maiz_qm_adj, "Digestible Protein (g/day)", "Maize"),
  compare_distributions(df_encovi_reference$lys_maiz, df_qm_adjusted$lys_maiz_qm_adj, "Lysine (mg/day)", "Maize"),
  compare_distributions(df_encovi_reference$trp_maiz, df_qm_adjusted$trp_maiz_qm_adj, "Tryptophan (mg/day)", "Maize"),
  compare_distributions(df_encovi_reference$kcal_maiz, df_qm_adjusted$kcal_maiz_qm_adj, "Energy (kcal/day)", "Maize"),
  # Non-maize components
  compare_distributions(df_encovi_reference$fe_no_maiz, df_qm_adjusted$fe_no_maiz_qm_adj, "Iron (mg/day)", "Non-Maize"),
  compare_distributions(df_encovi_reference$zn_no_maiz, df_qm_adjusted$zn_no_maiz_qm_adj, "Zinc (mg/day)", "Non-Maize"),
  compare_distributions(df_encovi_reference$prot_apro_no_maiz, df_qm_adjusted$prot_apro_no_maiz_qm_adj, "Digestible Protein (g/day)", "Non-Maize"),
  compare_distributions(df_encovi_reference$lys_no_maiz, df_qm_adjusted$lys_no_maiz_qm_adj, "Lysine (mg/day)", "Non-Maize"),
  compare_distributions(df_encovi_reference$trp_no_maiz, df_qm_adjusted$trp_no_maiz_qm_adj, "Tryptophan (mg/day)", "Non-Maize"),
  compare_distributions(df_encovi_reference$kcal_no_maiz, df_qm_adjusted$kcal_no_maiz_qm_adj, "Energy (kcal/day)", "Non-Maize")
)

# Pivot wider for cleaner table
comparison_wide <- comparison_results |>
  pivot_wider(
    id_cols = nutrient,
    names_from = component,
    values_from = c(median_encovi, median_sivesnu, ratio, ks_statistic, p_value),
    names_glue = "{component}_{.value}"
  ) |>
  select(
    nutrient,
    Maize_median_encovi, Maize_median_sivesnu, Maize_ratio, Maize_ks_statistic, Maize_p_value,
    `Non-Maize_median_encovi`, `Non-Maize_median_sivesnu`, `Non-Maize_ratio`, `Non-Maize_ks_statistic`, `Non-Maize_p_value`
  )

export(comparison_wide, here("01_data", "02_processed", "web_resources", "04_01_tbl_e3_distribution_comparison.rds"))

# Create GT table
comparison_wide |>
  mutate(
    # Format p-values as threshold categories
    Maize_p_formatted = case_when(
      Maize_p_value < 0.001 ~ "< 0.001",
      Maize_p_value < 0.01 ~ "< 0.01",
      Maize_p_value < 0.05 ~ "< 0.05",
      TRUE ~ sprintf("%.3f", Maize_p_value)
    ),
    `Non-Maize_p_formatted` = case_when(
      `Non-Maize_p_value` < 0.001 ~ "< 0.001",
      `Non-Maize_p_value` < 0.01 ~ "< 0.01",
      `Non-Maize_p_value` < 0.05 ~ "< 0.05",
      TRUE ~ sprintf("%.3f", `Non-Maize_p_value`)
    )
  ) |>
  select(-Maize_p_value, -`Non-Maize_p_value`) |>
  gt() |>
  tab_header(
    title = "Nutritional Distribution Calibration Assessment",
    subtitle = "Kolmogorov-Smirnov test comparing SIVESNU (predicted) vs ENCOVI (reference)"
  ) |>
  cols_label(
    nutrient = "Nutrient",
    Maize_median_encovi = "ENCOVI",
    Maize_median_sivesnu = "SIVESNU",
    Maize_ratio = "Ratio",
    Maize_ks_statistic = "KS",
    Maize_p_formatted = "p-value",
    `Non-Maize_median_encovi` = "ENCOVI",
    `Non-Maize_median_sivesnu` = "SIVESNU",
    `Non-Maize_ratio` = "Ratio",
    `Non-Maize_ks_statistic` = "KS",
    `Non-Maize_p_formatted` = "p-value"
  ) |>
  tab_spanner(label = "From Maize", columns = starts_with("Maize")) |>
  tab_spanner(label = "From Non-Maize", columns = starts_with("Non-Maize")) |>
  fmt_number(columns = contains("median"), decimals = 1) |>
  fmt_number(columns = contains("ratio"), decimals = 2) |>
  fmt_number(columns = contains("ks_statistic"), decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = Maize_ratio, rows = Maize_ratio >= 0.9 & Maize_ratio <= 1.1)
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = `Non-Maize_ratio`, rows = `Non-Maize_ratio` >= 0.9 & `Non-Maize_ratio` <= 1.1)
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = Maize_ks_statistic, rows = Maize_ks_statistic < 0.1)
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = `Non-Maize_ks_statistic`, rows = `Non-Maize_ks_statistic` < 0.1)
  ) |>
  tab_footnote(
    footnote = "Ratio = SIVESNU median / ENCOVI median. Values 0.9-1.1 indicate good calibration.",
    locations = cells_column_labels(columns = Maize_ratio)
  ) |>
  tab_footnote(
    footnote = "KS: Kolmogorov-Smirnov statistic (lower = better match). Green highlights: KS < 0.1 or Ratio 0.9-1.1.",
    locations = cells_column_labels(columns = Maize_ks_statistic)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(compare_distributions, comparison_results, comparison_wide)
```

---

# Phase F: Export

```{r}
#| label: f-save-output

# Save final dataset as parquet
write_parquet(
  df_sivesnu_nutritional,
  here("01_data", "02_processed", 
       "transfer", "04_01_sivesnu_nutritional_predicted.parquet")
)
```


# Notes

## Processing Summary

This script applies Module 3 transfer models to SIVESNU 2018 data to predict nutrient intake values for all individuals. The workflow includes:

1. **Data loading**: SIVESNU 2018 household and individual datasets merged
2. **Variable harmonization**: Recoding to match ENCOVI 2023 predictor structure
3. **Missing value imputation**: Random Forest-based multivariate imputation (via `impute_rf_parallel`) for predictors and FIES variables, with median/mode completion of any residual values
4. **Yeo-Johnson transformation**: Applied to continuous predictors (miembros_hogar, n_cuartos)
5. **Model prediction**: Module 3 gamma regression models applied with exp() back-transformation for log-link models. Plant percentage model uses identity link (no back-transformation).
6. **Target population filter**: children 6-59 months, applied before calibration so that source and reference are age-comparable
7. **FIES adjustment**: Food security-based fine-tuning (±2-5% based on FIES raw score)
8. **Quantile mapping**: Distribution calibration using filtered, age-matched ENCOVI reference distributions
9. **Maize alignment**: rank-based realignment of the maize block against the non-maize composite

## Methodology

**FIES for food security assessment**: Food security status is measured using the Food Insecurity Experience Scale (FIES), an 8-question instrument available in SIVESNU 2018. FIES follows FAO international methodology and provides comprehensive assessment of household food security across multiple dimensions (worry, quality, quantity, and severe deprivation).

**Population scope and age-matched calibration**: Prediction runs on all SIVESNU individuals, but the target population is filtered to children 6-59 months *before* the quantile mapping, and the ENCOVI reference is restricted to the same age range.

Quantile mapping transports the quantiles of a source distribution onto those of a reference, so the two describe comparable populations. Nutrient intake scales with body size, and the child medians sit well below the all-age medians in the ENCOVI reference, by a ratio that differs by nutrient and by source. Matching the age range on both sides keeps the level and the composition of any index built from the result on the child scale, and makes the section E3 validation a like-for-like comparison.

The mapping is estimated over the pooled age-matched range.

**Quantile mapping reference filtering**: Reference distributions from ENCOVI are filtered to remove zeros and Tukey outliers (Q3 + 1.5*IQR) to match the preprocessing applied during model training. Note that this trim is applied to the *reference* only; script 04_08 applies its own plausibility rules to the synthetic population so that the two ends of the pipeline describe comparable supports.

**Raw prediction preservation**: The exported dataset carries two parallel blocks of the twelve source-specific nutrient variables. The `_raw` block holds the transfer-model output before the FIES adjustment, the quantile mapping and the maize-source alignment; the `_adjusted` block holds the calibrated values consumed by the rest of the pipeline. Both blocks use the naming convention adopted from 04_02 onwards (`fe_mg_maiz`, `prot_g_maiz`, `ene_kcal_maiz`, ...), so each raw column pairs one-to-one with its calibrated counterpart. The `_raw` block is carried unchanged through 04_02 to 04_08 and is never used in any calculation; it exists so the effect of the calibration can be audited at any stage of the module.

## Limitations

Individual intake in ENCOVI was allocated from household consumption over a 15-day recall using an adult-male-equivalent scale, so the age gradient of the reference distribution partly reflects that allocation rule.
