---
title: "Calculate FIES Rasch Model Score"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script calculates the Food Insecurity Experience Scale (FIES) score using the Rasch model methodology recommended by FAO. The Rasch model transforms the ordinal raw score into an interval-scale measure that accounts for item severity and response patterns.

1. Load SIVESNU data with bioavailability calculations (from Script 04_03)
2. Verify FIES variables coding (Yes/No responses)
3. Prepare binary matrix for Rasch model (0/1 coding)
4. Fit Rasch model using FAO's RM.weights package
5. Extract individual food insecurity scores
6. Validate results against raw scores
7. Save processed dataset with Rasch-based FIES score

**Key methodological decisions**:

-   **Rasch model implementation**: Uses FAO's `RM.weights` package, specifically designed for FIES analysis with survey weights support.

-   **Item ordering**: FIES items follow increasing severity (worry → food deprivation), which is captured in the Rasch item difficulty parameters.

-   **Score interpretation**: Higher Rasch scores indicate greater food insecurity, calibrated to the FAO global reference scale.

**FIES questions** (C001-C008 in survey):

1. Worried about not having enough food
2. Unable to eat healthy/nutritious food
3. Ate only a few kinds of foods (low variety)
4. Had to skip a meal
5. Ate less than thought should
6. Household ran out of food
7. Felt hungry but did not eat
8. Went without eating for a whole day

# Setup

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  RM.weights,         # FAO Rasch Model for FIES
  survey,             # Complex survey methodology
  srvyr,              # Tidy wrapper for survey package
  gt,                 # Table formatting and presentation
  tidyverse           # General purpose data manipulation
)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter
drop   <- base::drop

# Set display options
options(scipen = 999, digits = 4)

source(here("02_code", "00_functions", "00_functions.R"))

# transfer/
ensure_output_dir(here("01_data", "02_processed", "transfer"))

# web_resources/
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

# Phase A: Data Loading and Verification

## A1. Load SIVESNU Data with Bioavailability

Load SIVESNU 2018 children's data with bioavailability calculations from Script 04_03.

```{r}
#| label: a1-load-data

df_input <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "04_03_sivesnu_bioavailable.parquet")
)
```

## A2. Verify FIES Item Response Distribution

Examine the current state of FIES variables to confirm proper coding before Rasch model application.

```{r}
#| label: tbl-a2-fies-distribution
#| tbl-cap: "FIES Item Response Distribution"

# Define FIES item descriptions (FAO standard ordering by severity)
fies_descriptions <- tibble(
  item = paste0("fies", 1:8),
  code = paste0("C00", 1:8),
  description = c(
    "Worried about not having enough food",
    "Unable to eat healthy/nutritious food",
    "Ate only a few kinds of foods",
    "Had to skip a meal",
    "Ate less than should",
    "Household ran out of food",
    "Felt hungry but did not eat",
    "Went without eating for a whole day"
  ),
  severity = c(
    "Mild", "Mild", "Mild",
    "Moderate", "Moderate", "Moderate",
    "Severe", "Severe"
  )
)

# Calculate response distribution for each FIES item
fies_summary <- df_input |>
  select(fies1:fies8) |>
  pivot_longer(everything(), names_to = "item", values_to = "response") |>
  summarise(
    n_yes = sum(response == "Yes", na.rm = TRUE),
    n_no = sum(response == "No", na.rm = TRUE),
    n_na = sum(is.na(response)),
    pct_yes = n_yes / (n_yes + n_no) * 100,
    .by = item
  ) |>
  left_join(fies_descriptions, by = "item") |>
  arrange(item) |>
  select(code, description, severity, n_yes, n_no, n_na, pct_yes)

# Export data for web resources
export(fies_summary, here("01_data", "02_processed", "web_resources", "04_04_tbl_a2_fies_distribution.rds"))

# Display as GT table
fies_summary |>
  gt() |>
  tab_header(
    title = "FIES Item Response Distribution",
    subtitle = "Food Insecurity Experience Scale (8 items)"
  ) |>
  cols_label(
    code = "Code",
    description = "Question",
    severity = "Severity",
    n_yes = "Yes",
    n_no = "No",
    n_na = "NA",
    pct_yes = "% Yes"
  ) |>
  fmt_number(columns = c(n_yes, n_no, n_na), decimals = 0) |>
  fmt_number(columns = pct_yes, decimals = 1) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(rows = severity == "Mild", columns = severity)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFE0B2"),
    locations = cells_body(rows = severity == "Moderate", columns = severity)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = severity == "Severe", columns = severity)
  ) |>
  tab_source_note(
    source_note = "Source: SIVESNU 2018. Items ordered by increasing severity per FAO FIES methodology."
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Clean auxiliary objects
rm(fies_descriptions, fies_summary)
```

---

# Phase B: Rasch Model Estimation

## B1. Fit Rasch Model

Convert FIES responses to binary format (0/1) and apply FAO's Rasch model implementation using the `RM.weights` package.

```{r}
#| label: b1-fit-rasch-model

# Define FIES item descriptions for later use
fies_descriptions <- tibble(
 item = paste0("fies", 1:8),
  code = paste0("C00", 1:8),
  description = c(
    "Worried about not having enough food",
    "Unable to eat healthy/nutritious food",
    "Ate only a few kinds of foods",
    "Had to skip a meal",
    "Ate less than should",
    "Household ran out of food",
    "Felt hungry but did not eat",
    "Went without eating for a whole day"
  ),
  severity = c("Mild", "Mild", "Mild", "Moderate", "Moderate", "Moderate", "Severe", "Severe")
)

# Create binary data frame for Rasch model (0 = No, 1 = Yes)
# Note: RM.w expects a data.frame with integer 0/1 values
fies_df <- df_input |>
  select(fies1:fies8) |>
  mutate(across(everything(), ~ if_else(.x == "Yes", 1L, 0L))) |>
  as.data.frame()

# Extract weights as numeric vector
fies_weights <- df_input$pesonino

# Fit weighted Rasch model (following FAO example pattern)
rasch_model <- RM.w(fies_df, fies_weights)
```

## B2. Item Difficulty Parameters

Extract and display item difficulty (severity) parameters from the Rasch model. Higher values indicate more severe food insecurity experiences.

```{r}
#| label: tbl-b2-item-parameters
#| tbl-cap: "FIES Item Difficulty Parameters (Rasch Model)"

# Extract item parameters
item_params <- tibble(
  item = paste0("fies", 1:8),
  difficulty = rasch_model$b
) |>
  left_join(fies_descriptions, by = "item") |>
  arrange(difficulty) |>
  mutate(rank = row_number()) |>
  select(rank, code, description, severity, difficulty)

# Export data for web resources
export(item_params, here("01_data", "02_processed", "web_resources", "04_04_tbl_b2_item_parameters.rds"))

# Display as GT table
item_params |>
  gt() |>
  tab_header(
    title = "FIES Item Difficulty Parameters",
    subtitle = "Rasch model estimates (ordered by difficulty)"
  ) |>
  cols_label(
    rank = "Rank",
    code = "Code",
    description = "Question",
    severity = "Severity",
    difficulty = "Difficulty (β)"
  ) |>
  fmt_number(columns = difficulty, decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(rows = severity == "Mild", columns = severity)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFE0B2"),
    locations = cells_body(rows = severity == "Moderate", columns = severity)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = severity == "Severe", columns = severity)
  ) |>
  tab_source_note(
    source_note = "Higher difficulty = more severe food insecurity. Negative values = more common experiences."
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(item_params)
```

## B3. Raw Score Severity Parameters

The Rasch model transforms ordinal raw scores (0-8) into interval-scale severity parameters. These parameters are used to assign individual food insecurity scores.

```{r}
#| label: tbl-b3-raw-score-parameters
#| tbl-cap: "Severity Parameters by Raw Score"

# Extract raw score parameters
raw_score_params <- tibble(
  raw_score = 0:8,
  severity = rasch_model$a,
  se = rasch_model$se.a
) |>
  mutate(
    classification = case_when(
      raw_score == 0 ~ "Food Secure",
      raw_score <= 3 ~ "Mild FI",
      raw_score <= 6 ~ "Moderate FI",
      raw_score >= 7 ~ "Severe FI"
    )
  )

# Export data for web resources
export(raw_score_params, here("01_data", "02_processed", "web_resources", "04_04_tbl_b3_raw_score_parameters.rds"))

raw_score_params |>
  gt() |>
  tab_header(
    title = "FIES Severity Parameters by Raw Score",
    subtitle = "Rasch model transformation"
  ) |>
  cols_label(
    raw_score = "Raw Score",
    severity = "Severity (θ)",
    se = "Std. Error",
    classification = "Classification"
  ) |>
  fmt_number(columns = c(severity, se), decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = classification == "Food Secure", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(rows = classification == "Mild FI", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFE0B2"),
    locations = cells_body(rows = classification == "Moderate FI", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = classification == "Severe FI", columns = classification)
  ) |>
  tab_source_note(
    source_note = "FI = Food Insecurity. Classification follows FAO thresholds."
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

---

# Phase C: Score Assignment and Validation

## C1. Assign Individual Rasch Scores

Map each individual's raw score to the corresponding Rasch severity parameter.

```{r}
#| label: c1-assign-individual-scores

# Create lookup table for raw score to Rasch severity mapping
score_lookup <- tibble(
  fies_raw_score = 0:8,
  fies_rasch_score = rasch_model$a
)

# Assign Rasch scores to individuals
df_rasch <- df_input |>
  left_join(score_lookup, by = "fies_raw_score")
```

## C2. Validation - Raw Score vs Rasch Score

Compare the distribution of raw scores (simple sum) with Rasch-transformed scores.

```{r}
#| label: tbl-c2-score-comparison
#| tbl-cap: "Distribution of FIES Scores: Raw vs Rasch"

# Create survey design
svy_rasch <- df_rasch |>
  as_survey_design(
    ids = conglomerado,
    weights = pesonino
  )

# Summarize by raw score
score_comparison <- svy_rasch |>
  group_by(fies_raw_score) |>
  summarise(
    n = n(),
    pct = survey_mean(vartype = NULL) * 100,
    mean_rasch = survey_mean(fies_rasch_score, vartype = NULL)
  ) |>
  left_join(
    raw_score_params |> select(raw_score, classification),
    by = c("fies_raw_score" = "raw_score")
  )

# Export data for web resources
export(score_comparison, here("01_data", "02_processed", "web_resources", "04_04_tbl_c2_score_comparison.rds"))

score_comparison |>
  gt() |>
  tab_header(
    title = "FIES Score Distribution",
    subtitle = "Raw scores mapped to Rasch severity parameters"
  ) |>
  cols_label(
    fies_raw_score = "Raw Score",
    n = "N",
    pct = "% (weighted)",
    mean_rasch = "Rasch Score (θ)",
    classification = "Classification"
  ) |>
  fmt_number(columns = pct, decimals = 1) |>
  fmt_number(columns = mean_rasch, decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = classification == "Food Secure", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(rows = classification == "Mild FI", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFE0B2"),
    locations = cells_body(rows = classification == "Moderate FI", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = classification == "Severe FI", columns = classification)
  ) |>
  tab_source_note(
    source_note = "Rasch scores provide interval-scale measurement; raw scores are ordinal."
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )

# Clean auxiliary objects
rm(svy_rasch, score_comparison)
```

## C3. Model Fit Statistics

Assess Rasch model quality through reliability and item fit statistics.

```{r}
#| label: tbl-c3-model-fit
#| tbl-cap: "Rasch Model Fit Statistics"

# Compile fit statistics
fit_stats <- tibble(
  statistic = c(
    "Rasch Reliability (observed)",
    "Rasch Reliability (flat)",
    "Complete Cases",
    "Convergence"
  ),
  value = c(
    sprintf("%.3f", rasch_model$reliab),
    sprintf("%.3f", rasch_model$reliab.fl),
    as.character(rasch_model$n.compl),
    rasch_model$converged
  )
)

# Export data for web resources
export(fit_stats, here("01_data", "02_processed", "web_resources", "04_04_tbl_c3_model_fit.rds"))

fit_stats |>
  gt() |>
  tab_header(
    title = "Rasch Model Fit Statistics"
  ) |>
  cols_label(
    statistic = "Statistic",
    value = "Value"
  ) |>
  tab_source_note(
    source_note = "Reliability > 0.7 indicates acceptable measurement precision."
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

## C4. Distribution Visualization

```{r}
#| label: fig-c4-rasch-distribution
#| fig-cap: "Distribution of FIES Rasch Scores"
#| fig-height: 7
#| fig-width: 10

# Prepare classification for coloring
df_plot <- df_rasch |>
  mutate(
    classification = factor(
      case_when(
        fies_raw_score == 0 ~ "Food Secure",
        fies_raw_score <= 3 ~ "Mild FI",
        fies_raw_score <= 6 ~ "Moderate FI",
        fies_raw_score >= 7 ~ "Severe FI"
      ),
      levels = c("Food Secure", "Mild FI", "Moderate FI", "Severe FI")
    )
  )

# Export data for web resources
export(df_plot, here("01_data", "02_processed", "web_resources", "04_04_fig_c4_rasch_distribution.rds"))

# Define color palette matching table colors
fi_colors <- c(
  "Food Secure" = "#D4EDDA",
  "Mild FI" = "#FFF3CD",
  "Moderate FI" = "#FFE0B2",
  "Severe FI" = "#FFCDD2"
)

# Histogram of Rasch scores
ggplot(df_plot, aes(x = fies_rasch_score, fill = classification)) +
  geom_histogram(bins = 9, color = "white", linewidth = 0.5) +
  scale_fill_manual(values = fi_colors, name = "Classification") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  annotate("text", x = -0.3, y = Inf, label = "← Less severe", 
           hjust = 1, vjust = 2, size = 3, color = "gray40") +
  annotate("text", x = 0.3, y = Inf, label = "More severe →", 
           hjust = 0, vjust = 2, size = 3, color = "gray40") +
  labs(
    title = "Distribution of Food Insecurity (Rasch Scale)",
    subtitle = "Children 6-59 months, SIVESNU 2018",
    x = "FIES Rasch Score (θ)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 13, color = "gray30", face = "bold"),
    plot.subtitle = element_text(size = 11, color = "gray50"),
    axis.title = element_text(size = 10, color = "gray40"),
    legend.position = "bottom"
  )

# Clean auxiliary object
rm(df_plot, fi_colors)
```

---

# Phase D: Export

```{r}
#| label: d-save-output

# Prepare final dataset
df_output <- df_rasch |>
  select(
    # === Identifiers ===
    nboletahogar, conglomerado, pesonino,
    
    # === Demographics ===
    parentesco, sexo, edad_an, edad_mes, edad_grupos,
    
    # === Geographic ===
    departamento, area,
    
    # === Household characteristics ===
    miembros_hogar_original, n_cuartos_original,
    grado_estudios, grado_estudios_hogar,
    propiedad, tipo_vivienda,
    material_paredes, material_techo, material_piso,
    
    # === Basic services ===
    cocina, tipo_sanitario, fuente_agua, recoleccion_basura,
    
    # === Household assets ===
    electricidad, televisor, telefonia_fija, telefonia_celular, computadora,
    
    # === Food security (FIES) ===
    fies1, fies2, fies3, fies4, fies5, fies6, fies7, fies8,
    fies_raw_score,
    fies_rasch_score,
    
    # === Nutritional intake by source ===
    fe_mg_maiz, fe_mg_nomaiz,
    zn_mg_maiz, zn_mg_nomaiz,
    prot_g_maiz, prot_g_nomaiz,
    lys_mg_maiz, lys_mg_nomaiz,
    trp_mg_maiz, trp_mg_nomaiz,
    ene_kcal_maiz, ene_kcal_nomaiz,

    # === Raw transfer-model predictions (preserved from 04_01) ===
    # Pre-FIES, pre-quantile-mapping, pre-alignment. Diagnostics only.
    fe_mg_maiz_raw, fe_mg_nomaiz_raw,
    zn_mg_maiz_raw, zn_mg_nomaiz_raw,
    prot_g_maiz_raw, prot_g_nomaiz_raw,
    lys_mg_maiz_raw, lys_mg_nomaiz_raw,
    trp_mg_maiz_raw, trp_mg_nomaiz_raw,
    ene_kcal_maiz_raw, ene_kcal_nomaiz_raw,

    # === Dietary composition ===
    per_vegetales,
    
    # === Bioavailable micronutrients ===
    zn_mg_total, taz_mg, zn_fractional_absorption,
    fe_mg_total, fe_absorption_factor, fe_absorbed_mg,
    
    # === Total intake (other nutrients) ===
    prot_g_total, lys_mg_total, trp_mg_total, ene_kcal_total,
    
    # === Chispitas supplementation ===
    receives_chispitas, adherence_factor_final,
    fe_chispitas_effective_mg, zn_chispitas_effective_mg
  )

# Save final dataset
write_parquet(
  df_output,
  here("01_data", "02_processed", 
       "transfer", "04_04_sivesnu_fies_rasch.parquet")
)

# Clean auxiliary objects
rm(fies_df, fies_weights, fies_descriptions, rasch_model, df_rasch, fit_stats)
```

# Notes

## Methodology

### Rasch Model for FIES

The Rasch model is a one-parameter Item Response Theory (IRT) model that transforms ordinal raw scores into interval-scale measurements. Key properties:

-   **Item difficulty parameters (β)**: Represent the severity level at which each item is typically reported. Items with negative β are reported by most respondents (less severe); items with positive β are reported only by severely food insecure respondents.

-   **Person severity parameters (θ)**: Represent individual food insecurity levels on the same scale as item difficulties. Higher θ indicates greater food insecurity.

-   **Interval-scale measurement**: Rasch scores have equal intervals, making differences comparable across the scale.

### Weighted Estimation

Survey weights were applied during model estimation so that item parameters reflect the Guatemalan population. This is appropriate because:

1. SIVESNU uses complex sampling with unequal selection probabilities
2. Population-representative item parameters enable valid comparisons with FAO global reference scales
3. Individual scores are assigned using population-calibrated parameters

### Classification Thresholds

FAO classification based on raw scores:

| Raw Score | Classification |
|-----------|----------------|
| 0         | Food Secure    |
| 1-3       | Mild FI        |
| 4-6       | Moderate FI    |
| 7-8       | Severe FI      |

## Limitations

1. **Extreme scores**: Raw scores 0 and 8 require assumptions for parameter estimation (handled by RM.weights default settings).

2. **Local independence**: Rasch model assumes conditional independence between items given the latent trait; residual correlations may indicate violation.

3. **Unidimensionality**: FIES assumes a single food insecurity dimension; model fit statistics should be checked.

## References

-   FAO (2016). Methods for estimating comparable rates of food insecurity experienced by adults throughout the world. Rome: Food and Agriculture Organization.

-   Cafiero C, Viviani S, Nord M (2018). Food security measurement in a global context: The Food Insecurity Experience Scale. *Measurement* 116:146-152.

-   Nord M (2014). Introduction to Item Response Theory applied to Food Security Measurement. FAO Technical Paper.


