---
title: "Data Sources and Integration"
subtitle: "Module 1: Nutrient Intakes"
description-meta: "Integrating ENCOVI 2023 consumption records with INCAP food composition tables and PDCAAS protein quality scores into one nutrient intake base."
---

## Overview

In the Guatemalan data context, estimating nutrient intake from household survey
data requires integrating multiple data sources: consumption records from a
nationally representative survey, food composition tables with nutrient content,
and expert-provided conversions to translate local measurement units into
standardized weights.

This module loads and integrates three primary sources to create a unified
analytical dataset:

1. **ENCOVI 2023** — Guatemala's National Survey of Living Conditions provides
   household food consumption records, demographic characteristics, and
   socioeconomic indicators for over 10,000 households.

2. **INCAP 2014** — The Instituto de Nutrición de Centro América y Panamá
   (INCAP) Food Composition Table for Central America contains
   nutrient content (energy, protein, iron, zinc) and amino acid profiles
   (lysine, tryptophan, methionine, cysteine, threonine) for foods common in
   the Guatemalan diet.

3. **Expert protein-quality and conversion data** — Two contributions that
   complete the INCAP base:

   - **Protein quality scores (PDCAAS)** — Amino acid profiles and Protein
     Digestibility Corrected Amino Acid Scores compiled from the literature
     by nutrition expert María José Soto Méndez (FINUT) — principally
     [@Boye2012_PDCAAS], [@SuarezLopez2006_PDCAAS] and FAO amino acid data,
     following the FAO/WHO/UNU reference methodology [@WHO2007_ProteinAminoAcid].
     Where no literature value was available for a food, the score was derived
     from comparable food groups. These scores are delivered as an additional
     column in the composition table for direct application in the framework.
   - **Unit conversion factors** — Weights in grams for non-standard units
     (bags, packages, cups) compiled through on-site market research in
     Guatemalan supermarkets and markets, addressing the gap left by the
     original composition tables where most entries lacked complete gram
     equivalences.

The integration process standardizes food quantities to grams and links
consumption records to nutritional composition. The distribution of household
intake to individuals is documented in [Intake Calculation](01_03_intake_calculation.qmd).

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
pacman::p_unload(all = TRUE)

# Load required packages (using pacman for automatic installation)
pacman::p_load(
  rio,            # Robust import/export for various formats
  here,           # Robust file path management
  tidyverse,      # General purpose data manipulation
  arrow,          # Parquet file handling
  gt              # Table formatting and presentation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script.
tbl_data_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_01_tbl_c1_data_summary.rds"),
  trust = TRUE
)
```


## ENCOVI 2023: Household Survey Data

ENCOVI (Encuesta Nacional de Condiciones de Vida) is Guatemala's primary source
for household-level socioeconomic data. The 2023 survey includes a detailed food
consumption module where households report foods purchased and obtained during
the previous 15 days, along with quantities and measurement units.

Three datasets are loaded from ENCOVI:

- **Food consumption** — Individual food items with quantities, units, and
  acquisition method (purchased vs. obtained without purchase).
- **Household characteristics** — Housing conditions, basic services, and assets
  used as socioeconomic predictors in transfer models.
- **Individual demographics** — Age, sex, education, and household relationship
  for each person, enabling individual-level intake calculations.


## INCAP 2014: Food Composition Tables

INCAP maintains the reference food composition database for the region. The
2014 edition provides
nutrient content per 100g for foods commonly consumed in Guatemala.

The integrated composition database combines three provenance layers, each
documented at the value level:

- **INCAP composition base** — Energy, protein, iron, zinc, and the
  edible-portion factor, transcribed from the INCAP food composition report
  for Central American foods.
- **Amino acid profiles** — Essential amino acids including lysine and
  tryptophan, critical for assessing protein quality in maize-based diets,
  compiled from the literature by María José Soto Méndez (FINUT).
- **PDCAAS scores** — Protein Digestibility Corrected Amino Acid Scores
  indicating protein quality relative to human requirements
  [@Boye2012_PDCAAS]. Literature-based scores were compiled by FINUT; where no
  literature value was available, the score was derived from comparable food
  groups [@SuarezLopez2006_PDCAAS].

The composition tables are keyed by food code (`ali`) and linked to the ENCOVI
survey food code (`codine`), which is the operative join key to consumption
records. All nutrient values are expressed on a per-gram basis to
facilitate calculations with varying portion sizes.

::: {.callout-tip}
## Protein Quality Score Assignment
Each food in the composition database carries a PDCAAS value assigned through
one of three routes:

- **Literature-based and derived scores** — Most foods receive a score
  compiled by María José Soto Méndez (FINUT) from the literature
  (Boye et al., 2012; Suárez López et al., 2006), or derived from comparable food
  groups where no literature value was available.
- **Negligible protein contribution** — Foods that provide no meaningful
  dietary protein (sugars, fats and oils, salt, soft drinks) are assigned a
  conservative value of 0.1, reflecting their negligible contribution to
  protein quality.
- **Insufficient composition data** — Where composition data are too sparse to
  assign a value, the same conservative 0.1 default is applied so that protein
  quality estimates are not inflated for foods with undetermined digestibility.
:::

```{r}
#| label: tbl-pdcaas-coverage
#| tbl-cap: "PDCAAS assignment across the composition database"

# Read from the composition table: the assignment routes are reported only here
# Deduplicate on `codine`, the unique operative key, since `ali` repeats across rows
incap_composition <- read_parquet(
  here("..", "01_data", "03_external", "incap_food_composition.parquet")
) |>
  select(codine, pdcaas) |>
  distinct()

n_total <- nrow(incap_composition)
n_default <- sum(is.na(incap_composition$pdcaas))
n_negligible <- sum(incap_composition$pdcaas == 0.1, na.rm = TRUE)
n_scored <- n_total - n_default - n_negligible

tibble(
  Category = c(
    "Literature-based or derived score",
    "Negligible protein contribution (0.1)",
    "Insufficient composition data (0.1 default)",
    "Total"
  ),
  N = c(n_scored, n_negligible, n_default, n_total),
  Percentage = c(
    n_scored / n_total,
    n_negligible / n_total,
    n_default / n_total,
    1
  )
) |>
  gt() |>
  fmt_percent(columns = Percentage, decimals = 1) |>
  cols_align(align = "right", columns = c(N, Percentage)) |>
  tab_options(
    table.font.size = px(14),
    heading.align = "left"
  )
```


## Food Unit Equivalences

Household surveys record food quantities in local measurement units — bags
(*bolsas*), packages (*paquetes*), trays (*bandejas*), dozens (*docenas*) — that
vary by food type and regional packaging conventions. Converting these to grams
requires local expertise.

The food equivalence table was compiled by María José Soto (FINUT) through
on-site market research in Guatemala during 2024. The table provides:

| Data Element | Purpose |
|--------------|---------|
| **Gram conversions** | Weight in grams for each food-unit combination |
| **Plant-based classification** | Identifies foods containing phytates that affect mineral bioavailability |
| **Survey frequency** | Number of times each unit appears in ENCOVI responses |

The conversions reflect the retail packaging that survey respondents report
against. A *bolsa* (bag) of rice, for example, has a different weight than a
*bolsa* of beans, and both differ from standardized INCAP reference weights.


## Household Socioeconomic Variables

Household characteristics from ENCOVI are recoded into interpretable factor
variables. These serve as predictors in the transfer models that estimate
nutrient intake for SIVESNU children (source of anthropometric data) based on
their household conditions.

The variables capture multiple dimensions of socioeconomic status:

- **Housing quality** — Wall, roof, and floor materials.
- **Basic services** — Water source, sanitation, electricity.
- **Assets** — Television, telephone, computer ownership.
- **Geographic context** — Department and urban/rural classification.

Individual nutrient intake is derived from household intake using Adult Male
Equivalent (AME) factors. See
[Intake Calculation](01_03_intake_calculation.qmd) for the methodology and
application.


## Summary

### ENCOVI 2023 Survey Data

```{r}
#| label: tbl-summary-encovi

tbl_data_summary$encovi_summary |>
  gt() |>
  tab_header(
    title = "ENCOVI 2023 Survey Data",
    subtitle = "Loaded datasets overview"
  ) |>
  fmt_number(
    columns = c(Observations, Variables),
    decimals = 0,
    use_seps = TRUE
  ) |>
  cols_label(
    Dataset = "Dataset",
    Observations = "N",
    Variables = "Variables"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  )
```

### INCAP 2014 Food Composition Database

```{r}
#| label: tbl-summary-incap

tbl_data_summary$incap_summary |>
  gt() |>
  tab_header(
    title = "INCAP 2014 Food Composition Database",
    subtitle = "Integrated composition tables with amino acids and PDCAAS"
  ) |>
  cols_label(
    Component = "Component",
    Status = "Available"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  )
```

### External Expert Data

```{r}
#| label: tbl-summary-external

tbl_data_summary$external_summary |>
  gt() |>
  tab_header(
    title = "External Expert Data",
    subtitle = "María José Soto (FINUT) contributions"
  ) |>
  fmt_number(
    columns = Count,
    decimals = 0,
    use_seps = TRUE
  ) |>
  cols_label(
    Dataset = "Dataset",
    Count = "N",
    Description = "Description"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  )
```

### Processed Base Data

```{r}
#| label: tbl-summary-processed

tbl_data_summary$processed_summary |>
  gt() |>
  tab_header(
    title = "Processed Base Data",
    subtitle = "Cleaned and prepared datasets for nutrient intake calculations"
  ) |>
  fmt_number(
    columns = Count,
    decimals = 0,
    use_seps = TRUE
  ) |>
  sub_missing(
    columns = Percentage,
    missing_text = "—"
  ) |>
  cols_label(
    Dataset = "Dataset",
    Count = "N",
    Percentage = "%"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  )
```
