---
title: "Estimate Age in Months"
subtitle: "Module 1: Nutrient Intakes"
---

## Why Estimate Age in Months?

ENCOVI 2023 records age as integer years, which creates two methodological
limitations for the biofortification impact analysis:

1. **Loss of the 6-11 month population** — Children reported as `edad == 0`
   cannot be separated into those under 6 months (excluded due to exclusive
   breastfeeding) and those aged 6-11 months (part of the study population).
   Without month-level resolution, as shown in
   @tbl-c1-age-distribution-pre, the entire age-0 cohort must be excluded,
   losing the children aged 6-11 months who should be included.

2. **Imprecision in anthropometric modeling** — WHO LMS growth reference tables
   require age in months. Using integer years introduces up to ±5.5 months of
   error in a range where growth curves have steep gradients, distorting
   height-for-age Z-score calculations.

This module resolves both limitations by generating a synthetic `edad_meses`
variable using empirical birth distributions from the Guatemala National
Statistics Institute (INE) vital statistics registry (2019-2023)
[@INE2024_NacimientosVitales]. Each birth-year cohort is matched to the 
corresponding age-in-years group. Month assignment uses a greedy quota-filling 
algorithm that distributes children to birth months so that the survey-weighted 
distribution matches the observed INE proportions by department and sex.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
pacman::p_unload(all = TRUE)

# Load required packages (using pacman for automatic installation)
pacman::p_load(
  rio,            # Robust import/export for various formats
  here,           # Robust file path management
  tidyverse,      # General purpose data manipulation
  gt,             # Table formatting and presentation
  scales          # Number formatting (percent_format)
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script.
tbl_a1_file_metadata <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_tbl_a1_file_metadata.rds"),
  trust = TRUE
)

tbl_a3_loading_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_tbl_a3_loading_summary.rds"),
  trust = TRUE
)

tbl_b1_proportion_validation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_tbl_b1_proportion_validation.rds"),
  trust = TRUE
)

fig_b2_seasonal_patterns <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_fig_b2_seasonal_patterns.rds"),
  trust = TRUE
)

fig_b2_sex_differential <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_fig_b2_sex_differential.rds"),
  trust = TRUE
)

tbl_c1_age_distribution_pre <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_tbl_c1_age_distribution_pre.rds"),
  trust = TRUE
)

tbl_c2_cohort_mapping <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_tbl_c2_cohort_mapping.rds"),
  trust = TRUE
)

tbl_d3_assignment_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_tbl_d3_assignment_summary.rds"),
  trust = TRUE
)

fig_e1_month_distribution_by_cohort <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_fig_e1_month_distribution_by_cohort.rds"),
  trust = TRUE
)

fig_e2_edad_meses_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_fig_e2_edad_meses_distribution.rds"),
  trust = TRUE
)

fig_e3_department_age_boxplot <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_fig_e3_department_age_boxplot.rds"),
  trust = TRUE
)

tbl_e4_weighted_deviation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_tbl_e4_weighted_deviation.rds"),
  trust = TRUE
)

tbl_e5_sibling_separation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_03_tbl_e5_sibling_separation.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Extract concrete figures used in narrative text.
n_recovered <- tbl_d3_assignment_summary |>
  filter(Category == "Children 6-11 months (synthetic assignment)") |>
  pull(N)

n_children_0_4 <- tbl_d3_assignment_summary |>
  filter(Category == "Children 0-4 years (synthetic assignment)") |>
  pull(N)

mean_dev_pp_max <- tbl_e4_weighted_deviation |>
  summarise(max_mean = max(mean_abs_dev) * 100) |>
  pull(max_mean) |>
  round(1)

twins_row <- tbl_e5_sibling_separation |>
  filter(separation_class == "Twins (0 months)")
n_twins <- twins_row |> pull(n_pairs)
pct_twins <- (twins_row |> pull(pct)) * 100
```


## Data Sources

The INE publishes annual vital statistics with births disaggregated by month
of occurrence, department of maternal residence, and sex. Five years of data
(2019-2023) are loaded and standardized to construct department- and
sex-specific birth seasonality profiles.

Each birth year maps to one ENCOVI age cohort: children aged 0 years were
born in 2023, aged 1 in 2022, and so on through age 4 (born 2019). This
one-to-one mapping allows the INE distributions to serve as population-level
reference weights for synthetic month assignment within each cohort.

::: {.callout-note}
## Data Structure
The unified registry consolidates five annual INE releases (2019-2023) into a
single table of births by year, month, department, and sex. Records classified
as "Extranjero" or "Ignorado" are excluded from proportion calculations, as
ENCOVI assigns all children to one of the 22 departments.
:::

```{r}
#| label: tbl-a1-file-metadata
#| tbl-cap: "INE birth registry file metadata"

tbl_a1_file_metadata |>
  gt() |>
  tab_header(
    title = "INE Birth Registry Source",
    subtitle = "Guatemala vital statistics, births by month, department, and sex"
  ) |>
  cols_label(
    year = "Year",
    source = "Source File",
    coverage = "Coverage"
  ) |>
  cols_align(align = "center", columns = year) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

The table above documents the documentary coverage of the unified birth
registry: five birth years (2019-2023), each mapping to one ENCOVI age cohort,
consolidated into a single source providing births by month, department, and sex.

```{r}
#| label: tbl-a3-loading-summary
#| tbl-cap: "Birth data loading summary by year"

tbl_a3_loading_summary |>
  gt() |>
  tab_header(
    title = "Birth Registry Data Summary",
    subtitle = "INE Guatemala vital statistics (2019-2023)"
  ) |>
  cols_label(
    birth_year = "Year",
    n_records = "Records",
    n_departments = "Departments",
    n_months = "Months",
    total_births = "Total Births"
  ) |>
  fmt_number(columns = c(n_records, total_births), decimals = 0, use_seps = TRUE) |>
  cols_align(align = "center") |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = n_departments, rows = n_departments != 22)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = n_months, rows = n_months != 12)
  ) |>
  tab_footnote(
    footnote = "Expected: 22 departments x 12 months x 2 sexes = 528 records per year.",
    locations = cells_column_labels(columns = n_records)
  ) |>
  tab_footnote(
    footnote = "Red cells indicate deviations from expected values.",
    locations = cells_title(groups = "subtitle")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Data Completeness
All five years contain exactly 528 records (22 departments × 12 months ×
2 sexes), matching the expected record grid for the vital statistics registry.
Total annual births range from approximately 331,000 to 366,000 across the
period.
:::


## Birth Seasonality Patterns

Before using INE birth proportions for month assignment, we verify that
birth timing carries meaningful departmental and sex-specific variation
beyond a uniform distribution. If births were uniformly distributed, each
month would contain 8.33% (1/12) of annual births.

```{r}
#| label: tbl-b1-proportion-validation
#| tbl-cap: "Proportion sum validation by stratum"

tbl_b1_proportion_validation |>
  gt() |>
  tab_header(
    title = "Birth Proportion Validation",
    subtitle = "Sum of monthly proportions by year (expected = 1.000 per stratum)"
  ) |>
  cols_label(
    birth_year = "Year",
    n_strata = "Strata (dept x sex)",
    min_sum = "Min Sum",
    max_sum = "Max Sum",
    mean_sum = "Mean Sum",
    all_valid = "All Valid"
  ) |>
  fmt_number(columns = c(min_sum, max_sum, mean_sum), decimals = 4) |>
  cols_align(align = "center") |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = all_valid, rows = all_valid == TRUE)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

Within each (year × department × sex) stratum the monthly proportions sum to
1, which is the structural condition the quota-based assignment algorithm
operates on.

### Departmental Patterns

```{r}
#| label: fig-b2-seasonal-patterns
#| fig-cap: "Monthly birth proportions by department (2019-2023 average)"
#| fig-alt: >
#|   Faceted line chart with 22 panels, one per department, ordered by
#|   seasonal variability. Each panel plots month (January to December) on
#|   the x-axis against the proportion of annual births on the y-axis, with
#|   a dashed reference line at the uniform 8.33%. Most departments dip in
#|   November and December and peak in the first quarter, showing systematic
#|   seasonal timing that departs from a uniform distribution.
#| fig-width: 10
#| fig-height: 12

uniform_ref <- 1 / 12

fig_b2_seasonal_patterns |>
  mutate(
    departamento = fct_reorder(departamento, mean_prop, .fun = sd, .desc = TRUE)
  ) |>
  ggplot(aes(x = mes, y = mean_prop)) +
  geom_hline(
    yintercept = uniform_ref,
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.5
  ) +
  geom_line(color = "#1B6CA8", linewidth = 0.6) +
  geom_point(color = "#1B6CA8", size = 1.3) +
  facet_wrap(~ departamento, ncol = 4, scales = "free_y") +
  scale_x_continuous(breaks = 1:12, labels = substr(month.abb, 1, 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    title = "Monthly Birth Proportions by Department",
    subtitle = "Average 2019-2023. Dashed line = uniform distribution (8.33%)",
    x = "Month",
    y = "Proportion of Annual Births"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 7, colour = "azure4"),
    strip.text = element_text(face = "bold", color = "azure4", size = 8),
    panel.grid.minor = element_blank()
  )
```

All 22 departments show seasonal patterns that depart from the uniform
distribution (dashed green line), with most departments exhibiting a trough
in November-December and peaks in the first quarter of the year. The month
assignment therefore draws on department-specific empirical proportions.

### Sex Differential

```{r}
#| label: fig-b2-sex-differential
#| fig-cap: "Male-to-female birth ratio by month (national average 2019-2023)"
#| fig-alt: >
#|   Lollipop chart with points and vertical stems for each month from
#|   January to December on the x-axis and the male-to-female birth ratio
#|   on the y-axis, zoomed to a range of 0.99 to 1.06 with a dashed
#|   reference line at parity (1.0). Every month sits above parity, between
#|   about 1.02 and 1.05, reflecting the biological excess of male births.
#| fig-width: 9
#| fig-height: 5

fig_b2_sex_differential |>
  ggplot(aes(x = mes_label, y = ratio_mf)) +
  geom_hline(
    yintercept = 1,
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.6
  ) +
  geom_segment(
    aes(xend = mes_label, yend = 1),
    color = "#1B6CA8",
    linewidth = 0.8
  ) +
  geom_point(color = "#1B6CA8", size = 3) +
  coord_cartesian(ylim = c(0.99, 1.06)) +
  labs(
    title = "Male-to-Female Birth Ratio by Month",
    subtitle = "National aggregate 2019-2023. Dashed line = parity (1.0)",
    x = "Month",
    y = "M/F Ratio"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    panel.grid.minor = element_blank()
  )
```

The male-to-female birth ratio exceeds parity in every month (range:
1.02-1.05), consistent with the documented global biological excess of male
births. This sex differential, combined with departmental seasonality, is why
the month assignment is conditioned on both department and sex.


## ENCOVI Children and Cohort Mapping

### Target Population

ENCOVI 2023 contains `r format(n_children_0_4, big.mark = ",")` children aged
0-4 years, distributed across five single-year age groups. These children form
the population requiring synthetic month assignment.

```{r}
#| label: tbl-c1-age-distribution-pre
#| tbl-cap: "Age distribution of children 0-4 years before month assignment"

tbl_c1_age_distribution_pre |>
  gt() |>
  tab_header(
    title = "Children Aged 0-4 Years in ENCOVI 2023",
    subtitle = "Distribution by age (integer years) and sex before month assignment (sampled children, unweighted)"
  ) |>
  cols_label(
    edad = "Age (years)",
    Hombre = "Male",
    Mujer = "Female",
    Total = "Total"
  ) |>
  fmt_number(columns = c(Hombre, Mujer, Total), decimals = 0, use_seps = TRUE) |>
  cols_align(align = "center") |>
  grand_summary_rows(
    fns = list(Total = ~ sum(.)),
    columns = c(Hombre, Mujer, Total),
    fmt = ~ fmt_number(., decimals = 0, use_seps = TRUE)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Cohort Mapping Logic

Each integer age maps to a single birth-year cohort. A child aged `n` years
was born in `2023 - n`. Within each cohort, the 12 possible birth months
define a 12-month age range.

```{r}
#| label: tbl-c2-cohort-mapping
#| tbl-cap: "Age-to-cohort mapping for synthetic month assignment"

tbl_c2_cohort_mapping |>
  gt() |>
  tab_header(
    title = "Cohort Mapping: ENCOVI Age to INE Birth Year",
    subtitle = "Each integer age maps to a birth year and 12-month range"
  ) |>
  cols_label(
    edad = "Age (years)",
    birth_year = "Birth Year",
    month_offset = "Month Offset",
    age_range_months = "Age Range (months)"
  ) |>
  cols_align(align = "center") |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Cohort Assignment Assumption
This mapping assumes children were interviewed during 2023 and that their
reported age reflects completed years. For children interviewed near their
birthday, the resulting imprecision is bounded at ±1 month within the cohort.
:::


## Synthetic Month Assignment

### Assignment Algorithm

Within each stratum (birth year × department × sex), the algorithm distributes
children to birth months using a greedy quota-filling approach:

1. Calculate target population per month from INE proportions and total
   survey-weighted population in the stratum.
2. Sort children by survey weight (descending), with random tie-breaking.
3. Assign each child to the month with the largest remaining quota,
   subtracting the child's weight from that month's quota.

The **survey-weighted** (population-represented) distribution of assigned
months therefore follows the INE reference proportions.

A child assigned birth month `m` within a given birth year receives
`edad_meses = month_offset + (12 - m)`, mapping January births to the oldest
position within the cohort and December births to the youngest.

::: {.callout-important}
## Population-Level Quota Matching
The algorithm operates on survey-weighted quotas. A department with 50
sampled children representing 5,000 individuals therefore produces a weighted
distribution that follows INE proportions at the population level — the level
at which results are reported. The greedy approach is quasi-deterministic:
stochasticity is limited to tie-breaking among children with identical survey
weights.
:::

### Biological Plausibility Correction

Within households containing multiple children aged 0-4 years, independent
stratum-based assignment can produce age-in-months values that are biologically
implausible for siblings. Two children from the same mother cannot be born
less than 9 months apart unless they are twins (age difference of exactly
0 months).

The post-assignment correction identifies all sibling pairs where the assigned
age separation falls in the invalid range (1-8 months) and reassigns the child
with the smaller survey weight to the valid birth month with the largest
remaining INE quota in that child's stratum. The correction is iterative:
after each reassignment, all sibling pairs within the household are
re-evaluated.

### Assignment Results

```{r}
#| label: tbl-d3-assignment-summary
#| tbl-cap: "Age in months assignment summary"

tbl_d3_assignment_summary |>
  gt() |>
  tab_header(
    title = "Synthetic Age-in-Months Assignment Summary",
    subtitle = "ENCOVI 2023 individual dataset"
  ) |>
  cols_label(
    Category = "Category",
    N = "N"
  ) |>
  fmt_number(columns = N, decimals = 0, use_seps = TRUE) |>
  cols_align(align = "left", columns = Category) |>
  cols_align(align = "right", columns = N) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      rows = Category == "Children 6-11 months (synthetic assignment)"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(
      rows = Category == "Children 6-11 months (synthetic assignment)"
    )
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Study Population Recovery
The synthetic month assignment identifies
`r format(n_recovered, big.mark = ",")` children aged 6-11 months, who cannot
be separated from infants under 6 months on integer age alone (both report
`edad == 0`). This extends the study population to the intended 6-59 month
range, covering the complementary feeding window.
:::


## Validation

### Distributional Match to INE Proportions

The primary validation compares the survey-weighted distribution of assigned
birth months against the INE reference proportions for each cohort. Grey
segments show the deviation between observed (blue dots) and expected (green
crosses) proportions.

```{r}
#| label: fig-e1-month-distribution-by-cohort
#| fig-cap: "Synthetic assignment vs INE expected birth proportions by cohort"
#| fig-alt: >
#|   Faceted chart with one panel per cohort. Each panel plots birth month
#|   (January to December) on the x-axis against proportion on the y-axis.
#|   Blue dots mark the observed survey-weighted proportions, a green dashed
#|   line with crosses marks the INE expected proportions, and grey vertical
#|   segments connect them to show the deviation. Observed and expected
#|   values track closely across cohorts, indicating the synthetic month
#|   assignment reproduces the reference distribution.
#| fig-width: 12
#| fig-height: 8

fig_e1_month_distribution_by_cohort |>
  ggplot(aes(x = birth_month_inferred)) +
  geom_segment(
    aes(xend = birth_month_inferred, y = prop_expected, yend = prop_observed),
    color = "grey70",
    linewidth = 0.4
  ) +
  geom_point(aes(y = prop_observed), color = "#1B6CA8", size = 2) +
  geom_line(
    aes(y = prop_expected),
    color = "#3AAF85",
    linewidth = 0.7,
    linetype = "dashed"
  ) +
  geom_point(aes(y = prop_expected), color = "#3AAF85", size = 1.5, shape = 4) +
  facet_wrap(~ cohort_label, ncol = 3, scales = "fixed") +
  scale_x_continuous(breaks = 1:12, labels = substr(month.abb, 1, 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    title = "Synthetic Assignment vs INE Expected Birth Proportions",
    subtitle = "Blue dots = observed (survey-weighted). Green dashed = INE expected. Grey segments = deviation",
    x = "Birth Month",
    y = "Proportion"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    strip.text = element_text(face = "bold", color = "azure4", size = 9),
    panel.grid.minor = element_blank()
  )
```

### Age Distribution in Study Population

```{r}
#| label: fig-e2-edad-meses-distribution
#| fig-cap: "Distribution of edad_meses for children 0-59 months"
#| fig-alt: >
#|   Histogram of synthetic age in months (0 to 59) on the x-axis against
#|   count of children on the y-axis, using one-month bins. A dashed
#|   vertical line at 5.5 months marks the boundary between the excluded
#|   under-6-months group and the 6-59 months study population. The bars
#|   are roughly uniform across the full range, consistent with comparable
#|   birth cohort sizes in ENCOVI.
#| fig-width: 10
#| fig-height: 6

fig_e2_edad_meses_distribution |>
  ggplot(aes(x = edad_meses)) +
  geom_histogram(
    binwidth = 1,
    fill = "#1B6CA8",
    color = "white",
    alpha = 0.7
  ) +
  geom_vline(
    xintercept = 5.5,
    color = "#E07A5F",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  annotate(
    "text",
    x = 2.7, y = Inf,
    label = "< 6 months\n(excluded)",
    vjust = 1.5,
    hjust = 0.5,
    color = "#E07A5F",
    size = 3.5,
    fontface = "italic"
  ) +
  annotate(
    "text",
    x = 32, y = Inf,
    label = "6-59 months\n(study population)",
    vjust = 1.5,
    hjust = 0.5,
    color = "#3AAF85",
    size = 3.5,
    fontface = "bold"
  ) +
  scale_x_continuous(breaks = seq(0, 59, by = 6)) +
  labs(
    title = "Distribution of Synthetic Age in Months",
    subtitle = "ENCOVI 2023 children 0-59 months. Dashed line = 6-month boundary",
    x = "Age (months)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    panel.grid.minor = element_blank()
  )
```

The histogram shows a relatively uniform distribution of age in months across
the 0-59 range, consistent with approximately equal birth cohort sizes in
ENCOVI. The 6-month boundary (dashed red line) clearly separates the excluded
breastfeeding population from the study population.

### Departmental Age Distribution

```{r}
#| label: fig-e3-department-age-boxplot
#| fig-cap: "Synthetic edad_meses distribution by department for children 6-59 months"
#| fig-alt: >
#|   Horizontal box-and-whisker chart with one box per department, ordered
#|   by median. Department runs along the y-axis and synthetic age in months
#|   (6 to 59) along the x-axis. Medians cluster around 30 to 40 months with
#|   comparable interquartile ranges and no department showing distributional
#|   anomalies, indicating no systematic assignment bias.
#| fig-width: 10
#| fig-height: 8

fig_e3_department_age_boxplot |>
  mutate(
    departamento = fct_reorder(departamento, edad_meses, .fun = median)
  ) |>
  ggplot(aes(x = departamento, y = edad_meses)) +
  geom_boxplot(
    fill = "#1B6CA8",
    alpha = 0.3,
    color = "azure4",
    outlier.size = 0.8,
    outlier.alpha = 0.4
  ) +
  coord_flip() +
  labs(
    title = "Synthetic Age Distribution by Department",
    subtitle = "Children 6-59 months, ENCOVI 2023",
    x = NULL,
    y = "Age (months)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    axis.text.y = element_text(size = 9),
    panel.grid.minor = element_blank()
  )
```

Departmental boxplots confirm that the synthetic age distribution is
well-behaved across all 22 departments, with median ages centered around
30-40 months and comparable interquartile ranges. No department shows
distributional anomalies that would suggest systematic assignment bias.

### Weighted Deviation from INE Proportions

The primary quality metric quantifies the departure of the synthetic
assignment from the target INE birth proportions. For each department, the
mean and maximum absolute deviation are calculated across all cohort-month
strata.

```{r}
#| label: tbl-e4-weighted-deviation
#| tbl-cap: "Weighted deviation from INE proportions by department"

tbl_e4_weighted_deviation |>
  gt() |>
  tab_header(
    title = "Weighted Deviation from INE Birth Proportions",
    subtitle = "By department, across all cohorts and months"
  ) |>
  cols_move(columns = n_children, after = departamento) |>
  cols_label(
    departamento = "Department",
    n_children = "N Children (sample)",
    n_strata = "Strata (cohort x month)",
    mean_abs_dev = "Mean |Deviation|",
    max_abs_dev = "Max |Deviation|"
  ) |>
  fmt_number(columns = c(mean_abs_dev, max_abs_dev), decimals = 4) |>
  cols_align(align = "left", columns = departamento) |>
  cols_align(align = "center", columns = c(n_children, n_strata, mean_abs_dev, max_abs_dev)) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = mean_abs_dev, rows = mean_abs_dev > 0.02)
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = mean_abs_dev, rows = mean_abs_dev <= 0.02)
  ) |>
  tab_footnote(
    footnote = "Green = mean deviation <= 2pp. Red = exceeds 2pp. Higher deviations correlate with smaller sample sizes and the discrete nature of the assignment algorithm. Max deviation captures the worst individual stratum (single month x cohort combination).",
    locations = cells_column_labels(columns = mean_abs_dev)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Assignment Quality
All 22 departments achieve mean absolute deviations below
`r mean_dev_pp_max` percentage points from the INE target proportions. The
departments with the highest deviations are those with smaller sample sizes
where the discrete nature of the assignment algorithm has a proportionally
larger impact.
:::

### Sibling Age Separation

Sibling age separations are reported after the biological plausibility
correction. Valid sibling pairs have an age difference of exactly 0 months
(twins) or at least 9 months (minimum full gestation period).

```{r}
#| label: tbl-e5-sibling-separation
#| tbl-cap: "Sibling age separation diagnostic after biological plausibility correction"

tbl_e5_sibling_separation |>
  gt() |>
  tab_header(
    title = "Sibling Age Separation Distribution",
    subtitle = "All sibling pairs in households with 2+ children aged 0-4 years, after biological plausibility correction"
  ) |>
  cols_label(
    separation_class = "Separation Category",
    n_pairs = "N Pairs",
    pct = "Proportion"
  ) |>
  fmt_number(columns = n_pairs, decimals = 0) |>
  fmt_percent(columns = pct, decimals = 1) |>
  cols_align(align = "left", columns = separation_class) |>
  cols_align(align = "center", columns = c(n_pairs, pct)) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(
      columns = n_pairs,
      rows = separation_class == "Invalid (1-8 months)"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = n_pairs,
      rows = separation_class != "Invalid (1-8 months)"
    )
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Biological Plausibility
After the correction, as shown in @tbl-e5-sibling-separation, no sibling
pairs remain in the invalid 1-8 month separation range. The majority of
pairs are separated by more than 24 months, consistent with typical
inter-birth intervals in Guatemala. Approximately
`r format(n_twins, big.mark = ",")` pairs (`r round(pct_twins, 1)`%) are
classified as twins (0-month separation), inferred from the algorithm
output rather than directly observed (ENCOVI does not record twin status).
:::


## Summary

### Key Findings

1. **Study population coverage**: The synthetic month assignment identifies
   `r format(n_recovered, big.mark = ",")` children aged 6-11 months, extending
   the study population to the specified 6-59 month range.

2. **Distributional accuracy**: Survey-weighted month distributions track INE
   birth proportions with mean deviations below `r mean_dev_pp_max`
   percentage points in each of the 22 departments.

3. **Biological consistency**: As documented in @tbl-e5-sibling-separation, no
   sibling pair remains in the impossible 1-8 month separation range after the
   plausibility correction.

4. **Seasonal information carried through**: The assignment reproduces
   departmental and sex-specific birth seasonality.

::: {.callout-warning}
## Limitations

1. **Interview timing assumption** — The lack of individual-level interview
   dates means cohort assignment is approximate. A child whose household was
   interviewed near the child's birthday may be misclassified by one cohort.
   This ambiguity affects approximately 1/12 of children at each boundary.

2. **Residual stochasticity** — The assignment is quasi-deterministic, with a
   random component limited to tie-breaking among children with identical
   survey weights. Individual-level assignments are synthetic and should not
   be interpreted as observed values.

3. **Twin identification** — The pairs classified as twins (0-month
   separation, see @tbl-e5-sibling-separation) are inferred rather than
   observed, as ENCOVI does not record twin status.
:::

### Application

This dataset provides month-level age resolution for all ENCOVI 2023
individuals, enabling precise WHO LMS Z-score calculations in downstream
anthropometric modeling. The synthetic `edad_meses` variable is used by
subsequent scripts in both the Covariate prediction pathway (individual
consumption allocation) and the Meta-analysis (Gunaratna) pathway
(height-for-age Z-score computation).
