---
title: "Intake Calculation"
subtitle: "Module 1: Nutrient Intakes"
description-meta: "Converting household food acquisition into individual daily nutrient intake using adult male equivalents, keeping maize and non-maize sources separate."
---

## Overview

Household consumption surveys like ENCOVI report food acquisition at the
household level, while nutrient intake analysis requires individual-level
estimates. This document transforms household food consumption into individual
nutrient intake through three key steps:

1. **Standardize maize products** — Convert tortillas and flour to grain
   equivalent weight to isolate the fraction affected by biofortification.
2. **Separate food sources** — Calculate intake from maize and non-maize
   sources independently, enabling scenario modeling where only maize nutrients
   change.
3. **Distribute to individuals** — Apply Adult Male Equivalent (AME) factors to
   allocate household consumption to each member proportionally to their energy
   requirements.

The separation of maize from other food sources is fundamental to
biofortification impact modeling. Because biofortification only affects
specific nutrient levels in maize, we need to know precisely how much of each
person's iron, zinc, and protein intake comes from maize versus other foods.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
pacman::p_unload(all = TRUE)

# Load required packages (using pacman for automatic installation)
pacman::p_load(
  rio,            # Robust import/export for various formats
  here,           # Robust file path management
  tidyverse,      # General purpose data manipulation
  gt              # Table formatting and presentation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script.
tbl_c_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_02_tbl_c_summary.rds"),
  trust = TRUE
)

tbl_d1_summary_sample_characteristics <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_04_tbl_d1_summary_sample_characteristics.rds"),
  trust = TRUE
)

tbl_d2a_summary_total_maize_intake <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_04_tbl_d2a_summary_total_maize_intake.rds"),
  trust = TRUE
)

tbl_d2b_summary_maize_contribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_04_tbl_d2b_summary_maize_contribution.rds"),
  trust = TRUE
)

tbl_d2c_summary_intake_by_age <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_04_tbl_d2c_summary_intake_by_age.rds"),
  trust = TRUE
)

tbl_d2d_summary_diet_composition <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_04_tbl_d2d_summary_diet_composition.rds"),
  trust = TRUE
)

tbl_d3a_quality_check_missing <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_04_tbl_d3a_quality_check_missing.rds"),
  trust = TRUE
)

tbl_d3b_quality_check_extremes <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "01_04_tbl_d3b_quality_check_extremes.rds"),
  trust = TRUE
)
```


## Household Food Processing

### Maize Standardization

In Guatemala, maize is rarely consumed as raw grain. Most households eat
tortillas — the staple food prepared through nixtamalization — or use
commercially milled maize flour. To assess the impact of biofortification, we
must convert these processed products back to their **maize grain equivalent**.

Following INCAP guidelines for dietary intake assessment in Guatemala, weight
conversion factors are applied:

| Product | Factor | Rationale |
|---------|--------|-----------|
| Tortillas | ×0.52 | Nixtamalization adds water and calcium hydroxide (*cal*); ~52% of final weight is maize grain |
| Maize flour | ×0.98 | Dry milling retains ~98% of grain weight |
| Raw grain | ×1.00 | No conversion needed |

After weight conversion, nutrient values are reassigned to match raw white
maize composition from the INCAP Food Composition Table (per 100g: 365 kcal,
9.42g protein, 2.71mg iron, 2.21mg zinc). Nutrient calculations then reflect
the maize fraction that biofortification modifies.

::: {.callout-tip}
## Raw Maize Composition After Weight Conversion
Tortilla composition carries nutrients contributed by the nixtamalization
process (calcium from *cal*) and water content that dilutes nutrient density.
Applying raw maize values to the converted grain-equivalent weight isolates
the component of a tortilla that biofortification modifies. Total dietary
energy intake is unchanged by this step: the conversion identifies the
maize-derived nutrient fraction within the household diet.

Source: Conversion factors derived from INCAP dietary assessment protocols.
:::


### Separating Maize from Non-Maize Sources

Biofortification increases the micronutrient content of food crops through
conventional plant breeding. When farmers adopt biofortified maize varieties,
only the maize-derived portion of dietary intake changes — nutrients from
beans, eggs, vegetables, and other foods remain constant.

To model intervention scenarios accurately, we calculate intake separately for:

- **Maize sources**: Grain, tortillas, and flour (codine 7, 17, 21).
- **Non-maize sources**: All other foods (calculated by subtraction).


### Plant-Based and Animal-Based Foods

The proportion of plant-based foods in the diet is a descriptor of dietary
composition at the household level. It is computed here and used in downstream
analyses to characterise the dietary patterns of the target population.

**Plant-based foods** (`vegetal_cereal == "si"`) include cereals (maize, rice,
wheat), legumes (black beans, lentils), vegetables, fruits, tubers, and oils.

**Animal-based foods** (`vegetal_cereal == "no"`) include meat (beef, pork,
chicken), dairy products (milk, cheese), eggs, and fish.

These datasets track weight only (grams per day), not full nutrient profiles:
they are used to derive the plant-based percentage of each household's diet,
which is reported as a descriptor of dietary composition.


### Household Daily Intake

```{r}
#| label: tbl-c-summary

tbl_c_summary |>
  gt() |>
  tab_header(
    title = "Household Daily Intake Summary",
    subtitle = "Median and interquartile range (IQR) by food group"
  ) |>
  cols_label(
    Metric = "Metric",
    Total = "Total Foods",
    Maize = "Maize Only",
    `Plant-based` = "Plant-based",
    `Animal-based` = "Animal-based"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Metric,
      rows = Metric %in% c("Households (N)", "Grams (median)", "Energy - kcal (median)",
                           "Protein - g (median)", "Iron - mg (median)", "Zinc - mg (median)")
    )
  ) |>
  tab_style(
    style = cell_text(style = "italic", color = "gray50"),
    locations = cells_body(
      columns = Metric,
      rows = grepl("IQR", Metric)
    )
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  ) |>
  tab_footnote(
    footnote = "Plant-based and animal-based datasets only track weight (grams), not full nutritional composition.",
    locations = cells_column_labels(columns = c(`Plant-based`, `Animal-based`))
  )
```

::: {.callout-note}
## Data Processing Details

- **Unit conversions**: All food quantities converted to grams using
  expert-provided equivalence table.
- **Recall period**: Daily intake calculated by dividing 15-day recall period
  by 15.
- **Edible portion**: All calculations include adjustment for edible portion
  (`pct_aprov`) to account for the inedible portion of food (peels, bones,
  etc.).
- **Food groupings**: Total (all foods), Maize (grain, tortillas, flour),
  Plant-based (cereals, grains, vegetables, fruits), Animal-based (meat,
  dairy, eggs, fish).
:::


## Individual Intake Distribution

ENCOVI reports what the household acquired, not what each person consumed.
Nutrient intake, requirements, and subsequent health outcomes, however, are
individual-level phenomena. Aggregate household consumption data masks
individual-level variation — a 3-year-old and an adult man have very different
nutrient requirements that simple per-capita division would not capture.


### Adult Male Equivalent (AME)

Individual nutrient intake distribution requires assigning each household
member a fraction of household intake proportional to their energy
requirements. This is done via Adult Male Equivalent (AME) factors, calculated
on the synthetic `edad_meses` variable produced by
[Estimate Age in Months](01_02_estimate_age_in_months.qmd). Month-level
resolution is what allows AME to be assigned within the 6-11 month range, the
complementary feeding window.

The **Adult Male Equivalent (AME)** method assigns each household member a
factor proportional to their energy requirements relative to an adult male
aged 18-30 years (AME = 1.0) [@Weisell2012_AME]. Household food intake is then
distributed according to each member's share of total household AME.

The AME factors used here follow INCAP's dietary recommendations for Guatemala
[@INCAP2012_AME]:

| Age Group | Males | Females |
|-----------|-------|---------|
| 0 – 5 months | 0 | 0 |
| 6 months – 1 year | 0.21 | 0.21 |
| 1 – 2 years | 0.27 | 0.27 |
| 2 – 4 years | 0.37 | 0.37 |
| 4 – 7 years | 0.44 | 0.44 |
| 7 – 10 years | 0.56 | 0.51 |
| 10 – 14 years | 0.73 | 0.65 |
| 14 – 18 years | 0.96 | 0.73 |
| 18 – 30 years | 1.00 | 0.74 |
| 30 – 60 years | 0.95 | 0.74 |
| 60+ years | 0.76 | 0.65 |

For example, in a household with a 25-year-old man (AME = 1.0), a 25-year-old
woman (AME = 0.74), and a 3-year-old child (AME = 0.37), the total household
AME is 2.11. The child's share is 0.37/2.11 = 17.5% of household food intake.

::: {.callout-note}
## Infants Under 6 Months
Infants under 6 months receive `ame_porcentaje = 0`, reflecting the WHO
recommendation for exclusive breastfeeding during this period. These infants
are not allocated any household food consumption in subsequent calculations,
as their nutrient requirements are assumed to be met through breastmilk.
The synthetic `edad_meses` variable enables this precise distinction between
infants under 6 months (excluded) and children aged 6-11 months (included with
AME = 0.21).
:::


### Maize vs Non-Maize Separation

Individual intake is calculated separately for maize and non-maize sources.

::: {.callout-important}
## Why This Separation Matters
This separation is essential for biofortification modeling — only
maize-derived nutrients will change in intervention scenarios, while non-maize
intake remains constant.
:::


### Survey Design

ENCOVI uses a complex sampling design with Primary Sampling Units (UPM) and
expansion weights that allow inference to the full Guatemalan population (see
@tbl-d1-summary-sample-characteristics for the weighted population estimate).
All population-level statistics require survey-weighted estimation, with
standard errors accounting for clustering within UPMs.


## Summary

### Sample Characteristics

```{r}
#| label: tbl-d1-summary-sample-characteristics

tbl_d1_summary_sample_characteristics |>
  gt() |>
  tab_header(
    title = "Sample Characteristics",
    subtitle = "Survey-weighted estimates for Guatemala"
  ) |>
  cols_label(
    Characteristic = "Characteristic",
    Value = "Value"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```


### Total and Maize Daily Intake

```{r}
#| label: tbl-d2a-summary-total-maize-intake

tbl_d2a_summary_total_maize_intake |>
  gt() |>
  tab_header(
    title = "Total and Maize Daily Intake",
    subtitle = "Survey-weighted population estimates with 95% confidence intervals"
  ) |>
  cols_label(
    Nutrient = "Nutrient/Metric",
    `Mean (95% CI)` = "Mean (95% CI)"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  ) |>
  tab_footnote(
    footnote = "Estimates account for complex survey design (clustering, weights).",
    locations = cells_title(groups = "subtitle")
  )
```


### Maize Contribution to Nutrient Intake

The table below shows mean daily intake for key nutrients, separated by source.
The "% from Maize" column is the biofortification-modifiable fraction: the
share of intake that would increase if households consumed biofortified
varieties. For protein, biofortification does not change the total protein
content of the grain — it changes the amino acid profile (lysine, tryptophan),
and with it the utilizable protein fraction.

```{r}
#| label: tbl-d2b-summary-maize-contribution

tbl_d2b_summary_maize_contribution |>
  gt() |>
  tab_header(
    title = "Maize Contribution to Total Nutrient Intake",
    subtitle = "Percentage of daily nutrients from maize products"
  ) |>
  cols_label(
    Nutrient = "Nutrient",
    `% from Maize (95% CI)` = "% from Maize (95% CI)"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```


### Intake by Age Group

The table below describes the distribution of mean daily nutrient intake across
age groups under the Adult Male Equivalent allocation, where household food is
distributed to each member in proportion to energy requirements.

```{r}
#| label: tbl-d2c-summary-intake-by-age

tbl_d2c_summary_intake_by_age |>
  gt() |>
  tab_header(
    title = "Daily Nutrient Intake by Age Group",
    subtitle = "Survey-weighted population estimates"
  ) |>
  fmt_number(
    columns = c(mean_kcal, mean_kcal_se, mean_prot, mean_prot_se,
                mean_fe, mean_fe_se, mean_zn, mean_zn_se,
                pct_maiz_kcal, pct_maiz_kcal_se),
    decimals = 1
  ) |>
  fmt_number(
    columns = n,
    decimals = 0,
    use_seps = TRUE
  ) |>
  cols_label(
    edad_grupos = "Age Group",
    n = "N",
    mean_kcal = "Energy (kcal)",
    mean_kcal_se = "Energy (SE)",
    mean_prot = "Protein (g)",
    mean_prot_se = "Protein (SE)",
    mean_fe = "Iron (mg)",
    mean_fe_se = "Iron (SE)",
    mean_zn = "Zinc (mg)",
    mean_zn_se = "Zinc (SE)",
    pct_maiz_kcal = "Maize Energy (%)",
    pct_maiz_kcal_se = "Maize Energy (SE)"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```


### Diet Composition {#diet-composition}

Plant-based diets predominate in Guatemala (see
@tbl-d2d-summary-diet-composition for the population distribution across
plant-based food share categories). The pattern is consistent with limited
access to animal-source foods in rural and low-income populations.

```{r}
#| label: tbl-d2d-summary-diet-composition

tbl_d2d_summary_diet_composition |>
  gt() |>
  tab_header(
    title = "Distribution of Plant-Based Diet Categories",
    subtitle = "Survey-weighted population proportions"
  ) |>
  fmt_number(
    columns = n,
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_percent(
    columns = c(proportion, proportion_low, proportion_upp),
    decimals = 1
  ) |>
  cols_label(
    per_vegetales_estratificado = "Plant-based Foods (%)",
    n = "Sample N",
    proportion = "Population %",
    proportion_low = "95% CI Lower",
    proportion_upp = "95% CI Upper"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  ) |>
  tab_footnote(
    footnote = "Categories based on percentage of grams from plant-based foods.",
    locations = cells_column_labels(columns = per_vegetales_estratificado)
  )
```

::: {.callout-tip}
## Plant Percentage Application
The distribution of the plant-based food percentage across the population is
reported here as a descriptor of dietary composition. This distribution is
also referenced in the
[iron absorption decision](../04_covariate_prediction/04_03_calculate_bioavailability_adjustments.qmd#imp-iron-absorption-rate),
where it supports the choice of a uniform 5% iron absorption factor for the
target population.
:::


### Data Quality

Data completeness is high across key variables. Infants under 6 months are
assigned `ame_porcentaje = 0` under the exclusive breastfeeding assumption
and are excluded from household-level food allocation — these are not treated
as missing values in the analysis.

```{r}
#| label: tbl-d3a-quality-check-missing

tbl_d3a_quality_check_missing |>
  gt() |>
  tab_header(
    title = "Missing Values by Key Variable",
    subtitle = "Data completeness assessment"
  ) |>
  fmt_number(
    columns = `Missing N`,
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    columns = `Missing %`,
    decimals = 2,
    pattern = "{x}%"
  ) |>
  cols_label(
    Variable = "Variable",
    `Missing N` = "N",
    `Missing %` = "%"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

Extreme values are present in a non-trivial share of records, consistent with
unusual household compositions or recall error. Handling of these values in
subsequent analyses is described in
[Modeling Data Preparation](../03_transfer_models/03_01_prepare_modeling_data.qmd).

```{r}
#| label: tbl-d3b-quality-check-extremes

tbl_d3b_quality_check_extremes |>
  gt() |>
  tab_header(
    title = "Extreme Value Detection",
    subtitle = "Potential data quality issues or outliers"
  ) |>
  fmt_number(
    columns = Count,
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    columns = Percentage,
    decimals = 2,
    pattern = "{x}%"
  ) |>
  cols_label(
    Check = "Quality Check",
    Count = "N",
    Percentage = "%"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  ) |>
  tab_footnote(
    footnote = "Extreme values may indicate data entry errors, unusual household compositions, or recall bias.",
    locations = cells_title(groups = "subtitle")
  )
```

::: {.callout-note}
## Data Quality Considerations

- **Infants under 6 months**: Have `ame_porcentaje = 0`, resulting in zero
  calculated intake from solid foods under the exclusive breastfeeding
  assumption.
- **Missing values**: Typically occur for individuals in households with no
  food consumption data reported. These are genuine missing data, not
  calculation errors.
- **Extreme values**: May reflect data entry errors, unusual household
  compositions (e.g., large families, guests), or reporting period issues
  (recall bias).
:::
