---
title: "Convert Raw Data to Parquet Format"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script converts original data files from Excel and Stata formats
to Parquet format for efficient processing by the analytical pipeline.
The workflow:

1. Reads `raw_sources.yml` to identify active source files (only
   files marked `active: true` are converted)
2. Converts each survey source file to Parquet (no data transformations)
3. Unifies multi-file reference sources into single Parquet tables:
   - INE birth registry (2019-2023) → one parquet with `anio` column
   - WHO Child Growth Standards (monthly, 0-5) → one parquet with `sexo`
   - WHO Child Growth Standards (daily expanded, 0-5) → one parquet with `sexo`
   - WHO Growth Reference 2007 (monthly, 5-19) → one parquet with `sexo`
4. Verifies conversion fidelity (row count, column count, column names)
5. Cross-references with the hash log from `00_01_download_raw_sources.qmd`
6. Packages active originals into `_originals/active_raw_sources.zip`
7. Removes the original files after successful conversion, verification,
   and archival packaging (controlled by `remove_originals_after_conversion`)

**Survey sources converted (one-to-one, manifest-driven):**

- ENCOVI 2023: 5 active Excel files (out of 35+ available)
- SIVESNU 2018: 4 Stata files

**Reference sources unified (many-to-one):**

- INE Nacimientos 2019-2023: 5 Excel files → 1 Parquet
- WHO Standards monthly: 4 Excel files → 1 Parquet
- WHO Standards daily (expanded): 2 Excel files → 1 Parquet
- WHO Reference 5-19: 2 Excel files → 1 Parquet

::: {.callout-note}
## Prerequisite
This script expects the original files to be present in
`01_data/01_raw/{SOURCE}/`. Run `00_01_download_raw_sources.qmd` first
if starting from a clean environment. If the Parquet files already
exist and `force_convert` is `FALSE` (default), the script skips
conversion and produces only the verification summary.
:::

# Setup

```{r}
#| label: setup

rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

pacman::p_load(
  arrow,          # Parquet file handling
  here,           # Robust file path management
  gt,             # Publication tables
  janitor,        # clean_names()
  yaml,           # Manifest parsing
  fs,             # Filesystem operations in IO helpers
  tidyverse       # General purpose data manipulation (includes purrr, stringr)
)

# Packages used via :: in provisioning functions (not loaded to avoid masking)
# haven    — read_dta() for Stata file import
# readxl   — read_excel() for Excel file import
# curl     — download functions (only needed by 00_01)
# digest   — SHA256 hashing (only needed by 00_01)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

options(scipen = 999, digits = 4)

# Load provisioning utility functions
source(here("02_code", "00_functions", "00_functions_provisioning.R"))
```

# Configuration

```{r}
#| label: configuration

# Force re-conversion even if Parquet files already exist
force_convert <- FALSE

# Remove original files after successful conversion, verification, and
# archival packaging. Originals are preserved in the ZIP archive and can
# be re-downloaded at any time by re-running 00_01.
remove_originals_after_conversion <- TRUE

# Directories
raw_root      <- here("01_data", "01_raw")
originals_dir <- here("01_data", "01_raw", "_originals")

# Hash log from 00_01 (used for cross-referencing, not recalculated)
hash_log_path <- here("01_data", "01_raw", "_hashes_active.csv")
```

# Phase A: Convert Survey Data (One-to-One)

Each active original file is converted to a Parquet file with the same
name in the same directory. No data transformations are applied. The
list of files to convert is driven by `raw_sources.yml` — only files
marked `active: true` are processed.

## A1. Build Conversion Manifest

```{r}
#| label: a1-build-conversion-manifest

manifest <- yaml::read_yaml(
  here("02_code", "_config", "raw_sources.yml")
)

manifest_flat <- flatten_manifest(manifest) |>
  filter(active)

# Build file-level conversion manifest from active entries only
conversion_manifest <- manifest_flat |>
  filter(source_label %in% c("ENCOVI_2023", "SIVESNU")) |>
  mutate(
    source_dir = case_when(
      source_label == "ENCOVI_2023" ~ here("01_data", "01_raw", "ENCOVI_2023"),
      source_label == "SIVESNU"     ~ here("01_data", "01_raw", "SIVESNU_2018")
    ),
    original_path = file.path(source_dir, filename),
    original_filename = filename,
    parquet_path = str_replace(original_path, "\\.(xlsx|dta)$", ".parquet"),
    parquet_exists = file.exists(parquet_path),
    needs_conversion = force_convert | !parquet_exists
  )
```

```{r}
#| label: tbl-a1-conversion-manifest
#| tbl-cap: "Active files identified for one-to-one conversion"

conversion_manifest |>
  select(source_label, original_filename, parquet_exists, needs_conversion) |>
  gt() |>
  tab_header(
    title = "Conversion manifest",
    subtitle = "Active files from raw_sources.yml and their Parquet status"
  ) |>
  cols_label(
    source_label = "Source",
    original_filename = "Original file",
    parquet_exists = "Parquet exists",
    needs_conversion = "Needs conversion"
  )
```

## A2. Execute Conversions

```{r}
#| label: a2-convert-files

files_to_convert <- conversion_manifest |> filter(needs_conversion)

if (nrow(files_to_convert) > 0) {
  conversion_results <- files_to_convert |>
    pmap_dfr(function(original_path, parquet_path, source_label, ...) {
      convert_one_file(original_path, parquet_path, source_label)
    })
} else {
  conversion_results <- tibble(
    source_label = character(), original_filename = character(),
    parquet_filename = character(), original_nrow = integer(),
    original_ncol = integer(), conversion_success = logical()
  )
}
```

```{r}
#| label: tbl-a2-conversion-results
#| tbl-cap: "One-to-one conversion results"

if (nrow(conversion_results) > 0) {
  conversion_results |>
    select(source_label, original_filename, original_nrow,
           original_ncol, conversion_success) |>
    gt() |>
    tab_header(
      title = "Phase A - Parquet conversion",
      subtitle = "Files converted in this execution"
    ) |>
    cols_label(
      source_label = "Source",
      original_filename = "Original file",
      original_nrow = "Rows",
      original_ncol = "Columns",
      conversion_success = "Success"
    ) |>
    fmt_number(columns = original_nrow, decimals = 0, use_seps = TRUE)
} else {
  tibble(status = "All Parquet files already exist. No conversions performed.") |>
    gt() |>
    tab_header(title = "Phase A - Parquet conversion")
}
```

# Phase B: Unify INE Birth Registry (2019-2023)

The INE birth registry is published as one Excel file per year, each
with different sheet names and header structures. This phase reads all
five files, harmonizes the format, and produces a single Parquet file
with an `anio` column identifying the source year.

```{r}
#| label: b1-unify-ine-nacimientos

ine_source_dir <- here("01_data", "01_raw", "INE_NACIMIENTOS")
ine_output_path <- here("01_data", "01_raw", "INE_NACIMIENTOS",
                         "ine_nacimientos_2019_2023.parquet")

ine_needs_conversion <- force_convert | !file.exists(ine_output_path)

if (ine_needs_conversion) {

  ine_files <- tibble(
    anio = 2019:2023,
    filename = paste0("INE_Guatemala_Nacimientos_", 2019:2023, ".xlsx"),
    sheet_name = case_when(
      anio %in% c(2019, 2020) ~ "Depto. residencia y mes ocu",
      anio %in% c(2021, 2022, 2023) ~ "2 Depto. residencia y mes ocu."
    ),
    skip_rows = case_when(
      anio == 2023 ~ 7L,
      TRUE         ~ 5L
    )
  )

  departamentos_gt <- c(
    "Guatemala", "El Progreso", "Sacatepequez", "Chimaltenango",
    "Escuintla", "Santa Rosa", "Solola", "Totonicapan",
    "Quetzaltenango", "Suchitepequez", "Retalhuleu", "San Marcos",
    "Huehuetenango", "Quiche", "Baja Verapaz", "Alta Verapaz",
    "Peten", "Izabal", "Zacapa", "Chiquimula",
    "Jalapa", "Jutiapa"
  )

  df_nacimientos <- ine_files |>
    pmap_dfr(function(anio, filename, sheet_name, skip_rows) {
      load_ine_birth_year(
        anio = anio,
        filename = filename,
        sheet_name = sheet_name,
        skip_rows = skip_rows,
        source_dir = ine_source_dir,
        departamentos_gt = departamentos_gt
      )
    }) |>
    arrange(anio, departamento, mes)

  write_parquet(df_nacimientos, ine_output_path)

  ine_summary <- tibble(
    output_file = basename(ine_output_path),
    description = "Birth registry 2019-2023",
    n_rows = nrow(df_nacimientos),
    detail = paste0(n_distinct(df_nacimientos$departamento),
                    " departments, ", n_distinct(df_nacimientos$anio),
                    " years"),
    success = file.exists(ine_output_path)
  )

  rm(df_nacimientos)

} else {
  ine_summary <- tibble(
    output_file = basename(ine_output_path),
    description = "Birth registry 2019-2023",
    n_rows = NA_integer_, detail = NA_character_, success = TRUE
  )
}
```

```{r}
#| label: tbl-b1-ine-summary
#| tbl-cap: "INE birth registry unification summary"

ine_summary |>
  gt() |>
  tab_header(
    title = "Phase B - INE birth registry",
    subtitle = "Five annual files unified into a single Parquet"
  ) |>
  cols_label(
    output_file = "Output file",
    description = "Description",
    n_rows = "Rows",
    detail = "Detail",
    success = "Success"
  ) |>
  fmt_number(columns = n_rows, decimals = 0, use_seps = TRUE) |>
  sub_missing(missing_text = "cached")
```

# Phase C: Unify WHO Growth Standards and Reference

The WHO publishes length/height-for-age data across multiple files
covering different age ranges, sexes, and resolutions. This phase
produces three unified Parquet files, each serving a distinct
analytical purpose in the framework.

## C1. WHO Standards — Monthly (0-60 months)

Four files (boys/girls x 0-2/2-5 years) are combined into a single
table with a `sexo` column. Month 24 appears in both the 0-2 and 2-5
files; the 0-2 version is retained and the 2-5 duplicate is removed.
This is consistent with the WHO measurement protocol: at 24 months the
standard transitions from recumbent length to standing height, and the
0-2 table uses the recumbent measurement convention.

```{r}
#| label: c1-who-standards-monthly

who_src <- here("01_data", "01_raw", "WHO_GROWTH")

who_monthly_path <- here(who_src, "who_standards_lhfa_monthly.parquet")
who_monthly_needs <- force_convert | !file.exists(who_monthly_path)

if (who_monthly_needs) {

  df_girls <- bind_rows(
    readxl::read_excel(
      here(who_src, "lhfa_girls_0_to_2_years_zscores.xlsx")
    ) |> clean_names(),
    readxl::read_excel(
      here(who_src, "lhfa_girls_2_to_5_years_zscores.xlsx")
    ) |> clean_names() |>
      filter(month > 24)
  ) |>
    mutate(sexo = "Mujer")

  df_boys <- bind_rows(
    readxl::read_excel(
      here(who_src, "lhfa_boys_0_to_2_years_zscores.xlsx")
    ) |> clean_names(),
    readxl::read_excel(
      here(who_src, "lhfa_boys_2_to_5_years_zscores.xlsx")
    ) |> clean_names() |>
      filter(month > 24)
  ) |>
    mutate(sexo = "Hombre")

  df_who_monthly <- bind_rows(df_girls, df_boys) |>
    arrange(sexo, month)

  write_parquet(df_who_monthly, who_monthly_path)

  who_monthly_summary <- tibble(
    output_file = basename(who_monthly_path),
    description = "Standards 0-60 months (monthly)",
    n_rows = nrow(df_who_monthly),
    age_range = paste0(min(df_who_monthly$month), "-",
                       max(df_who_monthly$month), " months"),
    success = TRUE
  )

  rm(df_girls, df_boys, df_who_monthly)

} else {
  who_monthly_summary <- tibble(
    output_file = basename(who_monthly_path),
    description = "Standards 0-60 months (monthly)",
    n_rows = NA_integer_, age_range = NA_character_, success = TRUE
  )
}
```

## C2. WHO Standards — Daily Expanded (0-1856 days)

Two expanded tables (boys/girls) with daily resolution. No
deduplication needed — the files cover 0-1856 days without overlap.

```{r}
#| label: c2-who-standards-daily

who_daily_path <- here(who_src, "who_standards_lhfa_daily.parquet")
who_daily_needs <- force_convert | !file.exists(who_daily_path)

if (who_daily_needs) {

  df_who_daily <- bind_rows(
    readxl::read_excel(
      here(who_src, "lhfa-boys-zscore-expanded-tables.xlsx")
    ) |> clean_names() |>
      mutate(sexo = "Hombre"),
    readxl::read_excel(
      here(who_src, "lhfa-girls-zscore-expanded-tables.xlsx")
    ) |> clean_names() |>
      mutate(sexo = "Mujer")
  ) |>
    arrange(sexo, day)

  write_parquet(df_who_daily, who_daily_path)

  who_daily_summary <- tibble(
    output_file = basename(who_daily_path),
    description = "Standards 0-1856 days (daily expanded)",
    n_rows = nrow(df_who_daily),
    age_range = paste0("0-", round(max(df_who_daily$day) / 30.4375),
                       " months (daily)"),
    success = TRUE
  )

  rm(df_who_daily)

} else {
  who_daily_summary <- tibble(
    output_file = basename(who_daily_path),
    description = "Standards 0-1856 days (daily expanded)",
    n_rows = NA_integer_, age_range = NA_character_, success = TRUE
  )
}
```

## C3. WHO Growth Reference 2007 — Monthly (5-19 years)

Two files (boys/girls) from the WHO 2007 Growth Reference for
school-age children and adolescents. Monthly resolution, months
61-228.

```{r}
#| label: c3-who-reference-monthly

who_ref_path <- here(who_src, "who_reference_hfa_monthly.parquet")
who_ref_needs <- force_convert | !file.exists(who_ref_path)

if (who_ref_needs) {

  df_who_ref <- bind_rows(
    readxl::read_excel(
      here(who_src, "hfa-boys-z-who-2007-exp.xlsx")
    ) |> clean_names() |>
      mutate(sexo = "Hombre"),
    readxl::read_excel(
      here(who_src, "hfa-girls-z-who-2007-exp.xlsx")
    ) |> clean_names() |>
      mutate(sexo = "Mujer")
  ) |>
    arrange(sexo, month)

  write_parquet(df_who_ref, who_ref_path)

  who_ref_summary <- tibble(
    output_file = basename(who_ref_path),
    description = "Reference 2007, 5-19 years (monthly)",
    n_rows = nrow(df_who_ref),
    age_range = paste0(min(df_who_ref$month), "-",
                       max(df_who_ref$month), " months"),
    success = TRUE
  )

  rm(df_who_ref)

} else {
  who_ref_summary <- tibble(
    output_file = basename(who_ref_path),
    description = "Reference 2007, 5-19 years (monthly)",
    n_rows = NA_integer_, age_range = NA_character_, success = TRUE
  )
}
```

```{r}
#| label: tbl-c3-who-summary
#| tbl-cap: "WHO growth data unification summary"

bind_rows(who_monthly_summary, who_daily_summary, who_ref_summary) |>
  gt() |>
  tab_header(
    title = "Phase C - WHO growth data",
    subtitle = "Multiple Excel files unified into Parquet tables"
  ) |>
  cols_label(
    output_file = "Output file",
    description = "Description",
    n_rows = "Rows",
    age_range = "Age range",
    success = "Success"
  ) |>
  fmt_number(columns = n_rows, decimals = 0, use_seps = TRUE) |>
  sub_missing(missing_text = "cached")
```

# Phase D: Verify Conversion Fidelity

Re-read each one-to-one Parquet file and compare its structure against
the original to confirm that the conversion preserved all data.

```{r}
#| label: d1-verify-fidelity

verifiable <- conversion_manifest |>
  filter(file.exists(original_path), file.exists(parquet_path))

if (nrow(verifiable) > 0) {
  verification <- verifiable |>
    pmap_dfr(function(original_path, parquet_path, source_label, ...) {
      verify_one_pair(original_path, parquet_path, source_label)
    })
} else {
  verification <- tibble(
    source_label = character(), filename = character(),
    nrow_original = integer(), nrow_parquet = integer(),
    ncol_original = integer(), ncol_parquet = integer(),
    cols_match = logical(), rows_match = logical(),
    fidelity_ok = logical()
  )
}
```

```{r}
#| label: tbl-d1-fidelity-verification
#| tbl-cap: "Conversion fidelity verification"

if (nrow(verification) > 0) {
  verification |>
    select(source_label, filename, nrow_original, nrow_parquet,
           ncol_original, ncol_parquet, fidelity_ok) |>
    gt() |>
    tab_header(
      title = "Phase D - Conversion fidelity",
      subtitle = "Structure comparison between original and Parquet files"
    ) |>
    cols_label(
      source_label = "Source",
      filename = "Parquet file",
      nrow_original = "Rows (original)",
      nrow_parquet = "Rows (parquet)",
      ncol_original = "Cols (original)",
      ncol_parquet = "Cols (parquet)",
      fidelity_ok = "Fidelity OK"
    ) |>
    fmt_number(
      columns = c(nrow_original, nrow_parquet),
      decimals = 0, use_seps = TRUE
    ) |>
    tab_style(
      style = cell_fill(color = "#D4EDDA"),
      locations = cells_body(columns = fidelity_ok,
                             rows = fidelity_ok == TRUE)
    ) |>
    tab_style(
      style = cell_fill(color = "#F8D7DA"),
      locations = cells_body(columns = fidelity_ok,
                             rows = fidelity_ok == FALSE)
    )
} else {
  tibble(
    status = "No original-parquet pairs available for verification."
  ) |>
    gt() |>
    tab_header(title = "Phase D - Conversion fidelity")
}
```

```{r}
#| label: d2-fidelity-guard

if (nrow(verification) > 0) {
  n_failed <- sum(!verification$fidelity_ok)
  if (n_failed > 0) {
    stop(sprintf(
      "Phase D halted: %d file(s) failed fidelity verification.",
      n_failed
    ))
  }
}
```

# Phase E: Cross-Reference with Download Hashes

If the hash log from `00_01_download_raw_sources.qmd` is available,
display it alongside the converted files for a complete provenance
chain.

```{r}
#| label: tbl-e1-hash-cross-reference
#| tbl-cap: "Provenance chain: download hashes linked to converted files"

if (file.exists(hash_log_path)) {
  hash_log <- read_csv(hash_log_path, show_col_types = FALSE)

  provenance <- conversion_manifest |>
    select(source_label, original_filename, parquet_exists) |>
    left_join(
      hash_log |> select(filename, sha256_observed, status),
      by = c("original_filename" = "filename")
    ) |>
    mutate(
      sha256_short = if_else(
        is.na(sha256_observed),
        NA_character_,
        paste0(substr(sha256_observed, 1, 16), "...")
      )
    )

  provenance |>
    select(source_label, original_filename, sha256_short,
           status, parquet_exists) |>
    gt() |>
    tab_header(
      title = "Provenance chain",
      subtitle = "Download hash verification linked to Parquet conversion"
    ) |>
    cols_label(
      source_label = "Source",
      original_filename = "Original file",
      sha256_short = "SHA256 (truncated)",
      status = "Hash status",
      parquet_exists = "Parquet available"
    ) |>
    sub_missing(missing_text = "not in hash log") |>
    tab_footnote(
      footnote = "Full hashes in 01_data/01_raw/_hashes_active.csv",
      locations = cells_column_labels(columns = sha256_short)
    )
} else {
  tibble(
    note = paste(
      "Hash log not found. Run 00_01_download_raw_sources.qmd to",
      "generate it. Conversion is valid regardless; the hash log",
      "provides additional provenance for audit purposes."
    )
  ) |>
    gt() |>
    tab_header(
      title = "Provenance chain",
      subtitle = "Hash log from 00_01 not available"
    )
}
```

# Phase F: Package Active Sources to ZIP

After conversion and fidelity verification, the original files are
bundled into a single ZIP archive for audit purposes. This archive is
the canonical artifact to share with auditors: it contains the exact
bytes used by the framework, accompanied by the hash log from `00_01`.

```{r}
#| label: f1-package-active

fs::dir_create(originals_dir)

active_manifest <- flatten_manifest(
  yaml::read_yaml(here("02_code", "_config", "raw_sources.yml"))
) |>
  filter(active) |>
  mutate(
    target_dir = case_when(
      source_label == "ENCOVI_2023"     ~ here(raw_root, "ENCOVI_2023"),
      source_label == "SIVESNU"         ~ here(raw_root, "SIVESNU_2018"),
      source_label == "INE_NACIMIENTOS" ~ here(raw_root, "INE_NACIMIENTOS"),
      source_label == "WHO_GROWTH"      ~ here(raw_root, "WHO_GROWTH"),
      TRUE                              ~ NA_character_
    ),
    target_path = file.path(target_dir, filename)
  )

active_files_on_disk <- active_manifest |>
  pull(target_path) |>
  keep(file.exists)

active_archive_path <- file.path(originals_dir, "active_raw_sources.zip")

active_pkg <- package_to_zip(
  files = active_files_on_disk,
  archive_path = active_archive_path,
  working_dir = raw_root
)
```

```{r}
#| label: tbl-f-packaging-summary
#| tbl-cap: "ZIP packaging summary"

active_pkg |>
  select(archive_path, n_files, archive_size_mb, success) |>
  gt() |>
  tab_header(
    title = "Phase F - ZIP packaging",
    subtitle = "Active originals bundled for audit retention"
  ) |>
  cols_label(
    archive_path = "Path",
    n_files = "Files",
    archive_size_mb = "Size (MB)",
    success = "Success"
  ) |>
  fmt_number(columns = archive_size_mb, decimals = 1) |>
  sub_missing(missing_text = "-")
```

# Phase G: Remove Original Files

After successful conversion, verification, and archival packaging, the
original Excel and Stata files are removed from the working directories.
The originals are preserved in `active_raw_sources.zip` and can be
re-downloaded at any time by re-running `00_01_download_raw_sources.qmd`.

```{r}
#| label: g1-remove-originals

# Removal runs if: (a) enabled, (b) ZIP archive created successfully, and
# (c) either verification passed or was skipped because parquets already
# existed (nrow == 0 means nothing needed converting this run).
fidelity_ok <- nrow(verification) == 0 || all(verification$fidelity_ok)

if (remove_originals_after_conversion && fidelity_ok && active_pkg$success) {

  # Survey originals (ENCOVI xlsx, SIVESNU dta) from conversion manifest
  survey_files <- conversion_manifest |> pull(original_path)

  # Reference source xlsx (INE births, WHO growth) from unification phases
  ine_xlsx <- list.files(ine_source_dir, pattern = "\\.xlsx$",
                         full.names = TRUE)
  who_xlsx <- list.files(who_src, pattern = "\\.xlsx$",
                         full.names = TRUE)

  files_to_remove <- c(survey_files, ine_xlsx, who_xlsx) |>
    keep(file.exists)

  if (length(files_to_remove) > 0) {
    files_removed <- file.remove(files_to_remove)

    tibble(
      filename = basename(files_to_remove),
      removed = files_removed
    ) |>
      gt() |>
      tab_header(
        title = "Phase G - Original file removal",
        subtitle = "Files removed after successful conversion and archival"
      ) |>
      cols_label(filename = "File", removed = "Removed")
  } else {
    tibble(status = "No original files found to remove (already cleaned up).") |>
      gt() |>
      tab_header(title = "Phase G - Original file removal")
  }

} else if (!remove_originals_after_conversion) {

  tibble(
    status = paste(
      "Original file removal is disabled.",
      "Set remove_originals_after_conversion <- TRUE to enable."
    )
  ) |>
    gt() |>
    tab_header(title = "Phase G - Original file removal")
}
```


# Summary

```{r}
#| label: tbl-summary
#| tbl-cap: "Script execution summary"

unified_files <- c(
  ine_output_path,
  who_monthly_path,
  who_daily_path,
  who_ref_path
)

summary_data <- tribble(
  ~metric,                              ~value,
  "Survey sources converted (1:1)",     as.character(n_distinct(conversion_manifest$source_label)),
  "Active files in manifest",           as.character(nrow(conversion_manifest)),
  "Files converted (this run)",         as.character(nrow(files_to_convert)),
  "Files already cached",              as.character(sum(!conversion_manifest$needs_conversion)),
  "Unified reference parquets",         as.character(sum(file.exists(unified_files))),
  "Fidelity verifications passed",      as.character(if (nrow(verification) > 0) sum(verification$fidelity_ok) else 0L),
  "Hash log available",                 as.character(file.exists(hash_log_path)),
  "Active sources archive",             as.character(active_pkg$success),
  "Originals removed",                  as.character(remove_originals_after_conversion)
)

summary_data |>
  gt() |>
  tab_header(
    title = "Execution summary",
    subtitle = "00_02_convert_to_parquet.qmd"
  ) |>
  cols_label(metric = "Metric", value = "Value")
```

# Notes

## Methodology

This script performs three types of operations:

**One-to-one conversions** (Phase A): Format conversion only — no data
transformations, filtering, recoding, or derived variables. The list
of files to convert is read from `raw_sources.yml` (only entries with
`active: true`), ensuring that only the files required by the pipeline
are converted. The Parquet files are byte-level representations of the
original tabular content read by `readxl::read_excel()` or
`haven::read_dta()`.

**Many-to-one unifications** (Phases B-C): Multiple source files are
combined into a single Parquet table with an identifying column (`anio`
for INE, `sexo` for WHO). Minimal harmonization is applied:

- INE: sheet names and header rows are normalized across years;
  department names are standardized to match the pipeline convention
- WHO monthly: month 24 is deduplicated (retained from the 0-2 file,
  removed from the 2-5 file)
- WHO daily and reference: combined without modification

**Archival packaging** (Phase F): After conversion and fidelity
verification, the original files are bundled into a ZIP archive for
audit purposes. The active sources archive contains the exact bytes
used by the framework. Packaging is performed here (rather than in
`00_01`) so that the audit bundle accompanies verified conversion
results. After packaging, the original files are removed from the
working directories (Phase G).

## Limitations

If the original files have been removed (Phase G or manual deletion),
the fidelity verification in Phase D cannot run. The hash log from
`00_01` remains the primary provenance artifact.

Stata `.dta` files may contain labeled values (factor levels with
numeric codes). `haven::read_dta()` preserves these as `haven_labelled`
vectors, which Arrow writes as their underlying numeric type with
labels stored in Parquet metadata. Downstream scripts handle this
correctly via `haven::as_factor()` or explicit recoding.
