---
title: "Download and Verify Raw Data Sources"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script provisions the `01_data/01_raw/` directory with the original
data sources required by the framework. The workflow:

1. Reads the source manifest from `02_code/_config/raw_sources.yml`
2. Downloads each file from its authoritative public URL
3. Extracts the SIVESNU 2018 datasets from the SIINSAN omnibus archive
4. Moves staged files to their final source-specific directories
5. Verifies the SHA256 hash of each active file against the manifest
6. Writes a CSV log of all hashes and download metadata
7. Cleans up the temporary download staging directory

Archival packaging of the active originals into a ZIP bundle is
performed by `00_02_convert_to_parquet.qmd` after conversion and
fidelity verification.

This script is intended to be executed once during initial setup, or
whenever the source manifest is updated. Subsequent analyses read the
unpacked originals directly from `01_data/01_raw/{SOURCE}/`.

::: {.callout-note}
## Hash verification mode
The default behaviour is strict: any mismatch between the downloaded
file's hash and the manifest aborts the script. Set `strict_hash <- FALSE`
in the configuration block below for a permissive run that warns instead
of aborting (use only for debugging or when the source has been verified
to publish the same content under a new technical identifier).
:::

# Setup

```{r}
#| label: setup

rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

pacman::p_load(
  yaml,           # Manifest parsing
  rio,            # Data import/export
  here,           # Robust file paths
  janitor,        # Data cleaning utilities
  gt,             # Publication tables
  curl,           # HTTP downloads in download_with_retry()
  digest,         # SHA256 hashing in compute_file_hash()
  fs,             # Filesystem operations in IO helpers
  tidyverse       # Data manipulation and visualisation (includes purrr, stringr)
)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

# Set display options
options(scipen = 999, digits = 4)

# Load IO utility functions
source(here("02_code", "00_functions", "00_functions_provisioning.R"))
```

# Configuration

```{r}
#| label: configuration

# Hash verification mode
#   TRUE  - abort if any file's hash does not match the manifest
#   FALSE - log mismatches and continue (debugging / known-OK content)
strict_hash <- TRUE

# Force re-download even if the file already exists locally
force_download <- FALSE

# Directories
raw_root        <- here("01_data", "01_raw")
tmp_download    <- here("01_data", "_tmp_download")

# Output: hash log
hash_log_path <- here("01_data", "01_raw", "_hashes_active.csv")
```

# Load Manifest

```{r}
#| label: load-manifest

manifest <- yaml::read_yaml(
  here("02_code", "_config", "raw_sources.yml")
)

manifest_flat <- flatten_manifest(manifest) |>
  mutate(
    target_dir = case_when(
      source_label == "ENCOVI_2023"     ~ here(raw_root, "ENCOVI_2023"),
      source_label == "SIVESNU"         ~ here(raw_root, "SIVESNU_2018"),
      source_label == "INE_NACIMIENTOS" ~ here(raw_root, "INE_NACIMIENTOS"),
      source_label == "WHO_GROWTH"      ~ here(raw_root, "WHO_GROWTH"),
      TRUE                              ~ NA_character_
    ),
    target_path = file.path(target_dir, filename)
  )
```

```{r}
#| label: tbl-manifest-summary
#| tbl-cap: "Source manifest summary by data source"

manifest_flat |>
  summarise(
    n_files = n(),
    n_active = sum(active),
    total_size_mb = sum(size_mb, na.rm = TRUE),
    .by = source_label
  ) |>
  gt() |>
  tab_header(
    title = "Manifest summary",
    subtitle = "Files declared per source, with active subset and total volume"
  ) |>
  fmt_number(columns = total_size_mb, decimals = 1) |>
  cols_label(
    source_label = "Source",
    n_files = "Total files",
    n_active = "Active",
    total_size_mb = "Total size (MB)"
  )
```

# Phase A: Download Direct Files

ENCOVI 2023, INE Nacimientos, and WHO Growth Reference are each composed
of multiple files with direct download URLs. Downloads occur into a
temporary staging directory and are moved to the final location only
after successful hash verification.

```{r}
#| label: a1-prepare-directories

fs::dir_create(tmp_download)
walk(unique(na.omit(manifest_flat$target_dir)), fs::dir_create)
```

```{r}
#| label: a2-direct-downloads

direct_files <- manifest_flat |>
  filter(source_label %in% c("ENCOVI_2023", "INE_NACIMIENTOS",
                              "WHO_GROWTH")) |>
  filter(active)

direct_results <- direct_files |>
  mutate(
    staged_path = file.path(tmp_download, filename),
    needs_download = force_download | !file.exists(target_path)
  ) |>
  mutate(
    download_outcome = pmap(
      list(url, staged_path, needs_download),
      function(u, p, n) {
        if (!n) {
          return(tibble(success = TRUE, attempts = 0L,
                        error_message = NA_character_))
        }
        download_with_retry(url = u, dest_path = p) |>
          select(success, attempts, error_message)
      }
    )
  ) |>
  unnest(download_outcome)
```

```{r}
#| label: tbl-a2-direct-download-status
#| tbl-cap: "Direct download status by file"

direct_results |>
  mutate(
    status = case_when(
      !needs_download         ~ "cached",
      success                 ~ "downloaded",
      TRUE                    ~ "failed"
    )
  ) |>
  select(source_label, filename, status, attempts, error_message) |>
  gt() |>
  tab_header(
    title = "Phase A - Direct downloads",
    subtitle = "Files retrieved from authoritative public URLs"
  ) |>
  cols_label(
    source_label = "Source",
    filename = "File",
    status = "Status",
    attempts = "Attempts",
    error_message = "Error (if any)"
  ) |>
  sub_missing(missing_text = "-") |>
  data_color(
    columns = status,
    palette = c("downloaded" = "#3AAF85",
                "cached"     = "#1B6CA8",
                "failed"     = "#C44441")
  )
```

```{r}
#| label: a3-direct-download-guard

n_failed <- direct_results |>
  filter(needs_download, !success) |>
  nrow()

if (n_failed > 0) {
  stop(sprintf(
    "Phase A halted: %d file(s) failed to download. Review the table above.",
    n_failed
  ))
}
```

# Phase B: Acquire SIVESNU Omnibus Archive

The SIINSAN portal ships SIVESNU 2013-2018 as a single ZIP archive. The
script downloads it into the staging directory, then extracts only the
years needed.

```{r}
#| label: b1-download-sivesnu-zip

sivesnu_zip_url <- manifest$sivesnu$archive_url
sivesnu_zip_staged <- file.path(tmp_download, "BASE-DE-DATOS-SIVESNU.zip")

sivesnu_zip_status <- if (!file.exists(sivesnu_zip_staged) ||
                          force_download) {
  download_with_retry(url = sivesnu_zip_url,
                      dest_path = sivesnu_zip_staged,
                      max_attempts = 3,
                      timeout_sec = 600)
} else {
  tibble(url = sivesnu_zip_url, dest_path = sivesnu_zip_staged,
         success = TRUE, attempts = 0L, error_message = NA_character_)
}
```

```{r}
#| label: tbl-b1-sivesnu-zip-status
#| tbl-cap: "SIVESNU omnibus ZIP acquisition"

sivesnu_zip_status |>
  mutate(
    archive_size_mb = if (file.exists(sivesnu_zip_staged)) {
      round(as.numeric(fs::file_size(sivesnu_zip_staged)) / 1024^2, 2)
    } else NA_real_
  ) |>
  select(url, success, attempts, archive_size_mb, error_message) |>
  gt() |>
  tab_header(
    title = "Phase B - SIVESNU archive download",
    subtitle = "Single ZIP containing years 2013-2018"
  ) |>
  cols_label(
    url = "URL",
    success = "Success",
    attempts = "Attempts",
    archive_size_mb = "Archive size (MB)",
    error_message = "Error"
  ) |>
  sub_missing(missing_text = "-")

if (!sivesnu_zip_status$success) {
  stop("Phase B halted: SIVESNU archive could not be downloaded.")
}
```

```{r}
#| label: b2-extract-sivesnu-2018

sivesnu_2018_dir <- here(raw_root, "SIVESNU_2018")

extract_2018 <- extract_zip_selective(
  zip_path = sivesnu_zip_staged,
  internal_filter = "2018/[Dd]atasets/.*\\.dta$",
  dest_dir = sivesnu_2018_dir
)
```

```{r}
#| label: tbl-b-extraction-summary
#| tbl-cap: "SIVESNU selective extraction (2018 active)"

extract_2018 |>
  mutate(filename = basename(dest_path)) |>
  select(filename, extracted) |>
  gt() |>
  tab_header(
    title = "Phase B - SIVESNU extraction",
    subtitle = "Selective unpack of 2018 .dta files"
  ) |>
  cols_label(
    filename = "File",
    extracted = "Extracted"
  )
```

# Phase C: Move Direct Files to Final Locations

Direct downloads from Phase A are validated and moved from the staging
directory to their final source-specific folders.

```{r}
#| label: c1-move-direct-files

stage_to_final <- direct_results |>
  filter(needs_download, success) |>
  mutate(
    moved = pmap_lgl(
      list(staged_path, target_path),
      function(src, dst) {
        if (!file.exists(src)) return(FALSE)
        fs::file_copy(src, dst, overwrite = TRUE)
        fs::file_delete(src)
        file.exists(dst)
      }
    )
  )
```

# Phase D: Verify Hashes

All active files (downloaded direct + extracted from SIVESNU) are
hashed and compared against the manifest expectations.

```{r}
#| label: d1-verify-active-hashes

active_for_check <- manifest_flat |> filter(active)

verification <- active_for_check |>
  mutate(
    check = pmap(
      list(target_path, sha256),
      function(p, h) verify_hash_sha256(p, h)
    )
  ) |>
  unnest(check, names_sep = "_") |>
  select(source_label, filename, sha256, check_observed_hash, check_status)
```

```{r}
#| label: tbl-d1-hash-verification
#| tbl-cap: "Hash verification status by file (active sources only)"

verification |>
  rename(
    expected = sha256,
    observed = check_observed_hash,
    status = check_status
  ) |>
  mutate(
    expected_short = if_else(is.na(expected),
                             NA_character_,
                             paste0(substr(expected, 1, 12), "...")),
    observed_short = if_else(is.na(observed),
                             NA_character_,
                             paste0(substr(observed, 1, 12), "..."))
  ) |>
  select(source_label, filename, expected_short, observed_short, status) |>
  gt() |>
  tab_header(
    title = "Phase D - Hash verification",
    subtitle = "SHA256 comparison against manifest expectations"
  ) |>
  cols_label(
    source_label = "Source",
    filename = "File",
    expected_short = "Expected (manifest)",
    observed_short = "Observed (downloaded)",
    status = "Status"
  ) |>
  sub_missing(missing_text = "-") |>
  data_color(
    columns = status,
    palette = c("match"     = "#3AAF85",
                "no_expect" = "#1B6CA8",
                "mismatch"  = "#C44441",
                "missing"   = "#C44441")
  )
```

```{r}
#| label: d2-strict-hash-guard

n_mismatch <- sum(verification$check_status == "mismatch")
n_missing  <- sum(verification$check_status == "missing")

if (n_missing > 0) {
  stop(sprintf(
    "Phase D halted: %d expected file(s) are missing on disk.",
    n_missing
  ))
}

if (n_mismatch > 0 && strict_hash) {
  stop(sprintf(paste(
    "Phase D halted: %d file(s) failed hash verification under strict mode.",
    "Review the table above and either (a) update the manifest after",
    "manually confirming the new content is intended, or (b) re-run with",
    "strict_hash <- FALSE if you have already confirmed the content is",
    "equivalent."), n_mismatch))
}
```

# Phase E: Cleanup Temporary Files

The staging directory is removed unconditionally. All active files have
already been moved to their final directories in Phase C.

```{r}
#| label: e1-cleanup-tmp

cleanup_tmp_download(tmp_download)
```

# Save Outputs

```{r}
#| label: save-hash-log

hash_log <- verification |>
  filter(check_status %in% c("match", "no_expect")) |>
  rename(
    sha256_expected = sha256,
    sha256_observed = check_observed_hash,
    status = check_status
  ) |>
  left_join(
    manifest_flat |> select(filename, size_mb, recorded_on, citation),
    by = "filename"
  ) |>
  select(source_label, filename, sha256_observed, sha256_expected,
         size_mb, status, recorded_on, citation)

rio::export(
  hash_log,
  hash_log_path
)
```

```{r}
#| label: tbl-save-final-summary
#| tbl-cap: "Final outputs persisted on disk"

tibble(
  output = "Hash log (CSV)",
  path = hash_log_path,
  exists = file.exists(hash_log_path)
) |>
  gt() |>
  tab_header(
    title = "Final outputs",
    subtitle = "Files persisted on disk after script execution"
  ) |>
  cols_label(
    output = "Output",
    path = "Path",
    exists = "Exists"
  )
```

# Notes

## Methodology

The manifest separates active sources (consumed by the framework
pipeline) from inactive sources (retained in the manifest for reference).
The ENCOVI 2023 active subset is restricted to the five files the
current pipeline reads; the SIVESNU omnibus ZIP is selectively extracted
to populate `01_data/01_raw/SIVESNU_2018/` with the four active .dta
files.

Hash verification operates in strict mode by default: any divergence
between the manifest's expected SHA256 and the observed value aborts
the script. This protects the analytical pipeline against silent
upstream changes (a publisher re-uploading a file with the same URL
but different content). The permissive mode exists for the rare case
where the publisher has been verified to publish identical content
under a new technical artefact (e.g., metadata-only change).

Archival packaging of the active originals into a ZIP bundle is handled
by `00_02_convert_to_parquet.qmd`, which creates the archive after
conversion and fidelity verification.

## Data Quality

All active hashes were recorded from the same machine on the dates
listed in the manifest's `recorded_on` field. WHO source URLs include
a `?sfvrsn=` cache-busting parameter that may rotate when WHO
republishes. Such rotation does not affect content as long as the
SHA256 hash matches the manifest; if it does not, the strict-mode
guard will surface the change.

INE Guatemala URLs use opaque hashes in their path. These URLs may
change if the INE re-uploads a file. The `landing_url` field of the
manifest provides the stable public page where current URLs can be
re-discovered if a download fails.

## Limitations

The SIVESNU omnibus archive is large (~95 MB) and downloads in a
single transfer. On unstable networks, the timeout in
`download_with_retry()` may need adjustment via the `timeout_sec`
argument.

## References

INE Guatemala data portal (CC BY-SA 4.0):
<https://www.ine.gob.gt/pobreza-menu/>,
<https://www.ine.gob.gt/vitales/>.

SIINSAN - SIVESNU databases:
<https://portal.siinsan.gob.gt/monitoreo-y-evaluacion/>.

WHO Child Growth Standards:
<https://www.who.int/tools/child-growth-standards/standards/length-height-for-age>.

WHO Growth Reference 5-19:
<https://www.who.int/tools/growth-reference-data-for-5to19-years>.
