---
title: "Build Nutrition Reference Tables"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script builds the two typed Parquet tables that the analytical
pipeline reads, from the single editable source workbook
`incap_finut_master.xlsx`:

1. Read the `food_composition` sheet (one row per food) and write it
   as a typed Parquet table
2. Read the `food_equivalences` sheet (one row per food and unit) and
   write it as a typed Parquet table
3. Validate row counts, key uniqueness, and column types for both
   tables

The Parquet tables are byte-for-byte regenerable from the workbook.
The workbook is the source of truth; values are edited there and this
script is re-run to refresh the Parquet outputs. This script applies
no normalization or imputation: it transfers the workbook values as
they are, preserving the per-100 g basis of the composition table.

# Setup

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,          # Parquet file handling
  readxl,         # Excel sheet reading
  here,           # Robust file path management
  janitor,        # Data cleaning utilities
  gt,             # Table formatting and presentation
  tidyverse       # General purpose data manipulation
)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

# Set display options
options(scipen = 999, digits = 4)

source(here("02_code", "00_functions", "00_functions.R"))

# external/
ensure_output_dir(here("01_data", "02_processed", "external"))
```

# Load Source Workbook

```{r}
#| label: load-master

master_path <- here("01_data", "03_external", "incap_finut_master.xlsx")

# Composition: one row per food (codine). Values per 100 g edible
# portion, as transcribed from the source tables.
df_composition <- read_excel(master_path, sheet = "food_composition") |>
  clean_names()

# Equivalences: one row per food and presentation unit. The numeric
# measurement code is already provided in the workbook.
df_equivalences <- read_excel(master_path, sheet = "food_equivalences") |>
  clean_names()
```

# Build Tables

```{r}
#| label: build-composition

# Enforce column types for the composition table
df_food_composition <- df_composition |>
  mutate(
    ali = as.integer(ali),
    codine = as.integer(codine),
    nombre_alimento = as.character(nombre_alimento),
    across(
      c(energ_kcal, proteina_g, hierro_mg, zinc_mg, pct_aprov,
        lys_g, met_g, cys_g, thr_g, trp_g,
        pdcaas, npu, pdcaas_qpm, npu_qpm),
      as.numeric
    )
  ) |>
  arrange(codine)
```

```{r}
#| label: build-equivalences

# Enforce column types for the equivalences table; preserve row order
df_food_equivalences <- df_equivalences |>
  mutate(
    codine = as.integer(codine),
    alimento = as.character(alimento),
    vegetal_cereal = as.character(vegetal_cereal),
    medida_alimento = as.character(medida_alimento),
    medida = as.integer(medida),
    gramos = as.numeric(gramos),
    n = as.integer(n)
  )
```

# Validation

```{r}
#| label: tbl-validation-summary
#| tbl-cap: "Built nutrition tables overview"

tbl_validation <- tibble(
  Table = c("food_composition", "food_equivalences"),
  Rows = c(nrow(df_food_composition), nrow(df_food_equivalences)),
  `Unique foods (codine)` = c(
    n_distinct(df_food_composition$codine),
    n_distinct(df_food_equivalences$codine)
  ),
  `Codine is unique` = c(
    !any(duplicated(df_food_composition$codine)),
    NA
  )
)

tbl_validation |>
  gt() |>
  tab_header(
    title = "Nutrition Reference Tables",
    subtitle = "Built from incap_finut_master.xlsx"
  ) |>
  fmt_number(columns = c(Rows, `Unique foods (codine)`),
             decimals = 0, use_seps = TRUE) |>
  sub_missing(missing_text = "—") |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-composition-coverage
#| tbl-cap: "Composition table column coverage (non-missing values)"

tbl_coverage <- df_food_composition |>
  summarise(across(
    c(energ_kcal, proteina_g, hierro_mg, zinc_mg, pct_aprov,
      lys_g, met_g, cys_g, thr_g, trp_g, pdcaas, npu,
      pdcaas_qpm, npu_qpm),
    ~ sum(!is.na(.x))
  )) |>
  pivot_longer(everything(), names_to = "Column", values_to = "Non-missing") |>
  mutate(Total = nrow(df_food_composition))

tbl_coverage |>
  gt() |>
  tab_header(
    title = "Composition Column Coverage",
    subtitle = "Non-missing value count per nutrient column"
  ) |>
  fmt_number(columns = c(`Non-missing`, Total), decimals = 0) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

# Save Outputs

```{r}
#| label: save-outputs

# Composition table (one row per food)
write_parquet(
  df_food_composition,
  here("01_data", "03_external", "incap_food_composition.parquet")
)

# Equivalences table (one row per food and unit)
write_parquet(
  df_food_equivalences,
  here("01_data", "03_external", "incap_food_equivalences.parquet")
)
```

# Notes

## Source and Lineage

The source of truth is `incap_finut_master.xlsx` in `03_external`. Its
`food_composition` and `food_equivalences` sheets are transferred here
to typed Parquet tables without modification of values. Provenance for
each protein-quality value is documented in the workbook's
`source_notes` sheet, and the directory `README.md`
describes the three provenance layers (INCAP composition base, FINUT
protein-quality layer, Azahar Data Insights derived layer).

## Build Behaviour

This script transfers values as stored in the workbook. Nutrient
values in `food_composition` remain on a per-100 g basis. The
per-gram normalization and the conservative PDCAAS imputation for
foods without a score are applied downstream in Script 01_01, where
they belong as analytical decisions of the pipeline.

## Outputs

- `incap_food_composition.parquet` — one row per food (`codine`),
  consumed by Script 01_01.
- `incap_food_equivalences.parquet` — one row per food and
  presentation unit, consumed by Script 01_01. Row order follows the
  workbook (foods in ascending `codine` order of first appearance);
  Script 01_02 derives food factor levels from this order.
