---
title: "Transfer Model: Digestible Protein Intake"
subtitle: "Module 3: Transfer Models"
description-meta: "Survey-weighted gamma regression for digestible protein intake, with PDCAAS applied at food level so protein is modeled directly, not adjusted after."
---

## Why Model Digestible Protein?

Protein quality, not just quantity, is an important determinant of child growth
outcomes. The outcome modeled here is **digestible protein** (PDCAAS-adjusted):
the PDCAAS (Protein Digestibility Corrected Amino Acid Score) accounts for both
amino acid composition and digestibility, so the adjusted figure represents the
protein fraction available for human use. By modeling digestible protein from
maize and non-maize sources separately, we can:

1. **Simulate biofortification scenarios** — Predict how higher-quality maize varieties 
   would change total digestible protein intake
2. **Preserve baseline non-maize intake** — Keep other dietary sources constant 
   when modeling interventions
3. **Account for protein quality** — PDCAAS is an intrinsic property of each food, 
   enabling direct impact assessment without separate bioavailability adjustments

::: {.callout-important}
## Methodological Note
Protein quality (PDCAAS) is incorporated at the food level during intake
calculation, which allows digestible protein to be modeled directly; for
micronutrients, bioavailability is applied after modeling. PDCAAS values are
approximate — they vary by maize variety, processing method, and dietary context.
Our analysis uses representative values for processed maize in Guatemalan diets. 
See [how PDCAAS was computed](../01_nutritional_intakes/01_01_load_and_prepare_base_data.qmd#incap-2014-food-composition-tables)
for the source and imputation rule applied to foods without computed values.
:::

## Modeling Approach

We use **survey-weighted gamma regression** because digestible protein intake data is:

- **Positive and continuous** — People consume some amount of protein daily
- **Right-skewed** — Most people have moderate intake, few have very high intake
- **Population-representative** — Survey weights carry the estimates to the
  national level

The models predict digestible protein intake based on demographic and socioeconomic 
characteristics that are available in both ENCOVI (training data) and 
SIVESNU (application target).

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  fitdistrplus,       # Distribution fitting for goodness-of-fit assessment
  ggpubr,             # Arrange diagnostic plots into a 2x2 grid
  survey,             # Required for svyglm method dispatch (model diagnostics)
  performance,        # Pseudo-R² (Nagelkerke / Cragg-Uhler) and diagnostics
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))

# --- Local helpers: distribution assessment (web formatting) ---------------
# Recreate the distribution-fit outputs from the raw outcome vector exported
# by the framework, returning table and diagnostic grid as separate objects
# so each renders in its own chunk with web-consistent formatting.

build_distribution_table <- function(outcome) {

  fit_gamma       <- fitdist(data = outcome, "gamma")
  fit_lognormal   <- fitdist(data = outcome, "lnorm")
  fit_exponential <- fitdist(data = outcome, "exp")

  data.frame(
    distribution = c("Gamma", "Lognormal", "Exponential"),
    loglik = c(fit_gamma$loglik, fit_lognormal$loglik, fit_exponential$loglik),
    aic    = c(fit_gamma$aic, fit_lognormal$aic, fit_exponential$aic),
    bic    = c(fit_gamma$bic, fit_lognormal$bic, fit_exponential$bic),
    ks_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$ks,
      gofstat(fit_lognormal, fitnames = "lnorm")$ks,
      gofstat(fit_exponential, fitnames = "exp")$ks
    ),
    cvm_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$cvm,
      gofstat(fit_lognormal, fitnames = "lnorm")$cvm,
      gofstat(fit_exponential, fitnames = "exp")$cvm
    ),
    ad_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$ad,
      gofstat(fit_lognormal, fitnames = "lnorm")$ad,
      gofstat(fit_exponential, fitnames = "exp")$ad
    )
  ) |>
    gt() |>
    tab_header(
      title = "Distribution Fit Comparison",
      subtitle = "Goodness-of-fit statistics for Gamma, Lognormal, and Exponential distributions"
    ) |>
    cols_label(
      distribution = "Distribution",
      loglik   = "Log-Likelihood",
      aic      = "AIC",
      bic      = "BIC",
      ks_stat  = "Kolmogorov-Smirnov",
      cvm_stat = "Cramer-von Mises",
      ad_stat  = "Anderson-Darling"
    ) |>
    fmt_number(
      columns = c(loglik, aic, bic, ks_stat, cvm_stat, ad_stat),
      decimals = 4
    ) |>
    tab_spanner(label = "Information Criteria", columns = c(aic, bic)) |>
    tab_spanner(label = "Goodness-of-Fit Statistics", columns = c(ks_stat, cvm_stat, ad_stat)) |>
    cols_align(align = "center", columns = everything()) |>
    cols_align(align = "left", columns = distribution) |>
    tab_options(
      table.font.size = 11,
      heading.title.font.size = 13,
      heading.subtitle.font.size = 10,
      table.width = pct(100)
    )
}

build_distribution_plots <- function(outcome) {

  fit_gamma <- fitdist(data = outcome, "gamma")

  shape <- fit_gamma$estimate["shape"]
  rate  <- fit_gamma$estimate["rate"]
  x_obs <- sort(fit_gamma$data)

  plot_density <- data.frame(x = fit_gamma$data) |>
    ggplot(aes(x = x)) +
    geom_histogram(aes(y = after_stat(density)), bins = 30,
                   color = "azure4", fill = "lightgray", alpha = 0.5) +
    stat_function(fun = dgamma, args = list(shape = shape, rate = rate),
                  color = "#3AAF85", linetype = "dashed", linewidth = 0.8) +
    labs(title = "Empirical vs Theoretical Density", x = "", y = "Density") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      plot.subtitle = element_text(color = "azure4", size = 10),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_qq <- data.frame(
    empirical = x_obs,
    theoretical = qgamma(ppoints(length(x_obs)), shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = theoretical, y = empirical)) +
    geom_point(color = "#1B6CA8", size = 1, alpha = 0.2) +
    geom_abline(intercept = 0, slope = 1, color = "#3AAF85",
                linetype = "dashed", linewidth = 0.8) +
    labs(title = "Quantile-Quantile Plot", x = "Theoretical Quantiles", y = "Empirical Quantiles") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_cdf <- data.frame(
    x = x_obs,
    empirical_cdf = ecdf(fit_gamma$data)(x_obs),
    theoretical_cdf = pgamma(x_obs, shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = x)) +
    geom_line(aes(y = empirical_cdf), color = "lightgray", linewidth = 2.5) +
    geom_line(aes(y = theoretical_cdf), color = "#3AAF85",
              linetype = "dashed", linewidth = 0.8) +
    labs(title = "Empirical vs Theoretical CDF",
         subtitle = "CDF: Cumulative Distribution Function", x = "Data", y = "CDF") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      plot.subtitle = element_text(color = "azure4", size = 10),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_pp <- data.frame(
    empirical_prob = ecdf(fit_gamma$data)(x_obs),
    theoretical_prob = pgamma(x_obs, shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = theoretical_prob, y = empirical_prob)) +
    geom_point(color = "#1B6CA8", size = 1, alpha = 0.2) +
    geom_abline(intercept = 0, slope = 1, color = "#3AAF85",
                linetype = "dashed", linewidth = 0.8) +
    labs(title = "P-P Plot", x = "Theoretical Probabilities", y = "Empirical Probabilities") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  ggarrange(plot_density, plot_qq, plot_cdf, plot_pp, ncol = 2, nrow = 2)
}
```

```{r}
#| label: load-rds
#| include: false

# (a) Web resources for visualizations (exported by framework 03_04)
fig_b1_evaluate_distribution_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_fig_b1_evaluate_distribution_maize.rds"),
  trust = TRUE
)

tbl_b2_coefficients_full_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_tbl_b2_coefficients_full_model_maize.rds"),
  trust = TRUE
)

fig_b2_diagnostic_plot_full_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_fig_b2_diagnostic_plot_full_model_maize.rds"),
  trust = TRUE
)

tbl_b3_wald_test_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_tbl_b3_wald_test_maize.rds"),
  trust = TRUE
)

tbl_b4_coefficients_refined_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_tbl_b4_coefficients_refined_model_maize.rds"),
  trust = TRUE
)

fig_b4_diagnostic_plot_refined_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_fig_b4_diagnostic_plot_refined_model_maize.rds"),
  trust = TRUE
)

fig_b5_evaluate_diagnostics_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_fig_b5_evaluate_diagnostics_maize.rds"),
  trust = TRUE
)

fig_c1_evaluate_distribution_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_fig_c1_evaluate_distribution_non_maize.rds"),
  trust = TRUE
)

tbl_c2_coefficients_full_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_tbl_c2_coefficients_full_model_non_maize.rds"),
  trust = TRUE
)

fig_c2_diagnostic_plot_full_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_fig_c2_diagnostic_plot_full_model_non_maize.rds"),
  trust = TRUE
)

tbl_c3_wald_test_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_tbl_c3_wald_test_non_maize.rds"),
  trust = TRUE
)

tbl_c4_coefficients_refined_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_tbl_c4_coefficients_refined_model_non_maize.rds"),
  trust = TRUE
)

fig_c4_diagnostic_plot_refined_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_fig_c4_diagnostic_plot_refined_model_non_maize.rds"),
  trust = TRUE
)

fig_c5_evaluate_diagnostics_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_04_fig_c5_evaluate_diagnostics_non_maize.rds"),
  trust = TRUE
)

# (b) Fitted models for performance metrics
models <- import(
  here("..", "01_data", "02_processed", "models",
       "03_04_digestible_protein_models.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Significant predictor counts from Wald tests (maize and non-maize)
n_predictors_maize <- tbl_b3_wald_test_maize |>
  dplyr::filter(significant == "*") |>
  nrow()

n_predictors_non_maize <- tbl_c3_wald_test_non_maize |>
  dplyr::filter(significant == "*") |>
  nrow()

# Collinearity of the refined maize model, on the degrees-of-freedom adjusted
# VIF scale reported by the diagnostic grid
vif_maize <- adjust_collinearity_df(models$model_prot_apro_maize)
max_vif_maize <- round(max(vif_maize$VIF), 1)
n_vif_over_5 <- sum(vif_maize$VIF > 5)
n_vif_terms <- nrow(vif_maize)
```


## Data Preparation

### Source Data

We use the cleaned ENCOVI dataset from the previous preparation step, containing 
individuals with complete predictor variables and nutrient intake estimates.

### Handling Extreme Values

Gamma regression requires positive values and is sensitive to extreme outliers. 
We apply conservative outlier detection:

::: {.callout-tip}
## Outlier Threshold
Upper outliers identified using the standard **Tukey method (Q3 + 1.5×IQR)**. 
The threshold suits macronutrients:

- Protein intake distributions are more stable
- Extreme values are consistent with measurement error
- Macronutrient intakes have natural biological limits
:::

### Survey Design

All models incorporate ENCOVI's complex survey design (clustering and
expansion weights), which places the estimates at the population level.


## Modeling Digestible Protein from Maize

### Distribution Selection

Before modeling, we assess which statistical distribution best fits the 
digestible protein intake data. Three candidate distributions are compared:

```{r}
#| label: tbl-b1-evaluate-distribution-maize
#| tbl-cap: "Distribution fit comparison for digestible protein intake from maize"

build_distribution_table(fig_b1_evaluate_distribution_maize)
```

```{r}
#| label: fig-b1-evaluate-distribution-maize
#| fig-cap: "Distribution assessment for digestible protein intake from maize"
#| fig-alt: >
#|   Four goodness-of-fit panels arranged in a 2-by-2 grid comparing observed
#|   digestible protein from maize against a fitted gamma distribution. The
#|   top-left panel overlays a dashed gamma density curve on a histogram of the
#|   data; the top-right is a quantile-quantile plot; the bottom-left compares
#|   empirical and theoretical cumulative distribution functions; the bottom-right
#|   is a probability-probability plot. Points and curves track the diagonal
#|   reference lines closely, indicating the gamma distribution provides the best
#|   fit to this positive, right-skewed intake.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

build_distribution_plots(fig_b1_evaluate_distribution_maize)
```

::: {.callout-important}
## Distribution Selection
**Gamma distribution** provides the best fit based on:

- Lowest AIC/BIC (information criteria)
- Highest Log-Likelihood

Gamma regression with log link is appropriate for positive, right-skewed 
continuous outcomes.
:::

### Full Model

We first train a model including all 20 candidate predictors to identify 
which variables significantly predict digestible protein intake from maize.

```{r}
#| label: tbl-b2-coefficients-full-model-maize
#| tbl-cap: "Model coefficients for digestible protein intake from maize (full model)"

tbl_b2_coefficients_full_model_maize
```

```{r}
#| label: fig-b2-diagnostic-plot-full-model-maize
#| fig-cap: "Observed vs predicted digestible protein intake from maize (full model)"
#| fig-alt: >
#|   Scatter chart with observed digestible protein from maize in grams per day
#|   along the x-axis and the full model's predicted values along the y-axis.
#|   Points cluster along an upward diagonal trend, showing predictions rise with
#|   observed intake, though scatter around the line reflects the model's limited
#|   precision for individual-level values.

fig_b2_diagnostic_plot_full_model_maize |>
  ggplot(aes(x = prot_apro_individuo_dia_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Digestible Protein from Maize (g/day)",
    y = "Predicted Digestible Protein from Maize (g/day)"
  )
```

### Variable Selection

Not all predictors contribute meaningfully to prediction. We use **Wald tests** 
to identify statistically significant variables and remove those that add 
complexity without improving predictions.

```{r}
#| label: tbl-b3-wald-test-maize
#| tbl-cap: "Wald test results for predictor significance (digestible protein from maize)"

tbl_b3_wald_test_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = "Wald F-test results for digestible protein intake from maize model"
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-important}
## Variables Retained
**`r n_predictors_maize` predictors** show significant relationships with 
digestible protein intake from maize, spanning demographic, geographic, 
socioeconomic, housing and asset categories. Variables that do not 
significantly predict digestible protein intake from maize after controlling 
for other factors are excluded from the refined model.
:::

### Refined Model

The final model is fitted on the significant predictors alone.

```{r}
#| label: tbl-b4-coefficients-refined-model-maize
#| tbl-cap: "Model coefficients for digestible protein intake from maize (refined model)"

tbl_b4_coefficients_refined_model_maize
```

```{r}
#| label: fig-b4-diagnostic-plot-refined-model-maize
#| fig-cap: "Observed vs predicted digestible protein intake from maize (refined model)"
#| fig-alt: >
#|   Scatter chart with observed digestible protein from maize in grams per day
#|   along the x-axis and the refined model's predicted values along the y-axis.
#|   Points follow the same upward diagonal trend as the full model, confirming
#|   that dropping non-significant predictors preserves predictive accuracy while
#|   simplifying the model.

fig_b4_diagnostic_plot_refined_model_maize |>
  ggplot(aes(x = prot_apro_individuo_dia_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Digestible Protein from Maize (g/day)",
    y = "Predicted Digestible Protein from Maize (g/day)"
  )
```

### Model Diagnostics

We verify the model meets statistical assumptions:

```{r}
#| label: fig-b5-evaluate-diagnostics-maize
#| fig-cap: "Diagnostic plots for digestible protein from maize model"
#| fig-alt: >
#|   Grid of regression diagnostic panels for the refined maize gamma model,
#|   covering residuals, leverage, and collinearity. Residuals show no systematic
#|   pattern and are approximately normal, no observations exert undue leverage,
#|   and the collinearity panel plots each term's degrees-of-freedom adjusted
#|   variance inflation factor against shaded low, moderate and high bands.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

evaluate_gamma_model(models$model_prot_apro_maize)
```

::: {.callout-tip}
## Diagnostic Assessment
- **Residuals**: No systematic patterns, approximately normal
- **Leverage**: No overly influential observations
- **Collinearity**: Degrees-of-freedom adjusted variance inflation factors (VIF)
  reach a maximum of `r max_vif_maize`, with `r n_vif_over_5` of
  `r n_vif_terms` terms above the conventional threshold of 5

Factors and polynomial terms contribute several columns to the model matrix, so
their generalized VIF is rescaled by the term's degrees of freedom. This places
every term on the single-coefficient scale that the thresholds of 5 and 10
refer to.
:::


## Modeling Digestible Protein from Non-Maize Sources

Digestible protein from non-maize sources includes animal products, legumes, 
and other plant proteins. This component remains **constant** in biofortification 
scenarios, representing the baseline dietary protein that would not change with 
maize interventions.

### Distribution Selection

```{r}
#| label: tbl-c1-evaluate-distribution-non-maize
#| tbl-cap: "Distribution fit comparison for digestible protein intake from non-maize sources"

build_distribution_table(fig_c1_evaluate_distribution_non_maize)
```

```{r}
#| label: fig-c1-evaluate-distribution-non-maize
#| fig-cap: "Distribution assessment for digestible protein intake from non-maize sources"
#| fig-alt: >
#|   Four goodness-of-fit panels arranged in a 2-by-2 grid comparing observed
#|   digestible protein from non-maize sources against a fitted gamma
#|   distribution. The panels show a histogram with an overlaid dashed gamma
#|   density curve, a quantile-quantile plot, an empirical versus theoretical
#|   cumulative distribution function plot, and a probability-probability plot.
#|   Points and curves align closely with the diagonal reference lines, confirming
#|   the gamma distribution again fits best.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

build_distribution_plots(fig_c1_evaluate_distribution_non_maize)
```

::: {.callout-important}
## Distribution Selection
**Gamma distribution** again provides the best fit for non-maize digestible 
protein intake.
:::

### Full Model

```{r}
#| label: tbl-c2-coefficients-full-model-non-maize
#| tbl-cap: "Model coefficients for digestible protein intake from non-maize sources (full model)"

tbl_c2_coefficients_full_model_non_maize
```

```{r}
#| label: fig-c2-diagnostic-plot-full-model-non-maize
#| fig-cap: "Observed vs predicted digestible protein intake from non-maize sources (full model)"
#| fig-alt: >
#|   Scatter chart with observed digestible protein from non-maize sources in
#|   grams per day along the x-axis and the full model's predicted values along
#|   the y-axis. Points cluster along an upward diagonal trend, showing predictions
#|   increase with observed intake, with scatter around the line reflecting
#|   individual-level variability the model does not capture.

fig_c2_diagnostic_plot_full_model_non_maize |>
  ggplot(aes(x = prot_apro_individuo_dia_no_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Digestible Protein from Non-Maize (g/day)",
    y = "Predicted Digestible Protein from Non-Maize (g/day)"
  )
```

### Variable Selection

```{r}
#| label: tbl-c3-wald-test-non-maize
#| tbl-cap: "Wald test results for predictor significance (digestible protein from non-maize)"

tbl_c3_wald_test_non_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = "Wald F-test results for digestible protein intake from non-maize model"
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-important}
## Variables Retained
**`r n_predictors_non_maize` predictors** significantly predict digestible 
protein intake from non-maize sources, drawing from the same set of demographic, 
geographic, socioeconomic and housing variables. The retained set differs from the maize model: the two outcomes are predicted
by different subsets of the same candidate variables.
:::

### Refined Model

```{r}
#| label: tbl-c4-coefficients-refined-model-non-maize
#| tbl-cap: "Model coefficients for digestible protein intake from non-maize sources (refined model)"

tbl_c4_coefficients_refined_model_non_maize
```

```{r}
#| label: fig-c4-diagnostic-plot-refined-model-non-maize
#| fig-cap: "Observed vs predicted digestible protein intake from non-maize sources (refined model)"
#| fig-alt: >
#|   Scatter chart with observed digestible protein from non-maize sources in
#|   grams per day along the x-axis and the refined model's predicted values along
#|   the y-axis. Points follow the same upward diagonal trend as the full model,
#|   showing that restricting to significant predictors retains predictive accuracy
#|   while simplifying the model.

fig_c4_diagnostic_plot_refined_model_non_maize |>
  ggplot(aes(x = prot_apro_individuo_dia_no_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Digestible Protein from Non-Maize (g/day)",
    y = "Predicted Digestible Protein from Non-Maize (g/day)"
  )
```

### Model Diagnostics

```{r}
#| label: fig-c5-evaluate-diagnostics-non-maize
#| fig-cap: "Diagnostic plots for digestible protein from non-maize sources model"
#| fig-alt: >
#|   Grid of regression diagnostic panels for the refined non-maize gamma model,
#|   covering residuals, leverage, and collinearity. Residuals show no systematic
#|   pattern and are approximately normal, no observations are overly influential,
#|   and the collinearity panel plots each term's degrees-of-freedom adjusted
#|   variance inflation factor against shaded low, moderate and high bands.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

evaluate_gamma_model(models$model_prot_apro_non_maize)
```


## Summary

### Final Model Performance

```{r}
#| label: tbl-model-performance
#| tbl-cap: "Final refined model performance metrics"

dplyr::bind_rows(
  extract_gamma_model_metrics(models$model_prot_apro_maize,     "Digestible protein from maize"),
  extract_gamma_model_metrics(models$model_prot_apro_non_maize, "Digestible protein from non-maize")
) |>
  # Reshape: metrics as rows, models as columns
  tidyr::pivot_longer(
    cols = -source,
    names_to = "metric",
    values_to = "value"
  ) |>
  tidyr::pivot_wider(
    names_from = source,
    values_from = value
  ) |>
  # Apply readable labels per metric
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      n_obs                 = "N observations",
      n_predictors          = "N predictors",
      pseudo_r2_mcfadden    = "Pseudo-R² (McFadden)",
      pseudo_r2_cragg_uhler = "Pseudo-R² (Cragg-Uhler)",
      aic                   = "AIC",
      dispersion            = "Dispersion"
    )
  ) |>
  gt() |>
  cols_label(metric = "Metric") |>
  fmt_number(
    rows    = metric == "N observations",
    columns = c(`Digestible protein from maize`, `Digestible protein from non-maize`),
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    rows    = metric == "N predictors",
    columns = c(`Digestible protein from maize`, `Digestible protein from non-maize`),
    decimals = 0
  ) |>
  fmt_number(
    rows    = metric %in% c("Pseudo-R² (McFadden)",
                            "Pseudo-R² (Cragg-Uhler)",
                            "Dispersion"),
    columns = c(`Digestible protein from maize`, `Digestible protein from non-maize`),
    decimals = 3
  ) |>
  fmt_number(
    rows    = metric == "AIC",
    columns = c(`Digestible protein from maize`, `Digestible protein from non-maize`),
    decimals = 0
  ) |>
  tab_header(
    title    = "Final Model Performance",
    subtitle = "Refined gamma regression with log link"
  ) |>
  tab_footnote(
    footnote  = "Pseudo-R² (McFadden) = 1 − (residual deviance / null deviance). Conservative measure; values above 0.2 are conventionally considered good for GLMs.",
    locations = cells_body(columns = metric, rows = metric == "Pseudo-R² (McFadden)")
  ) |>
  tab_footnote(
    footnote  = "Pseudo-R² (Cragg-Uhler), also known as Nagelkerke's R², is normalized to a [0, 1] range and generally yields higher values than McFadden. Both measures are reported for transparency.",
    locations = cells_body(columns = metric, rows = metric == "Pseudo-R² (Cragg-Uhler)")
  ) |>
  tab_options(
    table.font.size            = 11,
    heading.title.font.size    = 13,
    heading.subtitle.font.size = 10,
    table.width                = pct(100)
  )
```

### Model Comparison

| Outcome | Final Predictors | Retained predictors (examples) |
|---------|------------------|--------------------------------|
| Digestible protein from maize | `r n_predictors_maize` variables | Age, department, household size, area |
| Digestible protein from non-maize | `r n_predictors_non_maize` variables | Age, household size, area, education, wall materials |

### Key Findings

1. **More predictors for non-maize**: The non-maize model retains
   `r n_predictors_non_maize` predictors against `r n_predictors_maize` in the
   maize model, drawing more of the housing and service variables into the
   retained set

2. **Geographic variation**: Department is retained in both models, consistent
   with regional variation in dietary patterns

3. **Shared socioeconomic predictors**: Age, household size, area and mean
   household education are retained in both models, so the same socioeconomic
   axis carries predictive power for maize and non-maize digestible protein

### Application

These models will be applied to SIVESNU 2018 children to predict their digestible 
protein intake profiles based on their demographic and socioeconomic characteristics. 
The predicted values enable linking nutritional status to stunting outcomes.
