---
title: "Transfer Model: Iron Intake"
subtitle: "Module 3: Transfer Models"
description-meta: "Survey-weighted gamma regression predicting child iron intake from maize and non-maize sources separately, gamma selected over lognormal by fit."
---

## Why Model Iron?

Iron is an essential micronutrient for child development, and iron deficiency
may contribute to poor growth outcomes, including stunting, particularly when
combined with other forms of undernutrition. In Guatemala, maize is a
substantial dietary source of iron (see
[Daily Nutrient Intake by Source](../01_nutritional_intakes/01_03_intake_calculation.qmd)
in Module 1). By modeling iron intake from maize and non-maize sources
separately, we can:

1. **Simulate biofortification scenarios** — Predict how higher-content maize varieties
   would change total iron intake
2. **Preserve baseline non-maize intake** — Keep other dietary sources constant
   when modeling interventions
3. **Locate the modifiable fraction** — Identify where the maize-derived share
   of iron intake is largest

## Modeling Approach

We use **survey-weighted gamma regression** because iron intake data is:

- **Positive and continuous** — People consume some amount of iron daily
- **Right-skewed** — Most people have moderate intake, few have very high intake
- **Population-representative** — Survey weights carry the estimates to the
  national level

The models predict iron intake based on demographic and socioeconomic 
characteristics that are available in both ENCOVI (training data) and 
SIVESNU (application target).

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  fitdistrplus,       # Distribution fitting for goodness-of-fit assessment
  ggpubr,             # Arrange diagnostic plots into a 2x2 grid
  survey,             # Required for svyglm method dispatch (model diagnostics)
  performance,        # Pseudo-R² (Nagelkerke / Cragg-Uhler) and diagnostics
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))

# --- Local helpers: distribution assessment (web formatting) ---------------
# Recreate the distribution-fit outputs from the raw outcome vector exported
# by the framework, returning table and diagnostic grid as separate objects
# so each renders in its own chunk with web-consistent formatting. 

build_distribution_table <- function(outcome) {

  fit_gamma       <- fitdist(data = outcome, "gamma")
  fit_lognormal   <- fitdist(data = outcome, "lnorm")
  fit_exponential <- fitdist(data = outcome, "exp")

  data.frame(
    distribution = c("Gamma", "Lognormal", "Exponential"),
    loglik = c(fit_gamma$loglik, fit_lognormal$loglik, fit_exponential$loglik),
    aic    = c(fit_gamma$aic, fit_lognormal$aic, fit_exponential$aic),
    bic    = c(fit_gamma$bic, fit_lognormal$bic, fit_exponential$bic),
    ks_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$ks,
      gofstat(fit_lognormal, fitnames = "lnorm")$ks,
      gofstat(fit_exponential, fitnames = "exp")$ks
    ),
    cvm_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$cvm,
      gofstat(fit_lognormal, fitnames = "lnorm")$cvm,
      gofstat(fit_exponential, fitnames = "exp")$cvm
    ),
    ad_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$ad,
      gofstat(fit_lognormal, fitnames = "lnorm")$ad,
      gofstat(fit_exponential, fitnames = "exp")$ad
    )
  ) |>
    gt() |>
    tab_header(
      title = "Distribution Fit Comparison",
      subtitle = "Goodness-of-fit statistics for Gamma, Lognormal, and Exponential distributions"
    ) |>
    cols_label(
      distribution = "Distribution",
      loglik   = "Log-Likelihood",
      aic      = "AIC",
      bic      = "BIC",
      ks_stat  = "Kolmogorov-Smirnov",
      cvm_stat = "Cramer-von Mises",
      ad_stat  = "Anderson-Darling"
    ) |>
    fmt_number(
      columns = c(loglik, aic, bic, ks_stat, cvm_stat, ad_stat),
      decimals = 4
    ) |>
    tab_spanner(label = "Information Criteria", columns = c(aic, bic)) |>
    tab_spanner(label = "Goodness-of-Fit Statistics", columns = c(ks_stat, cvm_stat, ad_stat)) |>
    cols_align(align = "center", columns = everything()) |>
    cols_align(align = "left", columns = distribution) |>
    tab_options(
      table.font.size = 11,
      heading.title.font.size = 13,
      heading.subtitle.font.size = 10,
      table.width = pct(100)
    )
}

build_distribution_plots <- function(outcome) {

  fit_gamma <- fitdist(data = outcome, "gamma")

  shape <- fit_gamma$estimate["shape"]
  rate  <- fit_gamma$estimate["rate"]
  x_obs <- sort(fit_gamma$data)

  plot_density <- data.frame(x = fit_gamma$data) |>
    ggplot(aes(x = x)) +
    geom_histogram(aes(y = after_stat(density)), bins = 30,
                   color = "azure4", fill = "lightgray", alpha = 0.5) +
    stat_function(fun = dgamma, args = list(shape = shape, rate = rate),
                  color = "#3AAF85", linetype = "dashed", linewidth = 0.8) +
    labs(title = "Empirical vs Theoretical Density", x = "", y = "Density") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      plot.subtitle = element_text(color = "azure4", size = 10),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_qq <- data.frame(
    empirical = x_obs,
    theoretical = qgamma(ppoints(length(x_obs)), shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = theoretical, y = empirical)) +
    geom_point(color = "#1B6CA8", size = 1, alpha = 0.2) +
    geom_abline(intercept = 0, slope = 1, color = "#3AAF85",
                linetype = "dashed", linewidth = 0.8) +
    labs(title = "Quantile-Quantile Plot", x = "Theoretical Quantiles", y = "Empirical Quantiles") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_cdf <- data.frame(
    x = x_obs,
    empirical_cdf = ecdf(fit_gamma$data)(x_obs),
    theoretical_cdf = pgamma(x_obs, shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = x)) +
    geom_line(aes(y = empirical_cdf), color = "lightgray", linewidth = 2.5) +
    geom_line(aes(y = theoretical_cdf), color = "#3AAF85",
              linetype = "dashed", linewidth = 0.8) +
    labs(title = "Empirical vs Theoretical CDF",
         subtitle = "CDF: Cumulative Distribution Function", x = "Data", y = "CDF") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      plot.subtitle = element_text(color = "azure4", size = 10),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_pp <- data.frame(
    empirical_prob = ecdf(fit_gamma$data)(x_obs),
    theoretical_prob = pgamma(x_obs, shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = theoretical_prob, y = empirical_prob)) +
    geom_point(color = "#1B6CA8", size = 1, alpha = 0.2) +
    geom_abline(intercept = 0, slope = 1, color = "#3AAF85",
                linetype = "dashed", linewidth = 0.8) +
    labs(title = "P-P Plot", x = "Theoretical Probabilities", y = "Empirical Probabilities") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  ggarrange(plot_density, plot_qq, plot_cdf, plot_pp, ncol = 2, nrow = 2)
}
```

```{r}
#| label: load-rds
#| include: false

# (a) Web resources for visualizations (exported by framework 03_02)
fig_b1_evaluate_distribution_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_fig_b1_evaluate_distribution_maize.rds"),
  trust = TRUE
)

tbl_b2_coefficients_full_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_tbl_b2_coefficients_full_model_maize.rds"),
  trust = TRUE
)

fig_b2_diagnostic_plot_full_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_fig_b2_diagnostic_plot_full_model_maize.rds"),
  trust = TRUE
)

tbl_b3_wald_test_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_tbl_b3_wald_test_maize.rds"),
  trust = TRUE
)

tbl_b4_coefficients_refined_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_tbl_b4_coefficients_refined_model_maize.rds"),
  trust = TRUE
)

fig_b4_diagnostic_plot_refined_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_fig_b4_diagnostic_plot_refined_model_maize.rds"),
  trust = TRUE
)

fig_b5_evaluate_diagnostics_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_fig_b5_evaluate_diagnostics_maize.rds"),
  trust = TRUE
)

fig_c1_evaluate_distribution_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_fig_c1_evaluate_distribution_non_maize.rds"),
  trust = TRUE
)

tbl_c2_coefficients_full_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_tbl_c2_coefficients_full_model_non_maize.rds"),
  trust = TRUE
)

fig_c2_diagnostic_plot_full_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_fig_c2_diagnostic_plot_full_model_non_maize.rds"),
  trust = TRUE
)

tbl_c3_wald_test_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_tbl_c3_wald_test_non_maize.rds"),
  trust = TRUE
)

tbl_c4_coefficients_refined_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_tbl_c4_coefficients_refined_model_non_maize.rds"),
  trust = TRUE
)

fig_c4_diagnostic_plot_refined_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_fig_c4_diagnostic_plot_refined_model_non_maize.rds"),
  trust = TRUE
)

fig_c5_evaluate_diagnostics_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_02_fig_c5_evaluate_diagnostics_non_maize.rds"),
  trust = TRUE
)

# (b) Fitted models for performance metrics
models <- import(
  here("..", "01_data", "02_processed", "models",
       "03_02_iron_models.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Significant predictor counts from Wald tests (maize and non-maize)
n_predictors_maize <- tbl_b3_wald_test_maize |>
  dplyr::filter(significant == "*") |>
  nrow()

n_predictors_non_maize <- tbl_c3_wald_test_non_maize |>
  dplyr::filter(significant == "*") |>
  nrow()

# Collinearity of the refined maize model, on the degrees-of-freedom adjusted
# VIF scale reported by the diagnostic grid
vif_maize <- adjust_collinearity_df(models$model_fe_maize)
max_vif_maize <- round(max(vif_maize$VIF), 1)
n_vif_over_5 <- sum(vif_maize$VIF > 5)
n_vif_terms <- nrow(vif_maize)
```


## Data Preparation

### Source Data

We use the cleaned ENCOVI dataset from the previous preparation step, containing 
individuals with complete predictor variables and nutrient intake estimates.

### Handling Extreme Values

Gamma regression requires positive values and is sensitive to extreme outliers. 
We apply conservative outlier detection:

::: {.callout-tip}
## Outlier Threshold
Upper outliers identified as values exceeding **P95 + 1.5×IQR** (i.e., 95th percentile 
plus 1.5 times the interquartile range). The threshold suits micronutrients:

- Micronutrient intake distributions are markedly right-skewed
- Biologically plausible high values arise from consumption of
  micronutrient-dense or fortified foods
- Retaining the upper tail supports model generalizability to populations with
  varied dietary patterns
:::

### Survey Design

All models incorporate ENCOVI's complex survey design (clustering and
expansion weights), which places the estimates at the population level.


## Modeling Iron Intake from Maize

### Distribution Selection

Before modeling, we assess which statistical distribution best fits the iron 
intake data. Three candidate distributions are compared:

```{r}
#| label: tbl-b1-evaluate-distribution-maize
#| tbl-cap: "Distribution fit comparison for iron intake from maize"

build_distribution_table(fig_b1_evaluate_distribution_maize)
```

```{r}
#| label: fig-b1-evaluate-distribution-maize
#| fig-cap: "Distribution assessment for iron intake from maize"
#| fig-alt: >
#|   Four-panel goodness-of-fit grid for iron intake from maize. The top-left
#|   panel overlays a histogram of intake (density on the y-axis) with a fitted
#|   gamma curve; the top-right is a quantile-quantile plot of empirical against
#|   theoretical quantiles; the bottom-left compares the empirical and
#|   theoretical cumulative distributions; and the bottom-right is a
#|   probability-probability plot. Across all panels the points and curves track
#|   the gamma reference line closely, confirming that the right-skewed intake
#|   is well described by a gamma distribution.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

build_distribution_plots(fig_b1_evaluate_distribution_maize)
```

::: {.callout-important}
## Distribution Selection
**Gamma distribution** provides the best fit based on:

- Lowest AIC/BIC (information criteria)
- Highest Log-Likelihood

Gamma regression with log link is appropriate for positive, right-skewed 
continuous outcomes.
:::

### Full Model

We first train a model including all 20 candidate predictors to identify 
which variables significantly predict iron intake from maize.

```{r}
#| label: tbl-b2-coefficients-full-model-maize
#| tbl-cap: "Model coefficients for iron intake from maize (full model)"

tbl_b2_coefficients_full_model_maize
```

```{r}
#| label: fig-b2-diagnostic-plot-full-model-maize
#| fig-cap: "Observed vs predicted iron intake from maize (full model)"
#| fig-alt: >
#|   Scatter plot with observed iron from maize (mg/day) on the x-axis and
#|   predicted iron from maize (mg/day) on the y-axis for the full model with
#|   all 20 candidate predictors. Points follow a positive, upward-sloping
#|   cloud, showing predictions rise with observed intake and track the data
#|   across the range.

fig_b2_diagnostic_plot_full_model_maize |>
  ggplot(aes(x = fe_individuo_dia_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Iron from Maize (mg/day)",
    y = "Predicted Iron from Maize (mg/day)"
  )
```

### Variable Selection

Not all predictors contribute meaningfully to prediction. We use **Wald tests** 
to identify statistically significant variables and remove those that add 
complexity without improving predictions.

```{r}
#| label: tbl-b3-wald-test-maize
#| tbl-cap: "Wald test results for predictor significance (iron from maize)"

tbl_b3_wald_test_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = "Wald F-test results for iron intake from maize model"
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-important}
## Variables Retained
**`r n_predictors_maize` predictors** show significant relationships with iron 
intake from maize, spanning demographic, geographic, socioeconomic, housing 
and asset categories. Variables that do not significantly predict iron intake 
from maize after controlling for other factors are excluded from the refined 
model.
:::

### Refined Model

The final model is fitted on the significant predictors alone.

```{r}
#| label: tbl-b4-coefficients-refined-model-maize
#| tbl-cap: "Model coefficients for iron intake from maize (refined model)"

tbl_b4_coefficients_refined_model_maize
```

```{r}
#| label: fig-b4-diagnostic-plot-refined-model-maize
#| fig-cap: "Observed vs predicted iron intake from maize (refined model)"
#| fig-alt: >
#|   Scatter plot with observed iron from maize (mg/day) on the x-axis and
#|   predicted iron from maize (mg/day) on the y-axis for the refined model that
#|   keeps only the significant predictors. Points follow the same positive
#|   trend as the full model, indicating the simpler model preserves predictive
#|   accuracy.

fig_b4_diagnostic_plot_refined_model_maize |>
  ggplot(aes(x = fe_individuo_dia_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Iron from Maize (mg/day)",
    y = "Predicted Iron from Maize (mg/day)"
  )
```

### Model Diagnostics

We verify the model meets statistical assumptions:

```{r}
#| label: fig-b5-evaluate-diagnostics-maize
#| fig-cap: "Diagnostic plots for iron from maize model"
#| fig-alt: >
#|   Grid of regression diagnostic plots for the refined gamma model of iron
#|   from maize, including residuals versus fitted values, a normal
#|   quantile-quantile plot of the residuals, and leverage and collinearity
#|   checks. The residuals scatter without any systematic pattern and fall close
#|   to the reference line, no observations stand out as high-leverage, and the
#|   collinearity panel plots each term's degrees-of-freedom adjusted variance
#|   inflation factor against shaded low, moderate and high reference bands.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

evaluate_gamma_model(models$model_fe_maize)
```

::: {.callout-tip}
## Diagnostic Assessment
- **Residuals**: No systematic patterns, approximately normal
- **Leverage**: No overly influential observations
- **Collinearity**: Degrees-of-freedom adjusted variance inflation factors (VIF)
  reach a maximum of `r max_vif_maize`, with `r n_vif_over_5` of
  `r n_vif_terms` terms above the conventional threshold of 5

Factors and polynomial terms contribute several columns to the model matrix, so
their generalized VIF is rescaled by the term's degrees of freedom. This places
every term on the single-coefficient scale that the thresholds of 5 and 10
refer to.
:::


## Modeling Iron Intake from Non-Maize Sources

Iron from non-maize sources includes vegetables, fruits, legumes, and animal 
products. This component remains **constant** in biofortification scenarios, 
representing the baseline dietary iron that would not change with maize 
interventions.

### Distribution Selection

```{r}
#| label: tbl-c1-evaluate-distribution-non-maize
#| tbl-cap: "Distribution fit comparison for iron intake from non-maize sources"

build_distribution_table(fig_c1_evaluate_distribution_non_maize)
```

```{r}
#| label: fig-c1-evaluate-distribution-non-maize
#| fig-cap: "Distribution assessment for iron intake from non-maize sources"
#| fig-alt: >
#|   Four-panel goodness-of-fit grid for iron intake from non-maize sources. The
#|   top-left panel overlays a histogram of intake (density on the y-axis) with a
#|   fitted gamma curve; the top-right is a quantile-quantile plot of empirical
#|   against theoretical quantiles; the bottom-left compares the empirical and
#|   theoretical cumulative distributions; and the bottom-right is a
#|   probability-probability plot. The points and curves again follow the gamma
#|   reference line closely, confirming a gamma distribution best fits this
#|   right-skewed intake.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

build_distribution_plots(fig_c1_evaluate_distribution_non_maize)
```

::: {.callout-important}
## Distribution Selection
**Gamma distribution** again provides the best fit for non-maize iron intake.
:::

### Full Model

```{r}
#| label: tbl-c2-coefficients-full-model-non-maize
#| tbl-cap: "Model coefficients for iron intake from non-maize sources (full model)"

tbl_c2_coefficients_full_model_non_maize
```

```{r}
#| label: fig-c2-diagnostic-plot-full-model-non-maize
#| fig-cap: "Observed vs predicted iron intake from non-maize sources (full model)"
#| fig-alt: >
#|   Scatter plot with observed iron from non-maize sources (mg/day) on the
#|   x-axis and predicted iron from non-maize sources (mg/day) on the y-axis for
#|   the full model with all candidate predictors. Points form a positive,
#|   upward-sloping cloud, showing predictions increase with observed intake
#|   across the range.

fig_c2_diagnostic_plot_full_model_non_maize |>
  ggplot(aes(x = fe_individuo_dia_no_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Iron from Non-Maize (mg/day)",
    y = "Predicted Iron from Non-Maize (mg/day)"
  )
```

### Variable Selection

```{r}
#| label: tbl-c3-wald-test-non-maize
#| tbl-cap: "Wald test results for predictor significance (iron from non-maize)"

tbl_c3_wald_test_non_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = "Wald F-test results for iron intake from non-maize model"
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-important}
## Variables Retained
**`r n_predictors_non_maize` predictors** significantly predict iron intake from 
non-maize sources, drawing from the same set of demographic, geographic, 
socioeconomic and housing variables. The retained set differs from the maize model: the two outcomes are predicted
by different subsets of the same candidate variables.
:::

### Refined Model

```{r}
#| label: tbl-c4-coefficients-refined-model-non-maize
#| tbl-cap: "Model coefficients for iron intake from non-maize sources (refined model)"

tbl_c4_coefficients_refined_model_non_maize
```

```{r}
#| label: fig-c4-diagnostic-plot-refined-model-non-maize
#| fig-cap: "Observed vs predicted iron intake from non-maize sources (refined model)"
#| fig-alt: >
#|   Scatter plot with observed iron from non-maize sources (mg/day) on the
#|   x-axis and predicted iron from non-maize sources (mg/day) on the y-axis for
#|   the refined model that keeps only the significant predictors. Points follow
#|   the same positive trend as the full model, showing the simpler model
#|   retains predictive accuracy.

fig_c4_diagnostic_plot_refined_model_non_maize |>
  ggplot(aes(x = fe_individuo_dia_no_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Iron from Non-Maize (mg/day)",
    y = "Predicted Iron from Non-Maize (mg/day)"
  )
```

### Model Diagnostics

```{r}
#| label: fig-c5-evaluate-diagnostics-non-maize
#| fig-cap: "Diagnostic plots for iron from non-maize sources model"
#| fig-alt: >
#|   Grid of regression diagnostic plots for the refined gamma model of iron
#|   from non-maize sources, including residuals versus fitted values, a normal
#|   quantile-quantile plot of the residuals, and leverage and collinearity
#|   checks. The residuals scatter without systematic pattern and stay close to
#|   the reference line, with no high-leverage points, and the collinearity
#|   panel plots each term's degrees-of-freedom adjusted variance inflation
#|   factor against shaded low, moderate and high reference bands.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

evaluate_gamma_model(models$model_fe_non_maize)
```


## Summary

### Final Model Performance

```{r}
#| label: tbl-model-performance
#| tbl-cap: "Final refined model performance metrics"

dplyr::bind_rows(
  extract_gamma_model_metrics(models$model_fe_maize,     "Iron from maize"),
  extract_gamma_model_metrics(models$model_fe_non_maize, "Iron from non-maize")
) |>
  # Reshape: metrics as rows, models as columns
  tidyr::pivot_longer(
    cols = -source,
    names_to = "metric",
    values_to = "value"
  ) |>
  tidyr::pivot_wider(
    names_from = source,
    values_from = value
  ) |>
  # Apply readable labels per metric
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      n_obs                 = "N observations",
      n_predictors          = "N predictors",
      pseudo_r2_mcfadden    = "Pseudo-R² (McFadden)",
      pseudo_r2_cragg_uhler = "Pseudo-R² (Cragg-Uhler)",
      aic                   = "AIC",
      dispersion            = "Dispersion"
    )
  ) |>
  gt() |>
  cols_label(metric = "Metric") |>
  fmt_number(
    rows    = metric == "N observations",
    columns = c(`Iron from maize`, `Iron from non-maize`),
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    rows    = metric == "N predictors",
    columns = c(`Iron from maize`, `Iron from non-maize`),
    decimals = 0
  ) |>
  fmt_number(
    rows    = metric %in% c("Pseudo-R² (McFadden)",
                            "Pseudo-R² (Cragg-Uhler)",
                            "Dispersion"),
    columns = c(`Iron from maize`, `Iron from non-maize`),
    decimals = 3
  ) |>
  fmt_number(
    rows    = metric == "AIC",
    columns = c(`Iron from maize`, `Iron from non-maize`),
    decimals = 0
  ) |>
  tab_header(
    title    = "Final Model Performance",
    subtitle = "Refined gamma regression with log link"
  ) |>
  tab_footnote(
    footnote  = "Pseudo-R² (McFadden) = 1 − (residual deviance / null deviance). Conservative measure; values above 0.2 are conventionally considered good for GLMs.",
    locations = cells_body(columns = metric, rows = metric == "Pseudo-R² (McFadden)")
  ) |>
  tab_footnote(
    footnote  = "Pseudo-R² (Cragg-Uhler), also known as Nagelkerke's R², is normalized to a [0, 1] range and generally yields higher values than McFadden. Both measures are reported for transparency.",
    locations = cells_body(columns = metric, rows = metric == "Pseudo-R² (Cragg-Uhler)")
  ) |>
  tab_options(
    table.font.size            = 11,
    heading.title.font.size    = 13,
    heading.subtitle.font.size = 10,
    table.width                = pct(100)
  )
```

### Model Comparison

| Outcome | Final Predictors | Retained predictors (examples) |
|---------|------------------|--------------------------------|
| Iron from maize | `r n_predictors_maize` variables | Age, department, education, area |
| Iron from non-maize | `r n_predictors_non_maize` variables | Age, household size, dwelling type |

### Key Findings

1. **Geographic variation**: Department is retained in both models, consistent
   with regional variation in dietary patterns

2. **Differential role of area**: Urban/rural area is retained in the maize
   model and not in the non-maize model, so it carries predictive power for
   one intake source and not the other

3. **Household size**: Household size is retained in the non-maize model only

### Application

These models will be applied to SIVESNU 2018 children to predict their iron 
intake profiles based on their demographic and socioeconomic characteristics. 
The predicted values enable linking nutritional status to stunting outcomes.
