---
title: "Transfer Model: Energy Intake"
subtitle: "Module 3: Transfer Models"
description-meta: "Survey-weighted gamma regression for child energy intake, which conditions whether protein and micronutrients can be used for linear growth."
---

## Why Model Energy?

Energy intake (kilocalories) is fundamental for child growth and development.
Where energy intake is low, the utilization of dietary protein and
micronutrients for growth is constrained, so energy adequacy conditions the
interpretation of the other nutrient outcomes. By modeling energy intake from
maize and non-maize sources separately, we can:

1. **Simulate biofortification scenarios** — Understand how energy from maize 
   contributes to total energy intake
2. **Preserve baseline non-maize intake** — Keep other dietary sources constant 
   when modeling interventions
3. **Contextualize nutrient adequacy** — Energy intake bears on the
   interpretation of the other nutrient indicators

::: {.callout-note}
## Role of Energy in Stunting
Biofortification targets micronutrient content and protein quality, and energy
adequacy provides the context for interpreting those nutrient impacts. Where
energy intake is inadequate, the benefit a child draws from higher nutrient
density in the food supply is correspondingly limited.
:::

## Modeling Approach

We use **survey-weighted gamma regression** because energy intake data is:

- **Positive and continuous** — People consume some amount of calories daily
- **Right-skewed** — Most people have moderate intake, few have very high intake
- **Population-representative** — Survey weights carry the estimates to the
  national level

The models predict energy intake based on demographic and socioeconomic 
characteristics that are available in both ENCOVI (training data) and 
SIVESNU (application target).

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  fitdistrplus,       # Distribution fitting for goodness-of-fit assessment
  ggpubr,             # Arrange diagnostic plots into a 2x2 grid
  survey,             # Required for svyglm method dispatch (model diagnostics)
  performance,        # Pseudo-R² (Nagelkerke / Cragg-Uhler) and diagnostics
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))

# --- Local helpers: distribution assessment (web formatting) ---------------
# Recreate the distribution-fit outputs from the raw outcome vector exported
# by the framework, returning table and diagnostic grid as separate objects
# so each renders in its own chunk with web-consistent formatting.

build_distribution_table_energy <- function(outcome) {

  shape_start <- mean(outcome, na.rm = TRUE)^2 / var(outcome, na.rm = TRUE)
  rate_start  <- mean(outcome, na.rm = TRUE) / var(outcome, na.rm = TRUE)

  fit_gamma <- fitdist(
    data = outcome, "gamma",
    start = list(shape = shape_start, rate = rate_start),
    lower = c(0.01, 0.0001)
  )
  fit_lognormal <- fitdist(data = outcome, "lnorm")

  data.frame(
    distribution = c("Gamma", "Lognormal"),
    loglik = c(fit_gamma$loglik, fit_lognormal$loglik),
    aic    = c(fit_gamma$aic, fit_lognormal$aic),
    bic    = c(fit_gamma$bic, fit_lognormal$bic),
    ks_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$ks,
      gofstat(fit_lognormal, fitnames = "lnorm")$ks
    ),
    cvm_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$cvm,
      gofstat(fit_lognormal, fitnames = "lnorm")$cvm
    ),
    ad_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$ad,
      gofstat(fit_lognormal, fitnames = "lnorm")$ad
    )
  ) |>
    gt() |>
    tab_header(
      title = "Distribution Fit Comparison",
      subtitle = "Goodness-of-fit statistics for Gamma and Lognormal distributions"
    ) |>
    cols_label(
      distribution = "Distribution",
      loglik   = "Log-Likelihood",
      aic      = "AIC",
      bic      = "BIC",
      ks_stat  = "Kolmogorov-Smirnov",
      cvm_stat = "Cramer-von Mises",
      ad_stat  = "Anderson-Darling"
    ) |>
    fmt_number(
      columns = c(loglik, aic, bic, ks_stat, cvm_stat, ad_stat),
      decimals = 4
    ) |>
    tab_spanner(label = "Information Criteria", columns = c(aic, bic)) |>
    tab_spanner(label = "Goodness-of-Fit Statistics", columns = c(ks_stat, cvm_stat, ad_stat)) |>
    cols_align(align = "center", columns = everything()) |>
    cols_align(align = "left", columns = distribution) |>
    tab_options(
      table.font.size = 11,
      heading.title.font.size = 13,
      heading.subtitle.font.size = 10,
      table.width = pct(100)
    )
}

build_distribution_plots <- function(outcome) {

  shape_start <- mean(outcome, na.rm = TRUE)^2 / var(outcome, na.rm = TRUE)
  rate_start  <- mean(outcome, na.rm = TRUE) / var(outcome, na.rm = TRUE)

  fit_gamma <- fitdist(
    data = outcome, "gamma",
    start = list(shape = shape_start, rate = rate_start),
    lower = c(0.01, 0.0001)
  )

  shape <- fit_gamma$estimate["shape"]
  rate  <- fit_gamma$estimate["rate"]
  x_obs <- sort(fit_gamma$data)

  plot_density <- data.frame(x = fit_gamma$data) |>
    ggplot(aes(x = x)) +
    geom_histogram(aes(y = after_stat(density)), bins = 30,
                   color = "azure4", fill = "lightgray", alpha = 0.5) +
    stat_function(fun = dgamma, args = list(shape = shape, rate = rate),
                  color = "#3AAF85", linetype = "dashed", linewidth = 0.8) +
    labs(title = "Empirical vs Theoretical Density", x = "", y = "Density") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      plot.subtitle = element_text(color = "azure4", size = 10),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_qq <- data.frame(
    empirical = x_obs,
    theoretical = qgamma(ppoints(length(x_obs)), shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = theoretical, y = empirical)) +
    geom_point(color = "#1B6CA8", size = 1, alpha = 0.2) +
    geom_abline(intercept = 0, slope = 1, color = "#3AAF85",
                linetype = "dashed", linewidth = 0.8) +
    labs(title = "Quantile-Quantile Plot", x = "Theoretical Quantiles", y = "Empirical Quantiles") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_cdf <- data.frame(
    x = x_obs,
    empirical_cdf = ecdf(fit_gamma$data)(x_obs),
    theoretical_cdf = pgamma(x_obs, shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = x)) +
    geom_line(aes(y = empirical_cdf), color = "lightgray", linewidth = 2.5) +
    geom_line(aes(y = theoretical_cdf), color = "#3AAF85",
              linetype = "dashed", linewidth = 0.8) +
    labs(title = "Empirical vs Theoretical CDF",
         subtitle = "CDF: Cumulative Distribution Function", x = "Data", y = "CDF") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      plot.subtitle = element_text(color = "azure4", size = 10),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_pp <- data.frame(
    empirical_prob = ecdf(fit_gamma$data)(x_obs),
    theoretical_prob = pgamma(x_obs, shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = theoretical_prob, y = empirical_prob)) +
    geom_point(color = "#1B6CA8", size = 1, alpha = 0.2) +
    geom_abline(intercept = 0, slope = 1, color = "#3AAF85",
                linetype = "dashed", linewidth = 0.8) +
    labs(title = "P-P Plot", x = "Theoretical Probabilities", y = "Empirical Probabilities") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  ggarrange(plot_density, plot_qq, plot_cdf, plot_pp, ncol = 2, nrow = 2)
}
```

```{r}
#| label: load-rds
#| include: false

# (a) Web resources for visualizations (exported by framework 03_07)
fig_b1_evaluate_distribution_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_fig_b1_evaluate_distribution_maize.rds"),
  trust = TRUE
)

tbl_b2_coefficients_full_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_tbl_b2_coefficients_full_model_maize.rds"),
  trust = TRUE
)

fig_b2_diagnostic_plot_full_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_fig_b2_diagnostic_plot_full_model_maize.rds"),
  trust = TRUE
)

tbl_b3_wald_test_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_tbl_b3_wald_test_maize.rds"),
  trust = TRUE
)

tbl_b4_coefficients_refined_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_tbl_b4_coefficients_refined_model_maize.rds"),
  trust = TRUE
)

fig_b4_diagnostic_plot_refined_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_fig_b4_diagnostic_plot_refined_model_maize.rds"),
  trust = TRUE
)

fig_b5_evaluate_diagnostics_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_fig_b5_evaluate_diagnostics_maize.rds"),
  trust = TRUE
)

fig_c1_evaluate_distribution_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_fig_c1_evaluate_distribution_non_maize.rds"),
  trust = TRUE
)

tbl_c2_coefficients_full_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_tbl_c2_coefficients_full_model_non_maize.rds"),
  trust = TRUE
)

fig_c2_diagnostic_plot_full_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_fig_c2_diagnostic_plot_full_model_non_maize.rds"),
  trust = TRUE
)

tbl_c3_wald_test_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_tbl_c3_wald_test_non_maize.rds"),
  trust = TRUE
)

tbl_c4_coefficients_refined_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_tbl_c4_coefficients_refined_model_non_maize.rds"),
  trust = TRUE
)

fig_c4_diagnostic_plot_refined_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_fig_c4_diagnostic_plot_refined_model_non_maize.rds"),
  trust = TRUE
)

fig_c5_evaluate_diagnostics_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_07_fig_c5_evaluate_diagnostics_non_maize.rds"),
  trust = TRUE
)

# (b) Fitted models for performance metrics
models <- import(
  here("..", "01_data", "02_processed", "models",
       "03_07_energy_models.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Significant predictor counts from Wald tests (maize and non-maize)
n_predictors_maize <- tbl_b3_wald_test_maize |>
  dplyr::filter(significant == "*") |>
  nrow()

n_predictors_non_maize <- tbl_c3_wald_test_non_maize |>
  dplyr::filter(significant == "*") |>
  nrow()

# Collinearity of the refined maize model, on the degrees-of-freedom adjusted
# VIF scale reported by the diagnostic grid
vif_maize <- adjust_collinearity_df(models$model_kcal_maize)
max_vif_maize <- round(max(vif_maize$VIF), 1)
n_vif_over_5 <- sum(vif_maize$VIF > 5)
n_vif_terms <- nrow(vif_maize)
```


## Data Preparation

### Source Data

We use the cleaned ENCOVI dataset from the previous preparation step, containing 
individuals with complete predictor variables and nutrient intake estimates.

### Handling Extreme Values

Gamma regression requires positive values and is sensitive to extreme outliers. 
We apply conservative outlier detection:

::: {.callout-tip}
## Outlier Threshold
Upper outliers identified using the standard **Tukey method (Q3 + 1.5×IQR)**. 
The threshold suits macronutrients:

- Energy intake distributions are more stable
- Extreme values are consistent with measurement error
- Energy intakes have natural biological limits
:::

### Survey Design

All models incorporate ENCOVI's complex survey design (clustering and
expansion weights), which places the estimates at the population level.


## Modeling Energy Intake from Maize

### Distribution Selection

Before modeling, we assess which statistical distribution best fits the energy 
intake data. Three candidate distributions are compared:

```{r}
#| label: tbl-b1-evaluate-distribution-maize
#| tbl-cap: "Distribution fit comparison for energy intake from maize"

build_distribution_table_energy(fig_b1_evaluate_distribution_maize)
```

```{r}
#| label: fig-b1-evaluate-distribution-maize
#| fig-cap: "Distribution assessment for energy intake from maize"
#| fig-alt: >
#|   Four goodness-of-fit panels comparing observed energy intake from maize
#|   against a fitted Gamma distribution. Top left, a histogram of intake
#|   overlaid with the theoretical Gamma density curve. Top right, a
#|   quantile-quantile plot; bottom left, empirical versus theoretical
#|   cumulative distribution curves; bottom right, a probability-probability
#|   plot. Points and curves track the reference lines closely, indicating the
#|   Gamma distribution fits the right-skewed intake data well.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

build_distribution_plots(fig_b1_evaluate_distribution_maize)
```

::: {.callout-important}
## Distribution Selection
**Gamma distribution** provides the best fit based on:

- Lowest AIC/BIC (information criteria)
- Highest Log-Likelihood

Gamma regression with log link is appropriate for positive, right-skewed 
continuous outcomes.
:::

### Full Model

We first train a model including all 20 candidate predictors to identify 
which variables significantly predict energy intake from maize.

```{r}
#| label: tbl-b2-coefficients-full-model-maize
#| tbl-cap: "Model coefficients for energy intake from maize (full model)"

tbl_b2_coefficients_full_model_maize
```

```{r}
#| label: fig-b2-diagnostic-plot-full-model-maize
#| fig-cap: "Observed vs predicted energy intake from maize (full model)"
#| fig-alt: >
#|   Scatter chart with observed energy from maize in kilocalories per day on
#|   the x-axis and model-predicted energy on the y-axis, using all 20
#|   candidate predictors. Each point is a person. Predicted values span a
#|   narrower range than observed values, so the cloud rises gently rather than
#|   following the ideal one-to-one line, showing the model captures average
#|   intake but compresses the extremes.

fig_b2_diagnostic_plot_full_model_maize |>
  ggplot(aes(x = kcal_individuo_dia_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Energy from Maize (kcal/day)",
    y = "Predicted Energy from Maize (kcal/day)"
  )
```

### Variable Selection

Not all predictors contribute meaningfully to prediction. We use **Wald tests** 
to identify statistically significant variables and remove those that add 
complexity without improving predictions.

```{r}
#| label: tbl-b3-wald-test-maize
#| tbl-cap: "Wald test results for predictor significance (energy from maize)"

tbl_b3_wald_test_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = "Wald F-test results for energy intake from maize model"
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-important}
## Variables Retained
**`r n_predictors_maize` predictors** show significant relationships with energy 
intake from maize, spanning demographic, geographic, socioeconomic, housing 
and asset categories. Variables that do not significantly predict energy intake 
from maize after controlling for other factors are excluded from the refined 
model.
:::

### Refined Model

The final model is fitted on the significant predictors alone.

```{r}
#| label: tbl-b4-coefficients-refined-model-maize
#| tbl-cap: "Model coefficients for energy intake from maize (refined model)"

tbl_b4_coefficients_refined_model_maize
```

```{r}
#| label: fig-b4-diagnostic-plot-refined-model-maize
#| fig-cap: "Observed vs predicted energy intake from maize (refined model)"
#| fig-alt: >
#|   Scatter chart with observed energy from maize in kilocalories per day on
#|   the x-axis and model-predicted energy on the y-axis, using only the
#|   significant predictors retained after variable selection. Each point is a
#|   person. The pattern closely mirrors the full model, confirming that
#|   dropping non-significant variables preserves predictive accuracy while
#|   simplifying the model.

fig_b4_diagnostic_plot_refined_model_maize |>
  ggplot(aes(x = kcal_individuo_dia_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Energy from Maize (kcal/day)",
    y = "Predicted Energy from Maize (kcal/day)"
  )
```

### Model Diagnostics

We verify the model meets statistical assumptions:

```{r}
#| label: fig-b5-evaluate-diagnostics-maize
#| fig-cap: "Diagnostic plots for energy from maize model"
#| fig-alt: >
#|   A grid of regression diagnostic panels for the refined Gamma model of
#|   energy from maize, covering residual behaviour, influential observations,
#|   and predictor collinearity. Residuals scatter around their reference line
#|   with no systematic pattern, no points exert excessive leverage, and the
#|   collinearity panel plots each term's degrees-of-freedom adjusted variance
#|   inflation factor against shaded low, moderate and high bands.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

evaluate_gamma_model(models$model_kcal_maize)
```

::: {.callout-tip}
## Diagnostic Assessment
- **Residuals**: No systematic patterns, approximately normal
- **Leverage**: No overly influential observations
- **Collinearity**: Degrees-of-freedom adjusted variance inflation factors (VIF)
  reach a maximum of `r max_vif_maize`, with `r n_vif_over_5` of
  `r n_vif_terms` terms above the conventional threshold of 5

Factors and polynomial terms contribute several columns to the model matrix, so
their generalized VIF is rescaled by the term's degrees of freedom. This places
every term on the single-coefficient scale that the thresholds of 5 and 10
refer to.
:::


## Modeling Energy Intake from Non-Maize Sources

::: {.callout-note}
## Non-Maize Energy Sources
Non-maize energy sources contribute to dietary diversity and often provide 
additional protein and fat alongside energy. This model captures the portion 
of energy intake that may reflect economic access to a more varied diet.
:::

### Distribution Selection

```{r}
#| label: tbl-c1-evaluate-distribution-non-maize
#| tbl-cap: "Distribution fit comparison for energy intake from non-maize sources"

build_distribution_table_energy(fig_c1_evaluate_distribution_non_maize)
```

```{r}
#| label: fig-c1-evaluate-distribution-non-maize
#| fig-cap: "Distribution assessment for energy intake from non-maize sources"
#| fig-alt: >
#|   Four goodness-of-fit panels comparing observed energy intake from
#|   non-maize sources against a fitted Gamma distribution. Top left, a
#|   histogram of intake overlaid with the theoretical Gamma density curve. Top
#|   right, a quantile-quantile plot; bottom left, empirical versus theoretical
#|   cumulative distribution curves; bottom right, a probability-probability
#|   plot. Points and curves stay close to the reference lines, confirming the
#|   Gamma distribution again fits the right-skewed intake data well.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

build_distribution_plots(fig_c1_evaluate_distribution_non_maize)
```

::: {.callout-important}
## Distribution Selection
**Gamma distribution** again provides the best fit for non-maize energy intake.
:::

### Full Model

```{r}
#| label: tbl-c2-coefficients-full-model-non-maize
#| tbl-cap: "Model coefficients for energy intake from non-maize sources (full model)"

tbl_c2_coefficients_full_model_non_maize
```

```{r}
#| label: fig-c2-diagnostic-plot-full-model-non-maize
#| fig-cap: "Observed vs predicted energy intake from non-maize sources (full model)"
#| fig-alt: >
#|   Scatter chart with observed energy from non-maize sources in kilocalories
#|   per day on the x-axis and model-predicted energy on the y-axis, using all
#|   candidate predictors. Each point is a person. Predictions span a narrower
#|   range than the observed spread, reflecting the greater heterogeneity of
#|   non-maize food sources and a weaker fit than the maize model.

fig_c2_diagnostic_plot_full_model_non_maize |>
  ggplot(aes(x = kcal_individuo_dia_no_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Energy from Non-Maize (kcal/day)",
    y = "Predicted Energy from Non-Maize (kcal/day)"
  )
```

### Variable Selection

```{r}
#| label: tbl-c3-wald-test-non-maize
#| tbl-cap: "Wald test results for predictor significance (energy from non-maize)"

tbl_c3_wald_test_non_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = "Wald F-test results for energy intake from non-maize model"
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-important}
## Variables Retained
**`r n_predictors_non_maize` predictors** significantly predict energy intake from 
non-maize sources, drawing from the same set of demographic, geographic, 
socioeconomic and housing variables. The retained set differs from the maize model: the two outcomes are predicted
by different subsets of the same candidate variables.
:::

### Refined Model

```{r}
#| label: tbl-c4-coefficients-refined-model-non-maize
#| tbl-cap: "Model coefficients for energy intake from non-maize sources (refined model)"

tbl_c4_coefficients_refined_model_non_maize
```

```{r}
#| label: fig-c4-diagnostic-plot-refined-model-non-maize
#| fig-cap: "Observed vs predicted energy intake from non-maize sources (refined model)"
#| fig-alt: >
#|   Scatter chart with observed energy from non-maize sources in kilocalories
#|   per day on the x-axis and model-predicted energy on the y-axis, using only
#|   the significant predictors retained after variable selection. Each point
#|   is a person. The pattern matches the full model, confirming that the
#|   simpler refined model retains its predictive accuracy.

fig_c4_diagnostic_plot_refined_model_non_maize |>
  ggplot(aes(x = kcal_individuo_dia_no_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Energy from Non-Maize (kcal/day)",
    y = "Predicted Energy from Non-Maize (kcal/day)"
  )
```

### Model Diagnostics

```{r}
#| label: fig-c5-evaluate-diagnostics-non-maize
#| fig-cap: "Diagnostic plots for energy from non-maize sources model"
#| fig-alt: >
#|   A grid of regression diagnostic panels for the refined Gamma model of
#|   energy from non-maize sources, covering residual behaviour, influential
#|   observations, and predictor collinearity. Residuals scatter around their
#|   reference line without systematic pattern, no observations show excessive
#|   leverage, and the collinearity panel plots each term's degrees-of-freedom
#|   adjusted variance inflation factor against shaded low, moderate and high
#|   bands.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

evaluate_gamma_model(models$model_kcal_non_maize)
```


## Summary

### Final Model Performance

```{r}
#| label: tbl-model-performance
#| tbl-cap: "Final refined model performance metrics"

dplyr::bind_rows(
  extract_gamma_model_metrics(models$model_kcal_maize,     "Energy from maize"),
  extract_gamma_model_metrics(models$model_kcal_non_maize, "Energy from non-maize")
) |>
  # Reshape: metrics as rows, models as columns
  tidyr::pivot_longer(
    cols = -source,
    names_to = "metric",
    values_to = "value"
  ) |>
  tidyr::pivot_wider(
    names_from = source,
    values_from = value
  ) |>
  # Apply readable labels per metric
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      n_obs                 = "N observations",
      n_predictors          = "N predictors",
      pseudo_r2_mcfadden    = "Pseudo-R² (McFadden)",
      pseudo_r2_cragg_uhler = "Pseudo-R² (Cragg-Uhler)",
      aic                   = "AIC",
      dispersion            = "Dispersion"
    )
  ) |>
  gt() |>
  cols_label(metric = "Metric") |>
  fmt_number(
    rows    = metric == "N observations",
    columns = c(`Energy from maize`, `Energy from non-maize`),
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    rows    = metric == "N predictors",
    columns = c(`Energy from maize`, `Energy from non-maize`),
    decimals = 0
  ) |>
  fmt_number(
    rows    = metric %in% c("Pseudo-R² (McFadden)",
                            "Pseudo-R² (Cragg-Uhler)",
                            "Dispersion"),
    columns = c(`Energy from maize`, `Energy from non-maize`),
    decimals = 3
  ) |>
  fmt_number(
    rows    = metric == "AIC",
    columns = c(`Energy from maize`, `Energy from non-maize`),
    decimals = 0
  ) |>
  tab_header(
    title    = "Final Model Performance",
    subtitle = "Refined gamma regression with log link"
  ) |>
  tab_footnote(
    footnote  = "Pseudo-R² (McFadden) = 1 − (residual deviance / null deviance). Conservative measure; values above 0.2 are conventionally considered good for GLMs.",
    locations = cells_body(columns = metric, rows = metric == "Pseudo-R² (McFadden)")
  ) |>
  tab_footnote(
    footnote  = "Pseudo-R² (Cragg-Uhler), also known as Nagelkerke's R², is normalized to a [0, 1] range and generally yields higher values than McFadden. Both measures are reported for transparency.",
    locations = cells_body(columns = metric, rows = metric == "Pseudo-R² (Cragg-Uhler)")
  ) |>
  tab_options(
    table.font.size            = 11,
    heading.title.font.size    = 13,
    heading.subtitle.font.size = 10,
    table.width                = pct(100)
  )
```

### Model Comparison

| Outcome | Final Predictors | Retained predictors (examples) |
|---------|------------------|--------------------------------|
| Energy from maize | `r n_predictors_maize` variables | Age, department, area, household size |
| Energy from non-maize | `r n_predictors_non_maize` variables | Age, electricity, roof materials, household size |

### Key Findings

1. **Largest retained set of the module**: The non-maize energy model retains
   `r n_predictors_non_maize` predictors, the widest set across the six
   nutrient models

2. **Infrastructure variables**: Electricity and roof materials are retained in
   the non-maize model and not in the maize model

3. **Differential role of area**: Urban/rural area is retained in the maize
   model and not in the non-maize model, so it carries predictive power for
   one intake source and not the other

### Application

These models will be applied to SIVESNU 2018 children to predict their energy 
intake profiles based on their demographic and socioeconomic characteristics. 
The predicted values enable linking nutritional status to stunting outcomes.
