---
title: "Transfer Model: Lysine Intake"
subtitle: "Module 3: Transfer Models"
---

## Why Model Lysine?

Lysine is an essential amino acid that cannot be synthesized by the human body 
and must be obtained from the diet. It is the **first limiting amino acid in maize**, 
meaning maize protein quality is constrained by its lysine content. By modeling 
lysine intake from maize and non-maize sources separately, we can:

1. **Simulate biofortification scenarios** — Predict how quality protein maize (QPM) 
   with higher lysine would change total lysine intake
2. **Preserve baseline non-maize intake** — Keep other dietary sources constant 
   when modeling interventions
3. **Assess protein quality impacts** — As the first limiting amino acid in maize, 
   lysine availability can constrain the overall utilization of dietary protein

::: {.callout-note}
## Why Lysine Matters
In maize-dependent diets, the lysine content of the grain sets a ceiling on how
much of the dietary protein can be used for growth, even when total protein
intake appears adequate. Quality Protein Maize (QPM) varieties carry a higher
lysine content than conventional maize, which is the channel through which
biofortification acts on protein quality.
:::

## Modeling Approach

We use **survey-weighted gamma regression** because lysine intake data is:

- **Positive and continuous** — People consume some amount of lysine daily
- **Right-skewed** — Most people have moderate intake, few have very high intake
- **Population-representative** — Survey weights carry the estimates to the
  national level

The models predict lysine intake based on demographic and socioeconomic 
characteristics that are available in both ENCOVI (training data) and 
SIVESNU (application target).

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  fitdistrplus,       # Distribution fitting for goodness-of-fit assessment
  ggpubr,             # Arrange diagnostic plots into a 2x2 grid
  survey,             # Required for svyglm method dispatch (model diagnostics)
  performance,        # Pseudo-R² (Nagelkerke / Cragg-Uhler) and diagnostics
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))

# --- Local helpers: distribution assessment (web formatting) ---------------
# Recreate the distribution-fit outputs from the raw outcome vector exported
# by the framework, returning table and diagnostic grid as separate objects
# so each renders in its own chunk with web-consistent formatting.

build_distribution_table <- function(outcome) {

  fit_gamma       <- fitdist(data = outcome, "gamma")
  fit_lognormal   <- fitdist(data = outcome, "lnorm")
  fit_exponential <- fitdist(data = outcome, "exp")

  data.frame(
    distribution = c("Gamma", "Lognormal", "Exponential"),
    loglik = c(fit_gamma$loglik, fit_lognormal$loglik, fit_exponential$loglik),
    aic    = c(fit_gamma$aic, fit_lognormal$aic, fit_exponential$aic),
    bic    = c(fit_gamma$bic, fit_lognormal$bic, fit_exponential$bic),
    ks_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$ks,
      gofstat(fit_lognormal, fitnames = "lnorm")$ks,
      gofstat(fit_exponential, fitnames = "exp")$ks
    ),
    cvm_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$cvm,
      gofstat(fit_lognormal, fitnames = "lnorm")$cvm,
      gofstat(fit_exponential, fitnames = "exp")$cvm
    ),
    ad_stat = c(
      gofstat(fit_gamma, fitnames = "gamma")$ad,
      gofstat(fit_lognormal, fitnames = "lnorm")$ad,
      gofstat(fit_exponential, fitnames = "exp")$ad
    )
  ) |>
    gt() |>
    tab_header(
      title = "Distribution Fit Comparison",
      subtitle = "Goodness-of-fit statistics for Gamma, Lognormal, and Exponential distributions"
    ) |>
    cols_label(
      distribution = "Distribution",
      loglik   = "Log-Likelihood",
      aic      = "AIC",
      bic      = "BIC",
      ks_stat  = "Kolmogorov-Smirnov",
      cvm_stat = "Cramer-von Mises",
      ad_stat  = "Anderson-Darling"
    ) |>
    fmt_number(
      columns = c(loglik, aic, bic, ks_stat, cvm_stat, ad_stat),
      decimals = 4
    ) |>
    tab_spanner(label = "Information Criteria", columns = c(aic, bic)) |>
    tab_spanner(label = "Goodness-of-Fit Statistics", columns = c(ks_stat, cvm_stat, ad_stat)) |>
    cols_align(align = "center", columns = everything()) |>
    cols_align(align = "left", columns = distribution) |>
    tab_options(
      table.font.size = 11,
      heading.title.font.size = 13,
      heading.subtitle.font.size = 10,
      table.width = pct(100)
    )
}

build_distribution_plots <- function(outcome) {

  fit_gamma <- fitdist(data = outcome, "gamma")

  shape <- fit_gamma$estimate["shape"]
  rate  <- fit_gamma$estimate["rate"]
  x_obs <- sort(fit_gamma$data)

  plot_density <- data.frame(x = fit_gamma$data) |>
    ggplot(aes(x = x)) +
    geom_histogram(aes(y = after_stat(density)), bins = 30,
                   color = "azure4", fill = "lightgray", alpha = 0.5) +
    stat_function(fun = dgamma, args = list(shape = shape, rate = rate),
                  color = "#3AAF85", linetype = "dashed", linewidth = 0.8) +
    labs(title = "Empirical vs Theoretical Density", x = "", y = "Density") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      plot.subtitle = element_text(color = "azure4", size = 10),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_qq <- data.frame(
    empirical = x_obs,
    theoretical = qgamma(ppoints(length(x_obs)), shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = theoretical, y = empirical)) +
    geom_point(color = "#1B6CA8", size = 1, alpha = 0.2) +
    geom_abline(intercept = 0, slope = 1, color = "#3AAF85",
                linetype = "dashed", linewidth = 0.8) +
    labs(title = "Quantile-Quantile Plot", x = "Theoretical Quantiles", y = "Empirical Quantiles") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_cdf <- data.frame(
    x = x_obs,
    empirical_cdf = ecdf(fit_gamma$data)(x_obs),
    theoretical_cdf = pgamma(x_obs, shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = x)) +
    geom_line(aes(y = empirical_cdf), color = "lightgray", linewidth = 2.5) +
    geom_line(aes(y = theoretical_cdf), color = "#3AAF85",
              linetype = "dashed", linewidth = 0.8) +
    labs(title = "Empirical vs Theoretical CDF",
         subtitle = "CDF: Cumulative Distribution Function", x = "Data", y = "CDF") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      plot.subtitle = element_text(color = "azure4", size = 10),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  plot_pp <- data.frame(
    empirical_prob = ecdf(fit_gamma$data)(x_obs),
    theoretical_prob = pgamma(x_obs, shape = shape, rate = rate)
  ) |>
    ggplot(aes(x = theoretical_prob, y = empirical_prob)) +
    geom_point(color = "#1B6CA8", size = 1, alpha = 0.2) +
    geom_abline(intercept = 0, slope = 1, color = "#3AAF85",
                linetype = "dashed", linewidth = 0.8) +
    labs(title = "P-P Plot", x = "Theoretical Probabilities", y = "Empirical Probabilities") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "azure4", size = 14),
      axis.line = element_line(color = "azure4"),
      axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
      axis.text = element_text(size = 8, colour = "azure4")
    )

  ggarrange(plot_density, plot_qq, plot_cdf, plot_pp, ncol = 2, nrow = 2)
}
```

```{r}
#| label: load-rds
#| include: false

# (a) Web resources for visualizations (exported by framework 03_05)
fig_b1_evaluate_distribution_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_fig_b1_evaluate_distribution_maize.rds"),
  trust = TRUE
)

tbl_b2_coefficients_full_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_tbl_b2_coefficients_full_model_maize.rds"),
  trust = TRUE
)

fig_b2_diagnostic_plot_full_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_fig_b2_diagnostic_plot_full_model_maize.rds"),
  trust = TRUE
)

tbl_b3_wald_test_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_tbl_b3_wald_test_maize.rds"),
  trust = TRUE
)

tbl_b4_coefficients_refined_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_tbl_b4_coefficients_refined_model_maize.rds"),
  trust = TRUE
)

fig_b4_diagnostic_plot_refined_model_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_fig_b4_diagnostic_plot_refined_model_maize.rds"),
  trust = TRUE
)

fig_b5_evaluate_diagnostics_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_fig_b5_evaluate_diagnostics_maize.rds"),
  trust = TRUE
)

fig_c1_evaluate_distribution_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_fig_c1_evaluate_distribution_non_maize.rds"),
  trust = TRUE
)

tbl_c2_coefficients_full_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_tbl_c2_coefficients_full_model_non_maize.rds"),
  trust = TRUE
)

fig_c2_diagnostic_plot_full_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_fig_c2_diagnostic_plot_full_model_non_maize.rds"),
  trust = TRUE
)

tbl_c3_wald_test_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_tbl_c3_wald_test_non_maize.rds"),
  trust = TRUE
)

tbl_c4_coefficients_refined_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_tbl_c4_coefficients_refined_model_non_maize.rds"),
  trust = TRUE
)

fig_c4_diagnostic_plot_refined_model_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_fig_c4_diagnostic_plot_refined_model_non_maize.rds"),
  trust = TRUE
)

fig_c5_evaluate_diagnostics_non_maize <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_05_fig_c5_evaluate_diagnostics_non_maize.rds"),
  trust = TRUE
)

# (b) Fitted models for performance metrics
models <- import(
  here("..", "01_data", "02_processed", "models",
       "03_05_lysine_models.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Significant predictor counts from Wald tests (maize and non-maize)
n_predictors_maize <- tbl_b3_wald_test_maize |>
  dplyr::filter(significant == "*") |>
  nrow()

n_predictors_non_maize <- tbl_c3_wald_test_non_maize |>
  dplyr::filter(significant == "*") |>
  nrow()

# Collinearity of the refined maize model, on the degrees-of-freedom adjusted
# VIF scale reported by the diagnostic grid
vif_maize <- adjust_collinearity_df(models$model_lys_maize)
max_vif_maize <- round(max(vif_maize$VIF), 1)
n_vif_over_5 <- sum(vif_maize$VIF > 5)
n_vif_terms <- nrow(vif_maize)
```


## Data Preparation

### Source Data

We use the cleaned ENCOVI dataset from the previous preparation step, containing 
individuals with complete predictor variables and nutrient intake estimates.

### Handling Extreme Values

Gamma regression requires positive values and is sensitive to extreme outliers. 
We apply conservative outlier detection:

::: {.callout-tip}
## Outlier Threshold
Upper outliers identified using the standard **Tukey method (Q3 + 1.5×IQR)**. 
The threshold suits amino acids:

- Amino acid intake distributions are more stable
- Extreme values are consistent with measurement error
- Amino acid intakes have natural biological limits tied to protein consumption
:::

### Survey Design

All models incorporate ENCOVI's complex survey design (clustering and
expansion weights), which places the estimates at the population level.


## Modeling Lysine Intake from Maize

### Distribution Selection

Before modeling, we assess which statistical distribution best fits the lysine 
intake data. Three candidate distributions are compared:

```{r}
#| label: tbl-b1-evaluate-distribution-maize
#| tbl-cap: "Distribution fit comparison for lysine intake from maize"

build_distribution_table(fig_b1_evaluate_distribution_maize)
```

```{r}
#| label: fig-b1-evaluate-distribution-maize
#| fig-cap: "Distribution assessment for lysine intake from maize"
#| fig-alt: >
#|   Four-panel goodness-of-fit assessment for lysine intake from maize. The
#|   top-left panel overlays a right-skewed histogram of intake with a dashed
#|   gamma density curve that closely tracks the bars. The top-right
#|   quantile-quantile plot and bottom-right probability-probability plot both
#|   show points falling close to the diagonal reference line. The bottom-left
#|   panel compares the empirical and theoretical cumulative distribution
#|   functions, which nearly overlap. The close agreement across all panels
#|   supports the gamma distribution as the best fit.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

build_distribution_plots(fig_b1_evaluate_distribution_maize)
```

::: {.callout-important}
## Distribution Selection
**Gamma distribution** provides the best fit based on:

- Lowest AIC/BIC (information criteria)
- Highest Log-Likelihood

Gamma regression with log link is appropriate for positive, right-skewed 
continuous outcomes.
:::

### Full Model

We first train a model including all 20 candidate predictors to identify 
which variables significantly predict lysine intake from maize.

```{r}
#| label: tbl-b2-coefficients-full-model-maize
#| tbl-cap: "Model coefficients for lysine intake from maize (full model)"

tbl_b2_coefficients_full_model_maize
```

```{r}
#| label: fig-b2-diagnostic-plot-full-model-maize
#| fig-cap: "Observed vs predicted lysine intake from maize (full model)"
#| fig-alt: >
#|   Scatter chart with observed lysine intake from maize in grams per day on
#|   the x-axis and model-predicted intake in grams per day on the y-axis, for
#|   the full model with all 20 candidate predictors. Points concentrate in the
#|   lower intake range and follow a positive but dispersed trend, indicating
#|   moderate predictive accuracy.

fig_b2_diagnostic_plot_full_model_maize |>
  ggplot(aes(x = lys_individuo_dia_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Lysine from Maize (g/day)",
    y = "Predicted Lysine from Maize (g/day)"
  )
```

### Variable Selection

Not all predictors contribute meaningfully to prediction. We use **Wald tests** 
to identify statistically significant variables and remove those that add 
complexity without improving predictions.

```{r}
#| label: tbl-b3-wald-test-maize
#| tbl-cap: "Wald test results for predictor significance (lysine from maize)"

tbl_b3_wald_test_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = "Wald F-test results for lysine intake from maize model"
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-important}
## Variables Retained
**`r n_predictors_maize` predictors** show significant relationships with lysine 
intake from maize, spanning demographic, geographic, socioeconomic, housing 
and asset categories. Variables that do not significantly predict lysine intake 
from maize after controlling for other factors are excluded from the refined 
model.
:::

### Refined Model

The final model is fitted on the significant predictors alone.

```{r}
#| label: tbl-b4-coefficients-refined-model-maize
#| tbl-cap: "Model coefficients for lysine intake from maize (refined model)"

tbl_b4_coefficients_refined_model_maize
```

```{r}
#| label: fig-b4-diagnostic-plot-refined-model-maize
#| fig-cap: "Observed vs predicted lysine intake from maize (refined model)"
#| fig-alt: >
#|   Scatter chart with observed lysine intake from maize in grams per day on
#|   the x-axis and predicted intake in grams per day on the y-axis, for the
#|   refined model using only significant predictors. The positive, dispersed
#|   pattern closely matches the full-model plot, showing that dropping
#|   non-significant predictors preserves predictive accuracy.

fig_b4_diagnostic_plot_refined_model_maize |>
  ggplot(aes(x = lys_individuo_dia_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Lysine from Maize (g/day)",
    y = "Predicted Lysine from Maize (g/day)"
  )
```

### Model Diagnostics

We verify the model meets statistical assumptions:

```{r}
#| label: fig-b5-evaluate-diagnostics-maize
#| fig-cap: "Diagnostic plots for lysine from maize model"
#| fig-alt: >
#|   Grid of regression diagnostic panels for the gamma model of lysine intake
#|   from maize. The panels show residuals plotted against fitted values with no
#|   systematic pattern, residual quantiles that fall approximately along the
#|   normal reference line, a leverage plot with no overly influential
#|   observations, and degrees-of-freedom adjusted variance inflation factors
#|   plotted against shaded low, moderate and high reference bands.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

evaluate_gamma_model(models$model_lys_maize)
```

::: {.callout-tip}
## Diagnostic Assessment
- **Residuals**: No systematic patterns, approximately normal
- **Leverage**: No overly influential observations
- **Collinearity**: Degrees-of-freedom adjusted variance inflation factors (VIF)
  reach a maximum of `r max_vif_maize`, with `r n_vif_over_5` of
  `r n_vif_terms` terms above the conventional threshold of 5

Factors and polynomial terms contribute several columns to the model matrix, so
their generalized VIF is rescaled by the term's degrees of freedom. This places
every term on the single-coefficient scale that the thresholds of 5 and 10
refer to.
:::


## Modeling Lysine Intake from Non-Maize Sources

::: {.callout-note}
## Lysine in Non-Maize Foods
Non-maize sources (especially animal products and legumes) are typically 
lysine-rich. This model captures the higher-quality protein component of 
the diet that complements maize-based nutrition.
:::

### Distribution Selection

```{r}
#| label: tbl-c1-evaluate-distribution-non-maize
#| tbl-cap: "Distribution fit comparison for lysine intake from non-maize sources"

build_distribution_table(fig_c1_evaluate_distribution_non_maize)
```

```{r}
#| label: fig-c1-evaluate-distribution-non-maize
#| fig-cap: "Distribution assessment for lysine intake from non-maize sources"
#| fig-alt: >
#|   Four-panel goodness-of-fit assessment for lysine intake from non-maize
#|   sources. The top-left panel overlays a right-skewed histogram of intake
#|   with a dashed gamma density curve that closely tracks the bars. The
#|   top-right quantile-quantile plot and bottom-right probability-probability
#|   plot both show points falling close to the diagonal reference line. The
#|   bottom-left panel compares the empirical and theoretical cumulative
#|   distribution functions, which nearly overlap. The close agreement across
#|   all panels again supports the gamma distribution as the best fit.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

build_distribution_plots(fig_c1_evaluate_distribution_non_maize)
```

::: {.callout-important}
## Distribution Selection
**Gamma distribution** again provides the best fit for non-maize lysine intake.
:::

### Full Model

```{r}
#| label: tbl-c2-coefficients-full-model-non-maize
#| tbl-cap: "Model coefficients for lysine intake from non-maize sources (full model)"

tbl_c2_coefficients_full_model_non_maize
```

```{r}
#| label: fig-c2-diagnostic-plot-full-model-non-maize
#| fig-cap: "Observed vs predicted lysine intake from non-maize sources (full model)"
#| fig-alt: >
#|   Scatter chart with observed lysine intake from non-maize sources in grams
#|   per day on the x-axis and model-predicted intake in grams per day on the
#|   y-axis, for the full model with all candidate predictors. Points follow a
#|   positive but dispersed trend, indicating moderate predictive accuracy.

fig_c2_diagnostic_plot_full_model_non_maize |>
  ggplot(aes(x = lys_individuo_dia_no_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Lysine from Non-Maize (g/day)",
    y = "Predicted Lysine from Non-Maize (g/day)"
  )
```

### Variable Selection

```{r}
#| label: tbl-c3-wald-test-non-maize
#| tbl-cap: "Wald test results for predictor significance (lysine from non-maize)"

tbl_c3_wald_test_non_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = "Wald F-test results for lysine intake from non-maize model"
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-important}
## Variables Retained
**`r n_predictors_non_maize` predictors** significantly predict lysine intake from 
non-maize sources, drawing from the same set of demographic, geographic, 
socioeconomic and housing variables. The retained set differs from the maize model: the two outcomes are predicted
by different subsets of the same candidate variables.
:::

### Refined Model

```{r}
#| label: tbl-c4-coefficients-refined-model-non-maize
#| tbl-cap: "Model coefficients for lysine intake from non-maize sources (refined model)"

tbl_c4_coefficients_refined_model_non_maize
```

```{r}
#| label: fig-c4-diagnostic-plot-refined-model-non-maize
#| fig-cap: "Observed vs predicted lysine intake from non-maize sources (refined model)"
#| fig-alt: >
#|   Scatter chart with observed lysine intake from non-maize sources in grams
#|   per day on the x-axis and predicted intake in grams per day on the y-axis,
#|   for the refined model using only significant predictors. The positive,
#|   dispersed pattern closely matches the full-model plot, showing that
#|   dropping non-significant predictors preserves predictive accuracy.

fig_c4_diagnostic_plot_refined_model_non_maize |>
  ggplot(aes(x = lys_individuo_dia_no_maiz, y = .fitted)) +
  geom_point(alpha = 0.2, color = "steelblue4") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.caption = element_text(size = 8, hjust = 1, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4", hjust = 1),
    axis.text = element_text(size = 10, color = "azure4")
  ) +
  labs(
    x = "Observed Lysine from Non-Maize (g/day)",
    y = "Predicted Lysine from Non-Maize (g/day)"
  )
```

### Model Diagnostics

```{r}
#| label: fig-c5-evaluate-diagnostics-non-maize
#| fig-cap: "Diagnostic plots for lysine from non-maize sources model"
#| fig-alt: >
#|   Grid of regression diagnostic panels for the gamma model of lysine intake
#|   from non-maize sources. The panels show residuals plotted against fitted
#|   values with no systematic pattern, residual quantiles that fall
#|   approximately along the normal reference line, a leverage plot with no
#|   overly influential observations, and degrees-of-freedom adjusted variance
#|   inflation factors plotted against shaded low, moderate and high bands.
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

evaluate_gamma_model(models$model_lys_non_maize)
```


## Summary

### Final Model Performance

```{r}
#| label: tbl-model-performance
#| tbl-cap: "Final refined model performance metrics"

dplyr::bind_rows(
  extract_gamma_model_metrics(models$model_lys_maize,     "Lysine from maize"),
  extract_gamma_model_metrics(models$model_lys_non_maize, "Lysine from non-maize")
) |>
  # Reshape: metrics as rows, models as columns
  tidyr::pivot_longer(
    cols = -source,
    names_to = "metric",
    values_to = "value"
  ) |>
  tidyr::pivot_wider(
    names_from = source,
    values_from = value
  ) |>
  # Apply readable labels per metric
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      n_obs                 = "N observations",
      n_predictors          = "N predictors",
      pseudo_r2_mcfadden    = "Pseudo-R² (McFadden)",
      pseudo_r2_cragg_uhler = "Pseudo-R² (Cragg-Uhler)",
      aic                   = "AIC",
      dispersion            = "Dispersion"
    )
  ) |>
  gt() |>
  cols_label(metric = "Metric") |>
  fmt_number(
    rows    = metric == "N observations",
    columns = c(`Lysine from maize`, `Lysine from non-maize`),
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    rows    = metric == "N predictors",
    columns = c(`Lysine from maize`, `Lysine from non-maize`),
    decimals = 0
  ) |>
  fmt_number(
    rows    = metric %in% c("Pseudo-R² (McFadden)",
                            "Pseudo-R² (Cragg-Uhler)",
                            "Dispersion"),
    columns = c(`Lysine from maize`, `Lysine from non-maize`),
    decimals = 3
  ) |>
  fmt_number(
    rows    = metric == "AIC",
    columns = c(`Lysine from maize`, `Lysine from non-maize`),
    decimals = 0
  ) |>
  tab_header(
    title    = "Final Model Performance",
    subtitle = "Refined gamma regression with log link"
  ) |>
  tab_footnote(
    footnote  = "Pseudo-R² (McFadden) = 1 − (residual deviance / null deviance). Conservative measure; values above 0.2 are conventionally considered good for GLMs.",
    locations = cells_body(columns = metric, rows = metric == "Pseudo-R² (McFadden)")
  ) |>
  tab_footnote(
    footnote  = "Pseudo-R² (Cragg-Uhler), also known as Nagelkerke's R², is normalized to a [0, 1] range and generally yields higher values than McFadden. Both measures are reported for transparency.",
    locations = cells_body(columns = metric, rows = metric == "Pseudo-R² (Cragg-Uhler)")
  ) |>
  tab_options(
    table.font.size            = 11,
    heading.title.font.size    = 13,
    heading.subtitle.font.size = 10,
    table.width                = pct(100)
  )
```

### Model Comparison

| Outcome | Final Predictors | Retained predictors (examples) |
|---------|------------------|--------------------------------|
| Lysine from maize | `r n_predictors_maize` variables | Age, department, education, area |
| Lysine from non-maize | `r n_predictors_non_maize` variables | Age, household size, area, wall materials, rooms |

### Key Findings

1. **Same retained set as digestible protein**: The lysine models select the
   same predictors as the digestible protein models, which follows from lysine
   being one component of the protein intake being modeled

2. **Geographic variation**: Department is retained in both models, consistent
   with regional variation in dietary patterns

3. **Housing variables in the non-maize model**: Wall materials and number of
   rooms are retained for non-maize lysine and not for maize lysine

### Application

These models will be applied to SIVESNU 2018 children to predict their lysine 
intake profiles based on their demographic and socioeconomic characteristics. 
The predicted values enable linking nutritional status to stunting outcomes.
