---
title: "Prepare Nutrient Transfer Modeling Data"
subtitle: "Module 3: Transfer Models"
---


## Why Transfer Models?

Children in the SIVESNU 2018 data (the nutrition surveillance system) have anthropometric 
measurements (height, weight) but **no dietary intake data**. Meanwhile, ENCOVI 2023 
households have detailed food consumption records but **no child growth measurements**.

Transfer models bridge this gap: we use demographic and socioeconomic characteristics 
common to both datasets to predict nutrient intake for SIVESNU children based on 
patterns observed in ENCOVI.

## Overview

This document describes the data preparation steps required before building 
prediction models. High-quality models require:

1. **Complete data** — Missing values must be handled appropriately
2. **Independent predictors** — Highly correlated variables create instability
3. **Informative variables** — Variables with no variation provide no predictive value
4. **Normalized distributions** — Extreme skewness can bias predictions

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  car,                # Box-Cox/Yeo-Johnson profile plots
  gt,                 # Table formatting and presentation
  tidyverse           # General purpose data manipulation
)

# Packages used via :: (not loaded to minimize memory footprint)
# ggcorrplot — ggcorrplot() for correlation matrix visualization

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))
```

```{r}
#| label: load-rds
#| include: false

# Web resources for visualizations (exported by framework 03_01)
tbl_b1_diagnose_missing <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_01_tbl_b1_diagnose_missing.rds"),
  trust = TRUE
)

tbl_b2_impute_missing <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_01_tbl_b2_impute_missing.rds"),
  trust = TRUE
)

tbl_b2b_completion_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_01_tbl_b2b_completion_summary.rds"),
  trust = TRUE
)

fig_c1_correlation_analysis <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_01_fig_c1_correlation_analysis.rds"),
  trust = TRUE
)

tbl_c2_near_zero_variance <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_01_tbl_c2_near_zero_variance.rds"),
  trust = TRUE
)

fig_d1_evaluate_yj_miembros_hogar <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_01_fig_d1_evaluate_yj_miembros_hogar.rds"),
  trust = TRUE
)

fig_d1_evaluate_yj_n_cuartos <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_01_fig_d1_evaluate_yj_n_cuartos.rds"),
  trust = TRUE
)

fig_d2_verify_yj_miembros_hogar <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_01_fig_d2_verify_yj_miembros_hogar.rds"),
  trust = TRUE
)

fig_d2_verify_yj_n_cuartos <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "03_01_fig_d2_verify_yj_n_cuartos.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Imputation quality metrics from B2 table
oobe_row <- tbl_b2_impute_missing |>
  dplyr::filter(metric == "OOBE (Continuous Variables)")

oobe_value <- dplyr::pull(oobe_row, value)

# The imputation engine returns NA for OOBE when no continuous variable
# required imputation. Reporting the raw NA leaves an empty value in the
# prose, so the reason recorded in the table is reported instead.
oobe_display <- if (is.na(oobe_value)) {
  paste0("N/A (", tolower(dplyr::pull(oobe_row, interpretation)), ")")
} else {
  as.character(round(oobe_value, 2))
}

pfc_value <- tbl_b2_impute_missing |>
  dplyr::filter(metric == "PFC (Categorical Variables)") |>
  dplyr::pull(value) |>
  round(2)

# Missing value summary statistics
n_vars_missing <- nrow(tbl_b1_diagnose_missing)
max_pct_missing <- tbl_b1_diagnose_missing |>
  dplyr::pull(pct_missing) |>
  max() |>
  round(1)
worst_var_missing <- tbl_b1_diagnose_missing |>
  dplyr::arrange(dplyr::desc(pct_missing)) |>
  dplyr::slice(1) |>
  dplyr::pull(variable)
```


## Data Loading and Variable Selection

### Source Data

We start with the individual-level nutrient intake dataset created in Module 1, 
containing the full sample of individuals from ENCOVI 2023 across all 22 departments 
of Guatemala.

### Predictor Variables

Transfer models predict nutrient intake based on characteristics that are 
available in **both** ENCOVI and SIVESNU. The shared variables span 
six categories:

| Category | Variables |
|----------|-----------|
| **Demographic** | Age, sex, household relationship |
| **Geographic** | Department, urban/rural area |
| **Household composition** | Number of members, education level, property ownership |
| **Housing characteristics** | Dwelling type, construction materials, rooms, sanitation |
| **Basic services** | Water source, garbage collection |
| **Household assets** | Electricity, television, telephone, computer |

Variables unique to one dataset (e.g., detailed food consumption in ENCOVI,
anthropometric measurements in SIVESNU) cannot be used as predictors.

### Functional Form of the Predictors

The candidate set entering each nutrient model comprises 20 terms, with the
functional form set by the type of variable:

- **Continuous predictors** enter as orthogonal polynomials: age at degree 2,
  and household members, mean household education and number of rooms at
  degree 3. The polynomial terms let intake vary non-linearly across the range
  of each variable, which matters for age in particular, where intake rises
  steeply through childhood and flattens in adulthood.
- **Nominal predictors** (household relationship, sex, department, urban/rural
  area, property regime, dwelling type, construction materials, sanitation,
  water source, garbage collection, and the asset indicators) enter as factors,
  contributing one coefficient per level.



## Missing Value Treatment

### Why This Matters

Prediction models require complete data for all observations, and the
imputation method has to hold the joint structure of the predictors so that
later correlation and variance filtering operate on realistic values.

### Our Approach: Random Forest Imputation

We use a **Random Forest-based multivariate imputation procedure**, which
predicts missing values based on patterns in the complete data. The 
implementation uses the `missRanger` algorithm with a multi-threaded 
`ranger` engine for computational efficiency on large surveys. This method:

- Preserves correlations between variables
- Handles both numeric and categorical data
- Provides out-of-bag error metrics (OOBE for continuous, PFC for categorical)
  to verify imputation accuracy

### Missing Data Summary

```{r}
#| label: tbl-b1-diagnose-missing
#| tbl-cap: "Missing value summary for predictor variables"

tbl_b1_diagnose_missing |>
  gt() |>
  tab_header(
    title = "Missing Values in Predictor Variables",
    subtitle = "Variables requiring imputation before modeling"
  ) |>
  cols_label(
    variable = "Variable",
    n_missing = "N Missing",
    pct_missing = "% Missing"
  ) |>
  fmt_number(
    columns = n_missing,
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    columns = pct_missing,
    decimals = 2,
    pattern = "{x}%"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

Only `r n_vars_missing` variables have missing data, with the worst case affecting
less than `r max_pct_missing`% of observations (`r worst_var_missing`). Most variables 
are complete.

### Imputation Quality

```{r}
#| label: tbl-b2-impute-missing
#| tbl-cap: "Random Forest imputation quality metrics"

tbl_b2_impute_missing |>
  gt() |>
  tab_header(
    title = "Missing Value Imputation Quality Assessment",
    subtitle = "Out-of-bag error estimates from Random Forest imputation"
  ) |>
  cols_label(
    metric = "Metric",
    value = "Value",
    interpretation = "Quality Assessment"
  ) |>
  fmt_number(
    columns = value,
    decimals = 4
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "Acceptable"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "Review required"
    )
  ) |>
  tab_footnote(
    footnote = "OOBE: aggregated out-of-bag error across the continuous variables imputed, bounded in [0, 1] and invariant to their scale. Values < 0.5 indicate acceptable imputation quality for continuous variables.",
    locations = cells_body(columns = metric, rows = 1)
  ) |>
  tab_footnote(
    footnote = "PFC: Proportion of Falsely Classified entries. Values < 0.2 indicate acceptable imputation quality for categorical variables.",
    locations = cells_body(columns = metric, rows = 2)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

::: {.callout-tip}
## Quality Assessment
OOBE (for continuous variables) and PFC (for categorical variables) are
out-of-bag (OOB) error metrics — calculated from observations that were
randomly excluded during tree construction within the Random Forest
algorithm, providing an internal quality metric without requiring a separate
test set. The ranges conventionally treated as acceptable for this imputation
method are OOBE < 0.50 and PFC < 0.20.

- **OOBE = `r oobe_display`**
- **PFC = `r pfc_value`** (categorical variables) — within the acceptable range
:::

### Residual Missing Value Completion

The Random Forest imputation engine can, in rare cases, leave residual 
missing values for a small subset of rows when the out-of-bag prediction 
collapses inside a terminal node with zero within-leaf variance. This
affects a small fraction of the sample and is resolved by a final
completion step, which leaves the predictor matrix complete.

::: {.callout-note}
## Completion Rule
Residual missing values are completed with a variable-specific rule:

- **Numeric / integer variables**: median of the observed values, which 
  preserves the discrete support of ordinal scales (e.g., years of schooling 
  taking values such as 1, 1.5, 2, 2.5).
- **Factor variables**: mode of the observed values.
:::

```{r}
#| label: tbl-b2b-completion-summary
#| tbl-cap: "Residual missing value completion after Random Forest imputation"

tbl_b2b_completion_summary |>
  gt() |>
  tab_header(
    title = "Residual Missing Value Completion",
    subtitle = "Values completed by variable-specific rule after RF imputation"
  ) |>
  cols_label(
    variable = "Variable",
    n_completed = "N Completed",
    rule = "Completion Rule"
  ) |>
  fmt_number(
    columns = n_completed,
    decimals = 0,
    use_seps = TRUE
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      rows = rule == "Integrity check: passed"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(
      rows = rule == "Integrity check: review required"
    )
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

::: {.callout-tip}
## Integrity Check
The final row of the table reports the number of residual missing values
**after** the completion step. A value of zero means the predictor matrix is
complete before correlation analysis, near-zero variance filtering, and model
training.
:::


## Variable Quality Assessment

Before building models, we evaluate whether all candidate predictors provide 
useful information. Two common issues can degrade model performance:

### Multicollinearity

When two variables measure essentially the same thing, including both creates 
redundancy and statistical instability.

```{r}
#| label: fig-c1-correlation-analysis
#| fig-cap: "Correlation matrix of predictor variables"
#| fig-alt: >
#|   Correlation heatmap of the predictor variables, shown as an upper
#|   triangular matrix with the same variables along both axes, reordered by
#|   hierarchical clustering. Each cell is labelled with its correlation
#|   coefficient and colour-coded from orange for negative through white near
#|   zero to blue for positive. Most pairs are weakly correlated, but number of
#|   bedrooms and total rooms stand out as strongly positively correlated
#|   (r = 0.8), flagging the redundancy that leads to dropping one of them.
#| fig-width: 10
#| fig-height: 10
#| warning: false

fig_c1_correlation_analysis |>
  ggcorrplot::ggcorrplot(
    hc.order = TRUE,
    colors = c("#E46726", "white", "#6D9EC1"),
    type = "lower",
    lab = TRUE
  )
```

::: {.callout-important}
## Decision: Remove Redundant Variable
**Number of bedrooms** (`n_dormitorios`) correlates strongly (r > 0.8) with
**total rooms** (`n_cuartos`). Total rooms is retained as the measure of
overall dwelling size, and number of bedrooms is dropped from the candidate
set.
:::

### Near-Zero Variance

Variables where almost all observations fall in a single category provide 
minimal discriminatory power for prediction.

```{r}
#| label: tbl-c2-near-zero-variance
#| tbl-cap: "Near-zero variance analysis for predictor variables"
#| results: asis

analyze_variance(tbl_c2_near_zero_variance)
```

::: {.callout-important}
## Decision: Remove Low-Information Variables
Two variables flagged for near-zero variance:

- **Kitchen type** (`cocina`): nearly all households have the same type
- **Landline telephone** (`telefonia_fija`): nearly all households lack landlines

These variables cannot meaningfully distinguish between nutrient profiles
and are excluded from the model.
:::

## Variable Transformations

### Why Transform?

Many statistical models assume predictor variables follow approximately normal 
distributions. Highly skewed variables (e.g., household size, where most 
households are small but some are very large) can bias predictions.

### Yeo-Johnson Transformation

We apply the Yeo-Johnson transformation, which automatically determines the 
optimal adjustment for each variable. A lambda (λ) value near 1 indicates 
the variable is already approximately normal; values near 0 indicate 
logarithmic transformation is needed.

```{r}
#| label: fig-d1-evaluate-yj-miembros-hogar
#| fig-cap: "Yeo-Johnson transformation profile for household members"
#| fig-alt: >
#|   Profile likelihood curve for the Yeo-Johnson transformation of household
#|   members. The transformation parameter lambda is on the x-axis and the
#|   log-likelihood on the y-axis, with vertical dashed lines marking the 95%
#|   confidence interval for the best lambda. The curve peaks near lambda of 0,
#|   indicating a right-skewed variable that benefits from a near-logarithmic
#|   transformation.

boxCox(
  lm(miembros_hogar ~ 1, data = fig_d1_evaluate_yj_miembros_hogar),
  main = "Household Members - Transformation Profile",
  family = "yjPower"
)
```

```{r}
#| label: fig-d1-evaluate-yj-n-cuartos
#| fig-cap: "Yeo-Johnson transformation profile for number of rooms"
#| fig-alt: >
#|   Profile likelihood curve for the Yeo-Johnson transformation of number of
#|   rooms. The transformation parameter lambda is on the x-axis and the
#|   log-likelihood on the y-axis, with vertical dashed lines marking the 95%
#|   confidence interval for the best lambda. The curve peaks near lambda of 0,
#|   indicating a right-skewed variable that benefits from a near-logarithmic
#|   transformation.

boxCox(
  lm(n_cuartos ~ 1, data = fig_d1_evaluate_yj_n_cuartos),
  main = "Number of Rooms - Transformation Profile",
  family = "yjPower"
)
```

Both household members and number of rooms show λ ≈ 0, indicating right-skewed 
distributions that benefit from transformation.

```{r}
#| label: fig-d2-verify-yj-miembros-hogar
#| fig-cap: "Transformation profile for household members (post-transformation)"
#| fig-alt: >
#|   Profile likelihood curve for the already-transformed household members
#|   variable. The transformation parameter lambda is on the x-axis and the
#|   log-likelihood on the y-axis, with vertical dashed lines marking the 95%
#|   confidence interval for the best lambda. The curve now peaks near lambda of
#|   1, confirming that the Yeo-Johnson transformation has successfully
#|   normalized the distribution.

boxCox(
  lm(miembros_hogar_yj ~ 1, data = fig_d2_verify_yj_miembros_hogar),
  main = "Household Members - Post-Transformation",
  family = "yjPower"
)
```

```{r}
#| label: fig-d2-verify-yj-n-cuartos
#| fig-cap: "Transformation profile for number of rooms (post-transformation)"
#| fig-alt: >
#|   Profile likelihood curve for the already-transformed number of rooms
#|   variable. The transformation parameter lambda is on the x-axis and the
#|   log-likelihood on the y-axis, with vertical dashed lines marking the 95%
#|   confidence interval for the best lambda. The curve now peaks near lambda of
#|   1, confirming that the Yeo-Johnson transformation has successfully
#|   normalized the distribution.

boxCox(
  lm(n_cuartos_yj ~ 1, data = fig_d2_verify_yj_n_cuartos),
  main = "Number of Rooms - Post-Transformation",
  family = "yjPower"
)
```

::: {.callout-tip}
## Transformation Result
After transformation, both variables show λ ≈ 1, the value expected for an
approximately normal distribution. The transformation parameters are saved for
consistent application to SIVESNU data.
:::


## Summary of Data Preparation

This preparation module brings the modeling dataset to the state the transfer
models operate on:

| Step | Action | Result |
|------|--------|--------|
| Missing values | Random Forest imputation (`missRanger` / `ranger`) | Imputed predictor matrix with OOB diagnostics |
| Residual completion | Median (numeric) / mode (factor) safety net | 100% complete data |
| Multicollinearity | Remove `n_dormitorios` | Independent predictors |
| Near-zero variance | Remove `cocina`, `telefonia_fija` | Informative variables only |
| Normalization | Yeo-Johnson transformation | Improved distributions |

### Final ENCOVI Dataset

The prepared dataset retains the full sample of individuals from ENCOVI 2023, 
with predictor variables filtered for quality and continuous predictors 
normalized via Yeo-Johnson transformation. The transformation parameters 
are exported for consistent application to SIVESNU data during the modeling 
of individual nutrients documented in subsequent sections.
