---
title: "Transfer Height to ENCOVI"
subtitle: "Module 5: Meta-analysis"
---

## Why Transfer Height to ENCOVI?

ENCOVI 2023 does not include anthropometric measurements for children. To
estimate the stunting impact of biofortified maize, we need individual-level
height predictions that can be converted to height-for-age z-scores (HAZ).
This module bridges that gap through a four-stage pipeline:

1. **GAM prediction** — Apply the
   [trained height model](05_02_transfer_model_height.qmd) to ENCOVI 2023
   children using their socioeconomic profiles
2. **Quantile mapping** — Calibrate predicted heights against
   Guatemala-specific reference distributions derived from ENSMI stunting
   prevalence and WHO growth standards
3. **HAZ calculation** — Convert calibrated heights to height-for-age
   z-scores and classify stunting status
4. **Weight calibration** — Adjust survey weights to match ENSMI 2014-2015
   departmental stunting prevalence

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))
```

```{r}
#| label: load-rds
#| include: false

# Web resources for visualizations (exported by framework 05_03)
tbl_a3_lms_overview <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_a3_lms_overview.rds"),
  trust = TRUE
)

tbl_b1_population_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_b1_population_summary.rds"),
  trust = TRUE
)

tbl_b2_exclusion_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_b2_exclusion_summary.rds"),
  trust = TRUE
)

tbl_b5_missing_values <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_b5_missing_values.rds"),
  trust = TRUE
)

tbl_c2_prediction_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_c2_prediction_summary.rds"),
  trust = TRUE
)

fig_c2_predicted_height_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_fig_c2_predicted_height_distribution.rds"),
  trust = TRUE
)

tbl_d1_strata_counts <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_d1_strata_counts.rds"),
  trust = TRUE
)

tbl_d2_reference_params <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_d2_reference_params.rds"),
  trust = TRUE
)

tbl_d4_qm_height_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_d4_qm_height_comparison.rds"),
  trust = TRUE
)

tbl_d4_qm_by_stratum <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_d4_qm_by_stratum.rds"),
  trust = TRUE
)

fig_d4_height_density_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_fig_d4_height_density_comparison.rds"),
  trust = TRUE
)

tbl_e2_haz_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_e2_haz_summary.rds"),
  trust = TRUE
)

tbl_e2_haz_by_department <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_e2_haz_by_department.rds"),
  trust = TRUE
)

fig_e2_haz_density <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_fig_e2_haz_density.rds"),
  trust = TRUE
)

fig_e2_haz_by_age_sex <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_fig_e2_haz_by_age_sex.rds"),
  trust = TRUE
)

tbl_f3_calibration_diagnostics <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_f3_calibration_diagnostics.rds"),
  trust = TRUE
)

tbl_f3_departmental_validation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_f3_departmental_validation.rds"),
  trust = TRUE
)

tbl_f3_national_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_f3_national_summary.rds"),
  trust = TRUE
)

tbl_g1_final_validation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_03_tbl_g1_final_validation.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Final N children retained (from tbl-g1 — "Total children 6-59 months")
n_children_final <- tbl_g1_final_validation |>
  filter(Check == "Total children 6-59 months") |>
  pull(Value) |>
  as.integer()

# National stunting prevalence target (from tbl-f3-national-summary)
ensmi_national_stunting <- tbl_f3_national_summary |>
  filter(Metric == "ENSMI 2014-2015 official stunting prevalence") |>
  pull(Value)

# Departments matched exactly (from tbl-g1 — "Departments with ENSMI match (0.00 pp)")
n_departments_matched <- tbl_g1_final_validation |>
  filter(Check == "Departments with ENSMI match (0.00 pp)") |>
  pull(Value) |>
  as.integer()

# Calibrated weighted stunting prevalence (national) for use in narrative
calibrated_stunting <- tbl_f3_national_summary |>
  filter(Metric == "Stunting prevalence (HAZ < -2)") |>
  pull(Value)

# Calibrated weighted mean HAZ
calibrated_mean_haz <- tbl_f3_national_summary |>
  filter(Metric == "Mean HAZ (weighted)") |>
  pull(Value)

# Weight CV calibrated and adjustment ratio range from f3-calibration-diagnostics
cv_calibrated_value <- tbl_f3_calibration_diagnostics |>
  filter(Metric == "Weight CV (calibrated)") |>
  pull(Value)

adjustment_ratio_range <- tbl_f3_calibration_diagnostics |>
  filter(Metric == "Adjustment ratio range") |>
  pull(Value)

mean_adjustment_ratio <- tbl_f3_calibration_diagnostics |>
  filter(Metric == "Mean adjustment ratio") |>
  pull(Value)

cv_initial_value <- tbl_f3_calibration_diagnostics |>
  filter(Metric == "Weight CV (initial)") |>
  pull(Value)

deff_value <- tbl_f3_calibration_diagnostics |>
  filter(Metric == "Design Effect (DEFF)") |>
  pull(Value)

n_ratios_extreme_high <- tbl_f3_calibration_diagnostics |>
  filter(Metric == "Ratios > 3.0 (extreme high)") |>
  pull(Value)

# Missingness in the ENCOVI predictor set, ordered by share of records
tbl_b5_ordered <- tbl_b5_missing_values |>
  arrange(desc(Pct_Missing))

var_missing_max    <- tbl_b5_ordered$Variable[1]
pct_missing_max    <- tbl_b5_ordered$Pct_Missing[1]
var_missing_second <- tbl_b5_ordered$Variable[2]
pct_missing_second <- tbl_b5_ordered$Pct_Missing[2]

n_vars_with_missing <- nrow(tbl_b5_ordered)

n_missing_education <- tbl_b5_ordered |>
  filter(Variable == "grado_estudios_hogar") |>
  pull(N_Missing)

# Cohort restriction and quantile mapping effect
n_excluded_dietary <- tbl_b2_exclusion_summary |>
  filter(Metric == "Excluded: incomplete dietary record") |>
  pull(Value)

n_cohort_pre <- tbl_b2_exclusion_summary |>
  filter(Metric == "Children 6-59 months (pre-exclusion)") |>
  pull(Value)

# Formatted here so both ends of the comparison print on the same scale: the
# default drops a trailing zero and would render one as 10.7 and the other
# as 11.23.
sd_before_qm <- tbl_d4_qm_height_comparison |>
  filter(Stage == "Before QM (predicted)") |>
  pull(SD_Height) |>
  sprintf(fmt = "%.2f")

sd_after_qm <- tbl_d4_qm_height_comparison |>
  filter(Stage == "After QM (calibrated)") |>
  pull(SD_Height) |>
  sprintf(fmt = "%.2f")

n_strata_qm <- nrow(tbl_d4_qm_by_stratum)

# Departmental gaps against ENSMI before the weights are calibrated
n_dept_over_10pp <- sum(abs(tbl_e2_haz_by_department$`Diff (pp)`) > 10)
n_dept_5_to_10pp <- sum(abs(tbl_e2_haz_by_department$`Diff (pp)`) > 5 &
                        abs(tbl_e2_haz_by_department$`Diff (pp)`) <= 10)
n_dept_total     <- nrow(tbl_e2_haz_by_department)
max_dept_gap_pp  <- round(max(abs(tbl_e2_haz_by_department$`Diff (pp)`)), 1)

# National deviations reported alongside the ENSMI reference
stunting_deviation <- tbl_f3_national_summary |>
  filter(Metric == "Stunting deviation from ENSMI") |>
  pull(Value)

haz_deviation <- tbl_f3_national_summary |>
  filter(Metric == "HAZ deviation from ENSMI") |>
  pull(Value)
```


## Data Sources

Three external inputs are required: the trained GAM model with its
transformation objects, from the two preceding steps of this module; the WHO
LMS growth reference parameters; and the ENSMI 2014-2015 departmental stunting
targets.

```{r}
#| label: tbl-a3-lms-overview
#| tbl-cap: "WHO LMS reference table summary"

tbl_a3_lms_overview |>
  gt() |>
  tab_header(
    title = "WHO LMS Reference Parameters",
    subtitle = "Length/height-for-age by sex and age in months"
  ) |>
  cols_label(
    sexo = "Sex",
    N_ages = "Age Points",
    Age_range = "Age Range",
    M_range = "Median (SD0) Height Range"
  ) |>
  cols_align(align = "left") |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Target Population and Exclusions

### Population Identification

```{r}
#| label: tbl-b1-population-summary
#| tbl-cap: "ENCOVI 2023 children 6-59 months target population"

tbl_b1_population_summary |>
  gt() |>
  tab_header(
    title = "Target Population Identification",
    subtitle = "ENCOVI 2023 children aged 6-59 months, on the age assigned in Module 1"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Dietary Record Completeness

The analytical cohort is restricted to children whose household reports a
complete dietary record: the food consumption module supplies their
source-specific nutrient intake and the plant-source share of that intake. Both
quantities are required downstream by the bioavailability, inadequacy and
nutrient-profile calculations, and the twelve intake variables are recorded
together, so completeness is required on the full set rather than on any single
component.

The restriction is applied before weight calibration, so the departmental
stunting prevalence targets are calibrated over the population that reaches the
analysis. It removes `r n_excluded_dietary` of the `r n_cohort_pre` children
aged 6-59 months.

```{r}
#| label: tbl-b2-exclusion-summary
#| tbl-cap: "Records excluded for an incomplete dietary record"

tbl_b2_exclusion_summary |>
  gt() |>
  tab_header(
    title = "Dietary Record Completeness",
    subtitle = "Cohort restricted to children with nutrient intake and plant-source share"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Predictor Harmonization and Imputation

ENCOVI variables are recoded to match the coding structure used in
[SIVESNU](05_01_prepare_sivesnu_height_data.qmd), and the normalization objects
fitted there are applied to the continuous predictors, so both surveys enter
the model on the same scale. Missing values are imputed via Random Forest using
the `impute_rf_parallel()` utility function, which wraps
`missRanger::missRanger()` with a multi-threaded `ranger` engine.

```{r}
#| label: tbl-b5-missing-values
#| tbl-cap: "Missing values in GAM predictor variables"

tbl_b5_missing_values |>
  gt() |>
  tab_header(
    title = "Missing Values in Predictor Variables",
    subtitle = "ENCOVI 2023 children 6-59 months"
  ) |>
  fmt_number(columns = Pct_Missing, decimals = 1, pattern = "{x}%") |>
  fmt_number(columns = N_Missing, decimals = 0) |>
  cols_align(align = "left", columns = Variable) |>
  cols_align(align = "right", columns = c(N_Missing, Pct_Missing)) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Where the Missingness Sits
`r n_vars_with_missing` predictors carry missing values in the analytical
cohort. Two of them account for almost all of it: `r var_missing_max` at
`r round(pct_missing_max, 1)`% of records and `r var_missing_second` at
`r round(pct_missing_second, 1)`%, both service-provision indicators.
`grado_estudios_hogar` is missing for `r n_missing_education` records. Random
Forest imputation fills these values from the rest of the predictor set.
:::


## Height Prediction

The trained GAM model is applied to generate height predictions for each
ENCOVI child based on their socioeconomic profile.

```{r}
#| label: tbl-c2-prediction-summary
#| tbl-cap: "Predicted height summary by age group"

tbl_c2_prediction_summary |>
  gt() |>
  tab_header(
    title = "Predicted Height Distribution by Age Group",
    subtitle = "ENCOVI 2023 children 6-59 months — GAM predictions"
  ) |>
  cols_label(
    age_group = "Age Group",
    N = "N",
    Mean_Height = "Mean (cm)",
    SD_Height = "SD (cm)",
    Min_Height = "Min (cm)",
    Max_Height = "Max (cm)"
  ) |>
  cols_align(align = "left", columns = age_group) |>
  cols_align(align = "right", columns = -age_group) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-c2-predicted-height-distribution
#| fig-cap: "Predicted height distribution by age group and sex"
#| fig-alt: >
#|   Faceted density plot with one panel per age group, showing GAM-predicted
#|   height in centimetres on the x-axis and density on the y-axis, with
#|   separate overlaid curves for boys and girls. Within each panel the curves
#|   are narrow and concentrated, illustrating that predictions capture only
#|   the conditional mean and so compress variance relative to true population
#|   spread, a compression corrected by the later quantile-mapping step.
#| fig-width: 10
#| fig-height: 7

fig_c2_predicted_height_distribution |>
  ggplot(aes(x = talla_cm_predicted, fill = sexo)) +
  geom_density(alpha = 0.5, color = NA) +
  facet_wrap(~ age_group, scales = "free") +
  scale_fill_manual(
    values = c("Hombre" = "#1B6CA8",
               "Mujer" = "#3AAF85")
  ) +
  labs(
    title = "Predicted Height Distribution by Age Group and Sex",
    subtitle = "ENCOVI 2023 children 6-59 months — GAM transfer model",
    x = "Predicted Height (cm)",
    y = "Density",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    strip.text = element_text(face = "bold", color = "azure4"),
    legend.position = "bottom",
    legend.text = element_text(size = 8, color = "azure4")
  )
```

::: {.callout-note}
## Prediction Variance
GAM predictions capture the conditional mean of height given socioeconomic
predictors, producing narrower distributions than expected population
variability. This variance compression is corrected by quantile mapping in
the next step.
:::


## Quantile Mapping Calibration

### Stratification

Children are partitioned into 18 strata (9 six-month age bands × 2 sexes)
for quantile mapping. Each stratum receives its own reference distribution
parameters.

```{r}
#| label: tbl-d1-strata-counts
#| tbl-cap: "Sample size per age-sex stratum"

tbl_d1_strata_counts |>
  gt() |>
  tab_header(
    title = "Sample Size per Age-Sex Stratum",
    subtitle = "ENCOVI 2023 children 6-59 months — 18 strata for quantile mapping"
  ) |>
  tab_spanner(label = "Hombre", columns = starts_with("Hombre")) |>
  tab_spanner(label = "Mujer", columns = starts_with("Mujer")) |>
  cols_label(
    age_band = "Age Band",
    Hombre_N = "N",
    Hombre_N_weighted = "Weighted N",
    Mujer_N = "N",
    Mujer_N_weighted = "Weighted N"
  ) |>
  fmt_number(columns = contains("weighted"), decimals = 0) |>
  cols_align(align = "left", columns = age_band) |>
  cols_align(align = "right", columns = -age_band) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Guatemala-Specific Reference Distributions

For each stratum, the expected height distribution is derived from WHO LMS
parameters and ENSMI national stunting prevalence
(`r ensmi_national_stunting`). The Guatemala-specific mean (mu) is solved
from the equation P(height < threshold) = p_stunting, and sigma is set to
the WHO reference SD.

```{r}
#| label: tbl-d2-reference-params
#| tbl-cap: "Guatemala-specific reference distribution parameters by stratum"

tbl_d2_reference_params |>
  gt() |>
  tab_header(
    title = "Guatemala-Specific Reference Distributions",
    subtitle = paste0(
      "Derived from WHO LMS parameters and ENSMI 2014-2015 stunting prevalence (",
      ensmi_national_stunting, ")"
    )
  ) |>
  fmt_number(
    columns = c(`WHO Median (M)`, `Stunting Threshold (cm)`,
                `Guatemala mu (cm)`, `Guatemala sigma (cm)`),
    decimals = 2
  ) |>
  cols_align(align = "left", columns = c(`Age Band`, Sex)) |>
  cols_align(align = "right", columns = -c(`Age Band`, Sex)) |>
  tab_source_note(
    md("**mu_guatemala** = height_threshold - sigma × Φ⁻¹(p_stunting)<br>**sigma_guatemala** = M × S (WHO reference SD)")
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-important}
## Quantile Mapping Rationale
GAM predictions compress variance relative to the true population. Normal
quantile mapping corrects this by transforming each predicted height
through the source CDF and back through the Guatemala-specific inverse CDF,
preserving rank order while expanding the distribution to match the
expected population variability.
:::

### Quantile Mapping Application and Diagnostics

```{r}
#| label: tbl-d4-qm-height-comparison
#| tbl-cap: "Height distribution before and after quantile mapping"

tbl_d4_qm_height_comparison |>
  gt() |>
  tab_header(
    title = "Height Distribution: Before vs After Quantile Mapping",
    subtitle = "Normal QM within 6-month age bands by sex — Guatemala-specific reference"
  ) |>
  cols_label(
    Stage = "Stage",
    Mean_Height = "Mean (cm)",
    SD_Height = "SD (cm)",
    Median_Height = "Median (cm)",
    Min_Height = "Min (cm)",
    Max_Height = "Max (cm)"
  ) |>
  cols_align(align = "left", columns = Stage) |>
  cols_align(align = "right", columns = -Stage) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-d4-qm-by-stratum
#| tbl-cap: "Quantile mapping effect by age-sex stratum"

tbl_d4_qm_by_stratum |>
  gt() |>
  tab_header(
    title = "Quantile Mapping Effect by Stratum",
    subtitle = "Comparison of predicted, calibrated, and reference parameters"
  ) |>
  cols_align(align = "left", columns = c(`Age Band`, Sex)) |>
  cols_align(align = "right", columns = -c(`Age Band`, Sex)) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-d4-height-density-comparison
#| fig-cap: "Height distribution before and after quantile mapping by age band"
#| fig-alt: >
#|   Faceted density plot with one panel per six-month age band, showing height
#|   in centimetres on the x-axis and density on the y-axis. Each panel overlays
#|   two curves: the narrow predicted distribution before calibration and the
#|   wider calibrated distribution after quantile mapping. From the 12-month
#|   band onward the calibrated curve is visibly wider than the predicted one;
#|   in the 6-11 month band the two curves nearly coincide.
#| fig-width: 10
#| fig-height: 10

fig_d4_height_density_comparison |>
  ggplot(aes(x = height, fill = stage)) +
  geom_density(alpha = 0.4, color = NA) +
  facet_wrap(~ age_band, scales = "free", ncol = 3) +
  scale_fill_manual(values = c("Predicted" = "lightgray", "Calibrated" = "#1B6CA8")) +
  labs(
    title = "Height Distribution: Before vs After Quantile Mapping",
    subtitle = "Normal QM within 6-month age bands — Guatemala-specific reference",
    x = "Height (cm)",
    y = "Density",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    strip.text = element_text(face = "bold", color = "azure4"),
    legend.position = "bottom",
    legend.text = element_text(size = 8, color = "azure4")
  )
```

::: {.callout-note}
## Quantile Mapping Effect
Calibration moves the height distributions onto the Guatemala-specific
reference means and widens their spread, taking the pooled SD from
`r sd_before_qm` cm to `r sd_after_qm` cm. @tbl-d4-qm-by-stratum reports the
calibrated mean against the reference mean for each of the `r n_strata_qm`
strata.
:::


## HAZ Calculation and Stunting Classification

Calibrated heights are converted to height-for-age z-scores using the WHO
LMS formula. For length/height-for-age, L = 1 across all ages, simplifying
to HAZ = (height − M) / (M × S), where M is the median height-for-age from
the WHO reference and S is the coefficient of variation. Stunting is
classified as HAZ < −2.

```{r}
#| label: tbl-e2-haz-summary
#| tbl-cap: "HAZ distribution summary: raw vs calibrated"

tbl_e2_haz_summary |>
  gt() |>
  tab_header(
    title = "HAZ Distribution: Raw vs Calibrated",
    subtitle = "Raw HAZ from GAM-predicted height; calibrated HAZ from QM-adjusted height"
  ) |>
  cols_label(
    Stage = "Stage",
    Mean_HAZ = "Mean",
    SD_HAZ = "SD",
    Median_HAZ = "Median",
    Pct_Stunted = "Stunting %",
    Range = "Range"
  ) |>
  cols_align(align = "left", columns = Stage) |>
  cols_align(align = "right", columns = -Stage) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Departmental HAZ Before Weight Calibration

```{r}
#| label: tbl-e2-haz-by-department
#| tbl-cap: "Calibrated HAZ and stunting prevalence by department"

tbl_e2_haz_by_department |>
  gt() |>
  tab_header(
    title = "Departmental HAZ and Stunting Prevalence (Pre-Calibration)",
    subtitle = "Survey-weighted estimates before weight calibration — comparison with ENSMI targets"
  ) |>
  fmt_number(columns = c(`Mean HAZ`, `SD HAZ`), decimals = 3) |>
  fmt_number(columns = c(`Stunting %`, `ENSMI Target %`, `Diff (pp)`), decimals = 1) |>
  cols_align(align = "left", columns = Department) |>
  cols_align(align = "right", columns = -Department) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(
      columns = `Diff (pp)`,
      rows = abs(`Diff (pp)`) > 10
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(
      columns = `Diff (pp)`,
      rows = abs(`Diff (pp)`) > 5 & abs(`Diff (pp)`) <= 10
    )
  ) |>
  tab_footnote(
    footnote = "Red: > 10 pp deviation. Yellow: 5-10 pp deviation. Weight calibration corrects these differences.",
    locations = cells_column_labels(columns = `Diff (pp)`)
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-warning}
## Pre-Calibration Deviations
Before weight calibration, `r n_dept_over_10pp` of `r n_dept_total`
departments sit more than 10 percentage points away from their ENSMI stunting
target and a further `r n_dept_5_to_10pp` between 5 and 10, with the widest gap
at `r max_dept_gap_pp` pp. The gaps run in both directions, which places them
in the geographic distribution of the model's predictions rather than in a
uniform level shift. The weight calibration step closes them.
:::

### HAZ Distribution Visualization

```{r}
#| label: fig-e2-haz-density
#| fig-cap: "HAZ distribution: raw vs calibrated"
#| fig-alt: >
#|   Density plot of height-for-age z-score on the x-axis against density on
#|   the y-axis, overlaying two curves: raw HAZ from GAM-predicted height and
#|   calibrated HAZ from quantile-mapped height. A dashed vertical line marks
#|   the stunting threshold at HAZ equal to minus 2. The calibrated curve is
#|   wider than the compressed raw curve and its centre sits slightly higher,
#|   so it leaves less mass below the threshold and moves closer to Guatemala's
#|   reference prevalence.
#| fig-width: 10
#| fig-height: 6

fig_e2_haz_density |>
  ggplot(aes(x = haz, fill = stage)) +
  geom_density(alpha = 0.4, color = NA) +
  geom_vline(xintercept = -2, color = "#E07A5F", linetype = "dashed",
             linewidth = 0.6) +
  scale_fill_manual(
    values = c("Raw (predicted height)" = "lightgray",
               "Calibrated (QM height)" = "#1B6CA8")
  ) +
  annotate(
    "text", x = -2.1, y = Inf,
    label = "Stunting threshold (HAZ = -2)",
    hjust = 1, vjust = 1.5, size = 3, color = "#E07A5F"
  ) +
  labs(
    title = "HAZ Distribution: Raw vs Calibrated",
    subtitle = "Raw from GAM-predicted height; calibrated from QM-adjusted height",
    x = "Height-for-Age Z-score",
    y = "Density",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    legend.position = "bottom",
    legend.text = element_text(size = 8, color = "azure4")
  )
```

```{r}
#| label: fig-e2-haz-by-age-sex
#| fig-cap: "Calibrated HAZ distribution by age band and sex"
#| fig-alt: >
#|   Faceted density plot with one panel per six-month age band, showing
#|   calibrated height-for-age z-score on the x-axis and density on the y-axis,
#|   with separate overlaid curves for boys and girls and a dashed vertical
#|   line at the stunting threshold of HAZ equal to minus 2. Across panels the
#|   distributions for both sexes are broadly similar in location and spread,
#|   indicating consistent calibrated HAZ across age bands and sexes.
#| fig-width: 10
#| fig-height: 10

fig_e2_haz_by_age_sex |>
  ggplot(aes(x = haz_calibrated, fill = sexo)) +
  geom_density(alpha = 0.4, color = NA) +
  geom_vline(xintercept = -2, color = "#E07A5F", linetype = "dashed",
             linewidth = 0.5) +
  facet_wrap(~ age_band, ncol = 3) +
  scale_fill_manual(values = c("Hombre" = "#1B6CA8", "Mujer" = "#3AAF85")) +
  labs(
    title = "Calibrated HAZ by Age Band and Sex",
    subtitle = "After Guatemala-specific quantile mapping",
    x = "HAZ (calibrated)",
    y = "Density",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    strip.text = element_text(face = "bold", color = "azure4"),
    legend.position = "bottom",
    legend.text = element_text(size = 8, color = "azure4")
  )
```


## Weight Calibration

Survey weights are adjusted so that weighted stunting prevalence matches
ENSMI 2014-2015 departmental targets exactly. This corrects residual
discrepancies between the model-implied stunting rates and official
estimates, while preserving total weighted population per department.

### Calibration Diagnostics

```{r}
#| label: tbl-f3-calibration-diagnostics
#| tbl-cap: "Weight calibration diagnostics"

tbl_f3_calibration_diagnostics |>
  gt() |>
  tab_header(
    title = "Survey Weight Calibration Diagnostics",
    subtitle = "Calibrated to ENSMI 2014-2015 departmental stunting prevalence"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Departmental Stunting Validation

```{r}
#| label: tbl-f3-departmental-validation
#| tbl-cap: "Stunting prevalence by department: calibrated vs ENSMI targets"

tbl_f3_departmental_validation |>
  gt() |>
  tab_header(
    title = "Departmental Stunting Prevalence Validation",
    subtitle = "Calibrated weighted prevalence vs ENSMI 2014-2015 targets"
  ) |>
  fmt_number(columns = c(`ENSMI Target (%)`, `Calibrated (%)`), decimals = 1) |>
  fmt_number(columns = `Diff (pp)`, decimals = 2) |>
  fmt_number(columns = N, decimals = 0) |>
  cols_align(align = "left", columns = Department) |>
  cols_align(align = "right", columns = -Department) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(
      columns = `Diff (pp)`,
      rows = abs(`Diff (pp)`) > 1
    )
  ) |>
  tab_footnote(
    footnote = "Red: departments with > 1 pp deviation from target.",
    locations = cells_column_labels(columns = `Diff (pp)`)
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-important}
## What the Calibration Costs
The `r n_departments_matched` departments reach their ENSMI stunting target
to 0.00 pp. The mean adjustment ratio is `r mean_adjustment_ratio`, so the
total weight is preserved, but the adjustment is not evenly distributed:
individual ratios span `r adjustment_ratio_range`, and
`r n_ratios_extreme_high` units sit above 3.0.

The dispersion of the weights therefore rises. Their coefficient of variation
moves from `r cv_initial_value` to `r cv_calibrated_value`, and the design
effect stands at `r deff_value`, which means the calibrated sample carries the
precision of one about `r round(as.numeric(deff_value), 2)` times smaller.
Departmental alignment is obtained at that price.
:::

### National Indicators

```{r}
#| label: tbl-f3-national-summary
#| tbl-cap: "National stunting and HAZ indicators after weight calibration"

tbl_f3_national_summary |>
  gt() |>
  tab_header(
    title = "National Stunting and HAZ Indicators",
    subtitle = "ENCOVI 2023 children 6-59 months — after departmental weight calibration to ENSMI 2014-2015"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Metric %in% c("Stunting prevalence (HAZ < -2)",
                           "Mean HAZ (weighted)")
    )
  ) |>
  tab_style(
    style = cell_text(style = "italic", color = "azure4"),
    locations = cells_body(
      columns = everything(),
      rows = Metric %in% c("ENSMI 2014-2015 official stunting prevalence",
                           "ENSMI 2014-2015 official mean HAZ")
    )
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## National Indicators
After calibration the weighted stunting prevalence is
`r calibrated_stunting` against the ENSMI national figure of
`r ensmi_national_stunting`, a gap of `r stunting_deviation`. The calibration
targets are departmental, so the national aggregate also carries the
composition of ENCOVI's sample across departments, which need not match the
composition behind the ENSMI national figure. The weighted mean HAZ is
`r calibrated_mean_haz` against the official `r tbl_f3_national_summary |> filter(Metric == "ENSMI 2014-2015 official mean HAZ") |> pull(Value)`,
a difference of `r haz_deviation`.
:::


## Final Validation Summary

A final validation pass checks completeness of the output dataset and
verifies that the calibration procedure achieves its targets at the
departmental level.

```{r}
#| label: tbl-g1-final-validation
#| tbl-cap: "Final validation summary"

tbl_g1_final_validation |>
  gt() |>
  tab_header(
    title = "Final Validation Summary",
    subtitle = "Height transfer, quantile mapping, and weight adjustment"
  ) |>
  cols_align(align = "left", columns = Check) |>
  cols_align(align = "right", columns = c(Value, Status)) |>
  tab_style(
    style = cell_text(color = "#2E7D32", weight = "bold"),
    locations = cells_body(columns = Status, rows = Status == "PASS")
  ) |>
  tab_style(
    style = cell_text(color = "#C62828", weight = "bold"),
    locations = cells_body(columns = Status, rows = Status == "FAIL")
  ) |>
  tab_style(
    style = cell_text(color = "#F57F17", weight = "bold"),
    locations = cells_body(columns = Status, rows = Status == "WARN")
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Summary

### Key Findings

1. **Height prediction**: The GAM transfer model produces a height for each of
   `r format(n_children_final, big.mark = ",")` ENCOVI 2023 children aged 6-59
   months from their socioeconomic profile.

2. **Quantile mapping**: Normal quantile mapping within `r n_strata_qm`
   age-sex strata widens the predicted height distribution from an SD of
   `r sd_before_qm` cm to `r sd_after_qm` cm and aligns each stratum with its
   Guatemala-specific reference parameters.

3. **HAZ calculation**: Calibrated heights give a weighted national stunting
   prevalence of `r calibrated_stunting` and a mean HAZ of
   `r calibrated_mean_haz`, `r stunting_deviation` and `r haz_deviation` from
   the ENSMI 2014-2015 official figures.

4. **Weight calibration**: The `r n_departments_matched` departments reach
   their ENSMI stunting target to 0.00 pp, with the weight coefficient of
   variation moving from `r cv_initial_value` to `r cv_calibrated_value`.

::: {.callout-warning}
## Limitations

1. **Model-based heights** — Height values are predicted from socioeconomic
   proxies rather than measured, so the estimates hold at the level of the
   distribution and not of the individual child.

2. **ENSMI temporal gap** — Calibration targets come from ENSMI 2014-2015 and
   the ENCOVI data are from 2023, so any change in stunting between the two
   collection periods is absorbed into the calibration.

3. **Quantile mapping assumes normality** — The normal quantile mapping treats
   height as Gaussian within each age-sex stratum, which bounds how well the
   tails of the calibrated distribution are reproduced.

4. **Weight dispersion** — Calibration raises the weight coefficient of
   variation from `r cv_initial_value` to `r cv_calibrated_value` and leaves
   `r n_ratios_extreme_high` units with an adjustment ratio above 3.0, which
   widens the sampling variance of the weighted estimates.
:::

### Application

This dataset — containing predicted height, calibrated HAZ, stunting
classification, and calibrated survey weights for
`r format(n_children_final, big.mark = ",")` children — feeds the
[bioavailability and inadequacy calculations](05_04_calculate_bioavailability_inadequacy.qmd)
and the
[QPM stunting impact estimation](05_05_calculate_qpm_stunting_impact.qmd).
