---
title: "Bioavailability and Nutrient Inadequacy"
subtitle: "Module 5: Meta-analysis"
---

## Why Calculate Bioavailability and Inadequacy?

Raw nutrient intake values do not directly indicate nutrient adequacy. Iron
and zinc pass through absorption processes that reduce the fraction available
to the body, and the thresholds for adequacy vary by age. This module bridges
dietary intake and nutrient adequacy by applying bioavailability corrections
and classifying each child under two conditions: **baseline** (conventional
maize) and **QPM intervention** (biofortified maize at 100% coverage).

The QPM intervention pre-computes the full nutritional recalculation at the
individual level, so that downstream scenario scripts only need to select
which children are impacted by a given production coverage level.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  scales,             # Number formatting (comma) for log-scale density plot
  patchwork,          # Multi-panel plot composition (fig-h2)
  glue,               # String interpolation for diagnostic labels
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))
```

```{r}
#| label: load-rds
#| include: false

# Web resources for visualizations (exported by framework 05_04)

# Phase B — Chispitas integration (NEW)
tbl_b3_validate_chispitas_integration <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_b3_validate_chispitas_integration.rds"),
  trust = TRUE
)

# Phase C — Implausible intake detection and imputation
tbl_c1_detect_implausible_intake <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_c1_detect_implausible_intake.rds"),
  trust = TRUE
)

tbl_c1_sensitivity <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_c1_sensitivity.rds"),
  trust = TRUE
)

tbl_c4_imputation_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_c4_imputation_comparison.rds"),
  trust = TRUE
)

fig_c4_density_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_fig_c4_density_comparison.rds"),
  trust = TRUE
)

# Phase C5 — Non-maize protein outlier treatment (NEW)
tbl_c5_prot_nomaiz_outlier_detection <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_c5_prot_nomaiz_outlier_detection.rds"),
  trust = TRUE
)

tbl_c5_prot_imputation_diagnostics <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_c5_prot_imputation_diagnostics.rds"),
  trust = TRUE
)

# Phase F — Baseline diagnostics
tbl_f1_bioavailability_by_age <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_f1_bioavailability_by_age.rds"),
  trust = TRUE
)

tbl_f2_inadequacy_prevalence <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_f2_inadequacy_prevalence.rds"),
  trust = TRUE
)

fig_f3a_iron_adequacy_ratio <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_fig_f3a_iron_adequacy_ratio.rds"),
  trust = TRUE
)

fig_f3b_zinc_adequacy_ratio <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_fig_f3b_zinc_adequacy_ratio.rds"),
  trust = TRUE
)

fig_f3c_protein_adequacy_ratio <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_fig_f3c_protein_adequacy_ratio.rds"),
  trust = TRUE
)

tbl_f4_stunting_inadequacy <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_f4_stunting_inadequacy.rds"),
  trust = TRUE
)

# Phase H — QPM comparison
tbl_h1_baseline_vs_qpm_inadequacy <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_h1_baseline_vs_qpm_inadequacy.rds"),
  trust = TRUE
)

fig_h2_delta_absorbed_nutrients <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_fig_h2_delta_absorbed_nutrients.rds"),
  trust = TRUE
)

tbl_h3_bioavailability_by_age_qpm <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_04_tbl_h3_bioavailability_by_age_qpm.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Selected maize ceiling (fixed pipeline constant, kcal/day) and its
# equivalent in grams at the INCAP conversion factor for nixtamalized maize
MAIZE_KCAL_CEILING <- 2500
KCAL_PER_G_MAIZE   <- 3.65
MAIZE_G_EQUIV      <- round(MAIZE_KCAL_CEILING / KCAL_PER_G_MAIZE)

# Supplementation adherence, as used by the framework. The factor is a scalar
# carried on every row of the integration table.
chispitas_adherence <- tbl_b3_validate_chispitas_integration |>
  pull(adherence_national) |>
  first()

# National flagging stats (from tbl-c1-detect-implausible-intake row "NATIONAL")
national_row <- tbl_c1_detect_implausible_intake |>
  filter(departamento == "NATIONAL")

n_flagged    <- national_row |> pull(n_flagged)
pct_flagged  <- national_row |> pull(pct_flagged)

# Non-maize protein outlier diagnostics (from tbl-c5 outlier detection)
prot_nomaiz_fence_val <- tbl_c5_prot_nomaiz_outlier_detection |>
  filter(Metric == "Tukey fence (g/day)") |>
  pull(Value)

n_prot_flagged <- tbl_c5_prot_nomaiz_outlier_detection |>
  filter(Metric == "Children exceeding fence") |>
  pull(Value)

pct_prot_flagged <- tbl_c5_prot_nomaiz_outlier_detection |>
  filter(Metric == "Percentage flagged") |>
  pull(Value)

# Departmental spread of the maize flagging rate
dept_flagging <- tbl_c1_detect_implausible_intake |>
  filter(departamento != "NATIONAL")

max_dept_flagged_pct <- max(dept_flagging$pct_flagged)
min_dept_flagged_pct <- min(dept_flagging$pct_flagged)
dept_max_flagged <- dept_flagging |>
  slice_max(pct_flagged, n = 1) |>
  pull(departamento)

# Maize intake means used to compare the imputed values against valid records
maize_means <- tbl_c4_imputation_comparison |>
  filter(variable == "gr_individuo_dia_maiz")

mean_valid_maize   <- maize_means |> filter(source == "Valid") |> pull(mean)
mean_outlier_maize <- maize_means |> filter(source == "Original outlier") |> pull(mean)
mean_imputed_maize <- maize_means |> filter(source == "Imputed (k-NN)") |> pull(mean)

# Which nutrient leads the inadequacy ranking in each age-sex stratum
inadequacy_leader <- tbl_f2_inadequacy_prevalence |>
  mutate(leader = case_when(
    pct_fe_inadeq >= pct_zn_inadeq & pct_fe_inadeq >= pct_prot_inadeq ~ "Iron",
    pct_zn_inadeq >= pct_prot_inadeq ~ "Zinc",
    TRUE ~ "Protein"
  ))

n_strata_iron_leads <- sum(inadequacy_leader$leader == "Iron")
n_strata_total <- nrow(inadequacy_leader)

# Sensitivity grid around the selected ceiling
sens_grid <- tbl_c1_sensitivity |>
  arrange(ceiling_kcal)

sens_low  <- sens_grid |> slice_min(ceiling_kcal, n = 1)
sens_high <- sens_grid |> slice_max(ceiling_kcal, n = 1)

# National baseline inadequacy, by nutrient (from tbl-h1)
inadeq_baseline <- tbl_h1_baseline_vs_qpm_inadequacy |>
  select(Nutrient,
         baseline = `Baseline Inad. (%)`,
         qpm = `QPM Inad. (%)`,
         change = `Change (pp)`)

get_inadeq <- function(nutrient, column) {
  inadeq_baseline[[column]][inadeq_baseline$Nutrient == nutrient]
}

fe_inadeq_baseline   <- get_inadeq("Iron (absorbed)", "baseline")
zn_inadeq_baseline   <- get_inadeq("Zinc (TAZ)", "baseline")
prot_inadeq_baseline <- get_inadeq("Protein (PDCAAS)", "baseline")
any_inadeq_baseline  <- get_inadeq("Any nutrient", "baseline")
any_inadeq_qpm       <- get_inadeq("Any nutrient", "qpm")

fe_inadeq_drop   <- abs(get_inadeq("Iron (absorbed)", "change"))
zn_inadeq_drop   <- abs(get_inadeq("Zinc (TAZ)", "change"))
prot_inadeq_drop <- abs(get_inadeq("Protein (PDCAAS)", "change"))

# Stunting by number of inadequacies (from tbl-f4)
f4 <- tbl_f4_stunting_inadequacy

f4_highest <- f4 |> slice_max(stunting_prev, n = 1)
f4_lowest  <- f4 |> slice_min(stunting_prev, n = 1)
f4_no_inad <- f4 |> filter(inadequacy_group == "No inadequacy")
f4_others_max_maize <- f4 |>
  filter(inadequacy_group != "No inadequacy") |>
  pull(maize_g) |>
  max()
```


## Parameters and Data

The analysis requires four sets of parameters: the Miller equation
coefficients for zinc absorption in children, the iron absorption factor
for plant-based diets, age-specific nutrient inadequacy thresholds (H-AR),
and the QPM biofortification factors from the Semilla Nueva Standard Report
(2020).

::: {.callout-note}
## Bioavailability Models
Iron uses a uniform 5% absorption factor (Allen et al., 2006) appropriate
for the predominantly plant-based dietary patterns in this population. Zinc
uses the age-dependent Miller equation (Miller et al., 2015), which captures
saturation kinetics where fractional absorption decreases at higher intakes.
Protein quality is assessed via PDCAAS adjustment applied at the food level
during the [intake calculation](../01_nutritional_intakes/01_03_intake_calculation.qmd).
:::


## Chispitas Supplementation Integration

Guatemala's Chispitas program distributes Multiple Micronutrient Powder
(MNP) sachets to children aged 6-59 months through the public health
system. Each sachet contains 10 mg iron and 4.1 mg zinc, administered in
60-day cycles every ~180 days. To convert episodic supplementation to
average daily contribution, a 1/3 adjustment factor is applied (60 active
days / 180 day cycle), yielding 3.33 mg Fe/day and 1.37 mg Zn/day during
active periods.

ENCOVI 2023 records dietary intake and does not capture supplementation, so the
expected Chispitas contribution is estimated from external sources. The
integration uses:

- **Departmental coverage rates**: Official INE population projections
  (2022), representing the proportion of children 6-59 months receiving
  first Chispitas delivery by department.
- **National adherence factor**: `r round(chispitas_adherence, 3)`, the
  survey-weighted mean of the adherence estimated in
  [Chispitas Supplementation](../04_covariate_prediction/04_02_integrate_chispitas_supplementation.qmd),
  taken over the SIVESNU 2018 children recorded as programme recipients. The
  adherence cascade reflects sequential filtering through access,
  administration, and dosing compliance.

The expected daily contribution per child is computed as the product of
composition, departmental coverage, and national adherence, and is added to
each child's non-maize iron and zinc intake. Chispitas contains no protein,
so protein intake variables remain unchanged.

**Source of the adherence factor**: SIVESNU 2018 is the nationally
representative source of Chispitas adherence data for Guatemala.

```{r}
#| label: tbl-b3-validate-chispitas-integration
#| tbl-cap: "Chispitas supplementation integration by department"

tbl_b3_validate_chispitas_integration |>
  select(-adherence_national) |>
  gt() |>
  tab_header(
    title = "Chispitas Integration: Non-Maize Intake Before and After Adjustment",
    subtitle = "Expected delta by department (Fe and Zn only)"
  ) |>
  cols_label(
    departamento = "Department",
    chispitas_coverage = "Coverage",
    mean_fe_original = "Fe Original",
    mean_fe_adjusted = "Fe Adjusted",
    delta_fe = "Δ Fe",
    mean_zn_original = "Zn Original",
    mean_zn_adjusted = "Zn Adjusted",
    delta_zn = "Δ Zn"
  ) |>
  fmt_percent(columns = chispitas_coverage, decimals = 1) |>
  fmt_number(
    columns = c(mean_fe_original, mean_fe_adjusted, delta_fe,
                mean_zn_original, mean_zn_adjusted, delta_zn),
    decimals = 3
  ) |>
  tab_spanner(label = "Iron (mg/day)",
              columns = c(mean_fe_original, mean_fe_adjusted, delta_fe)) |>
  tab_spanner(label = "Zinc (mg/day)",
              columns = c(mean_zn_original, mean_zn_adjusted, delta_zn)) |>
  tab_source_note(
    glue("Coverage: INE 2022. Adherence: {round(chispitas_adherence, 3)} (SIVESNU 2018). Composition: 3.33 mg Fe, 1.37 mg Zn per day.")
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```


## Implausible Intake Detection and Treatment

Two screens flag a record as implausible.

**Maize ceiling.** Maize energy intake above `r format(MAIZE_KCAL_CEILING, big.mark = ",")`
kcal/day is physiologically implausible for children aged 6-59 months. The
ceiling corresponds to approximately `r MAIZE_G_EQUIV` g/day of nixtamalized
maize at `r KCAL_PER_G_MAIZE` kcal/g (INCAP) and exceeds the total daily energy
requirement for any child in this age range. The FAO Food Balance Sheet for
Guatemala reports 226 g/capita/day of maize for adults.

**Non-maize screen.** The upper tail of the non-maize distribution reaches
values far above the median, so each non-maize component — iron, zinc, lysine,
tryptophan and energy — is screened against its own upper Tukey fence at
$Q3 + 3 \times IQR$. The multiplier is 3 rather than the usual 1.5 because the
rule declares a record implausible and triggers imputation: on non-maize energy
it places the cut at a level that, from non-maize sources alone, already
exceeds the total energy requirement for this age range.

The ENCOVI 2023 food consumption module records intake at the household level
and distributes it to individuals via AME coefficients, so an implausible value
on either side signals unreliable data across the entire nutrient profile. The
treatment invalidates all twelve source-specific nutrient variables plus maize
and non-maize grams for the flagged records, and imputes plausible values via
k-nearest neighbours (RANN k-d tree, k = 10, inverse-distance weighting).

**Predictors.** The imputation is conditioned on the child's own ENCOVI record:
age, sex, department, area, household education and the plant-source share of
the diet. The quantity being reconstructed is an ENCOVI dietary value, so it is
rebuilt from variables observed for the same child in the same survey.

Five of the six are collected outside the food consumption module. The
plant-source share is a composition ratio derived from it, and is retained on
that basis: the reporting error the screen targets scales the reported
quantities of a household and leaves their relative composition intact, so the
ratio carries dietary-pattern signal that survives the error invalidating the
amounts.

The set carries one variable per dimension of the matching space — demographics,
geography, socioeconomic position and diet composition. All are continuous or
binary, which keeps the scaled Euclidean metric well defined and leaves each
dimension weighted once.

```{r}
#| label: tbl-c1-detect-implausible-intake
#| tbl-cap: "Implausible intake detection summary"

tbl_c1_detect_implausible_intake |>
  gt() |>
  tab_header(
    title = "Implausible Intake Detection",
    subtitle = glue(
      "Maize ceiling: {format(MAIZE_KCAL_CEILING, big.mark = ',')} kcal/day ",
      "(= {MAIZE_G_EQUIV} g/day) | non-maize: Q3 + 3·IQR per component"
    )
  ) |>
  fmt_number(columns = c(n_total, n_flagged), decimals = 0, use_seps = TRUE) |>
  fmt_number(columns = pct_flagged, decimals = 1) |>
  cols_label(
    departamento = "Department",
    n_total = "N Total",
    n_flagged = "N Flagged",
    pct_flagged = "% Flagged"
  ) |>
  tab_row_group(label = "By Department", rows = departamento != "NATIONAL") |>
  tab_row_group(label = "National", rows = departamento == "NATIONAL") |>
  tab_source_note(
    glue("Conversion: {KCAL_PER_G_MAIZE} kcal/g (INCAP, nixtamalized maize tortilla). FAO FBS Guatemala: 226 g/capita/day (adults).")
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-c1-sensitivity
#| tbl-cap: "Sensitivity analysis across ceiling thresholds"

tbl_c1_sensitivity |>
  gt() |>
  tab_header(
    title = "Sensitivity Analysis: Maize Energy Ceiling",
    subtitle = "Impact of threshold selection on flagging rate"
  ) |>
  fmt_number(columns = c(equiv_g), decimals = 0) |>
  fmt_number(columns = ceiling_kcal, decimals = 0, use_seps = TRUE) |>
  fmt_number(columns = n_flagged, decimals = 0, use_seps = TRUE) |>
  fmt_number(columns = pct_flagged, decimals = 1) |>
  cols_label(
    ceiling_kcal = "Ceiling (kcal/day)",
    equiv_g = "Equiv. g/day",
    n_flagged = "N Flagged",
    pct_flagged = "% Flagged"
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F4FD"),
    locations = cells_body(rows = ceiling_kcal == MAIZE_KCAL_CEILING)
  ) |>
  tab_source_note(
    glue("Highlighted row: selected ceiling. Conversion: {KCAL_PER_G_MAIZE} kcal/g nixtamalized maize (INCAP).")
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-important}
## Threshold Selection
The `r format(MAIZE_KCAL_CEILING, big.mark = ",")` kcal/day ceiling, together
with the non-maize screen, flags `r format(n_flagged, big.mark = ",")` records
nationally (`r round(pct_flagged, 1)`%). The sensitivity analysis places the
maize component in the range of available options: a
`r format(sens_low$ceiling_kcal, big.mark = ",")` kcal/day ceiling would flag
`r round(sens_low$pct_flagged, 1)`% of records on that criterion alone and a
`r format(sens_high$ceiling_kcal, big.mark = ",")` kcal/day ceiling
`r round(sens_high$pct_flagged, 1)`%. The selected value sits above the total
daily energy requirement of a child in this age range, which is what makes it
implausible rather than merely high. Flagging rates reach
`r round(max_dept_flagged_pct, 1)`% in `r dept_max_flagged`, against
`r round(min_dept_flagged_pct, 1)`% at the lower end, so the affected records
are concentrated geographically.
:::

### Imputation Diagnostics

```{r}
#| label: tbl-c4-imputation-comparison
#| tbl-cap: "Nutritional variable comparison: valid vs original outliers vs imputed"

tbl_c4_imputation_comparison |>
  gt(groupname_col = "variable") |>
  tab_header(
    title = "Post-Imputation Diagnostic: Valid vs Outlier vs Imputed",
    subtitle = glue("N flagged: {n_flagged} ({round(pct_flagged, 1)}%) | k = 10, inverse-distance weighted")
  ) |>
  fmt_number(columns = c(mean, median, sd, max, p95), decimals = 2) |>
  cols_label(
    source = "Source",
    mean = "Mean",
    median = "Median",
    sd = "SD",
    max = "Max",
    p95 = "P95"
  ) |>
  tab_source_note("Valid: non-flagged records. Original: pre-invalidation values. Imputed: k-NN RANN replacement.") |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-c4-density-comparison
#| fig-cap: "Density comparison: valid (blue) vs original outliers (red) vs imputed (green)"
#| fig-alt: >
#|   Faceted density plot with one panel per nutrient variable, showing intake
#|   value on a log10 x-axis and density on the y-axis. Each panel overlays three
#|   distributions: valid non-flagged records, original outlier values before
#|   invalidation, and k-nearest-neighbour imputed replacements. The imputed
#|   distributions fall within the range of the valid records but are narrower
#|   and more sharply peaked, the concentration around the conditional mean that
#|   k-nearest-neighbour imputation produces. The original outlier distributions
#|   sit roughly one order of magnitude further right on the logarithmic axis.
#| fig-width: 10
#| fig-height: 8
#| warning: false

fig_c4_density_comparison |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = 0.45, color = NA) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50",
             linewidth = 0.3) +
  facet_wrap(~ variable, scales = "free", ncol = 3) +
  scale_fill_manual(
    values = c("Valid" = "#1B6CA8",
               "Original outlier" = "#E74C3C",
               "Imputed (k-NN)" = "#3AAF85"),
    name = ""
  ) +
  scale_x_log10(labels = comma) +
  coord_cartesian(xlim = c(1, NA)) +
  labs(
    title = "Outlier Treatment — Density Comparison",
    subtitle = glue(
      "N flagged: {n_flagged} | Maize ceiling: ",
      "{format(MAIZE_KCAL_CEILING, big.mark = ',')} kcal/day | k-NN RANN (k = 10)"
    ),
    x = "Intake value",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 11, color = "azure4"),
    strip.text = element_text(size = 10, face = "bold", color = "azure4"),
    axis.title = element_text(size = 9, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    legend.position = "bottom",
    legend.text = element_text(size = 10),
    panel.grid.minor = element_blank()
  )
```

::: {.callout-note}
## Where the Imputed Values Land
The k-NN imputation (k = 10, inverse-distance weighted) places the replaced
records inside the range of the valid ones. Mean maize intake is
`r round(mean_imputed_maize, 0)` g/day among the imputed records against
`r round(mean_valid_maize, 0)` g/day among the valid ones, where the original
flagged values averaged `r format(round(mean_outlier_maize, 0), big.mark = ",")`
g/day. The same ordering holds for the other four variables. Imputed values are
less dispersed than valid ones, which is the expected behaviour of a
neighbour average.
:::

### Non-Maize Protein Outlier Treatment

Digestible protein from non-maize sources carries a heavy right tail in the
dietary intake distribution. A small fraction of children report non-maize
protein intake above its upper Tukey fence, and those values propagate into the
protein totals.

Protein is screened on its own axis, once the block treatment above has settled
the rest of the nutrient profile. The criterion targets the tail of a single
component rather than a systematically unreliable record, so the fence
multiplier is 1.5 rather than the 3 used for the block screen. Records above
the fence are replaced via k-NN imputation with the same predictor set and the
same weighting; only the non-maize protein variable is invalidated and imputed,
and every other nutrient variable is carried through unchanged.

```{r}
#| label: tbl-c5-prot-nomaiz-outlier-detection
#| tbl-cap: "Non-maize protein outlier detection (Tukey fence)"

tbl_c5_prot_nomaiz_outlier_detection |>
  gt() |>
  tab_header(
    title = "Non-Maize Protein Outlier Detection",
    subtitle = "Tukey criterion (Q3 + 1.5 × IQR) on non-maize digestible protein"
  ) |>
  cols_label(Metric = "Metric", Value = "Value") |>
  tab_source_note(
    "Upper Tukey fence on the non-maize digestible protein distribution of the analytical cohort."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-c5-prot-imputation-diagnostics
#| tbl-cap: "Non-maize protein imputation: valid vs original vs imputed"

tbl_c5_prot_imputation_diagnostics |>
  gt() |>
  tab_header(
    title = "Non-Maize Protein: Imputation Diagnostics",
    subtitle = glue("{n_prot_flagged} records imputed ({pct_prot_flagged}%) | Fence: {prot_nomaiz_fence_val} g/day")
  ) |>
  fmt_number(columns = c(Mean, Median, Max), decimals = 2) |>
  fmt_number(columns = N, decimals = 0, use_seps = TRUE) |>
  tab_source_note("k-NN RANN (k = 10, inverse-distance weighted). Same predictors as the block imputation.") |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Baseline Bioavailability Calculations

Total nutrient intake is calculated by aggregating maize and non-maize
sources. The Miller equation estimates Total Absorbed Zinc (TAZ) as a
function of age and total dietary zinc, capturing saturation kinetics.
Iron bioavailability applies a uniform 5% absorption factor.


## Baseline Nutrient Inadequacy Classification

Each child's bioavailable nutrient intake is compared against age-specific
H-AR thresholds. Iron uses a uniform threshold (0.5 mg/day absorbed), zinc
uses three EFSA age bands [@EFSA2014_Zinc], and protein uses two WHO/FAO age bands [@WHO2007_ProteinAminoAcid]. A child
is classified as adequate or inadequate for each nutrient independently.


## Baseline Diagnostic Results

### Bioavailability by Age Group

```{r}
#| label: tbl-f1-bioavailability-by-age
#| tbl-cap: "Micronutrient bioavailability by age group (baseline)"

tbl_f1_bioavailability_by_age |>
  gt() |>
  tab_header(
    title = "Micronutrient Intake and Bioavailability by Age Group — Baseline",
    subtitle = "ENCOVI 2023, children 6-59 months (survey-weighted)"
  ) |>
  cols_label(
    age_group_5 = "Age Group",
    n = "N",
    fe_intake = "Fe Intake",
    fe_intake_low = "95% CI",
    fe_intake_upp = "",
    fe_absorbed = "Fe Absorbed",
    fe_absorbed_low = "95% CI",
    fe_absorbed_upp = "",
    zn_intake = "Zn Intake",
    zn_intake_low = "95% CI",
    zn_intake_upp = "",
    taz = "TAZ",
    taz_low = "95% CI",
    taz_upp = "",
    zn_abs_pct = "Zn Abs %",
    zn_abs_pct_low = "95% CI",
    zn_abs_pct_upp = "",
    prot_intake = "Protein",
    prot_intake_low = "95% CI",
    prot_intake_upp = ""
  ) |>
  cols_merge(columns = c(fe_intake_low, fe_intake_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(fe_absorbed_low, fe_absorbed_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(zn_intake_low, zn_intake_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(taz_low, taz_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(zn_abs_pct_low, zn_abs_pct_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(prot_intake_low, prot_intake_upp), pattern = "[{1}, {2}]") |>
  fmt_number(
    columns = c(fe_intake, fe_absorbed, zn_intake, taz, prot_intake,
                fe_intake_low, fe_absorbed_low, zn_intake_low, taz_low,
                prot_intake_low),
    decimals = 2
  ) |>
  fmt_number(columns = c(zn_abs_pct, zn_abs_pct_low), decimals = 1) |>
  tab_spanner(label = "Iron (mg/d)",
              columns = c(fe_intake, fe_intake_low, fe_absorbed, fe_absorbed_low)) |>
  tab_spanner(label = "Zinc (mg/d)",
              columns = c(zn_intake, zn_intake_low, taz, taz_low,
                          zn_abs_pct, zn_abs_pct_low)) |>
  tab_spanner(label = "Protein (g/d)",
              columns = c(prot_intake, prot_intake_low)) |>
  tab_source_note("N: unweighted sample size. All estimates survey-weighted.") |>
  tab_source_note("Iron: 5% uniform absorption (Allen et al., 2006). Zinc: Miller equation (age-dependent).") |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

### Inadequacy Prevalence by Age Group and Sex

```{r}
#| label: tbl-f2-inadequacy-prevalence
#| tbl-cap: "Nutrient inadequacy prevalence by age group and sex (baseline)"

tbl_f2_inadequacy_prevalence |>
  gt(groupname_col = "age_group_5") |>
  tab_header(
    title = "Nutrient Inadequacy Prevalence — Baseline",
    subtitle = "ENCOVI 2023, children 6-59 months by age group and sex (survey-weighted)"
  ) |>
  cols_label(
    sex_label = "Sex",
    n = "N",
    pct_fe_inadeq = "Iron",
    pct_fe_inadeq_low = "95% CI",
    pct_fe_inadeq_upp = "",
    pct_zn_inadeq = "Zinc",
    pct_zn_inadeq_low = "95% CI",
    pct_zn_inadeq_upp = "",
    pct_prot_inadeq = "Protein",
    pct_prot_inadeq_low = "95% CI",
    pct_prot_inadeq_upp = "",
    pct_any_inadeq = "Any",
    pct_any_inadeq_low = "95% CI",
    pct_any_inadeq_upp = "",
    pct_all_inadeq = "All 3",
    pct_all_inadeq_low = "95% CI",
    pct_all_inadeq_upp = ""
  ) |>
  cols_merge(columns = c(pct_fe_inadeq_low, pct_fe_inadeq_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(pct_zn_inadeq_low, pct_zn_inadeq_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(pct_prot_inadeq_low, pct_prot_inadeq_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(pct_any_inadeq_low, pct_any_inadeq_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(pct_all_inadeq_low, pct_all_inadeq_upp), pattern = "[{1}, {2}]") |>
  fmt_number(
    columns = c(pct_fe_inadeq, pct_zn_inadeq, pct_prot_inadeq,
                pct_any_inadeq, pct_all_inadeq,
                pct_fe_inadeq_low, pct_zn_inadeq_low, pct_prot_inadeq_low,
                pct_any_inadeq_low, pct_all_inadeq_low),
    decimals = 1
  ) |>
  tab_spanner(label = "Inadequacy Prevalence (%)",
              columns = c(pct_fe_inadeq, pct_fe_inadeq_low,
                          pct_zn_inadeq, pct_zn_inadeq_low,
                          pct_prot_inadeq, pct_prot_inadeq_low)) |>
  tab_spanner(label = "Combined (%)",
              columns = c(pct_any_inadeq, pct_any_inadeq_low,
                          pct_all_inadeq, pct_all_inadeq_low)) |>
  tab_source_note("N: unweighted sample size. All estimates survey-weighted.") |>
  tab_source_note("Inadequacy: intake below age-specific threshold.") |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

::: {.callout-important}
## Baseline Inadequacy
Nationally the three nutrients sit close together: iron at
`r round(fe_inadeq_baseline, 1)`%, zinc at `r round(zn_inadeq_baseline, 1)`%
and protein at `r round(prot_inadeq_baseline, 1)`%
(see @tbl-h1-baseline-vs-qpm-inadequacy), with
`r round(any_inadeq_baseline, 1)`% of children inadequate in at least one of
them. The ordering is not stable across strata: in
@tbl-f2-inadequacy-prevalence iron is the most prevalent in
`r n_strata_iron_leads` of the `r n_strata_total` age-sex strata, and the
remaining strata are led by zinc or protein.
:::

### Adequacy Ratio Distributions

```{r}
#| label: fig-f3a-iron-adequacy-ratio
#| fig-cap: "Iron adequacy ratio distribution by age group and sex (baseline)"
#| fig-alt: >
#|   Faceted density plot with one panel per age group in a single row, showing
#|   the iron adequacy ratio (absorbed iron divided by the age-specific
#|   requirement) on the x-axis from 0 to 5 and density on the y-axis. Within
#|   each panel two distributions are overlaid for male and female children. A
#|   red dashed reference line marks the adequacy threshold of 1. Mass
#|   concentrates below the threshold in the 6-11 and 12-23 month groups and
#|   shifts toward the threshold from 24 months onward, where the mode sits
#|   close to 1. The youngest children are furthest from adequacy.
#| fig-width: 10
#| fig-height: 5

fig_f3a_iron_adequacy_ratio |>
  ggplot(aes(x = ratio_fe_baseline, fill = sex_label)) +
  geom_density(alpha = 0.5, color = NA) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "#E74C3C",
             linewidth = 0.7) +
  facet_wrap(~ age_group_5, nrow = 1) +
  scale_x_continuous(limits = c(0, 5), breaks = 0:5) +
  scale_fill_manual(
    values = c("Male" = "#1B6CA8", "Female" = "#3AAF85"),
    name = "Sex"
  ) +
  labs(
    title = "Iron (absorbed) Adequacy Ratio by Age Group and Sex — Baseline",
    subtitle = "ENCOVI 2023 — red dashed line = adequacy threshold (ratio = 1)",
    x = "Adequacy Ratio (Fe absorbed / H-AR)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 13, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    strip.text = element_text(size = 9, face = "bold"),
    legend.position = "bottom"
  )
```

```{r}
#| label: fig-f3b-zinc-adequacy-ratio
#| fig-cap: "Zinc adequacy ratio distribution by age group and sex (baseline)"
#| fig-alt: >
#|   Faceted density plot with one panel per age group in a single row, showing
#|   the zinc adequacy ratio (total absorbed zinc divided by the age-specific
#|   requirement) on the x-axis from 0 to 2 and density on the y-axis. Within
#|   each panel two distributions are overlaid for male and female children. A
#|   red dashed reference line marks the adequacy threshold of 1. Zinc
#|   inadequacy sits at the same level as iron and affects a substantial share
#|   of children, most of all in the youngest group, whose distribution
#|   straddles the threshold on both sides.
#| fig-width: 10
#| fig-height: 5

fig_f3b_zinc_adequacy_ratio |>
  ggplot(aes(x = ratio_zn_baseline, fill = sex_label)) +
  geom_density(alpha = 0.5, color = NA) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "#E74C3C",
             linewidth = 0.7) +
  facet_wrap(~ age_group_5, nrow = 1) +
  scale_x_continuous(limits = c(0, 2), breaks = 0:2) +
  scale_fill_manual(
    values = c("Male" = "#1B6CA8", "Female" = "#3AAF85"),
    name = "Sex"
  ) +
  labs(
    title = "Zinc (TAZ) Adequacy Ratio by Age Group and Sex — Baseline",
    subtitle = "ENCOVI 2023 — red dashed line = adequacy threshold (ratio = 1)",
    x = "Adequacy Ratio (TAZ / H-AR)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 13, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    strip.text = element_text(size = 9, face = "bold"),
    legend.position = "bottom"
  )
```

```{r}
#| label: fig-f3c-protein-adequacy-ratio
#| fig-cap: "Protein adequacy ratio distribution by age group and sex (baseline)"
#| fig-alt: >
#|   Faceted density plot with one panel per age group in a single row, showing
#|   the protein adequacy ratio (PDCAAS-adjusted protein divided by the
#|   age-specific requirement) on the x-axis from 0 to 6 and density on the
#|   y-axis. Within each panel two distributions are overlaid for male and
#|   female children. A red dashed reference line marks the adequacy threshold
#|   of 1. The distribution centres on the threshold: the mode sits at or below
#|   1 in all five age groups, and at 6-11 months most of the mass falls below
#|   1. The right tail extends to ratios of 3 to 4.
#| fig-width: 10
#| fig-height: 5

fig_f3c_protein_adequacy_ratio |>
  ggplot(aes(x = ratio_prot_baseline, fill = sex_label)) +
  geom_density(alpha = 0.5, color = NA) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "#E74C3C",
             linewidth = 0.7) +
  facet_wrap(~ age_group_5, nrow = 1) +
  scale_x_continuous(limits = c(0, 6), breaks = 0:6) +
  scale_fill_manual(
    values = c("Male" = "#1B6CA8", "Female" = "#3AAF85"),
    name = "Sex"
  ) +
  labs(
    title = "Protein (PDCAAS) Adequacy Ratio by Age Group and Sex — Baseline",
    subtitle = "ENCOVI 2023 — red dashed line = adequacy threshold (ratio = 1)",
    x = "Adequacy Ratio (Protein / H-AR)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 13, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    strip.text = element_text(size = 9, face = "bold"),
    legend.position = "bottom"
  )
```

### Stunting and Inadequacy Relationship

```{r}
#| label: tbl-f4-stunting-inadequacy
#| tbl-cap: "Stunting prevalence by nutrient inadequacy status (baseline)"

tbl_f4_stunting_inadequacy |>
  gt() |>
  tab_header(
    title = "Stunting Prevalence by Number of Nutrient Inadequacies — Baseline",
    subtitle = "ENCOVI 2023, children 6-59 months (survey-weighted)"
  ) |>
  cols_label(
    inadequacy_group = "Inadequacy Status",
    n = "N",
    maize_g = "Maize (g/day)",
    pct_rural = "% Rural",
    stunting_prev = "Stunting %",
    stunting_prev_low = "95% CI",
    stunting_prev_upp = "",
    mean_haz = "Mean HAZ",
    mean_haz_low = "95% CI",
    mean_haz_upp = ""
  ) |>
  cols_merge(columns = c(stunting_prev_low, stunting_prev_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(mean_haz_low, mean_haz_upp), pattern = "[{1}, {2}]") |>
  fmt_number(columns = c(maize_g), decimals = 0) |>
  fmt_number(columns = c(pct_rural), decimals = 1) |>
  fmt_number(columns = c(stunting_prev, stunting_prev_low), decimals = 1) |>
  fmt_number(columns = c(mean_haz, mean_haz_low), decimals = 2) |>
  tab_source_note("Stunting: HAZ < -2 (WHO classification). Inadequacy: intake below H-AR. Maize intake and rural share shown to characterise the socioeconomic profile of each group.") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

::: {.callout-note}
## Non-Monotonic Pattern
Stunting prevalence does not increase monotonically with the number of current
nutrient inadequacies (see @tbl-f4-stunting-inadequacy): the highest prevalence,
`r round(f4_highest$stunting_prev, 1)`%, falls on the
“`r as.character(f4_highest$inadequacy_group)`” group and the lowest,
`r round(f4_lowest$stunting_prev, 1)`%, on the
“`r as.character(f4_lowest$inadequacy_group)`” group.

The children with no current inadequacy are the heaviest maize consumers, at a
mean of `r round(f4_no_inad$maize_g, 0)` g/day against
`r round(f4_others_max_maize, 0)` g/day or less in the other groups. Maize is
the main dietary source of iron, zinc and protein in this population, so a
maize-dominated diet clears the adequacy thresholds, which is what places these
children in the “no inadequacy” group. The same group is the most rural, at
`r round(f4_no_inad$pct_rural, 1)`%, and carries the lowest mean
height-for-age, at `r round(f4_no_inad$mean_haz, 2)`. Chronic undernutrition is
associated with conditions that current nutrient adequacy does not measure,
among them sanitation, infection burden and dietary diversity. Maize dependence
sits on both sides at once, as a source of nutrient adequacy and as a marker of
that profile, which is what produces the ordering in the table.
:::


## QPM Intervention (100% Coverage)

Biofortification factors are applied to maize-source nutrients
(Fe ×1.19, Zn ×1.38, Protein ×1.94), and the entire bioavailability and
inadequacy pipeline is recalculated under the intervention scenario. This
represents the maximum individual-level benefit assuming all maize is
replaced by QPM varieties.

The protein factor of 1.94 carries **protein quality**, not protein quantity.
Conventional plant breeding leaves total protein content per gram of maize
unchanged and increases the proportion of lysine and tryptophan within that
protein, which raises the protein digestibility-corrected amino acid score
(PDCAAS) of QPM-source protein from 0.317 to 0.616. This factor is
applied to the PDCAAS-adjusted protein intake variable already computed at the
food level in the
[intake calculation](../01_nutritional_intakes/01_03_intake_calculation.qmd),
so the protein values entering the inadequacy pipeline carry the quality
change. Lysine and tryptophan enter this pathway through that same index and
carry no separate factor of their own, which keeps the amino acid profile
counted once.

### Baseline vs QPM Inadequacy

```{r}
#| label: tbl-h1-baseline-vs-qpm-inadequacy
#| tbl-cap: "Nutrient inadequacy prevalence: baseline vs QPM intervention (maximum impact)"

tbl_h1_baseline_vs_qpm_inadequacy |>
  gt() |>
  tab_header(
    title = "Nutrient Inadequacy: Baseline vs. QPM Intervention (100% Coverage)",
    subtitle = "ENCOVI 2023, children 6-59 months — Maximum potential impact (all maize biofortified)"
  ) |>
  fmt_number(columns = c(`Baseline Inad. (%)`, `QPM Inad. (%)`,
                         `Change (pp)`, `Cases Resolved (%)`), decimals = 1) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = `Change (pp)`, rows = `Change (pp)` < 0)
  ) |>
  tab_source_note("QPM factors: Fe ×1.19, Zn ×1.38, Protein ×1.94 (PDCAAS). All estimates survey-weighted.") |>
  tab_source_note("Cases resolved: children moving from inadequate (baseline) to adequate (QPM).") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

::: {.callout-note}
## QPM Impact at Full Coverage
At 100% biofortified maize coverage, protein inadequacy falls by
`r round(prot_inadeq_drop, 1)` percentage points, zinc by
`r round(zn_inadeq_drop, 1)` and iron by `r round(fe_inadeq_drop, 1)`, taking
the share of children inadequate in at least one nutrient from
`r round(any_inadeq_baseline, 1)`% to `r round(any_inadeq_qpm, 1)`% (see
@tbl-h1-baseline-vs-qpm-inadequacy). The protein figure follows the PDCAAS
change in QPM, from 0.317 to 0.616, a factor of 1.94 on the quality-adjusted
protein reaching the child.
:::

### Absorbed Nutrient Increase Distributions

```{r}
#| label: fig-h2-delta-absorbed-nutrients
#| fig-cap: "Distribution of absorbed nutrient increase under QPM intervention by age group"
#| fig-alt: >
#|   Three side-by-side density panels for iron, zinc, and protein, showing the
#|   per-child increase in absorbed nutrient under the QPM intervention on each
#|   x-axis (delta absorbed iron in mg/day, delta total absorbed zinc in mg/day,
#|   and delta protein in g/day) and density on the y-axis. Within each panel the
#|   distribution is split by age group. A dashed reference line marks zero
#|   change. Each x-axis starts at zero and is clipped at an upper quantile, so
#|   the right tail of every panel is not shown.
#| fig-width: 12
#| fig-height: 5

p_delta_fe <- fig_h2_delta_absorbed_nutrients |>
  filter(!is.na(delta_fe_absorbed_mg) & is.finite(delta_fe_absorbed_mg)) |>
  ggplot(aes(x = delta_fe_absorbed_mg, fill = age_group_5)) +
  geom_density(alpha = 0.5, color = NA) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50",
             linewidth = 0.5) +
  coord_cartesian(
    xlim = c(0, quantile(fig_h2_delta_absorbed_nutrients$delta_fe_absorbed_mg,
                         0.99, na.rm = TRUE) * 1.1)
  ) +
  scale_fill_manual(
    values = c("#1B6CA8", "#3AAF85", "#F4A261", "#E76F51", "#264653"),
    name = "Age Group"
  ) +
  labs(
    title = "Iron (absorbed) increase",
    x = "Delta Fe absorbed (mg/day)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 11, face = "bold", color = "azure4"),
    axis.title = element_text(size = 9, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    legend.position = "none"
  )

p_delta_zn <- fig_h2_delta_absorbed_nutrients |>
  filter(!is.na(delta_taz_mg) & is.finite(delta_taz_mg)) |>
  ggplot(aes(x = delta_taz_mg, fill = age_group_5)) +
  geom_density(alpha = 0.5, color = NA) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50",
             linewidth = 0.5) +
  coord_cartesian(
    xlim = c(0, quantile(fig_h2_delta_absorbed_nutrients$delta_taz_mg,
                         0.99, na.rm = TRUE) * 1.1)
  ) +
  scale_fill_manual(
    values = c("#1B6CA8", "#3AAF85", "#F4A261", "#E76F51", "#264653"),
    name = "Age Group"
  ) +
  labs(
    title = "Zinc (TAZ) increase",
    x = "Delta TAZ (mg/day)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 11, face = "bold", color = "azure4"),
    axis.title = element_text(size = 9, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    legend.position = "none"
  )

p_delta_prot <- fig_h2_delta_absorbed_nutrients |>
  filter(!is.na(delta_prot_g_total) & is.finite(delta_prot_g_total)) |>
  ggplot(aes(x = delta_prot_g_total, fill = age_group_5)) +
  geom_density(alpha = 0.5, color = NA) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50",
             linewidth = 0.5) +
  coord_cartesian(
    xlim = c(0, quantile(fig_h2_delta_absorbed_nutrients$delta_prot_g_total,
                         0.99, na.rm = TRUE) * 1.1)
  ) +
  scale_fill_manual(
    values = c("#1B6CA8", "#3AAF85", "#F4A261", "#E76F51", "#264653"),
    name = "Age Group"
  ) +
  labs(
    title = "Protein (PDCAAS) increase",
    x = "Delta Protein (g/day)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 11, face = "bold", color = "azure4"),
    axis.title = element_text(size = 9, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    legend.position = "bottom"
  )

p_delta_fe + p_delta_zn + p_delta_prot +
  plot_layout(ncol = 3, guides = "collect") &
  theme(legend.position = "bottom")
```

### Bioavailability by Age Group (QPM)

```{r}
#| label: tbl-h3-bioavailability-by-age-qpm
#| tbl-cap: "Micronutrient bioavailability by age group (QPM intervention)"

tbl_h3_bioavailability_by_age_qpm |>
  gt() |>
  tab_header(
    title = "Micronutrient Intake and Bioavailability by Age Group — QPM Intervention",
    subtitle = "ENCOVI 2023, children 6-59 months (survey-weighted, 100% coverage)"
  ) |>
  cols_label(
    age_group_5 = "Age Group",
    n = "N",
    fe_intake = "Fe Intake",
    fe_intake_low = "95% CI",
    fe_intake_upp = "",
    fe_absorbed = "Fe Absorbed",
    fe_absorbed_low = "95% CI",
    fe_absorbed_upp = "",
    zn_intake = "Zn Intake",
    zn_intake_low = "95% CI",
    zn_intake_upp = "",
    taz = "TAZ",
    taz_low = "95% CI",
    taz_upp = "",
    zn_abs_pct = "Zn Abs %",
    zn_abs_pct_low = "95% CI",
    zn_abs_pct_upp = "",
    prot_intake = "Protein",
    prot_intake_low = "95% CI",
    prot_intake_upp = ""
  ) |>
  cols_merge(columns = c(fe_intake_low, fe_intake_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(fe_absorbed_low, fe_absorbed_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(zn_intake_low, zn_intake_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(taz_low, taz_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(zn_abs_pct_low, zn_abs_pct_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(prot_intake_low, prot_intake_upp), pattern = "[{1}, {2}]") |>
  fmt_number(
    columns = c(fe_intake, fe_absorbed, zn_intake, taz, prot_intake,
                fe_intake_low, fe_absorbed_low, zn_intake_low, taz_low,
                prot_intake_low),
    decimals = 2
  ) |>
  fmt_number(columns = c(zn_abs_pct, zn_abs_pct_low), decimals = 1) |>
  tab_spanner(label = "Iron (mg/d)",
              columns = c(fe_intake, fe_intake_low, fe_absorbed, fe_absorbed_low)) |>
  tab_spanner(label = "Zinc (mg/d)",
              columns = c(zn_intake, zn_intake_low, taz, taz_low,
                          zn_abs_pct, zn_abs_pct_low)) |>
  tab_spanner(label = "Protein (g/d)",
              columns = c(prot_intake, prot_intake_low)) |>
  tab_source_note("N: unweighted sample size. All estimates survey-weighted.") |>
  tab_source_note("QPM factors: Fe ×1.19, Zn ×1.38, Protein ×1.94 (PDCAAS). Iron: 5% absorption. Zinc: Miller equation.") |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```


## Summary

### Key Findings

1. **Chispitas supplementation integration**: Departmental Chispitas
   contributions (Fe and Zn) are added to non-maize intake variables before
   the bioavailability pipeline, so the Meta-analysis (Gunaratna) pathway
   baseline includes current micronutrient supplementation policy in
   Guatemala.

2. **Outlier treatment**: Records above the maize energy ceiling
   (`r format(MAIZE_KCAL_CEILING, big.mark = ",")` kcal/day) or above the
   non-maize Tukey fence are flagged and replaced via k-NN imputation
   conditioned on the child's own ENCOVI record. Non-maize protein is screened
   separately against its own Tukey fence and imputed the same way.

3. **Baseline inadequacy**: Nationally, iron reaches
   `r round(fe_inadeq_baseline, 1)`%, zinc `r round(zn_inadeq_baseline, 1)`%
   and protein `r round(prot_inadeq_baseline, 1)`%. The ordering varies by
   age-sex stratum (see @tbl-f2-inadequacy-prevalence and
   @tbl-h1-baseline-vs-qpm-inadequacy).

4. **QPM intervention**: At 100% coverage, protein inadequacy falls furthest,
   by `r round(prot_inadeq_drop, 1)` percentage points, followed by zinc at
   `r round(zn_inadeq_drop, 1)` and iron at `r round(fe_inadeq_drop, 1)`
   (see @tbl-h1-baseline-vs-qpm-inadequacy). The protein figure follows the
   PDCAAS change (0.317 → 0.616), with total protein quantity unchanged.

5. **Pre-computed intervention profiles**: Both baseline and QPM nutrient
   indicators are computed at the individual level, so downstream scenario
   simulation selects children rather than recalculating them.

::: {.callout-warning}
## Limitations

1. **Iron absorption uniformity** — A single 5% factor is applied to all
   children, without modeling individual variation from iron status or
   dietary enhancers/inhibitors.

2. **Cross-sectional design** — Nutrient inadequacy measures current intake
   and stunting measures cumulative growth history, so the two are observed
   over different time horizons.

3. **Full substitution assumption** — The QPM columns are computed with all
   maize replaced by biofortified varieties, and partial coverage is applied
   downstream through binary assignment.

4. **Chispitas adherence estimate** — The national adherence factor of
   `r round(chispitas_adherence, 3)` is derived from SIVESNU 2018 and applied
   uniformly across departments.

5. **Imputed record dispersion** — The k-NN replacements for the
   `r format(n_flagged, big.mark = ",")` flagged records are neighbour
   averages, so they carry less spread than the valid records they sit
   among.

6. **Predictor set of the imputation** — The six predictors span demographics,
   geography, socioeconomic position and diet composition, one variable per
   dimension. Determinants of dietary intake outside those four dimensions are
   not represented in the matching space, so the replacements reproduce the
   central tendency of that space rather than the full heterogeneity of
   intake.
:::

### Application

This enriched dataset — containing baseline and QPM nutrient profiles,
bioavailability indicators, and inadequacy classifications for the analytical
cohort — feeds the
[QPM stunting impact calculation](05_05_calculate_qpm_stunting_impact.qmd) and
the scenario simulation pipeline in Module 6.
