---
title: "Height Transfer Model"
subtitle: "Module 5: Meta-analysis"
---

## Why a Height Transfer Model?

The Meta-analysis (Gunaratna) pathway requires predicted height for ENCOVI
2023 children to calculate individual-level impacts of biofortified maize on
stunting. Since ENCOVI does not include anthropometric measurements, we train
a predictive model on SIVESNU 2018 — which contains direct height data — and
apply it across surveys using the shared set of socioeconomic predictors
identified when
[preparing the SIVESNU height data](05_01_prepare_sivesnu_height_data.qmd).

We use a Generalized Additive Model (GAM) because the age-height relationship
in children under 5 is inherently non-linear: growth velocity is highest in
infancy and decelerates progressively. GAMs capture this curvilinear growth
trajectory through smooth terms while retaining parametric estimation for
categorical predictors. The Gaussian family is appropriate because height
conditional on age and covariates is approximately normally distributed.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  mgcv,               # Required for gam() method dispatch (geom_smooth, gratia::draw)
  gratia,             # GAM partial effects visualization (fig-e2)
  ggpubr,             # ggarrange for 2x2 diagnostic grid (fig-e1)
  patchwork,          # Combining plots (fig-a3, fig-c3)
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))
```

```{r}
#| label: load-rds
#| include: false

# Web resources for visualizations (exported by framework 05_02)
tbl_a1_data_overview <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_a1_data_overview.rds"),
  trust = TRUE
)

tbl_a2_predictor_types <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_a2_predictor_types.rds"),
  trust = TRUE
)

fig_a3_outcome_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_fig_a3_outcome_distribution.rds"),
  trust = TRUE
)

tbl_b1_modeling_data_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_b1_modeling_data_summary.rds"),
  trust = TRUE
)

tbl_b3_smooth_terms <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_b3_smooth_terms.rds"),
  trust = TRUE
)

tbl_b3_parametric_terms <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_b3_parametric_terms.rds"),
  trust = TRUE
)

tbl_b3_overall_fit <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_b3_overall_fit.rds"),
  trust = TRUE
)

tbl_c2_cv_overall <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_c2_cv_overall.rds"),
  trust = TRUE
)

tbl_c2_cv_by_fold <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_c2_cv_by_fold.rds"),
  trust = TRUE
)

fig_c3_cv_observed_vs_predicted <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_fig_c3_cv_observed_vs_predicted.rds"),
  trust = TRUE
)

tbl_d1_cv_by_age <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_d1_cv_by_age.rds"),
  trust = TRUE
)

fig_d1_residuals_by_age <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_fig_d1_residuals_by_age.rds"),
  trust = TRUE
)

tbl_d2_cv_by_department <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_d2_cv_by_department.rds"),
  trust = TRUE
)

tbl_d3_stunting_classification <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_d3_stunting_classification.rds"),
  trust = TRUE
)

fig_e1_residual_diagnostics <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_fig_e1_residual_diagnostics.rds"),
  trust = TRUE
)

tbl_e3_concurvity <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_tbl_e3_concurvity.rds"),
  trust = TRUE
)

fig_e4_observed_vs_fitted <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_02_fig_e4_observed_vs_fitted.rds"),
  trust = TRUE
)

# GAM model object (required by fig-e2 for gratia::draw() and by fig-e4
# subtitle for R² / deviance explained)
gam_objects <- import(
  here("..", "01_data", "02_processed", "models",
       "05_02_height_transfer_model.rds"),
  trust = TRUE
)
gam_height <- gam_objects$gam_height
```

```{r}
#| label: inline-values
#| include: false

# CV performance metrics for fig-c3 subtitle (from tbl-c2)
cv_rmse <- tbl_c2_cv_overall |> pull(RMSE)
cv_r2   <- tbl_c2_cv_overall |> pull(R2)

# Full-model fit metrics for fig-e4 subtitle (from gam_height summary)
gam_summary    <- summary(gam_height)
full_r2        <- gam_summary$r.sq
full_dev_expl  <- gam_summary$dev.expl

# Counts used in narrative (Boruta-selected predictors)
n_predictors <- tbl_a1_data_overview |>
  filter(Item == "Boruta-selected predictors") |>
  pull(Value) |>
  as.integer()

n_observations <- tbl_a1_data_overview |>
  filter(Item == "Observations") |>
  pull(Value) |>
  parse_number()

n_expanded <- tbl_b1_modeling_data_summary |>
  filter(Metric == "Expanded observations (post-uncount)") |>
  pull(Value)

# Stunting prevalence from tbl-a1 (already formatted as e.g. "42.2%")
stunting_prevalence <- tbl_a1_data_overview |>
  filter(Item == "Stunting prevalence (HAZ < -2)") |>
  pull(Value)

# Height range from tbl-a1 (already formatted as e.g. "60.5 - 112.1")
height_range <- tbl_a1_data_overview |>
  filter(Item == "Height range (cm)") |>
  pull(Value)

# Departments the table flags: RMSE above one SD from the departmental mean
dept_flagged <- tbl_d2_cv_by_department |>
  filter(RMSE > mean(RMSE) + sd(RMSE)) |>
  arrange(desc(RMSE))

n_departments  <- nrow(tbl_d2_cv_by_department)
n_dept_flagged <- nrow(dept_flagged)
dept_flagged_list <- paste(dept_flagged$departamento, collapse = ", ")

# Residual asymmetry between stunted and non-stunted children (observed minus
# predicted, so a negative value is an overprediction of height)
resid_stunted <- tbl_d3_stunting_classification |>
  filter(Metric == "Mean residual - stunted children (cm)") |>
  pull(Value)

resid_not_stunted <- tbl_d3_stunting_classification |>
  filter(Metric == "Mean residual - non-stunted children (cm)") |>
  pull(Value)

mae_stunted <- tbl_d3_stunting_classification |>
  filter(Metric == "MAE - stunted children (cm)") |>
  pull(Value)

mae_not_stunted <- tbl_d3_stunting_classification |>
  filter(Metric == "MAE - non-stunted children (cm)") |>
  pull(Value)

# Highest pairwise concurvity among the smooth terms
max_concurvity <- max(tbl_e3_concurvity$Concurvity)
```


## Input Data

The modeling-ready dataset from the
[previous step](05_01_prepare_sivesnu_height_data.qmd) contains
`r format(n_observations, big.mark = ",")` SIVESNU 2018 children aged 6-59
months with `r n_predictors` Boruta-selected predictors and the outcome
variable (`talla_cm`).

```{r}
#| label: tbl-a1-data-overview
#| tbl-cap: "Input data overview"

tbl_a1_data_overview |>
  mutate(
    Value = str_replace_all(
      Value,
      "[0-9]+\\.[0-9]+",
      \(x) formatC(as.numeric(x), digits = 1, format = "f")
    )
  ) |>
  gt() |>
  tab_header(
    title = "Input Data Summary",
    subtitle = "SIVESNU 2018 modeling-ready dataset"
  ) |>
  cols_align(align = "left", columns = Item) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### GAM Specification

Each predictor receives a model term appropriate to its data type. Continuous
predictors with expected non-linear effects use smooth terms (`s()`), while
binary and unordered categorical predictors enter as parametric (linear)
terms. Department enters through a penalized random-effect term, which lets
the estimation shrink the geographic contribution according to the support the
data provide for it.

```{r}
#| label: tbl-a2-predictor-types
#| tbl-cap: "Predictor variable types and roles in the GAM"

tbl_a2_predictor_types |>
  gt() |>
  tab_header(
    title = "Predictor Variables and GAM Specification",
    subtitle = paste0(n_predictors,
                      " Boruta-selected predictors with assigned model terms")
  ) |>
  cols_label(
    Variable = "Variable",
    Class = "R Class",
    Unique = "Unique Values",
    GAM_Term = "GAM Term"
  ) |>
  cols_align(align = "left") |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = str_starts(GAM_Term, "s\\("))
  ) |>
  tab_footnote(
    footnote = "Blue rows: smooth (non-linear) terms. White rows: parametric (linear) terms.",
    locations = cells_column_labels(columns = GAM_Term)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Outcome Distribution

```{r}
#| label: fig-a3-outcome-distribution
#| fig-cap: "Distribution of height (talla_cm) in the training population"
#| fig-alt: >
#|   Two panels side by side. Left panel is a histogram of child height in
#|   centimetres, showing a broad, roughly bell-shaped spread with a slight
#|   left skew that reflects Guatemala's high stunting prevalence. Right panel
#|   is a scatter plot of height on the y-axis against age in months on the
#|   x-axis, overlaid with a dashed GAM smooth curve that rises steeply up to
#|   about 24 months and then flattens, confirming the non-linear,
#|   decelerating growth trajectory the smooth terms are designed to capture.
#| fig-width: 10
#| fig-height: 6

# Height distribution
p_hist <- fig_a3_outcome_distribution |>
  ggplot(aes(x = talla_cm)) +
  geom_histogram(
    bins = 40,
    fill = "lightgray",
    color = "azure4",
    alpha = 0.5
  ) +
  labs(
    title = "Height Distribution",
    subtitle = paste0("SIVESNU 2018 children aged 6-59 months (N = ",
                      format(n_observations, big.mark = ","), ")"),
    x = "Height (cm)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Height vs Age non-linear relation
p_age <- fig_a3_outcome_distribution |>
  ggplot(aes(x = edad_mes, y = talla_cm)) +
  geom_point(alpha = 0.3, color = "#1B6CA8", size = 1) +
  geom_smooth(
    method = "gam",
    formula = y ~ s(x),
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Height vs. Age",
    subtitle = "Non-linear growth curve captured by GAM smooth",
    x = "Age (months)",
    y = "Height (cm)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

p_hist + p_age
```

The height distribution spans `r height_range` cm with a slight left skew,
consistent with a stunting prevalence of `r stunting_prevalence` in this
population. The age-height scatterplot shows the non-linear growth trajectory,
with rapid gains in the first two years easing into a more gradual increase
after 24 months, which is the pattern the GAM smooth terms are specified to
capture.


## GAM Model Training

The training pipeline first expands the SIVESNU dataset by survey weights
(`pesonino`) to incorporate the sampling design, then collapses rare factor
levels (n < 10) into existing categories to prevent empty-level issues in
cross-validation folds. Ordered factors are converted to numeric for smooth
term estimation, and an age group variable is added for stratified
cross-validation.

```{r}
#| label: tbl-b1-modeling-data-summary
#| tbl-cap: "Modeling data dimensions after preparation"

tbl_b1_modeling_data_summary |>
  gt() |>
  tab_header(
    title = "Modeling Data Summary",
    subtitle = "After type preparation for mgcv"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

The model is then trained with **REML** (Restricted Maximum Likelihood) smooth
parameter estimation.

### Smooth Term Significance

The approximate F-tests are significant for all three smooth terms. The
effective degrees of freedom (EDF) reported alongside them say how much
curvature each one takes up.

```{r}
#| label: tbl-b3-smooth-terms
#| tbl-cap: "GAM smooth term significance"

tbl_b3_smooth_terms |>
  mutate(
    significant = if_else(p_value < 0.05, "*", ""),
    p_value_fmt = case_when(
      p_value < 0.001 ~ "< 0.001",
      TRUE ~ as.character(round(p_value, 4))
    )
  ) |>
  select(Term, edf, ref_df, f_statistic, p_value_fmt, significant) |>
  gt() |>
  tab_header(
    title = "Smooth Term Significance",
    subtitle = "Approximate F-tests for non-linear effects"
  ) |>
  cols_label(
    Term = "Smooth Term",
    edf = "EDF",
    ref_df = "Ref. df",
    f_statistic = "F",
    p_value_fmt = "p-value",
    significant = ""
  ) |>
  fmt_number(columns = c(edf, ref_df, f_statistic), decimals = 2) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = significant == "*")
  ) |>
  tab_footnote(
    footnote = "EDF = effective degrees of freedom. EDF close to 1 indicates near-linear relationship.",
    locations = cells_column_labels(columns = edf)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {#imp-grado-estudios-definition .callout-note}
## Smooth Term Interpretation
The age smooth takes 4.24 effective degrees of freedom, which is the
non-linear growth curve. The `grado_estudios_hogar` smooth takes 2.01,
a mild curvature in the education-height relationship. This variable encodes
the household's parental educational attainment as the mean of the head of
household and their spouse or partner on a six-point ordinal scale mapped to
numeric values (1 = Ninguno/Desconocido, 2 = Solo sabe leer, 4 = Primaria,
7 = Secundaria, 10 = Superior/universitaria). The mean over the conjugal unit
is the summary used, and other household members do not enter it. The
department term takes 0.98 of its single available degree of freedom, so the
penalty leaves it close to unshrunk.
:::

### Parametric Coefficients

```{r}
#| label: tbl-b3-parametric-terms
#| tbl-cap: "GAM parametric term coefficients"

tbl_b3_parametric_terms |>
  mutate(
    significant = if_else(p_value < 0.05, "*", ""),
    p_value_fmt = case_when(
      p_value < 0.001 ~ "< 0.001",
      TRUE ~ as.character(round(p_value, 4))
    )
  ) |>
  select(Term, estimate, se, t_value, p_value_fmt, significant) |>
  gt() |>
  tab_header(
    title = "Parametric Term Coefficients",
    subtitle = "Linear effects of categorical predictors"
  ) |>
  cols_label(
    Term = "Term",
    estimate = "Estimate",
    se = "SE",
    t_value = "t",
    p_value_fmt = "p-value",
    significant = ""
  ) |>
  fmt_number(columns = c(estimate, se, t_value), decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = significant == "*")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Overall Model Fit

```{r}
#| label: tbl-b3-overall-fit
#| tbl-cap: "Overall model fit statistics"

tbl_b3_overall_fit |>
  gt() |>
  tab_header(
    title = "Overall Model Fit",
    subtitle = "GAM with Gaussian family and REML estimation"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric %in% c("Deviance explained",
                                                  "Adjusted R-squared"))
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Model Fit
The GAM accounts for `r round(full_dev_expl * 100, 1)`% of deviance
(Adj. R² = `r round(full_r2, 3)`) on the training data. Age carries most of
this, as expected for child height. Beyond it, the parametric block contributes
significant coefficients for child sex, water source, television ownership,
mobile phone access, area of residence, floor and wall materials, and refuse
collection.
:::


## Cross-Validation

### 10-Fold Stratified CV

We evaluate out-of-sample prediction performance using stratified 10-fold
cross-validation. Stratification by age group gives each fold a representative
spread along the growth curve.

```{r}
#| label: tbl-c2-cv-overall
#| tbl-cap: "10-fold cross-validation performance"

tbl_c2_cv_overall |>
  pivot_longer(everything(), names_to = "Metric", values_to = "Value") |>
  mutate(
    Metric = case_when(
      Metric == "RMSE" ~ "Root Mean Squared Error (cm)",
      Metric == "MAE" ~ "Mean Absolute Error (cm)",
      Metric == "R2" ~ "R-squared (out-of-sample)",
      Metric == "Mean_Bias" ~ "Mean Bias (cm)",
      Metric == "SD_Residual" ~ "SD of Residuals (cm)",
      Metric == "Pct_within_5cm" ~ "Predictions within +/- 5 cm (%)",
      Metric == "Pct_within_3cm" ~ "Predictions within +/- 3 cm (%)"
    ),
    Value = round(Value, 3)
  ) |>
  gt() |>
  tab_header(
    title = "Cross-Validation Performance Summary",
    subtitle = "10-fold stratified CV (stratified by age group)"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = str_detect(Metric, "R-squared|RMSE"))
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-c2-cv-by-fold
#| tbl-cap: "Performance by CV fold"

tbl_c2_cv_by_fold |>
  gt() |>
  tab_header(
    title = "Performance by Cross-Validation Fold",
    subtitle = "Stability assessment across folds"
  ) |>
  cols_label(
    fold_id = "Fold",
    N = "N",
    RMSE = "RMSE (cm)",
    MAE = "MAE (cm)",
    R2 = "R²",
    Mean_Bias = "Bias (cm)"
  ) |>
  fmt_number(columns = c(RMSE, MAE, Mean_Bias), decimals = 2) |>
  fmt_number(columns = R2, decimals = 3) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Cross-Validation Results
Out-of-sample performance stays close to the in-sample fit:
RMSE = `r round(cv_rmse, 2)` cm, R² = `r round(cv_r2, 3)`, with a mean bias of
`r round(tbl_c2_cv_overall$Mean_Bias, 2)` cm. Across the ten folds RMSE ranges
from `r round(min(tbl_c2_cv_by_fold$RMSE), 2)` to
`r round(max(tbl_c2_cv_by_fold$RMSE), 2)` cm and R² from
`r round(min(tbl_c2_cv_by_fold$R2), 3)` to
`r round(max(tbl_c2_cv_by_fold$R2), 3)`, so the fit does not hinge on which
subset is held out. `r round(tbl_c2_cv_overall$Pct_within_5cm, 1)`% of
predictions fall within ±5 cm and
`r round(tbl_c2_cv_overall$Pct_within_3cm, 1)`% within ±3 cm of observed
height.
:::

### CV Diagnostic Plots

```{r}
#| label: fig-c3-cv-observed-vs-predicted
#| fig-cap: "Cross-validation: observed vs predicted height"
#| fig-alt: >
#|   Two panels side by side. Left panel is a scatter plot of predicted height
#|   on the y-axis against observed height on the x-axis, with a dashed
#|   one-to-one reference line; points cluster tightly along this line,
#|   indicating close agreement out of sample. Right panel plots residuals
#|   against predicted height with a dashed line at zero and a loess smooth
#|   that stays close to flat and horizontal, showing no systematic bias
#|   across the range of predicted heights.
#| fig-width: 10
#| fig-height: 6

# Prediction vs observed
p_scatter <- fig_c3_cv_observed_vs_predicted |>
  ggplot(aes(x = talla_cm, y = predicted)) +
  geom_point(alpha = 0.3, color = "#1B6CA8", size = 1.2) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "#3AAF85",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  labs(
    title = "Observed vs Predicted Height (CV)",
    subtitle = paste0("RMSE = ", round(cv_rmse, 2), " cm, R² = ",
                      round(cv_r2, 3)),
    x = "Observed Height (cm)",
    y = "Predicted Height (cm)"
  ) +
  coord_equal() +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Residuals
p_residual <- fig_c3_cv_observed_vs_predicted |>
  ggplot(aes(x = predicted, y = residual)) +
  geom_point(alpha = 0.3, color = "#1B6CA8", size = 1.2) +
  geom_hline(
    yintercept = 0,
    color = "azure4",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  geom_smooth(
    method = "loess",
    se = TRUE,
    color = "#3AAF85",
    linewidth = 0.6,
    alpha = 0.2,
    formula = 'y ~ x'
  ) +
  labs(
    title = "Residuals vs Predicted",
    subtitle = "Loess smooth for systematic bias detection",
    x = "Predicted Height (cm)",
    y = "Residual (cm)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

p_scatter + p_residual
```


## Performance by Subgroup

### By Age Group

```{r}
#| label: tbl-d1-cv-by-age
#| tbl-cap: "Cross-validation performance by age group"

tbl_d1_cv_by_age |>
  gt() |>
  tab_header(
    title = "Prediction Performance by Age Group",
    subtitle = "10-fold CV results stratified by age"
  ) |>
  cols_label(
    age_group = "Age Group",
    N = "N",
    Mean_Height = "Mean Height (cm)",
    RMSE = "RMSE (cm)",
    MAE = "MAE (cm)",
    R2 = "R²",
    Mean_Bias = "Bias (cm)",
    Pct_within_3cm = "Within ±3 cm (%)"
  ) |>
  fmt_number(columns = c(Mean_Height, RMSE, MAE, Mean_Bias), decimals = 2) |>
  fmt_number(columns = R2, decimals = 3) |>
  fmt_number(columns = Pct_within_3cm, decimals = 1) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-d1-residuals-by-age
#| fig-cap: "Residual distribution by age group"
#| fig-alt: >
#|   Box plots of prediction residuals (observed minus predicted height) for
#|   five age groups from 6-11 months to 48-59 months along the x-axis, with a
#|   dashed reference line at zero. Every box is centred close to zero,
#|   indicating little bias, but the boxes and whiskers widen from the
#|   youngest to the oldest group, showing that residual spread and prediction
#|   error grow as children age.
#| fig-width: 10
#| fig-height: 5

fig_d1_residuals_by_age |>
  ggplot(aes(x = age_group, y = residual, fill = age_group)) +
  geom_boxplot(alpha = 0.5, outlier.alpha = 0.3, outlier.size = 1,
               color = "azure4") +
  geom_hline(
    yintercept = 0,
    color = "#3AAF85",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  scale_fill_manual(
    values = c(
      "6-11m"  = "#B3D9E8",
      "12-23m" = "#80BFD4",
      "24-35m" = "#4DA6C0",
      "36-47m" = "#2688A8",
      "48-59m" = "#1B6CA8"
    )
  ) +
  labs(
    title = "Prediction Residuals by Age Group",
    subtitle = "Boxplots of (observed - predicted) height",
    x = "Age Group",
    y = "Residual (cm)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    legend.position = "none"
  )
```

::: {.callout-note}
## Age-Dependent Accuracy
RMSE rises from 3.25 cm in the 6-11 month group to 4.69 cm in the 48-59 month
group, as shown in @tbl-d1-cv-by-age, which follows the wider spread of height
at older ages. Within-group R² values sit between 0.28 and 0.49 because age,
the leading predictor, varies little inside each group, leaving the
socioeconomic variables and child sex to account for the rest.
:::

### By Department

```{r}
#| label: tbl-d2-cv-by-department
#| tbl-cap: "Cross-validation performance by department"

tbl_d2_cv_by_department |>
  gt() |>
  tab_header(
    title = "Prediction Performance by Department",
    subtitle = "10-fold CV results by geographic unit"
  ) |>
  cols_label(
    departamento = "Department",
    N = "N",
    RMSE = "RMSE (cm)",
    MAE = "MAE (cm)",
    Mean_Bias = "Bias (cm)"
  ) |>
  fmt_number(columns = c(RMSE, MAE, Mean_Bias), decimals = 2) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = RMSE, rows = RMSE > mean(RMSE) + sd(RMSE))
  ) |>
  tab_footnote(
    footnote = "Red cells: departments with RMSE > mean + 1 SD.",
    locations = cells_column_labels(columns = RMSE)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-warning}
## Departmental Variation
`r n_dept_flagged` of `r n_departments` departments exceed the mean RMSE by
more than one standard deviation and are flagged in red in
@tbl-d2-cv-by-department: `r dept_flagged_list`. Two patterns sit behind the
flag. Sacatepéquez, Santa Rosa and Sololá hold 4 to 12 children each, so their
RMSE is estimated on very few cases. Escuintla and Zacapa hold 76 and 30 and
carry a mean bias above 1.7 cm, an underprediction of height that the national
figures absorb. Predictions for these five departments carry higher
uncertainty.
:::

### Stunting Classification Quality

```{r}
#| label: tbl-d3-stunting-classification
#| tbl-cap: "Stunting classification accuracy from predicted heights (CV)"

tbl_d3_stunting_classification |>
  gt() |>
  tab_header(
    title = "Prediction Quality by Stunting Status",
    subtitle = "Differential accuracy for stunted vs non-stunted children"
  ) |>
  cols_label(Metric = "Metric", Value = "Value") |>
  fmt_number(columns = Value, decimals = 2) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-important}
## Shrinkage Toward the Conditional Mean
Absolute accuracy is comparable for both groups: MAE is
`r round(mae_stunted, 2)` cm for stunted children against
`r round(mae_not_stunted, 2)` cm for the rest. The residuals are not
symmetric, though. Observed minus predicted height averages
`r round(resid_stunted, 2)` cm for stunted children and
`r round(resid_not_stunted, 2)` cm for the others, so the model predicts
stunted children taller than they are and non-stunted children shorter.

This is the shrinkage a conditional-mean estimator produces when the outcome
carries variation the predictors do not reach: the fitted values compress
toward the centre and the tails of the height distribution lose mass. Applying
these predictions directly would understate stunting prevalence. The
distribution is restored by the quantile mapping and the weight calibration in
[transferring height to ENCOVI](05_03_transfer_height_to_encovi.qmd), which is
where the population-level figures are produced.
:::


## Model Diagnostics

### Residual Analysis

```{r}
#| label: fig-e1-residual-diagnostics
#| fig-cap: "GAM residual diagnostic plots"
#| fig-alt: >
#|   Four-panel diagnostic grid. Top left plots deviance residuals against
#|   fitted values with a loess curve that stays flat, showing no systematic
#|   curvature. Top right is a Q-Q plot of standardised residuals following the
#|   diagonal reference line closely, with only minor departures at the tails.
#|   Bottom left is a scale-location plot of the square root of standardised
#|   residuals against fitted values with a flat smooth, indicating roughly
#|   constant variance. Bottom right is a histogram of standardised residuals
#|   that closely tracks the overlaid standard normal curve.
#| fig-width: 10
#| fig-height: 8

# Plot 1: Residuals vs Fitted (Linearity)
plot_resid_fitted <- fig_e1_residual_diagnostics |>
  ggplot(aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.3, size = 1, color = "#1B6CA8") +
  geom_smooth(
    formula = 'y ~ x',
    se = FALSE,
    method = "loess",
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Residuals vs Fitted",
    subtitle = "Reference line should be flat and horizontal",
    x = "Fitted Values",
    y = "Deviance Residuals"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Plot 2: Q-Q Plot of Residuals (Normality)
plot_qq <- fig_e1_residual_diagnostics |>
  ggplot(aes(x = theoretical_q, y = std_residuals)) +
  geom_point(alpha = 0.3, size = 1, color = "#1B6CA8") +
  geom_abline(
    intercept = 0,
    slope = 1,
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Q-Q Plot of Residuals",
    subtitle = "Points should follow the diagonal reference line",
    x = "Theoretical Quantiles",
    y = "Standardized Residuals"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Plot 3: Scale-Location (Homoscedasticity)
plot_scale_loc <- fig_e1_residual_diagnostics |>
  ggplot(aes(x = fitted, y = sqrt_abs_std)) +
  geom_point(alpha = 0.3, size = 1, color = "#1B6CA8") +
  geom_smooth(
    formula = 'y ~ x',
    se = FALSE,
    method = "loess",
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Homogeneity of Variance",
    subtitle = "Reference line should be flat and horizontal",
    x = "Fitted Values",
    y = expression(sqrt("Standardized Residuals"))
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Plot 4: Histogram of Residuals (Distribution)
plot_hist_resid <- fig_e1_residual_diagnostics |>
  ggplot(aes(x = std_residuals)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 40,
    fill = "lightgray",
    color = "azure4",
    alpha = 0.5
  ) +
  stat_function(
    fun = dnorm,
    args = list(mean = 0, sd = 1),
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Distribution of Residuals",
    subtitle = "Dashed line = standard normal reference",
    x = "Standardized Residuals",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Combine into 2x2 grid
ggarrange(
  plot_resid_fitted,
  plot_qq,
  plot_scale_loc,
  plot_hist_resid,
  ncol = 2, nrow = 2
)
```

::: {.callout-note}
## Diagnostic Assessment
The four panels read as follows: residuals against fitted values show no
systematic curvature, the Q-Q plot follows the diagonal with minor departures
at the tails, the scale-location plot is approximately flat, and the residual
histogram tracks the standard normal reference.
:::

### Smooth Term Partial Effects

```{r}
#| label: fig-e2-smooth-effects
#| fig-cap: "Partial effects of smooth terms"
#| fig-alt: >
#|   Panels showing the estimated partial effect of each smooth term on height,
#|   each with a fitted curve, a shaded confidence band, partial residual
#|   points and a data rug. The age panel shows a steep positive slope from 6
#|   to about 24 months that flattens progressively through 59 months, the
#|   characteristic decelerating growth curve. The parental education panel
#|   shows a weak positive gradient. The department random-effect panel is a
#|   Q-Q plot whose points fall near the diagonal, indicating approximately
#|   normal department-level variation.
#| fig-width: 10
#| fig-height: 8

draw(gam_height, residuals = TRUE, rug = TRUE) &
  theme_minimal() &
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 9),
    plot.caption = element_text(color = "azure4", size = 8.5),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    strip.text = element_text(color = "azure4", size = 10)
  ) &
  scale_colour_manual(values = c("#1B6CA8"), guide = "none") &
  scale_fill_manual(values = c("#1B6CA8"), guide = "none")
```

The age partial effect shows the decelerating growth curve: a steep positive
slope from 6 to around 24 months that flattens through 59 months. The parental
education smooth shows a weak positive gradient. The department panel is a Q-Q
plot of the random-effect term against the normal distribution.

### Concurvity Assessment

```{r}
#| label: tbl-e3-concurvity
#| tbl-cap: "Concurvity assessment for smooth terms"

tbl_e3_concurvity |>
  gt() |>
  tab_header(
    title = "Pairwise Concurvity (values > 0.3)",
    subtitle = "GAM analogue of multicollinearity between smooth terms"
  ) |>
  cols_label(
    Term = "Term",
    Against = "Against",
    Concurvity = "Concurvity"
  ) |>
  fmt_number(columns = Concurvity, decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = Concurvity, rows = Concurvity > 0.8)
  ) |>
  tab_footnote(
    footnote = "Values > 0.8 (red) indicate high concurvity requiring attention.",
    locations = cells_column_labels(columns = Concurvity)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Concurvity
The highest pairwise concurvity is `r round(max_concurvity, 3)`, between the
department term and the parametric block, below the 0.8 value at which the
table flags a term. Geographic location and socioeconomic indicators are
correlated in Guatemala, so some overlap between the two is expected. No term
is removed on concurvity grounds.
:::

### Observed vs Fitted (Full Model)

```{r}
#| label: fig-e4-observed-vs-fitted
#| fig-cap: "Observed vs fitted height (full model, not CV)"
#| fig-alt: >
#|   Scatter plot of fitted height on the y-axis against observed height on the
#|   x-axis for the full training model, with a dashed one-to-one reference
#|   line and equal axis scaling. Points fall tightly along the diagonal across
#|   the full height range, in line with the adjusted R-squared and deviance
#|   explained reported in the subtitle.
#| fig-width: 7
#| fig-height: 6

fig_e4_observed_vs_fitted |>
  ggplot(aes(x = talla_cm, y = fitted)) +
  geom_point(alpha = 0.3, color = "#1B6CA8", size = 1.2) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "#3AAF85",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  labs(
    title = "Observed vs Fitted Height (Full Model)",
    subtitle = paste0(
      "Adj. R² = ", round(full_r2, 3),
      ", Deviance explained = ",
      round(full_dev_expl * 100, 1), "%"
    ),
    x = "Observed Height (cm)",
    y = "Fitted Height (cm)"
  ) +
  coord_equal() +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )
```


## Summary

### Key Findings

1. **Model fit**: The GAM accounts for `r round(full_dev_expl * 100, 1)`% of
   deviance (Adj. R² = `r round(full_r2, 3)`) with three significant smooth
   terms and parametric predictors covering demographic, geographic and
   household-asset dimensions.

2. **Out-of-sample accuracy**: Cross-validation gives
   RMSE = `r round(cv_rmse, 2)` cm and R² = `r round(cv_r2, 3)`, with a mean
   bias of `r round(tbl_c2_cv_overall$Mean_Bias, 2)` cm and fold-level RMSE
   between `r round(min(tbl_c2_cv_by_fold$RMSE), 2)` and
   `r round(max(tbl_c2_cv_by_fold$RMSE), 2)` cm.

3. **Age dominance**: The age smooth carries the non-linear growth curve,
   while sex and the socioeconomic predictors add discriminating information
   on top of it.

4. **Shrinkage toward the conditional mean**: Height is overpredicted by
   `r abs(round(resid_stunted, 2))` cm on average for stunted children and
   underpredicted by `r round(resid_not_stunted, 2)` cm for the rest. The
   quantile mapping applied downstream restores the spread of the
   distribution.

::: {.callout-warning}
## Limitations

1. **Training sample size** — `r format(n_observations, big.mark = ",")`
   original observations, `r format(n_expanded, big.mark = ",")` after
   expansion by survey weights, is a modest base for a GAM with
   `r n_predictors` predictors and three smooth terms.

2. **Age-dependent accuracy** — RMSE rises from 3.25 cm in the 6-11 month
   group to 4.69 cm in the 48-59 month group, following the greater spread of
   height at older ages.

3. **Departmental coverage** — `r n_dept_flagged` departments carry an RMSE
   more than one standard deviation above the departmental mean, in two cases
   with a mean bias above 1.7 cm.

4. **Cross-survey transfer assumption** — The model assumes that the
   height-socioeconomic relationship in SIVESNU 2018, a health centre sample,
   holds in ENCOVI 2023, a household survey. The sampling frame difference is
   addressed by the calibration in
   [transferring height to ENCOVI](05_03_transfer_height_to_encovi.qmd).
:::

### Application

The trained GAM is applied in
[transferring height to ENCOVI](05_03_transfer_height_to_encovi.qmd) to predict
height for ENCOVI 2023 children from their socioeconomic profiles. Those
predictions are calibrated and converted into individual height-for-age
z-scores through the WHO LMS reference tables, which is what the Meta-analysis
(Gunaratna) pathway needs to estimate the change in stunting under
biofortified maize adoption.
