---
title: "Biofortification Priority Assignment"
subtitle: "Module 5: Meta-analysis"
---

## Why Assign Biofortification Priority?

Biofortified maize reaches different populations through different
channels. Children in households that produce and consume their own maize
are the most direct beneficiaries; children in non-agricultural urban
households access biofortified varieties through progressively more
indirect market mechanisms. This module assigns a five-level priority
hierarchy to each child based on their household's agricultural profile,
determining the order in which children receive biofortified maize in
scenario simulations.

### Priority Hierarchy

| Level | Label | Definition |
|-------|-------|------------|
| 1 | Maize farm + self-consumption | Produces white maize, retains for own consumption |
| 2 | Maize farm, no self-consumption | Produces white maize, sells entire harvest |
| 3 | Non-maize agricultural household | Agricultural production but no white maize |
| 4 | Rural non-agricultural | Rural household, no reported production |
| 5 | Urban | Urban household |

The hierarchy is constructed at the household level: children in the same
household receive the same priority, since biofortified maize enters the
household food supply as a whole.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Professional table formatting
  scales,             # Number formatting (percent_format) for barplot
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))
```

```{r}
#| label: load-rds
#| include: false

# Web resources for visualizations (exported by framework 05_06)

# Phase B — Household vector validation
tbl_b4_household_vector_validation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_06_tbl_b4_household_vector_validation.rds"),
  trust = TRUE
)

tbl_b4_assert_mutual_exclusivity <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_06_tbl_b4_assert_mutual_exclusivity.rds"),
  trust = TRUE
)

# Phase C — Completeness check
tbl_c2_completeness_check <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_06_tbl_c2_completeness_check.rds"),
  trust = TRUE
)

# Phase D — Distribution analysis
tbl_d1_national_distribution_unweighted <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_06_tbl_d1_national_distribution_unweighted.rds"),
  trust = TRUE
)

tbl_d2_national_distribution_weighted <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_06_tbl_d2_national_distribution_weighted.rds"),
  trust = TRUE
)

tbl_d3_departmental_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_06_tbl_d3_departmental_distribution.rds"),
  trust = TRUE
)

fig_d4_priority_distribution_barplot <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_06_fig_d4_priority_distribution_barplot.rds"),
  trust = TRUE
)

tbl_d5_stunting_by_priority <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_06_tbl_d5_stunting_by_priority.rds"),
  trust = TRUE
)

# Phase E — Output structure (NEW)
tbl_e1_output_structure <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_06_tbl_e1_output_structure.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Total agricultural households (sum of levels 1, 2, 3 from tbl-b4-household)
n_total_agri <- tbl_b4_household_vector_validation |>
  filter(level == "Total agricultural households") |>
  pull(n_households)

n_level_1 <- tbl_b4_household_vector_validation |>
  filter(level == "Level 1: Maize + self-consumption") |>
  pull(n_households)

n_level_2 <- tbl_b4_household_vector_validation |>
  filter(level == "Level 2: Maize, no self-consumption") |>
  pull(n_households)

# Analytical cohort size (from tbl-d1)
n_children_total <- tbl_d1_national_distribution_unweighted |>
  summarise(n = sum(n_children)) |>
  pull(n)

# Population structure (weighted % from tbl-d2)
pct_level_1 <- tbl_d2_national_distribution_weighted |>
  filter(priority_level == 1L) |>
  pull(pct_weighted)

pct_level_2 <- tbl_d2_national_distribution_weighted |>
  filter(priority_level == 2L) |>
  pull(pct_weighted)

pct_level_3 <- tbl_d2_national_distribution_weighted |>
  filter(priority_level == 3L) |>
  pull(pct_weighted)

pct_levels_123 <- pct_level_1 + pct_level_2 + pct_level_3
```


## Data Sources

Three datasets are integrated in the framework: the
[ENCOVI children with stunting profiles](05_05_calculate_qpm_stunting_impact.qmd),
the white maize farmer dataset with self-consumption flags built in the
economic impact module for the 2022 seed, and the raw ENCOVI agricultural
production module, covering all 82 crop types, which identifies the
non-agricultural and non-maize agricultural households.

The hierarchy is built on the 2022 seed farmer dataset. Seed generation enters
the economic projections and the adoption scores, and leaves the
household-level agricultural profile unchanged: the two variables read here,
`no_hogar` and `has_household_consumption`, take the same values for the 2022,
2026 and 2028 seeds. The priority assignment therefore serves every seed
scenario.


## Household Agricultural Classification

Three mutually exclusive household vectors are constructed based on
agricultural activity. Level 1 identifies maize-producing households that
retain harvest for consumption; Level 2 identifies maize producers who
sell everything; Level 3 identifies households with other agricultural
production but no white maize.

### Vector Validation

```{r}
#| label: tbl-b4-household-vector-validation
#| tbl-cap: "Household agricultural vector validation: mutual exclusivity check"

tbl_b4_household_vector_validation |>
  gt() |>
  tab_header(
    title = "Household Agricultural Vector Validation",
    subtitle = "Mutual exclusivity check: Levels 1 + 2 + 3 = Total agricultural"
  ) |>
  cols_label(
    level = "Level",
    n_households = "N Households"
  ) |>
  fmt_number(columns = n_households, decimals = 0, use_seps = TRUE) |>
  tab_footnote(
    footnote = "Counts refer to sampled households in ENCOVI 2023.",
    locations = cells_column_labels(columns = n_households)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14
  )
```

::: {.callout-note}
## Mutual Exclusivity
The three agricultural household vectors sum to the total number of
agricultural households in the raw production data,
`r format(n_total_agri, big.mark = ",")`, and their pairwise intersections are
empty, so the classification covers the agricultural households once each.
Level 1, maize with self-consumption, holds
`r format(n_level_1, big.mark = ",")` households against
`r format(n_level_2, big.mark = ",")` in Level 2, so
`r round(100 * n_level_1 / (n_level_1 + n_level_2), 1)`% of maize-producing
households retain part of their harvest.
:::

```{r}
#| label: tbl-b4-assert-mutual-exclusivity
#| tbl-cap: "Mutual exclusivity verification between priority levels"

tbl_b4_assert_mutual_exclusivity |>
  gt() |>
  tab_header(
    title = "Mutual Exclusivity Verification",
    subtitle = "All intersections should be zero"
  ) |>
  cols_label(
    comparison = "Comparison",
    n_overlap = "N Overlap",
    status = "Status"
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = everything(), rows = status == "PASS")
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(columns = everything(),
                           rows = str_detect(status, "FAIL"))
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14
  )
```


## Priority Assignment

Each child receives a priority level via strict hierarchical precedence: a
child is assigned to the highest applicable level. Levels 4 and 5 are
residual categories based on the `area` variable (Rural/Urbana) from the
ENCOVI household module.

```{r}
#| label: tbl-c2-completeness-check
#| tbl-cap: "Assignment completeness verification for all sampled children"

tbl_c2_completeness_check |>
  gt() |>
  tab_header(
    title = "Assignment Completeness Verification",
    subtitle = "All checks should show 0 (except total children)"
  ) |>
  cols_label(
    check = "Check",
    value = "Value"
  ) |>
  fmt_number(columns = value, decimals = 0, use_seps = TRUE) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14
  )
```


## Distribution Analysis

### National Distribution (Unweighted)

```{r}
#| label: tbl-d1-national-distribution-unweighted
#| tbl-cap: "Child distribution by biofortification priority level (unweighted sample counts)"

tbl_d1_national_distribution_unweighted |>
  gt() |>
  tab_header(
    title = "Child Distribution by Biofortification Priority Level",
    subtitle = "Unweighted sample counts — ENCOVI 2023 children aged 6–59 months"
  ) |>
  fmt_number(columns = c(n_children, n_households),
             decimals = 0, use_seps = TRUE) |>
  fmt_percent(columns = c(pct_children, pct_cumulative), decimals = 1) |>
  cols_label(
    priority_level = "Level",
    priority_label = "Description",
    n_children = "Children",
    n_households = "Households",
    pct_children = "% of Total",
    pct_cumulative = "Cumulative %"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14
  )
```

### National Distribution (Survey-Weighted)

```{r}
#| label: tbl-d2-national-distribution-weighted
#| tbl-cap: "Child distribution by biofortification priority level (survey-weighted population estimates)"

tbl_d2_national_distribution_weighted |>
  gt() |>
  tab_header(
    title = "Child Distribution by Biofortification Priority Level",
    subtitle = "Survey-weighted population estimates — ENCOVI 2023 children aged 6–59 months"
  ) |>
  fmt_number(columns = c(n_weighted, n_weighted_low, n_weighted_upp),
             decimals = 0, use_seps = TRUE) |>
  fmt_percent(columns = c(pct_weighted, pct_cumulative), decimals = 1) |>
  cols_label(
    priority_level = "Level",
    priority_label = "Description",
    n_weighted = "Population Estimate",
    n_weighted_low = "95% CI Lower",
    n_weighted_upp = "95% CI Upper",
    pct_weighted = "% of Total",
    pct_cumulative = "Cumulative %"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14
  )
```

::: {.callout-important}
## Population Structure
At the population level, `r scales::percent(pct_level_1, accuracy = 0.1)`
of children aged 6-59 months live in households that produce and consume
their own maize (Level 1). An additional
`r scales::percent(pct_level_2, accuracy = 0.1)` live in maize-producing
households without self-consumption (Level 2), and
`r scales::percent(pct_level_3, accuracy = 0.1)` in non-maize agricultural
households (Level 3). The first three levels, which carry the most direct
access to biofortified varieties, cover
`r scales::percent(pct_levels_123, accuracy = 0.1)` of the child population,
so around four in five children are reached through the market channels of
Levels 4 and 5 (see @tbl-d2-national-distribution-weighted).
:::

### Departmental Distribution

```{r}
#| label: tbl-d3-departmental-distribution
#| tbl-cap: "Departmental distribution of children by priority level (survey-weighted)"

tbl_d3_departmental_distribution |>
  gt() |>
  tab_header(
    title = "Departmental Distribution of Children by Priority Level",
    subtitle = "Survey-weighted percentage within each department"
  ) |>
  fmt_percent(columns = starts_with("level_"), decimals = 1) |>
  cols_label(
    departamento = "Department",
    level_1 = "L1: Maize + SC",
    level_2 = "L2: Maize, no SC",
    level_3 = "L3: Non-maize Ag.",
    level_4 = "L4: Rural",
    level_5 = "L5: Urban"
  ) |>
  tab_footnote(
    footnote = "SC = self-consumption. Ag. = agricultural household.",
    locations = cells_column_labels(columns = starts_with("level_"))
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14
  )
```

```{r}
#| label: fig-d4-priority-distribution-barplot
#| fig-cap: "Distribution of children across biofortification priority levels by department"
#| fig-alt: >
#|   Stacked bar chart with one bar per department on the x-axis and the
#|   survey-weighted proportion of children on the y-axis. Each bar is divided
#|   into the five priority levels, from maize farm with self-consumption through
#|   maize farm without self-consumption, non-maize agricultural, rural
#|   non-agricultural, and urban. Departments are ordered from left to right by
#|   their combined agricultural household share (Levels 1 to 3). The chart shows
#|   wide geographic variation: rural maize-dependent departments on the left are
#|   dominated by the agricultural levels, while urban-dominant departments on the
#|   right are almost entirely the urban level.
#| fig-width: 10
#| fig-height: 8

# Order departments by agricultural household share (Levels 1-3 combined)
dept_order <- fig_d4_priority_distribution_barplot |>
  filter(priority_label %in% c(
    "Maize farm + self-consumption",
    "Maize farm, no self-consumption",
    "Non-maize agricultural household"
  )) |>
  summarise(pct_agri = sum(pct), .by = departamento) |>
  arrange(desc(pct_agri)) |>
  pull(departamento)

fig_d4_priority_distribution_barplot |>
  mutate(departamento = factor(departamento, levels = dept_order)) |>
  ggplot(aes(x = departamento, y = pct, fill = priority_label)) +
  geom_col(width = 0.4) +
  scale_y_continuous(labels = percent_format(),
                     expand = expansion(mult = c(0, 0.02))) +
  scale_fill_manual(
    values = c(
      "Maize farm + self-consumption"    = "#2E7D32",
      "Maize farm, no self-consumption"  = "#66BB6A",
      "Non-maize agricultural household" = "#A5D6A7",
      "Rural non-agricultural"           = "#FFB74D",
      "Urban"                            = "#90A4AE"
    ),
    name = "Priority Level"
  ) +
  labs(
    title = "Child Distribution by Biofortification Priority Level",
    subtitle = "Departments ordered by agricultural household share (Levels 1–3)",
    x = NULL,
    y = "Proportion of children (survey-weighted)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    legend.position = "bottom",
    legend.text = element_text(size = 9),
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey40")
  ) +
  guides(fill = guide_legend(nrow = 3, byrow = TRUE))
```

The agricultural household share varies widely across departments (see
@fig-d4-priority-distribution-barplot). It reaches 39.7% in Huehuetenango and
37.0% in Chiquimula at one end, and falls to 2.0% in Guatemala and 2.2% in El
Progreso at the other, where Level 5 covers 89.3% and 62.0% of children
respectively.

### Stunting by Priority Level

```{r}
#| label: tbl-d5-stunting-by-priority
#| tbl-cap: "Baseline stunting prevalence by biofortification priority level"

tbl_d5_stunting_by_priority |>
  gt() |>
  tab_header(
    title = "Baseline Stunting Prevalence by Priority Level",
    subtitle = "Survey-weighted proportion of children with HAZ < -2 SD"
  ) |>
  fmt_percent(columns = c(stunting_prev, stunting_prev_low, stunting_prev_upp),
              decimals = 1) |>
  fmt_number(columns = n_total, decimals = 0, use_seps = TRUE) |>
  cols_label(
    priority_level = "Level",
    priority_label = "Description",
    stunting_prev = "Stunting %",
    stunting_prev_low = "95% CI Lower",
    stunting_prev_upp = "95% CI Upper",
    n_total = "Population"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14
  )
```

::: {.callout-note}
## Stunting by Level
The separation runs between rural and urban rather than along the five levels
in order (see @tbl-d5-stunting-by-priority). Levels 1, 3 and 4 sit together at
65.3%, 62.0% and 66.6%, and Level 5 stands apart at 24.1%. The highest
prevalence is Level 4, rural non-agricultural households, not Level 1.

Level 2 reads 36.2%, but it rests on 10 sampled children and its confidence
interval runs from 4.1% to 68.2%, so the point estimate carries little
information. The hierarchy orders access to biofortified maize; it is not
ordered by stunting prevalence.
:::


## Output Dataset Structure

The output dataset carries the variables of the
[QPM stunting impact dataset](05_05_calculate_qpm_stunting_impact.qmd) and adds
two: `priority_level`, an integer from 1 to 5, and `priority_label`, a factor
with the descriptive labels. The existing columns pass through unchanged.

```{r}
#| label: tbl-e1-output-structure
#| tbl-cap: "Output dataset structure summary"

tbl_e1_output_structure |>
  gt() |>
  tab_header(
    title = "Output Dataset Structure",
    subtitle = "Priority assignment added to the QPM stunting impact dataset"
  ) |>
  cols_label(
    metric = "Metric",
    value = "Value"
  ) |>
  fmt_number(columns = value, decimals = 0, use_seps = TRUE) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14
  )
```


## Summary

### Key Findings

1. **Complete assignment**: Each of the
   `r format(n_children_total, big.mark = ",")` children receives a priority
   classification, with no missing values in `priority_level`, `area` or
   `no_hogar` (see @tbl-c2-completeness-check and
   @tbl-d1-national-distribution-unweighted).

2. **Agricultural household share**:
   `r scales::percent(pct_levels_123, accuracy = 0.1)` of the child
   population (Levels 1-3) lives in households with direct or indirect
   agricultural connections to the biofortification supply chain
   (see @tbl-d2-national-distribution-weighted).

3. **Maize self-consumption dominance**:
   `r round(100 * n_level_1 / (n_level_1 + n_level_2), 1)`% of
   maize-producing households retain part of their production for own
   consumption (Level 1 against Level 2 in
   @tbl-b4-household-vector-validation), which places most of the direct
   channel inside the household itself.

4. **Stunting by level**: Levels 1, 3 and 4 carry baseline stunting between
   62.0% and 66.6% and Level 5 carries 24.1% (see
   @tbl-d5-stunting-by-priority), so the contrast that shows up is rural
   against urban rather than a gradient along the hierarchy.

::: {.callout-warning}
## Limitations

1. **Self-consumption as proxy** — The flag indicates retention of maize
   production but does not quantify the amount consumed. A household
   retaining a small fraction receives the same Level 1 priority as one
   consuming most of its harvest.

2. **Agricultural module coverage** — ENCOVI may undercount subsistence
   farming, particularly among indigenous communities. Some Level 4-5
   households may have unregistered agricultural activity.

3. **Static classification** — The hierarchy is built on ENCOVI 2023
   cross-sectional data, so seasonal variation in agricultural employment
   falls outside it.

4. **Level 2 sample size** — Level 2 rests on 10 sampled children, and its
   stunting estimate carries a confidence interval from 4.1% to 68.2%.
:::

### Application

This priority-enriched dataset is the final input for the scenario
simulation pipeline (Module 6), which uses the priority hierarchy to
assign biofortified maize to children in order of decreasing agricultural
proximity at each production coverage level.
