---
title: "QPM Stunting Impact"
subtitle: "Module 5: Meta-analysis"
description-meta: "Meta-analysis pathway: growth velocity by baseline HAZ and a protein adequacy sigmoid turn QPM intake gains into height and stunting change."
---

## Why Calculate QPM Stunting Impact?

The Meta-analysis (Gunaratna) pathway quantifies the stunting reduction
achievable through biofortified maize consumption. Using the 9% height
growth velocity increase from @Gunaratna2010_QPM — a meta-analysis of 9
community-based RCTs, of which 7 contributed to the height effect estimate
after excluding Mexico (no height data reported) and Nicaragua (statistical
outlier) — this module calculates the individual-level height increment
attributable to quality protein maize (QPM) for each child based on their
baseline HAZ, age, sex, and protein adequacy.

The QPM height increment is modulated by a protein adequacy factor that
carries individual-level heterogeneity: children meeting their protein
requirement take the full 9% effect, and children far below their requirement
take an attenuated one, following a rescaled logistic sigmoid. The modulation
follows the mechanism QPM acts through, a higher protein quality (PDCAAS),
which needs protein quantity in the diet to reach growth.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  scales,             # Number formatting for visualizations
  patchwork,          # Multi-panel plot composition
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))
```

```{r}
#| label: load-rds
#| include: false

# Web resources for visualizations (exported by framework 05_05)

# Phase A — Growth velocity reference
fig_a3_growth_velocity_curves <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_05_fig_a3_growth_velocity_curves.rds"),
  trust = TRUE
)

# Phase B — Growth velocity assignment
tbl_b2_growth_velocity_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_05_tbl_b2_growth_velocity_summary.rds"),
  trust = TRUE
)

tbl_b3_haz_below_neg4 <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_05_tbl_b3_haz_below_neg4.rds"),
  trust = TRUE
)

# Phase C — Protein adequacy factor
tbl_c2_factor_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_05_tbl_c2_factor_summary.rds"),
  trust = TRUE
)

# Phase E — Impact diagnostics
tbl_e1_height_increment_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_05_tbl_e1_height_increment_summary.rds"),
  trust = TRUE
)

tbl_e2_stunting_baseline_vs_qpm <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_05_tbl_e2_stunting_baseline_vs_qpm.rds"),
  trust = TRUE
)

tbl_e3_stunting_by_department <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_05_tbl_e3_stunting_by_department.rds"),
  trust = TRUE
)

fig_e4_delta_height_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_05_fig_e4_delta_height_distribution.rds"),
  trust = TRUE
)

fig_e5_delta_haz_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_05_fig_e5_delta_haz_distribution.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Sigmoid parameters (fixed pipeline constants from framework 05_05)
ADEQUACY_K  <- 8     # Sigmoid slope parameter
ADEQUACY_R0 <- 0.5   # Inflection point (ratio where factor = 0.5)

# Analytical cohort size (sum across tbl-b2 strata)
n_children_total <- tbl_b2_growth_velocity_summary |>
  summarise(n = sum(n)) |>
  pull(n)

# Children with HAZ below -4 SD (from tbl-b3)
n_below_neg4_total <- tbl_b3_haz_below_neg4 |>
  summarise(n = sum(n_below_neg4)) |>
  pull(n)

pct_below_neg4_overall <- tbl_b3_haz_below_neg4 |>
  summarise(
    n_total = sum(n_total),
    n_below = sum(n_below_neg4)
  ) |>
  mutate(pct = n_below / n_total * 100) |>
  pull(pct)

# National stunting and HAZ change under full QPM coverage
e2_value <- function(label) {
  tbl_e2_stunting_baseline_vs_qpm |>
    filter(metric_label == label) |>
    pull(value)
}

stunting_baseline_pct <- e2_value("Stunting baseline (%)")
stunting_qpm_pct      <- e2_value("Stunting with QPM (%)")
stunting_reduction_pp <- abs(e2_value("Stunting reduction (pp)"))
haz_baseline_mean     <- e2_value("Mean HAZ baseline")
haz_qpm_mean          <- e2_value("Mean HAZ with QPM")
haz_delta_mean        <- e2_value("Mean delta HAZ (QPM effect)")

# Departmental spread of the modelled reduction
dept_change_max <- max(abs(tbl_e3_stunting_by_department$change_pp))
dept_change_min <- min(abs(tbl_e3_stunting_by_department$change_pp))
dept_change_cor <- cor(tbl_e3_stunting_by_department$stunting_baseline,
                       tbl_e3_stunting_by_department$change_pp)

# Mean HAZ change by baseline stunting status, for the density comparison
delta_haz_by_status <- fig_e5_delta_haz_distribution |>
  group_by(stunting_label) |>
  summarise(mean_delta = mean(delta_haz_qpm), .groups = "drop")

mean_delta_haz_stunted <- delta_haz_by_status |>
  filter(stunting_label == "Stunted (baseline)") |>
  pull(mean_delta)

mean_delta_haz_not_stunted <- delta_haz_by_status |>
  filter(stunting_label == "Not stunted (baseline)") |>
  pull(mean_delta)
```


## Data and Growth Velocity Reference

The framework loads the
[bioavailability-enriched dataset](05_04_calculate_bioavailability_inadequacy.qmd)
and WHO Child Growth Standards (0-60 months, daily resolution) plus
WHO 2007 Reference [@deOnis2007_GrowthReference] (61-71 months, monthly 
resolution). The combined reference table provides complete coverage for 
calculating 12-month growth velocity for children aged 6-59 months at 
baseline (endline ages 18-71 months).

### Growth Velocity by Baseline HAZ

```{r}
#| label: fig-a3-growth-velocity-curves
#| fig-cap: "12-month growth velocity by baseline age and HAZ category"
#| fig-alt: >
#|   Line chart with two panels, one for boys and one for girls, showing baseline
#|   age in months on the x-axis and expected 12-month growth velocity in
#|   centimetres on the y-axis. Each panel draws one line per baseline
#|   height-for-age category, from minus 4 to plus 3 standard deviations. All
#|   lines slope downward, so growth velocity decreases with age, consistent with
#|   the deceleration of linear growth under age 5. The lines are clearly
#|   separated by category, with lower height-for-age children showing lower
#|   absolute velocities.
#| fig-width: 10
#| fig-height: 7

fig_a3_growth_velocity_curves |>
  ggplot(aes(x = edad_meses, y = velocity_cm, color = sd_label)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ sex_label) +
  scale_color_manual(
    values = c(
      "SD -4" = "#67001F", "SD -3" = "#B2182B", "SD -2" = "#D6604D",
      "SD -1" = "#F4A582", "SD 0 (median)" = "#333333",
      "SD +1" = "#92C5DE", "SD +2" = "#4393C3", "SD +3" = "#2166AC"
    ),
    name = "Baseline HAZ Category"
  ) +
  labs(
    title = "Expected 12-Month Growth Velocity by Baseline Age and HAZ Category",
    subtitle = "Computed from WHO Child Growth Standards (0-60m) and WHO 2007 Reference (61-71m)",
    x = "Baseline Age (months)",
    y = "Growth Velocity (cm / 12 months)"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 13, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    strip.text = element_text(size = 10, face = "bold"),
    legend.position = "bottom"
  ) +
  guides(color = guide_legend(nrow = 2, byrow = TRUE))
```

Growth velocity decreases with age in every HAZ category, consistent with the
deceleration of linear growth in children under 5. Children with lower baseline
HAZ carry lower absolute growth velocities: at 12 months the reference gives
9.6 cm per year at −4 SD against 12.2 cm at the median. The curves are
separated across categories, which is the variation the interpolation reads
each child's velocity from.


## Meta-Analysis and Modulation Parameters

The effect size from @Gunaratna2010_QPM is a 9% increase in height growth
velocity (multiplicative ratio = 1.09; 95% CI: 6-15%, p < 0.0001), derived
from a meta-analysis of 7 community-based RCTs (main dataset) in children
with mild-to-moderate undernutrition and maize-dependent diets. The
intervention is projected over 12 months for the cohort of children aged
6-59 months at baseline.

The protein adequacy factor is a rescaled logistic sigmoid with slope
parameter k = `r ADEQUACY_K` and inflection point r₀ = `r ADEQUACY_R0`.
The biological basis is a dose-response relationship with a homeostatic
tolerance zone near full adequacy and accelerated decline below metabolic
breakpoints (Scheffler et al., 2021; @Gunaratna2010_QPM).


## Growth Velocity Assignment

### Interpolation and Validation

For each child, the expected 12-month growth velocity is interpolated
linearly between the two adjacent HAZ categories in the WHO reference
table. A child with HAZ = -2.3 receives 70% weight from the SD(-2)
velocity and 30% from the SD(-3) velocity, avoiding artificial
discontinuities.

```{r}
#| label: tbl-b2-growth-velocity-summary
#| tbl-cap: "Growth velocity assignment summary by age group and sex"

tbl_b2_growth_velocity_summary |>
  gt(groupname_col = "age_group_5") |>
  tab_header(
    title = "12-Month Growth Velocity Assignment",
    subtitle = "Interpolated from WHO growth velocity tables by baseline HAZ, age, and sex"
  ) |>
  cols_label(
    sexo = "Sex",
    n = "N",
    n_assigned = "Assigned",
    pct_assigned = "% Assigned",
    mean_gv = "Mean (cm)",
    sd_gv = "SD (cm)",
    min_gv = "Min (cm)",
    max_gv = "Max (cm)"
  ) |>
  fmt_number(columns = c(pct_assigned), decimals = 1) |>
  fmt_number(columns = c(mean_gv, sd_gv, min_gv, max_gv), decimals = 2) |>
  tab_source_note("Growth velocity: expected cm of growth in 12 months, interpolated between adjacent HAZ categories.") |>
  tab_source_note("All children aged 6-59 months receive a velocity assignment. WHO under-5 standards (0-60m) and WHO 2007 reference (61-71m) provide complete endline coverage.") |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

::: {.callout-note}
## Complete Assignment
Each of the `r format(n_children_total, big.mark = ",")` children in the cohort
receives a growth velocity assignment. Mean velocity falls from 12.1 cm per 12
months at 6-11 months to 5.4 cm at 48-59 months (see
@tbl-b2-growth-velocity-summary), following the deceleration in the reference.
:::

### HAZ Clamping Below -4 SD

Children with baseline HAZ below -4 SD receive the growth velocity at the
-4 SD boundary. The WHO expanded tables extend coverage to -4 SD,
improving coverage for Guatemala's left-shifted distribution.

```{r}
#| label: tbl-b3-haz-below-neg4
#| tbl-cap: "Children with baseline HAZ below -4 SD (growth velocity clamped to -4 SD boundary)"

tbl_b3_haz_below_neg4 |>
  gt(groupname_col = "age_group_5") |>
  tab_header(
    title = "Children with Baseline HAZ Below -4 SD",
    subtitle = "Growth velocity assigned at the -4 SD boundary value (clamped, not extrapolated)"
  ) |>
  cols_label(
    sexo = "Sex",
    n_total = "N (with GV)",
    n_below_neg4 = "N Below -4 SD",
    pct_below_neg4 = "% Below -4 SD",
    mean_haz_clamped_group = "Mean HAZ (clamped group)",
    min_haz = "Min HAZ"
  ) |>
  fmt_number(columns = c(pct_below_neg4), decimals = 1) |>
  fmt_number(columns = c(mean_haz_clamped_group, min_haz), decimals = 2) |>
  sub_missing(missing_text = "---") |>
  tab_source_note("These children receive the same growth velocity as a child at HAZ = -4.0.") |>
  tab_source_note("WHO expanded tables extend coverage to -4 SD. Values beyond this boundary are conservatively assigned the -4 SD velocity.") |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

::: {.callout-note}
## Boundary Treatment
`r round(pct_below_neg4_overall, 1)`% of children,
`r format(n_below_neg4_total, big.mark = ",")` in total, have a baseline HAZ
below -4 SD, with the lowest observed at -5.24. They take the velocity at the
-4 SD boundary, which the reference tables cover, and the QPM increment for
these children is computed on that value.
:::


## Protein Adequacy Factor

### Rescaled Logistic Sigmoid

The protein adequacy factor modulates the QPM effect based on the child's
protein adequacy ratio under the intervention. The function satisfies
three anchor conditions: f(0) = 0, f(r₀) = 0.5, and f(1) = 1.

$$f(r) = \frac{\sigma(r) - \sigma(0)}{\sigma(1) - \sigma(0)}, \quad \sigma(r) = \frac{1}{1 + e^{-k(r - r_0)}}$$

Where k = `r ADEQUACY_K` (slope) and r₀ = `r ADEQUACY_R0` (inflection
point). Near full adequacy, between ratios 0.8 and 1.0, the factor stays close
to 1, which is the homeostatic tolerance zone. Below a ratio of 0.4 it declines
sharply, at the metabolic breakpoints where protein intake constrains growth.

```{r}
#| label: fig-c1-sigmoid-curve
#| fig-cap: "Theoretical protein adequacy factor: rescaled logistic sigmoid"
#| fig-alt: >
#|   Line chart showing the protein adequacy ratio (QPM intake divided by
#|   requirement) on the x-axis from 0 to 1.2 and the modulating adequacy factor
#|   on the y-axis from 0 to 1. The curve is an S-shaped logistic sigmoid rising
#|   from 0 at a ratio of 0, passing through 0.5 at the inflection point marked by
#|   dashed reference lines at ratio 0.5, and levelling off near 1 as the ratio
#|   approaches and exceeds full adequacy. The steep central rise and flat upper
#|   plateau show that children near full protein adequacy receive close to the
#|   full QPM effect while the effect declines sharply below a ratio of about 0.4.
#| fig-width: 8
#| fig-height: 5

# Precompute sigmoid anchor values for rescaling
sigmoid_f0 <- 1 / (1 + exp(-ADEQUACY_K * (0 - ADEQUACY_R0)))
sigmoid_f1 <- 1 / (1 + exp(-ADEQUACY_K * (1 - ADEQUACY_R0)))

df_sigmoid_curve <- tibble(
  ratio = seq(0, 1.2, by = 0.005)
) |>
  mutate(
    sigmoid_raw = 1 / (1 + exp(-ADEQUACY_K * (ratio - ADEQUACY_R0))),
    adequacy_factor = pmin(1, pmax(0,
      (sigmoid_raw - sigmoid_f0) / (sigmoid_f1 - sigmoid_f0)
    ))
  )

df_sigmoid_curve |>
  ggplot(aes(x = ratio, y = adequacy_factor)) +
  geom_line(color = "#3AAF85", linewidth = 1.0) +
  geom_vline(xintercept = 0.5, linetype = "dashed",
             color = "grey70", linewidth = 0.3) +
  geom_hline(yintercept = 0.5, linetype = "dashed",
             color = "grey70", linewidth = 0.3) +
  annotate("text", x = 0.52, y = 0.47,
           label = "Inflection (r0 = 0.5)",
           hjust = 0, size = 3, color = "grey40") +
  coord_cartesian(xlim = c(0, 1.2), ylim = c(0, 1.05)) +
  scale_x_continuous(breaks = seq(0, 1.2, 0.2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  labs(
    title = "Protein Adequacy Factor: Theoretical Sigmoid Curve",
    subtitle = "Rescaled logistic sigmoid (k = 8, r0 = 0.5) — anchor conditions: f(0) = 0, f(r0) = 0.5, f(1) = 1",
    x = "Protein Adequacy Ratio (QPM intake / requirement)",
    y = "Adequacy Factor"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 13, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4")
  )
```

```{r}
#| label: tbl-c2-factor-summary
#| tbl-cap: "Distribution of protein adequacy factor by age group"

tbl_c2_factor_summary |>
  gt() |>
  tab_header(
    title = "Protein Adequacy Factor by Age Group",
    subtitle = "Mean values for children with assigned growth velocity"
  ) |>
  cols_label(
    age_group_5 = "Age Group",
    n = "N",
    mean_adequacy = "Mean",
    sd_adequacy = "SD",
    median_adequacy = "Median",
    min_adequacy = "Min"
  ) |>
  fmt_number(columns = c(mean_adequacy, sd_adequacy,
                         median_adequacy, min_adequacy), decimals = 3) |>
  tab_source_note("Adequacy factor: rescaled sigmoid (k=8, r0=0.5) applied to QPM protein adequacy ratio.") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

::: {.callout-important}
## Adequacy Factor Distribution
The mean factor ranges from 0.777 at 6-11 months, the lowest of the five
bands, to 0.895 at 24-35 months (see @tbl-c2-factor-summary). Medians sit at
1.000 in four bands and 0.984 at 6-11 months, so most children take the
unattenuated effect and the sigmoid acts on the lower tail of the protein
adequacy distribution. The minimum is 0 in every band.
:::


## QPM Height Increment and Stunting Reclassification

The QPM-attributable height increment combines three components:

$$\Delta h_{QPM} = v_{12m} \times E \times A$$

Where $v_{12m}$ is the 12-month growth velocity, $E$ = 0.09 (meta-analysis
effect size), and $A$ is the protein adequacy factor. The final height
under QPM is $h_{QPM} = h_{baseline} + v_{12m} + \Delta h_{QPM}$, and HAZ
is recalculated using WHO LMS parameters at the endline age.


## Results

### Height Increment by Age Group

```{r}
#| label: tbl-e1-height-increment-summary
#| tbl-cap: "QPM-attributable height increment by age group"

tbl_e1_height_increment_summary |>
  gt() |>
  tab_header(
    title = "QPM-Attributable Height Increment by Age Group",
    subtitle = "ENCOVI 2023, children 6-59 months — 100% coverage scenario (survey-weighted)"
  ) |>
  cols_label(
    age_group_5 = "Age Group",
    n = "N",
    mean_gv = "Growth Vel. (cm/12m)",
    mean_gv_low = "95% CI",
    mean_gv_upp = "",
    mean_delta_h = "QPM Increment (cm)",
    mean_delta_h_low = "95% CI",
    mean_delta_h_upp = "",
    mean_delta_haz = "Delta HAZ",
    mean_delta_haz_low = "95% CI",
    mean_delta_haz_upp = ""
  ) |>
  cols_merge(columns = c(mean_gv_low, mean_gv_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(mean_delta_h_low, mean_delta_h_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(mean_delta_haz_low, mean_delta_haz_upp), pattern = "[{1}, {2}]") |>
  fmt_number(columns = c(mean_gv, mean_gv_low), decimals = 2) |>
  fmt_number(columns = c(mean_delta_h, mean_delta_h_low), decimals = 3) |>
  fmt_number(columns = c(mean_delta_haz, mean_delta_haz_low), decimals = 4) |>
  tab_source_note("Effect size: 9% (Gunaratna et al., 2010). Modulation: sigmoid protein adequacy factor (k=8, r0=0.5).") |>
  tab_source_note("N: unweighted sample size. All estimates survey-weighted.") |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

::: {.callout-note}
## Age-Dependent Impact
The QPM height increment falls from 0.854 cm at 6-11 months to 0.424 cm at
48-59 months (see @tbl-e1-height-increment-summary), which follows the growth
velocity it is a proportion of. The corresponding change in HAZ moves in the
same direction, from 0.292 to 0.088.
:::

### National Stunting Impact

```{r}
#| label: tbl-e2-stunting-baseline-vs-qpm
#| tbl-cap: "Stunting impact: baseline vs. QPM intervention (100% coverage)"

tbl_e2_stunting_baseline_vs_qpm |>
  gt() |>
  tab_header(
    title = "Stunting Impact: Baseline vs. QPM Intervention (100% Coverage)",
    subtitle = "ENCOVI 2023, children 6-59 months (survey-weighted)"
  ) |>
  cols_label(
    n = "N",
    metric_label = "Metric",
    value = "Value"
  ) |>
  fmt_number(columns = value, decimals = 2) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = value,
                           rows = metric_label == "Stunting reduction (pp)")
  ) |>
  tab_source_note("Stunting: HAZ < -2 (WHO). QPM effect: 9% growth velocity increase (Gunaratna, 2010).") |>
  tab_source_note("N: unweighted sample size. All estimates survey-weighted.") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

::: {.callout-important}
## Stunting Reduction
At 100% QPM coverage the modelled national stunting prevalence moves from
`r round(stunting_baseline_pct, 1)`% to `r round(stunting_qpm_pct, 1)`%, a
reduction of `r round(stunting_reduction_pp, 1)` percentage points, and mean
HAZ from `r round(haz_baseline_mean, 3)` to `r round(haz_qpm_mean, 3)`, a
change of `r round(haz_delta_mean, 3)`. This is the upper bound of the model,
computed with all maize replaced by biofortified varieties.
:::

### Departmental Impact

```{r}
#| label: tbl-e3-stunting-by-department
#| tbl-cap: "Stunting impact by department (100% coverage)"

tbl_e3_stunting_by_department |>
  gt() |>
  tab_header(
    title = "Stunting Impact by Department — QPM Intervention (100% Coverage)",
    subtitle = "ENCOVI 2023 (survey-weighted)"
  ) |>
  cols_label(
    departamento = "Department",
    n = "N",
    stunting_baseline = "Baseline (%)",
    stunting_qpm = "QPM (%)",
    change_pp = "Change (pp)",
    mean_haz_baseline = "HAZ Baseline",
    mean_haz_qpm = "HAZ QPM",
    mean_delta_haz = "Delta HAZ"
  ) |>
  fmt_number(columns = c(stunting_baseline, stunting_qpm, change_pp),
             decimals = 1) |>
  fmt_number(columns = c(mean_haz_baseline, mean_haz_qpm, mean_delta_haz),
             decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = change_pp, rows = change_pp < 0)
  ) |>
  tab_source_note("QPM effect: 9% growth velocity increase × sigmoid protein adequacy factor.") |>
  tab_source_note("N: unweighted sample size. All estimates survey-weighted.") |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

The modelled reduction ranges from `r round(dept_change_max, 1)` percentage
points in Chiquimula to `r round(dept_change_min, 1)` in Guatemala (see
@tbl-e3-stunting-by-department). It is negatively correlated with baseline
prevalence, at `r round(dept_change_cor, 2)`, so departments starting higher
tend to move further, though the relation is loose: Totonicapán carries the
highest baseline at 70.0% and a reduction of 5.8 pp, while Chiquimula starts at
55.6% and reaches 14.5. The mean HAZ change is close to uniform across
departments, between 0.128 and 0.167, so what varies is how many children the
change carries across the −2 threshold rather than the size of the change
itself.

### QPM Height Increment Distribution

```{r}
#| label: fig-e4-delta-height-distribution
#| fig-cap: "Distribution of QPM-attributable height increment by age group and sex"
#| fig-alt: >
#|   Faceted density plot with one panel per age group in a single row, showing
#|   the QPM-attributable height increment in centimetres on the x-axis and
#|   density on the y-axis. Within each panel two distributions are overlaid for
#|   male and female children. The increment is largest for the youngest children
#|   and its distribution shifts toward smaller values in older age groups,
#|   reflecting the higher growth velocity of younger children.
#| fig-width: 10
#| fig-height: 6

fig_e4_delta_height_distribution |>
  ggplot(aes(x = delta_height_qpm, fill = sex_label)) +
  geom_density(alpha = 0.5, color = NA) +
  facet_wrap(~ age_group_5, nrow = 1) +
  scale_fill_manual(
    values = c("Male" = "#1B6CA8", "Female" = "#3AAF85"),
    name = "Sex"
  ) +
  labs(
    title = "QPM-Attributable Height Increment by Age Group and Sex",
    subtitle = "ENCOVI 2023 — Effect size 9% × sigmoid protein adequacy factor",
    x = "QPM Height Increment (cm)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 13, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    strip.text = element_text(size = 9, face = "bold"),
    legend.position = "bottom"
  )
```

```{r}
#| label: fig-e5-delta-haz-distribution
#| fig-cap: "Distribution of HAZ change attributable to QPM by baseline stunting status"
#| fig-alt: >
#|   Density plot showing the change in height-for-age z-score attributable to
#|   QPM on the x-axis and density on the y-axis, with two overlaid distributions
#|   for children stunted versus not stunted at baseline. A dashed reference line
#|   marks zero change. Both distributions lie to the right of zero and overlap
#|   closely, with the not-stunted group sitting marginally further right. Both
#|   are bimodal, separating children near full protein adequacy from those the
#|   sigmoid factor attenuates.
#| fig-width: 10
#| fig-height: 5

fig_e5_delta_haz_distribution |>
  ggplot(aes(x = delta_haz_qpm, fill = stunting_label)) +
  geom_density(alpha = 0.5, color = NA) +
  geom_vline(xintercept = 0, linetype = "dashed",
             color = "gray50", linewidth = 0.5) +
  scale_fill_manual(
    values = c("Stunted (baseline)" = "#E76F51",
               "Not stunted (baseline)" = "#3AAF85"),
    name = "Baseline Status"
  ) +
  labs(
    title = "Distribution of HAZ Change Attributable to QPM",
    subtitle = "ENCOVI 2023 — Stratified by baseline stunting status",
    x = "Delta HAZ (QPM - Baseline endline)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 13, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 8, color = "azure4"),
    legend.position = "bottom"
  )
```

The two groups receive HAZ changes of similar magnitude, with a mean of
`r round(mean_delta_haz_not_stunted, 3)` among children not stunted at baseline
against `r round(mean_delta_haz_stunted, 3)` among those who are. The ordering
follows the growth velocity reference: a lower baseline HAZ carries a lower
expected 12-month velocity, and the QPM increment is a fixed proportion of that
velocity. The bimodal structure corresponds to the protein adequacy factor,
which separates children near full adequacy from those the sigmoid attenuates.


## Summary

### Key Findings

1. **Growth velocity**: All `r format(n_children_total, big.mark = ",")`
   children receive WHO-based growth velocity assignments interpolated by
   baseline HAZ, age, and sex. Only `r round(pct_below_neg4_overall, 1)`%
   require clamping below -4 SD.

2. **Protein adequacy modulation**: The mean adequacy factor ranges from 0.777
   to 0.895 across age groups (see @tbl-c2-factor-summary), with medians at or
   just below 1.0, so the sigmoid acts on the lower tail.

3. **National impact**: At 100% coverage, modelled stunting prevalence moves
   from `r round(stunting_baseline_pct, 1)`% to
   `r round(stunting_qpm_pct, 1)`% and mean HAZ by
   `r round(haz_delta_mean, 3)` (see @tbl-e2-stunting-baseline-vs-qpm).

4. **Departmental variation**: The reduction ranges from
   `r round(dept_change_min, 1)` to `r round(dept_change_max, 1)` percentage
   points (see @tbl-e3-stunting-by-department), while the mean HAZ change stays
   between 0.128 and 0.167, so the spread comes from how many children each
   department holds near the -2 threshold.

::: {.callout-warning}
## Limitations

1. **Effect size baseline and modulation** — The 9% growth velocity
   increase from Gunaratna et al. (2010) is applied as the baseline
   effect, with individual modulation through each child's protein
   adequacy ratio under the intervention (sigmoid factor described in the
   protein adequacy section above). Children meeting their protein
   requirement take close to the full 9% effect and children far below it take
   an attenuated one. The meta-analysis confidence interval of 6-15% marks
   heterogeneity across populations that this single mechanism does not
   represent.

2. **Cross-sectional to longitudinal** — Each child is projected forward 12
   months with their current characteristics held fixed.

3. **Protein as sole modulator** — Protein adequacy is the only quantity that
   modulates the QPM effect; energy, micronutrients and infection burden do
   not enter the modulation.

4. **Age range extension** — The Gunaratna meta-analysis is calibrated
   primarily on children 6-24 months, and the extension to 25-59 months
   applies the protein quality mechanism across the full 6-59 month range.

5. **Boundary clamping** — The
   `r format(n_below_neg4_total, big.mark = ",")` children below -4 SD take the
   velocity at that boundary, which sets a floor on the increment assigned to
   the most affected children.
:::

### Application

This dataset — containing baseline and QPM stunting profiles for the
analytical cohort — is the primary input for the scenario simulation
pipeline (Module 6), which combines the individual-level QPM impact with
production coverage scenarios to estimate population-level stunting
reduction under different adoption levels.
