---
title: "Prepare SIVESNU Height Data"
subtitle: "Module 5: Meta-analysis"
---

## Why Prepare SIVESNU Height Data?

The Meta-analysis (Gunaratna) pathway requires a model that predicts child
height from socioeconomic and demographic variables. This model will be
trained on SIVESNU 2018 — which contains direct anthropometric measurements —
and applied to children in the ENCOVI 2023 dataset, who lack height data but
have comparable socioeconomic profiles.

The predictor variable set comprises demographic, socioeconomic, and
geographic indicators present in both surveys, constituting the bridge for
cross-dataset prediction. The hypothesis is that the same variables predicting
nutrient intakes ([transfer models](../03_transfer_models/03_01_prepare_modeling_data.qmd))
also predict child stature, since socioeconomic conditions are linked to both
dietary quality and growth outcomes.

This document describes the preparation of the SIVESNU training dataset
through six stages: population filtering, variable harmonization, missing
value imputation, predictor quality assessment, normalization, and feature
selection.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  ggcorrplot,         # Correlation matrix visualization (used by fig-d1)
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file (single source of truth)
source(here("..", "02_code", "00_functions", "00_functions.R"))
```

```{r}
#| label: load-rds
#| include: false

# Web resources for visualizations (exported by framework 05_01)
tbl_a2_population_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_tbl_a2_population_summary.rds"),
  trust = TRUE
)

tbl_b2_merge_quality <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_tbl_b2_merge_quality.rds"),
  trust = TRUE
)

tbl_c1_diagnose_missing <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_tbl_c1_diagnose_missing.rds"),
  trust = TRUE
)

tbl_c2_impute_missing <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_tbl_c2_impute_missing.rds"),
  trust = TRUE
)

fig_d1_correlation_analysis <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_fig_d1_correlation_analysis.rds"),
  trust = TRUE
)

tbl_d2_near_zero_variance <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_tbl_d2_near_zero_variance.rds"),
  trust = TRUE
)

tbl_e1_yj_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_tbl_e1_yj_summary.rds"),
  trust = TRUE
)

tbl_f1_boruta_results <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_tbl_f1_boruta_results.rds"),
  trust = TRUE
)

fig_f1_boruta_importance <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_fig_f1_boruta_importance.rds"),
  trust = TRUE
)

tbl_g1_final_dataset_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_tbl_g1_final_dataset_summary.rds"),
  trust = TRUE
)

fig_g2_height_distributions <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "05_01_fig_g2_height_distributions.rds"),
  trust = TRUE
)

# Boruta model object (required to recompute shadowMax threshold for fig-f1)
boruta_objects <- import(
  here("..", "01_data", "02_processed", "models",
       "05_01_boruta_results.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Final population size (Nivel 1 from tbl-a2)
n_final_population <- tbl_a2_population_summary |>
  filter(Metric == "Final target population") |>
  pull(N)

# Boruta variable counts (Nivel 1 from tbl-f1)
n_boruta_confirmed <- tbl_f1_boruta_results |>
  filter(Metric == "Confirmed (relevant)") |>
  pull(Value) |>
  as.integer()

n_boruta_tentative <- tbl_f1_boruta_results |>
  filter(Metric == "Tentative (borderline)") |>
  pull(Value) |>
  as.integer()

n_boruta_selected <- tbl_f1_boruta_results |>
  filter(Metric == "Variables selected (Confirmed + Tentative)") |>
  pull(Value) |>
  as.integer()

n_boruta_evaluated <- tbl_f1_boruta_results |>
  filter(Metric == "Variables evaluated") |>
  pull(Value) |>
  as.integer()

# Shadow maximum threshold for Boruta importance reference line
# (Recomputed from the Boruta model object — not redundantly exported as RDS)
shadow_max_median <- boruta_objects$boruta_result$ImpHistory |>
  as.data.frame() |>
  pull(shadowMax) |>
  median(na.rm = TRUE)

# Imputation quality: categorical OOB error and the cells it applies to
pfc_value <- tbl_c2_impute_missing |>
  filter(Metric == "PFC (Categorical Variables)") |>
  pull(Value)

n_missing_cells <- sum(tbl_c1_diagnose_missing$n_missing)

# Strongest off-diagonal correlation pair, for the multicollinearity screen
cor_offdiag <- fig_d1_correlation_analysis
diag(cor_offdiag) <- 0
cor_max_cell <- which(abs(cor_offdiag) == max(abs(cor_offdiag)), arr.ind = TRUE)[1, ]
cor_max_value <- round(max(abs(cor_offdiag)), 2)
cor_max_pair <- paste0("`", rownames(cor_offdiag)[cor_max_cell[1]], "` with `",
                       colnames(cor_offdiag)[cor_max_cell[2]], "`")
rm(cor_offdiag, cor_max_cell)
```


## Target Population

SIVESNU 2018 includes anthropometric measurements for children under 5 years.
We filter to the study population (6-59 months) and require valid height and
height-for-age z-score data, resulting in `r format(n_final_population, big.mark = ",")`
children available for model training.

Children under 6 months are excluded as they should be exclusively breastfed
per WHO guidelines, making their anthropometric patterns non-comparable with
older children's dietary intake profiles.

```{r}
#| label: tbl-a2-population-summary
#| tbl-cap: "Target population summary after filtering"

tbl_a2_population_summary |>
  gt() |>
  tab_header(
    title = "Target Population Filtering Summary",
    subtitle = "SIVESNU 2018 children aged 6-59 months with valid anthropometry"
  ) |>
  fmt_number(columns = N, decimals = 0, sep_mark = ",") |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = N) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "Final target population")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Metric == "Final target population")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Population Yield
The filtering pipeline retains `r format(n_final_population, big.mark = ",")`
of 902 children (85.7%). The age restriction (6-59 months) accounts for most of
the attrition, removing 90 children, and a further 39 records lack a height
measurement. Every child with a valid height also carries a valid
height-for-age z-score, so the two criteria select the same records.
:::


## Variable Harmonization

SIVESNU and ENCOVI use different variable naming conventions and coding
schemes. We harmonize SIVESNU variables to match the coding structure used in
ENCOVI 2023, so that the trained height model applies across surveys without
recoding. The harmonization covers geographic classification,
household characteristics, housing materials, basic services, and household
assets.

```{r}
#| label: tbl-b2-merge-quality
#| tbl-cap: "Merge quality assessment"

tbl_b2_merge_quality |>
  gt() |>
  tab_header(
    title = "Child-Household Merge Quality",
    subtitle = "Matching children to their household socioeconomic profiles"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Merge Quality
The `r format(n_final_population, big.mark = ",")` children match to their
household socioeconomic profiles at a rate of 100%. Department and area
classification are close to complete, at 771 and 770 records respectively,
leaving 2 and 3 records without a geographic identifier.
:::


## Missing Value Treatment

### Missing Value Diagnosis

Missing data is minimal across predictor variables. The four variables with
any missing values — `fuente_agua`, `area`, `departamento`, and
`tipo_vivienda` — each affect fewer than 1% of records.

```{r}
#| label: tbl-c1-diagnose-missing
#| tbl-cap: "Missing value summary for predictor variables"

tbl_c1_diagnose_missing |>
  gt() |>
  tab_header(
    title = "Missing Values in Predictor Variables",
    subtitle = "Assessment prior to imputation"
  ) |>
  cols_label(
    variable = "Variable",
    n_missing = "N Missing",
    pct_missing = "% Missing"
  ) |>
  fmt_number(columns = n_missing, decimals = 0, sep_mark = ",") |>
  fmt_number(columns = pct_missing, decimals = 2, pattern = "{x}%") |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = pct_missing, rows = pct_missing > 10)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = pct_missing,
                           rows = pct_missing > 0 & pct_missing <= 10)
  ) |>
  tab_style(
    style = cell_fill(color = "#C8E6C9"),
    locations = cells_body(columns = pct_missing, rows = pct_missing == 0)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Random Forest Imputation

Missing values are imputed via Random Forest using the `impute_rf_parallel()`
utility function, which wraps `missRanger::missRanger()` with a multi-threaded
`ranger` engine. The procedure preserves inter-variable correlations and
handles mixed data types, continuous and categorical, simultaneously.
Imputation quality is summarized via out-of-bag (OOB) error metrics — OOBE for
continuous variables and the proportion of falsely classified entries (PFC) for
categorical variables.

```{r}
#| label: tbl-c2-impute-missing
#| tbl-cap: "Random Forest imputation quality metrics"

tbl_c2_impute_missing |>
  gt() |>
  tab_header(
    title = "Missing Value Imputation Quality Assessment",
    subtitle = "Out-of-bag error estimates from Random Forest imputation (missRanger)"
  ) |>
  cols_label(
    Metric = "Metric",
    Value = "Value",
    Assessment = "Quality Assessment"
  ) |>
  tab_style(
    style = cell_fill(color = "#C8E6C9"),
    locations = cells_body(columns = Assessment,
                           rows = Assessment == "Acceptable")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = Assessment,
                           rows = Assessment == "Review required")
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(columns = Assessment,
                           rows = Assessment == "No continuous variables required imputation")
  ) |>
  tab_footnote(
    footnote = "OOBE: aggregated out-of-bag error across the continuous variables imputed, bounded in [0, 1] and invariant to their scale. Values < 0.5 indicate acceptable imputation quality for continuous variables. N/A when no continuous variables had missing values.",
    locations = cells_body(columns = Metric, rows = 1)
  ) |>
  tab_footnote(
    footnote = "PFC: Proportion of Falsely Classified entries. Values < 0.2 indicate acceptable imputation quality for categorical variables.",
    locations = cells_body(columns = Metric, rows = 2)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Reading the Two Metrics
Out-of-bag (OOB) error is computed from observations randomly excluded during
tree construction, which gives an internal quality metric without a separate
test set.

The continuous metric is reported as N/A: the two continuous predictors
(`miembros_hogar`, `n_cuartos`) had no missing values, so no continuous
imputation was required.

The categorical metric is `r pfc_value`, above the 0.2 reference value, and the
table flags it for review. It is the OOB misclassification rate of the forests
fitted for the four variables that carry missing values, so it measures how
predictable those variables are from the rest of the predictor set, not the
accuracy of the specific cells filled. Those cells number
`r n_missing_cells` across `r format(n_final_population, big.mark = ",")`
records.
:::


## Predictor Quality Assessment

### Correlation Structure

We assess pairwise correlation among predictor variables to identify potential
multicollinearity. Pairs correlating above 0.8 are candidates for removal.

```{r}
#| label: fig-d1-correlation-analysis
#| fig-cap: "Correlation matrix of predictor variables for height modeling"
#| fig-alt: >
#|   Correlation heatmap showing the lower triangle of pairwise correlations
#|   among the predictor variables, with variables hierarchically clustered
#|   along both axes and each cell labelled with its coefficient. Color runs
#|   from orange for negative through white at zero to blue for positive
#|   correlations. Most pairs are weak; housing quality indicators (materials,
#|   assets, services) form a block of moderate positive correlations while
#|   demographic variables stay largely independent. The strongest pair is 0.6,
#|   recoleccion_basura with area, below the 0.8 multicollinearity screen.
#| fig-width: 10
#| fig-height: 10
#| warning: false

fig_d1_correlation_analysis |>
  ggcorrplot(
    hc.order = TRUE,
    colors = c("#E46726", "white", "#6D9EC1"),
    type = "lower",
    lab = TRUE
  )
```

::: {.callout-note}
## Correlation Patterns
The highest pairwise correlation is `r cor_max_value` (`r cor_max_pair`), below
the 0.8 value above which pairs become candidates for removal. Housing quality
indicators cluster together (materials, assets, services), while demographic
variables (`edad`, `sexo`, `parentesco`) stay largely independent of the
socioeconomic indicators. No variable pair is removed on correlation grounds.
:::

### Near-Zero Variance Analysis

Variables with near-zero variance provide minimal discriminating information
for the transfer model and can introduce noise in feature selection.

```{r}
#| label: tbl-d2-near-zero-variance
#| tbl-cap: "Near-zero variance analysis for predictor variables"

tbl_d2_near_zero_variance |>
  gt() |>
  tab_header(
    title = "Near-Zero Variance Analysis",
    subtitle = "Variables with insufficient variation for modeling"
  ) |>
  cols_label(
    variable = "Variable",
    freqRatio = "Frequency Ratio",
    percentUnique = "% Unique Values",
    zeroVar = "Zero Variance",
    nzv = "Near-Zero Variance"
  ) |>
  fmt_number(columns = c(freqRatio, percentUnique), decimals = 2) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = nzv, rows = nzv == TRUE)
  ) |>
  tab_source_note(
    source_note = "Variables flagged as near-zero variance will be removed from the modeling dataset."
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-important}
## Variables Removed
Two variables are flagged as near-zero variance and removed from the modeling
dataset: `tipo_vivienda` (frequency ratio 56.62 — dominated by "Casa formal")
and `telefonia_fija` (frequency ratio 29.92 — nearly all households lack
landlines). The remaining 20 variables proceed to normalization and feature
selection.
:::


## Continuous Variable Normalization

The two continuous predictors (`miembros_hogar` and `n_cuartos`) are normalized
to reduce skewness. `bestNormalize()` evaluates a family of candidate
transformations for each variable and keeps the one that comes closest to
normality; the Transformation Type column reports which one it selected. The
fitted objects are saved and reapplied to ENCOVI data in the
[transfer step](05_03_transfer_height_to_encovi.qmd), so both surveys enter the
model on the same scale.

```{r}
#| label: tbl-e1-yj-summary
#| tbl-cap: "Yeo-Johnson transformation summary"

tbl_e1_yj_summary |>
  gt() |>
  tab_header(
    title = "Yeo-Johnson Transformation Results",
    subtitle = "Normalization of continuous predictor variables"
  ) |>
  fmt_number(
    columns = c(`Pre-transform Skewness`, `Post-transform Skewness`),
    decimals = 3
  ) |>
  cols_align(align = "left", columns = Variable) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Transformation Outcomes
Skewness in `n_cuartos` falls from 1.724 to 0.097 under the transformation.
For `miembros_hogar` the procedure selects a centre-scale normalization, which
leaves the skewness at its original value of 1.278: household size is a count
with a short right tail, and the location-scale shift does not change its
shape.
:::


## Feature Selection

### Boruta Algorithm

The Boruta algorithm evaluates each candidate predictor against random shadow
variables to determine statistical relevance for height prediction. Feature
selection uses raw height in centimetres (`talla_cm`) as the outcome, which is
the variable transferred to ENCOVI, and keeps the selection independent of age:
height-for-age z-scores are age-standardized, and age is itself a predictor.
The implementation uses `Boruta::getImpRfZ` as the importance source, which
delegates the variable importance computation to the multi-threaded `ranger`
backend.

```{r}
#| label: tbl-f1-boruta-results
#| tbl-cap: "Boruta feature selection results for height prediction"

tbl_f1_boruta_results |>
  gt() |>
  tab_header(
    title = "Boruta Feature Selection Results",
    subtitle = "Variable importance assessment for height (talla_cm) prediction"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      rows = Metric == "Variables selected (Confirmed + Tentative)"
    )
  ) |>
  tab_source_note(
    md("**Method:** Boruta with ranger importance source (multi-threaded), 150 iterations against shadow variables")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-f1-boruta-importance
#| fig-cap: "Variable importance from Boruta feature selection for height prediction"
#| fig-alt: >
#|   Horizontal boxplot chart ranking the selected predictor variables by
#|   Boruta importance, with importance on the x-axis and one box per variable
#|   on the y-axis. Boxes are colored by decision: green for confirmed and
#|   yellow for tentative. A dashed vertical line marks the shadow-maximum
#|   threshold that variables must exceed to be retained. Age (edad) dominates
#|   with importance an order of magnitude above all others, followed by a
#|   secondary cluster of geographic and housing quality indicators.
#| fig-height: 8
#| fig-width: 10

ggplot(fig_f1_boruta_importance,
       aes(x = importance, y = variable, fill = decision)) +
  geom_boxplot(alpha = 0.8, outlier.size = 0.5, color = "grey40") +
  geom_vline(
    xintercept = shadow_max_median,
    linetype = "dashed",
    color = "azure4",
    linewidth = 0.8
  ) +
  scale_fill_manual(
    values = c("Confirmed" = "#3AAF85", "Tentative" = "#F9A825")
  ) +
  labs(
    title = "Boruta Variable Importance for Height Prediction",
    subtitle = "Selected variables only. Dashed line = shadow maximum threshold",
    x = "Importance",
    y = NULL,
    fill = "Decision"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    axis.text.y = element_text(size = 9),
    legend.position = "bottom",
    legend.text = element_text(size = 8, color = "azure4"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )
```

::: {.callout-important}
## Feature Selection Results
Of `r n_boruta_evaluated` candidate predictors, Boruta confirms
`r n_boruta_confirmed` as relevant and classifies `r n_boruta_tentative` as
tentative (borderline), yielding `r n_boruta_selected` selected variables. Age
(`edad`) reaches a median importance of 90.8, an order of magnitude above the
next variable, which is consistent with the developmental relationship between
age and height. Geographic (`area`, `departamento`) and housing quality
indicators (`recoleccion_basura`, `fuente_agua`, `material_piso`) form a
secondary cluster in the 5 to 9 range, carrying the socioeconomic gradient
associated with growth beyond what age accounts for.
:::


## Final Modeling Dataset

The selected predictors, outcome variable, and validation references are
consolidated into a single modeling-ready dataset. Three protected
variables — `edad`, `departamento`, and `sexo` — are retained regardless of
their Boruta decision because
[transferring height to ENCOVI](05_03_transfer_height_to_encovi.qmd) requires
them: `edad` and `sexo` for the WHO LMS lookup, and `departamento` for survey
weight recalibration.

```{r}
#| label: tbl-g1-final-dataset-summary
#| tbl-cap: "Final modeling dataset summary"

tbl_g1_final_dataset_summary |>
  gt() |>
  tab_header(
    title = "Final Modeling Dataset Structure",
    subtitle = "Variables retained in the SIVESNU height modeling dataset"
  ) |>
  cols_label(
    Category = "Category",
    Variables = "Variables",
    Count = "N"
  ) |>
  cols_align(align = "left", columns = c(Category, Variables)) |>
  cols_align(align = "right", columns = Count) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Category == "TOTAL")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Category == "TOTAL")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Height Distribution by Stunting Status

The final visualization shows how height is distributed by stunting status
within each age group, which is the variation the transfer model is asked to
reproduce in ENCOVI.

```{r}
#| label: fig-g2-height-distributions
#| fig-cap: "Height distribution by age group and stunting status"
#| fig-alt: >
#|   Faceted density plot with one panel per age group, each showing height in
#|   centimetres on the x-axis. Within every panel two overlaid curves compare
#|   children who are not stunted (green) against those who are stunted with
#|   HAZ below -2 (orange). In each age group the stunted curve sits at shorter
#|   heights than the not-stunted curve, and both shift rightward with older
#|   age groups, with the distance between the two curves widening from the
#|   youngest panel to the oldest.
#| fig-width: 10
#| fig-height: 7

ggplot(fig_g2_height_distributions,
       aes(x = talla_cm, fill = stunting_status)) +
  geom_density(alpha = 0.5, color = NA) +
  facet_wrap(~ age_group, scales = "free") +
  scale_fill_manual(
    values = c("Not stunted" = "#3AAF85",
               "Stunted (HAZ < -2)" = "#E07A5F")
  ) +
  labs(
    title = "Height Distribution by Age Group and Stunting Status",
    subtitle = "SIVESNU 2018 children aged 6-59 months",
    x = "Height (cm)",
    y = "Density",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    strip.text = element_text(face = "bold", color = "azure4"),
    legend.position = "bottom",
    legend.text = element_text(size = 8, color = "azure4")
  )
```

The two curves are separated in every age group, and the distance between their
means grows from 4.4 cm at 6-11 months to 8.4 cm at 48-59 months. This pattern
is consistent with the cumulative character of chronic undernutrition, and it
is the signal the transfer model carries into ENCOVI.


## Summary

### Key Findings

1. **Training dataset**: `r format(n_final_population, big.mark = ",")` SIVESNU
   2018 children aged 6-59 months with valid anthropometric measurements form
   the training population for the height transfer model.

2. **Variable harmonization**: 22 predictor variables are harmonized between
   the SIVESNU and ENCOVI coding schemes, and every child matches to a
   household socioeconomic profile.

3. **Data quality**: `r n_missing_cells` cells are missing across the predictor
   set, each affected variable below 1% of records, and all of them
   categorical. They are filled by Random Forest imputation, whose categorical
   OOB error stands at `r pfc_value`.

4. **Feature selection**: `r n_boruta_selected` of `r n_boruta_evaluated`
   candidate predictors are selected by Boruta. Age leads the importance
   ranking by an order of magnitude, and geographic and housing indicators form
   a secondary cluster.

::: {.callout-warning}
## Limitations

1. **SIVESNU sampling bias** — SIVESNU 2018 uses a health center-based
   sampling frame, which may over-represent children who access public health
   services. The bias is addressed by the weight recalibration against ENSMI
   stunting prevalence in
   [transferring height to ENCOVI](05_03_transfer_height_to_encovi.qmd).

2. **Cross-sectional design** — Height is measured at a single point in time,
   so the transfer model estimates the association between socioeconomic
   profiles and height.

3. **Sample size** — `r format(n_final_population, big.mark = ",")` children
   supports the `r n_boruta_selected` selected predictors, and bounds how much
   interaction structure the model can resolve, particularly in the smaller
   departments.

4. **Categorical imputation error** — The OOB misclassification rate for the
   four imputed variables is `r pfc_value`, above the 0.2 reference value.
:::

### Application

This modeling-ready dataset, along with the saved transformation objects and
the Boruta selection results, feeds the
[height transfer model](05_02_transfer_model_height.qmd), trained on SIVESNU
and then applied to ENCOVI 2023 children in
[transferring height to ENCOVI](05_03_transfer_height_to_encovi.qmd).
