---
title: "Prepare SIVESNU Height Data for Transfer Modeling"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script prepares the SIVESNU 2018 dataset for building a height transfer model that will predict child stature in ENCOVI 2023. 

The predictor variable set comprises demographic, socioeconomic, and geographic variables present in both ENCOVI 2023 and SIVESNU 2018, which constitute the bridge between the two surveys for cross-dataset prediction. Feature selection via Boruta confirms or refines this subset specifically for height prediction.

Main operations:

1. Load SIVESNU 2018 raw data (household, individual, and child tables)
2. Filter to target population (children aged 6-59 months with valid anthropometric data)
3. Merge tables and harmonize variables to match ENCOVI coding structure
4. Impute missing values using Random Forest (missRanger via `impute_rf_parallel()`)
5. Evaluate predictor quality (correlation, near-zero variance)
6. Apply Yeo-Johnson transformations to continuous predictors
7. Run Boruta feature selection with multi-threaded ranger engine using height (talla_cm) as outcome
8. Save modeling-ready dataset and transformation objects

**Target population**: Children aged 6-59 months. Children under 6 months are excluded as they should be exclusively breastfed per WHO guidelines, making their anthropometric patterns non-comparable with older children's dietary intake profiles.

# Setup

## Load Packages

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  janitor,            # Data cleaning utilities
  haven,              # Manage Stata files
  gt,                 # Table formatting and presentation
  bestNormalize,      # Variable normalization
  moments,            # Skewness calculation for YJ assessment
  tidyverse           # General purpose data manipulation
)

# Packages used via :: (not loaded to avoid namespace bloat):
# ggcorrplot  — ggcorrplot() for correlation matrix visualization
# caret       — nearZeroVar() for variable selection
# Boruta      — Boruta(), getImpRfZ(), attStats() for RF-based feature selection
# missRanger  — loaded internally by impute_rf_parallel() in 00_functions.R

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file
source(here("02_code", "00_functions", "00_functions.R"))

# transfer/
ensure_output_dir(here("01_data", "02_processed", "transfer"))

# models/
ensure_output_dir(here("01_data", "02_processed", "models"))

# web_resources/
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

---

# Phase A: Data Loading

## A1. Load SIVESNU Raw Data

Load the three SIVESNU 2018 tables required for this script: household characteristics, individual demographics, and child anthropometric measurements.

```{r}
#| label: a1-load-sivesnu-data

# Load SIVESNU 2018 household data
sivesnu_hogar <- read_parquet(
  here("01_data", "01_raw", "SIVESNU_2018", "sivesnu_gt18_hogar_publico.parquet")
) |>
  clean_names() |>
  zap_label() |>
  zap_labels() |>
  zap_formats()

# Load SIVESNU 2018 individual/members data
sivesnu_individuos <- read_parquet(
  here("01_data", "01_raw", "SIVESNU_2018", "sivesnu_gt18_miembros_publico.parquet")
) |>
  clean_names() |>
  zap_label() |>
  zap_labels() |>
  zap_formats()

# Load SIVESNU 2018 child data (contains anthropometric measurements)
sivesnu_nino <- read_parquet(
  here("01_data", "01_raw", "SIVESNU_2018", "sivesnu_gt18_nino_publico.parquet")
) |>
  clean_names() |>
  zap_label() |>
  zap_labels() |>
  zap_formats()
```

## A2. Extract Child Anthropometric Data

Extract height measurements and age in months from the child table. Filter to the target population (6-59 months) and remove records with missing or implausible anthropometric values.

```{r}
#| label: a2-extract-child-data

# Extract relevant child-level variables
df_children <- sivesnu_nino |>
  transmute(
    # Identifiers
    nboletahogar,
    conglomerado,
    # Survey weight
    pesonino = pesonino,
    # Age in months
    edad_mes = edad_mes,
    # Sex (recode to match ENCOVI labels)
    sexo_nino = case_when(
      sexonino == 1 ~ "Hombre",
      sexonino == 2 ~ "Mujer",
      TRUE ~ NA_character_
    ),
    # Anthropometric measurements
    talla_cm = talla_cm,
    # WHO z-scores (for validation, not used as predictors)
    zlen = zlen,
    hfa = hfa
  ) |>
  # Filter to target population: 6-59 months
  filter(
    edad_mes >= 6,
    edad_mes <= 59
  ) |>
  # Remove records with missing or implausible height
  filter(
    !is.na(talla_cm),
    talla_cm > 40,   # Physiological minimum for 6+ month children
    talla_cm < 130   # Physiological maximum for under-5 children
  ) |>
  # Remove records with missing z-scores
  drop_na(zlen)
```

```{r}
#| label: tbl-a2-population-summary
#| tbl-cap: "Target population summary after filtering"

# Summary of filtered population
tbl_a2_population_summary <- tibble(
  Metric = c(
    "Total children in SIVESNU child table",
    "Children aged 6-59 months",
    "With valid height measurement",
    "With valid ZLEN z-score",
    "Final target population"
  ),
  N = c(
    nrow(sivesnu_nino),
    sivesnu_nino |> filter(edad_mes >= 6, edad_mes <= 59) |> nrow(),
    sivesnu_nino |>
      filter(edad_mes >= 6, edad_mes <= 59, !is.na(talla_cm),
             talla_cm > 40, talla_cm < 130) |>
      nrow(),
    df_children |> nrow(),
    nrow(df_children)
  )
)

export(tbl_a2_population_summary,
            here("01_data", "02_processed", "web_resources",
                 "05_01_tbl_a2_population_summary.rds"))

tbl_a2_population_summary |>
  gt() |>
  tab_header(
    title = "Target Population Filtering Summary",
    subtitle = "SIVESNU 2018 children aged 6-59 months with valid anthropometry"
  ) |>
  fmt_number(columns = N, decimals = 0, sep_mark = ",") |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = N) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "Final target population")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Metric == "Final target population")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_a2_population_summary)
```

---

# Phase B: Variable Harmonization

Merge SIVESNU tables and recode variables to ensure coding compatibility with ENCOVI 2023, which is required for the transfer model to generate predictions across both surveys. The harmonization covers geographic, socioeconomic, housing, service, and asset variables using standardized factor levels.

## B1. Merge and Recode SIVESNU Variables

```{r}
#| label: b1-merge-and-recode

# Merge household and individual tables, then recode to match ENCOVI structure
df_merge <- sivesnu_individuos |>
  left_join(sivesnu_hogar, by = join_by(nboletahogar, conglomerado)) |>
  transmute(
    # --- Identifiers ---
    nboletahogar,
    conglomerado,

    # --- Geographic ---
    departamento = factor(
      departamento,
      levels = c("Guatemala", "", "Sacatepéquez", "Chimaltenango",
                 "Escuintla", "Santa Rosa", "Sololá", "Totonicapán",
                 "Quezaltenango", "Suchitepéquez", "Retalhuleu", "San Marcos",
                 "Huehuetenango", "Quiché", "Baja Verapaz", "Alta Verapaz",
                 "Petén", "Izabal", "Zacapa", "Chiquimula",
                 "Jalapa", "Jutiapa"),
      labels = c("Guatemala", "El Progreso", "Sacatepéquez", "Chimaltenango",
                 "Escuintla", "Santa Rosa", "Sololá", "Totonicapán",
                 "Quetzaltenango", "Suchitepéquez", "Retalhuleu", "San Marcos",
                 "Huehuetenango", "Quiché", "Baja Verapaz", "Alta Verapaz",
                 "Petén", "Izabal", "Zacapa", "Chiquimula",
                 "Jalapa", "Jutiapa"),
      ordered = TRUE
    ),
    area = factor(
      ruralurbano,
      levels = c(1, 2),
      labels = c("Urbana", "Rural")
    ),

    # --- Household characteristics ---
    miembros_hogar = num_miembros_hogar,
    propiedad = factor(
      ifelse(vivienda_propriedad == "", 96, vivienda_propriedad),
      levels = c(1, 2, 3, 4, 96),
      labels = c("Propia y totalmente pagada", "Propia y pagandola a plazos",
                 "Alquilada", "Cedida o prestada", "Otro")
    ),

    # --- Housing characteristics ---
    material_paredes = factor(
      ifelse(vivienda_material_pared == "" |
               vivienda_material_pared == 8 |
               is.na(vivienda_material_pared),
             96, vivienda_material_pared),
      levels = c(6, 3, 2, 5, 4, 1, 7, 96),
      labels = c("Ladrillo", "Block", "Adobe", "Madera",
                 "Lamina metalica", "Bajareque", "Lepa, palo o caña", "Otro")
    ),
    material_techo = factor(
      ifelse(vivienda_material_techo == "" | is.na(vivienda_material_techo),
             96, vivienda_material_techo),
      levels = c(5, 3, 4, 2, 1, 96),
      labels = c("Concreto", "Lamina metalica", "Asbesto cemento",
                 "Teja", "Paja, palma o similar", "Otro")
    ),
    material_piso = factor(
      ifelse(vivienda_material_piso == "" | is.na(vivienda_material_piso), 
             96, vivienda_material_piso),
      levels = c(6, 5, 2, 7, 4, 3, 1, 96), 
      labels = c("Ladrillo ceramico", "Ladrillo de cemento", "Ladrillo de barro", 
                 "Torta de cemento", "Tierra", "Madera", "Tierra", "Tierra")
    ),
    tipo_vivienda = factor(
      ifelse(vivienda_tipo == "", 96, vivienda_tipo),
      levels = c(1, 2, 3, 4, 5, 96),
      labels = c("Casa formal", "Apartamento", "Cuarto en casa de vecindad",
                 "Rancho", "Casa improvisada", "Otro")
    ),
    n_cuartos = vivienda_cuartos,
    tipo_sanitario = factor(
      case_when(
        serv_san_tipo == 1 ~ "Uso exclusivo",
        serv_san_tipo == 2 ~ "Uso compartido",
        serv_san_tipo == 3 ~ "Uso compartido",
        serv_san_tipo == 4 ~ "Uso exclusivo",
        serv_san_tipo == 5 ~ "Uso exclusivo",
        serv_san_tipo == 6 ~ "Uso compartido",
        serv_san_tipo == 96 ~ "Uso compartido",
        TRUE ~ "Uso compartido"
      ),
      levels = c("Uso exclusivo", "Uso compartido")
    ),
    cocina = factor(
      cocina_tiene, 
      levels = c(1, 2), 
      labels = c("De uso exclusivo", "De uso compartido")
    ),

    # --- Basic services ---
    fuente_agua = factor(
      ifelse(agua_fuente %in% c("", 9), 96, agua_fuente),
      levels = c(1, 3, 2, 5, 6, 8, 7, 96),
      labels = c("Tubería dentro de la vivienda",
                 "Tubería fuera de la vivienda",
                 "Chorro publico", "Pozo perforado",
                 "Río, lago, manantial", "Camion cisterna",
                 "Agua de lluvia", "Otro")
    ),
    recoleccion_basura = factor(
      ifelse(basura_deshacer == "", 96, basura_deshacer),
      levels = c(3, 5, 4, 2, 1, 96),
      labels = c("Servicio municipal", "Servicio privado",
                 "Servicio municipal", "La queman o la entierran",
                 "La tiran en cualquier lugar",
                 "La tiran en cualquier lugar")
    ),

    # --- Household assets ---
    electricidad = factor(
      bienes_luz, levels = c("Yes", "No"), labels = c("Si", "No")
    ),
    televisor = factor(
      bienes_television, levels = c("Yes", "No"), labels = c("Si", "No")
    ),
    telefonia_fija = factor(
      bienes_telefonofija, levels = c("Yes", "No"), labels = c("Si", "No")
    ),
    telefonia_celular = factor(
      bienes_celular, levels = c("Yes", "No"), labels = c("Si", "No")
    ),
    computadora = factor(
      bienes_computadora, levels = c("Yes", "No"), labels = c("Si", "No")
    ),

    # --- Individual demographics ---
    parentesco = factor(
      case_when(
        parentesco == 1 ~ "Jefe del hogar",
        parentesco == 2 ~ "Esposo(a) o compañero(a)",
        parentesco %in% c(3, 10) ~ "Hijo(a)",
        parentesco == 4 ~ "Yerno o nuera",
        parentesco == 5 ~ "Nieto(a)",
        parentesco == 6 ~ "Padre o madre",
        parentesco == 7 ~ "Suegro(a)",
        parentesco == 8 ~ "Hermano(a)",
        parentesco == 9 ~ "Otro pariente",
        parentesco == 11 ~ "Empleado de casa particular",
        parentesco == 12 ~ "Otro no pariente",
        TRUE ~ "Otro pariente"
      )
    ),
    sexo = factor(sexo, levels = c(1, 2), labels = c("Hombre", "Mujer")),
    edad = edad,

    # --- Education ---
    grado_estudios = fct_relevel(
      factor(
        case_when(
          grado_estudio %in% c("", "9-No sabe") ~ "Desconocido",
          grado_estudio == "0-Ninguno" ~ "Ninguno",
          grado_estudio == "4-Alfabetización" ~ "Solo sabe leer",
          grado_estudio == "1-Primaria" ~ "Primaria",
          grado_estudio == "2-Secundaria" ~ "Secundaria",
          grado_estudio == "3-Superior/universitaria" ~ "Superior/universitaria"
        ),
        ordered = TRUE
      ),
      c("Desconocido", "Ninguno", "Solo sabe leer",
        "Primaria", "Secundaria", "Superior/universitaria")
    ),
    grado_estudios_num = case_when(
      grado_estudios == "Desconocido" ~ 1,
      grado_estudios == "Ninguno" ~ 1,
      grado_estudios == "Solo sabe leer" ~ 2,
      grado_estudios == "Primaria" ~ 4,
      grado_estudios == "Secundaria" ~ 7,
      grado_estudios == "Superior/universitaria" ~ 10
    ),
    grado_estudios_num = if_else(
      parentesco %in% c("Jefe del hogar", "Esposo(a) o compañero(a)"),
      grado_estudios_num,
      NA
    )
  ) |>
  mutate(
    grado_estudios_hogar = mean(grado_estudios_num, na.rm = TRUE),
    .by = "nboletahogar",
    .keep = "unused"
  ) |>
  zap_label() |>
  zap_labels() |>
  zap_formats()
```

## B2. Join with Child Anthropometric Data

Merge the harmonized household/individual data with child-level anthropometric measurements. Each child is matched to their household's socioeconomic profile.

```{r}
#| label: b2-join-with-children

# Join children with household/individual data
# Children are identified by household + age + sex match
df_joined <- df_children |>
  left_join(
    df_merge,
    by = join_by(nboletahogar, conglomerado),
    relationship = "many-to-many"
  ) |>
  # Resolve many-to-many: keep the match closest in age
  mutate(
    age_diff = abs(edad_mes / 12 - edad)
  ) |>
  slice_min(age_diff, n = 1, by = c(nboletahogar, conglomerado,
                                      edad_mes, sexo_nino)) |>
  select(-age_diff, -edad) |>
  # Use child's sex from the child table (more reliable)
  mutate(
    sexo = factor(sexo_nino, levels = c("Hombre", "Mujer"))
  ) |>
  select(-sexo_nino) |>
  # Convert edad_mes to years for compatibility with transfer model predictors
  mutate(
    edad = edad_mes / 12
  ) |>
  # Remove duplicates if any remain
  distinct()
```

```{r}
#| label: tbl-b2-merge-quality
#| tbl-cap: "Merge quality assessment"

tbl_b2_merge_quality <- tibble(
  Metric = c(
    "Children before merge",
    "Children after merge",
    "Match rate",
    "Children with complete department",
    "Children with complete area"
  ),
  Value = c(
    as.character(nrow(df_children)),
    as.character(nrow(df_joined)),
    paste0(round(nrow(df_joined) / nrow(df_children) * 100, 1), "%"),
    as.character(sum(!is.na(df_joined$departamento))),
    as.character(sum(!is.na(df_joined$area)))
  )
)

export(tbl_b2_merge_quality,
            here("01_data", "02_processed", "web_resources",
                 "05_01_tbl_b2_merge_quality.rds"))

tbl_b2_merge_quality |>
  gt() |>
  tab_header(
    title = "Child-Household Merge Quality",
    subtitle = "Matching children to their household socioeconomic profiles"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_b2_merge_quality)
```

---

# Phase C: Missing Value Treatment

## C1. Diagnose Missing Values

Assess the extent of missing data in predictor variables to determine imputation requirements.

```{r}
#| label: tbl-c1-diagnose-missing
#| tbl-cap: "Missing value summary for predictor variables"

# Calculate missing value statistics for predictors
missing_summary <- df_joined |>
  select(
    parentesco, sexo, edad, departamento, area,
    miembros_hogar, grado_estudios_hogar, propiedad,
    tipo_vivienda, material_paredes, material_techo, material_piso,
    n_cuartos, cocina, tipo_sanitario,
    fuente_agua, recoleccion_basura,
    electricidad, televisor, telefonia_fija, telefonia_celular, computadora
  ) |>
  summarise(
    across(
      everything(),
      list(
        n_missing = \(x) sum(is.na(x)),
        pct_missing = \(x) mean(is.na(x)) * 100
      )
    )
  ) |>
  pivot_longer(
    everything(),
    names_to = c("variable", ".value"),
    names_pattern = "(.+)_(n_missing|pct_missing)"
  ) |>
  arrange(desc(pct_missing))

export(missing_summary,
            here("01_data", "02_processed", "web_resources",
                 "05_01_tbl_c1_diagnose_missing.rds"))

# Format as GT table
missing_summary |>
  gt() |>
  tab_header(
    title = "Missing Values in Predictor Variables",
    subtitle = "Assessment prior to imputation"
  ) |>
  cols_label(
    variable = "Variable",
    n_missing = "N Missing",
    pct_missing = "% Missing"
  ) |>
  fmt_number(columns = n_missing, decimals = 0, sep_mark = ",") |>
  fmt_number(columns = pct_missing, decimals = 2, pattern = "{x}%") |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = pct_missing, rows = pct_missing > 10)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = pct_missing,
                           rows = pct_missing > 0 & pct_missing <= 10)
  ) |>
  tab_style(
    style = cell_fill(color = "#C8E6C9"),
    locations = cells_body(columns = pct_missing, rows = pct_missing == 0)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

rm(missing_summary)
```

## C2. Impute Missing Values

Apply Random Forest imputation using the `impute_rf_parallel()` utility function to handle missing values in predictor variables. The function wraps `missRanger::missRanger()` with a multi-threaded `ranger` engine, preserving inter-variable correlations and handling mixed data types. Output follows the same contract (`ximp`, `OOBerror`) used across preparation scripts.

```{r}
#| label: tbl-c2-impute-missing
#| tbl-cap: "Random Forest imputation quality metrics"

# Separate predictors for imputation
predictors_to_impute <- df_joined |>
  select(
    parentesco, sexo, departamento, area,
    miembros_hogar, grado_estudios_hogar, propiedad,
    tipo_vivienda, material_paredes, material_techo, material_piso,
    n_cuartos, cocina, tipo_sanitario,
    fuente_agua, recoleccion_basura,
    electricidad, televisor, telefonia_fija, telefonia_celular, computadora
  ) |>
  # Convert ordered factors for imputation compatibility
  mutate(
    departamento = factor(departamento, ordered = FALSE)
  ) |>
  # Drop unused factor levels to avoid ranger warnings
  droplevels() |>
  as.data.frame()

# Apply Random Forest imputation via missRanger (multi-threaded ranger engine)
set.seed(789)
imputed_result <- impute_rf_parallel(
  data = predictors_to_impute,
  maxiter = 10,
  ntree = 100,
  verbose = FALSE
)

# Extract imputed data
predictors_imputed <- imputed_result$ximp

# Reconstruct complete dataset
df_imputed <- df_joined |>
  select(nboletahogar, conglomerado, pesonino, edad_mes, edad,
         talla_cm, zlen, hfa, sexo) |>
  bind_cols(
    predictors_imputed |> select(-sexo)
  )

# Imputation quality table
imputation_metrics <- imputed_result$OOBerror

# Report each metric only when a variable of that type required imputation
n_continuous <- imputed_result$n_continuous
n_categorical <- imputed_result$n_categorical

oobe_display <- if (n_continuous > 0) {
  formatC(imputation_metrics["OOBE"], digits = 4, format = "f")
} else {
  "N/A"
}
oobe_assessment <- if (n_continuous > 0) {
  if_else(imputation_metrics["OOBE"] < 0.5, "Acceptable", "Review required")
} else {
  "No continuous variables required imputation"
}

pfc_display <- if (n_categorical > 0) {
  formatC(imputation_metrics["PFC"], digits = 4, format = "f")
} else {
  "N/A"
}
pfc_assessment <- if (n_categorical > 0) {
  if_else(imputation_metrics["PFC"] < 0.2, "Acceptable", "Review required")
} else {
  "No categorical variables required imputation"
}

tbl_c2_impute_missing <- tibble(
  Metric = c(
    "OOBE (Continuous Variables)",
    "PFC (Categorical Variables)",
    "Number of Iterations",
    "Trees per Iteration"
  ),
  Value = c(
    oobe_display,
    pfc_display,
    "10",
    "100"
  ),
  Assessment = c(
    oobe_assessment,
    pfc_assessment,
    "Fixed parameter",
    "Fixed parameter"
  )
)

export(tbl_c2_impute_missing,
            here("01_data", "02_processed", "web_resources",
                 "05_01_tbl_c2_impute_missing.rds"))

tbl_c2_impute_missing |>
  gt() |>
  tab_header(
    title = "Missing Value Imputation Quality Assessment",
    subtitle = "Out-of-bag error estimates from Random Forest imputation (missRanger)"
  ) |>
  cols_label(
    Metric = "Metric",
    Value = "Value",
    Assessment = "Quality Assessment"
  ) |>
  tab_style(
    style = cell_fill(color = "#C8E6C9"),
    locations = cells_body(columns = Assessment,
                           rows = Assessment == "Acceptable")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = Assessment,
                           rows = Assessment == "Review required")
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(columns = Assessment,
                           rows = Assessment == "No continuous variables required imputation")
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(columns = Assessment,
                           rows = Assessment == "No categorical variables required imputation")
  ) |>
  tab_footnote(
    footnote = "OOBE: out-of-bag error across continuous variables, the square root of the variance-weighted mean of per-variable unexplained variance. Values < 0.5 indicate acceptable imputation quality. N/A when no continuous variables had missing values.",
    locations = cells_body(columns = Metric, rows = 1)
  ) |>
  tab_footnote(
    footnote = "PFC: Proportion of Falsely Classified entries. Values < 0.2 indicate acceptable imputation quality for categorical variables. N/A when no categorical variables had missing values.",
    locations = cells_body(columns = Metric, rows = 2)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Clean auxiliary objects
rm(predictors_to_impute, imputed_result, predictors_imputed, imputation_metrics,
   n_continuous, n_categorical,
   oobe_display, oobe_assessment, pfc_display, pfc_assessment,
   tbl_c2_impute_missing)
```

---

# Phase D: Variable Quality Assessment

## D1. Evaluate Predictor Correlation

Assess pairwise correlation among predictor variables to identify potential multicollinearity issues. Variables with correlation exceeding r > 0.8 are candidates for removal.

```{r}
#| label: fig-d1-correlation-analysis
#| fig-cap: "Correlation matrix of predictor variables for height modeling"
#| fig-width: 10
#| fig-height: 10
#| warning: false

# Calculate correlation matrix for all predictors
fig_d1_correlation_analysis <- df_imputed |>
  select(
    parentesco, sexo, edad, departamento, area, miembros_hogar,
    grado_estudios_hogar, propiedad, tipo_vivienda, material_paredes,
    material_techo, material_piso, n_cuartos, cocina, tipo_sanitario,
    fuente_agua, recoleccion_basura, electricidad, televisor,
    telefonia_fija, telefonia_celular, computadora
  ) |>
  mutate(across(everything(), \(x) as.numeric(as.factor(x)))) |>
  cor() |>
  round(1)

export(fig_d1_correlation_analysis,
            here("01_data", "02_processed", "web_resources",
                 "05_01_fig_d1_correlation_analysis.rds"))

fig_d1_correlation_analysis |>
  ggcorrplot::ggcorrplot(
    hc.order = TRUE,
    colors = c("#E46726", "white", "#6D9EC1"),
    type = "lower",
    lab = TRUE
  )

# Cleanup auxiliary objects
rm(fig_d1_correlation_analysis)
```

## D2. Identify Near-Zero Variance Predictors

Evaluate variables with near-zero variance that provide minimal information for modeling.

```{r}
#| label: tbl-d2-near-zero-variance
#| tbl-cap: "Near-zero variance analysis for predictor variables"

# Analyze variance for all predictors
nzv_results <- df_imputed |>
  select(
    parentesco, sexo, edad, departamento, area, miembros_hogar,
    grado_estudios_hogar, propiedad, tipo_vivienda, material_paredes,
    material_techo, material_piso, n_cuartos, cocina, tipo_sanitario,
    fuente_agua, recoleccion_basura, electricidad, televisor,
    telefonia_fija, telefonia_celular, computadora
  ) |>
  caret::nearZeroVar(saveMetrics = TRUE) |>
  rownames_to_column("variable") |>
  arrange(desc(freqRatio))

export(nzv_results,
            here("01_data", "02_processed", "web_resources",
                 "05_01_tbl_d2_near_zero_variance.rds"))

# Display results
nzv_results |>
  gt() |>
  tab_header(
    title = "Near-Zero Variance Analysis",
    subtitle = "Variables with insufficient variation for modeling"
  ) |>
  cols_label(
    variable = "Variable",
    freqRatio = "Frequency Ratio",
    percentUnique = "% Unique Values",
    zeroVar = "Zero Variance",
    nzv = "Near-Zero Variance"
  ) |>
  fmt_number(columns = c(freqRatio, percentUnique), decimals = 2) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = nzv, rows = nzv == TRUE)
  ) |>
  tab_source_note(
    source_note = "Variables flagged as near-zero variance will be removed from the modeling dataset."
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Identify NZV variables to remove
vars_nzv <- nzv_results |>
  filter(nzv == TRUE) |>
  pull(variable)

rm(nzv_results)
```

## D3. Remove Problematic Variables

Remove variables flagged as near-zero variance by the NZV analysis. Variables with insufficient variation provide no discriminating information for the transfer model and can introduce noise in downstream feature selection.

```{r}
#| label: d3-remove-variables

# Remove near-zero variance variables and high-correlation redundancies
df_clean <- df_imputed |>
  select(-any_of(vars_nzv))
```

---

# Phase E: Yeo-Johnson Transformations

Apply Yeo-Johnson transformations to normalize continuous predictors. These transformations must be saved and reapplied identically when preparing ENCOVI data for prediction in script 05_03.

## E1. Transform Continuous Predictors

```{r}
#| label: e1-yeo-johnson-transforms

# Apply Yeo-Johnson transformation to miembros_hogar
yj_miembros_hogar <- bestNormalize(df_clean$miembros_hogar,
                                    allow_orderNorm = FALSE)

# Apply Yeo-Johnson transformation to n_cuartos
yj_n_cuartos <- bestNormalize(df_clean$n_cuartos,
                               allow_orderNorm = FALSE)

# Apply transformations to create final modeling variables
df_transformed <- df_clean |>
  mutate(
    miembros_hogar_original = miembros_hogar,
    n_cuartos_original = n_cuartos,
    miembros_hogar = predict(yj_miembros_hogar),
    n_cuartos = predict(yj_n_cuartos)
  )
```

```{r}
#| label: tbl-e1-yj-summary
#| tbl-cap: "Yeo-Johnson transformation summary"

tbl_e1_yj_summary <- tibble(
  Variable = c("miembros_hogar", "n_cuartos"),
  `Pre-transform Skewness` = c(
    skewness(df_clean$miembros_hogar, na.rm = TRUE),
    skewness(df_clean$n_cuartos, na.rm = TRUE)
  ),
  `Post-transform Skewness` = c(
    skewness(df_transformed$miembros_hogar, na.rm = TRUE),
    skewness(df_transformed$n_cuartos, na.rm = TRUE)
  ),
  `Transformation Type` = c(
    class(yj_miembros_hogar$chosen_transform)[1],
    class(yj_n_cuartos$chosen_transform)[1]
  )
)

export(tbl_e1_yj_summary,
            here("01_data", "02_processed", "web_resources",
                 "05_01_tbl_e1_yj_summary.rds"))

tbl_e1_yj_summary |>
  gt() |>
  tab_header(
    title = "Yeo-Johnson Transformation Results",
    subtitle = "Normalization of continuous predictor variables"
  ) |>
  fmt_number(
    columns = c(`Pre-transform Skewness`, `Post-transform Skewness`),
    decimals = 3
  ) |>
  cols_align(align = "left", columns = Variable) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_e1_yj_summary)
```

---

# Phase F: Feature Selection

Apply Boruta algorithm to identify the predictor variables with confirmed importance for height prediction. The Boruta algorithm evaluates each candidate predictor against random shadow variables to determine statistical relevance for height prediction. The ranger engine is used as importance source, enabling multi-threaded tree construction within each Boruta iteration. With ~20 candidate predictors, the feature space is manageable without dimensionality reduction techniques, preserving full interpretability of the selected variable set.

## F1. Run Boruta Feature Selection

Feature selection uses height in cm (`talla_cm`) as the outcome variable. This is the actual measured variable that will be predicted in ENCOVI, providing a direct assessment of which socioeconomic/demographic variables best predict child stature.

```{r}
#| label: f1-run-boruta

# Prepare data for Boruta
# Exclude identifiers, age_months, z-scores, and original untransformed variables
df_for_boruta <- df_transformed |>
  select(
    # Outcome
    talla_cm,
    # Predictors (exclude NZV-removed and keep only modeling variables)
    any_of(c(
      "parentesco", "sexo", "edad", "departamento", "area",
      "miembros_hogar", "grado_estudios_hogar", "propiedad",
      "tipo_vivienda", "material_paredes", "material_techo",
      "material_piso", "n_cuartos", "cocina", "tipo_sanitario",
      "fuente_agua", "recoleccion_basura",
      "electricidad", "televisor", "telefonia_celular", "computadora"
    ))
  ) |>
  # Convert all to numeric for Boruta
  mutate(across(everything(), \(x) as.numeric(as.factor(x))))

# Run Boruta algorithm with ranger as importance source (multi-threaded)
set.seed(789)
boruta_result <- Boruta::Boruta(
  talla_cm ~ .,
  data = df_for_boruta,
  doTrace = 0,
  maxRuns = 150,
  getImp = Boruta::getImpRfZ,
  num.threads = parallel::detectCores() - 2
)

# Extract decision statistics
boruta_stats <- Boruta::attStats(boruta_result) |>
  rownames_to_column("variable") |>
  mutate(
    decision = case_when(
      decision == "Confirmed" ~ "Confirmed",
      decision == "Tentative" ~ "Tentative",
      decision == "Rejected" ~ "Rejected"
    )
  ) |>
  arrange(desc(medianImp))
```

```{r}
#| label: tbl-f1-boruta-results
#| tbl-cap: "Boruta feature selection results for height prediction"

tbl_f1_boruta_results <- tibble(
  Metric = c(
    "Variables evaluated",
    "Confirmed (relevant)",
    "Tentative (borderline)",
    "Rejected (not relevant)",
    "Variables selected (Confirmed + Tentative)"
  ),
  Value = c(
    as.character(nrow(boruta_stats)),
    as.character(sum(boruta_stats$decision == "Confirmed")),
    as.character(sum(boruta_stats$decision == "Tentative")),
    as.character(sum(boruta_stats$decision == "Rejected")),
    as.character(sum(boruta_stats$decision %in%
                       c("Confirmed", "Tentative")))
  )
)

export(tbl_f1_boruta_results,
            here("01_data", "02_processed", "web_resources",
                 "05_01_tbl_f1_boruta_results.rds"))

tbl_f1_boruta_results |>
  gt() |>
  tab_header(
    title = "Boruta Feature Selection Results",
    subtitle = "Variable importance assessment for height (talla_cm) prediction"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      rows = Metric == "Variables selected (Confirmed + Tentative)"
    )
  ) |>
  tab_source_note(
    md("**Method:** Boruta with ranger importance source (multi-threaded), 150 iterations against shadow variables")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_f1_boruta_results)
```

```{r}
#| label: fig-f1-boruta-importance
#| fig-cap: "Variable importance from Boruta feature selection for height prediction"
#| fig-height: 8
#| fig-width: 10

# Prepare importance data for selected variables
df_boruta_plot <- boruta_result$ImpHistory |>
  as.data.frame() |>
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "importance"
  ) |>
  filter(is.finite(importance), !str_detect(variable, "^shadow")) |>
  left_join(
    boruta_stats |> select(variable, decision, medianImp),
    by = "variable"
  ) |>
  filter(decision %in% c("Confirmed", "Tentative")) |>
  mutate(
    decision = factor(decision, levels = c("Confirmed", "Tentative")),
    variable = fct_reorder(variable, medianImp)
  )

# Shadow max threshold for reference line
shadow_max_median <- boruta_result$ImpHistory |>
  as.data.frame() |>
  pull(shadowMax) |>
  median(na.rm = TRUE)

export(df_boruta_plot,
            here("01_data", "02_processed", "web_resources",
                 "05_01_fig_f1_boruta_importance.rds"))

# Plot with corporate aesthetic
ggplot(df_boruta_plot,
       aes(x = importance, y = variable, fill = decision)) +
  geom_boxplot(alpha = 0.8, outlier.size = 0.5, color = "grey40") +
  geom_vline(
    xintercept = shadow_max_median,
    linetype = "dashed",
    color = "azure4",
    linewidth = 0.8
  ) +
  scale_fill_manual(
    values = c("Confirmed" = "#3AAF85", "Tentative" = "#F9A825")
  ) +
  labs(
    title = "Boruta Variable Importance for Height Prediction",
    subtitle = "Selected variables only. Dashed line = shadow maximum threshold",
    x = "Importance",
    y = NULL,
    fill = "Decision"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    axis.text.y = element_text(size = 9),
    legend.position = "bottom",
    legend.text = element_text(size = 8, color = "azure4"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )

rm(df_boruta_plot, shadow_max_median)
```

## F2. Extract Selected Variables

```{r}
#| label: f2-extract-selected-variables

# Extract selected variables (Confirmed + Tentative)
variables_selected <- boruta_stats |>
  filter(decision %in% c("Confirmed", "Tentative")) |>
  pull(variable)

# Ensure critical variables are included even if Boruta rejects them
# edad and departamento are required for downstream processing
# (WHO LMS lookup and weight recalibration respectively)
variables_protected <- c("edad", "departamento", "sexo")
variables_final <- union(variables_selected, variables_protected)
```

---

# Phase G: Build Final Dataset and Export

## G1. Build Modeling-Ready Dataset

Construct the final dataset containing only the selected predictors, outcome variable, and required identifiers.

```{r}
#| label: g1-build-final-dataset

# Build modeling-ready dataset
df_modeling_ready <- df_transformed |>
  select(
    # Identifiers and metadata
    nboletahogar, conglomerado, pesonino, edad_mes,
    # Outcome
    talla_cm,
    # Validation reference (not used as predictors)
    zlen, hfa,
    # Selected predictors
    all_of(variables_final),
    # Keep original continuous values for reference
    miembros_hogar_original, n_cuartos_original
  )
```

```{r}
#| label: tbl-g1-final-dataset-summary
#| tbl-cap: "Final modeling dataset summary"

tbl_g1_final_dataset_summary <- tibble(
  Category = c(
    "Identifiers & metadata",
    "Outcome variable",
    "Validation references (not predictors)",
    "Selected predictors",
    "Original continuous values (reference)",
    "TOTAL"
  ),
  Variables = c(
    "nboletahogar, conglomerado, pesonino, edad_mes",
    "talla_cm",
    "zlen, hfa",
    paste(variables_final, collapse = ", "),
    "miembros_hogar_original, n_cuartos_original",
    ""
  ),
  Count = c(
    4,
    1,
    2,
    length(variables_final),
    2,
    ncol(df_modeling_ready)
  )
)

export(tbl_g1_final_dataset_summary,
            here("01_data", "02_processed", "web_resources",
                 "05_01_tbl_g1_final_dataset_summary.rds"))

tbl_g1_final_dataset_summary |>
  gt() |>
  tab_header(
    title = "Final Modeling Dataset Structure",
    subtitle = paste0(nrow(df_modeling_ready), " children x ",
                      ncol(df_modeling_ready), " variables")
  ) |>
  cols_label(
    Category = "Category",
    Variables = "Variables",
    Count = "N"
  ) |>
  cols_align(align = "left", columns = c(Category, Variables)) |>
  cols_align(align = "right", columns = Count) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Category == "TOTAL")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Category == "TOTAL")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_g1_final_dataset_summary)
```

## G2. Descriptive Statistics of Height by Key Variables

```{r}
#| label: fig-g2-height-distributions
#| fig-cap: "Height distribution by age group and stunting status"
#| fig-width: 10
#| fig-height: 7

# Create age groups for visualization
fig_g2_height_distributions <- df_modeling_ready |>
  mutate(
    age_group = cut(
      edad_mes,
      breaks = c(5, 11, 23, 35, 47, 59),
      labels = c("6-11m", "12-23m", "24-35m", "36-47m", "48-59m")
    ),
    stunting_status = factor(
      if_else(hfa == 1, "Stunted (HAZ < -2)", "Not stunted"),
      levels = c("Not stunted", "Stunted (HAZ < -2)")
    )
  )

export(fig_g2_height_distributions,
            here("01_data", "02_processed", "web_resources",
                 "05_01_fig_g2_height_distributions.rds"))

fig_g2_height_distributions |>
  ggplot(aes(x = talla_cm, fill = stunting_status)) +
  geom_density(alpha = 0.5, color = NA) +
  facet_wrap(~ age_group, scales = "free") +
  scale_fill_manual(
    values = c("Not stunted" = "#3AAF85",
               "Stunted (HAZ < -2)" = "#E07A5F")
  ) +
  labs(
    title = "Height Distribution by Age Group and Stunting Status",
    subtitle = "SIVESNU 2018 children aged 6-59 months",
    x = "Height (cm)",
    y = "Density",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    strip.text = element_text(face = "bold", color = "azure4"),
    legend.position = "bottom",
    legend.text = element_text(size = 8, color = "azure4")
  )

# Cleanup auxiliary objects
rm(fig_g2_height_distributions)
```

## G3. Save Outputs

```{r}
#| label: g3-save-outputs

# Save modeling-ready dataset
write_parquet(
  df_modeling_ready,
  here("01_data", "02_processed", 
       "transfer", "05_01_sivesnu_height_modeling_data.parquet")
)

# Save Yeo-Johnson transformation objects (for application to ENCOVI in 05_03)
export(
  list(
    yj_miembros_hogar = yj_miembros_hogar,
    yj_n_cuartos = yj_n_cuartos
  ),
  here("01_data", "02_processed", 
       "models", "05_01_yj_transforms.rds")
)

# Save Boruta results and selected variables
export(
  list(
    boruta_result = boruta_result,
    boruta_stats = boruta_stats,
    variables_selected = variables_selected,
    variables_final = variables_final,
    variables_protected = variables_protected
  ),
  here("01_data", "02_processed", 
       "models", "05_01_boruta_results.rds")
)

# Clean auxiliary objects
rm(yj_miembros_hogar, yj_n_cuartos, boruta_result, boruta_stats,
   variables_selected, variables_final, variables_protected,
   df_joined, df_clean, df_transformed, vars_nzv)
```


---

# Notes

## Methodological Decisions

**Variable set**: The predictor variables comprise demographic, socioeconomic, and geographic indicators present in both ENCOVI 2023 and SIVESNU 2018, constituting the bridge between surveys for cross-dataset prediction. Variable coding follows the structure used in the Module 3 transfer models (scripts 03_02 through 03_08), so that the trained height model applies to ENCOVI data without recoding. The premise is that the variables predicting nutrient intake also predict child stature, since socioeconomic conditions are linked to both dietary quality and growth outcomes.

**Boruta on talla_cm**: Feature selection uses raw height in centimeters as the outcome variable. This directly targets the variable that will be transferred to ENCOVI and avoids circularity, since HAZ z-scores depend on age (which is already a predictor). The age variable captures the strong developmental relationship between age and height, while other predictors capture socioeconomic determinants of growth beyond what age explains.

**Variable inclusion strategy**: All candidate predictor variables (including those with high missing rates or expected low variance) enter the pipeline and are evaluated algorithmically. Near-zero variance detection and Boruta feature selection determine which variables are retained, so that inclusion and exclusion decisions are data-driven and auditable.

**Yeo-Johnson transformations**: Applied to `miembros_hogar` and `n_cuartos` to normalize continuous predictors. Transformation objects are saved separately for consistent application to ENCOVI data in script 05_03.

## Imputation Engine

Missing value imputation uses the `impute_rf_parallel()` utility function (defined in `00_functions.R`), which wraps `missRanger::missRanger()` with a multi-threaded `ranger` engine. Output follows the `list(ximp, OOBerror = c(OOBE, PFC))` contract consumed by diagnostic tables across the framework.

Reproducibility note: with `num.threads > 1`, ranger introduces non-deterministic thread scheduling. Imputed values are statistically equivalent across runs but not bit-identical. For strict audit-grade reproducibility, set `num.threads = 1` in the `impute_rf_parallel()` call at the cost of runtime.

## Boruta Configuration

Boruta is executed with `maxRuns = 150` and `getImp = Boruta::getImpRfZ`, which uses `ranger::ranger()` as the importance source. Multi-threaded tree construction is enabled via `num.threads = parallel::detectCores() - 2`.

Reproducibility note: with `num.threads > 1`, Confirmed/Tentative/Rejected decisions for variables with clear importance signals are stable across runs; decisions for variables at the margin of the shadow maximum threshold may oscillate. For strict audit-grade reproducibility, set `num.threads = 1` at the cost of runtime.

## Data Quality

**Population size**: The target population comprises the SIVESNU 2018 children aged 6-59 months with valid anthropometric measurements. Records with missing height values were excluded; the counts are reported in the Phase A extraction table.

**Merge strategy**: Children are matched to household characteristics via household identifiers. The many-to-many join (multiple household members per child record) is resolved by selecting the closest age match, prioritizing the child's own demographic profile.

## Limitations

**SIVESNU sampling**: SIVESNU 2018 uses a health center-based sampling frame, which may introduce selection bias toward children who access public health services. This bias is addressed in script 05_03 through weight recalibration against ENSMI stunting prevalence targets.

**Cross-sectional design**: Height is measured at a single point in time. The transfer model captures the association between socioeconomic profiles and height, not causal effects of specific interventions.
