---
title: "Train Height Transfer Model"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script trains and validates a Generalized Additive Model (GAM) to predict child height (talla_cm) from demographic, socioeconomic, and geographic predictors. The trained model will be applied in script 05_03 to predict height for ENCOVI 2023 children, enabling the meta-analysis pathway to calculate individual-level stunting effects from biofortified maize.

The workflow:

1. Load modeling-ready data from script 05_01 (Boruta-selected predictors and the height outcome)
2. Explore outcome distribution and predictor-outcome relationships
3. Specify GAM structure with smooth terms for continuous and ordered predictors
4. Train the GAM model and evaluate overall fit
5. Perform 10-fold cross-validation for out-of-sample performance estimation
6. Assess departmental prediction quality and age-group accuracy
7. Run comprehensive model diagnostics (residuals, concurvity, partial effects)
8. Save trained model for application to ENCOVI in script 05_03

**Modeling approach**: GAM with Gaussian family, using thin-plate regression splines for continuous predictors (edad, miembros_hogar, n_cuartos, grado_estudios_hogar) and random effects for categorical predictors with many levels (departamento). The smooth terms accommodate the non-linear age-height growth curve, and the Gaussian family is appropriate because height in children follows a normal distribution conditional on age and covariates.

# Setup

## Load Packages

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  mgcv,               # Generalized Additive Models
  gt,                 # Table formatting and presentation
  patchwork,          # Plot composition
  tidyverse           # General purpose data manipulation
)

# Packages used via :: (not loaded to minimize memory footprint)
# rio     — import(), export() for model I/O
# gratia  — draw() for GAM diagnostics
# ggpubr  — ggpubr::ggarrange() for plot composition

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file
source(here("02_code", "00_functions", "00_functions.R"))

# transfer/
ensure_output_dir(here("01_data", "02_processed", "transfer"))

# models/
ensure_output_dir(here("01_data", "02_processed", "models"))

# web_resources/
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

---

# Phase A: Data Loading and Exploration

## A1. Load Modeling Data

Load the modeling-ready dataset produced by script 05_01, containing the children aged 6-59 months with the Boruta-selected predictors and the outcome variable (talla_cm).

```{r}
#| label: a1-load-data

# Load modeling-ready dataset from 05_01
df_raw <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "05_01_sivesnu_height_modeling_data.parquet")
)

# Load Boruta results to confirm selected variables
boruta_results <- import(
  here("01_data", "02_processed", 
       "models", "05_01_boruta_results.rds"),
  trust = TRUE 
)

# Extract the final predictor set from Boruta
predictor_vars <- boruta_results$variables_final
```

```{r}
#| label: tbl-a1-data-overview
#| tbl-cap: "Input data overview"

# Overview initial table
tbl_a1_data_overview <- tibble(
  Item = c(
    "Observations",
    "Boruta-selected predictors",
    "Outcome variable",
    "Age range (months)",
    "Height range (cm)",
    "Stunting prevalence (HAZ < -2)"
  ),
  Value = c(
    format(nrow(df_raw), big.mark = ","),
    as.character(length(predictor_vars)),
    "talla_cm",
    paste0(min(df_raw$edad_mes), " - ", max(df_raw$edad_mes)),
    paste0(round(min(df_raw$talla_cm), 1), " - ", round(max(df_raw$talla_cm), 1)),
    paste0(round(mean(df_raw$hfa == 1) * 100, 1), "%")
  )
)

export(
  tbl_a1_data_overview,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_a1_data_overview.rds")
)

tbl_a1_data_overview |>
  gt() |>
  tab_header(
    title = "Input Data Summary",
    subtitle = "SIVESNU 2018 modeling-ready dataset from script 05_01"
  ) |>
  cols_align(align = "left", columns = Item) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_a1_data_overview)
```

## A2. Predictor Summary

```{r}
#| label: tbl-a2-predictor-types
#| tbl-cap: "Predictor variable types and roles in the GAM"

# Table with the GAM roles
tbl_a2_predictor_types <- tibble(
  Variable = predictor_vars,
  Class = map_chr(predictor_vars, \(v) class(df_raw[[v]])[1]),
  Unique = map_int(predictor_vars, \(v) n_distinct(df_raw[[v]], na.rm = TRUE))
) |>
  mutate(
    GAM_Term = case_when(
      Variable == "edad" ~ "s(edad, k = 10)",
      Variable == "grado_estudios_hogar" ~ "s(grado_estudios_hogar, k = 5)",
      Variable == "departamento" ~ "s(departamento, bs = 're')",
      TRUE ~ paste0("parametric: ", Variable)
    )
  )

export(
  tbl_a2_predictor_types,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_a2_predictor_types.rds")
)

tbl_a2_predictor_types |>
  gt() |>
  tab_header(
    title = "Predictor Variables and GAM Specification",
    subtitle = "14 Boruta-selected predictors with assigned model terms"
  ) |>
  cols_label(
    Variable = "Variable",
    Class = "R Class",
    Unique = "Unique Values",
    GAM_Term = "GAM Term"
  ) |>
  cols_align(align = "left") |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = str_starts(GAM_Term, "s\\("))
  ) |>
  tab_footnote(
    footnote = "Blue rows: smooth (non-linear) terms. White rows: parametric (linear) terms.",
    locations = cells_column_labels(columns = GAM_Term)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_a2_predictor_types)
```

## A3. Outcome Distribution

```{r}
#| label: fig-a3-outcome-distribution
#| fig-cap: "Distribution of height (talla_cm) in the training population"
#| fig-width: 10
#| fig-height: 6

fig_a3_data <- df_raw |>
  select(talla_cm, edad_mes)

export(
  fig_a3_data,
  here("01_data", "02_processed", "web_resources", "05_02_fig_a3_outcome_distribution.rds")
)

# Height distribution
p_hist <- df_raw |>
  ggplot(aes(x = talla_cm)) +
  geom_histogram(
    bins = 40,
    fill = "lightgray",
    color = "azure4",
    alpha = 0.5
  ) +
  labs(
    title = "Height Distribution",
    subtitle = "SIVESNU 2018 children aged 6-59 months (N = 773)",
    x = "Height (cm)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Height vs Age non-linear relation
p_age <- df_raw |>
  ggplot(aes(x = edad_mes, y = talla_cm)) +
  geom_point(alpha = 0.3, color = "#1B6CA8", size = 1) +
  geom_smooth(
    method = "gam",
    formula = y ~ s(x),
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Height vs. Age",
    subtitle = "Non-linear growth curve captured by GAM smooth",
    x = "Age (months)",
    y = "Height (cm)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Show the results
p_hist + p_age

rm(fig_a3_data, p_hist, p_age)
```

---

# Phase B: GAM Model Training

## B1. Prepare Data for Modeling

Prepare the modeling dataframe with appropriate variable types for mgcv. Ordered factors need to be converted to numeric for smooth term estimation. Unordered categorical variables remain as factors for parametric estimation.

```{r}
#| label: b1-prepare-modeling-data

# Build modeling dataframe with correct types for mgcv
# Apply uncount using survey weights (pesonino) to incorporate sampling design
df_model <- df_raw |>
  uncount(weights = pesonino) |>
  mutate(
    # Collapse rare factor levels (n < 10) into existing categories
    # to prevent empty-level issues in cross-validation folds
    material_piso = fct_lump_min(material_piso, 
                                 min = 10, other_level = "Otro"),
    propiedad = fct_lump_min(propiedad, 
                             min = 10, other_level = "Otro"),
    fuente_agua = fct_lump_min(fuente_agua, 
                               min = 10, other_level = "Otro"),
    material_paredes = fct_lump_min(material_paredes, 
                                    min = 10, other_level = "Otro"),
    # Convert ordered factors to numeric for smooth terms
    departamento_num = as.numeric(departamento),
    # Ensure binary/unordered factors remain as factors
    across(
      all_of(predictor_vars[map_lgl(predictor_vars, 
                                    \(v) is.factor(df_raw[[v]]))]),
      \(x) factor(x, ordered = FALSE)
    ),
    # Age group for stratified cross-validation
    age_group = cut(
      edad_mes,
      breaks = c(5, 11, 23, 35, 47, 59),
      labels = c("6-11m", "12-23m", "24-35m", "36-47m", "48-59m")
    )
  ) |>
  # Select only what the model needs
  select(
    # Outcome
    talla_cm,
    # Continuous predictors (for smooth terms)
    edad, grado_estudios_hogar,
    # Departamento as numeric (for random effect smooth)
    departamento_num,
    # Categorical predictors (parametric)
    all_of(predictor_vars[!predictor_vars %in% c("edad",
                                                   "grado_estudios_hogar",
                                                   "departamento")]),
    # Keep identifiers for diagnostics
    edad_mes, zlen, hfa,
    # CV stratification variable
    age_group
  ) |>
  drop_na()
```

```{r}
#| label: tbl-b1-modeling-data-summary
#| tbl-cap: "Modeling data dimensions after preparation"

# Final data dimensions for the model
tbl_b1_modeling_data_summary <- tibble(
  Metric = c("Original observations (pre-uncount)",
             "Expanded observations (post-uncount)",
             "Total columns", "Predictor variables",
             "Complete cases"),
  Value = c(nrow(df_raw),
            nrow(df_model), ncol(df_model), length(predictor_vars),
            sum(complete.cases(df_model)))
)

export(
  tbl_b1_modeling_data_summary,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_b1_modeling_data_summary.rds")
)

tbl_b1_modeling_data_summary |>
  gt() |>
  tab_header(
    title = "Modeling Data Summary",
    subtitle = "After type preparation for mgcv"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_b1_modeling_data_summary)
```

## B2. Specify and Train GAM

The GAM specification uses:

- **Smooth terms** (`s()`) for continuous predictors to capture non-linear relationships, particularly the age-height growth curve which is inherently non-linear in children under 5
- **Random effect smooth** (`bs = "re"`) for departamento to account for geographic variation without overfitting to 22 department-level means
- **Parametric terms** for binary and unordered categorical predictors where non-linear relationships are not expected
- **Gaussian family** because height conditional on covariates is approximately normal
- **REML** (Restricted Maximum Likelihood) for smooth parameter estimation

```{r}
#| label: b2-train-gam

# Define the GAM formula
# Smooth terms for continuous variables; parametric for categorical
gam_formula <- talla_cm ~
  s(edad, k = 10) +
  s(grado_estudios_hogar, k = 5) +
  s(departamento_num, bs = "re") +
  area +
  sexo +
  recoleccion_basura +
  fuente_agua +
  material_piso +
  televisor +
  material_paredes +
  telefonia_celular +
  electricidad +
  computadora +
  propiedad

# Train GAM with REML
gam_height <- gam(
  gam_formula,
  data = df_model,
  family = gaussian(),
  method = "REML"
)
```

## B3. Model Summary

```{r}
#| label: tbl-b3-smooth-terms
#| tbl-cap: "GAM smooth term significance"

# Extract smooth term statistics
smooth_summary <- summary(gam_height)$s.table |>
  as.data.frame() |>
  rownames_to_column("Term") |>
  as_tibble() |>
  rename(
    edf = edf,
    ref_df = Ref.df,
    f_statistic = F,
    p_value = `p-value`
  )

export(
  smooth_summary,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_b3_smooth_terms.rds")
)

smooth_summary |>
  mutate(
    significant = if_else(p_value < 0.05, "*", ""),
    p_value_fmt = case_when(
      p_value < 0.001 ~ "< 0.001",
      TRUE ~ as.character(round(p_value, 4))
    )
  ) |>
  select(Term, edf, ref_df, f_statistic, p_value_fmt, significant) |>
  gt() |>
  tab_header(
    title = "Smooth Term Significance",
    subtitle = "Approximate F-tests for non-linear effects"
  ) |>
  cols_label(
    Term = "Smooth Term",
    edf = "EDF",
    ref_df = "Ref. df",
    f_statistic = "F",
    p_value_fmt = "p-value",
    significant = ""
  ) |>
  fmt_number(columns = c(edf, ref_df, f_statistic), decimals = 2) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = significant == "*")
  ) |>
  tab_footnote(
    footnote = "EDF = effective degrees of freedom. EDF close to 1 indicates near-linear relationship.",
    locations = cells_column_labels(columns = edf)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(smooth_summary)
```

```{r}
#| label: tbl-b3-parametric-terms
#| tbl-cap: "GAM parametric term coefficients"

# Extract parametric coefficients
param_summary <- summary(gam_height)$p.table |>
  as.data.frame() |>
  rownames_to_column("Term") |>
  as_tibble() |>
  rename(
    estimate = Estimate,
    se = `Std. Error`,
    t_value = `t value`,
    p_value = `Pr(>|t|)`
  )

export(
  param_summary,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_b3_parametric_terms.rds")
)

param_summary |>
  mutate(
    significant = if_else(p_value < 0.05, "*", ""),
    p_value_fmt = case_when(
      p_value < 0.001 ~ "< 0.001",
      TRUE ~ as.character(round(p_value, 4))
    )
  ) |>
  select(Term, estimate, se, t_value, p_value_fmt, significant) |>
  gt() |>
  tab_header(
    title = "Parametric Term Coefficients",
    subtitle = "Linear effects of categorical predictors"
  ) |>
  cols_label(
    Term = "Term",
    estimate = "Estimate",
    se = "SE",
    t_value = "t",
    p_value_fmt = "p-value",
    significant = ""
  ) |>
  fmt_number(columns = c(estimate, se, t_value), decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = significant == "*")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(param_summary)
```

```{r}
#| label: tbl-b3-overall-fit
#| tbl-cap: "Overall model fit statistics"

# Show overall model fit
tbl_b3_overall_fit <- tibble(
  Metric = c(
    "Deviance explained",
    "Adjusted R-squared",
    "GCV score",
    "REML score",
    "Scale estimate (residual variance)",
    "N observations"
  ),
  Value = c(
    paste0(round(summary(gam_height)$dev.expl * 100, 1), "%"),
    round(summary(gam_height)$r.sq, 4),
    round(gam_height$gcv.ubre, 2),
    round(gam_height$reml.scale, 2),
    round(gam_height$sig2, 2),
    as.character(nrow(df_model))
  )
)

export(
  tbl_b3_overall_fit,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_b3_overall_fit.rds")
)

tbl_b3_overall_fit |>
  gt() |>
  tab_header(
    title = "Overall Model Fit",
    subtitle = "GAM with Gaussian family and REML estimation"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric %in% c("Deviance explained",
                                                  "Adjusted R-squared"))
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_b3_overall_fit)
```

---

# Phase C: Cross-Validation

## C1. 10-Fold Cross-Validation

Evaluate out-of-sample prediction performance using stratified 10-fold cross-validation. Stratification by age group ensures that each fold contains a representative distribution of the age-height growth curve.

```{r}
#| label: c1-cross-validation

# Assign stratified folds
set.seed(789)
df_cv <- df_model |>
  mutate(
    fold = sample(rep(1:10, length.out = n())),
    .by = age_group
  )

# Run 10-fold CV
cv_results <- map_dfr(1:10, \(k) {
  # Split data
  df_train <- df_cv |> filter(fold != k)
  df_test  <- df_cv |> filter(fold == k)

  # Train GAM on training fold
  gam_fold <- gam(
    gam_formula,
    data = df_train,
    family = gaussian(),
    method = "REML"
  )

  # Predict on test fold
  df_test |>
    mutate(
      predicted = predict(gam_fold, newdata = df_test, type = "response"),
      residual = talla_cm - predicted,
      fold_id = k
    ) |>
    select(fold_id, edad_mes, age_group, talla_cm, predicted, residual,
           zlen, hfa, departamento_num)
})
```

## C2. Cross-Validation Performance

```{r}
#| label: tbl-c2-cv-overall
#| tbl-cap: "10-fold cross-validation performance"

# Calculate overall CV metrics
cv_metrics <- cv_results |>
  summarise(
    RMSE = sqrt(mean(residual^2)),
    MAE = mean(abs(residual)),
    R2 = 1 - sum(residual^2) / sum((talla_cm - mean(talla_cm))^2),
    Mean_Bias = mean(residual),
    SD_Residual = sd(residual),
    Pct_within_5cm = mean(abs(residual) <= 5) * 100,
    Pct_within_3cm = mean(abs(residual) <= 3) * 100
  )

export(
  cv_metrics,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_c2_cv_overall.rds")
)

cv_metrics |>
  pivot_longer(everything(), names_to = "Metric", values_to = "Value") |>
  mutate(
    Metric = case_when(
      Metric == "RMSE" ~ "Root Mean Squared Error (cm)",
      Metric == "MAE" ~ "Mean Absolute Error (cm)",
      Metric == "R2" ~ "R-squared (out-of-sample)",
      Metric == "Mean_Bias" ~ "Mean Bias (cm)",
      Metric == "SD_Residual" ~ "SD of Residuals (cm)",
      Metric == "Pct_within_5cm" ~ "Predictions within +/- 5 cm (%)",
      Metric == "Pct_within_3cm" ~ "Predictions within +/- 3 cm (%)"
    ),
    Value = round(Value, 3)
  ) |>
  gt() |>
  tab_header(
    title = "Cross-Validation Performance Summary",
    subtitle = "10-fold stratified CV (stratified by age group)"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = str_detect(Metric, "R-squared|RMSE"))
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-c2-cv-by-fold
#| tbl-cap: "Performance by CV fold"

tbl_c2_cv_by_fold <- cv_results |>
  summarise(
    N = n(),
    RMSE = sqrt(mean(residual^2)),
    MAE = mean(abs(residual)),
    R2 = 1 - sum(residual^2) / sum((talla_cm - mean(talla_cm))^2),
    Mean_Bias = mean(residual),
    .by = fold_id
  ) |>
  arrange(fold_id)

export(
  tbl_c2_cv_by_fold,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_c2_cv_by_fold.rds")
)

tbl_c2_cv_by_fold |>
  gt() |>
  tab_header(
    title = "Performance by Cross-Validation Fold",
    subtitle = "Stability assessment across folds"
  ) |>
  cols_label(
    fold_id = "Fold",
    N = "N",
    RMSE = "RMSE (cm)",
    MAE = "MAE (cm)",
    R2 = "R²",
    Mean_Bias = "Bias (cm)"
  ) |>
  fmt_number(columns = c(RMSE, MAE, Mean_Bias), decimals = 2) |>
  fmt_number(columns = R2, decimals = 3) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_c2_cv_by_fold)
```

## C3. CV Diagnostic Plots

```{r}
#| label: fig-c3-cv-observed-vs-predicted
#| fig-cap: "Cross-validation: observed vs predicted height"
#| fig-width: 10
#| fig-height: 6

export(
  cv_results,
  here("01_data", "02_processed", "web_resources", "05_02_fig_c3_cv_observed_vs_predicted.rds")
)

# Prediction vs observed
p_scatter <- cv_results |>
  ggplot(aes(x = talla_cm, y = predicted)) +
  geom_point(alpha = 0.3, color = "#1B6CA8", size = 1.2) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "#3AAF85",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  labs(
    title = "Observed vs Predicted Height (CV)",
    subtitle = paste0("RMSE = ", round(cv_metrics$RMSE, 2), " cm, R² = ",
                       round(cv_metrics$R2, 3)),
    x = "Observed Height (cm)",
    y = "Predicted Height (cm)"
  ) +
  coord_equal() +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Residuals
p_residual <- cv_results |>
  ggplot(aes(x = predicted, y = residual)) +
  geom_point(alpha = 0.3, color = "#1B6CA8", size = 1.2) +
  geom_hline(
    yintercept = 0,
    color = "azure4",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  geom_smooth(
    method = "loess",
    se = TRUE,
    color = "#3AAF85",
    linewidth = 0.6,
    alpha = 0.2,
    formula = 'y ~ x'
  ) +
  labs(
    title = "Residuals vs Predicted",
    subtitle = "Loess smooth for systematic bias detection",
    x = "Predicted Height (cm)",
    y = "Residual (cm)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

p_scatter + p_residual

rm(p_scatter, p_residual)
```

---

# Phase D: Performance by Subgroup

## D1. Performance by Age Group

```{r}
#| label: tbl-d1-cv-by-age
#| tbl-cap: "Cross-validation performance by age group"

tbl_d1_cv_by_age <- cv_results |>
  summarise(
    N = n(),
    Mean_Height = mean(talla_cm),
    RMSE = sqrt(mean(residual^2)),
    MAE = mean(abs(residual)),
    R2 = 1 - sum(residual^2) / sum((talla_cm - mean(talla_cm))^2),
    Mean_Bias = mean(residual),
    Pct_within_3cm = mean(abs(residual) <= 3) * 100,
    .by = age_group
  ) |>
  arrange(age_group)

export(
  tbl_d1_cv_by_age,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_d1_cv_by_age.rds")
)

tbl_d1_cv_by_age |>
  gt() |>
  tab_header(
    title = "Prediction Performance by Age Group",
    subtitle = "10-fold CV results stratified by age"
  ) |>
  cols_label(
    age_group = "Age Group",
    N = "N",
    Mean_Height = "Mean Height (cm)",
    RMSE = "RMSE (cm)",
    MAE = "MAE (cm)",
    R2 = "R²",
    Mean_Bias = "Bias (cm)",
    Pct_within_3cm = "Within ±3 cm (%)"
  ) |>
  fmt_number(columns = c(Mean_Height, RMSE, MAE, Mean_Bias), decimals = 2) |>
  fmt_number(columns = R2, decimals = 3) |>
  fmt_number(columns = Pct_within_3cm, decimals = 1) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_d1_cv_by_age)
```

```{r}
#| label: fig-d1-residuals-by-age
#| fig-cap: "Residual distribution by age group"
#| fig-width: 10
#| fig-height: 5

fig_d1_data <- cv_results |>
  select(age_group, residual)

export(
  fig_d1_data,
  here("01_data", "02_processed", "web_resources", "05_02_fig_d1_residuals_by_age.rds")
)

cv_results |>
  ggplot(aes(x = age_group, y = residual, fill = age_group)) +
  geom_boxplot(alpha = 0.5, outlier.alpha = 0.3, outlier.size = 1,
               color = "azure4") +
  geom_hline(
    yintercept = 0,
    color = "#3AAF85",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  scale_fill_manual(
    values = c(
      "6-11m"  = "#B3D9E8",
      "12-23m" = "#80BFD4",
      "24-35m" = "#4DA6C0",
      "36-47m" = "#2688A8",
      "48-59m" = "#1B6CA8"
    )
  ) +
  labs(
    title = "Prediction Residuals by Age Group",
    subtitle = "Boxplots of (observed - predicted) height",
    x = "Age Group",
    y = "Residual (cm)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    legend.position = "none"
  )

# Cleanup auxiliary objects
rm(fig_d1_data)
```

## D2. Performance by Department

```{r}
#| label: tbl-d2-cv-by-department
#| tbl-cap: "Cross-validation performance by department"

# Map department numbers back to names
dept_levels <- levels(df_raw$departamento)

tbl_d2_cv_by_department <- cv_results |>
  mutate(
    departamento = factor(
      departamento_num,
      levels = seq_along(dept_levels),
      labels = dept_levels
    )
  ) |>
  summarise(
    N = n(),
    RMSE = sqrt(mean(residual^2)),
    MAE = mean(abs(residual)),
    Mean_Bias = mean(residual),
    .by = departamento
  ) |>
  arrange(departamento)

export(
  tbl_d2_cv_by_department,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_d2_cv_by_department.rds")
)

tbl_d2_cv_by_department |>
  gt() |>
  tab_header(
    title = "Prediction Performance by Department",
    subtitle = "10-fold CV results by geographic unit"
  ) |>
  cols_label(
    departamento = "Department",
    N = "N",
    RMSE = "RMSE (cm)",
    MAE = "MAE (cm)",
    Mean_Bias = "Bias (cm)"
  ) |>
  fmt_number(columns = c(RMSE, MAE, Mean_Bias), decimals = 2) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = RMSE, rows = RMSE > mean(RMSE) + sd(RMSE))
  ) |>
  tab_footnote(
    footnote = "Red cells: departments with RMSE > mean + 1 SD.",
    locations = cells_column_labels(columns = RMSE)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(dept_levels, tbl_d2_cv_by_department)
```

## D3. Stunting Classification Accuracy

Evaluate how well predicted height translates to stunting classification. This is relevant because the downstream application (script 05_05) uses predicted height to compute HAZ scores.

```{r}
#| label: tbl-d3-stunting-classification
#| tbl-cap: "Stunting classification accuracy from predicted heights (CV)"

# Use observed HAZ classification as reference
tbl_d3_stunting_classification <- cv_results |>
  mutate(
    observed_stunted = hfa == 1,
    # Approximate: children with large negative residuals on already-short
    # children may flip stunting status
    predicted_short = predicted < talla_cm  # just a proxy direction check
  ) |>
  summarise(
    N = n(),
    Observed_Stunting_Pct = mean(observed_stunted) * 100,
    Mean_Residual_Stunted = mean(residual[observed_stunted]),
    Mean_Residual_NonStunted = mean(residual[!observed_stunted]),
    MAE_Stunted = mean(abs(residual[observed_stunted])),
    MAE_NonStunted = mean(abs(residual[!observed_stunted]))
  ) |>
  pivot_longer(everything(), names_to = "Metric", values_to = "Value") |>
  mutate(
    Metric = case_when(
      Metric == "N" ~ "Total observations",
      Metric == "Observed_Stunting_Pct" ~ "Observed stunting prevalence (%)",
      Metric == "Mean_Residual_Stunted" ~ "Mean residual - stunted children (cm)",
      Metric == "Mean_Residual_NonStunted" ~ "Mean residual - non-stunted children (cm)",
      Metric == "MAE_Stunted" ~ "MAE - stunted children (cm)",
      Metric == "MAE_NonStunted" ~ "MAE - non-stunted children (cm)"
    )
  )

export(
  tbl_d3_stunting_classification,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_d3_stunting_classification.rds")
)

tbl_d3_stunting_classification |>
  gt() |>
  tab_header(
    title = "Prediction Quality by Stunting Status",
    subtitle = "Differential accuracy for stunted vs non-stunted children"
  ) |>
  cols_label(Metric = "Metric", Value = "Value") |>
  fmt_number(columns = Value, decimals = 2) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_d3_stunting_classification)
```

---

# Phase E: Model Diagnostics

## E1. Residual Analysis

```{r}
#| label: fig-e1-residual-diagnostics
#| fig-cap: "GAM residual diagnostic plots"
#| fig-width: 10
#| fig-height: 8

# =========================================================================
# Extract model diagnostics data
# =========================================================================
diag_data <- tibble(
  fitted    = fitted(gam_height),
  residuals = residuals(gam_height, type = "deviance"),
  response  = df_model$talla_cm
) |>
  mutate(
    std_residuals  = residuals / sd(residuals),
    sqrt_abs_std   = sqrt(abs(std_residuals)),
    theoretical_q  = qqnorm(std_residuals, plot.it = FALSE)$x
  )

export(
  diag_data,
  here("01_data", "02_processed", "web_resources", "05_02_fig_e1_residual_diagnostics.rds")
)

# =========================================================================
# Plot 1: Residuals vs Fitted (Linearity)
# =========================================================================
plot_resid_fitted <- diag_data |>
  ggplot(aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.3, size = 1, color = "#1B6CA8") +
  geom_smooth(
    formula = 'y ~ x',
    se = FALSE,
    method = "loess",
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Residuals vs Fitted",
    subtitle = "Reference line should be flat and horizontal",
    x = "Fitted Values",
    y = "Deviance Residuals"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# =========================================================================
# Plot 2: Q-Q Plot of Residuals (Normality)
# =========================================================================
qq_obj <- qqnorm(diag_data$std_residuals, plot.it = FALSE)

plot_qq <- tibble(
  theoretical = qq_obj$x,
  sample      = qq_obj$y
) |>
  ggplot(aes(x = theoretical, y = sample)) +
  geom_point(alpha = 0.3, size = 1, color = "#1B6CA8") +
  geom_abline(
    intercept = 0,
    slope = 1,
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Q-Q Plot of Residuals",
    subtitle = "Points should follow the diagonal reference line",
    x = "Theoretical Quantiles",
    y = "Standardized Residuals"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# =========================================================================
# Plot 3: Scale-Location (Homoscedasticity)
# =========================================================================
plot_scale_loc <- diag_data |>
  ggplot(aes(x = fitted, y = sqrt_abs_std)) +
  geom_point(alpha = 0.3, size = 1, color = "#1B6CA8") +
  geom_smooth(
    formula = 'y ~ x',
    se = FALSE,
    method = "loess",
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Homogeneity of Variance",
    subtitle = "Reference line should be flat and horizontal",
    x = "Fitted Values",
    y = expression(sqrt("Standardized Residuals"))
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# =========================================================================
# Plot 4: Histogram of Residuals (Distribution)
# =========================================================================
plot_hist_resid <- diag_data |>
  ggplot(aes(x = std_residuals)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 40,
    fill = "lightgray",
    color = "azure4",
    alpha = 0.5
  ) +
  stat_function(
    fun = dnorm,
    args = list(mean = 0, sd = 1),
    color = "#3AAF85",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    title = "Distribution of Residuals",
    subtitle = "Dashed line = standard normal reference",
    x = "Standardized Residuals",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# =========================================================================
# Combine into 2x2 grid
# =========================================================================
ggpubr::ggarrange(
  plot_resid_fitted,
  plot_qq,
  plot_scale_loc,
  plot_hist_resid,
  ncol = 2, nrow = 2
)

rm(diag_data, plot_resid_fitted, plot_qq, plot_scale_loc, plot_hist_resid, qq_obj)
```

## E2. Smooth Term Visualization

```{r}
#| label: fig-e2-smooth-effects
#| fig-cap: "Partial effects of smooth terms"
#| fig-width: 10
#| fig-height: 8

fig_e2_data <- gratia::smooth_estimates(gam_height)

export(
  fig_e2_data,
  here("01_data", "02_processed", "web_resources", "05_02_fig_e2_smooth_effects.rds")
)

gratia::draw(gam_height, residuals = TRUE, rug = TRUE) &
  theme_minimal() &
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 9),
    plot.caption = element_text(color = "azure4", size = 8.5),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    strip.text = element_text(color = "azure4", size = 10)
  ) &
  scale_colour_manual(values = c("#1B6CA8"), guide = "none") &
  scale_fill_manual(values = c("#1B6CA8"), guide = "none")

# Cleanup auxiliary objects
rm(fig_e2_data)
```

## E3. Concurvity Assessment

Concurvity is the GAM analogue of multicollinearity. High concurvity between smooth terms can inflate standard errors and destabilize the model.

```{r}
#| label: tbl-e3-concurvity
#| tbl-cap: "Concurvity assessment for smooth terms"

tbl_e3_concurvity <- concurvity(gam_height, full = FALSE)$estimate |>
  as.data.frame() |>
  rownames_to_column("Term") |>
  as_tibble() |>
  pivot_longer(-Term, names_to = "Against", values_to = "Concurvity") |>
  filter(Term != Against, Concurvity > 0.3) |>
  arrange(desc(Concurvity))

export(
  tbl_e3_concurvity,
  here("01_data", "02_processed", "web_resources", "05_02_tbl_e3_concurvity.rds")
)

tbl_e3_concurvity |>
  gt() |>
  tab_header(
    title = "Pairwise Concurvity (values > 0.3)",
    subtitle = "GAM analogue of multicollinearity between smooth terms"
  ) |>
  cols_label(
    Term = "Term",
    Against = "Against",
    Concurvity = "Concurvity"
  ) |>
  fmt_number(columns = Concurvity, decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = Concurvity, rows = Concurvity > 0.8)
  ) |>
  tab_footnote(
    footnote = "Values > 0.8 (red) indicate high concurvity requiring attention.",
    locations = cells_column_labels(columns = Concurvity)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(tbl_e3_concurvity)
```

## E4. Observed vs Fitted (Full Model)

```{r}
#| label: fig-e4-observed-vs-fitted
#| fig-cap: "Observed vs fitted height (full model, not CV)"
#| fig-width: 7
#| fig-height: 6

fig_e4_data <- df_model |>
  mutate(
    fitted = fitted(gam_height),
    residual = residuals(gam_height)
  ) |>
  select(talla_cm, fitted, residual)

export(
  fig_e4_data,
  here("01_data", "02_processed", "web_resources", "05_02_fig_e4_observed_vs_fitted.rds")
)

fig_e4_data |>
  ggplot(aes(x = talla_cm, y = fitted)) +
  geom_point(alpha = 0.3, color = "#1B6CA8", size = 1.2) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "#3AAF85",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  labs(
    title = "Observed vs Fitted Height (Full Model)",
    subtitle = paste0(
      "Adj. R² = ", round(summary(gam_height)$r.sq, 3),
      ", Deviance explained = ",
      round(summary(gam_height)$dev.expl * 100, 1), "%"
    ),
    x = "Observed Height (cm)",
    y = "Fitted Height (cm)"
  ) +
  coord_equal() +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4")
  )

# Cleanup auxiliary objects
rm(fig_e4_data)
```

---

# Phase F: Export

## F1. Train Final Model on Full Dataset

The final model is trained on the complete dataset for maximum predictive power when applied to ENCOVI in script 05_03. Cross-validation performance metrics from Phase C provide the unbiased estimate of out-of-sample accuracy.

```{r}
#| label: f1-final-model

# The model trained in Phase B (gam_height) already uses the full dataset.
# No retraining needed — we export this model directly.

# Store CV metrics alongside the model for reference
cv_summary <- list(
  rmse = cv_metrics$RMSE,
  mae = cv_metrics$MAE,
  r2 = cv_metrics$R2,
  mean_bias = cv_metrics$Mean_Bias,
  pct_within_5cm = cv_metrics$Pct_within_5cm,
  pct_within_3cm = cv_metrics$Pct_within_3cm
)
```

## F2. Save Model and Diagnostics

```{r}
#| label: f2-save-outputs

# Save trained GAM model
export(
  list(
    gam_height = gam_height,
    cv_summary = cv_summary,
    predictor_vars = predictor_vars,
    gam_formula = gam_formula
  ),
  here("01_data", "02_processed", 
       "models", "05_02_height_transfer_model.rds")
)

# Save CV predictions for downstream validation
write_parquet(
  cv_results,
  here("01_data", "02_processed", 
       "transfer", "05_02_cv_predictions.parquet")
)
```

---

# Notes

## Methodological Decisions

**GAM specification**: The smooth terms of the Generalized Additive Model accommodate the non-linear age-height growth curve. The Gaussian family is appropriate because child height conditional on age and covariates is approximately normally distributed.

**Survey weight integration**: Sampling weights (pesonino) are incorporated via `uncount()` to expand the sampled children into weighted observations. This allows using `mgcv::gam()` directly on the expanded frame, while preserving the representativeness of the SIVESNU sampling structure.

**Smooth term structure**: Age uses a thin-plate regression spline (k = 10) to accommodate the curvilinear growth trajectory in children 6-59 months. Department is modeled as a random effect smooth (`bs = "re"`) to borrow strength across departments with few observations and provide shrinkage estimation. REML is used for smooth parameter selection following Wood (2011).

**Stratified CV**: Cross-validation folds were stratified by age group to ensure each fold contains a representative sample of the growth curve, preventing fold-specific estimates dominated by a narrow age range.

## Model Performance

Deviance explained and adjusted R² on the full dataset are reported in the model summary (B3). Out-of-sample performance — RMSE, MAE, R² and the share of predictions within ±3 cm and ±5 cm — is reported in the cross-validation summary (C2), together with the mean bias.

The effective degrees of freedom of the age smooth are consistent with the non-linear growth curve. The grado_estudios_hogar smooth shows a mildly non-linear relationship, and the department random effect is close to linear, so geographic heterogeneity beyond the parametric predictors is limited. Pairwise concurvity is reported in the concurvity assessment (E3).

Prediction error increases with age (D1), consistent with greater height variability in older children. Stunted and non-stunted children show comparable mean absolute error (D3).

## Limitations

**Training sample size**: The training set is modest for a GAM with this number of predictors and smooth terms. The effective degrees of freedom remain below their k limits, indicating adequate flexibility without overfitting.

**Cross-survey transfer**: The model assumes that the height-socioeconomic profile relationship in SIVESNU 2018 (health center-based sample) generalizes to ENCOVI 2023 (household survey). Systematic sampling frame differences may introduce bias, addressed in script 05_03 through calibration procedures.


