---
title: "Generating Synthetic Population"
subtitle: "Module 4: Covariate Prediction"
description-meta: "Donating dietary profiles from ENCOVI to SIVESNU children through a bridge model with regional candidate cells and departmental anchoring."
---

## Overview

Scenario simulation needs a population that carries three things at once: the
demographic and geographic structure of Guatemala, the dietary record of each
child, and an anthropometric outcome to move. No single survey carries all
three.

ENCOVI 2023 carries the first two. It is nationally representative, it holds the
household and geographic attributes that scenario assignment works on, and it
records food consumption in enough detail to derive each child's nutrient
intake. What it does not record is child height.

SIVESNU 2018 records height. It provides height-for-age z-scores for children in
the same 6–59 month range, but its sample is small and its departmental coverage
uneven, so it cannot itself serve as the population of a departmental
simulation.

This module builds the simulation population from the ENCOVI side and supplies
the missing outcome by **statistical matching** against SIVESNU:

1. **Receptor cohort** — Every record is an ENCOVI child aged 6–59 months with a
   complete dietary record, keeping its household, department and design weight.
2. **Bridge model** — A model of the height-for-age z-score estimated on the
   variables both surveys collect, which places donors and receptors on a single
   comparable metric.
3. **Donation** — Each receptor is matched to its nearest donors on that metric,
   within its region and age band, and receives the anthropometric block of one
   selected donor.
4. **Departmental anchoring** — Donor selection is tilted, department by
   department, so that weighted stunting prevalence approaches the Encuesta
   Nacional de Salud Materno Infantil (ENSMI) 2014–2015 reference
   [@MSPAS2017_ENSMI].
5. **Nutrient profile** — Supplementation is added and the derived nutrient
   variables are recomputed from the child's own ENCOVI intake.
6. **Weight calibration** — Design weights are calibrated to that same
   departmental reference, holding each department's weighted population fixed.

The resulting dataset is the input to the biofortification scenarios.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  scales,             # Number formatting
  glue,               # String interpolation for table subtitles
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script (exported by
# framework 04_07)

# --- Phase B: Cohort construction and harmonization ---
tbl_b3_receptor_cohort <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_b3_receptor_cohort.rds"),
  trust = TRUE
)

tbl_b4_donor_pool <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_b4_donor_pool.rds"),
  trust = TRUE
)

tbl_b4_donated_coverage <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_b4_donated_coverage.rds"),
  trust = TRUE
)

tbl_b5_harmonization <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_b5_harmonization.rds"),
  trust = TRUE
)

# --- Phase C: Bridge model and donation ---
tbl_c1_bridge_model <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_c1_bridge_model.rds"),
  trust = TRUE
)

fig_c1_smooth_terms <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_fig_c1_smooth_terms.rds"),
  trust = TRUE
)

# --- Phase D: Donation diagnostics ---
tbl_d1_donor_usage <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_d1_donor_usage.rds"),
  trust = TRUE
)

tbl_d2_departmental_tilt <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_d2_departmental_tilt.rds"),
  trust = TRUE
)

tbl_d3_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_d3_distribution.rds"),
  trust = TRUE
)

fig_d3_zscore_density <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_fig_d3_zscore_density.rds"),
  trust = TRUE
)

# --- Phase E: Nutrient profile ---
tbl_e1_chispitas <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_e1_chispitas.rds"),
  trust = TRUE
)

tbl_e2_nutritional_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_e2_nutritional_summary.rds"),
  trust = TRUE
)

# --- Phase F: Block PCA transformations ---
tbl_f_pca_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_f_pca_summary.rds"),
  trust = TRUE
)

# --- Phase G: Weight calibration ---
tbl_h_calibration_diagnostics <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_h_calibration_diagnostics.rds"),
  trust = TRUE
)

tbl_h_calibration_results <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_h_calibration_results.rds"),
  trust = TRUE
)

tbl_h_distribution_calibrated <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_h_distribution_calibrated.rds"),
  trust = TRUE
)

fig_h_density_calibrated <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_fig_h_density_calibrated.rds"),
  trust = TRUE
)

# --- Phase H: Export ---
tbl_i_output_inventory <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_07_tbl_i_output_inventory.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Figures used in the narrative text

n_receptors <- sum(tbl_b3_receptor_cohort$n_children)
n_households <- sum(tbl_b3_receptor_cohort$n_households)
n_weighted <- sum(tbl_b3_receptor_cohort$n_weighted)

n_donors <- tbl_c1_bridge_model |>
  filter(Metric == "Donors") |>
  pull(Value) |>
  parse_number()

# Supplementation adherence, as used by the framework. The factor is a scalar
# carried on every row of the contribution table.
adherence_national <- tbl_e1_chispitas |>
  pull(adherence_national) |>
  first()

bridge_deviance <- tbl_c1_bridge_model |>
  filter(Metric == "Deviance explained") |>
  pull(Value)

bridge_correlation <- tbl_c1_bridge_model |>
  filter(Metric == "Fitted-observed correlation") |>
  pull(Value)

# Departmental tilt: residual departure from the reference before calibration
mean_gap_pp <- round(mean(abs(tbl_d2_departmental_tilt$gap_pp)), 2)
max_gap_pp <- round(max(abs(tbl_d2_departmental_tilt$gap_pp)), 2)
n_within_1se <- sum(abs(tbl_d2_departmental_tilt$gap_se) <= 1)
n_departments <- nrow(tbl_d2_departmental_tilt)
n_dept_boundary <- sum(tbl_d2_departmental_tilt$boundary)

# Calibration: residual departure after calibration
max_gap_calibrated <- round(max(abs(tbl_h_calibration_results$gap_pp)), 2)

calib_ratio_range <- tbl_h_calibration_diagnostics |>
  filter(Metric == "Adjustment ratio range") |>
  pull(Value)

calib_covariate_shift <- tbl_h_calibration_diagnostics |>
  filter(Metric == "Maximum covariate shift") |>
  pull(Value)

n_output_variables <- sum(tbl_i_output_inventory$N_variables)
```


## Building the Two Cohorts

### Receptor Cohort

The receptors are the children the simulation acts on: ENCOVI 2023 children aged
6–59 months with a complete dietary record. Age in months is the value estimated
in
[Estimate Age in Months](../01_nutritional_intakes/01_02_estimate_age_in_months.qmd)
from the National Statistics Institute (INE) birth registry distributions.

Each receptor keeps its household identifier, its department and its design
weight, so the departmental composition of the synthetic population and the
joint structure of household, geographic and dietary attributes are those of the
survey.

One attribute is completed at this point. Household education is missing for a
small number of records and is required later, both as a bridge variable and as
a predictor of the nutrient imputation in
[Baseline and QPM Profiles](04_08_precompute_baseline_and_qpm_profiles.qmd).
Missing values are filled with the weighted median of the child's own
department and area, which keeps the completion inside the cell the variable is
used to distinguish.

```{r}
#| label: tbl-b3-receptor-cohort
#| tbl-cap: "Receptor cohort by department"

tbl_b3_receptor_cohort |>
  gt() |>
  tab_header(
    title = "Receptor Cohort",
    subtitle = "ENCOVI 2023 children 6-59 months with a complete dietary record"
  ) |>
  cols_label(
    region = "Region",
    departamento = "Department",
    n_children = "Children",
    n_households = "Households",
    n_weighted = "Weighted population"
  ) |>
  fmt_number(
    columns = c(n_children, n_households, n_weighted),
    decimals = 0,
    use_seps = TRUE
  ) |>
  cols_align(align = "left", columns = c(region, departamento)) |>
  cols_align(align = "right", columns = c(n_children, n_households, n_weighted)) |>
  grand_summary_rows(
    fns = list(Total = ~ sum(.)),
    columns = c(n_children, n_households, n_weighted),
    fmt = ~ fmt_number(., decimals = 0, use_seps = TRUE)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Donor Pool

The donors are the SIVESNU 2018 children carrying a valid height-for-age
z-score, together with the variables the matching needs and the anthropometric
block that is transferred. Donors are organised by region and age band, which
are the cells the matching operates within.

```{r}
#| label: tbl-b4-donor-pool
#| tbl-cap: "Donor pool by region and age band"

tbl_b4_donor_pool |>
  gt() |>
  tab_header(
    title = "Donor Pool",
    subtitle = "SIVESNU 2018 children with a valid height-for-age z-score, by donation cell"
  ) |>
  cols_label(region = "Region") |>
  tab_spanner(label = "Age band (months)", columns = -region) |>
  cols_align(align = "left", columns = region) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Asymmetry Between the Two Cohorts
There are `r format(n_receptors, big.mark = ",")` receptors and
`r format(n_donors, big.mark = ",")` donors, so donors are used by more than one
receptor. Donor usage and the reach of the candidate sets are reported in the
diagnostics below.
:::

```{r}
#| label: tbl-b4-donated-coverage
#| tbl-cap: "Availability of the donated variable block"

tbl_b4_donated_coverage |>
  gt() |>
  tab_header(
    title = "Donated Block Coverage",
    subtitle = "Variables transferred from the donor record"
  ) |>
  cols_label(
    Category = "Category",
    n_variables = "Variables",
    n_available = "Available in donor pool"
  ) |>
  fmt_number(columns = c(n_variables, n_available), decimals = 0) |>
  cols_align(align = "left", columns = Category) |>
  cols_align(align = "right", columns = c(n_variables, n_available)) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Harmonizing the Bridge Variables

The matching only works if a variable means the same thing on both sides. The
bridge variables are therefore brought to a common encoding: factor levels are
aligned to a single ordered set, levels observed in fewer than 15 donors are
collapsed into a residual category, and the resulting encoding is applied
identically to donors and receptors.

```{r}
#| label: tbl-b5-harmonization
#| tbl-cap: "Bridge variable harmonization"

tbl_b5_harmonization |>
  gt() |>
  tab_header(
    title = "Bridge Variable Harmonization",
    subtitle = "Level alignment applied identically to donors and receptors"
  ) |>
  cols_label(
    Variable = "Variable",
    Levels_retained = "Levels retained",
    Levels_collapsed = "Levels collapsed",
    Pct_rare_receptors = "% rare (receptors)",
    Pct_missing_receptors = "% missing (receptors)",
    Assigned_level = "Assigned level",
    In_bridge_model = "In bridge model"
  ) |>
  fmt_number(
    columns = c(Pct_rare_receptors, Pct_missing_receptors),
    decimals = 1
  ) |>
  cols_align(align = "left", columns = c(Variable, Assigned_level)) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Dietary Position
The nutritional index enters the bridge model as the child's **survey-weighted
percentile** within the intake distribution of its own survey. A percentile is
free of the scale of the underlying measurements, which is what lets the dietary
dimension inform the matching across two sources whose absolute intake levels
are not directly comparable.
:::


## The Bridge Model

The bridge model relates the height-for-age z-score to the variables both
surveys collect. Its purpose is not prediction: it defines a **one-dimensional
metric** on which donors and receptors are comparable. Two children with the
same predicted value share the same expected anthropometric position given
everything both surveys observe.

Age and the two continuous covariates enter through penalized smooth terms; the
harmonized factors enter parametrically.

```{r}
#| label: tbl-c1-bridge-model
#| tbl-cap: "Bridge model fit"

tbl_c1_bridge_model |>
  gt() |>
  tab_header(
    title = "Bridge Model",
    subtitle = "Height-for-age z-score on the variables shared by both surveys"
  ) |>
  cols_label(Metric = "Metric", Value = "Value") |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-c1-smooth-terms
#| fig-cap: "Estimated smooth terms of the bridge model"
#| fig-alt: >
#|   Faceted line chart with three panels, one per smooth term of the bridge
#|   model, in this order: age in months, household education level, and
#|   nutritional index percentile. Each panel plots the value of the covariate
#|   on the x-axis against its estimated contribution to the height-for-age
#|   z-score on the y-axis, with a dashed horizontal line at zero. Only the age
#|   curve departs clearly from a straight line, with a minimum around 27
#|   months; the education and index curves are close to linear. All three
#|   contributions sit below the zero line across the whole range.
#| fig-width: 10
#| fig-height: 4

fig_c1_smooth_terms |>
  ggplot(aes(x = x_value, y = fit)) +
  geom_hline(yintercept = 0, color = "grey70", linetype = "dashed",
             linewidth = 0.4) +
  geom_line(color = "#1B6CA8", linewidth = 1) +
  facet_wrap(~ term, scales = "free_x") +
  labs(
    title = "Bridge Model Smooth Terms",
    subtitle = "Estimated contribution to the height-for-age z-score",
    x = NULL,
    y = "Contribution to z-score"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    strip.text = element_text(face = "bold", color = "azure4", size = 9),
    panel.grid.minor = element_blank()
  )
```

::: {.callout-important}
## What the Bridge Model Carries
The model accounts for `r bridge_deviance` of the deviance in the donor pool,
with a fitted-observed correlation of `r bridge_correlation`. That is the share
of anthropometric position the shared variables locate; the remainder travels
across as the selected donor's residual, described next.
:::


## Donation

### Candidate Sets and Residual Transfer

Each receptor is matched to the five donors closest to it on the bridge metric,
within its donation cell of region and age band. Cells holding fewer donors than
the candidate set draw from the full pool, so every receptor is matched.

The transferred z-score is the receptor's **own predicted value plus the
residual of the selected donor**. The predicted component follows the receptor's
characteristics; the residual component reproduces the empirical dispersion of
the outcome. Together they retain the variance and shape observed in the donor
pool without assuming a parametric error form.

::: {.callout-note}
## Why Cells Are Regional
Candidate donors are drawn within region and age band rather than within
department. Departmental information reaches the population through the
receptor's own attributes and through the departmental anchoring described
next, so the donor pool is asked only for the conditional structure its sample
size supports.
:::

### Departmental Anchoring

Selection within each candidate set is governed by a departmental parameter that
shifts the probability of drawing a donor whose transferred z-score falls below
the stunting threshold. The expected weighted prevalence is monotone in that
parameter, so the value reproducing the ENSMI reference is unique and is found
by root finding.

The magnitude of the parameter is itself informative: it measures how far the
departmental level sits from what the shared variables predict on their own.

```{r}
#| label: tbl-d2-departmental-tilt
#| tbl-cap: "Departmental tilt and achieved prevalence"

tbl_d2_departmental_tilt |>
  gt() |>
  tab_header(
    title = "Departmental Anchoring",
    subtitle = "Tilt parameter and achieved prevalence against the ENSMI 2014-2015 reference"
  ) |>
  cols_label(
    departamento = "Department",
    n_children = "Children",
    prevalence_achieved = "Achieved",
    se_expected = "SE (pp)",
    stunting_rate_official = "ENSMI reference",
    lambda = "Tilt",
    boundary = "At boundary",
    gap_pp = "Gap (pp)",
    gap_se = "Gap (SE)"
  ) |>
  fmt_number(columns = n_children, decimals = 0, use_seps = TRUE) |>
  fmt_percent(
    columns = c(prevalence_achieved, stunting_rate_official),
    decimals = 1
  ) |>
  fmt_number(columns = c(se_expected, lambda, gap_pp, gap_se), decimals = 2) |>
  cols_align(align = "left", columns = departamento) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(columns = boundary, rows = boundary == TRUE)
  ) |>
  tab_footnote(
    footnote = "Tilt is the departmental parameter governing donor selection, searched within the range -8 to 8. A value at the boundary would mean the reference could not be reached within that range.",
    locations = cells_column_labels(columns = lambda)
  ) |>
  tab_footnote(
    footnote = "Gap (SE) expresses the departure from the reference in units of its own standard error.",
    locations = cells_column_labels(columns = gap_se)
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Anchoring Result
After the tilt, departmental prevalence departs from the ENSMI reference by
`r mean_gap_pp` percentage points on average and by `r max_gap_pp` at the widest
departure, with `r n_within_1se` of the `r n_departments` departments inside one
standard error. No department reaches the boundary of the admissible tilt range.

The tilt sets prevalence in expectation, over the random draw of donors. The
weight calibration below fixes it on the realised sample.
:::

### Donation Diagnostics

Because the donor pool is smaller than the receptor cohort, how often each donor
is reused and how far the candidate sets reach are properties worth reporting.

```{r}
#| label: tbl-d1-donor-usage
#| tbl-cap: "Donor usage and matching distance by region"

tbl_d1_donor_usage |>
  gt() |>
  tab_header(
    title = "Donor Usage",
    subtitle = "Reuse and matching distance within each region"
  ) |>
  cols_label(
    region = "Region",
    n_receptors = "Receptors",
    n_donors_used = "Donors used",
    max_reuse = "Max reuse",
    median_reuse = "Median reuse",
    top1_pct_weighted = "Top donor (% weighted)",
    mean_distance = "Mean distance",
    pct_own_cell = "% from own cell"
  ) |>
  fmt_number(
    columns = c(n_receptors, n_donors_used, max_reuse),
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(columns = c(median_reuse, mean_distance), decimals = 2) |>
  fmt_number(columns = c(top1_pct_weighted, pct_own_cell), decimals = 1) |>
  cols_align(align = "left", columns = region) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-d3-distribution
#| tbl-cap: "Transferred z-score distribution against the donor pool"

tbl_d3_distribution |>
  gt() |>
  tab_header(
    title = "Transferred Distribution",
    subtitle = "Survey-weighted on the scale of each source"
  ) |>
  cols_label(Statistic = "Statistic") |>
  fmt_number(columns = -Statistic, decimals = 2) |>
  cols_align(align = "left", columns = Statistic) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-d3-zscore-density
#| fig-cap: "Transferred height-for-age z-score distribution"
#| fig-alt: >
#|   Density curve of the transferred height-for-age z-score across the synthetic
#|   population, with the z-score on the x-axis and density on the y-axis,
#|   weighted by the design weight. A dashed vertical line marks the stunting
#|   threshold at minus 2. The distribution is smooth, with a main mode near
#|   minus 1.8 and an additional shoulder around minus 2.7, and a substantial
#|   share of its mass to the left of the threshold.
#| fig-width: 9
#| fig-height: 5

fig_d3_zscore_density |>
  ggplot(aes(x = zlen_predicted, weight = factor)) +
  geom_density(fill = "#1B6CA8", alpha = 0.35, color = "#1B6CA8") +
  geom_vline(xintercept = -2, linetype = "dashed", color = "#E53935",
             linewidth = 0.8) +
  labs(
    title = "Transferred Height-for-Age Z-Score",
    subtitle = "Synthetic population weighted by design weight. Dashed line = stunting threshold",
    x = "Height-for-age z-score",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    panel.grid.minor = element_blank()
  )
```


## Nutrient Profile

### Supplementation Contribution

The micronutrient powder programme delivers iron and zinc through daily sachets.
Its expected contribution is added to the non-maize component of each child's
intake, scaled by the departmental coverage published by INE and by a national
adherence factor of `r round(adherence_national, 3)`. That factor is the
survey-weighted mean of the adherence estimated in
[Chispitas Supplementation](04_02_integrate_chispitas_supplementation.qmd),
taken over the SIVESNU 2018 children recorded as programme recipients, which is
the full set of recipients in that survey rather than the subset selected as
donors.

```{r}
#| label: tbl-e1-chispitas
#| tbl-cap: "Expected supplementation contribution by department"

tbl_e1_chispitas |>
  select(-adherence_national) |>
  gt() |>
  tab_header(
    title = "Chispitas Contribution",
    subtitle = glue(
      "Expected daily addition to non-maize iron and zinc | ",
      "Adherence among recipients: {round(adherence_national, 3)}"
    )
  ) |>
  cols_label(
    departamento = "Department",
    chispitas_coverage_official = "Coverage",
    delta_fe_mg = "Iron (mg/day)",
    delta_zn_mg = "Zinc (mg/day)"
  ) |>
  fmt_percent(columns = chispitas_coverage_official, decimals = 1) |>
  fmt_number(columns = c(delta_fe_mg, delta_zn_mg), decimals = 3) |>
  cols_align(align = "left", columns = departamento) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Derived Variables

Total intake, absorbed fractions and the composite nutritional index are
computed from the source-specific components of the child's own ENCOVI record.
Iron absorption uses the fixed fraction adopted for predominantly plant-based
diets and zinc absorption follows the Miller saturation model with its age term,
as documented in
[Bioavailability Adjustments](04_03_calculate_bioavailability_adjustments.qmd).

The derivation runs twice:

- The **whole-diet pass** uses maize and non-maize together and produces the
  descriptive nutrient profile of the synthetic population.
- The **non-maize pass** uses only the non-maize sources and produces the index
  at which the GAM effect model is evaluated. It applies the Miller model to the
  non-maize zinc pool alone and uses its own normalization parameters.

The absorption constants are identical in both passes, so the axis the effect
model was estimated on and the axis it is evaluated on are constructed the same
way.

```{r}
#| label: tbl-e2-nutritional-summary
#| tbl-cap: "Nutrient profile of the synthetic population"

tbl_e2_nutritional_summary |>
  gt() |>
  tab_header(
    title = "Nutrient Profile",
    subtitle = "Synthetic population after supplementation and derivation"
  ) |>
  cols_label(
    Variable = "Variable",
    Mean = "Mean",
    P25 = "P25",
    Median = "Median",
    P75 = "P75"
  ) |>
  fmt_number(columns = c(Mean, P25, Median, P75), decimals = 2) |>
  cols_align(align = "left", columns = Variable) |>
  cols_align(align = "right", columns = c(Mean, P25, Median, P75)) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Block Components

The block components the response model references are recomputed from the
donated variables using the transformations fitted in
[Stunting Modeling Data](04_05_prepare_stunting_modeling_data.qmd).

```{r}
#| label: tbl-f-pca-summary
#| tbl-cap: "Block PCA components carried into the synthetic population"

tbl_f_pca_summary |>
  gt() |>
  tab_header(
    title = "Block PCA Components",
    subtitle = "Transformations applied to the donated variables"
  ) |>
  cols_label(
    Block = "Block",
    Description = "Description",
    Input_variables = "Input variables",
    Components = "Components"
  ) |>
  fmt_number(columns = c(Input_variables, Components), decimals = 0) |>
  cols_align(align = "left", columns = c(Block, Description)) |>
  cols_align(align = "right", columns = c(Input_variables, Components)) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Weight Calibration

Design weights are calibrated so that weighted departmental prevalence
reproduces the ENSMI reference, holding each department's weighted population
total fixed.

```{r}
#| label: tbl-h-calibration-diagnostics
#| tbl-cap: "Weight calibration diagnostics"

tbl_h_calibration_diagnostics |>
  gt() |>
  tab_header(
    title = "Calibration Diagnostics",
    subtitle = "Effect of the calibration on the weight distribution"
  ) |>
  cols_label(Metric = "Metric", Value = "Value") |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Scale of the Adjustment
Individual weights move within a ratio range of `r calib_ratio_range`, and the
largest shift induced in any covariate mean is `r calib_covariate_shift`. The
calibration reaches the departmental targets without displacing the composition
of the population it acts on.
:::

```{r}
#| label: tbl-h-calibration-results
#| tbl-cap: "Departmental prevalence after calibration"

tbl_h_calibration_results |>
  gt() |>
  tab_header(
    title = "Calibration Results",
    subtitle = "Weighted prevalence by department against the ENSMI reference"
  ) |>
  cols_label(
    departamento = "Department",
    n_children = "Children",
    n_weighted = "Weighted population",
    prevalence = "Prevalence",
    mean_zscore = "Mean z-score",
    stunting_rate_official = "ENSMI reference",
    gap_pp = "Gap (pp)"
  ) |>
  fmt_number(columns = c(n_children, n_weighted), decimals = 0, use_seps = TRUE) |>
  fmt_percent(columns = c(prevalence, stunting_rate_official), decimals = 1) |>
  fmt_number(columns = c(mean_zscore, gap_pp), decimals = 2) |>
  cols_align(align = "left", columns = departamento) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-h-distribution-calibrated
#| tbl-cap: "Z-score distribution under design and calibrated weights"

tbl_h_distribution_calibrated |>
  gt() |>
  tab_header(
    title = "Distributional Effect of the Calibration",
    subtitle = "Donor pool, design weights and calibrated weights"
  ) |>
  cols_label(Statistic = "Statistic") |>
  fmt_number(columns = -Statistic, decimals = 2) |>
  cols_align(align = "left", columns = Statistic) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-h-density-calibrated
#| fig-cap: "Z-score density under design and calibrated weights"
#| fig-alt: >
#|   Two overlaid density curves of the height-for-age z-score across the
#|   synthetic population, one weighted by the design weight and one by the
#|   calibrated weight, with the z-score on the x-axis and density on the y-axis.
#|   A dashed vertical line marks the stunting threshold at minus 2. The two
#|   curves track each other closely across the range.
#| fig-width: 9
#| fig-height: 5

fig_h_density_calibrated |>
  pivot_longer(
    cols = c(factor, weight_calibrated),
    names_to = "weight_type",
    values_to = "weight_value"
  ) |>
  mutate(
    weight_type = recode(
      weight_type,
      factor = "Design weights",
      weight_calibrated = "Calibrated weights"
    )
  ) |>
  ggplot(aes(x = zlen_predicted, weight = weight_value, fill = weight_type,
             color = weight_type)) +
  geom_density(alpha = 0.3) +
  geom_vline(xintercept = -2, linetype = "dashed", color = "#E53935",
             linewidth = 0.8) +
  scale_fill_manual(values = c("Design weights" = "#1B6CA8",
                               "Calibrated weights" = "#3AAF85")) +
  scale_color_manual(values = c("Design weights" = "#1B6CA8",
                                "Calibrated weights" = "#3AAF85")) +
  labs(
    title = "Effect of Calibration on the Z-Score Distribution",
    subtitle = "Dashed line = stunting threshold (HAZ = -2)",
    x = "Height-for-age z-score",
    y = "Density",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "azure4", size = 14),
    plot.subtitle = element_text(color = "azure4", size = 10),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, colour = "azure4", hjust = 0),
    axis.text = element_text(size = 8, colour = "azure4"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```


## Output

```{r}
#| label: tbl-i-output-inventory
#| tbl-cap: "Variable inventory of the synthetic population"

tbl_i_output_inventory |>
  gt() |>
  tab_header(
    title = "Output Inventory",
    subtitle = "Variables carried by the synthetic population dataset"
  ) |>
  cols_label(
    Category = "Category",
    N_variables = "Variables",
    Source = "Source"
  ) |>
  fmt_number(columns = N_variables, decimals = 0) |>
  cols_align(align = "left", columns = c(Category, Source)) |>
  cols_align(align = "right", columns = N_variables) |>
  grand_summary_rows(
    fns = list(Total = ~ sum(.)),
    columns = N_variables,
    fmt = ~ fmt_number(., decimals = 0)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

## Summary

### Key Findings

1. **Bridge model coverage**: The variables shared by both surveys account for
   `r bridge_deviance` of the deviance in the donor pool, with a fitted-observed
   correlation of `r bridge_correlation`. The rest of each child's anthropometric
   position travels across as the selected donor's residual.

2. **Departmental anchoring**: After the tilt, weighted departmental prevalence
   departs from the ENSMI reference by `r mean_gap_pp` percentage points on
   average, with `r n_within_1se` of `r n_departments` departments inside one
   standard error and none at the boundary of the admissible range.

3. **Calibration closes the residual gap**: After weight calibration the
   departmental gap is `r max_gap_calibrated` percentage points at its widest,
   with adjustment ratios inside `r calib_ratio_range` and a largest covariate
   mean shift of `r calib_covariate_shift`.

4. **Output**: `r format(n_receptors, big.mark = ",")` children across
   `r format(n_households, big.mark = ",")` households, representing
   `r format(round(n_weighted), big.mark = ",", scientific = FALSE)` children
   nationally, carrying `r n_output_variables` variables.

::: {.callout-warning}
## Limitations

1. **Conditional independence** — The donation assumes that, given the shared
   variables, the anthropometric block is independent of the variables observed
   only in ENCOVI. This assumption is not verifiable from the data.

2. **Donor reuse** — The donor pool is smaller than the receptor cohort, so
   donors are used by more than one receptor. Usage and mean matching distance
   are reported by region.

3. **Reference period** — The departmental reference predates the receptor
   survey. The departmental anchoring absorbs changes in level occurring between
   the two collection periods.
:::

### Application

This population is the input to
[Baseline and QPM Profiles](04_08_precompute_baseline_and_qpm_profiles.qmd),
which computes each child's baseline and biofortified nutrient profile, and
through it to the
[scenario simulation](../06_scenarios/06_01_calculate_market_baseline.qmd).
