---
title: "Integrating Micronutrient Supplementation"
subtitle: "Module 4: Covariate Prediction"
---

## Overview

Guatemala's Chispitas program distributes Multiple Micronutrient Powders (MNP) to
children 6–59 months through the public health system. These sachets provide
iron, zinc, and other micronutrients that complement dietary intake — a critical
intervention in a population where maize-based diets often fall short of
requirements.

To accurately model the relationship between nutrient intake and stunting risk,
we must account for supplementation alongside dietary intake. This requires
understanding not just program coverage, but actual consumption patterns: Did
the household receive Chispitas? Did caregivers administer them? How much of
each sachet was consumed?

This module integrates Chispitas contributions through a **three-level adherence
cascade**:

1. **Access** — Did the household obtain Chispitas?
2. **Administration** — Did caregivers give sachets to the child?
3. **Dosing compliance** — What proportion of each sachet was consumed?

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  scales,             # Number formatting for visualizations
  ggsankey,           # Sankey diagram visualizations
  cowplot,            # Combine multiple plots
  ggpubr,             # Extract legends from ggplot
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script (exported by framework 04_02).

tbl_c1_access_ever <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_02_tbl_c1_access_ever.rds"),
  trust = TRUE
)

tbl_c2_access_6months <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_02_tbl_c2_access_6months.rds"),
  trust = TRUE
)

tbl_c3_administration <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_02_tbl_c3_administration.rds"),
  trust = TRUE
)

tbl_c4_dosing_compliance <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_02_tbl_c4_dosing_compliance.rds"),
  trust = TRUE
)

fig_c5_sankey_adherence <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_02_fig_c5_sankey_adherence.rds"),
  trust = TRUE
)

tbl_d2_chispitas_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_02_tbl_d2_chispitas_summary.rds"),
  trust = TRUE
)

fig_e_distribution_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_02_fig_e_distribution_comparison.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Cascade retention rates (from c1-c4 weighted distributions)
pct_access_ever <- tbl_c1_access_ever |>
  filter(access_ever == "Yes, ever obtained") |>
  pull(percentage)

pct_access_6months <- tbl_c2_access_6months |>
  filter(access_6months == "Access in 6 months") |>
  pull(percentage)

pct_administration <- tbl_c3_administration |>
  filter(administration == "Gave part or all of the sachets") |>
  pull(percentage)

# Dosing compliance breakdown (% by category among administered)
pct_dosing_all <- tbl_c4_dosing_compliance |>
  filter(dosing_compliance == "All") |>
  pull(percentage)

pct_dosing_more_half <- tbl_c4_dosing_compliance |>
  filter(dosing_compliance == "More than half") |>
  pull(percentage)

pct_dosing_half <- tbl_c4_dosing_compliance |>
  filter(dosing_compliance == "Half") |>
  pull(percentage)

pct_dosing_less_half <- tbl_c4_dosing_compliance |>
  filter(dosing_compliance == "Less than half") |>
  pull(percentage)

# Adherence factors from summary
mean_adherence_recipients <- tbl_d2_chispitas_summary |>
  pull(mean_adherence_recipients)

mean_adherence_all <- tbl_d2_chispitas_summary |>
  pull(mean_adherence_all)

pct_recipients <- tbl_d2_chispitas_summary |>
  pull(pct_recipients)

# Administration gap = 100 - pct_administration (among those with 6m access)
pct_admin_gap <- 100 - pct_administration
```


## Chispitas Program Context

### Nutritional Composition

Each Chispitas sachet contains 15 micronutrients following WHO/UNICEF MNP
specifications. For this analysis, we focus on iron and zinc — the micronutrients
most relevant to stunting and directly comparable to biofortified maize
improvements [@INCAP2016_Chispitas].

| Micronutrient | Content per Sachet |
|---------------|-------------------|
| Iron (ferrous fumarate) | 10 mg |
| Zinc (zinc gluconate) | 4.1 mg |

### Distribution Protocol

The program distributes 60 sachets every ~180 days, intended for daily consumption
during 60 consecutive days. To convert this episodic pattern to average daily
intake (compatible with 24-hour recall methodology), we apply a 1/3 adjustment:

- **Effective daily Fe**: 10 mg × 1/3 = 3.33 mg/day
- **Effective daily Zn**: 4.1 mg × 1/3 = 1.37 mg/day


## Population and Variable Preparation

### Target Population

Chispitas are indicated for children 6–59 months — old enough for complementary
feeding (where sachets are mixed into food) but within the critical window for
micronutrient intervention. Children under 6 months are excluded as exclusive
breastfeeding is assumed during this period.

### Adherence Variables

SIVESNU 2018 [@INCAP2020_SIVESNU] captures supplementation behavior through four 
key variables:

| Variable | Question | Cascade Level |
|----------|----------|---------------|
| `chispitas_vez` | Ever obtained Chispitas? | Level 1: Access |
| `chispitas_6mes` | Obtained in last 6 months? | Level 1b: Recent access |
| `chispitas_todo_parte` | Gave sachets to child? | Level 2: Administration |
| `chispitas_sobresdelpaquete` | How much of sachet consumed? | Level 3: Dosing |

::: {.callout-note}
## Variable Interpretation
For `chispitas_6mes`, NA values indicate **no barrier reported** (access
maintained), not missing data. Only explicit "no" responses indicate access
interruption. This interpretation follows the official SIVESNU 2018 report
coding (INCAP, 2020).
:::


## Adherence Cascade Analysis

The cascade quantifies attrition at each stage from program access to actual
consumption of micronutrient powder. Each level filters the population, and the
final effective coverage is the one that remains after cumulative dropout.

### Level 1: Lifetime Access

```{r}
#| label: tbl-c1-access-ever
#| tbl-cap: "Ever Obtained Chispitas"

tbl_c1_access_ever |>
  gt() |>
  tab_header(
    title = "Ever Obtained Chispitas",
    subtitle = "Variable: chispitas_vez | Lifetime exposure to supplementation program"
  ) |>
  cols_label(
    access_ever = "Response",
    n_survey = "N (survey)",
    percentage = "Percentage"
  ) |>
  fmt_number(columns = n_survey, decimals = 0) |>
  fmt_number(columns = percentage, decimals = 1, pattern = "{x}%") |>
  tab_source_note(
    source_note = "Source: SIVESNU 2018 | Note: Survey-weighted estimates using expanded sample."
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

### Level 1b: Recent Access (Last 6 Months)

```{r}
#| label: tbl-c2-access-6months
#| tbl-cap: "Barriers to Access in Last 6 Months"

tbl_c2_access_6months |>
  gt() |>
  tab_header(
    title = "Access Status in Last 6 Months",
    subtitle = "Variable: chispitas_6mes | Among households that ever obtained Chispitas"
  ) |>
  cols_label(
    access_6months = "Access Status",
    n_survey = "N (survey)",
    percentage = "Percentage"
  ) |>
  fmt_number(columns = n_survey, decimals = 0) |>
  fmt_number(columns = percentage, decimals = 1, pattern = "{x}%") |>
  tab_source_note(
    source_note = "Source: SIVESNU 2018 | Note: NA values interpreted as 'access maintained' per official SIVESNU report coding."
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

### Level 2: Administration

```{r}
#| label: tbl-c3-administration
#| tbl-cap: "Dosing Compliance Status"

tbl_c3_administration |>
  gt() |>
  tab_header(
    title = "Dosing Compliance Status",
    subtitle = "Variable: chispitas_todo_parte | Among households with Chispitas access"
  ) |>
  cols_label(
    administration = "Dosing Behavior",
    n_survey = "N (survey)",
    percentage = "Percentage"
  ) |>
  fmt_number(columns = n_survey, decimals = 0) |>
  fmt_number(columns = percentage, decimals = 1, pattern = "{x}%") |>
  tab_source_note(
    source_note = "Source: SIVESNU 2018 | Note: Analysis restricted to households with confirmed Chispitas access."
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

### Level 3: Dosing Compliance

```{r}
#| label: tbl-c4-dosing-compliance
#| tbl-cap: "Sachet Consumption Quantity"

tbl_c4_dosing_compliance |>
  gt() |>
  tab_header(
    title = "Sachet Consumption Quantity",
    subtitle = "Variable: chispitas_sobresdelpaquete | Among households that administered Chispitas"
  ) |>
  cols_label(
    dosing_compliance = "Quantity Consumed",
    n_survey = "N (survey)",
    percentage = "Percentage"
  ) |>
  fmt_number(columns = n_survey, decimals = 0) |>
  fmt_number(columns = percentage, decimals = 1, pattern = "{x}%") |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = dosing_compliance == "All")
  ) |>
  tab_source_note(
    source_note = "Source: SIVESNU 2018 | Note: Full compliance (highlighted) indicates child received 100% of intended micronutrient dose."
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

### Cascade Visualization

```{r}
#| label: fig-c5-sankey-adherence
#| fig-cap: "Adherence Journey: From Access to Compliance Patterns"
#| fig-alt: >
#|   Sankey diagram tracing households across four left-to-right stages of the
#|   Chispitas adherence cascade: lifetime access, access in the last 6 months,
#|   administration to the child, and final adherence level. Nodes at each stage
#|   are labelled (for example "Ever obtained", "Access in 6 months", "Gave part
#|   or all", "Full adherence" versus "Partial adherence"), and the width of each
#|   flowing band is proportional to the number of households following that path.
#|   The bands narrow and split at successive stages, showing progressive attrition
#|   from broad program access down to full dosing compliance.
#| fig-height: 8
#| fig-width: 10

# Create long format for ggsankey from the exported stage dataframe
sankey_long <- fig_c5_sankey_adherence |>
  make_long(stage1, stage2, stage3, stage4)

# Define color palette (consistent Azahar blues)
sankey_colors <- c(
  "Ever obtained" = "#1a3a5c",
  "Access in 6 months" = "#4a90c2",
  "No access in 6 months" = "#a8d4f0",
  "Unknown" = "#d3d3d3",
  "Gave part or all" = "#4a90c2",
  "Did NOT give any" = "#a8d4f0",
  "No data" = "#d3d3d3",
  "Full adherence" = "#1a3a5c",
  "Partial adherence" = "#7ab8e8"
)

# Create Sankey plot
ggplot(sankey_long, aes(
  x = x,
  next_x = next_x,
  node = node,
  next_node = next_node,
  fill = factor(node),
  label = node
)) +
  geom_sankey(
    flow.alpha = 0.5,
    node.color = "gray30",
    smooth = 8,
    show.legend = FALSE
  ) +
  geom_sankey_label(
    size = 3.5,
    color = "black",
    fill = "white",
    alpha = 0.9,
    na.rm = TRUE
  ) +
  scale_fill_manual(values = sankey_colors, na.value = "#d3d3d3") +
  labs(
    title = "Adherence Journey: From Access to Compliance Patterns",
    subtitle = "Flow restricted to households that 'Ever obtained' Chispitas, showing dosing behavior and consolidated reasons",
    x = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, size = 11)
  )
```


## Nutritional Integration

### Adherence Factor Assignment

Children who received Chispitas are assigned adherence factors based on reported
dosing compliance:

```{r}
#| label: tbl-adherence-factor-mapping

tbl_c4_dosing_compliance |>
  filter(dosing_compliance %in% c("All", "More than half", "Half", "Less than half")) |>
  mutate(
    adherence_factor = case_when(
      dosing_compliance == "All" ~ 1.00,
      dosing_compliance == "More than half" ~ 0.75,
      dosing_compliance == "Half" ~ 0.50,
      dosing_compliance == "Less than half" ~ 0.25
    )
  ) |>
  select(dosing_compliance, percentage, adherence_factor) |>
  gt() |>
  cols_label(
    dosing_compliance = "Dosing Compliance",
    percentage = "% of Recipients",
    adherence_factor = "Adherence Factor"
  ) |>
  fmt_number(columns = percentage, decimals = 1, pattern = "{x}%") |>
  fmt_number(columns = adherence_factor, decimals = 2) |>
  tab_options(
    table.font.size = 12,
    table.width = pct(100)
  )
```

::: {.callout-note}
## Interpreting Adherence Factors
These factors represent the proportion of each sachet consumed per administration
event, not temporal adherence. A child receiving "half the sachet" daily receives
approximately 50% of the intended micronutrient dose during supplementation periods.
:::

The effective daily contribution combines the protocol adjustment (1/3) with
individual adherence. For example:

- **Full compliance**: 10 mg Fe × 1/3 × 1.00 = 3.33 mg Fe/day.
- **Half compliance**: 10 mg Fe × 1/3 × 0.50 = 1.67 mg Fe/day.

### Integration Summary

```{r}
#| label: tbl-d2-chispitas-summary
#| tbl-cap: "Chispitas Supplementation Summary"

tbl_d2_chispitas_summary |>
  pivot_longer(everything(), names_to = "Metric", values_to = "Value") |>
  mutate(
    Metric = case_when(
      Metric == "total_children" ~ "Total children (6-59 months)",
      Metric == "n_recipients" ~ "Chispitas recipients",
      Metric == "pct_recipients" ~ "Recipients (%)",
      Metric == "mean_adherence_all" ~ "Mean adherence factor (all children)",
      Metric == "mean_adherence_recipients" ~ "Mean adherence factor (recipients only)",
      Metric == "mean_fe_contribution" ~ "Mean effective Fe contribution (mg/day)",
      Metric == "mean_zn_contribution" ~ "Mean effective Zn contribution (mg/day)"
    )
  ) |>
  gt() |>
  tab_header(
    title = "Chispitas Supplementation Summary",
    subtitle = "Effective coverage and nutrient contribution (children 6-59 months)"
  ) |>
  cols_label(
    Metric = "Indicator",
    Value = "Value"
  ) |>
  fmt_number(columns = Value, decimals = 2) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  ) |>
  tab_source_note(
    source_note = "Source: SIVESNU 2018 | Note: Recipients = children with access who received part or all of sachets."
  )
```

::: {.callout-tip}
## Why Integrate into Non-Maize?
Chispitas are added to the non-maize intake component to maintain separation
between dietary sources. This structure enables direct comparison: biofortification
modifies the nutrient composition of maize, while supplementation modifies the
non-maize totals.
:::


## Validation

```{r}
#| label: fig-e-distribution-comparison
#| fig-cap: "Distribution Comparison: Non-Maize Intake Before and After Chispitas Integration"
#| fig-alt: >
#|   Two side-by-side panels of overlaid violin and box plots, one for iron and
#|   one for zinc, each with intake in milligrams per day on the y-axis. Within
#|   every panel the "Before Chispitas" and "After Chispitas" distributions are
#|   compared side by side. The after-integration violins and box medians sit
#|   higher, showing that supplementation raises non-maize iron and zinc intake.
#| fig-height: 6
#| fig-width: 10

# Define color palette
color_palette <- c("Before Chispitas" = "#a8d4f0", "After Chispitas" = "#5C946E")

# Calculate 95th percentile for axis limits
fe_upper <- quantile(
  c(fig_e_distribution_comparison$fe_before, fig_e_distribution_comparison$fe_after),
  0.95, na.rm = TRUE
)
zn_upper <- quantile(
  c(fig_e_distribution_comparison$zn_before, fig_e_distribution_comparison$zn_after),
  0.95, na.rm = TRUE
)

# Prepare long format for Iron
df_fe <- fig_e_distribution_comparison |>
  select(fe_before, fe_after) |>
  filter(fe_before > 0 & fe_before <= fe_upper,
         fe_after > 0 & fe_after <= fe_upper) |>
  pivot_longer(everything(), names_to = "stage", values_to = "value") |>
  mutate(stage = if_else(stage == "fe_before", "Before Chispitas", "After Chispitas"))

# Prepare long format for Zinc
df_zn <- fig_e_distribution_comparison |>
  select(zn_before, zn_after) |>
  filter(zn_before > 0 & zn_before <= zn_upper,
         zn_after > 0 & zn_after <= zn_upper) |>
  pivot_longer(everything(), names_to = "stage", values_to = "value") |>
  mutate(stage = if_else(stage == "zn_before", "Before Chispitas", "After Chispitas"))

# Iron plot
plot_fe <- df_fe |>
  ggplot(aes(x = stage, y = value, fill = stage)) +
  geom_violin(alpha = 0.6, trim = TRUE) +
  geom_boxplot(width = 0.1, alpha = 0.8, outlier.alpha = 0.3) +
  scale_fill_manual(values = color_palette) +
  labs(title = "Iron (Fe) - mg/day", x = NULL, y = "Intake (mg/day)") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, color = "azure4", face = "bold", hjust = 0.5),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4"),
    legend.position = "none",
    panel.grid.minor = element_blank()
  )

# Zinc plot
plot_zn <- df_zn |>
  ggplot(aes(x = stage, y = value, fill = stage)) +
  geom_violin(alpha = 0.6, trim = TRUE) +
  geom_boxplot(width = 0.1, alpha = 0.8, outlier.alpha = 0.3) +
  scale_fill_manual(values = color_palette) +
  labs(title = "Zinc (Zn) - mg/day", x = NULL, y = "Intake (mg/day)") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, color = "azure4", face = "bold", hjust = 0.5),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4"),
    legend.position = "none",
    panel.grid.minor = element_blank()
  )

# Create legend plot
legend_plot <- df_fe |>
  ggplot(aes(x = stage, y = value, fill = stage)) +
  geom_violin(alpha = 0.6) +
  scale_fill_manual(values = color_palette) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 11, color = "azure4")
  ) +
  guides(fill = guide_legend(override.aes = list(alpha = 0.6)))

# Extract legend
legend_grob <- ggpubr::get_legend(legend_plot)

# Arrange plots in grid
plot_grid_plots <- plot_grid(plot_fe, plot_zn, ncol = 2, nrow = 1)

# Combine with legend
plot_grid(plot_grid_plots, legend_grob, ncol = 1, rel_heights = c(1, 0.08))
```


## Summary

### Cascade Results

| Level | Filter | Retention |
|-------|--------|-----------|
| Access (lifetime) | Ever obtained Chispitas | `r sprintf("%.1f%%", pct_access_ever)` |
| Access (6 months) | Maintained access | `r sprintf("%.1f%%", pct_access_6months)` of above |
| Administration | Gave sachets | `r sprintf("%.1f%%", pct_administration)` of above |
| Dosing | Full compliance | `r sprintf("%.1f%%", pct_dosing_all)` of recipients |

The weighted average adherence factor among recipients is
`r sprintf("%.2f", mean_adherence_recipients)`. At the population level
(including non-recipients), the mean adherence factor is
`r sprintf("%.2f", mean_adherence_all)`.

### Key Findings

1. **High program reach** — `r sprintf("%.0f%%", pct_access_ever)` of children
   6–59 months have accessed Chispitas at some point, indicating broad program
   reach.

2. **Administration gap** — `r sprintf("%.0f%%", pct_admin_gap)` of households
   with access did not administer sachets to children, representing an
   intervention opportunity.

3. **Partial compliance common** — Among recipients,
   `r sprintf("%.1f%%", pct_dosing_all)` consumed full sachets,
   `r sprintf("%.1f%%", pct_dosing_more_half)` more than half,
   `r sprintf("%.1f%%", pct_dosing_half)` half, and
   `r sprintf("%.1f%%", pct_dosing_less_half)` less than half.

4. **SIVESNU vs official rates** — Observed coverage in our sample exceeds
   official MSPAS reports [@MSPAS2023_Chispitas], likely due to SIVESNU sampling
   characteristics. This discrepancy will be adjusted during synthetic
   population generation.
