---
title: "Pre-computing Baseline and QPM Profiles"
subtitle: "Module 4: Covariate Prediction"
description-meta: "Pre-computing each child's baseline and QPM nutrient profile and its GAM stunting effect, so scenario simulation becomes lookup and aggregation."
---

## Overview

This module pre-computes the complete **baseline (conventional maize)** and
**QPM (biofortified maize)** nutritional and stunting profiles for every
child in the synthetic population. Both nutrient indicators (intake,
bioavailability, inadequacy) and stunting outcomes (nutritional index,
GAM-derived HAZ change, stunting classification) are calculated upfront.

The result is a **scenario-ready dataset** where downstream scripts need only
select between pre-computed `*_baseline` and `*_qpm` columns based on each
child's biofortification assignment flag. This pre-computation is possible
because:

- The **baseline index position** depends only on the child's observed
  non-maize intake, which is fixed.
- The **QPM index position** depends only on that intake, the child's maize
  intake and the QPM biofortification factors, all fixed.

Therefore `delta_zlen_qpm`, the difference between the GAM fitted values at the
two positions, is a fixed per-child parameter that can be calculated once and
reused across every scenario in Module 6. The marginal effect at the baseline
position is computed alongside it and reported as a diagnostic, but does not
enter the impact calculation.

Main operations:

1. Load the synthetic population from
   [Generating Synthetic Population](04_07_generate_synthetic_population.html).
2. Load the GAM model and z-score normalization parameters from
   [Continuous Stunting Model](04_06_model_stunting_continuous_zlen.html).
3. Detect implausible consumption records, against a maize energy ceiling and
   a Tukey screen on each non-maize component, and impute the affected
   nutritional variables using k-Nearest Neighbours. Screen non-maize
   digestible protein separately and impute the records it flags.
4. Define QPM biofortification factors, absorption parameters, and H-AR
   thresholds for inadequacy classification.
5. Calculate the baseline nutritional profile: total intake, bioavailability,
   inadequacy classification, and nutritional index.
6. Calculate the QPM nutritional profile: biofortified intake,
   bioavailability, inadequacy classification, and nutritional index.
7. Express both nutritional indices on the estimation scale of the GAM, then
   calculate the delta NI and evaluate the smooth at both index positions.
8. Derive QPM stunting outcomes: delta HAZ, updated HAZ, and stunting
   classification.
9. Generate diagnostic tables comparing baseline and QPM profiles.
10. Export the enriched dataset with both profiles for scenario simulation.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  scales,             # Number formatting
  glue,               # String interpolation
  patchwork,          # Multi-panel plot composition
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script (exported by
# framework 04_08).

# --- Phase B: Outlier detection and imputation ---
tbl_b1_outlier_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_b1_outlier_summary.rds"),
  trust = TRUE
)

tbl_b4_imputation_quality <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_b4_imputation_quality.rds"),
  trust = TRUE
)

fig_b5_imputation_density <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_fig_b5_imputation_density.rds"),
  trust = TRUE
)

tbl_b6_protein_imputation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_b6_protein_imputation.rds"),
  trust = TRUE
)

# --- Phase E: GAM-based stunting impact ---
tbl_e1_index_profiles <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_e1_index_profiles.rds"),
  trust = TRUE
)

tbl_e3_stunting_impact_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_e3_stunting_impact_summary.rds"),
  trust = TRUE
)

tbl_e3_verify_delta_ni_by_age_group <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_e3_verify_delta_ni_by_age_group.rds"),
  trust = TRUE
)

# --- Phase G: Diagnostic comparison tables ---
tbl_g1_intake_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_g1_intake_comparison.rds"),
  trust = TRUE
)

tbl_g2_bioavailability_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_g2_bioavailability_comparison.rds"),
  trust = TRUE
)

tbl_g3_inadequacy_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_g3_inadequacy_comparison.rds"),
  trust = TRUE
)

fig_g4_delta_distributions <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_fig_g4_delta_distributions.rds"),
  trust = TRUE
)

# --- Phase H: Output coverage ---
tbl_h1_output_coverage <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_08_tbl_h1_output_coverage.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Outlier detection (NATIONAL row of tbl_b1)
national_outlier_row <- tbl_b1_outlier_summary |>
  filter(departamento == "NATIONAL")

n_outliers <- national_outlier_row |> pull(n_outliers)
pct_outliers <- national_outlier_row |> pull(pct_outliers) |> round(1)

# Non-maize protein screen (rows of tbl_b6)
n_prot_valid <- tbl_b6_protein_imputation |>
  filter(Source == "Valid (non-flagged)") |>
  pull(N)

n_prot_flagged <- tbl_b6_protein_imputation |>
  filter(Source == "Original outliers") |>
  pull(N)

pct_prot_flagged <- round(
  n_prot_flagged / (n_prot_valid + n_prot_flagged) * 100, 1
)

mean_prot_valid <- tbl_b6_protein_imputation |>
  filter(Source == "Valid (non-flagged)") |>
  pull(Mean)

mean_prot_original <- tbl_b6_protein_imputation |>
  filter(Source == "Original outliers") |>
  pull(Mean)

mean_prot_imputed <- tbl_b6_protein_imputation |>
  filter(Source == "Imputed (k-NN)") |>
  pull(Mean)

# Stunting impact (rows of tbl_e3)
mean_zlen_baseline <- tbl_e3_stunting_impact_summary |>
  filter(Metric == "Mean ZLEN") |>
  pull(Baseline) |>
  round(2)

mean_zlen_qpm <- tbl_e3_stunting_impact_summary |>
  filter(Metric == "Mean ZLEN") |>
  pull(QPM) |>
  round(2)

stunting_prev_baseline <- tbl_e3_stunting_impact_summary |>
  filter(Metric == "Stunting prevalence (%)") |>
  pull(Baseline) |>
  round(1)

stunting_prev_qpm <- tbl_e3_stunting_impact_summary |>
  filter(Metric == "Stunting prevalence (%)") |>
  pull(QPM) |>
  round(1)

stunting_reduction_pp <- round(stunting_prev_baseline - stunting_prev_qpm, 1)

mean_delta_zlen <- tbl_e3_stunting_impact_summary |>
  filter(Metric == "Mean delta ZLEN (QPM effect)") |>
  pull(QPM) |>
  round(3)

stunting_resolved_pct <- tbl_e3_stunting_impact_summary |>
  filter(Metric == "Stunting resolved (%)") |>
  pull(QPM) |>
  round(1)

# Intake changes
fe_pct_change <- tbl_g1_intake_comparison |>
  filter(Nutrient == "Iron (mg/day)") |>
  pull(Pct_Change) |>
  round(1)

zn_pct_change <- tbl_g1_intake_comparison |>
  filter(Nutrient == "Zinc (mg/day)") |>
  pull(Pct_Change) |>
  round(1)

prot_pct_change <- tbl_g1_intake_comparison |>
  filter(Nutrient == "Protein (g/day)") |>
  pull(Pct_Change) |>
  round(1)

# Inadequacy reductions
fe_inadeq_reduction <- tbl_g3_inadequacy_comparison |>
  filter(Nutrient == "Iron") |>
  pull(Reduction_pp) |>
  round(1)

zn_inadeq_reduction <- tbl_g3_inadequacy_comparison |>
  filter(Nutrient == "Zinc") |>
  pull(Reduction_pp) |>
  round(1)

prot_inadeq_reduction <- tbl_g3_inadequacy_comparison |>
  filter(Nutrient == "Protein") |>
  pull(Reduction_pp) |>
  round(1)
```


## Data Sources

### Synthetic Population

The input dataset is the calibrated synthetic population from
[Generating Synthetic Population](04_07_generate_synthetic_population.html).
It contains synthetic children with maize and non-maize nutrient intake
components, survey weights, baseline stunting predictions
(`zlen_predicted`, `hfa_calculated`), and all covariates needed for scenario
simulation.

### GAM Model and Normalization Parameters

The trained GAM for the nutritional-index → HAZ relationship and the z-score
normalization parameters come from
[Continuous Stunting Model](04_06_model_stunting_continuous_zlen.qmd). Two
parameter sets are loaded: those of the calibrated whole-diet index, and those
of the non-maize axis the GAM is fitted on. Both the baseline position and the
QPM increment are built on the non-maize axis with its own parameters, so the
increment is read off a curve whose unit means the same thing on both sides.
All are fixed objects and remain constant across scenarios.


## QPM Biofortification Factors

Nutrient content increases in Quality Protein Maize (QPM) biofortified
varieties relative to conventional maize, following the reference values
reported for the deployed varieties [@SemillaNueva2020_NutrientContent].
Biofortification is achieved through traditional plant breeding, not genetic
modification.

**PDCAAS calculation method.** The Protein Digestibility Corrected Amino Acid
Score (PDCAAS) follows FAO/WHO methodology [@WHO2007_ProteinAminoAcid] for children aged 1–2 years,
the most vulnerable segment of the target population:

$$PDCAAS = AAS \times Digestibility = \frac{Lys_{maize}}{Lys_{requirement}} \times D$$

| Parameter | Conventional | Biofortified | Source |
|-----------|--------------|--------------|--------|
| Lysine (mg/g maize) | 20.0 | 39.6 | Semilla Nueva Standard Report (2020) |
| FAO requirement (1–2 yr) | 52 mg/g | 52 mg/g | FAO/WHO 2007, Table 36 |
| Amino Acid Score (AAS) | 0.385 | 0.762 | = Lys_maize / 52 |
| Digestibility | 0.824 | 0.809 | Boye et al. 2012 |
| **PDCAAS** | **0.317** | **0.616** | = AAS × Digestibility |
| **Biofortification factor** | — | **1.94** | = 0.616 / 0.317 |

**Biofortification factors** applied to the maize-source nutrient
components:

| Nutrient | Factor | Source |
|----------|--------|--------|
| Iron | 1.19 | Semilla Nueva Standard Report (2020), +19% |
| Zinc | 1.38 | Semilla Nueva Standard Report (2020), +38% |
| Protein | 1.94 | PDCAAS increase from 0.317 to 0.616 (+94% utilizable protein) |
| Lysine | 1.98 | Semilla Nueva Standard Report (2020), 2.00% → 3.96% of protein |
| Tryptophan | 1.425 | Semilla Nueva Standard Report (2020), 0.40% → 0.57% of protein |
| Energy | 1.00 | Unchanged in QPM varieties |

Total protein quantity is unchanged in QPM varieties; what changes is the
amino acid profile. Lysine and tryptophan carry that change directly, and the
PDCAAS rises from 0.317 to 0.616 as its consequence.

::: {.callout-important}
## Each Series Receives the Change Once
The lysine and tryptophan factors apply to the **amino acid series**, which feed
the nutritional index. The protein factor applies to the **PDCAAS-adjusted
protein series**, which feeds the inadequacy classification against the protein
H-AR threshold. The two series are separate, so applying a factor to each
counts the same underlying change once on each side, and not twice on either.
:::


## Absorption Parameters

**Iron absorption** is set to a uniform 5% based on Allen et al. [-@Allen2020_HAR] for
predominantly plant-based, high-phytate diets — the dietary pattern of the
target population (>80% plant foods).

**Zinc absorption** follows the Miller equation [@Miller2015_Zinc] with
parameters $K_T = 2.42$ (saturation constant), $A = 0.633$ (age coefficient),
and $E = 0.311$ (age exponent). The equation computes total absorbed zinc
(TAZ) as a non-linear function of dietary zinc and age:

$$TAZ = \frac{1}{2}\left(K_T + A \cdot Age^{E} + TDZ - \sqrt{(K_T + A \cdot Age^{E} + TDZ)^2 - 4 \cdot A \cdot Age^{E} \cdot TDZ}\right)$$


## H-AR Thresholds for Inadequacy Classification

Harmonized Average Requirements (H-AR) define the thresholds below which a
child's absorbed nutrient intake is considered inadequate. Thresholds are
expressed in absorbed (iron, zinc) or bioavailable (protein, PDCAAS-adjusted)
units, matching the scale of the calculated nutrient values.

| Nutrient | Threshold | Age band | Source |
|----------|-----------|----------|--------|
| Iron (absorbed) | 0.5 mg/day | All ages (6–59 months) | Allen et al. (2020) |
| Zinc (absorbed via Miller) | 0.732 mg/day | 6–11 months | EFSA (2014) |
| Zinc (absorbed via Miller) | 1.074 mg/day | 12–35 months | EFSA (2014) |
| Zinc (absorbed via Miller) | 1.390 mg/day | 36–59 months | EFSA (2014) |
| Protein (digestible, PDCAAS) | 13 g/day | 6–35 months | WHO/FAO safe intake |
| Protein (digestible, PDCAAS) | 19 g/day | 36–59 months | WHO/FAO safe intake |


## Outlier Detection and Imputation

Two screens flag a record as implausible.

**Maize ceiling.** Children with daily maize energy intake exceeding 2,500 kcal
— equivalent to approximately 685 g/day of nixtamalized maize at 3.65 kcal/g
(INCAP conversion factor) — are flagged. That level exceeds plausible
consumption for children aged 6–59 months and points to a systematic error in
the household food consumption module the nutrient variables derive from.

**Non-maize screen.** The upper tail of the non-maize distribution reaches
values far above the median, placing a fraction of the population outside the
support of the SIVESNU distributions the models were fitted on. Each non-maize
component — iron, zinc, lysine, tryptophan and energy — is therefore screened
against its own upper Tukey fence at $Q3 + 3 \times IQR$. The multiplier is 3
rather than the usual 1.5 because the rule declares a record implausible and
triggers imputation: on non-maize energy it places the cut at a level that,
from non-maize sources alone, already exceeds the total energy requirement for
this age range.

All twelve source-specific nutrient variables (six maize-source and six
non-maize-source) originate from the same ENCOVI household consumption module,
so an implausible value on either side signals unreliable data across the
entire nutrient profile. For flagged records, all twelve variables are set to
NA and imputed using **k-Nearest Neighbours** (RANN k-d tree, k = 10,
inverse-distance weighting).

**Predictors.** The imputation is conditioned on the child's own ENCOVI record:
age, sex, department, area, household education and the plant-source share of
the diet. The quantity being reconstructed is an ENCOVI dietary value, so it is
rebuilt from variables observed for the same child in the same survey, which
keeps the reconstruction inside one measurement frame.

Five of the six are collected outside the food consumption module. The
plant-source share is a composition ratio derived from it, and is retained on
that basis: the reporting error the screen targets scales the reported
quantities of a household and leaves their relative composition intact, so the
ratio carries dietary-pattern signal that survives the error invalidating the
amounts.

The set carries one variable per dimension of the matching space — demographics,
geography, socioeconomic position and diet composition. All are continuous or
binary, which keeps the scaled Euclidean metric well defined and leaves each
dimension weighted once.

Records are imputed, never dropped, so the synthetic population stays
representative.

Imputed values are floored at zero to prevent negative nutrient intakes.

### Outlier Detection Summary

```{r}
#| label: tbl-b1-outlier-summary
#| tbl-cap: "Records flagged as implausible by either screen, by department"

tbl_b1_outlier_summary |>
  gt() |>
  tab_header(
    title = "Outlier Detection: Implausible Intake Records",
    subtitle = "Maize ceiling 2,500 kcal/day, plus the non-maize Tukey screen at Q3 + 3 x IQR"
  ) |>
  cols_label(
    departamento = "Department",
    n_total = "N Total",
    n_outliers = "N Outliers",
    pct_outliers = "% Outliers",
    mean_ene_maiz_outliers = "Mean kcal (outliers)",
    mean_ene_maiz_valid = "Mean kcal (valid)"
  ) |>
  fmt_number(columns = c(n_total, n_outliers), decimals = 0) |>
  fmt_number(columns = c(pct_outliers), decimals = 1) |>
  fmt_number(
    columns = c(mean_ene_maiz_outliers, mean_ene_maiz_valid),
    decimals = 0
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = departamento == "NATIONAL")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(columns = pct_outliers, rows = pct_outliers > 10)
  ) |>
  sub_missing(missing_text = "—") |>
  tab_source_note(
    source_note = "Conversion: 3.65 kcal/g (INCAP, nixtamalized maize tortilla)."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### k-NN Imputation Quality

Imputation quality is evaluated by comparing the distribution of imputed
values against valid records within the same departments where outliers
concentrate. Since outlier records are disproportionately located in
high-maize-consumption departments (Alta Verapaz, Baja Verapaz, Totonicapán),
a national-level comparison would underestimate the expected intake for
these children. The department-stratified comparison provides a more
appropriate benchmark.

Mean ratios close to 1.0 indicate that imputed values are consistent with
the intake patterns observed among valid children in the same geographic
and demographic context.

```{r}
#| label: tbl-b4-imputation-quality
#| tbl-cap: "k-NN imputation quality: department-stratified comparison of imputed vs valid distributions"

tbl_b4_imputation_quality |>
  gt() |>
  tab_header(
    title = "k-NN Imputation Quality Assessment",
    subtitle = "k = 10, inverse-distance weighted, RANN k-d tree"
  ) |>
  cols_label(
    variable = "Variable",
    Valid_mean = "Valid Mean",
    Imputed_mean = "Imputed Mean",
    Valid_sd = "Valid SD",
    Imputed_sd = "Imputed SD",
    mean_ratio = "Mean Ratio",
    sd_ratio = "SD Ratio"
  ) |>
  fmt_number(
    columns = c(Valid_mean, Imputed_mean, Valid_sd, Imputed_sd),
    decimals = 2
  ) |>
  fmt_number(columns = c(mean_ratio, sd_ratio), decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = mean_ratio,
      rows = mean_ratio > 0.8 & mean_ratio < 1.2
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(
      columns = mean_ratio,
      rows = mean_ratio <= 0.8 | mean_ratio >= 1.2
    )
  ) |>
  tab_source_note(
    source_note = "Comparison stratified to departments containing outlier records."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### Pre- and Post-Imputation Density Comparison

Density plots comparing the distribution of key nutritional variables before
and after imputation. The imputed values (outlier records only) fall within
the range of valid observations, concentrated around the conditional mean that
k-NN imputation produces.

```{r}
#| label: fig-b5-imputation-density
#| fig-cap: "Pre- and post-imputation density comparison for key nutritional variables"
#| fig-alt: >
#|   Grid of four overlaid density plots, one per maize-derived nutrient: iron
#|   (mg/day), zinc (mg/day), protein (g/day), and energy (kcal/day), each on
#|   its own x-axis. Within every panel three curves are compared: original
#|   outlier records, their imputed replacements, and valid non-outlier
#|   observations. The imputed curves stay within the range of the valid
#|   observations but are narrower and more sharply peaked than them, the
#|   concentration around the conditional mean that k-nearest-neighbour
#|   imputation produces. The original outlier curve sits well to the right of
#|   both.
#| fig-width: 10
#| fig-height: 10

colors_comparison <- c(
  "Original (outliers)" = "#E74C3C",
  "Imputed" = "#3AAF85",
  "Valid (non-outlier)" = "#1B6CA8"
)

theme_density <- theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 11, face = "bold", color = "gray30"),
    axis.title = element_text(size = 9, color = "gray30"),
    legend.title = element_blank()
  )

p_fe <- ggplot(
  fig_b5_imputation_density,
  aes(x = fe_mg_maiz, fill = source, color = source)
) +
  geom_density(alpha = 0.3, linewidth = 0.5) +
  scale_fill_manual(values = colors_comparison) +
  scale_color_manual(values = colors_comparison) +
  labs(title = "Iron from maize (mg/day)",
       x = "Fe maize (mg/day)", y = "Density") +
  coord_cartesian(
    xlim = c(0, quantile(fig_b5_imputation_density$fe_mg_maiz, 0.99, na.rm = TRUE))
  ) +
  theme_density

p_zn <- ggplot(
  fig_b5_imputation_density,
  aes(x = zn_mg_maiz, fill = source, color = source)
) +
  geom_density(alpha = 0.3, linewidth = 0.5) +
  scale_fill_manual(values = colors_comparison) +
  scale_color_manual(values = colors_comparison) +
  labs(title = "Zinc from maize (mg/day)",
       x = "Zn maize (mg/day)", y = "Density") +
  coord_cartesian(
    xlim = c(0, quantile(fig_b5_imputation_density$zn_mg_maiz, 0.99, na.rm = TRUE))
  ) +
  theme_density

p_prot <- ggplot(
  fig_b5_imputation_density,
  aes(x = prot_g_maiz, fill = source, color = source)
) +
  geom_density(alpha = 0.3, linewidth = 0.5) +
  scale_fill_manual(values = colors_comparison) +
  scale_color_manual(values = colors_comparison) +
  labs(title = "Protein from maize (g/day)",
       x = "Prot maize (g/day)", y = "Density") +
  coord_cartesian(
    xlim = c(0, quantile(fig_b5_imputation_density$prot_g_maiz, 0.99, na.rm = TRUE))
  ) +
  theme_density

p_ene <- ggplot(
  fig_b5_imputation_density,
  aes(x = ene_kcal_maiz, fill = source, color = source)
) +
  geom_density(alpha = 0.3, linewidth = 0.5) +
  scale_fill_manual(values = colors_comparison) +
  scale_color_manual(values = colors_comparison) +
  labs(title = "Energy from maize (kcal/day)",
       x = "Energy maize (kcal/day)", y = "Density") +
  coord_cartesian(
    xlim = c(0, quantile(fig_b5_imputation_density$ene_kcal_maiz, 0.99, na.rm = TRUE))
  ) +
  theme_density

(p_fe + p_zn) / (p_prot + p_ene) +
  plot_annotation(
    title = "Distribution Comparison: Original Outliers vs Imputed Values",
    subtitle = glue("{n_outliers} records imputed ({pct_outliers}% of population)"),
    theme = theme(
      plot.title = element_text(size = 14, face = "bold", color = "gray30"),
      plot.subtitle = element_text(size = 11, color = "gray50")
    )
  )
```


### Non-Maize Protein Outlier Treatment

Digestible protein from non-maize sources carries a heavy right tail in the
dietary intake distribution. A small fraction of children report non-maize
protein intake above its upper Tukey fence, and those values propagate into the
protein totals.

Protein is screened on its own axis, once the block treatment above has settled
the rest of the nutrient profile. The criterion targets the tail of a single
component rather than a systematically unreliable record, so the fence
multiplier is 1.5 rather than the 3 used for the block screen. Records above
the fence are replaced by k-NN imputation with the same predictor set and the
same weighting; only the non-maize protein variable is invalidated and imputed,
and every other nutrient variable is carried through unchanged.

```{r}
#| label: tbl-b6-protein-imputation
#| tbl-cap: "Non-maize protein imputation: valid, original and imputed values"

tbl_b6_protein_imputation |>
  gt() |>
  tab_header(
    title = "Non-Maize Protein: Imputation Diagnostics",
    subtitle = glue(
      "{format(n_prot_flagged, big.mark = ',')} records imputed ",
      "({pct_prot_flagged}% of the synthetic population)"
    )
  ) |>
  cols_label(
    Source = "Source",
    N = "N",
    Mean = "Mean",
    Median = "Median",
    Max = "Max"
  ) |>
  fmt_number(columns = c(Mean, Median, Max), decimals = 2) |>
  fmt_number(columns = N, decimals = 0, use_seps = TRUE) |>
  cols_align(align = "left", columns = Source) |>
  tab_source_note(
    "Tukey upper fence at Q3 + 1.5 x IQR. k-NN RANN (k = 10, inverse-distance weighted), same predictors as the block imputation."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-note}
## Where the Protein Replacements Land
The flagged records averaged
`r format(round(mean_prot_original, 1), big.mark = ",")` g/day of non-maize
digestible protein against `r round(mean_prot_valid, 1)` g/day among the valid
records. Their replacements average `r round(mean_prot_imputed, 1)` g/day, so
they sit within the range the valid records occupy rather than at its centre.
:::


## Baseline Nutritional Profile

The baseline profile represents each child's nutritional status under
**conventional maize** consumption. For every child, the following are
computed:

1. **Total intake**: maize + non-maize components for iron, zinc, protein,
   lysine, tryptophan, and energy.
2. **Iron bioavailability**: uniform 5% absorption applied to total iron
   intake.
3. **Zinc bioavailability**: total absorbed zinc (TAZ) via the Miller
   equation, age-dependent.
4. **Nutritional index**: four-component biofortification-responsive index on
   the non-maize axis, built from z-score-normalized absorbed iron, TAZ,
   lysine and tryptophan, using the fixed normalization parameters of that
   axis:

   $$NI_{bio}^{nomaiz} = \frac{1}{4}\left(Fe_{absorbed,z} + Zn_{absorbed,z} + Lys_{z} + Trp_{z}\right)$$

5. **Inadequacy classification**: comparison of absorbed iron, TAZ, and
   PDCAAS-adjusted protein against the age-specific H-AR thresholds defined
   above.

The result is a per-child baseline vector of intake totals, absorbed
fractions, inadequacy flags, and the nutritional index used by the
downstream GAM impact calculation.


## QPM Nutritional Profile

The QPM profile is computed analogously to the baseline, but applies the
biofortification factors to the maize-source components before summing:

$$Nutrient_{total,QPM} = Nutrient_{maize} \times Factor_{QPM} + Nutrient_{non\text{-}maize}$$

Non-maize sources remain unchanged, since biofortification modifies only the
maize component of the diet. The same Miller equation, z-score parameters,
and H-AR thresholds are reused so that baseline and QPM are strictly
comparable. The output is a parallel per-child vector of QPM intake totals,
absorbed fractions, inadequacy flags, and `nutritional_index_bio_qpm`.


## GAM-Based Stunting Impact

The stunting impact of full biofortification is obtained by evaluating the GAM
at each child's baseline and biofortified position on the index.

**Delta NI** captures the change in the four-component non-maize nutritional
index attributable to full biofortification:

$$\Delta NI = NI_{bio,QPM}^{nomaiz} - NI_{bio,baseline}^{nomaiz}$$

This value is fixed per child and invariant across production coverage levels —
it is the whole distance the index travels under complete adoption.

**Delta HAZ** is the difference between the two fitted values, with age and sex
held at each child's observed values:

$$\Delta HAZ_{QPM} = f\left(NI_{bio,QPM}^{nomaiz}\right) - f\left(NI_{bio,baseline}^{nomaiz}\right)$$

Taking the difference of fitted values, rather than a first-order term, lets the
estimate follow the curvature of the smooth across the whole interval the index
traverses. The marginal effect at the baseline position is also computed and
retained as a diagnostic.

### Index Distribution Under Both Profiles

The table below reports where the index sits before and after full
biofortification, on both axes. The non-maize axis is the one the GAM is
evaluated on; the whole-diet axis is reported alongside it as the descriptive
counterpart on the scale the rest of the pipeline uses.

```{r}
#| label: tbl-e1-index-profiles
#| tbl-cap: "Nutritional index distribution under baseline and QPM profiles"

tbl_e1_index_profiles |>
  gt() |>
  tab_header(
    title = "Nutritional Index Profiles",
    subtitle = "Survey-weighted quantiles across the synthetic population"
  ) |>
  cols_label(
    Quantile = "Quantile",
    Baseline = "Baseline",
    QPM = "QPM",
    Delta = "Delta",
    Baseline_raw = "Baseline",
    QPM_raw = "QPM",
    Delta_raw = "Delta"
  ) |>
  tab_spanner(
    label = "Whole-diet axis (descriptive)",
    columns = c(Baseline, QPM, Delta)
  ) |>
  tab_spanner(
    label = "Non-maize axis (evaluated by the GAM)",
    columns = c(Baseline_raw, QPM_raw, Delta_raw)
  ) |>
  fmt_number(columns = -Quantile, decimals = 3) |>
  cols_align(align = "left", columns = Quantile) |>
  cols_align(align = "right", columns = -Quantile) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

**Updated HAZ under QPM** and the corresponding stunting classification:

$$HAZ_{QPM} = HAZ_{baseline} + \Delta HAZ_{QPM}$$
$$\text{Stunted}_{QPM} = \mathbb{1}\{HAZ_{QPM} < -2\}$$

A child whose stunting status flips from stunted at baseline to not stunted
under QPM is flagged as `stunting_resolved_qpm = 1`.

### Pre-Computed Stunting Impact Summary

```{r}
#| label: tbl-e3-stunting-impact-summary
#| tbl-cap: "Pre-computed stunting impact: baseline vs QPM (full biofortification)"

tbl_e3_stunting_impact_summary |>
  gt() |>
  tab_header(
    title = "Stunting Impact: Baseline vs QPM (Full Biofortification)",
    subtitle = "Survey-weighted estimates for all synthetic children"
  ) |>
  fmt_number(
    columns = c(Baseline, QPM, Change),
    decimals = 2, rows = c(1, 2, 3, 5)
  ) |>
  fmt_number(
    columns = c(Baseline, QPM, Change),
    decimals = 0, rows = 4
  ) |>
  sub_missing(missing_text = "—") |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = -Metric) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = Change,
      rows = !is.na(Change) & Change > 0
    )
  ) |>
  tab_source_note(
    source_note = "GAM: s(nutritional_index_bio) → HAZ. Floor constraint: delta HAZ ≥ 0."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### Stunting Impact by Age Band

The table below disaggregates delta NI, marginal effect, and delta HAZ by
6-month age bands. Two elements of the construction vary with age: the
age-dependent AME granularity underlying the baseline nutrient intake
estimates, and the position of each age group on the GAM smooth.

```{r}
#| label: tbl-e3-verify-delta-ni-by-age-group
#| tbl-cap: "Delta NI and stunting impact by 6-month age band"

tbl_e3_verify_delta_ni_by_age_group |>
  gt() |>
  tab_header(
    title = "Delta NI and Stunting Impact by 6-Month Age Band",
    subtitle = "Survey-weighted estimates (full biofortification)"
  ) |>
  cols_label(
    age_group = "Age (months)",
    n = "N",
    mean_delta_ni = "Mean Δ NI",
    mean_marginal = "Mean Marginal Effect",
    mean_delta_zlen = "Mean Δ HAZ",
    pct_stunted_baseline = "Stunting Baseline (%)",
    stunting_reduction_pp = "Reduction (pp)"
  ) |>
  fmt_number(
    columns = c(mean_delta_ni, mean_marginal, mean_delta_zlen),
    decimals = 4
  ) |>
  fmt_number(
    columns = c(pct_stunted_baseline, stunting_reduction_pp),
    decimals = 2
  ) |>
  fmt_integer(columns = n) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = mean_delta_ni,
      rows = age_group == "6-11"
    )
  ) |>
  tab_source_note(
    source_note = "Larger delta NI in the youngest band reflects the AME granularity for children aged 6–11 months."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Baseline vs QPM Diagnostic Comparison

The following tables and visualizations compare the population-level nutrient
intake, bioavailability, inadequacy prevalence, and individual-level deltas
between the baseline and QPM profiles under **full biofortification**
(every child consuming QPM).

### Mean Nutrient Intake

```{r}
#| label: tbl-g1-intake-comparison
#| tbl-cap: "Mean nutrient intake: baseline vs QPM (full biofortification)"

tbl_g1_intake_comparison |>
  gt() |>
  tab_header(
    title = "Nutrient Intake Comparison: Baseline vs QPM",
    subtitle = "Survey-weighted mean intake (full biofortification, all children)"
  ) |>
  fmt_number(columns = c(Baseline, QPM, Delta), decimals = 2) |>
  fmt_number(columns = Pct_Change, decimals = 1) |>
  cols_label(Pct_Change = "% Change") |>
  cols_align(align = "left", columns = Nutrient) |>
  cols_align(align = "right", columns = -Nutrient) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = Pct_Change, rows = Pct_Change > 0)
  ) |>
  tab_source_note(
    source_note = "Biofortification factors: Fe ×1.19, Zn ×1.38, Protein ×1.94 (PDCAAS). Lys, Trp, Energy unchanged."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### Bioavailability and Nutritional Index

```{r}
#| label: tbl-g2-bioavailability-comparison
#| tbl-cap: "Absorbed nutrient and NI comparison: baseline vs QPM"

tbl_g2_bioavailability_comparison |>
  gt() |>
  tab_header(
    title = "Absorbed Nutrients and Nutritional Index: Baseline vs QPM",
    subtitle = "Survey-weighted mean values (full biofortification)"
  ) |>
  fmt_number(columns = c(Baseline, QPM, Delta), decimals = 3) |>
  fmt_number(columns = Pct_Change, decimals = 1) |>
  cols_label(Pct_Change = "% Change") |>
  cols_align(align = "left", columns = Nutrient) |>
  cols_align(align = "right", columns = -Nutrient) |>
  tab_source_note(
    source_note = "Iron: 5% absorption. Zinc: Miller equation. NI: z-score normalized (SIVESNU training params)."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### Inadequacy Prevalence

The table below compares the prevalence of inadequate intake (absorbed
nutrient below the age-specific H-AR threshold) under baseline and QPM, and
reports the share of inadequate children whose status is resolved by QPM.

```{r}
#| label: tbl-g3-inadequacy-comparison
#| tbl-cap: "Inadequacy prevalence: baseline vs QPM (full biofortification)"

tbl_g3_inadequacy_comparison |>
  gt() |>
  tab_header(
    title = "Inadequacy Prevalence: Baseline vs QPM",
    subtitle = "Survey-weighted prevalence (%, full biofortification scenario)"
  ) |>
  fmt_number(
    columns = c(Baseline_pct, QPM_pct, Resolved_pct, Reduction_pp),
    decimals = 1
  ) |>
  cols_label(
    Baseline_pct = "Baseline (%)",
    QPM_pct = "QPM (%)",
    Resolved_pct = "Resolved (%)",
    Reduction_pp = "Reduction (pp)"
  ) |>
  sub_missing(missing_text = "—") |>
  cols_align(align = "left", columns = Nutrient) |>
  cols_align(align = "right", columns = -Nutrient) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = Reduction_pp, rows = Reduction_pp > 0)
  ) |>
  tab_source_note(
    source_note = "H-AR: Fe 0.5 mg/d absorbed (Allen 2020), Zn age-dependent (EFSA 2014), Protein 13/19 g/d (WHO/FAO)."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### Delta Distributions

Density plots of the individual-level QPM − baseline deltas: absorbed iron,
absorbed zinc (TAZ), and the GAM-derived stunting impact (delta HAZ). All
distributions are bounded below at zero by construction.

```{r}
#| label: fig-g4-delta-distributions
#| fig-cap: "Distribution of QPM − baseline deltas across the synthetic population"
#| fig-alt: >
#|   Three density plots stacked vertically, each showing the individual-level
#|   QPM minus baseline delta for one measure: absorbed iron (mg/day), absorbed
#|   zinc or TAZ (mg/day), and GAM-derived stunting impact (delta HAZ in SD
#|   units). A dashed reference line marks zero in every panel. All three
#|   distributions are bounded at zero by construction and lie to its right,
#|   showing that biofortification yields uniformly positive improvements in
#|   nutrient absorption and predicted height-for-age across the population.
#| fig-width: 10
#| fig-height: 10

theme_delta <- theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 12, face = "bold", color = "gray30"),
    axis.title = element_text(size = 10, color = "gray30"),
    axis.text = element_text(size = 9, color = "gray40")
  )

p_delta_fe <- ggplot(
  fig_g4_delta_distributions,
  aes(x = delta_fe_absorbed_mg)
) +
  geom_density(fill = "#3E7CB1", alpha = 0.4, color = "#3E7CB1", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  labs(
    title = "Delta iron absorbed (mg/day)",
    x = "Δ Fe absorbed", y = "Density"
  ) +
  theme_delta

p_delta_zn <- ggplot(
  fig_g4_delta_distributions,
  aes(x = delta_taz_mg)
) +
  geom_density(fill = "#3E7CB1", alpha = 0.4, color = "#3E7CB1", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  labs(
    title = "Delta zinc absorbed — TAZ (mg/day)",
    x = "Δ TAZ", y = "Density"
  ) +
  theme_delta

p_delta_zlen <- ggplot(
  fig_g4_delta_distributions,
  aes(x = delta_zlen_qpm)
) +
  geom_density(fill = "#6BAF6B", alpha = 0.4, color = "#6BAF6B", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  labs(
    title = "Delta HAZ (GAM-derived stunting impact)",
    x = "Δ HAZ (SD units)", y = "Density"
  ) +
  theme_delta

p_delta_fe / p_delta_zn / p_delta_zlen +
  plot_annotation(
    title = "QPM − Baseline Deltas: Nutritional and Stunting Impact",
    subtitle = "Individual-level improvements under full biofortification (all children)",
    theme = theme(
      plot.title = element_text(size = 14, face = "bold", color = "gray30"),
      plot.subtitle = element_text(size = 11, color = "gray50")
    )
  )
```


## Summary

### Processing Pipeline

| Phase | Method | Purpose |
|-------|--------|---------|
| A. Parameters | Fixed constants | QPM factors, absorption parameters, H-AR thresholds |
| B. Outlier treatment | Maize ceiling + non-maize Tukey screen + k-NN imputation, then a separate protein screen | Reconstruct the nutrient profile for implausible records |
| C. Baseline profile | Intake aggregation, Miller equation, z-score normalization | Per-child baseline vector under conventional maize |
| D. QPM profile | Biofortification factors applied to maize sources | Per-child QPM vector under biofortification |
| E. Stunting impact | Difference of GAM fitted values at the two index positions | Per-child delta HAZ and QPM stunting classification |
| F. Deltas | QPM − baseline | Individual-level improvements and resolution flags |
| G. Diagnostics | Population-level comparisons | Validate aggregate impact of full biofortification |
| H. Export | Split parquets | Scenario-ready dataset for Module 6 |

### Key Results

- **Outlier treatment**: `r n_outliers` records (`r pct_outliers`% of the
  synthetic population) imputed via k-NN, plus
  `r format(n_prot_flagged, big.mark = ",")` records (`r pct_prot_flagged`%)
  on the separate non-maize protein screen.
- **Mean HAZ**: `r mean_zlen_baseline` (baseline) → `r mean_zlen_qpm`
  (QPM under full biofortification).
- **Stunting prevalence**: `r stunting_prev_baseline`% (baseline) →
  `r stunting_prev_qpm`% (QPM); reduction of
  `r stunting_reduction_pp` percentage points under full biofortification.
- **Mean delta HAZ**: `r mean_delta_zlen` SD units per child.
- **Stunting resolved**: `r stunting_resolved_pct`% of all synthetic
  children flip from stunted at baseline to not stunted under QPM.
- **Mean nutrient intake change**: iron `r fe_pct_change`%, zinc
  `r zn_pct_change`%, PDCAAS-adjusted protein `r prot_pct_change`%.
- **Inadequacy reduction**: iron `r fe_inadeq_reduction` pp, zinc
  `r zn_inadeq_reduction` pp, protein `r prot_inadeq_reduction` pp.

### Application

The exported dataset is the input to the **scenario simulation** in
Module 6. For every production coverage level, the scenario scripts select between
the pre-computed `*_baseline` and `*_qpm` columns based on each child's
biofortification assignment flag. Since both vectors are calculated upfront,
the scenario simulation reduces to a per-child indexing operation with no
additional GAM evaluation, model prediction, or bioavailability calculation
needed at runtime.

::: {.callout-warning}
## Limitations

1. **Extrapolation beyond the fitted range** — The QPM index position can fall
   above the range the smooth was estimated on, where the GAM extends its fit
   without observations to support it.

2. **Biofortification factor assumption** — The QPM factors (Fe ×1.19,
   Zn ×1.38, lysine ×1.98, tryptophan ×1.425, and protein ×1.94 via PDCAAS)
   are derived from the Semilla Nueva Standard Report (2020) and FAO/WHO
   methodology for children aged 1–2 years. Realized nutrient gains in
   deployment may vary with seed variety, soil conditions, cooking practice,
   and the age composition of the consuming population.

3. **k-NN imputation scope** — Imputed values reflect the central tendency
   of demographically similar valid children. Individual variation within
   the neighbourhood is compressed by the inverse-distance weighting,
   leading to slightly narrower SDs for imputed records than for valid
   records (visible in the imputation quality table).

4. **Predictor set of the imputation** — The six predictors span demographics,
   geography, socioeconomic position and diet composition, one variable per
   dimension. Determinants of dietary intake outside those four dimensions are
   not represented in the matching space, so the replacements reproduce the
   central tendency of that space rather than the full heterogeneity of intake.

5. **Iron absorption uniformity** — The 5% absorption factor is uniform
   across all children, following Allen et al. (2020) for plant-based
   diets. Heterogeneity in heme iron consumption and inhibitor exposure
   (phytates, polyphenols, tannins) is not captured at the individual level.
:::
