---
title: "Calculating Food Security Scores"
subtitle: "Module 4: Covariate Prediction"
---

## Overview

Food security — the consistent access to sufficient, safe, and nutritious food —
is a key determinant of child nutritional status. The Food Insecurity Experience
Scale (FIES) captures household food security through eight questions of
increasing severity, from worry about food availability to going entire days
without eating.

FIES produces raw scores from 0 (food secure) to 8 (severe food insecurity).
However, raw scores are ordinal — the difference between scores 1 and 2 is not
necessarily equal to the difference between 7 and 8. To enable valid statistical
modeling, we apply the **Rasch model** recommended by FAO, which transforms
responses into interval-scale measurements where differences are comparable
across the entire range.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script (exported by framework 04_04).

tbl_a2_fies_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_04_tbl_a2_fies_distribution.rds"),
  trust = TRUE
)

tbl_b2_item_parameters <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_04_tbl_b2_item_parameters.rds"),
  trust = TRUE
)

tbl_b3_raw_score_parameters <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_04_tbl_b3_raw_score_parameters.rds"),
  trust = TRUE
) |>
  # Recode framework's short labels to the descriptive labels reviewed
  # and approved for the methodological web.
  mutate(
    classification = recode(
      classification,
      "Mild FI"     = "Mild food insecurity",
      "Moderate FI" = "Moderate food insecurity",
      "Severe FI"   = "Severe food insecurity"
    )
  )

tbl_c2_score_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_04_tbl_c2_score_comparison.rds"),
  trust = TRUE
) |>
  mutate(
    classification = recode(
      classification,
      "Mild FI"     = "Mild food insecurity",
      "Moderate FI" = "Moderate food insecurity",
      "Severe FI"   = "Severe food insecurity"
    )
  )

tbl_c3_model_fit <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_04_tbl_c3_model_fit.rds"),
  trust = TRUE
)

fig_c4_rasch_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_04_fig_c4_rasch_distribution.rds"),
  trust = TRUE
) |>
  mutate(
    classification = fct_recode(
      classification,
      "Mild food insecurity"     = "Mild FI",
      "Moderate food insecurity" = "Moderate FI",
      "Severe food insecurity"   = "Severe FI"
    )
  )
```

```{r}
#| label: inline-values
#| include: false

# Weighted prevalence by classification (from tbl-c2-score-comparison)
prevalence_by_class <- tbl_c2_score_comparison |>
  summarise(prevalence = sum(pct), .by = classification)

pct_food_secure <- prevalence_by_class |>
  filter(classification == "Food Secure") |>
  pull(prevalence)

pct_mild <- prevalence_by_class |>
  filter(classification == "Mild food insecurity") |>
  pull(prevalence)

pct_moderate <- prevalence_by_class |>
  filter(classification == "Moderate food insecurity") |>
  pull(prevalence)

pct_severe <- prevalence_by_class |>
  filter(classification == "Severe food insecurity") |>
  pull(prevalence)

# Headline cifras for the opening narrative
pct_yes_item1 <- tbl_a2_fies_distribution |>
  filter(code == "C001") |>
  pull(pct_yes)

pct_yes_item8 <- tbl_a2_fies_distribution |>
  filter(code == "C008") |>
  pull(pct_yes)
```


## FIES Item Response Distribution

The eight FIES items follow FAO's standard ordering by severity. The expected
pattern — decreasing affirmative responses as severity increases — is confirmed
in SIVESNU data: `r sprintf("%.0f%%", pct_yes_item1)` report worrying about food
(mild), while only `r sprintf("%.0f%%", pct_yes_item8)` report going a full day
without eating (severe).

```{r}
#| label: tbl-a2-fies-distribution
#| tbl-cap: "FIES Item Response Distribution"

tbl_a2_fies_distribution |>
  gt() |>
  tab_header(
    title = "FIES Item Response Distribution",
    subtitle = "Food Insecurity Experience Scale (8 items)"
  ) |>
  cols_label(
    code = "Code",
    description = "Question",
    severity = "Severity",
    n_yes = "Yes",
    n_no = "No",
    n_na = "NA",
    pct_yes = "% Yes"
  ) |>
  fmt_number(columns = c(n_yes, n_no, n_na), decimals = 0) |>
  fmt_number(columns = pct_yes, decimals = 1) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(rows = severity == "Mild", columns = severity)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFE0B2"),
    locations = cells_body(rows = severity == "Moderate", columns = severity)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = severity == "Severe", columns = severity)
  ) |>
  tab_source_note(
    source_note = "Source: SIVESNU 2018. Items ordered by increasing severity per FAO FIES methodology."
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```


## Rasch Model Estimation

The Rasch model is fit on the eight binary FIES responses using survey weights,
following the FAO methodology for the Voices of the Hungry project. Two sets of
parameters emerge from the estimation: **item difficulty** parameters (one per
question, characterizing how severe the experience reported by each item is)
and **person severity** parameters (one per raw score, mapping each total to an
interval-scale food insecurity level).

### Item Difficulty Parameters

Item difficulty (β) represents the severity threshold at which each experience is
typically reported. Negative values indicate experiences reported by most
respondents; positive values indicate experiences reported only by severely
food insecure households.

```{r}
#| label: tbl-b2-item-parameters
#| tbl-cap: "FIES Item Difficulty Parameters (Rasch Model)"

tbl_b2_item_parameters |>
  gt() |>
  tab_header(
    title = "FIES Item Difficulty Parameters",
    subtitle = "Rasch model estimates (ordered by difficulty)"
  ) |>
  cols_label(
    rank = "Rank",
    code = "Code",
    description = "Question",
    severity = "Severity",
    difficulty = "Difficulty (β)"
  ) |>
  fmt_number(columns = difficulty, decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(rows = severity == "Mild", columns = severity)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFE0B2"),
    locations = cells_body(rows = severity == "Moderate", columns = severity)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = severity == "Severe", columns = severity)
  ) |>
  tab_source_note(
    source_note = "Higher difficulty = more severe food insecurity. Negative values = more common experiences."
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

### Raw Score to Rasch Transformation

The model transforms each raw score (0-8) into an interval-scale severity
parameter (θ). These parameters are used to assign individual food insecurity
scores.

```{r}
#| label: tbl-b3-raw-score-parameters
#| tbl-cap: "Severity Parameters by Raw Score"

tbl_b3_raw_score_parameters |>
  gt() |>
  tab_header(
    title = "FIES Severity Parameters by Raw Score",
    subtitle = "Rasch model transformation"
  ) |>
  cols_label(
    raw_score = "Raw Score",
    severity = "Severity (θ)",
    se = "Std. Error",
    classification = "Classification"
  ) |>
  fmt_number(columns = c(severity, se), decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = classification == "Food Secure", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(rows = classification == "Mild food insecurity", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFE0B2"),
    locations = cells_body(rows = classification == "Moderate food insecurity", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = classification == "Severe food insecurity", columns = classification)
  ) |>
  tab_source_note(
    source_note = "Classification follows FAO thresholds."
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```


## Validation

### Score Distribution Comparison

The following table shows the population distribution across raw scores and their
corresponding Rasch transformation. Most children live in households experiencing
some degree of food insecurity.

```{r}
#| label: tbl-c2-score-comparison
#| tbl-cap: "Distribution of FIES Scores: Raw vs Rasch"

tbl_c2_score_comparison |>
  gt() |>
  tab_header(
    title = "FIES Score Distribution",
    subtitle = "Raw scores mapped to Rasch severity parameters"
  ) |>
  cols_label(
    fies_raw_score = "Raw Score",
    n = "N",
    pct = "% (weighted)",
    mean_rasch = "Rasch Score (θ)",
    classification = "Classification"
  ) |>
  fmt_number(columns = pct, decimals = 1) |>
  fmt_number(columns = mean_rasch, decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = classification == "Food Secure", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(rows = classification == "Mild food insecurity", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFE0B2"),
    locations = cells_body(rows = classification == "Moderate food insecurity", columns = classification)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = classification == "Severe food insecurity", columns = classification)
  ) |>
  tab_source_note(
    source_note = "Rasch scores provide interval-scale measurement; raw scores are ordinal."
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

### Model Fit Statistics

```{r}
#| label: tbl-c3-model-fit
#| tbl-cap: "Rasch Model Fit Statistics"

tbl_c3_model_fit |>
  gt() |>
  tab_header(
    title = "Rasch Model Fit Statistics"
  ) |>
  cols_label(
    statistic = "Statistic",
    value = "Value"
  ) |>
  tab_source_note(
    source_note = "Reliability > 0.7 indicates acceptable measurement precision."
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

### Distribution Visualization

The histogram shows the distribution of Rasch-transformed food insecurity scores.
The bimodal pattern reflects distinct food secure and food insecure subpopulations.

```{r}
#| label: fig-c4-rasch-distribution
#| fig-cap: "Distribution of FIES Rasch Scores"
#| fig-alt: >
#|   Histogram of Rasch-transformed food insecurity scores (theta) on the
#|   x-axis and frequency on the y-axis, with bars shaded by food security
#|   classification from food secure through severe food insecurity. A dashed
#|   vertical line marks zero, separating less severe scores on the left from
#|   more severe scores on the right. The distribution is bimodal, with one
#|   peak of food secure children and a second peak spanning the food insecure
#|   classifications, consistent with two distinct subpopulations.
#| fig-height: 7
#| fig-width: 10

# Define color palette matching table colors
fi_colors <- c(
  "Food Secure" = "#D4EDDA",
  "Mild food insecurity" = "#FFF3CD",
  "Moderate food insecurity" = "#FFE0B2",
  "Severe food insecurity" = "#FFCDD2"
)

# Histogram of Rasch scores
ggplot(fig_c4_rasch_distribution, aes(x = fies_rasch_score, fill = classification)) +
  geom_histogram(bins = 9, color = "white", linewidth = 0.5) +
  scale_fill_manual(values = fi_colors, name = "Classification") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  annotate("text", x = -0.3, y = Inf, label = "← Less severe",
           hjust = 1, vjust = 2, size = 3, color = "gray40") +
  annotate("text", x = 0.3, y = Inf, label = "More severe →",
           hjust = 0, vjust = 2, size = 3, color = "gray40") +
  labs(
    title = "Distribution of Food Insecurity (Rasch Scale)",
    subtitle = "Children 6-59 months, SIVESNU 2018",
    x = "FIES Rasch Score (θ)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 13, color = "gray30", face = "bold"),
    plot.subtitle = element_text(size = 11, color = "gray50"),
    axis.title = element_text(size = 10, color = "gray40"),
    legend.position = "bottom"
  )
```


## Summary

### Food Security Classification

| Classification | Raw Score | Prevalence |
|----------------|-----------|------------|
| Food Secure | 0 | `r sprintf("%.1f%%", pct_food_secure)` |
| Mild food insecurity | 1-3 | `r sprintf("%.1f%%", pct_mild)` |
| Moderate food insecurity | 4-6 | `r sprintf("%.1f%%", pct_moderate)` |
| Severe food insecurity | 7-8 | `r sprintf("%.1f%%", pct_severe)` |

### Methodological Notes

1. **Rasch model advantages** — Transforms ordinal raw scores into interval-scale
   measurements with equal spacing. This enables valid arithmetic operations
   (means, regression coefficients) in subsequent modeling.

2. **Survey-weighted estimation** — Item parameters were estimated using survey
   weights to ensure population representativeness, following FAO methodology.

3. **Item ordering validated** — The estimated difficulty parameters follow the
   expected severity gradient: mild items (worry, diet quality) have negative β,
   while severe items (hunger, day without food) have positive β.

### Limitations

1. **Extreme scores** — Raw scores 0 and 8 require assumptions for parameter
   estimation (handled by RM.weights default settings).

2. **Local independence** — The Rasch model assumes conditional independence
   between items given the latent trait; residual correlations may indicate
   violation.

### Application

The Rasch-transformed FIES score enters the stunting models as a continuous
predictor, carrying the full gradient of food security into the model rather
than a small number of categorical strata.
