---
title: "Applying Transfer Models to SIVESNU"
subtitle: "Module 4: Covariate Prediction"
---

## Overview

The transfer models developed in Module 3 enable us to predict nutrient intake
for populations where detailed food consumption data is unavailable. SIVESNU 2018
(Sistema de Vigilancia Epidemiológica de Salud y Nutrición) provides anthropometric
measurements and food security indicators for Guatemalan children, but lacks the
detailed dietary recall needed to calculate nutrient intake directly.

By applying our ENCOVI-trained models to the SIVESNU data, we can:

1. **Estimate individual nutrient intake** — Predict iron, zinc, protein, lysine,
   tryptophan, and energy intake for each child based on their demographic
   and socioeconomic characteristics.

2. **Link nutrition to stunting outcomes** — Connect predicted intake with measured
   height-for-age z-scores to model the nutrition–stunting relationship.

3. **Enable biofortification simulation** — Create a baseline for comparing
   conventional vs. biofortified maize scenarios.

This module bridges the gap between ENCOVI's food consumption data and SIVESNU's
anthropometric measurements, creating a unified analytical dataset for impact
modeling.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  car,                # Yeo-Johnson likelihood profile (boxCox)
  ggpubr,             # Legend extraction utility
  cowplot,            # Multi-panel plot composition
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script (exported by framework 04_01).

tbl_b2_impute_missing <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_01_tbl_b2_impute_missing.rds"),
  trust = TRUE
)

tbl_b2b_completion_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_01_tbl_b2b_completion_summary.rds"),
  trust = TRUE
)

fig_c_yj_miembros_hogar_pre <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_01_fig_c_yj_miembros_hogar_pre.rds"),
  trust = TRUE
)

fig_c_yj_miembros_hogar_post <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_01_fig_c_yj_miembros_hogar_post.rds"),
  trust = TRUE
)

fig_c_yj_n_cuartos_pre <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_01_fig_c_yj_n_cuartos_pre.rds"),
  trust = TRUE
)

fig_c_yj_n_cuartos_post <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_01_fig_c_yj_n_cuartos_post.rds"),
  trust = TRUE
)

tbl_e2_final_validation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_01_tbl_e2_final_validation.rds"),
  trust = TRUE
)

fig_e3_density_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_01_fig_e3_density_comparison.rds"),
  trust = TRUE
)

tbl_e3_distribution_comparison <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_01_tbl_e3_distribution_comparison.rds"),
  trust = TRUE
)
```


## Data Sources

### SIVESNU 2018

SIVESNU is Guatemala's national nutrition surveillance system, collecting data on
child health and nutritional status [@INCAP2020_SIVESNU]. The 2018 round provides:

- **Anthropometric data** — Height-for-age z-scores for children 6–59 months.
- **Household characteristics** — Socioeconomic indicators matching ENCOVI structure.
- **Food security assessment** — FIES (Food Insecurity Experience Scale) responses.

SIVESNU does not include the detailed food consumption records that ENCOVI
carries, so nutrient intake cannot be calculated directly from it. This is the
gap the transfer models address.

### ENCOVI Reference Distributions

To calibrate our predictions, we use the observed nutrient distributions from
ENCOVI 2023 as reference targets, restricted to children aged 6–59 months so
that the reference describes the same population as the target. Quantile mapping
transports the predicted SIVESNU distributions onto the empirical patterns of
that reference.

Nutrient intake scales with body size, and the ENCOVI children in this age range
consume roughly half of what the all-age population does — median ratios of 0.57
for the maize-source components and 0.47–0.49 for the non-maize components.
Restricting the reference to the same age range keeps both the level and the
composition of the calibrated values on the child scale.

### Transfer Models

The 13 gamma regression models from Module 3 predict nutrient intake separately
for maize and non-maize sources. This separation enables biofortification
simulation by modifying only the maize component while preserving baseline
non-maize intake.


## Variable Harmonization

### Aligning Survey Structures

SIVESNU and ENCOVI use different variable names and coding schemes for similar
concepts. Before applying transfer models, we must harmonize SIVESNU variables
to match the exact structure expected by the ENCOVI-trained models.

This harmonization includes:

- **Factor level alignment** — Ensuring categorical variables have identical
  levels and ordering.
- **Variable renaming** — Matching ENCOVI naming conventions.
- **Derived variables** — Creating composite indicators (e.g.,
  [household parental education level](../05_meta_analysis/05_02_transfer_model_height.qmd#imp-grado-estudios-definition))
  using the same methodology.


### Missing Value Imputation

Survey data inevitably contains missing values. We apply Random Forest-based
multivariate imputation because it:

- Handles mixed data types (continuous and categorical) simultaneously.
- Preserves variable correlations and multivariate structure.
- Provides out-of-bag error estimates for quality assessment.

Implementation uses the `impute_rf_parallel()` utility function, which wraps
`missRanger` with a multi-threaded `ranger` engine and returns imputed values
alongside out-of-bag error estimates (OOBE for continuous variables, PFC for
categorical). The multi-threaded engine accelerates the iterative imputation
process without altering the underlying methodology, which remains the standard
chained Random Forest approach.

::: {.callout-note}
## About out-of-bag error
Out-of-bag (OOB) error estimates are calculated from observations that were
randomly excluded during tree construction within the Random Forest algorithm,
providing an internal quality metric for imputed values without requiring a
separate test set.
:::

```{r}
#| label: tbl-b2-impute-missing
#| tbl-cap: "Random Forest imputation quality metrics"

tbl_b2_impute_missing |>
  gt() |>
  tab_header(
    title = "Missing Value Imputation Quality Assessment",
    subtitle = "Out-of-bag error estimates from Random Forest imputation (includes FIES variables)"
  ) |>
  cols_label(
    metric = "Metric",
    value = "Value",
    interpretation = "Quality Assessment"
  ) |>
  fmt_number(
    columns = value,
    decimals = 4
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "Acceptable"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "Review required"
    )
  ) |>
  tab_footnote(
    footnote = "OOBE: aggregated out-of-bag error across the continuous variables imputed, bounded in [0, 1] and invariant to their scale. Values < 0.5 indicate acceptable imputation quality for continuous variables.",
    locations = cells_body(columns = metric, rows = 1)
  ) |>
  tab_footnote(
    footnote = "PFC: Proportion of Falsely Classified entries. Values < 0.2 indicate acceptable imputation quality for categorical variables.",
    locations = cells_body(columns = metric, rows = 2)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

::: {.callout-tip}
## Imputation Quality
OOBE < 0.5 and PFC < 0.2 are the ranges conventionally treated as acceptable
for this imputation method.
:::


### Completion of Residual Missing Values

The Random Forest imputation engine can leave a residual `NaN` for a small
subset of rows when the out-of-bag prediction collapses inside a terminal node
with zero within-leaf variance. This affects continuous discrete-scale
predictors.

Downstream analyses (Yeo-Johnson transformation, model prediction, quantile
mapping) operate on a complete predictor matrix, so residual missing values are
completed with a variable-specific rule:

- **Numeric / integer variables**: median of the observed values, which
  preserves the support of discrete-scale ordinal variables (e.g. years of
  schooling take values such as 1, 1.5, 2, 2.5, ...).
- **Factor variables**: mode of the observed values.

Completing rather than discarding keeps every record in the predicted
distributions. The count of values completed by each rule is reported in the
table below.

```{r}
#| label: tbl-b2b-completion-summary
#| tbl-cap: "Residual missing value completion after Random Forest imputation"

tbl_b2b_completion_summary |>
  gt() |>
  tab_header(
    title = "Residual Missing Value Completion",
    subtitle = "Values completed by variable-specific rule after RF imputation"
  ) |>
  cols_label(
    variable = "Variable",
    n_completed = "N Completed",
    rule = "Completion Rule"
  ) |>
  fmt_number(
    columns = n_completed,
    decimals = 0,
    use_seps = TRUE
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      rows = rule == "Integrity check: passed"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(
      rows = rule == "Integrity check: review required"
    )
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```


## Applying Transformations

### Yeo-Johnson Normalization

The transfer models were trained on Yeo-Johnson transformed continuous predictors.
We apply the same transformations (using parameters fitted on ENCOVI) to ensure
SIVESNU data enters the models in the expected scale.

```{r}
#| label: fig-c-yj-miembros-hogar-pre
#| fig-cap: "Yeo-Johnson transformation profile for household members (pre-transformation)"
#| fig-alt: >
#|   Profile log-likelihood curve for the Yeo-Johnson power parameter lambda,
#|   with lambda on the x-axis and log-likelihood on the y-axis, computed on the
#|   untransformed household-members variable. Dashed vertical lines mark the
#|   maximum-likelihood lambda and its 95% confidence interval, indicating the
#|   power needed to normalize the raw, right-skewed distribution.

boxCox(
  lm(miembros_hogar ~ 1, data = fig_c_yj_miembros_hogar_pre),
  main = "Household Members - Pre-Transformation",
  family = "yjPower"
)
```

```{r}
#| label: fig-c-yj-miembros-hogar-post
#| fig-cap: "Yeo-Johnson transformation profile for household members (post-transformation)"
#| fig-alt: >
#|   Profile log-likelihood curve for the Yeo-Johnson power parameter lambda,
#|   with lambda on the x-axis and log-likelihood on the y-axis, computed on the
#|   transformed household-members variable. The maximum-likelihood lambda now
#|   sits close to 1, and its 95% confidence interval brackets 1, the value
#|   expected for an approximately normal distribution.

boxCox(
  lm(miembros_hogar_transformed ~ 1, data = fig_c_yj_miembros_hogar_post),
  main = "Household Members - Post-Transformation",
  family = "yjPower"
)
```

```{r}
#| label: fig-c-yj-n-cuartos-pre
#| fig-cap: "Yeo-Johnson transformation profile for number of rooms (pre-transformation)"
#| fig-alt: >
#|   Profile log-likelihood curve for the Yeo-Johnson power parameter lambda,
#|   with lambda on the x-axis and log-likelihood on the y-axis, computed on the
#|   untransformed number-of-rooms variable. Dashed vertical lines mark the
#|   maximum-likelihood lambda and its 95% confidence interval, identifying the
#|   power needed to normalize the raw, skewed distribution.

boxCox(
  lm(n_cuartos ~ 1, data = fig_c_yj_n_cuartos_pre),
  main = "Number of Rooms - Pre-Transformation",
  family = "yjPower"
)
```

```{r}
#| label: fig-c-yj-n-cuartos-post
#| fig-cap: "Yeo-Johnson transformation profile for number of rooms (post-transformation)"
#| fig-alt: >
#|   Profile log-likelihood curve for the Yeo-Johnson power parameter lambda,
#|   with lambda on the x-axis and log-likelihood on the y-axis, computed on the
#|   transformed number-of-rooms variable. The maximum-likelihood lambda now sits
#|   close to 1, and its 95% confidence interval brackets 1, the value expected
#|   for an approximately normal distribution.

boxCox(
  lm(n_cuartos_transformed ~ 1, data = fig_c_yj_n_cuartos_post),
  main = "Number of Rooms - Post-Transformation",
  family = "yjPower"
)
```


## Nutrient Prediction

### Model Application

Each transfer model predicts nutrient intake on the log scale (via log link in
gamma regression). We apply **exponential back-transformation** to obtain
predictions in natural units (mg/day or g/day).

::: {.callout-note}
## Link Functions
All nutrient models use log link, so their predictions require exponential
back-transformation.
:::

### Food Security Adjustment

SIVESNU includes the Food Insecurity Experience Scale (FIES), an 8-question
instrument measuring household food security. FIES scores are used to apply a
fine adjustment to nutrient intake predictions before quantile mapping:

| FIES Score | Food Security Status | Description | Adjustment |
|------------|---------------------|-------------|------------|
| 0 | Food secure | Adequate food access | +5% |
| 1–3 | Mild food insecurity | Slight decrease in quantity and/or quality of food consumed | +2% |
| 4–6 | Moderate food insecurity | Moderate reduction in quantity and/or quality of food consumed | −2% |
| 7–8 | Severe food insecurity | Have run out of food for a day or more | −5% |

**Why the adjustment is applied.** The transfer models predict nutrient intake
from demographic and socioeconomic characteristics. Two children sharing the
same socioeconomic profile therefore receive identical predicted intake values,
regardless of differences in their actual food security situation. Quantile
mapping alone cannot separate these two children, because it preserves the ranks
it receives — if the inputs are identical, the outputs remain identical.

FIES is used as an independent proxy for the household food environment. Children
in more food-insecure households start from a slightly lower intake value (−2% or
−5%) than children in more food-secure households (+2% or +5%), even when the
transfer model assigned them the same starting point. The adjustment is applied
before quantile mapping, so that the subsequent rank-based calibration operates
on already-differentiated values rather than on identical ones.

**Origin of the adjustment factors.** The factors (+5% / +2% / −2% / −5%) are a
modeling parameter of this framework. They were set by inspecting the plausible
intake distributions across food security strata and choosing modest, symmetric
adjustments that introduce the intended differentiation without distorting the
overall distribution.

**Validation.** What is validated is the final intake distribution, after the
FIES adjustment, the quantile mapping and the maize-source alignment have all
been applied, against the age-matched ENCOVI reference. That comparison is
reported at the end of this page.

::: {.callout-note}
## Role of the Adjustment
The FIES adjustment is a fine-tuning step that introduces a food-security
gradient into predictions otherwise driven only by demographic and socioeconomic
covariates. Its modest magnitude (at most ±5%) reflects this secondary,
corrective role. Its purpose is to improve discrimination between children with
similar socioeconomic profiles but different food security status, while avoiding
the circularity of using the stunting outcome as an input.
:::

### Target Population Filtering

The stunting analysis focuses on children aged 6–59 months, the standard window
for assessing chronic malnutrition. Prediction runs on all SIVESNU individuals,
and the filter to this age range is applied **before** the calibration step that
follows.

The order matters. Quantile mapping transports the quantiles of a source
distribution onto those of a reference, so the two need to describe comparable
populations. Filtering first leaves the source on the child scale, matching the
ENCOVI reference that was already restricted to the same age range.

### Quantile Mapping Calibration

Model predictions do not match the observed ENCOVI distributions exactly.
Quantile mapping aligns the predicted distribution to the reference
distribution, preserving ranks while correcting systematic level differences.
The mapping is estimated over the pooled age-matched range.

### Maize-Source Alignment

A rank-based alignment is then applied to the maize-source variables, which
realigns each individual's maize-source values against their overall non-maize
nutrient profile.

::: {.callout-note}
## Assumption Behind the Alignment
The alignment rests on intra-individual nutrient coherence: an individual's
maize-source intake is expected to sit consistently with the rest of their
nutrient profile. Maize dependency is itself a marker of household resources,
so before the alignment the maize-source block and the stunting outcome share
that socioeconomic gradient.
:::

### Preserved Raw Predictions

The exported dataset carries two parallel blocks of the twelve source-specific
nutrient variables. The `_raw` block holds the transfer-model output before the
FIES adjustment, the quantile mapping and the maize-source alignment; the
adjusted block holds the calibrated values that the rest of the pipeline
consumes. Each raw column pairs one-to-one with its calibrated counterpart. The
raw block is carried unchanged through the remaining scripts of this module and
takes no part in any calculation: it exists so that the effect of the
calibration can be audited at any stage.


## Final Dataset

### Variable Selection

The final dataset contains all variables needed for stunting modeling and
biofortification scenario simulation.

```{r}
#| label: tbl-e2-final-validation
#| tbl-cap: "Final dataset structure"

tbl_e2_final_validation |>
  gt() |>
  tab_header(
    title = "Final Dataset Structure",
    subtitle = "SIVESNU 2018 with predicted nutrient intake values"
  ) |>
  cols_label(
    category = "Variable Category",
    n_variables = "N Variables",
    observations = "N Observations"
  ) |>
  fmt_number(
    columns = c(n_variables, observations),
    decimals = 0,
    use_seps = TRUE
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = category == "TOTAL")
  ) |>
  tab_options(
    table.font.size = 11,
    table.width = pct(100)
  )
```


### Distribution Validation Against the Reference

The figure and table below compare the distribution of calibrated nutrient
intakes for SIVESNU children (6–59 months) against the ENCOVI children they
were mapped onto. Both sides cover the same age range, so this is a
like-for-like comparison.

The figure compares total intake per nutrient, where the two populations track
each other closely (median ratios between 0.90 and 1.00). The table separates
the two source blocks, which behave differently by construction: the non-maize
components leave the quantile mapping untouched and sit within 3–8% of the
reference medians, with Kolmogorov-Smirnov statistics below 0.075, while the
maize-source components sit 14–20% above the reference medians, being the only
block modified after the mapping by the rank-based alignment described above.

```{r}
#| label: fig-e3-density-comparison
#| fig-cap: "Distribution comparison: SIVESNU children (6-59m, calibrated) vs ENCOVI children (6-59m, observed)"
#| fig-alt: >
#|   Grid of six overlaid density plots, one panel per nutrient: iron, zinc,
#|   digestible protein, lysine, tryptophan, and energy, each with intake on the
#|   x-axis and density on the y-axis. Two filled curves per panel compare the
#|   age-matched ENCOVI 2023 children's reference against the SIVESNU 2018
#|   calibrated children's intakes. The paired curves overlap closely in every
#|   panel, with the SIVESNU curves sitting marginally lower for digestible
#|   protein and lysine.
#| fig-width: 10
#| fig-height: 8

# Define color palette and labels
color_palette <- c(
  "ENCOVI 2023 (Reference)" = "#5C946E",
  "SIVESNU 2018 (Predicted)" = "#E07A5F"
)

nutrients <- c("fe_total", "zn_total", "prot_total", "lys_total", "trp_total", "kcal_total")
nutrient_labels <- c(
  "fe_total" = "Iron (mg/day)",
  "zn_total" = "Zinc (mg/day)",
  "prot_total" = "Digestible Protein (g/day)",
  "lys_total" = "Lysine (mg/day)",
  "trp_total" = "Tryptophan (mg/day)",
  "kcal_total" = "Energy (kcal/day)"
)

# Create individual density plots
plot_list <- map(nutrients, ~ {
  upper_limit <- quantile(fig_e3_density_comparison[[.x]], 0.95, na.rm = TRUE)

  fig_e3_density_comparison |>
    filter(.data[[.x]] > 0, .data[[.x]] <= upper_limit) |>
    ggplot(aes(x = .data[[.x]], fill = source)) +
    geom_density(alpha = 0.6, color = NA) +
    scale_fill_manual(values = color_palette, drop = FALSE) +
    labs(title = nutrient_labels[.x], x = NULL, y = "Density") +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 11, color = "azure4", face = "bold"),
      axis.title = element_text(size = 9, color = "azure4"),
      axis.text = element_text(size = 8, color = "azure4"),
      legend.position = "none",
      panel.grid.minor = element_blank()
    )
})

# Create a plot just to extract the legend
legend_plot <- fig_e3_density_comparison |>
  ggplot(aes(x = fe_total, fill = source)) +
  geom_density(alpha = 0.6) +
  scale_fill_manual(values = color_palette, drop = FALSE) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 11, color = "azure4")
  ) +
  guides(fill = guide_legend(override.aes = list(alpha = 0.6)))

# Extract legend and arrange grid
legend_grob <- ggpubr::get_legend(legend_plot)
plot_grid <- cowplot::plot_grid(plotlist = plot_list, ncol = 3, nrow = 2)
cowplot::plot_grid(plot_grid, legend_grob, ncol = 1, rel_heights = c(1, 0.08))
```

```{r}
#| label: tbl-e3-distribution-comparison
#| tbl-cap: "Distribution comparison by source block: SIVESNU children (6-59m) vs age-matched ENCOVI children"

tbl_e3_distribution_comparison |>
  mutate(
    Maize_p_formatted = case_when(
      Maize_p_value < 0.001 ~ "< 0.001",
      Maize_p_value < 0.01 ~ "< 0.01",
      Maize_p_value < 0.05 ~ "< 0.05",
      TRUE ~ sprintf("%.3f", Maize_p_value)
    ),
    `Non-Maize_p_formatted` = case_when(
      `Non-Maize_p_value` < 0.001 ~ "< 0.001",
      `Non-Maize_p_value` < 0.01 ~ "< 0.01",
      `Non-Maize_p_value` < 0.05 ~ "< 0.05",
      TRUE ~ sprintf("%.3f", `Non-Maize_p_value`)
    )
  ) |>
  select(-Maize_p_value, -`Non-Maize_p_value`) |>
  gt() |>
  tab_header(
    title = "Nutrient Intake Distributions — Children 6-59 months, SIVESNU vs ENCOVI",
    subtitle = "Medians and Kolmogorov-Smirnov statistic against the age-matched calibration reference"
  ) |>
  cols_label(
    nutrient = "Nutrient",
    Maize_median_encovi = "ENCOVI",
    Maize_median_sivesnu = "SIVESNU",
    Maize_ratio = "Ratio",
    Maize_ks_statistic = "KS",
    Maize_p_formatted = "p-value",
    `Non-Maize_median_encovi` = "ENCOVI",
    `Non-Maize_median_sivesnu` = "SIVESNU",
    `Non-Maize_ratio` = "Ratio",
    `Non-Maize_ks_statistic` = "KS",
    `Non-Maize_p_formatted` = "p-value"
  ) |>
  tab_spanner(label = "From Maize", columns = starts_with("Maize")) |>
  tab_spanner(label = "From Non-Maize", columns = starts_with("Non-Maize")) |>
  fmt_number(columns = contains("median"), decimals = 1) |>
  fmt_number(columns = contains("ratio"), decimals = 2) |>
  fmt_number(columns = contains("ks_statistic"), decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = Maize_ratio, rows = Maize_ratio >= 0.9 & Maize_ratio <= 1.1)
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = `Non-Maize_ratio`, rows = `Non-Maize_ratio` >= 0.9 & `Non-Maize_ratio` <= 1.1)
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = Maize_ks_statistic, rows = Maize_ks_statistic < 0.1)
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = `Non-Maize_ks_statistic`, rows = `Non-Maize_ks_statistic` < 0.1)
  ) |>
  tab_footnote(
    footnote = "Ratio = SIVESNU median / ENCOVI median, both for children 6-59 months. Green marks ratios within 0.9 to 1.1. The maize-source block sits above that band because the rank-based alignment is applied after the quantile mapping.",
    locations = cells_column_labels(columns = Maize_ratio)
  ) |>
  tab_footnote(
    footnote = "KS: Kolmogorov-Smirnov statistic on the distribution shape. Green marks values below 0.1.",
    locations = cells_column_labels(columns = Maize_ks_statistic)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

::: {.callout-tip}
## Reading the Validation
Both sides of the comparison are children aged 6–59 months, so the medians and
the Kolmogorov-Smirnov statistics measure how closely the calibrated SIVESNU
distributions reproduce the reference they were mapped onto.

The non-maize block is the one that reaches the reference directly: its ratios
fall inside the 0.9–1.1 band and its Kolmogorov-Smirnov statistics stay below
0.1. The maize-source block is realigned after the mapping, so its ratios sit
above that band by design.
:::


## Summary

### Processing Pipeline

| Step | Method | Purpose |
|------|--------|---------|
| Variable harmonization | Factor recoding | Match ENCOVI predictor structure |
| Missing imputation | Random Forest (`impute_rf_parallel`) | Handle incomplete records |
| Residual completion | Median / mode | Complete the predictor matrix |
| Transformation | Yeo-Johnson | Normalize continuous predictors |
| Prediction | Gamma GLM | Estimate nutrient intake |
| Target population filter | Children 6–59 months | Match the age range of the reference |
| FIES adjustment | Score-based scaling | Incorporate food security |
| Quantile mapping | Distribution alignment | Calibrate to the age-matched ENCOVI reference |
| Maize-source alignment | Rank-based realignment | Place the maize block on the individual's nutrient profile |

### Key Findings

1. **Age-matched calibration** — With the target population filtered before the
   mapping and the reference restricted to the same age range, the non-maize
   block reproduces the reference medians within 3–8%, with Kolmogorov-Smirnov
   statistics below 0.075.

2. **Maize block offset by design** — The maize-source components sit 14–20%
   above the reference medians, the footprint of the rank-based alignment
   applied after the mapping.

3. **Food security differentiation** — The FIES adjustment separates children
   whose demographic and socioeconomic profiles are identical, which the
   transfer models alone map to the same predicted intake.

### Application

This dataset provides the nutrient intake foundation for:

- **[Chispitas integration](04_02_integrate_chispitas_supplementation.qmd)** — Adding micronutrient supplementation effects.
- **[Bioavailability calculations](04_03_calculate_bioavailability_adjustments.qmd)** — Converting intake to absorbed nutrients.
- **[Stunting model development](04_06_model_stunting_continuous_zlen.qmd)** — Linking intake to HAZ outcomes.
- **[Scenario simulation](../06_scenarios/06_01_calculate_market_baseline.qmd)** — Modeling biofortified maize impact.
