---
title: "Preparing Stunting Modeling Data"
subtitle: "Module 4: Covariate Prediction"
description-meta: "Merging SIVESNU anthropometry with predicted intake, then block PCA, correlation filtering, Boruta and VIF checks to select stunting predictors."
---

## Overview

Predicting child stunting requires combining anthropometric outcomes with
demographic, socioeconomic, and health-related predictors. Survey datasets like
SIVESNU contain hundreds of potential variables, but effective modeling requires
a curated subset that balances predictive power with parsimony.

The challenge is threefold: (1) many variables have missing values that could
bias results, (2) groups of related variables are highly correlated and
redundant, and (3) not all variables contribute meaningful predictive
information. This module implements a systematic, transparent selection pipeline
that addresses each challenge while keeping every transformation traceable.

Before the merge, each SIVESNU source table is independently audited against an
operational data dictionary that declares the universe of valid values for every
variable. Out-of-domain values are recoded to NA so that all downstream stages
operate on data whose semantic domain is fully under control.

By combining multiple SIVESNU 2018 tables — child anthropometrics, household
characteristics, maternal factors, and the nutrient intake estimates from
previous modules — we build a unified analytical dataset optimized for stunting
prediction.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  scales,             # Number formatting (percent, comma)
  patchwork,          # Multi-panel plot composition for scree plots
  tidyverse           # General purpose data manipulation
)

# Set display options
options(
  scipen = 999,
  digits = 4,
  dplyr.summarise.inform = FALSE
)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script
# (exported by framework 04_05).

# --- Phase A: Data integration audit ---
tbl_a3_audit_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_a3_audit_summary.rds"),
  trust = TRUE
)

tbl_a3_audit_incidents <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_a3_audit_incidents.rds"),
  trust = TRUE
)

tbl_a4_merge_validation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_a4_merge_validation.rds"),
  trust = TRUE
)

# --- Phase B: Candidate variables ---
tbl_b6_candidates_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_b6_candidates_summary.rds"),
  trust = TRUE
)

# --- Phase C: Quality filters ---
tbl_c1_missing_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_c1_missing_summary.rds"),
  trust = TRUE
)

tbl_c2_nzv_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_c2_nzv_summary.rds"),
  trust = TRUE
)

tbl_c3_phase_c_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_c3_phase_c_summary.rds"),
  trust = TRUE
)

# --- Phase D: Imputation ---
tbl_d1_pre_imputation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_d1_pre_imputation.rds"),
  trust = TRUE
)

tbl_d2_imputation_results <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_d2_imputation_results.rds"),
  trust = TRUE
)

tbl_d2b_completion_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_d2b_completion_summary.rds"),
  trust = TRUE
)

tbl_d3_validation_imputation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_d3_validation_imputation.rds"),
  trust = TRUE
)

tbl_d4_phase_d_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_d4_phase_d_summary.rds"),
  trust = TRUE
)

# --- Phase E: Block PCA ---
tbl_e1_pca_blocks <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_e1_pca_blocks.rds"),
  trust = TRUE
)

tbl_e2_pca_results <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_e2_pca_results.rds"),
  trust = TRUE
)

fig_e2_scree_plots <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_fig_e2_scree_plots.rds"),
  trust = TRUE
)

tbl_e3_pca_update_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_e3_pca_update_summary.rds"),
  trust = TRUE
)

# --- Phase F: Feature selection ---
tbl_f1_correlation_filter <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_f1_correlation_filter.rds"),
  trust = TRUE
)

tbl_f2_boruta_results <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_f2_boruta_results.rds"),
  trust = TRUE
)

fig_f2_boruta_importance <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_fig_f2_boruta_importance.rds"),
  trust = TRUE
)

tbl_f3_vif_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_f3_vif_summary.rds"),
  trust = TRUE
)

tbl_f3_vif_results <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_f3_vif_results.rds"),
  trust = TRUE
)

tbl_f3_vif_dropped <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_f3_vif_dropped.rds"),
  trust = TRUE
)

# --- Phase G: Output summary ---
tbl_g4_output_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_g4_output_summary.rds"),
  trust = TRUE
)

tbl_g5_pipeline_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_05_tbl_g5_pipeline_summary.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Cifras extracted for narrative use throughout the document.

# --- A3 Audit aggregates ---
n_clean_total <- sum(tbl_a3_audit_summary$n_clean)
n_recoded_total <- sum(tbl_a3_audit_summary$n_recoded)
n_passthrough_total <- sum(tbl_a3_audit_summary$n_passthrough)
n_not_in_dict_total <- sum(tbl_a3_audit_summary$n_not_in_dict)
n_values_recoded_total <- sum(tbl_a3_audit_summary$values_recoded)

# --- A4 Merge ---
n_observations_final <- tbl_a4_merge_validation |>
  filter(Stage == "With valid outcomes (HFA & ZLEN)") |>
  pull(Records)

# --- D2 Imputation OOB ---
oob_oobe <- tbl_d2_imputation_results |>
  filter(str_detect(Metric, "OOBE")) |>
  pull(Value)

oob_pfc <- tbl_d2_imputation_results |>
  filter(str_detect(Metric, "PFC")) |>
  pull(Value)

# --- F2 Boruta selected count ---
n_boruta_selected <- tbl_f2_boruta_results |>
  filter(Metric == "Variables selected (Confirmed + Tentative)") |>
  pull(Value) |>
  as.integer()

# --- F3 VIF on the retained set, and the variables the pruning removed ---
n_vif_retained <- nrow(tbl_f3_vif_results)
max_vif_retained <- round(max(tbl_f3_vif_results$vif), 2)
n_vif_dropped <- nrow(tbl_f3_vif_dropped)
vars_vif_dropped <- paste0("`", tbl_f3_vif_dropped$variable, "`", collapse = ", ")

# --- F3 VIF acceptable ---
n_vif_acceptable <- tbl_f3_vif_summary |>
  filter(Metric == "Acceptable (VIF ≤ 5)") |>
  pull(Value) |>
  as.integer()

n_vif_evaluated <- tbl_f3_vif_summary |>
  filter(Metric == "Variables evaluated") |>
  pull(Value) |>
  as.integer()
```


## Target Population

Children aged 6-59 months with valid anthropometric measurements. Children
under 6 months are excluded because exclusive breastfeeding is recommended
during this period, making dietary intake patterns non-comparable to older
children who consume complementary foods.


## Data Integration Strategy

The base dataset — containing nutrient intake estimates, bioavailability
adjustments, and FIES scores from previous pipeline steps — is merged with
additional SIVESNU 2018 tables to create a comprehensive predictor set:

| Table | Contents | Purpose |
|-------|----------|---------|
| Child | stunting, HAZ, vaccination, health controls, illness history | Outcomes and child-level predictors |
| Household | Assets, infrastructure, cooking fuel, altitude | Socioeconomic predictors |
| Women | Maternal education, age, physiological state | Maternal predictors |
| Members | Occupation classification | Agricultural exposure indicators |


### Data Dictionary Audit

Each source table is audited independently against an operational data
dictionary that declares, for every variable in the table, its type (factor,
binary, numeric continuous, numeric integer, identifier, date, text) and its
universe of valid values (factor levels for categorical variables, plausible
numeric bounds for continuous and integer variables). Character columns are
loaded with whitespace trimming applied uniformly, normalizing fixed-width
string padding inherited from the original survey exports.

The audit applies a single rule: any value outside the declared universe is
replaced with NA. Empty strings in character columns are normalized to NA
before the comparison. The rule is applied uniformly regardless of whether
out-of-domain values correspond to documented non-response codes, sentinel
values, or artefacts from upstream transformations. Each table is audited
against its own dictionary subset, so variables that share names across tables
but use different encodings are validated against the universe corresponding
to their actual source.

#### Audit Summary by Source Table

The audit returns four outcomes per variable:

- **Clean**: no out-of-domain values, no recoding required.
- **Recoded**: variable in the dictionary, at least one value fell outside the declared domain and was replaced with NA.
- **Passthrough**: identifier, date, or text variable explicitly exempted from domain validation.
- **Not in dictionary**: variable present in the data but absent from the dictionary, passed through unchanged and flagged for review.

```{r}
#| label: tbl-a3-audit-summary
#| tbl-cap: "Audit outcome summary aggregated across the four SIVESNU tables"

tbl_a3_audit_summary |>
  gt() |>
  tab_header(
    title = "Variable Audit Summary by Source Table",
    subtitle = "Outcome distribution after audit against operational dictionary"
  ) |>
  fmt_number(
    columns = c(n_clean, n_recoded, n_passthrough, n_not_in_dict, values_recoded),
    use_seps = TRUE, decimals = 0
  ) |>
  cols_label(
    source_table = "Table",
    n_clean = "Clean",
    n_recoded = "Recoded",
    n_passthrough = "Passthrough",
    n_not_in_dict = "Not in dict.",
    values_recoded = "Values recoded"
  ) |>
  cols_align(align = "left", columns = source_table) |>
  cols_align(align = "right",
             columns = c(n_clean, n_recoded, n_passthrough, n_not_in_dict, values_recoded)) |>
  tab_source_note(
    md("**Audit rule:** values outside the declared domain (factor levels for categorical, [min, max] for numeric) are replaced with NA. Variables flagged 'Not in dict.' should be reviewed for inclusion.")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```

#### Variables With Recoding Incidents

This table lists variables where the audit recoded at least one value, sorted
by number of values replaced. Variables not in the dictionary appear in a
separate block at the end and are highlighted for review.

```{r}
#| label: tbl-a3-audit-incidents
#| tbl-cap: "Per-variable audit incidents across the four SIVESNU tables"

tbl_a3_audit_incidents |>
  gt() |>
  tab_header(
    title = "Per-Variable Audit Incidents",
    subtitle = "Variables with out-of-domain recodings and variables absent from the dictionary"
  ) |>
  fmt_number(columns = c(n_total, n_pre_na, n_recoded, n_post_na),
             use_seps = TRUE, decimals = 0) |>
  fmt_number(columns = pct_recoded, decimals = 2, pattern = "{x}%") |>
  sub_missing(missing_text = "—") |>
  cols_label(
    source_table = "Table",
    variable = "Variable",
    type = "Type",
    n_total = "N Total",
    n_pre_na = "NA (pre)",
    n_recoded = "Recoded",
    n_post_na = "NA (post)",
    pct_recoded = "% Recoded",
    audit_status = "Status"
  ) |>
  cols_align(align = "left", columns = c(source_table, variable, type, audit_status)) |>
  cols_align(align = "right",
             columns = c(n_total, n_pre_na, n_recoded, n_post_na, pct_recoded)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = variable)
  ) |>
  tab_style(
    style = cell_fill(color = "#ffffcc"),
    locations = cells_body(rows = audit_status == "not_in_dictionary")
  ) |>
  tab_source_note(
    md("**Recoded:** variable in the dictionary, the indicated count of values fell outside the declared domain and was replaced with NA.<br>**Not in dictionary:** variable present in the data but absent from the operational dictionary, passed through unchanged.")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```


### Merge Validation

Validate the merge process by examining record counts and match rates at each
integration stage. The base dataset already contains the target population
(children 6-59 months), so no additional age filtering is required.

```{r}
#| label: tbl-a4-merge-validation
#| tbl-cap: "Merge process validation summary"

tbl_a4_merge_validation |>
  gt() |>
  tab_header(
    title = "Phase A: Data Integration Summary",
    subtitle = "Record counts at each merge stage (audited tables)"
  ) |>
  fmt_number(columns = Records, use_seps = TRUE, decimals = 0) |>
  fmt_number(columns = Change, use_seps = TRUE, decimals = 0) |>
  sub_missing(missing_text = "—") |>
  cols_label(
    Stage = "Integration Stage",
    Records = "N Records",
    Change = "Change",
    Pct_of_base = "% of Base"
  ) |>
  cols_align(align = "left", columns = Stage) |>
  cols_align(align = "right", columns = c(Records, Change, Pct_of_base)) |>
  tab_source_note(
    md("**Note:** Base dataset duplicates (same nboletahogar/conglomerado/edad_an/sexo) are resolved during child-level merge. Source tables are audited individually before this merge.")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```


## Variable Selection Pipeline

Survey datasets often contain hundreds of variables, many of which are
redundant, have excessive missing values, or provide minimal predictive value.
The selection pipeline narrows that pool to a parsimonious, informative
predictor set.

The pipeline follows seven sequential steps:

| Step | Method | Purpose |
|------|--------|---------|
| 1 | Missing filter (>35%) | Exclude unreliable variables |
| 2 | Near-zero variance | Remove constant/near-constant variables |
| 3 | Random Forest imputation | Handle remaining missing values |
| 4 | Block PCA | Reduce dimensionality of correlated groups |
| 5 | Correlation filter (r>0.85) | Remove redundant predictors |
| 6 | Boruta selection | Identify relevant predictors |
| 7 | VIF verification | Confirm no multicollinearity (VIF < 5) |

::: {.callout-note}
## Protected Variables
Two variable sets are protected throughout the selection process: **assignment
variables** (required for biofortified maize targeting: department,
urban/rural, occupation, FIES score) and **nutritional variables** (the
intake estimates that will be modified in biofortification scenarios). These
participate in selection for assessment but are always retained regardless
of results.
:::


### Candidate Variables Summary

```{r}
#| label: tbl-b6-candidates-summary
#| tbl-cap: "Candidate variables for selection pipeline"

tbl_b6_candidates_summary |>
  gt() |>
  tab_header(
    title = "Phase B: Candidate Variable Summary",
    subtitle = "Variable counts before and after exclusions"
  ) |>
  fmt_number(columns = Count, use_seps = TRUE, decimals = 0) |>
  cols_align(align = "left", columns = Category) |>
  cols_align(align = "right", columns = Count) |>
  tab_row_group(
    label = "Exclusions",
    rows = 2:4
  ) |>
  tab_row_group(
    label = "Final",
    rows = c(1, 5)
  ) |>
  row_group_order(groups = c("Final", "Exclusions")) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```


## Missing Values Analysis

Variables with excessive missing values (>35%) are excluded before imputation.
This threshold balances information retention with imputation reliability —
variables with higher missingness cannot be reliably imputed without
introducing substantial uncertainty.

```{r}
#| label: tbl-c1-missing-summary
#| tbl-cap: "Missing values analysis summary"

tbl_c1_missing_summary |>
  gt() |>
  tab_header(
    title = "Missing Values Analysis",
    subtitle = "Distribution of candidate variables by missing rate"
  ) |>
  cols_align(align = "left", columns = Category) |>
  cols_align(align = "right", columns = Count) |>
  tab_style(
    style = cell_fill(color = "#ffcccc"),
    locations = cells_body(rows = Category == "High missing (>35%)")
  ) |>
  tab_source_note(
    md("**Threshold:** Variables with >35% missing will be excluded from selection pipeline")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```


## Near-Zero Variance Filter

Variables with minimal variation provide little discriminatory power for
prediction. The near-zero variance filter identifies variables where one value
dominates (frequency ratio > 19) or there are very few unique values relative
to sample size (< 10%).

```{r}
#| label: tbl-c2-nzv-summary
#| tbl-cap: "Near-zero variance analysis summary"

tbl_c2_nzv_summary |>
  gt() |>
  tab_header(
    title = "Near-Zero Variance Analysis",
    subtitle = "Distribution of variables by variance characteristics"
  ) |>
  cols_align(align = "left", columns = Category) |>
  cols_align(align = "right", columns = Count) |>
  tab_style(
    style = cell_fill(color = "#ffcccc"),
    locations = cells_body(rows = Category %in% c("Zero variance", "Near-zero variance"))
  ) |>
  tab_source_note(
    md("**Criteria:** freqRatio > 19 and percentUnique < 10%")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```

```{r}
#| label: tbl-c3-phase-c-summary
#| tbl-cap: "Phase C filtering summary"

tbl_c3_phase_c_summary |>
  gt() |>
  tab_header(
    title = "Phase C: Variable Quality Filtering",
    subtitle = "Sequential application of missing value and variance filters"
  ) |>
  fmt_number(columns = Count, use_seps = TRUE, decimals = 0) |>
  cols_align(align = "left", columns = Stage) |>
  cols_align(align = "right", columns = Count) |>
  tab_row_group(
    label = md("*Missing Values Filter*"),
    rows = 1:3
  ) |>
  tab_row_group(
    label = md("*Near-Zero Variance Filter*"),
    rows = 4:5
  ) |>
  row_group_order(groups = c("*Missing Values Filter*", "*Near-Zero Variance Filter*")) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Final candidate variables for modeling")
  ) |>
  tab_source_note(
    md("**Thresholds:** Missing >35% excluded; NZV per caret::nearZeroVar defaults")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```


## Missing Value Imputation

Variables that passed the quality filters may still have missing values (up to
35%). Random Forest-based multivariate imputation handles these gaps by
predicting missing values from observed patterns across all variables,
preserving complex relationships in the data.

Implementation uses the `impute_rf_parallel()` utility function, which wraps
`missRanger` with a multi-threaded `ranger` engine and returns imputed values
alongside out-of-bag error estimates (OOBE for continuous variables, PFC for
categorical).

```{r}
#| label: tbl-d1-pre-imputation
#| tbl-cap: "Missing values summary before imputation"

tbl_d1_pre_imputation |>
  gt() |>
  tab_header(
    title = "Pre-Imputation Missing Values Assessment",
    subtitle = "Candidate variables after Phase C filtering"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_source_note(
    md("**Method:** Random Forest-based multivariate imputation")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```

```{r}
#| label: tbl-d2-imputation-results
#| tbl-cap: "Random Forest imputation results"

tbl_d2_imputation_results |>
  gt() |>
  tab_header(
    title = "Random Forest Imputation Results",
    subtitle = "Out-of-bag error estimates"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_source_note(
    md("**OOBE:** aggregated out-of-bag error, continuous variables; acceptable below 0.50<br>**PFC:** Proportion Falsely Classified, categorical variables; acceptable below 0.20")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```


### Completion of Residual Missing Values

The Random Forest imputation engine can leave a residual `NaN` for a small
subset of rows when the out-of-bag prediction collapses inside a terminal node
with zero within-leaf variance. This affects continuous discrete-scale
predictors.

Downstream analyses (block PCA, correlation filter, Boruta, VIF) operate on a
complete predictor matrix, so residual missing values are completed with a
variable-specific rule:

- **Numeric / integer variables**: median of the observed values. The median preserves the support of discrete-scale ordinal variables, which the mean does not.
- **Factor variables**: mode of the observed values.

Completing rather than discarding keeps every record in the modeling dataset.
The count of values completed by each rule is reported in the table below.

```{r}
#| label: tbl-d2b-completion-summary
#| tbl-cap: "Residual missing value completion after Random Forest imputation"

tbl_d2b_completion_summary |>
  gt() |>
  tab_header(
    title = "Residual Missing Value Completion",
    subtitle = "Values completed by variable-specific rule after RF imputation"
  ) |>
  cols_label(
    variable = "Variable",
    n_completed = "N Completed",
    rule = "Completion Rule"
  ) |>
  fmt_number(
    columns = n_completed,
    decimals = 0,
    use_seps = TRUE
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = rule == "Integrity check: passed")
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(rows = rule == "Integrity check: review required")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


::: {.callout-tip}
## Imputation Quality
The out-of-bag error estimates indicate acceptable imputation quality.
Out-of-bag (OOB) error estimates are calculated from observations that were
randomly excluded during tree construction within the Random Forest algorithm,
providing an internal quality metric for imputed values without requiring a
separate test set. Random Forest imputation preserves variable relationships
better than simpler methods like mean/mode imputation, which is essential for
accurate downstream modeling.
:::

```{r}
#| label: tbl-d4-phase-d-summary
#| tbl-cap: "Phase D imputation summary"

tbl_d4_phase_d_summary |>
  gt() |>
  tab_header(
    title = "Phase D: Imputation Summary",
    subtitle = "Random Forest-based missing value imputation"
  ) |>
  cols_align(align = "left", columns = Stage) |>
  cols_align(align = "right", columns = Value) |>
  tab_source_note(
    md("Imputation method: `impute_rf_parallel()` (missRanger + ranger engine)")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(14)
  )
```


## Block PCA: Dimensionality Reduction

Many SIVESNU variables capture related constructs — vaccination records, wealth
indicators, health monitoring visits, and morbidity symptoms. These variable
groups exhibit high internal correlation, which can destabilize regression
models and obscure individual variable importance.

Block principal components analysis (PCA) transforms each conceptual group into
orthogonal principal components, reducing dimensionality while preserving the
majority of within-block variance. This approach yields interpretable composite
scores (e.g., "vaccination coverage", "household wealth") rather than dozens of
intercorrelated individual variables.

```{r}
#| label: tbl-e1-pca-blocks
#| tbl-cap: "PCA block definitions and variable counts"

tbl_e1_pca_blocks |>
  gt() |>
  tab_header(
    title = "PCA Block Definitions",
    subtitle = "Variable groups for dimensionality reduction"
  ) |>
  cols_align(align = "left", columns = c(Block, Description, `Sample Variables`)) |>
  cols_align(align = "center", columns = `N Variables`) |>
  tab_style(
    style = cell_fill(color = "#ffffcc"),
    locations = cells_body(rows = `N Variables` < 3)
  ) |>
  tab_source_note(
    md("**Note:** Blocks with <3 variables (highlighted) will not undergo PCA")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-e2-pca-results
#| tbl-cap: "Block PCA dimensionality reduction results"

tbl_e2_pca_results |>
  gt() |>
  tab_header(
    title = "Block PCA Dimensionality Reduction",
    subtitle = "Variance preserved while reducing variable count"
  ) |>
  cols_align(align = "left", columns = c(Block, Description)) |>
  cols_align(align = "center",
             columns = c(`Original Variables`, `Components Retained`,
                         `Variance Retained (%)`, `Reduction (%)`)) |>
  fmt_number(columns = `Variance Retained (%)`, decimals = 1) |>
  fmt_number(columns = `Reduction (%)`, decimals = 0) |>
  sub_missing(missing_text = "—") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Block == "TOTAL")
  ) |>
  tab_style(
    style = cell_fill(color = "#e8f5e9"),
    locations = cells_body(
      columns = `Variance Retained (%)`,
      rows = `Variance Retained (%)` >= 70
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#fff3e0"),
    locations = cells_body(
      columns = `Variance Retained (%)`,
      rows = `Variance Retained (%)` < 70 & `Variance Retained (%)` >= 60
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#ffebee"),
    locations = cells_body(
      columns = `Variance Retained (%)`,
      rows = `Variance Retained (%)` < 60
    )
  ) |>
  tab_source_note(
    md("**Method:** Principal Component Analysis with Kaiser criterion (eigenvalue > 1)<br>
        **Color coding:** Green ≥70%, Orange 60-70%, Red <60% variance retained")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-e2-scree-plots
#| fig-cap: "Scree plots showing eigenvalues and component retention by block"
#| fig-alt: >
#|   Grid of small scree plots, one panel per variable block, each with
#|   component number on the x-axis and eigenvalue on the y-axis. A line
#|   connects the points, which descend from left to right, and a dashed
#|   horizontal line marks the Kaiser cutoff at an eigenvalue of one. Points
#|   above the cutoff are highlighted as retained components and points below
#|   are greyed out, showing that only the first few components in each block
#|   are kept for dimensionality reduction.
#| fig-width: 12
#| fig-height: 10

scree_plots <- fig_e2_scree_plots |>
  group_split(Block) |>
  map(function(df_block) {
    block_name <- unique(df_block$Block)
    ggplot(df_block, aes(x = Component, y = Eigenvalue)) +
      geom_line(color = "grey50") +
      geom_point(aes(color = Retained), size = 3) +
      geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
      scale_color_manual(values = c("FALSE" = "grey70", "TRUE" = "steelblue")) +
      scale_x_continuous(
        breaks = function(x) seq(1, max(x), by = max(1, floor(max(x) / 10)))
      ) +
      labs(
        title = block_name,
        x = "Component",
        y = "Eigenvalue"
      ) +
      theme_minimal(base_size = 10) +
      theme(
        legend.position = "none",
        plot.title = element_text(face = "bold", size = 11)
      )
  })

wrap_plots(scree_plots, ncol = 3)
```

```{r}
#| label: tbl-e3-pca-update-summary
#| tbl-cap: "Dataset update after Block PCA"

tbl_e3_pca_update_summary |>
  gt() |>
  tab_header(
    title = "Phase E: Candidate Variable Transformation",
    subtitle = "Block PCA dimensionality reduction on candidate variables"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric %in% c("Candidate variables after PCA",
                                                  "Assignment variables preserved"))
  ) |>
  tab_style(
    style = cell_text(indent = px(20)),
    locations = cells_body(rows = str_starts(Metric, "—"))
  ) |>
  tab_source_note(
    md("**Criterion:** Kaiser rule (eigenvalue > 1) with minimum 1 component per block")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Dimensionality Reduction
Block PCA achieved substantial dimensionality reduction while preserving the
majority of variance within each conceptual domain. The vaccination and wealth
blocks show particularly high variance retention, indicating that these
constructs are well-captured by their principal components.
:::


## Feature Selection

With dimensionality reduced, we now identify which variables — both individual
predictors and PCA components — contribute meaningfully to stunting prediction.
This three-step process removes redundancy and identifies the most informative
predictors.

### Correlation Filter

Highly correlated variables (r > 0.85) provide redundant information and can
inflate coefficient standard errors. The correlation filter removes the
variable with higher mean absolute correlation from each highly correlated
pair.

```{r}
#| label: tbl-f1-correlation-filter
#| tbl-cap: "Correlation filter results"

tbl_f1_correlation_filter |>
  gt() |>
  tab_header(
    title = "Correlation Filter Results",
    subtitle = "Removal of redundant variables (high pairwise correlation)"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "Variables retained")
  ) |>
  tab_source_note(
    md("**Method:** caret::findCorrelation removes variable with higher mean absolute correlation from each pair")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Boruta Feature Selection

Boruta identifies variables with confirmed predictive importance by comparing
each variable's Random Forest importance against "shadow" variables (permuted
copies of the original predictors). Variables consistently outperforming their
shadows are classified as Confirmed; those consistently worse are Rejected.

The implementation runs with a multi-threaded `ranger` engine (via
`Boruta::getImpRfZ`) for efficient computation. Feature selection uses HAZ
(continuous z-score) rather than binary stunting status to preserve full
distributional information — since stunting is simply HAZ dichotomized at -2,
selecting on HAZ captures identical predictive relationships without
information loss.

```{r}
#| label: tbl-f2-boruta-results
#| tbl-cap: "Boruta feature selection results"

tbl_f2_boruta_results |>
  gt() |>
  tab_header(
    title = "Boruta Feature Selection Results",
    subtitle = "Variable importance assessment for HAZ prediction"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "Variables selected (Confirmed + Tentative)")
  ) |>
  tab_source_note(
    md("**Method:** Boruta with ranger Random Forest engine (Z-score importance)")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-f2-boruta-importance
#| fig-cap: "Variable importance from Boruta feature selection"
#| fig-alt: >
#|   Horizontal box plots ranking selected predictors, with Random Forest
#|   importance on the x-axis and one variable per row on the y-axis. Each box
#|   summarises the spread of importance scores across Boruta iterations, shaded
#|   by decision as either Confirmed or Tentative. Boxes sit at positive
#|   importance values well above their permuted shadow features, identifying
#|   the variables that meaningfully predict height-for-age z-score, including
#|   child age, maternal characteristics, food security, and PCA components.
#| fig-height: 12
#| fig-width: 10

ggplot(fig_f2_boruta_importance,
       aes(x = importance, y = variable, fill = decision)) +
  geom_boxplot(alpha = 0.4, outlier.size = 0.5) +
  scale_fill_manual(
    values = c(
      "Confirmed" = "#2E7D32",
      "Tentative" = "#F9A825"
    ),
    labels = c("Confirmed", "Tentative")
  ) +
  labs(
    title = "Boruta Variable Importance",
    subtitle = "Selected variables only (Confirmed + Tentative)",
    x = "Importance",
    y = NULL,
    fill = "Decision"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position = "bottom",
    axis.text.y = element_text(size = 8),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )
```

::: {.callout-important}
## Variables Selected
Boruta identified the most relevant predictors for the continuous HAZ outcome,
including child age, maternal characteristics, food security indicators, and
several PCA components capturing vaccination coverage, household wealth, and
child health monitoring patterns.
:::


### VIF Verification

Variance Inflation Factor (VIF) assesses multicollinearity across the selected
predictors. The correlation filter addressed pairwise correlations; VIF detects
multicollinearity arising from linear combinations of several variables. For
categorical variables the generalized VIF is adjusted by degrees of freedom, so
every variable is reported on a comparable scale.

Variables above the high threshold of 10 are pruned iteratively: at each step
the one with the lowest Boruta median importance among them is dropped and the
VIF is recomputed on the remaining set. A construct captured through two
separate survey items passes the pairwise correlation filter while still
inflating the variance of both coefficients, and this step resolves those
cases, keeping whichever member carries more signal on the outcome.

```{r}
#| label: tbl-f3-vif-summary
#| tbl-cap: "VIF Summary"

tbl_f3_vif_summary |>
  gt() |>
  tab_header(
    title = "Phase F3: VIF Verification Summary",
    subtitle = "Multicollinearity check on Boruta-selected variables"
  ) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = Metric == "High (VIF > 10)" & Value != "0")
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "Variables evaluated")
  ) |>
  tab_source_note(
    source_note = "Method: VIF with GVIF adjustment for categorical variables"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-f3-vif-results
#| tbl-cap: "VIF Analysis by Variable"

tbl_f3_vif_results |>
  gt() |>
  tab_header(
    title = "Variance Inflation Factor Analysis",
    subtitle = "Multicollinearity assessment for selected predictors"
  ) |>
  fmt_number(columns = vif, decimals = 2) |>
  cols_label(
    variable = "Variable",
    vif = "VIF",
    status = "Status"
  ) |>
  cols_align(align = "left", columns = variable) |>
  cols_align(align = "right", columns = vif) |>
  cols_align(align = "center", columns = status) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = status == "High")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(rows = status == "Moderate")
  ) |>
  tab_style(
    style = cell_fill(color = "#C8E6C9"),
    locations = cells_body(rows = status == "Acceptable")
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = vif > 10, columns = vif)
  ) |>
  tab_source_note(
    source_note = "Thresholds: VIF > 10 high multicollinearity; VIF > 5 moderate concern"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-f3-vif-dropped
#| tbl-cap: "Variables removed by the VIF pruning step"

tbl_f3_vif_dropped |>
  gt() |>
  tab_header(
    title = "Variables Removed by VIF Pruning",
    subtitle = "Lowest Boruta median importance among those above the threshold"
  ) |>
  fmt_number(columns = c(vif, median_imp), decimals = 2) |>
  cols_label(
    variable = "Variable",
    vif = "VIF at removal",
    median_imp = "Boruta median importance"
  ) |>
  cols_align(align = "left", columns = variable) |>
  cols_align(align = "right", columns = c(vif, median_imp)) |>
  tab_source_note(
    source_note = "Removed in the order shown; the VIF is the value held at the iteration of removal"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

::: {.callout-tip}
## Multicollinearity Assessment
The pruning removed `r n_vif_dropped` variables: `r vars_vif_dropped`. Each was
the less important member of a pair measuring the same construct — the
urban/rural classification, recorded in two separate survey items, and two
altitude measurements.

The `r n_vif_retained` variables that remain all sit at or below the acceptable
threshold of 5, with a maximum of `r max_vif_retained`.
:::


## Final Output

The pipeline produces two key datasets and transformation objects required for
downstream modeling and scenario simulation.

```{r}
#| label: tbl-g4-output-summary
#| tbl-cap: "Output files generated by this script"

tbl_g4_output_summary |>
  gt() |>
  tab_header(
    title = "Phase G: Output Files Summary",
    subtitle = "Files generated for downstream analysis"
  ) |>
  cols_align(align = "left", columns = c(File, Location, Description, Purpose)) |>
  cols_align(align = "center", columns = Contents) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = File)
  ) |>
  tab_source_note(
    source_note = "All files saved relative to project root via here()"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-g5-pipeline-summary
#| tbl-cap: "Complete variable selection pipeline summary"

tbl_g5_pipeline_summary |>
  gt() |>
  tab_header(
    title = "Variable Selection Pipeline Summary",
    subtitle = "Complete processing workflow from raw data to final predictors"
  ) |>
  cols_align(align = "left", columns = c(Phase, Input, Output, Method)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Phase == "G. Final Output")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Phase == "G. Final Output")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Summary

### Pipeline Overview

This module transformed raw SIVESNU 2018 data into a modeling-ready dataset
through a systematic, reproducible pipeline. Starting from multiple survey
tables, we audited each source independently, integrated child, household,
maternal, and occupation data, then applied rigorous variable selection to
identify the most relevant predictors for stunting.

### Key Results

| Stage | Output |
|-------|--------|
| Data integration | `r format(n_observations_final, big.mark = ",")` children with valid outcomes |
| Dictionary audit | `r format(n_recoded_total, big.mark = ",")` variables with out-of-domain recodings; `r format(n_values_recoded_total, big.mark = ",")` values recoded to NA |
| Imputation OOB error | OOBE = `r oob_oobe`; PFC = `r oob_pfc` |
| Boruta selection | `r n_boruta_selected` variables retained for HAZ prediction |
| VIF verification | `r n_vif_acceptable` of `r n_vif_evaluated` variables with acceptable multicollinearity |

### Methodological Notes

1. **Dictionary audit** — Each SIVESNU source table is validated independently
   against an operational data dictionary before the merge. Out-of-domain
   values are recoded to NA so downstream stages operate on data whose semantic
   domain is fully under control.

2. **Missing value threshold (35%)** — Variables exceeding this threshold
   cannot be reliably imputed without introducing substantial uncertainty. The
   threshold follows Van Buuren (2018) and Madley-Dowd et al. (2019) and
   balances information retention with imputation reliability.

3. **Boruta on HAZ** — Feature selection used the continuous outcome rather
   than the binary classification to preserve full distributional information.
   Since stunting is HAZ dichotomized at -2 SD, selecting on HAZ captures
   identical predictive relationships.

4. **Block PCA** — Conceptually related variables (vaccination, wealth, health
   monitoring) were reduced to principal components. This approach preserves
   construct-level interpretability while eliminating within-block
   multicollinearity.

5. **Protected variables** — Assignment variables (department, urban/rural,
   occupation, FIES) and nutritional estimates are always retained regardless
   of selection results, as they are required for biofortification targeting
   and scenario simulation.

### Application

The outputs from this module enable:

- **[Stunting model development](04_06_model_stunting_continuous_zlen.qmd)** — Training the continuous HAZ regression model, with the binary stunting classification derived directly from the predicted HAZ values.
- **[Synthetic population generation](04_07_generate_synthetic_population.qmd)** — Creating representative populations for scenario modeling using the pre-PCA dataset.
- **[Biofortification simulation](../06_scenarios/06_02_precompute_scenarios.qmd)** — Modifying nutritional variables while preserving covariate structure.
