---
title: "Calculating Micronutrient Bioavailability"
subtitle: "Module 4: Covariate Prediction"
description-meta: "Absorbed rather than consumed iron and zinc: Miller equation for zinc by age, and a flat 5% iron absorption justified by plant-dominant diets."
---

## Overview

Not all nutrients consumed are absorbed. Iron and zinc — the micronutrients most
relevant to stunting — have absorption rates that vary dramatically based on
dietary composition and physiological factors. A child consuming 10 mg of iron
from a plant-based diet may absorb only 0.5 mg, while the same amount from meat
could yield 2-3 mg of absorbed iron.

This module converts dietary intake estimates into **bioavailable nutrients** —
the amounts actually absorbed and available for physiological use. This
transformation is essential for comparing intake against requirements and
modeling nutritional adequacy.

We apply two established methodologies:

- **Zinc**: Miller equation — a saturable absorption model that accounts for age
  and total zinc intake [@Miller2015_Zinc].
- **Iron**: Uniform 5% absorption factor — based on the predominantly plant-based
  dietary patterns observed in this population [@Allen2006_Fortification].

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  survey,             # Survey-weighted quantiles for histogram reference lines
  srvyr,              # Tidy wrapper for survey package
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script (exported by framework 04_03).

fig_a3_plant_percentage_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_03_fig_a3_plant_percentage_distribution.rds"),
  trust = TRUE
)

tbl_a3_plant_percentage_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_03_tbl_a3_plant_percentage_summary.rds"),
  trust = TRUE
)

tbl_c1_miller_by_age <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_03_tbl_c1_miller_by_age.rds"),
  trust = TRUE
)

tbl_c2_iron_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_03_tbl_c2_iron_summary.rds"),
  trust = TRUE
)

tbl_c3_bioavailability_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_03_tbl_c3_bioavailability_summary.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Extract summary results for narrative cifras
zn_intake_mean <- tbl_c3_bioavailability_summary |>
  filter(measure == "Total Intake (mg/d)") |>
  pull(Zinc)

zn_absorbed_mean <- tbl_c3_bioavailability_summary |>
  filter(measure == "Absorbed (mg/d)") |>
  pull(Zinc)

zn_absorption_pct <- tbl_c3_bioavailability_summary |>
  filter(measure == "Absorption Rate (%)") |>
  pull(Zinc)

fe_intake_mean <- tbl_c3_bioavailability_summary |>
  filter(measure == "Total Intake (mg/d)") |>
  pull(Iron)

fe_absorbed_mean <- tbl_c3_bioavailability_summary |>
  filter(measure == "Absorbed (mg/d)") |>
  pull(Iron)

fe_absorption_pct <- tbl_c3_bioavailability_summary |>
  filter(measure == "Absorption Rate (%)") |>
  pull(Iron)
```


## Data and Parameters

### Integrated Dataset

We use the SIVESNU dataset with Chispitas supplementation already integrated.
Total zinc and iron intake therefore includes both dietary sources
and micronutrient powder contributions.


### Miller Equation Parameters

The Miller equation estimates Total Absorbed Zinc (TAZ) as a saturable function
of age and total dietary zinc (TDZ). Younger children have lower absorption
capacity, which increases with age [@Miller2015_Zinc].

$$TAZ = 0.5 \times \left( K_T + A \cdot Age^E + TDZ - \sqrt{(K_T + A \cdot Age^E + TDZ)^2 - 4 \cdot (A \cdot Age^E) \cdot TDZ} \right)$$

Where:

- $TAZ$ = Total Absorbed Zinc (mg/day).
- $TDZ$ = Total Daily Zinc from all sources (mg/day).
- $Age$ = Age in months.
- $K_T$ = 2.42 (saturation constant).
- $A$ = 0.633 (age coefficient).
- $E$ = 0.311 (age exponent).

::: {.callout-note}
## Equation Applicability
The Miller equation was derived from studies including micronutrient powder trials,
making it appropriate for estimating absorption from combined diet + Chispitas intake.
:::


## Iron Bioavailability Decision

Iron absorption varies dramatically with dietary composition — from ~5% in
plant-based diets (high phytate inhibition) to ~15-25% in mixed diets with
animal-source foods. Rather than applying individual-level adjustments, we first
characterize the population's dietary patterns.

### Dietary Composition Analysis

```{r}
#| label: fig-a3-plant-percentage-distribution
#| fig-cap: "Distribution of plant-based food percentage in children's diets (SIVESNU 2018)"
#| fig-alt: >
#|   Histogram of the percentage of plant-based food in each child's diet, with
#|   percentage from 0 to 100 on the x-axis and frequency on the y-axis. A solid
#|   vertical line marks the survey-weighted median and two dashed lines mark the
#|   tercile cutoffs. The distribution is concentrated toward high percentages,
#|   with most children clustered at the upper end, so that even the lower tercile
#|   cutoff falls within the very-high-phytate range that justifies a uniform 5%
#|   iron absorption rate.
#| fig-width: 9
#| fig-height: 6

# Define Azahar color palette
color_azul_primary <- "#1f78b4"

# Compute survey-weighted reference lines from the exported dataframe
svy_temp <- fig_a3_plant_percentage_distribution |>
  as_survey_design(ids = conglomerado, weights = pesonino)

weighted_median <- svyquantile(
  ~per_vegetales, design = svy_temp,
  quantiles = 0.50, na.rm = TRUE
) |>
  coef() * 100

weighted_terciles <- svyquantile(
  ~per_vegetales, design = svy_temp,
  quantiles = c(1/3, 2/3), na.rm = TRUE
) |>
  coef() * 100

# Create histogram
ggplot(fig_a3_plant_percentage_distribution, aes(x = per_vegetales * 100)) +
  geom_histogram(
    binwidth = 2.5,
    fill = color_azul_primary,
    color = "white",
    alpha = 0.8
  ) +
  geom_vline(
    xintercept = weighted_terciles[1],
    linetype = "dashed",
    color = "#3AAF85",
    linewidth = 0.8
  ) +
  geom_vline(
    xintercept = weighted_terciles[2],
    linetype = "dashed",
    color = "#3AAF85",
    linewidth = 0.8
  ) +
  geom_vline(
    xintercept = weighted_median,
    linetype = "solid",
    color = "#1a3a5c",
    linewidth = 1
  ) +
  annotate(
    "text",
    x = weighted_terciles[1] - 2,
    y = Inf,
    label = paste0("T1: ", round(weighted_terciles[1], 1), "%"),
    hjust = 1,
    vjust = 1.5,
    size = 3,
    color = "#236C51"
  ) +
  annotate(
    "text",
    x = weighted_terciles[2] + 2,
    y = Inf,
    label = paste0("T2: ", round(weighted_terciles[2], 1), "%"),
    hjust = 0,
    vjust = 1.5,
    size = 3,
    color = "#236C51"
  ) +
  labs(
    title = "Distribution of Plant-Based Food Consumption",
    subtitle = "Children 6-59 months, SIVESNU 2018",
    x = "Percentage of Plant-Based Food Consumption",
    y = "Frequency",
    caption = "Vertical lines: median (solid), tercile cutoffs (dashed). Source: SIVESNU 2018."
  ) +
  scale_x_continuous(
    breaks = seq(0, 100, 10),
    labels = function(x) paste0(x, "%")
  ) +
  coord_cartesian(xlim = c(0, 100)) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 13, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    plot.caption = element_text(size = 9, color = "grey50", hjust = 0),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4")
  )
```

```{r}
#| label: tbl-a3-plant-percentage-summary
#| tbl-cap: "Survey-weighted distribution of plant-based food percentage (SIVESNU 2018)"

tbl_a3_plant_percentage_summary |>
  gt() |>
  tab_header(
    title = "Plant-Based Food Percentage Distribution",
    subtitle = "Children 6-59 months, SIVESNU 2018"
  ) |>
  cols_label(
    statistic = "Statistic",
    value = "Value",
    interpretation = "Interpretation"
  ) |>
  tab_spanner(
    label = "Survey-Weighted Estimates",
    columns = c(value, interpretation)
  ) |>
  tab_footnote(
    footnote = "Survey weights applied (pesonino). Tercile cutoffs based on the plant-based diet classification of Allen et al. (2006).",
    locations = cells_title(groups = "subtitle")
  ) |>
  tab_source_note(
    source_note = md("**Source**: SIVESNU 2018. SD estimated from unweighted data.")
  ) |>
  cols_width(
    statistic ~ px(180),
    value ~ px(180),
    interpretation ~ px(280)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )
```

::: {#imp-iron-absorption-rate .callout-important}
## Methodological Decision: Iron Absorption Rate

**Finding**: A large majority of children have diets with high plant-based food
consumption (see @tbl-a3-plant-percentage-summary). Even the lower tercile cutoff
falls within the "very high phytate" category. The population distribution of the
plant-based diet percentage is also documented in
[Diet Composition](../01_nutritional_intakes/01_03_intake_calculation.qmd#diet-composition).

**Decision**: We apply a uniform 5% iron absorption rate, appropriate for the unrefined, 
high-phytate diets characteristic of this population [@Allen2006_Fortification]. 
Combined with the harmonized average requirement (H-AR) for iron [@Allen2020_HAR], 
this yields an absorbed iron requirement of 0.5 mg/day.

**Absorbed iron requirement**: With 5% absorption and 10 mg/day recommended intake,
the absorbed iron requirement is **0.5 mg/day**.
:::


## Validation

### Zinc Absorption by Age

The Miller equation produces age-dependent absorption rates. Younger children
(6-11 months) show lower fractional absorption than older children
(48-59 months), which is consistent with developing intestinal absorption
capacity.

```{r}
#| label: tbl-c1-miller-by-age
#| tbl-cap: "Zinc Absorption by Age Group (Miller Equation)"

tbl_c1_miller_by_age |>
  gt() |>
  tab_header(
    title = "Zinc Absorption by Age Group",
    subtitle = "Miller equation results for children 6-59 months"
  ) |>
  cols_label(
    age_group = "Age Group",
    n = "N",
    mean_zn_intake = "Zn Intake",
    mean_zn_intake_low = "95% CI",
    mean_zn_intake_upp = "",
    mean_taz = "TAZ",
    mean_taz_low = "95% CI",
    mean_taz_upp = "",
    mean_absorption_pct = "Absorption",
    mean_absorption_pct_low = "95% CI",
    mean_absorption_pct_upp = ""
  ) |>
  cols_merge(columns = c(mean_zn_intake_low, mean_zn_intake_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(mean_taz_low, mean_taz_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(mean_absorption_pct_low, mean_absorption_pct_upp), pattern = "[{1}, {2}]") |>
  fmt_number(columns = c(mean_zn_intake, mean_taz), decimals = 2) |>
  fmt_number(columns = c(mean_zn_intake_low, mean_taz_low), decimals = 2) |>
  fmt_number(columns = c(mean_absorption_pct, mean_absorption_pct_low), decimals = 1) |>
  tab_spanner(label = "Zinc Intake (mg/d)", columns = c(mean_zn_intake, mean_zn_intake_low)) |>
  tab_spanner(label = "Absorbed Zinc (mg/d)", columns = c(mean_taz, mean_taz_low)) |>
  tab_spanner(label = "Fractional (%)", columns = c(mean_absorption_pct, mean_absorption_pct_low)) |>
  tab_source_note("TAZ = Total Absorbed Zinc. Absorption increases with age per Miller equation.") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

### Iron Absorption Summary

```{r}
#| label: tbl-c2-iron-summary
#| tbl-cap: "Iron Intake and Absorption Summary"

tbl_c2_iron_summary |>
  gt() |>
  tab_header(
    title = "Iron Intake and Absorption",
    subtitle = "Children 6-59 months, uniform 5% absorption factor"
  ) |>
  cols_label(
    n = "N",
    mean_fe_intake = "Mean Intake",
    mean_fe_intake_low = "95% CI",
    mean_fe_intake_upp = "",
    mean_fe_absorbed = "Mean Absorbed",
    mean_fe_absorbed_low = "95% CI",
    mean_fe_absorbed_upp = "",
    absorption_factor = "Absorption Factor",
    absorbed_requirement = "Requirement (absorbed)"
  ) |>
  cols_merge(columns = c(mean_fe_intake_low, mean_fe_intake_upp), pattern = "[{1}, {2}]") |>
  cols_merge(columns = c(mean_fe_absorbed_low, mean_fe_absorbed_upp), pattern = "[{1}, {2}]") |>
  fmt_number(columns = c(mean_fe_intake, mean_fe_intake_low, mean_fe_absorbed, mean_fe_absorbed_low), decimals = 2) |>
  tab_spanner(label = "Iron Intake (mg/d)", columns = c(mean_fe_intake, mean_fe_intake_low)) |>
  tab_spanner(label = "Absorbed Iron (mg/d)", columns = c(mean_fe_absorbed, mean_fe_absorbed_low)) |>
  tab_spanner(label = "Methodology", columns = c(absorption_factor, absorbed_requirement)) |>
  tab_source_note("Absorption factor based on Allen et al. (2006) classification of diets by iron bioavailability; 
  absorbed requirement based on Allen et al. (2020) H-AR.") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```

### Combined Bioavailability Summary

```{r}
#| label: tbl-c3-bioavailability-summary
#| tbl-cap: "Micronutrient Bioavailability Summary"

tbl_c3_bioavailability_summary |>
  gt() |>
  tab_header(
    title = "Micronutrient Bioavailability Summary",
    subtitle = "Survey-weighted estimates for children 6-59 months"
  ) |>
  cols_label(
    measure = "Metric",
    Zinc = "Zinc",
    Iron = "Iron"
  ) |>
  fmt_number(columns = c(Zinc, Iron), decimals = 2) |>
  tab_source_note("Zinc: Miller equation (age-dependent). Iron: Assumed uniform 5% absorption based on dietary patterns") |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )
```


## Summary

### Key Results

| Micronutrient | Total Intake | Absorbed | Absorption Rate |
|---------------|--------------|----------|-----------------|
| Zinc | `r sprintf("%.2f mg/d", zn_intake_mean)` | `r sprintf("%.2f mg/d", zn_absorbed_mean)` | `r sprintf("%.1f%%", zn_absorption_pct)` (age-dependent) |
| Iron | `r sprintf("%.2f mg/d", fe_intake_mean)` | `r sprintf("%.2f mg/d", fe_absorbed_mean)` | `r sprintf("%.1f%%", fe_absorption_pct)` (uniform) |

### Methodological Notes

1. **Zinc absorption** — The Miller equation captures the saturable nature of
   zinc absorption: fractional absorption decreases at higher intakes while
   absolute absorption increases. Age-dependency reflects developing intestinal
   absorption capacity.

2. **Iron absorption** — The uniform 5% rate is conservative but appropriate for
   this population's dietary patterns. Individual variation exists, but
   population-level estimates justify this approach for modeling purposes.

3. **Chispitas contribution** — Supplemental micronutrients are included in total
   intake before absorption calculations. The Miller equation was validated with
   MNP studies, supporting this approach.

### Limitations

- **Mineral form differences**: Absorption calculations do not account for
  potential differences in absorption of the mineral forms delivered through
  Chispitas (zinc gluconate, ferrous fumarate) and dietary sources.

- **Population heterogeneity**: Uniform iron absorption may underestimate
  absorption for the small fraction of children with more diverse diets.

### Application

The bioavailable nutrient estimates enable:

- Comparison against absorbed nutrient requirements (not just intake requirements).
- More accurate adequacy assessment for stunting risk modeling.
- Scenario modeling of biofortification impact on absorbed micronutrients.
