---
title: "Continuous Stunting Model"
subtitle: "Module 4: Covariate Prediction"
---

## Overview

The continuous Z-score model predicts the **magnitude** of height-for-age
deviation (HAZ) from the WHO reference. Modeling HAZ directly rather than the
binary stunting classification (stunted = HAZ < -2) preserves the full
distributional information about child growth, since the WHO threshold is a
direct transformation of the same z-score. The continuous outcome serves two
purposes:

1. **Distributional analysis** — Understand how far children fall below growth
   standards, not just whether they cross the −2 SD threshold.
2. **Impact quantification** — Estimate the expected Z-score improvement from
   biofortification, enabling comparison with other nutritional interventions.

This approach is particularly valuable for biofortification impact assessment
because improvements in nutrient intake may shift children's Z-scores without
necessarily crossing the stunting threshold. Capturing these sub-threshold
improvements provides a more complete picture of intervention effects.

The binary stunting classification is derived afterwards by thresholding both
the observed and the predicted z-scores at −2, which lets us evaluate the
continuous model as a binary classifier of stunting status without fitting a
separate logistic regression.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  scales,             # Number formatting
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script (exported by
# framework 04_06).

# --- Phase A: Data preparation ---
tbl_a3_normalization_params <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_a3_normalization_params.rds"),
  trust = TRUE
)

tbl_a5_multicollinearity <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_a5_multicollinearity.rds"),
  trust = TRUE
)

# --- Phase B: Model specification ---
tbl_b2_formula_audit <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_b2_formula_audit.rds"),
  trust = TRUE
)

fig_b5_observed_predicted <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_fig_b5_observed_predicted.rds"),
  trust = TRUE
)

# --- Phase C: Residual diagnostics ---
fig_c1_residuals_histogram <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_fig_c1_residuals_histogram.rds"),
  trust = TRUE
)

fig_c2_residuals_fitted <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_fig_c2_residuals_fitted.rds"),
  trust = TRUE
)

tbl_c3_performance_metrics <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_c3_performance_metrics.rds"),
  trust = TRUE
)

# --- Phase D: Binary stunting classification (derived) ---
tbl_d1_prevalence <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_d1_prevalence.rds"),
  trust = TRUE
)

tbl_d2_confusion_matrix <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_d2_confusion_matrix.rds"),
  trust = TRUE
)

tbl_d3_classification_metrics <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_d3_classification_metrics.rds"),
  trust = TRUE
)

tbl_d4_auc_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_d4_auc_summary.rds"),
  trust = TRUE
)

fig_d4_roc_curve <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_fig_d4_roc_curve_data.rds"),
  trust = TRUE
)

# --- Variable importance ---
fig_d_variable_importance <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_fig_d_variable_importance.rds"),
  trust = TRUE
)

# --- Phase E: Performance summary ---
tbl_e2_performance_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_e2_performance_summary.rds"),
  trust = TRUE
)

# --- Phase G: GAM for impact simulation ---
fig_f3_gam_effect <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_fig_f3_gam_effect.rds"),
  trust = TRUE
)

fig_f4_marginal_effect <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_fig_f4_marginal_effect.rds"),
  trust = TRUE
)

tbl_f5_gam_verification <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_f5_gam_verification.rds"),
  trust = TRUE
)

tbl_f2_gam_fit_coverage <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "04_06_tbl_f2_gam_fit_coverage.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Performance metrics
rsq_value <- tbl_c3_performance_metrics |>
  filter(Metric == "R²") |>
  pull(Value)

n_obs_model <- tbl_c3_performance_metrics |>
  filter(Metric == "Observations") |>
  pull(Value)

# Binary classification headline cifras
prev_observed_weighted <- tbl_d1_prevalence |>
  filter(Metric == "Observed stunting prevalence (survey-weighted)") |>
  pull(Value)

prev_predicted_weighted <- tbl_d1_prevalence |>
  filter(Metric == "Predicted stunting prevalence (survey-weighted)") |>
  pull(Value)

# AUC
auc_value <- tbl_d4_auc_summary |>
  filter(Metric == "AUC (survey-weighted)") |>
  pull(Value)

sensitivity_who <- tbl_d4_auc_summary |>
  filter(Metric == "Sensitivity at WHO threshold") |>
  pull(Value)

specificity_who <- tbl_d4_auc_summary |>
  filter(Metric == "Specificity at WHO threshold") |>
  pull(Value)

# Formula audit counts
n_formula_terms <- nrow(
  tbl_b2_formula_audit |>
    filter(action %in% c("gam_inferred", "forced_linear", "kept_categorical"))
)

n_dropped_correlation <- sum(tbl_b2_formula_audit$action == "dropped_correlation")

# GAM ratio at stunted median
gam_ratio_stunted <- tbl_f5_gam_verification |>
  filter(Category == "Ratio GAM/Linear at Stunted Median") |>
  pull(Value)
```


## Data Preparation

### Training Data and Normalization Parameters

The model uses the prepared SIVESNU dataset from the previous module. The
nutritional index components are normalized using z-score parameters
(mean and SD) calculated on the outlier-treated training data; these
parameters are exported and reused downstream in scenario simulation, so that
biofortified values are projected onto the exact same scale as the training
data.


### Outlier Treatment

Outlier treatment is directional and stratified by household food insecurity.
Children are grouped into terciles of the FIES Rasch score, the measure of food
access collected directly in SIVESNU and documented in
[FIES Food Security Score](04_04_calculate_fies_food_security_score.qmd).
Within the most food-insecure tercile, upper outliers are pulled down to the
stratum median; within the most food-secure tercile, lower outliers are pulled
up to the stratum median. The intermediate tercile is left untouched.

An intake estimate that is extreme relative to a household's observed food
access carries the uncertainty of the transfer model prediction that produced
it. Bounding those values limits their influence on the z-score normalization
parameters, which set the scale of the composite index.

::: {.callout-note}
## Stratifying Variable
The stratification uses food access, which is observed independently of the
anthropometric outcome. The outcome itself takes no part in the treatment, so
the corrected values carry no imprint of the height-for-age z-scores the model
goes on to predict.
:::


### Nutritional Index Construction

The composite nutritional index combines six bioavailable nutrient
measurements into a single predictor using z-score normalization with equal
weights:

| Component | Units | Role |
|-----------|-------|------|
| Iron (absorbed) | mg/day | Micronutrient — biofortification target |
| Zinc (TAZ, Miller equation) | mg/day | Micronutrient — biofortification target |
| Protein (PDCAAS-adjusted) | g/day | Macronutrient quality — biofortification target |
| Lysine | mg/day | Amino acid quality |
| Tryptophan | mg/day | Amino acid quality |
| Energy | kcal/day | Overall caloric adequacy |

::: {.callout-note}
## Why a Composite Index?
Individual nutrients are highly correlated — children with adequate iron
intake typically have adequate zinc, protein, and energy intake. Including
all six nutrients as separate predictors would cause multicollinearity,
attenuating coefficient estimates and reducing sensitivity to biofortification
effects. The composite index captures overall nutrient adequacy while
maintaining a single coefficient that responds predictably to nutrient
improvements.
:::

```{r}
#| label: tbl-a3-normalization-params
#| tbl-cap: "Z-score normalization parameters for nutritional index components"

tbl_a3_normalization_params |>
  gt() |>
  tab_header(
    title = "Nutritional Index: Z-Score Normalization Parameters",
    subtitle = "Calculated from outlier-treated SIVESNU training data"
  ) |>
  fmt_number(columns = c(Mean, SD), decimals = 4) |>
  cols_label(
    Component = "Nutrient Component",
    Mean = "Mean (μ)",
    SD = "Std. Dev. (σ)"
  ) |>
  cols_align(align = "left", columns = Component) |>
  cols_align(align = "right", columns = c(Mean, SD)) |>
  tab_source_note(
    source_note = "Parameters saved to: 04_06_ni_zscore_normalization_params.rds"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### Multicollinearity Assessment

Before model fitting, we assess the correlation between the nutritional index
and every Boruta-selected predictor present in the dataset. Variables highly
correlated with the index (|r| > 0.45) could attenuate its coefficient
through multicollinearity, reducing the model's sensitivity to nutritional
changes in simulation scenarios. The table below shows the ten predictors most
correlated with the nutritional index; those exceeding the threshold are
dropped from the model formula in the next phase. Mandatory predictors,
including `nutritional_index` itself, bypass this filter.

```{r}
#| label: tbl-a5-multicollinearity
#| tbl-cap: "Correlation of Boruta-selected predictors with nutritional_index"

tbl_a5_multicollinearity |>
  head(10) |>
  gt() |>
  tab_header(
    title = "Multicollinearity with Nutritional Index",
    subtitle = "Top 10 correlated predictors (exclusion threshold: |r| > 0.45)"
  ) |>
  cols_label(
    variable = "Variable",
    correlation = "Correlation (r)",
    exclude = "Exclude"
  ) |>
  fmt_number(columns = correlation, decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = exclude, rows = exclude == "Yes")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = exclude, rows = exclude == "No")
  ) |>
  tab_source_note(
    source_note = "Variables flagged 'Yes' are dropped from the model formula in Phase B."
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Model Specification

### Predictor Structure

The model formula is constructed **data-driven** from the current state of the
pipeline. No predictor list is hard-coded in this script; every term is the
result of upstream selection and a univariate GAM evaluation:

- **Starting set**: the Boruta-selected variables exported by the previous module.
- **Correlation filter**: variables flagged as `Exclude = Yes` in the
  multicollinearity table are dropped to preserve the sensitivity of
  `nutritional_index` in simulation scenarios.
- **Mandatory inclusion**: `nutritional_index` is always kept, bypassing the
  correlation filter.
- **Polynomial degree**: for every numeric predictor, a univariate GAM is
  fitted against `zlen` and its effective degrees of freedom (edf) translate
  into the nearest admissible polynomial degree in {1, 2, 3}. Principal
  components (`pc_*`) are treated as linear by design, since polynomials on
  a linear combination of variables are not semantically meaningful.

Univariate GAM fits are parallelized through a local mirai pool for efficient
execution.


### Formula Construction Audit

The table below traces, variable by variable, the decision made by the formula
builder. Two columns deserve special attention:

**Action**: the bucket the variable falls into.

- `dropped_correlation`: Boruta-selected but dropped because its correlation with `nutritional_index` exceeds the exclusion threshold (`|r| > 0.45`). Flagged in pink.
- `forced_linear`: variables forced to linear (polynomial degree 1) by the `linear_only` argument. Principal components are handled this way. Flagged in blue.
- `gam_inferred`: numeric variables whose polynomial degree (1, 2 or 3) was determined from the effective degrees of freedom of a univariate GAM fitted against the outcome.
- `kept_categorical`: factor variables, entered as categorical terms in the formula (no polynomial applicable).

**GAM Status**: the outcome of the univariate GAM evaluation.

- `ok`: the GAM fitted cleanly and the polynomial degree was assigned from its edf.
- `forced`: the variable bypassed the GAM (listed in `linear_only`); degree fixed at 1.
- `categorical`: the variable is a factor; the GAM is not applicable.
- `fit_failed`: the GAM call aborted (typically because a numeric predictor has too few distinct values for the spline basis). Degree falls back to 1. Variables that are conceptually categorical but stored as numeric codes (e.g. binary 1/2 codings) end up here; their treatment as linear is functionally equivalent to their categorical counterpart and does not affect the model.
- `edf_unavailable`: the GAM fitted but returned a non-finite edf. Degree falls back to 1. Rare.

```{r}
#| label: tbl-b2-formula-audit
#| tbl-cap: "Audit trail of formula construction decisions"

tbl_b2_formula_audit |>
  arrange(action, desc(!is.na(edf)), variable) |>
  gt() |>
  tab_header(
    title = "Formula Construction Audit",
    subtitle = "Data-driven from Boruta + correlation filter + GAM edf"
  ) |>
  cols_label(
    variable = "Variable",
    source   = "Source",
    action   = "Action",
    degree   = "Polynomial Degree",
    edf      = "GAM edf",
    status   = "GAM Status"
  ) |>
  fmt_number(columns = edf, decimals = 3) |>
  sub_missing(columns = c(degree, edf), missing_text = "—") |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(rows = action == "dropped_correlation")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3E0"),
    locations = cells_body(rows = action == "dropped_not_in_data")
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = action == "forced_linear")
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### Model Fitting

We use survey-weighted linear regression (Gaussian family) to predict
continuous z-scores. The survey design accounts for SIVESNU's cluster sampling
with child-level weights (`pesonino`), so the coefficient estimates refer to the
national population of children aged 6–59 months.


### Model Discrimination

Good model performance appears as points clustering around the diagonal,
indicating predicted values closely match observed Z-scores. Systematic
deviations from the diagonal would suggest model misspecification or missing
predictors.

```{r}
#| label: fig-b5-observed-predicted
#| fig-cap: "Observed vs predicted Z-scores"
#| fig-alt: >
#|   Scatter plot with predicted height-for-age z-score on the x-axis and
#|   observed z-score on the y-axis, overlaid with a dashed diagonal reference
#|   line of perfect agreement. Points cluster in a broad cloud around the
#|   diagonal without systematic curvature or drift, indicating the survey-
#|   weighted linear model captures the overall relationship with no obvious
#|   misspecification, though scatter around the line reflects residual
#|   variability.
#| fig-width: 7
#| fig-height: 5

ggplot(fig_b5_observed_predicted, aes(x = .fitted, y = zlen)) +
  geom_point(alpha = 0.4, color = "gray50") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "#3AAF85") +
  labs(
    title = "Observed vs Predicted HAZ",
    subtitle = "Survey-weighted linear regression model",
    x = "Predicted Z-score",
    y = "Observed Z-score"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 12, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4")
  )
```


## Model Evaluation

### Residual Diagnostics

Linear regression assumes residuals are approximately normal and
homoscedastic. The histogram should show a symmetric distribution centered at
zero, while the residuals vs. fitted plot should display random scatter
without systematic patterns. A flat loess smooth is what the linear
specification predicts for the residual structure.

```{r}
#| label: fig-c1-residuals-histogram
#| fig-cap: "Distribution of model residuals"
#| fig-alt: >
#|   Histogram of model residuals (observed minus predicted z-score) on the
#|   x-axis and count on the y-axis, with a dashed vertical line at zero. The
#|   bars form a single symmetric bell-shaped mound centred on zero, consistent
#|   with the approximately normal, unbiased residuals that linear regression
#|   assumes.
#| fig-width: 7
#| fig-height: 5

ggplot(fig_c1_residuals_histogram, aes(x = .resid)) +
  geom_histogram(bins = 40, fill = "#3AAF85", color = "white", alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  labs(
    title = "Distribution of Residuals",
    subtitle = "Should be approximately normal and centered at zero",
    x = "Residual (Observed - Predicted)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 12, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4")
  )
```

```{r}
#| label: fig-c2-residuals-fitted
#| fig-cap: "Residuals vs fitted values"
#| fig-alt: >
#|   Scatter plot with predicted z-score on the x-axis and residual on the
#|   y-axis, a dashed horizontal reference line at zero, and a loess smooth
#|   curve through the points. The points spread in a roughly even band around
#|   zero and the smooth stays close to flat, showing random scatter without
#|   systematic pattern and supporting the homoscedasticity and linearity
#|   assumptions.
#| fig-width: 7
#| fig-height: 5

ggplot(fig_c2_residuals_fitted, aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.4, color = "gray50") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#E53935") +
  geom_smooth(method = "loess", se = FALSE, color = "#3AAF85", linewidth = 0.8) +
  labs(
    title = "Residuals vs Fitted Values",
    subtitle = "Should show random scatter without systematic patterns",
    x = "Predicted Z-score",
    y = "Residual"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 12, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4")
  )
```


### Performance Metrics

```{r}
#| label: tbl-c3-performance-metrics
#| tbl-cap: "Model performance metrics"

tbl_c3_performance_metrics |>
  gt() |>
  tab_header(
    title = "HAZ Linear Regression Model Performance",
    subtitle = "Survey-weighted linear regression with polynomial terms"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "R²")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Metric == "R²")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Binary Stunting Classification

The continuous HAZ model produces predictions on the height-for-age z-score
scale. Stunting is defined by the WHO threshold HAZ < −2. By dichotomising
both the observed and the predicted z-scores at this threshold, the continuous
model is evaluated as a binary classifier of stunting status.

All metrics in this section are survey-weighted (`pesonino`) so that they
reflect the population of children aged 6–59 months in Guatemala rather than
the unweighted sample.

### Derived Stunting Prevalence

```{r}
#| label: tbl-d1-prevalence
#| tbl-cap: "Observed vs predicted stunting prevalence"

tbl_d1_prevalence |>
  gt() |>
  tab_header(
    title = "Stunting Prevalence: Observed vs Predicted",
    subtitle = "WHO threshold HAZ < −2"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### Survey-Weighted Confusion Matrix

```{r}
#| label: tbl-d2-confusion-matrix
#| tbl-cap: "Survey-weighted confusion matrix at WHO threshold HAZ < −2"

tbl_d2_confusion_matrix |>
  select(stunting_observed, stunting_predicted, weighted_count) |>
  pivot_wider(
    names_from = stunting_predicted,
    values_from = weighted_count,
    values_fill = 0
  ) |>
  rename(`Observed \\ Predicted` = stunting_observed) |>
  gt() |>
  tab_header(
    title = "Survey-Weighted Confusion Matrix",
    subtitle = "Cell values are weighted counts (population estimates)"
  ) |>
  fmt_number(
    columns = c(Stunted, `Not stunted`),
    decimals = 0,
    use_seps = TRUE
  ) |>
  cols_align(align = "left", columns = `Observed \\ Predicted`) |>
  cols_align(align = "right", columns = c(Stunted, `Not stunted`)) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(
      columns = Stunted,
      rows = `Observed \\ Predicted` == "Stunted"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(
      columns = `Not stunted`,
      rows = `Observed \\ Predicted` == "Not stunted"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(
      columns = `Not stunted`,
      rows = `Observed \\ Predicted` == "Stunted"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(
      columns = Stunted,
      rows = `Observed \\ Predicted` == "Not stunted"
    )
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### Classification Metrics

```{r}
#| label: tbl-d3-classification-metrics
#| tbl-cap: "Survey-weighted classification metrics at WHO threshold"

tbl_d3_classification_metrics |>
  gt() |>
  tab_header(
    title = "Survey-Weighted Classification Metrics",
    subtitle = "Binary stunting classification at threshold HAZ < −2"
  ) |>
  cols_align(align = "left", columns = c(Metric, Interpretation)) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric %in% c("Sensitivity (Recall)", "Specificity"))
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Metric %in% c("Sensitivity (Recall)", "Specificity"))
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


### ROC Curve and AUC

```{r}
#| label: tbl-d4-auc-summary
#| tbl-cap: "Survey-weighted AUC and operating point comparison"

tbl_d4_auc_summary |>
  gt() |>
  tab_header(
    title = "ROC Analysis Summary",
    subtitle = "Survey-weighted AUC and operating point comparison"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "AUC (survey-weighted)")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Metric == "AUC (survey-weighted)")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: fig-d4-roc-curve
#| fig-cap: "Survey-weighted ROC curve for stunting classification"
#| fig-alt: >
#|   ROC curve with false positive rate (1 minus specificity) on the x-axis and
#|   true positive rate (sensitivity) on the y-axis, plotted against a dashed
#|   diagonal chance line. The curve bows above the diagonal toward the top-left
#|   corner, giving an area under the curve reported in the annotation, which
#|   indicates good discrimination between stunted and non-stunted children. Two
#|   marked operating points show the WHO threshold (HAZ below minus 2) and the
#|   Youden-optimal cutoff.
#| fig-width: 8
#| fig-height: 6

# Identify operating points from the ROC curve
roc_curve_annotated <- fig_d4_roc_curve |>
  as_tibble() |>
  mutate(
    distance_to_threshold = abs(threshold - 2),
    youden_j = TPR - FPR
  )

operating_point_who <- roc_curve_annotated |>
  slice_min(distance_to_threshold, n = 1, with_ties = FALSE)

operating_point_youden <- roc_curve_annotated |>
  slice_max(youden_j, n = 1, with_ties = FALSE)

# Extract AUC numeric from the summary table for annotation
auc_numeric <- tbl_d4_auc_summary |>
  filter(Metric == "AUC (survey-weighted)") |>
  pull(Value) |>
  as.numeric()

ggplot(fig_d4_roc_curve, aes(x = FPR, y = TPR)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed",
              color = "azure4", linewidth = 0.4) +
  geom_path(color = "#1B6CA8", linewidth = 1) +
  geom_point(
    data = operating_point_who,
    aes(x = FPR, y = TPR),
    color = "#3AAF85", size = 3
  ) +
  geom_point(
    data = operating_point_youden,
    aes(x = FPR, y = TPR),
    color = "#E53935", size = 3
  ) +
  annotate(
    "text",
    x = operating_point_who$FPR + 0.04,
    y = operating_point_who$TPR - 0.04,
    label = "WHO threshold\n(HAZ < −2)",
    size = 3.2, color = "azure4", hjust = 0
  ) +
  annotate(
    "text",
    x = operating_point_youden$FPR + 0.04,
    y = operating_point_youden$TPR + 0.04,
    label = "Youden optimum",
    size = 3.2, color = "azure4", hjust = 0
  ) +
  annotate(
    "text",
    x = 0.65, y = 0.15,
    label = paste0("AUC = ", round(auc_numeric, 3)),
    size = 4.5, color = "#1B6CA8", fontface = "bold"
  ) +
  scale_x_continuous(limits = c(0, 1), expand = c(0.005, 0.005)) +
  scale_y_continuous(limits = c(0, 1), expand = c(0.005, 0.005)) +
  coord_equal() +
  labs(
    title = "Survey-Weighted ROC Curve",
    subtitle = "Binary stunting classification (HAZ < −2)",
    x = "False Positive Rate (1 − Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 12, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4")
  )
```

::: {.callout-note}
## Scope of Evaluation
All metrics in this section are computed in-sample on
the training data and represent an upper bound on the model's true
classification performance. The metrics are survey-weighted using `pesonino`
so that they reflect the population of children aged 6–59 months in Guatemala.
The WHO threshold (HAZ < −2) is the operationally relevant cutoff for
stunting classification; the Youden-optimal threshold is reported as a
reference for the maximum classification capacity available from the
predicted scores.
:::

## Model Performance Summary

```{r}
#| label: tbl-e2-performance-summary
#| tbl-cap: "Final model performance summary"

tbl_e2_performance_summary |>
  gt() |>
  tab_header(
    title = "HAZ Continuous Model Performance",
    subtitle = "Survey-weighted linear regression with polynomial terms"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "right", columns = Value) |>
  tab_row_group(
    label = "Model Fit",
    rows = Metric %in% c("R²", "RMSE", "MAE")
  ) |>
  tab_row_group(
    label = "Model Configuration",
    rows = Metric == "Observations Used"
  ) |>
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_row_groups()
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "R²")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Metric == "R²")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## GAM Effect Model for Impact Simulation

The survey-weighted linear regression provides population-level predictions,
but biofortification impact may vary across the nutritional distribution.
Children with lower baseline nutrition might benefit more from nutrient
improvements than those already well-nourished.

To capture this heterogeneity, we fit a **Generalized Additive Model (GAM)**
that estimates local marginal effects — how much the z-score changes per unit
increase in the nutritional index at different points in the distribution —
conditional on child age and sex.

::: {.callout-note}
## Biofortification-Specific Index
Impact simulation uses a nutritional index built from the four nutrients whose
intake biofortified maize modifies:

- **Iron** (absorbed)
- **Zinc** (TAZ from Miller equation)
- **Lysine**
- **Tryptophan**

Biofortified varieties raise the lysine and tryptophan content of the grain,
and that amino acid profile is what sets the protein quality of a maize-based
diet. Representing the mechanism through the two amino acids assigns the change
to the quantities that carry it and gives each nutrient a single point of entry
into the index.

$$NI_{bio}^{nomaiz} = \frac{1}{4} \times (Fe_{absorbed,z} + Zn_{absorbed,z} + Lys_{z} + Trp_{z})$$
:::

::: {.callout-important}
## Non-Maize Basis
The index is built exclusively from the **non-maize** sources of the calibrated
block. The full model above uses the six-component `nutritional_index`; the two
are independent objects.

The basis is a question of identification. In these data the observable
variation in maize-source nutrients is variation in *maize quantity*, and maize
dependency is a marker of household resources: within the same socioeconomic
stratum, a diet more skewed to maize accompanies a lower height-for-age
z-score. The non-maize channel carries nutrients that arrive without that
marker, and its association with the outcome is positive and of a magnitude
consistent with the external evidence.

Quality Protein Maize does not change how much maize a child eats; it changes
the nutrient content of the maize already eaten. The increment is therefore
treated downstream as additional nutrient arriving in the non-maize pool,
which rests on the assumption that a milligram of iron acts the same whatever
food carried it. Bioavailability is handled separately through the iron
fraction and the Miller model, so that assumption applies to the quantity and
not to its absorption.

Two consequences follow. The axis has **its own normalization parameters**,
computed here and exported separately, so that the scripts rebuilding the index
from a child's own non-maize intake work in the units the smooth was estimated
in. And the **Chispitas contribution sits inside the axis**: it was added to
non-maize iron and zinc in
[Chispitas Supplementation](04_02_integrate_chispitas_supplementation.qmd),
and its departmental equivalent is added to the synthetic population, so the
channel is present on both sides by construction.
:::

::: {.callout-warning}
## Scope of the Estimate
The smooth loses statistical significance once socioeconomic covariates enter
the specification. The fitted slope is a descriptive dose-response, not an
identified causal effect.
:::


### GAM Effect Curve

The smooth curve shows how predicted Z-score changes across the nutritional
index distribution. The stunting threshold (HAZ = −2) is marked for reference.

```{r}
#| label: fig-f3-gam-effect
#| fig-cap: "Effect of biofortification-specific nutritional index on HAZ"
#| fig-alt: >
#|   Line chart with the biofortification-responsive nutritional index on the
#|   x-axis and predicted height-for-age z-score on the y-axis, showing a GAM
#|   smooth curve inside a shaded 95% confidence band. The curve rises as the
#|   nutritional index increases, so better nutrient adequacy predicts higher
#|   z-scores. A dashed horizontal line marks the stunting threshold at HAZ
#|   equals minus 2 for reference.
#| fig-width: 8
#| fig-height: 5

ggplot(fig_f3_gam_effect, aes(x = nutritional_index_bio_nomaiz, y = zlen_pred)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "grey70") +
  geom_line(color = "#3AAF85", linewidth = 1.2) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "#E53935",
             linewidth = 0.8) +
  annotate(
    "text",
    x = max(fig_f3_gam_effect$nutritional_index_bio_nomaiz) - 1,
    y = -1.85,
    label = "Stunting threshold (HAZ = −2)",
    color = "#E53935", size = 3
  ) +
  labs(
    title = "Effect of Nutritional Index on HAZ",
    subtitle = "GAM smooth with 95% CI — 4-component non-maize index",
    x = "Nutritional Index (Biofortification-Responsive)",
    y = "Predicted HAZ"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 12, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4")
  )
```


### Marginal Effect Analysis

The derivative of the GAM smooth shows how the marginal effect of nutrition on
the z-score varies across the distribution. The dashed reference line shows the
linear-model coefficient for comparison.

```{r}
#| label: fig-f4-marginal-effect
#| fig-cap: "Marginal effect of NI on HAZ across the distribution"
#| fig-alt: >
#|   Line chart with the biofortification-responsive nutritional index on the
#|   x-axis and the marginal effect (derivative of HAZ with respect to the
#|   index) on the y-axis, showing the GAM derivative with a shaded 95%
#|   confidence band and a dashed horizontal line at the constant linear-model
#|   coefficient. The GAM curve is essentially flat across the whole index
#|   range and sits just above the linear coefficient, so the marginal effect
#|   estimated by the smooth is close to constant.
#| fig-width: 8
#| fig-height: 5

# Linear-model reference coefficient is captured in tbl_f5_gam_verification
coef_linear <- tbl_f5_gam_verification |>
  filter(Category == "Linear Model Coefficient (NI)") |>
  pull(Value) |>
  as.numeric()

ggplot(fig_f4_marginal_effect, aes(x = nutritional_index_bio_nomaiz, y = .derivative)) +
  geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci),
              alpha = 0.2, fill = "grey70") +
  geom_line(color = "#3AAF85", linewidth = 1.2) +
  geom_hline(yintercept = coef_linear, linetype = "dashed",
             color = "#E53935", linewidth = 0.8) +
  annotate(
    "text",
    x = min(fig_f4_marginal_effect$nutritional_index_bio_nomaiz) + 0.5,
    y = coef_linear + 0.1,
    label = paste0("Linear coef = ", round(coef_linear, 3)),
    color = "#E53935", size = 3
  ) +
  labs(
    title = "Marginal Effect of NI on HAZ",
    subtitle = "GAM derivative with 95% CI vs linear model coefficient",
    x = "Nutritional Index (Biofortification-Responsive)",
    y = "Marginal Effect (dHAZ/dNI)"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 12, face = "bold", color = "azure4"),
    plot.subtitle = element_text(size = 10, color = "azure4"),
    axis.line = element_line(color = "azure4"),
    axis.title = element_text(size = 10, color = "azure4"),
    axis.text = element_text(size = 9, color = "azure4")
  )
```

::: {.callout-note}
## Shape of the Smooth
On this axis the estimated marginal effect is close to constant: the derivative
stays between 0.284 and 0.285 across an index range spanning −1.7 to 3.9, and
its ratio to the linear coefficient at the stunted median is
`r gam_ratio_stunted`. The specification admits curvature and the fitted smooth
returns little of it, so the GAM and the linear coefficient give nearly the same
marginal effect at every point of the distribution.
:::


### GAM Model Verification

```{r}
#| label: tbl-f5-gam-verification
#| tbl-cap: "GAM model verification and comparison with linear specification"

tbl_f5_gam_verification |>
  gt() |>
  tab_header(
    title = "GAM Model Verification",
    subtitle = "4-component non-maize NI (Fe, Zn, lysine, tryptophan) — GAM vs linear"
  ) |>
  tab_row_group(label = "Model Fit", rows = 1:2) |>
  tab_row_group(label = "Coefficient Comparison", rows = 3:4) |>
  tab_row_group(label = "NI Distribution by Stunting Status", rows = 5:6) |>
  tab_row_group(label = "Impact at Target Population", rows = 7:8) |>
  cols_align(align = "left", columns = Category) |>
  cols_align(align = "right", columns = Value) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Category == "Ratio GAM/Linear at Stunted Median")
  ) |>
  tab_source_note(
    source_note = "The ratio compares the GAM marginal effect at the stunted median against the constant linear coefficient."
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

```{r}
#| label: tbl-f2-gam-fit-coverage
#| tbl-cap: "Complete-case coverage of the GAM variables"

tbl_f2_gam_fit_coverage |>
  gt() |>
  tab_header(
    title = "GAM Fit Coverage",
    subtitle = "Observations available for the outcome, the index and the adjustment terms"
  ) |>
  fmt_number(columns = c(Observed, Total), decimals = 0, use_seps = TRUE) |>
  fmt_number(columns = Percent, decimals = 1) |>
  cols_label(
    Variable = "Variable",
    Observed = "Observed",
    Total = "Total",
    Percent = "% Complete"
  ) |>
  cols_align(align = "left", columns = Variable) |>
  cols_align(align = "right", columns = c(Observed, Total, Percent)) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Summary

### Model Overview

This module developed a survey-weighted continuous HAZ model with a
**data-driven formula** (`r n_formula_terms` final terms; `r n_dropped_correlation`
predictors dropped by the correlation filter), and a complementary GAM for
impact simulation:

| Model | Purpose | Key Feature |
|-------|---------|-------------|
| Survey-weighted linear regression | Population prediction | Data-driven formula with polynomial degrees inferred from univariate GAMs |
| Binary classification (derived) | Stunting status evaluation | HAZ-thresholded at −2; survey-weighted ROC AUC = `r auc_value` |
| GAM with biofortification index | Impact simulation | Non-maize index with age and sex adjustment; GAM/linear ratio `r gam_ratio_stunted` at the stunted median |

### Key Results

- **In-sample R²**: `r rsq_value` over `r n_obs_model` observations.
- **Observed vs predicted stunting prevalence** (survey-weighted): `r prev_observed_weighted` vs `r prev_predicted_weighted`.
- **Classification performance at WHO threshold**: sensitivity `r sensitivity_who`, specificity `r specificity_who`.
- **Discrimination**: survey-weighted ROC AUC = `r auc_value`.

### Application in Impact Simulation

For each child in the [synthetic population](04_07_generate_synthetic_population.qmd),
the GAM-based simulation:

1. Calculates the new non-maize index post-biofortification:
   $NI_{new} = NI_{baseline} + \Delta NI$
2. Evaluates the GAM at both index values, holding age and sex at their
   observed values: $f(NI_{baseline})$ and $f(NI_{new})$
3. Takes the z-score change as the difference of the two fitted values:
   $\Delta HAZ = f(NI_{new}) - f(NI_{baseline})$
4. Classifies new stunting status: stunted if $(HAZ_{baseline} + \Delta HAZ) < -2$

Taking the difference of fitted values, rather than a first-order term, lets the
estimate follow the curvature of the smooth across the whole interval the index
traverses.

::: {.callout-warning}
## Limitations

1. **Impact simulation scope** — The GAM is intended for aggregate impact
   estimation, not individual-level HAZ prediction.

2. **Extrapolation uncertainty** — Confidence intervals widen at distribution
   extremes where data is sparse.

3. **Proportionality assumption** — Impact estimates assume biofortification
   affects index components proportionally to established biofortification
   factors.

4. **In-sample classification metrics** — Sensitivity, specificity, and AUC
   are computed on the training data and represent an upper bound on
   out-of-sample performance.
:::
