---
title: "MAGA Weight Calibration"
subtitle: "Module 2: Economic Impact"
---

## Why Calibrate to MAGA?

Survey data and administrative records often produce different estimates of
agricultural activity. ENCOVI 2023 captures detailed farmer-level information
but uses a general population sampling frame that may under- or over-represent
certain agricultural regions. MAGA (Ministerio de Agricultura, Ganadería y
Alimentación) publishes official departmental statistics based on comprehensive
agricultural monitoring.

By calibrating ENCOVI survey weights to match MAGA totals, we achieve:

1. **Administrative consistency** — Weighted estimates align with official
   government statistics used in policy planning
2. **Improved representativeness** — Correct for differential sampling rates
   across agricultural departments
3. **Credibility for stakeholders** — Results match published figures that
   donors and policymakers reference

This calibration step is particularly important for economic impact projections,
where departmental production volumes directly affect estimated benefits from
biofortified maize adoption.

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  scales,             # Number formatting for visualizations
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script (exported by framework 02_02).

tbl_a2_scaling_factor <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_02_tbl_a2_scaling_factor.rds"),
  trust = TRUE
)

tbl_a5_encovi_vs_maga <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_02_tbl_a5_encovi_vs_maga.rds"),
  trust = TRUE
)

tbl_b2_calibration_validation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_02_tbl_b2_calibration_validation.rds"),
  trust = TRUE
)

tbl_c1_calibration_diagnostics <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_02_tbl_c1_calibration_diagnostics.rds"),
  trust = TRUE
)

fig_c1_weight_distribution <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_02_fig_c1_weight_distribution.rds"),
  trust = TRUE
)

tbl_d1_final_validation <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_02_tbl_d1_final_validation.rds"),
  trust = TRUE
)

tbl_d2_transformation_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_02_tbl_d2_transformation_summary.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Expansion configuration constants (defined in framework script 02_02, Phase C1).
# Reproduced here as constants — these are pipeline configuration parameters,
# not computed outputs.
expansion_threshold      <- 500
target_weight_per_record <- 100

# Derived counts from the weight distribution dataframe (df_farmers_calibrated)
n_to_expand <- fig_c1_weight_distribution |>
  dplyr::summarise(n = sum(needs_expansion, na.rm = TRUE)) |>
  dplyr::pull(n)

n_final_records <- fig_c1_weight_distribution |>
  dplyr::summarise(n = sum(n_copies, na.rm = TRUE)) |>
  dplyr::pull(n)

n_total_calibrated <- nrow(fig_c1_weight_distribution)

n_departments_calibrated <- nrow(tbl_b2_calibration_validation)

# Scaling factor applied to the ENCOVI weights before departmental calibration
scaling_factor_value <- tbl_a2_scaling_factor |>
  dplyr::filter(metric == "Scaling factor") |>
  dplyr::pull(value)

# Size of the pre-calibration gap against MAGA, in ratio terms
ratio_prod_min <- min(tbl_a5_encovi_vs_maga$ratio_production)
ratio_prod_max <- max(tbl_a5_encovi_vs_maga$ratio_production)
n_dept_outside_half_double <- sum(tbl_a5_encovi_vs_maga$ratio_production < 0.5 |
                                  tbl_a5_encovi_vs_maga$ratio_production > 2)

# Weight dispersion before and after the departmental adjustment
diag_value <- function(label) {
  tbl_c1_calibration_diagnostics |>
    dplyr::filter(Metric == label) |>
    dplyr::pull(Value)
}

cv_weights_initial    <- diag_value("Weight CV (initial)")
cv_weights_calibrated <- diag_value("Weight CV (calibrated)")
deff_weights          <- diag_value("Design Effect (DEFF)")
ratio_range_weights   <- diag_value("Adjustment ratio range")
ratios_extreme_low    <- diag_value("Ratios < 0.3 (extreme low)")
ratios_extreme_high   <- diag_value("Ratios > 3.0 (extreme high)")
```


## Calibration Approach

The calibration process adjusts individual survey weights so that weighted
departmental totals for **production** (quintales) and **cultivated land**
(manzanas) match MAGA 2023 official statistics.

### Data Sources

| Source | Purpose |
|--------|---------|
| ENCOVI 2023 | Farmer-level characteristics, original survey weights |
| MAGA 2023 | Departmental targets for production and land area |
| INE 2011 | Total agricultural household reference (1,299,377) |

### Calibration Variables

- **Production** (`total_production`) — Total white maize harvested in quintales
- **Land** (`land_wtcorn`) — Cultivated white maize area in manzanas

Both variables are calibrated simultaneously within each department using linear
calibration, ensuring weighted totals match MAGA targets exactly.


## Initial Weight Scaling

Before departmental calibration, we apply a scaling factor to align ENCOVI's
total agricultural household estimate with official INE census figures.

```{r}
#| label: tbl-a2-scaling-factor
#| tbl-cap: "Initial weight scaling to match INE agricultural census"

tbl_a2_scaling_factor |>
  gt() |>
  tab_header(
    title = "Farmer Population Scaling Factor",
    subtitle = "Alignment of ENCOVI estimates with INE agricultural census"
  ) |>
  cols_label(metric = "Metric", value = "Value") |>
  fmt_number(columns = value, rows = 1:2, use_seps = TRUE, decimals = 0) |>
  fmt_number(columns = value, rows = 3, decimals = 4) |>
  tab_source_note(
    source_note = "Source: INE (2011) — Distribución de hogares agropecuarios según tipología"
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-note}
## Scaling Factor
At `r round(scaling_factor_value, 3)`, the factor places ENCOVI's original
weights at around half the agricultural population the INE census estimates.
ENCOVI draws on a general population sampling frame, not an
agriculture-specific one, and the factor closes that gap before the
departmental calibration runs.
:::


## Pre-Calibration Diagnostic

Before applying calibration, we compare ENCOVI weighted estimates against MAGA
departmental targets to quantify the magnitude of discrepancies.

```{r}
#| label: tbl-a5-encovi-vs-maga
#| tbl-cap: "ENCOVI vs MAGA comparison before calibration"

tbl_a5_encovi_vs_maga |>
  select(
    departamento, n_farmers,
    production_encovi, production_maga_qq, ratio_production,
    land_encovi, land_maga_mz, ratio_land
  ) |>
  gt() |>
  tab_header(
    title = "Pre-Calibration: ENCOVI vs MAGA Comparison",
    subtitle = "Weighted estimates before departmental calibration"
  ) |>
  cols_label(
    departamento = "Department",
    n_farmers = "N (sample)",
    production_encovi = "ENCOVI",
    production_maga_qq = "MAGA",
    ratio_production = "Ratio",
    land_encovi = "ENCOVI",
    land_maga_mz = "MAGA",
    ratio_land = "Ratio"
  ) |>
  tab_spanner(
    label = "Production (qq)",
    columns = c(production_encovi, production_maga_qq, ratio_production)
  ) |>
  tab_spanner(
    label = "Land (mz)",
    columns = c(land_encovi, land_maga_mz, ratio_land)
  ) |>
  fmt_number(
    columns = c(production_encovi, production_maga_qq, land_encovi, land_maga_mz),
    use_seps = TRUE,
    decimals = 0
  ) |>
  fmt_number(
    columns = c(ratio_production, ratio_land),
    decimals = 2
  ) |>
  cols_align(align = "left", columns = departamento) |>
  cols_align(align = "right", columns = -departamento) |>
  tab_style(
    style = cell_fill(color = "#FFCCCC"),
    locations = cells_body(
      columns = ratio_production,
      rows = ratio_production < 0.5 | ratio_production > 2.0
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCCCC"),
    locations = cells_body(
      columns = ratio_land,
      rows = ratio_land < 0.5 | ratio_land > 2.0
    )
  ) |>
  grand_summary_rows(
    columns = c(n_farmers, production_encovi, production_maga_qq,
                land_encovi, land_maga_mz),
    fns = list(Total = ~ sum(.)),
    fmt = ~ fmt_number(., use_seps = TRUE, decimals = 0)
  ) |>
  tab_source_note(
    source_note = md("**Ratio** = ENCOVI / MAGA. Values far from 1.0 indicate large discrepancies requiring calibration.<br>
                      Red highlight: ratio < 0.5 or > 2.0")
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-warning}
## Substantial Discrepancies
Several departments show ratios far from 1.0, indicating that ENCOVI weighted
estimates diverge substantially from MAGA administrative data. Calibration
corrects these discrepancies departmentally.
:::


## Post-Calibration Validation

After applying departmental calibration, we verify that weighted totals match
MAGA targets exactly.

```{r}
#| label: tbl-b2-calibration-validation
#| tbl-cap: "Validation of calibrated weights against MAGA targets"

tbl_b2_calibration_validation |>
  select(
    departamento, n_farmers,
    production_calibrated, production_maga_qq, ratio_production,
    land_calibrated, land_maga_mz, ratio_land
  ) |>
  gt() |>
  tab_header(
    title = "Post-Calibration Validation: ENCOVI vs MAGA",
    subtitle = "Calibrated weighted estimates should match MAGA targets (ratio ≈ 1.00)"
  ) |>
  cols_label(
    departamento = "Department",
    n_farmers = "N (sample)",
    production_calibrated = "Calibrated",
    production_maga_qq = "MAGA",
    ratio_production = "Ratio",
    land_calibrated = "Calibrated",
    land_maga_mz = "MAGA",
    ratio_land = "Ratio"
  ) |>
  tab_spanner(
    label = "Production (qq)",
    columns = c(production_calibrated, production_maga_qq, ratio_production)
  ) |>
  tab_spanner(
    label = "Land (mz)",
    columns = c(land_calibrated, land_maga_mz, ratio_land)
  ) |>
  fmt_number(
    columns = c(production_calibrated, production_maga_qq,
                land_calibrated, land_maga_mz),
    use_seps = TRUE,
    decimals = 0
  ) |>
  fmt_number(
    columns = c(ratio_production, ratio_land),
    decimals = 4
  ) |>
  cols_align(align = "left", columns = departamento) |>
  cols_align(align = "right", columns = -departamento) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = c(ratio_production, ratio_land),
      rows = abs(ratio_production - 1) < 0.001 & abs(ratio_land - 1) < 0.001
    )
  ) |>
  grand_summary_rows(
    columns = c(n_farmers, production_calibrated, production_maga_qq,
                land_calibrated, land_maga_mz),
    fns = list(Total = ~ sum(.)),
    fmt = ~ fmt_number(., use_seps = TRUE, decimals = 0)
  ) |>
  tab_source_note(
    source_note = md("**Ratio** = Calibrated / MAGA. Values of 1.0000 confirm exact calibration.<br>
                      Green highlight: both ratios within ±0.1% of target.")
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

::: {.callout-note}
## Calibration Result
The `r n_departments_calibrated` departments reach a ratio of 1.0000 against
their MAGA targets for both production and land. The calibration is exact by
construction: it solves for the departmental weight adjustment that reproduces
both totals. What the diagnostics below report is the size of the adjustment
this required.
:::


## Record Expansion with Jitter

Calibration can produce very large individual weights, which create artificial
"steps" in cumulative distributions. These steps interfere with threshold-based
farmer segmentation, where cutpoints may fall on weight discontinuities.

We address this by **expanding** high-weight farmers into multiple records with
controlled random variation (jitter). The following variables are jittered in
the expanded records (the original base record of each farmer is left untouched):

- `qty_sold_qq`, `qty_household_qq`, `qty_animal_seed_qq` — Destination quantities
  (sales, household consumption, animal feed and seed reserves)
- `total_production` — Recalculated as the sum of the jittered destination
  quantities, so that allocation coherence is preserved
- `land_wtcorn` — Cultivated land (manzanas)
- `exp_seeds` — Seed expenditure (Q)
- `revenue` — Sales revenue (Q)
- `farmer_age` — Age of the farmer

Jitter is applied with the following parameters:

- **Intensity**: 3% of each variable's standard deviation
- **Bounds**: non-negative values enforced post-jitter
- **Allocation coherence**: `total_production` is derived from the jittered
  destination quantities rather than jittered on its own, which keeps production
  equal to the sum of its uses in every expanded record

```{r}
#| label: tbl-c1-calibration-diagnostics
#| tbl-cap: "Weight diagnostics and expansion parameters"

tbl_c1_calibration_diagnostics |>
  gt() |>
  tab_header(
    title = "Calibration Diagnostics and Expansion Plan",
    subtitle = "Weight validation metrics and jitter expansion parameters"
  ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "center", columns = c(Value, Threshold, Status)) |>
  tab_row_group(label = "Expansion Parameters", rows = 7:10) |>
  tab_row_group(label = "Calibration Metrics", rows = 1:6) |>
  row_group_order(groups = c("Calibration Metrics", "Expansion Parameters")) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(rows = str_detect(Status, "Pass"))
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3E0"),
    locations = cells_body(rows = str_detect(Status, "Warning"))
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = Metric %in% c("Expansion threshold", "Target weight per record",
                                                 "Farmers to expand", "Final record count"))
  ) |>
  tab_source_note(
    source_note = md("**DEFF** = Design Effect (1 + CV²). Values > 2 indicate efficiency loss.<br>
                      **Blue rows**: Expansion parameters to smooth distribution steps.")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    row_group.font.weight = "bold"
  )
```

```{r}
#| label: fig-c1-weight-distribution
#| fig-cap: "Distribution of calibrated weights with expansion threshold"
#| fig-alt: >
#|   Histogram of calibrated farmer weights along the x-axis against the number
#|   of farmers on the y-axis. The distribution is right-skewed, with most
#|   farmers concentrated at low weights and a long tail of high weights. A
#|   dashed vertical line marks the expansion threshold of 500; farmers above
#|   it are split into multiple records so that oversized weights, which create
#|   artificial steps in the distribution, are smoothed out.
#| fig-width: 10
#| fig-height: 6

ggplot(fig_c1_weight_distribution, aes(x = weight_calibrated)) +
  geom_histogram(
    bins = 50,
    fill = "#3E7CB1",
    alpha = 0.3,
    color = "white"
  ) +
  geom_vline(
    xintercept = expansion_threshold,
    linetype = "dashed",
    color = "#3AAF85",
    linewidth = 0.8
  ) +
  annotate(
    "text",
    x = expansion_threshold,
    y = Inf,
    label = paste0("Expansion threshold: ", comma(expansion_threshold)),
    vjust = 2,
    hjust = -0.05,
    color = "#3AAF85",
    size = 3.5
  ) +
  annotate(
    "text",
    x = max(fig_c1_weight_distribution$weight_calibrated) * 0.7,
    y = Inf,
    label = paste0(n_to_expand, " farmers (",
                   round(n_to_expand / n_total_calibrated * 100, 1),
                   "%) will be\nexpanded to ", comma(n_final_records), " records"),
    vjust = 3,
    hjust = 0.5,
    color = "#2C3E50",
    size = 3.5
  ) +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Distribution of Calibrated Weights",
    subtitle = paste0("Farmers with weight > ", comma(expansion_threshold),
                      " will be split into multiple records with target weight ≈ ",
                      comma(target_weight_per_record)),
    x = "Calibrated Weight",
    y = "Number of Farmers"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 12, face = "bold", color = "#2C3E50"),
    plot.subtitle = element_text(size = 10, color = "#566573"),
    axis.title = element_text(size = 10, color = "#2C3E50"),
    axis.text = element_text(size = 9, color = "#2C3E50")
  )
```

::: {.callout-important}
## Expansion Strategy
Farmers with calibrated weight above the expansion threshold are split into
multiple records, each with a target weight of approximately one fifth of the
threshold. This brings the maximum weight down by an order of magnitude and
smooths the large steps out of the distribution. A 3% standard deviation jitter
is applied to base variables to introduce controlled variation among copies.
:::

::: {.callout-important}
## What the Adjustment Costs
The departmental targets are met exactly, and the adjustment needed to meet
them is large. Individual weights are multiplied by factors spanning
`r ratio_range_weights`, against the conventional working range of [0.3, 3.0]:
`r ratios_extreme_low` of records fall below 0.3 and
`r ratios_extreme_high` above 3.0.

The dispersion of the weights rises accordingly, with their coefficient of
variation moving from `r cv_weights_initial` to `r cv_weights_calibrated` and a
design effect of `r deff_weights`. This is the arithmetic consequence of the
pre-calibration gap: the ENCOVI-to-MAGA production ratio runs from
`r round(ratio_prod_min, 2)` to `r round(ratio_prod_max, 2)` across
departments, so the weights carry the whole of that correction.
:::


## Final Validation

After expansion, jitter application, and re-calibration (to correct small
jitter-induced bias), we verify the final dataset still matches MAGA targets.

```{r}
#| label: tbl-d1-final-validation
#| tbl-cap: "Final validation: Expanded dataset weighted totals vs MAGA targets"

tbl_d1_final_validation |>
  select(
    departamento, n_records, n_households, n_farmers_weighted,
    production_final, production_maga_qq, ratio_production,
    land_final, land_maga_mz, ratio_land
  ) |>
  gt() |>
  tab_header(
    title = "Final Validation: Expanded Dataset vs MAGA Targets",
    subtitle = "Weighted totals after expansion and re-calibration"
  ) |>
  cols_label(
    departamento = "Department",
    n_records = "Records",
    n_households = "Households",
    n_farmers_weighted = "Farmers",
    production_final = "Final",
    production_maga_qq = "MAGA",
    ratio_production = "Ratio",
    land_final = "Final",
    land_maga_mz = "MAGA",
    ratio_land = "Ratio"
  ) |>
  tab_spanner(
    label = "Sample",
    columns = c(n_records, n_households, n_farmers_weighted)
  ) |>
  tab_spanner(
    label = "Production (qq)",
    columns = c(production_final, production_maga_qq, ratio_production)
  ) |>
  tab_spanner(
    label = "Land (mz)",
    columns = c(land_final, land_maga_mz, ratio_land)
  ) |>
  fmt_number(
    columns = c(production_final, production_maga_qq, land_final, land_maga_mz,
                n_farmers_weighted),
    use_seps = TRUE,
    decimals = 0
  ) |>
  fmt_number(
    columns = c(ratio_production, ratio_land),
    decimals = 4
  ) |>
  fmt_number(
    columns = c(n_records, n_households),
    use_seps = TRUE,
    decimals = 0
  ) |>
  cols_align(align = "left", columns = departamento) |>
  cols_align(align = "right", columns = -departamento) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = c(ratio_production, ratio_land),
      rows = abs(ratio_production - 1) < 0.0001 & abs(ratio_land - 1) < 0.0001
    )
  ) |>
  grand_summary_rows(
    columns = c(n_records, n_households, n_farmers_weighted,
                production_final, production_maga_qq,
                land_final, land_maga_mz),
    fns = list(Total = ~ sum(.)),
    fmt = ~ fmt_number(., use_seps = TRUE, decimals = 0)
  ) |>
  tab_source_note(
    source_note = md("**Ratio** = Final / MAGA. Values of 1.0000 confirm exact calibration.<br>
                      **Farmers** = Weighted total of white maize farmers represented.<br>
                      Green highlight: both ratios within ±0.01% of target.")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Transformation Pipeline

```{r}
#| label: tbl-d2-transformation-summary
#| tbl-cap: "Complete transformation from ENCOVI sample to calibrated dataset"

tbl_d2_transformation_summary |>
  gt() |>
  tab_header(
    title = "Dataset Transformation Pipeline",
    subtitle = "From ENCOVI sample to MAGA-calibrated expanded dataset"
  ) |>
  cols_align(align = "left", columns = c(Stage, Description)) |>
  cols_align(align = "right", columns = c(Records, `Weight Range`)) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = Stage == "D. Final output")
  ) |>
  tab_source_note(
    source_note = "Pipeline progresses from cleaned ENCOVI input through calibration, expansion with jitter, and re-calibration to the final dataset ready for segmentation modeling"
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```


## Summary

### Key Findings

1. **Initial discrepancies**: Before calibration the ENCOVI-to-MAGA production
   ratio ranges from `r round(ratio_prod_min, 2)` to
   `r round(ratio_prod_max, 2)`, with `r n_dept_outside_half_double` of the
   `r n_departments_calibrated` departments outside the [0.5, 2.0] range.

2. **Calibration**: Linear calibration brings both production and land area
   onto the MAGA departmental targets, with ratios at 1.0000.

3. **Cost in weight dispersion**: Meeting those targets moves the weight
   coefficient of variation from `r cv_weights_initial` to
   `r cv_weights_calibrated`, with a design effect of `r deff_weights` and
   adjustment ratios spanning `r ratio_range_weights`.

4. **Weight smoothing**: Record expansion with controlled jitter removes the
   discontinuities from the distribution while holding the weighted totals,
   and a re-calibration pass corrects the small bias the non-negativity
   constraint introduces.

### Application

The calibrated and expanded dataset provides:

- **Representative farmer population** — Weighted totals match official MAGA
  statistics for economic impact projections
- **Smooth distributions** — No artificial steps for threshold-based
  segmentation in farmer classification
- **Internal consistency** — Allocation coherence between total production and
  destination quantities preserved through the jitter procedure
