---
title: "Agricultural Data Preparation"
subtitle: "Module 2: Economic Impact"
---

## Overview

Economic impact assessment requires detailed information about maize farmers: their
land holdings, production volumes, input costs, and market participation. This
information enables calculation of potential economic gains from biofortified
maize adoption and identification of farmers most likely to benefit.

ENCOVI 2023 provides a comprehensive agricultural module, but the data require
substantial preparation before analysis. Raw survey responses contain heterogeneous
measurement units, missing values, and logical inconsistencies that must be
systematically addressed.

This document describes the preparation of agricultural data through three stages:

1. **Unit corrections** — Standardize land area to manzanas and production to
   quintales, resolving mixed units in survey responses
2. **Data integration** — Combine information from multiple survey modules into
   a unified household-level dataset
3. **Quality treatment** — Apply systematic cleaning including allocation
   normalization, structural NA handling, outlier detection, and imputation

```{r}
#| label: setup
#| include: false

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  scales,             # Number formatting for visualizations
  tidyverse           # General purpose data manipulation
)

# Set display options
options(scipen = 999, digits = 4)
```

```{r}
#| label: load-rds
#| include: false

# Centralized load of web_resources for this script (exported by framework 02_01).

tbl_a4_land_corrections <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_01_tbl_a4_land_corrections.rds"),
  trust = TRUE
)

tbl_a4_yield_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_01_tbl_a4_yield_summary.rds"),
  trust = TRUE
)

tbl_b6_integration_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_01_tbl_b6_integration_summary.rds"),
  trust = TRUE
)

tbl_c1_allocation_diagnostic <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_01_tbl_c1_allocation_diagnostic.rds"),
  trust = TRUE
)

tbl_c2_structural_nas <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_01_tbl_c2_structural_nas.rds"),
  trust = TRUE
)

tbl_c3_impossible_values <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_01_tbl_c3_impossible_values.rds"),
  trust = TRUE
)

tbl_c4_imputation_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_01_tbl_c4_imputation_summary.rds"),
  trust = TRUE
)

tbl_c6_quality_summary <- import(
  here("..", "01_data", "02_processed", "web_resources",
       "02_01_tbl_c6_quality_summary.rds"),
  trust = TRUE
)
```

```{r}
#| label: inline-values
#| include: false

# Counts for the narrative summary

# Number of base variables imputed
n_vars_imputed <- nrow(tbl_c4_imputation_summary)

# Total imputed values (sum across imputed base variables)
n_values_imputed <- tbl_c4_imputation_summary |>
  dplyr::summarise(total = sum(imputed, na.rm = TRUE)) |>
  dplyr::pull(total)

# Flagged impossible values (sum across criteria)
n_values_flagged <- tbl_c3_impossible_values |>
  dplyr::summarise(total = sum(n_cases, na.rm = TRUE)) |>
  dplyr::pull(total)
```


## Data Sources

ENCOVI 2023 includes a comprehensive agricultural module that captures farm-level
information across multiple survey sections. We integrate data from:

- **Agricultural unit registry** — Farm identification and poverty status
- **Crop production module** — Cultivation practices, harvest quantities, input costs, and sales
- **Household demographics** — Composition, location, and socioeconomic characteristics

The production data is filtered to white maize only (cod_cultivo = 142), the staple
crop targeted by biofortification efforts in Guatemala.


## Unit Corrections

ENCOVI respondents report land area and production quantities in heterogeneous
units, and the recorded unit field is not always consistent with the magnitude
of the reported value. Corrections are therefore applied on a plausibility
basis rather than uniformly: each value is first assessed as if it were already
expressed in the standard unit (manzanas for land, quintales for production),
and a conversion is applied only when the reported magnitude is implausible
under that assumption. Values that are already plausible are kept unchanged.

This conditional logic is reflected in the correction outcomes below: the large
majority of land records (around 90%) require no correction, while the remaining
records are reassigned to their most likely original unit based on the magnitude
of the reported value. The "likely unit" terminology denotes this inference:
the most plausible unit given the value, not a unit recorded as such in the
source data.

### Land Area Standardization

Values are standardized to **manzanas** (approximately 0.7 hectares)
using range-based detection:

- **< 0.05** — Likely caballerías (×45)
- **40–60** — Likely acres (×0.7)
- **60–100** — Likely cuerdas (×0.25)
- **100–350** — Likely cuadras (×0.025)
- **≥ 350** — Likely square meters (÷7,000)
- **Zero/NA** — Assigned minimum viable plot (0.05 mz)

```{r}
#| label: tbl-a4-land-corrections
#| tbl-cap: "Land area unit corrections applied to white maize records"

tbl_a4_land_corrections |>
  gt() |>
  tab_header(
    title = "Land Area Unit Corrections",
    subtitle = "White maize records (cod_cultivo = 142)"
  ) |>
  cols_label(
    correction_type = "Correction Applied",
    n_records = "Records",
    pct = "Percentage"
  ) |>
  cols_align(align = "left", columns = correction_type) |>
  fmt_number(columns = n_records, use_seps = TRUE, decimals = 0) |>
  fmt_percent(columns = pct, decimals = 1) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = correction_type == "No correction (manzanas)")
  ) |>
  tab_source_note(
    source_note = "Source: ENCOVI 2023, agricultural production module"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

### Yield Validation

After unit corrections, yield distribution is validated against expected ranges
for Guatemalan white maize production.

```{r}
#| label: tbl-a4-yield-summary
#| tbl-cap: "White maize yield distribution after unit corrections"

tbl_a4_yield_summary |>
  gt() |>
  tab_header(
    title = "White Maize Yield Distribution",
    subtitle = "After unit corrections (qq/mz)"
  ) |>
  cols_label(
    metric = "Metric",
    value = "Value"
  ) |>
  fmt_number(
    columns = value,
    rows = metric %in% c("Total records", "Records with valid yield",
                         "Records with yield < 3 qq/mz",
                         "Records with yield > 150 qq/mz"),
    decimals = 0
  ) |>
  fmt_number(
    columns = value,
    rows = !metric %in% c("Total records", "Records with valid yield",
                          "Records with yield < 3 qq/mz",
                          "Records with yield > 150 qq/mz"),
    decimals = 1
  ) |>
  tab_row_group(label = md("*Distribution*"), rows = 3:9) |>
  tab_row_group(label = md("*Sample Size*"), rows = 1:2) |>
  tab_row_group(label = md("*Extreme Values*"), rows = 10:11) |>
  row_group_order(groups = c("*Sample Size*", "*Distribution*", "*Extreme Values*")) |>
  tab_source_note(
    source_note = "Expected range for Guatemala: 20-40 qq/mz for non-hybrid varieties (locally known as criollo or OPV), up to 80-100 qq/mz for commercial hybrids"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    row_group.font.size = px(13)
  )
```

::: {.callout-note}
## Yield Validation
The median yield of approximately 25 qq/mz sits in the range reported for
traditional Guatemalan non-hybrid maize varieties, which is the reference point
against which the unit conversion is read.
:::


## Data Integration

Variables are derived from multiple ENCOVI modules and integrated at the
household level. The integration includes land metrics, input expenditures,
revenue and self-consumption, land characteristics, and farmer demographics.

```{r}
#| label: tbl-b6-integration-summary
#| tbl-cap: "Data integration summary"

tbl_b6_integration_summary |>
  gt() |>
  fmt_markdown(columns = Source) |>
  cols_align(align = "right", columns = Variables) |>
  cols_align(align = "left", columns = c(Source, Description)) |>
  tab_header(
    title = "Data Integration Summary",
    subtitle = md("*White maize farmers from ENCOVI 2023*")
  ) |>
  tab_source_note(
    source_note = "All dataframes joined by no_hogar (household ID)"
  ) |>
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(rows = 7)
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```


## Data Quality Treatment

This stage addresses data quality issues through systematic processes: allocation
normalization, structural NA handling, impossible value detection, and imputation.

### Production Allocation Consistency

Survey respondents often report destinations (sales, household consumption,
animal feed/seed) that do not sum to total production. All destination
quantities are scaled proportionally to match `total_production`.

The scaling preserves the relative distribution reported by each farmer. Every
destination quantity (sales, household consumption, animal feed and seed) is
multiplied by the ratio between total reported production and the sum of the
reported destinations, so the adjusted destinations sum to total production
while the share of each destination stays at its reported value.

$$
\text{qty\_adj}_i = \frac{\text{qty\_raw}_i}{\sum \text{qty\_raw}} \times \text{total\_production}
$$

```{r}
#| label: tbl-c1-allocation-diagnostic
#| tbl-cap: "Production allocation consistency diagnostic"

tbl_c1_allocation_diagnostic |>
  gt() |>
  tab_header(
    title = "Production Allocation Consistency Diagnostic",
    subtitle = "Distribution of farmers by allocation status"
  ) |>
  cols_label(
    allocation_status = "Status",
    n_records = "Records",
    pct_records = "% Records"
  ) |>
  fmt_number(columns = n_records, decimals = 0, sep_mark = ",") |>
  fmt_number(columns = pct_records, decimals = 1, pattern = "{x}%") |>
  tab_style(
    style = cell_fill(color = "#ffebee"),
    locations = cells_body(rows = str_detect(allocation_status, "Excess"))
  ) |>
  tab_style(
    style = cell_fill(color = "#fff3e0"),
    locations = cells_body(rows = str_detect(allocation_status, "Deficit"))
  ) |>
  tab_style(
    style = cell_fill(color = "#e8f5e9"),
    locations = cells_body(rows = str_detect(allocation_status, "Balanced"))
  ) |>
  tab_source_note(
    source_note = md("**Excess:** Reported destinations exceed production. **Deficit:** Destinations below production. Status compares the sum of destinations (sales, household consumption, animal feed and seed) against total production.")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    table.width = pct(100)
  )
```

### Structural NA Handling

Variables with NA representing "no activity" are converted to explicit zeros.
Price variables remain NA for non-sellers since price is undefined without a
transaction.

```{r}
#| label: tbl-c2-structural-nas
#| tbl-cap: "Structural NAs converted to zero"

tbl_c2_structural_nas |>
  gt() |>
  cols_label(
    variable = "Variable",
    na_before = "NAs Before",
    na_after = "NAs After",
    converted = "Converted",
    action = "Action"
  ) |>
  cols_align(align = "left", columns = variable) |>
  cols_align(align = "right", columns = c(na_before, na_after, converted)) |>
  cols_align(align = "center", columns = action) |>
  tab_header(
    title = "Structural NAs Converted to Zero",
    subtitle = md("*NAs representing 'no activity' replaced with explicit zeros*")
  ) |>
  tab_spanner(
    label = "NA Count",
    columns = c(na_before, na_after)
  ) |>
  tab_style(
    style = cell_fill(color = "#e8f5e9"),
    locations = cells_body(rows = action == "Converted to 0", columns = action)
  ) |>
  tab_style(
    style = cell_fill(color = "#fff3e0"),
    locations = cells_body(rows = action == "Remains NA (price/rate)", columns = action)
  ) |>
  tab_source_note(
    source_note = "Annual rates convert to 0 when no land; price variables remain NA when undefined"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    table.width = pct(100)
  )
```

### Impossible Value Detection

Domain-based rules identify logically impossible values, converting them to NA
for subsequent imputation:

- **Harvest frequency > 3/year** — Exceeds Guatemala's agricultural calendar
- **Price outside 50–500 Q/qq** — Outside typical Guatemalan range (150–250 Q/qq)
- **Seed expenditure > 5,000 Q/mz** — Exceeds premium hybrid costs (~1,500 Q/mz max)
- **Household consumption > 15 qq/person/year** — Physiologically implausible (>680 kg)
- **Animal/seed reserves > 500 qq** — Exceeds plausible storage (>22 tons)
- **Yield > 160 qq/mz** — Exceeds biological limits for white maize (>7.3 tons/mz)

Additionally, `farmer_score` — a composite indicator of likelihood of using
non-hybrid varieties, computed from seed expenditure and yield patterns
(see [farmer segmentation](02_03_farmer_segmentation.qmd) for the full
scoring methodology) — has existing NAs from the scoring process that will
also be imputed.

```{r}
#| label: tbl-c3-impossible-values
#| tbl-cap: "Impossible values flagged for imputation"

tbl_c3_impossible_values |>
  gt() |>
  cols_label(
    criterion = "Detection Criterion",
    base_variable = "Variable Set to NA",
    n_cases = "Cases"
  ) |>
  tab_header(
    title = "Impossible Values Flagged for Imputation",
    subtitle = "Domain-based outlier detection"
  ) |>
  fmt_number(columns = n_cases, decimals = 0) |>
  tab_source_note(
    source_note = "These base variables will be imputed using Random Forest in the next step"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    table.width = pct(100)
  )
```

### Random Forest Imputation

Missing values in base variables are imputed using a Random Forest-based
multivariate imputation procedure. The method is non-parametric, handles 
mixed continuous and categorical data, and captures complex non-linear 
relationships between variables. Implementation uses the
`impute_rf_parallel()` utility function, which runs a multi-threaded `ranger`
engine and reports out-of-bag (OOB) error estimates separately for continuous
variables (OOBE) and categorical variables (PFC).

Variables imputed: `n_harvests_total`, `qty_sold_qq`, `exp_seeds`,
`qty_household_qq`, `qty_animal_seed_qq`, `total_production`, and `farmer_score`.

After imputation, structural values are restored for non-sellers
(`is_seller = 0`): `qty_sold_qq` is set to zero (non-sellers have no sales by
definition), and price and gain variables remain NA (undefined without a
transaction).

```{r}
#| label: tbl-c4-imputation-summary
#| tbl-cap: "Random Forest imputation results"

tbl_c4_imputation_summary |>
  gt() |>
  cols_label(
    variable = "Variable",
    na_before = "NAs Before",
    na_after = "NAs After",
    imputed = "Values Imputed",
    status = "Status"
  ) |>
  cols_align(align = "left", columns = variable) |>
  cols_align(align = "right", columns = c(na_before, na_after, imputed)) |>
  cols_align(align = "center", columns = status) |>
  tab_header(
    title = "Random Forest Imputation Results",
    subtitle = md("*Multivariate imputation via `impute_rf_parallel()` (ranger engine)*")
  ) |>
  grand_summary_rows(
    columns = c(na_before, imputed),
    fns = list(Total = ~sum(.)),
    fmt = ~fmt_number(., decimals = 0)
  ) |>
  tab_style(
    style = cell_fill(color = "#e8f5e9"),
    locations = cells_body(rows = status == "Complete", columns = status)
  ) |>
  tab_source_note(
    source_note = "Imputation parameters: maxiter = 10, ntree = 100, seed = 789"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    table.width = pct(100)
  )
```

::: {.callout-tip}
## Imputation Quality
Out-of-bag (OOB) error metrics — OOBE for continuous variables and PFC for
categorical variables — provide internal quality estimates calculated from
observations randomly excluded during tree construction within the Random Forest
algorithm. This avoids the need for a separate test set. OOBE values below 0.50
are considered acceptable; PFC values below 0.20 indicate acceptable categorical
accuracy.
:::

### Quality Treatment Summary

```{r}
#| label: tbl-c6-quality-summary
#| tbl-cap: "Complete data quality treatment pipeline"

tbl_c6_quality_summary |>
  gt() |>
  cols_label(
    step = "Step",
    action = "Action",
    description = "Description",
    records_affected = "Records Affected"
  ) |>
  cols_align(align = "center", columns = step) |>
  cols_align(align = "left", columns = c(action, description, records_affected)) |>
  tab_header(
    title = "Data Quality Treatment Summary",
    subtitle = md("*Systematic cleaning of agricultural dataset*")
  ) |>
  tab_row_group(label = "Recalculation & Correction", rows = 7:9) |>
  tab_row_group(label = "Imputation", rows = 5:6) |>
  tab_row_group(label = "Error Detection", rows = 4) |>
  tab_row_group(label = "Structural Values", rows = 2:3) |>
  tab_row_group(label = "Allocation Normalization", rows = 1) |>
  row_group_order(groups = c("Allocation Normalization", "Structural Values",
                             "Error Detection", "Imputation",
                             "Recalculation & Correction")) |>
  tab_source_note(
    source_note = "Final dataset: white maize farmers ready for MAGA weight calibration and downstream segmentation"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    table.width = pct(100),
    row_group.font.weight = "bold"
  )
```


## Summary

### Key Findings

1. **Unit heterogeneity**: A proportion of land area and production records
   required unit conversion, highlighting the importance of systematic
   standardization

2. **Yield distribution**: Median yield aligns with expected values for
   traditional Guatemalan maize varieties, validating the conversion approach

3. **Quality treatment**: The cleaning pipeline detected approximately
   `r n_values_flagged` impossible values across `r n_vars_imputed` base variables
   and imputed `r format(n_values_imputed, big.mark = ",")` values via Random
   Forest, while preserving structural NAs for undefined metrics
   (price and gain for non-sellers)

### Application

This cleaned dataset provides the foundation for subsequent analyses in this
module:

- **MAGA weight calibration** — Calibrate survey weights to official
  agricultural statistics
- **Farmer segmentation** — Identify farmer typologies based on land
  holdings, input use, and market orientation
- **Economic impact calculation** — Estimate potential gains from
  biofortified maize adoption
- **Adoption modeling** — Predict which farmers are most likely to
  adopt biofortified varieties
