---
title: "Calculate Market Baseline (`r params$seed_label`, `r params$subsidy_label`)"
subtitle: "Module 6: Scenario Precomputation — Production, consumption, and gravity-based redistribution"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
params:
  # Seed identification
  seed_year: "2022"
  seed_label: "2022 Seed"
  seed_suffix: "s22"
  # Subsidy identification
  subsidy_label: "Current subsidy"
  subsidy_suffix: "sub00"
---

```{r}
#| label: setup

# =============================================================================
# Setup: Load packages, configure environment, resolve parameters
# =============================================================================

# Clean environment (preserve params injected by Quarto)
rm(list = setdiff(ls(), "params"))
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  survey,             # Complex survey methodology
  srvyr,              # Tidy wrapper for survey package
  gt,                 # Professional table formatting
  tidyverse           # General purpose data manipulation
)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter
drop   <- base::drop

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file
source(here("02_code", "00_functions", "00_functions.R"))

# --- Resolve parameters into local variables -------------------------------

# Seed
seed_year   <- params$seed_year
seed_label  <- params$seed_label
seed_suffix <- params$seed_suffix

# Subsidy
subsidy_label  <- params$subsidy_label
subsidy_suffix <- params$subsidy_suffix

# Seed code for plot/table titles (e.g., "S22"). Derived from seed_suffix
# to keep seed_variants.yml unchanged: no additional field required.
seed_code <- toupper(seed_suffix)

# Composite file prefix for all output artefacts (parquet + web_resources rds)
file_prefix <- paste0("06_01_", seed_suffix, "_", subsidy_suffix)

# scenarios/
ensure_output_dir(here("01_data", "02_processed", "scenarios"))

# web_resources/
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

# Overview

**Seed Variant: `r paste0(seed_year, " (", seed_code, ")")` — Subsidy Regime: `r subsidy_label`** — This script uses farmer adoption scores from the `r seed_code` × `r toupper(subsidy_suffix)` economic impact pathway. For alternative combinations, see the companion variants rendered from this parametrized template.

This script establishes the market baseline metrics and redistribution infrastructure required for scenario modeling in Module 6. It calculates departmental production capacity, consumption requirements, and applies a gravity-based redistribution model to simulate inter-departmental surplus flows across the eleven simulation levels of national biofortified maize land share.

The simulation presents `penetration` as the national share of maize land cultivated with biofortified seed, evaluated at eleven levels from 0% to 100% in 10-percentage-point increments. The continuous adoption model operates on an internal adoption index `p_internal`, the position along the production-weighted adoption ranking defined in Script 02_06. Each farmer has a per-farmer entry point (`p_entry`) and a behavioural profile (`S`, `M`, `L`), which together determine the conversion ratio `r_i(p_internal) = sigmoid_curve(p_internal, p_entry_i, floor_profile, k_profile)`. At the farmer's entry point, the conversion ratio equals the profile-specific floor; it then rises along a sigmoid curve towards 1 as `p_internal` increases. At `p_internal = 1`, all farmers have `r_i(1) = 1` by construction. The effective biofortified production of farmer `i` is `production_biofortified_i × weight_calibrated_i × r_i(p_internal)`, and the effective biofortified maize land is `land_wtcorn_i × weight_calibrated_i × r_i(p_internal)`.

A fine internal grid maps the continuous adoption index to the resulting national biofortified land share. The eleven simulation levels are obtained by associating the internal adoption index with national biofortified maize land targets from 0% to 100% in 10-percentage-point increments. All farmer-level adoption calculations are evaluated at `p_internal`, and all outputs are labelled with the external `penetration` value.

The three outputs of this script behave differently with respect to the subsidy regime:

- **`06_01_{seed}_{sub}_market_baseline.parquet`** captures structural departmental quantities — farmer counts, land allocation, FSI consumption demand, current production, and maximum biofortified production capacity at full adoption. These are physical and demographic in nature and are therefore **invariant to the subsidy regime** for a given seed: the subsidy affects seed prices, producer absorption, and farmer extra costs, but does not alter yields, segment composition, consumption demand, or the road network.

- **`06_01_{seed}_{sub}_saturation_scenarios.parquet`** captures departmental coverage at the eleven levels of national biofortified maize land share (0–100%) after gravity-based redistribution. Adoption progresses as a continuous process driven by the Likely Adopters score (Script 02_06), which incorporates the economic-benefit component for the active subsidy regime. As the subsidy regime shifts the relative attractiveness of biofortified seed across farmers, the **order in which farmers enter the adoption pool changes**, and therefore both the internal adoption index associated with each land share level and the spatial distribution of biofortified production at intermediate levels also change. The endpoints (0% and 100% biofortified maize land) converge across subsidy regimes by construction, since they correspond respectively to an empty adoption pool and the full farmer set with `r_i(1) = 1`.

- **`06_01_{seed}_{sub}_penetration_lookup.parquet`** records the association between each external `penetration` level and its internal adoption index `p_internal`, together with the biofortified maize land share achieved at that index. Downstream scenario scripts evaluate farmer-level adoption at the same internal adoption index that generates the departmental coverage stored in the saturation scenarios. The association is **subsidy-sensitive** at intermediate levels for the same reason as the saturation scenarios.

## Main Operations

1. Load farmer data with adoption ranking and continuous adoption parameters (Module 2 output)
2. Load ENCOVI 2023 consumption data for departmental maize demand (Module 1 output)
3. Load OSRM-derived travel time matrix and gravity weights (06_00 output)
4. Load calibrated model configuration for continuous adoption curves (02_05 output)
5. Calculate baseline farmer counts by segment (total, hybrid, OPV) per department
6. Calculate baseline land allocation for maize production per department
7. Calculate annual maize consumption by department using survey weights
8. Calculate maximum current and biofortified production capacity per department
9. Build the internal adoption grid and the national biofortified maize land share at each grid value
10. Associate the eleven simulation levels with their internal adoption index
11. Validate the land share scale and report the verification table and curve
12. Generate saturation diagnostic curves before redistribution
13. Apply gravity-based redistribution across the eleven simulation levels
14. Validate production conservation through redistribution
15. Export market baseline, saturation scenarios, and land share lookup

## Methodological Notes

**Biofortified Land Share Scale**: The national biofortified maize land share at internal adoption index `p_internal` is `Σ(land_wtcorn_i × r_i(p_internal) × weight_calibrated_i) / Σ(land_wtcorn_i × weight_calibrated_i)`. Numerator and denominator are expressed in manzanas, so the ratio requires no unit conversion. The share is non-decreasing in `p_internal`, equals 0 at `p_internal = 0` (empty adoption pool) and equals 1 at `p_internal = 1` (full adoption).

**Gravity-Based Redistribution**: Surplus production flows from departments with excess to all departments with deficit, weighted proportionally to the deficit at the destination and inversely proportional to the real travel time between departments. The formulation follows Anderson & van Wincoop (2003), analogous to IMPLAN inter-regional commodity flow estimation.

**Travel Time Matrix**: A 22×22 asymmetric matrix of driving times between departmental capitals, computed via OSRM over OpenStreetMap road network data. Asymmetry reflects real directional differences in Guatemala's road infrastructure (terrain gradients, one-way routing, etc.).

**Continuous Adoption Model**: The conversion ratio `r_i(p_internal)` determines the fraction of each farmer's potential biofortified production and maize land that is realised at internal adoption index `p_internal`. The model parameters (per-profile floors and steepness coefficients) are loaded from the calibrated model configuration produced by Script 02_05.

# Phase A: Data Loading

This phase loads the four data sources required for market baseline calculations:

1. **Farmer adoption data** (02_06): Contains farmer rankings, economic metrics, production capacity, continuous adoption parameters (`p_entry`, `profile`), and survey weights for the active seed × subsidy combination
2. **Individual nutrient intake** (01_04): Contains maize consumption data at individual level with survey design variables
3. **Travel time matrix** (06_00): Contains OSRM-derived gravity weights for inter-departmental redistribution
4. **Model configuration** (02_05): Contains calibrated continuous adoption curve parameters (per-profile floors and steepness coefficients)

## A1. Load Farmer Adoption Data

Load the complete farmer dataset with adoption scores, economic impact calculations, and segment assignments. This parquet is produced by `02_06x_apply_likely_adopters_model.qmd` for the `r seed_suffix` × `r subsidy_suffix` combination and integrates upstream data from segmentation (02_03) and economic impact (02_04).

```{r}
#| label: a1-load-farmer-data

# =============================================================================
# A1. Load Farmer Adoption Data
# =============================================================================

df_farmers_raw <- read_parquet(
  here("01_data", "02_processed",
       "scenarios", paste0("02_06_", seed_suffix, "_", subsidy_suffix,
                           "_farmers_adoption_scores.parquet"))
)
```

## A2. Load Model Configuration for Continuous Adoption Curves

Load the calibrated model configuration from Script 02_05 and extract the six continuous adoption parameters. These parameters govern the sigmoid conversion curves that determine the fraction of biofortified production realised by each farmer at each value of the internal adoption index.

```{r}
#| label: a2-load-model-config

# =============================================================================
# A2. Load Model Configuration and Extract Continuous Adoption Parameters
# =============================================================================

model_config <- readRDS(
  here("01_data", "02_processed", "models",
       "02_05_likely_adopters_model_config.rds")
)

# Extract continuous adoption curve parameters (per-profile floors and steepness)
floor_S <- model_config$continuous$floor_S
floor_M <- model_config$continuous$floor_M
floor_L <- model_config$continuous$floor_L
k_S     <- model_config$continuous$k_S
k_M     <- model_config$continuous$k_M
k_L     <- model_config$continuous$k_L

# Clean up: only the six scalar parameters are needed downstream
rm(model_config)
```

## A3. Load Individual Nutrient Intake Data

Load the individual-level nutrient intake dataset containing daily maize consumption and survey design variables. This data will be aggregated to departmental level using survey weights.

```{r}
#| label: a3-load-nutritional-data

# =============================================================================
# A3. Load Individual Nutritional Intake Data
# =============================================================================

df_individuals_raw <- read_parquet(
  here("01_data", "02_processed",
       "transfer", "01_04_individual_nutritional_intake_complete.parquet")
)
```

## A4. Load Travel Time Matrix and Gravity Weights

Load the pre-computed OSRM travel time matrix and gravity weights for the 22 departmental capitals. This matrix is computed once by `06_00_compute_travel_time_matrix.qmd` and used across all seed variants.

**Source of coordinates**: Departmental capitals (cabeceras departamentales) are the administrative seats of each of Guatemala's 22 departments. Coordinates were obtained from the GeoNames geographical database (www.geonames.org, Creative Commons Attribution 4.0 License), which aggregates data from official national mapping agencies and the OpenStreetMap community. Each coordinate corresponds to the populated place record classified as `feature class P` (populated place) and `feature code PPLA` (seat of a first-order administrative division) in GeoNames for the respective departmental capital. GeoNames data for Guatemala can be accessed at https://download.geonames.org/export/dump/GT.zip. Coordinates are in WGS84 (EPSG:4326).

```{r}
#| label: a4-load-travel-time

# =============================================================================
# A4. Load Travel Time Matrix and Gravity Weights
# =============================================================================

df_travel_time <- read_parquet(
  here("01_data", "02_processed",
       "external", "06_00_osrm_travel_time_matrix.parquet")
)
```

## A5. Data Dimensions Summary

Verify the loaded datasets contain expected records and display key dimensions for audit trail.

```{r}
#| label: tbl-a5-data-dimensions
#| tbl-cap: "Input datasets for market baseline calculations"

# =============================================================================
# A5. Summarize Data Dimensions
# =============================================================================

# Create dimensions summary
tbl_dimensions <- tibble(
  dataset = c(paste0("Farmers (02_06_", seed_suffix, "_", subsidy_suffix, ")"),
              "Model configuration (02_05)",
              "Individuals (01_04)",
              "Travel time matrix (06_00)"),
  n_records = c(nrow(df_farmers_raw), NA_integer_, nrow(df_individuals_raw),
                nrow(df_travel_time)),
  n_variables = c(ncol(df_farmers_raw), NA_integer_, ncol(df_individuals_raw),
                  ncol(df_travel_time)),
  notes = c("Per-farmer adoption scores, p_entry, profile",
            paste0("6 continuous params: floor_S=", floor_S,
                   ", floor_M=", floor_M, ", floor_L=", floor_L,
                   ", k_S=", k_S, ", k_M=", k_M, ", k_L=", k_L),
            "Individual nutritional intake with survey weights",
            "22x22 asymmetric OSRM travel time matrix")
)

# Export for web resources
export(tbl_dimensions, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_a5_data_dimensions.rds")))

# Display as GT table
tbl_dimensions |>
  gt() |>
  tab_header(
    title = "Phase A: Data Loading Summary",
    subtitle = "Input datasets for market baseline calculations"
  ) |>
  cols_label(
    dataset = "Dataset",
    n_records = "Records",
    n_variables = "Variables",
    notes = "Notes"
  ) |>
  cols_align(align = "left", columns = c(dataset, notes)) |>
  fmt_number(
    columns = c(n_records, n_variables),
    use_seps = TRUE,
    decimals = 0
  ) |>
  sub_missing(missing_text = "config") |>
  tab_source_note(
    source_note = "Source: ENCOVI 2023 dataset + OSRM travel time matrix + 02_05 model configuration"
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Clean up
rm(tbl_dimensions)
```

## A6. Travel Time Matrix Visualization

Display the OSRM-derived travel time matrix as a heatmap. This visualization confirms that the geographic distance structure is consistent with Guatemala's road network topology.

```{r}
#| label: fig-a6-travel-time-heatmap
#| fig-cap: "Driving travel times between departmental capitals (minutes) — OSRM asymmetric matrix"
#| fig-width: 10
#| fig-height: 10

# =============================================================================
# A6. Travel Time Matrix Heatmap
# =============================================================================

# Department ordering (INE standard)
dept_order <- c("Guatemala", "El Progreso", "Sacatep\u00e9quez", "Chimaltenango",
                "Escuintla", "Santa Rosa", "Solol\u00e1", "Totonicap\u00e1n",
                "Quetzaltenango", "Suchitep\u00e9quez", "Retalhuleu", "San Marcos",
                "Huehuetenango", "Quich\u00e9", "Baja Verapaz", "Alta Verapaz",
                "Pet\u00e9n", "Izabal", "Zacapa", "Chiquimula", "Jalapa", "Jutiapa")

# Prepare plot data
df_travel_plot <- df_travel_time |>
  select(origin, destination, travel_time_min) |>
  mutate(
    origin = factor(origin, levels = dept_order),
    destination = factor(destination, levels = rev(dept_order))
  )

# Export for web resources
export(df_travel_plot, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_fig_a6_travel_time_heatmap.rds")))

# Heatmap
ggplot(df_travel_plot, aes(x = origin, y = destination, fill = travel_time_min)) +
  geom_tile(color = "white", linewidth = 0.3) +
  geom_text(
    aes(label = round(travel_time_min, 0)),
    size = 2.2,
    color = "gray20"
  ) +
  scale_fill_gradient2(
    low = "#E8F0F7",
    high = "#1B6CA8",
    midpoint = median(df_travel_plot$travel_time_min),
    name = "Travel Time\n(minutes)"
  ) +
  labs(
    title = "Driving Travel Times Between Departmental Capitals",
    subtitle = "OSRM routing over OpenStreetMap road network \u2014 asymmetric (directional)",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.text.y = element_text(size = 8),
    plot.title = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "gray40", size = 11),
    legend.position = "bottom",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.4, "cm"),
    panel.grid = element_blank()
  )

# Clean up
rm(df_travel_plot)
```

## A7. Gravity Weights Summary

```{r}
#| label: tbl-a7-gravity-summary
#| tbl-cap: "Gravity weight matrix summary statistics"

# =============================================================================
# A7. Gravity Weights Summary
# =============================================================================

df_gravity_off_diag <- df_travel_time |>
  filter(origin != destination)

tbl_gravity_summary <- tibble(
  metric = c(
    "Beta (decay parameter)",
    "Number of department pairs",
    "Min gravity weight",
    "Max gravity weight",
    "Mean gravity weight",
    "Min travel time (min)",
    "Max travel time (min)",
    "Mean travel time (min)"
  ),
  value = c(
    as.character(unique(df_travel_time$beta)),
    nrow(df_gravity_off_diag),
    format(min(df_gravity_off_diag$gravity_weight), scientific = TRUE, digits = 3),
    format(max(df_gravity_off_diag$gravity_weight), scientific = TRUE, digits = 3),
    format(mean(df_gravity_off_diag$gravity_weight), scientific = TRUE, digits = 3),
    round(min(df_gravity_off_diag$travel_time_min), 1),
    round(max(df_gravity_off_diag$travel_time_min), 1),
    round(mean(df_gravity_off_diag$travel_time_min), 1)
  )
)

# Export for web resources
export(tbl_gravity_summary, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_a7_gravity_summary.rds")))

tbl_gravity_summary |>
  gt() |>
  tab_header(
    title = "Gravity Weight Matrix Summary",
    subtitle = "w(i,j) = 1 / travel_time(i,j)\u00b2 \u2014 Anderson & van Wincoop (2003)"
  ) |>
  cols_label(metric = "Metric", value = "Value") |>
  cols_align(align = "left", columns = metric) |>
  cols_align(align = "right", columns = value) |>
  tab_source_note(
    source_note = "Source: 06_00_osrm_travel_time_matrix.parquet"
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Clean up
rm(df_gravity_off_diag, tbl_gravity_summary)
```

---

# Phase B: Baseline Metrics Calculation

This phase calculates departmental baseline metrics required for scenario modeling. All calculations use survey weights to produce population-representative estimates. Biofortified production aggregations apply the continuous adoption model explicitly: at full adoption (`p_internal = 1`), every farmer has `r_i(1) = 1` by construction of the sigmoid curve, so `production_biofortified × weight_calibrated × r_i(1) = production_biofortified × weight_calibrated`. The formulas below include the `r_i(1)` factor for methodological transparency.

## B1. Calculate Farmer Totals by Department

Calculate the total number of maize farmers by department, segmented by technology adoption level (OPV/Criollo, Low, Mid, High). These totals serve as denominators for penetration percentage calculations.

```{r}
#| label: b1-farmer-totals

# =============================================================================
# B1. Calculate Farmer Totals by Department
# =============================================================================

# Create survey design for farmers
svy_farmers <- df_farmers_raw |>
  as_survey_design(
    ids = 1,
    weights = weight_calibrated,
    strata = NULL
  )

# Calculate farmer counts by department and segment
df_farmers_by_segment <- svy_farmers |>
  summarise(
    n_farmers = survey_total(vartype = NULL, na.rm = TRUE),
    .by = c(departamento, segment_final)
  ) |>
  as_tibble()

# Pivot to wide format for easier use
df_baseline_farmers <- df_farmers_by_segment |>
  pivot_wider(
    names_from = segment_final,
    values_from = n_farmers,
    names_prefix = "n_farmers_",
    values_fill = 0
  ) |>
  mutate(
    n_farmers_total = n_farmers_High + n_farmers_Mid +
                      n_farmers_Low + `n_farmers_OPV/Criollo`,
    n_farmers_hybrid = n_farmers_High + n_farmers_Mid + n_farmers_Low,
    n_farmers_opv = `n_farmers_OPV/Criollo`
  ) |>
  select(
    departamento,
    n_farmers_total,
    n_farmers_hybrid,
    n_farmers_opv,
    n_farmers_high = n_farmers_High,
    n_farmers_mid = n_farmers_Mid,
    n_farmers_low = n_farmers_Low
  )

# Clean up
rm(df_farmers_by_segment)
```

## B2. Calculate Land Totals by Department

Calculate total maize cultivation area by department and segment. Land is measured in manzanas (mz) and converted to hectares (ha) using the standard conversion factor (1 mz = 0.7 ha).

```{r}
#| label: b2-land-totals

# =============================================================================
# B2. Calculate Land Totals by Department
# =============================================================================

# Calculate land by department and segment
df_land_by_segment <- svy_farmers |>
  summarise(
    land_mz = survey_total(land_wtcorn, vartype = NULL, na.rm = TRUE),
    .by = c(departamento, segment_final)
  ) |>
  as_tibble()

# Pivot to wide format
df_baseline_land <- df_land_by_segment |>
  pivot_wider(
    names_from = segment_final,
    values_from = land_mz,
    names_prefix = "land_mz_",
    values_fill = 0
  ) |>
  mutate(
    land_mz_total = land_mz_High + land_mz_Mid +
                    land_mz_Low + `land_mz_OPV/Criollo`,
    land_mz_hybrid = land_mz_High + land_mz_Mid + land_mz_Low,
    land_mz_opv = `land_mz_OPV/Criollo`,
    # Convert to hectares (1 mz = 0.7 ha)
    land_ha_total = land_mz_total * 0.7,
    land_ha_hybrid = land_mz_hybrid * 0.7,
    land_ha_opv = land_mz_opv * 0.7
  ) |>
  select(
    departamento,
    land_mz_total,
    land_mz_hybrid,
    land_mz_opv,
    land_ha_total,
    land_ha_hybrid,
    land_ha_opv
  )

# Clean up
rm(df_land_by_segment)
```

## B3. Calculate Annual Maize Consumption by Department

Calculate total annual maize consumption by department using survey-weighted population estimates and calibrated per capita consumption from official sources.

**Consumption estimates from two sources:**

1. **Direct human consumption**: 110 kg/capita/year based on FAO/MAGA (2008), which states: *"El consumo per capita de ma\u00edz en Guatemala es de 110 kg/a\u00f1o (utilizaci\u00f3n directa)"*. This covers only tortillas and direct food use.

2. **Total FSI consumption** (Food, Seed, Industrial): 174 kg/capita/year based on USDA (2019) Grain and Feed Annual report. This includes all food-related uses: direct human consumption, backyard animal feed, seed retention, and post-harvest losses.

Coverage calculations use the **FSI total** as the denominator, since production must meet all domestic demand, not just direct human consumption.

**Conversion factors:**

- Kilograms to quintales: \u00f7 45.36 (1 qq = 100 lb = 45.36 kg)
- Quintales to metric tons: \u00d7 0.04536 (1 qq \u2248 0.04536 MT)

```{r}
#| label: b3-annual-consumption

# =============================================================================
# B3. Calculate Annual Maize Consumption by Department
# =============================================================================

# Define consumption parameters from official sources
CONSUMPTION_HUMAN_KG <- 110
CONSUMPTION_FSI_KG <- 174

# Conversion factor
KG_TO_QQ <- 45.36

# Create survey design for individuals
svy_individuals <- df_individuals_raw |>
  as_survey_design(
    ids = upm,
    weights = factor,
    strata = NULL
  )

# Calculate population by department
df_population_by_dept <- svy_individuals |>
  summarise(
    population = survey_total(vartype = NULL, na.rm = TRUE),
    .by = departamento
  ) |>
  as_tibble()

# Calculate consumption estimates (human and FSI)
df_baseline_consumption <- df_population_by_dept |>
  mutate(
    # Human consumption: 110 kg/capita/year (FAO/MAGA 2008)
    consumption_human_qq = population * CONSUMPTION_HUMAN_KG / KG_TO_QQ,
    consumption_human_mt = consumption_human_qq * 0.04536,

    # FSI consumption: 174 kg/capita/year (USDA 2019)
    consumption_fsi_qq = population * CONSUMPTION_FSI_KG / KG_TO_QQ,
    consumption_fsi_mt = consumption_fsi_qq * 0.04536
  )

# Clean up
rm(df_population_by_dept, CONSUMPTION_HUMAN_KG, CONSUMPTION_FSI_KG, KG_TO_QQ)
```

## B4. Calculate Maximum Current Production by Department

Calculate the maximum current maize production capacity by department if all farmers continue with their current practices. This represents the baseline production scenario (0% biofortified maize land).

```{r}
#| label: b4-current-production

# =============================================================================
# B4. Calculate Maximum Current Production by Department
# =============================================================================

# Calculate current production totals by department
df_baseline_current_prod <- svy_farmers |>
  summarise(
    current_production_qq = survey_total(
      total_production,
      vartype = NULL,
      na.rm = TRUE
    ),
    .by = departamento
  ) |>
  as_tibble() |>
  mutate(
    current_production_mt = current_production_qq * 0.04536
  )
```

## B5. Calculate Maximum Biofortified Production by Department

Calculate the maximum potential biofortified maize production by department at full adoption. This represents the ceiling scenario (100% biofortified maize land), where every farmer has `r_i(1) = 1` by the convergence property of the sigmoid curve, so the effective biofortified production equals `production_biofortified × weight_calibrated × 1`.

```{r}
#| label: b5-biofortified-production

# =============================================================================
# B5. Calculate Maximum Biofortified Production by Department
# =============================================================================

# At p = 1, r_i(1) = 1 for all farmers by construction of the sigmoid curve.
# The survey_total below is therefore equivalent to:
#   sum(production_biofortified * weight_calibrated * r_i(1))
# = sum(production_biofortified * weight_calibrated * 1)
# = sum(production_biofortified * weight_calibrated)

# Calculate biofortified production totals by department
df_baseline_bio_prod <- svy_farmers |>
  summarise(
    max_bio_production_qq = survey_total(
      production_biofortified,
      vartype = NULL,
      na.rm = TRUE
    ),
    .by = departamento
  ) |>
  as_tibble() |>
  mutate(
    max_bio_production_mt = max_bio_production_qq * 0.04536
  )
```

## B6. Consolidate Baseline Metrics

Join all calculated metrics into a single departmental baseline dataset and generate the summary table for validation.

```{r}
#| label: b6-consolidate-baseline

# =============================================================================
# B6. Consolidate All Baseline Metrics
# =============================================================================

# Convert departamento to character in all datasets before joining
df_baseline_farmers_chr <- df_baseline_farmers |>
  mutate(departamento = as.character(departamento))

df_baseline_land_chr <- df_baseline_land |>
  mutate(departamento = as.character(departamento))

df_baseline_consumption_chr <- df_baseline_consumption |>
  mutate(departamento = as.character(departamento))

df_baseline_current_prod_chr <- df_baseline_current_prod |>
  mutate(departamento = as.character(departamento))

df_baseline_bio_prod_chr <- df_baseline_bio_prod |>
  mutate(departamento = as.character(departamento))

# Join all baseline metrics by department
df_market_baseline <- df_baseline_farmers_chr |>
  left_join(df_baseline_land_chr, by = "departamento") |>
  left_join(df_baseline_consumption_chr, by = "departamento") |>
  left_join(df_baseline_current_prod_chr, by = "departamento") |>
  left_join(df_baseline_bio_prod_chr, by = "departamento") |>
  # Calculate coverage ratios at 100% penetration (against FSI consumption)
  mutate(
    max_coverage_ratio = max_bio_production_qq / consumption_fsi_qq,
    max_coverage_pct = pmin(max_coverage_ratio * 100, 100),
    has_excess = max_bio_production_qq > consumption_fsi_qq,
    excess_production_qq = pmax(max_bio_production_qq - consumption_fsi_qq, 0)
  ) |>
  # Restore departamento as ordered factor for consistency
  mutate(
    departamento = factor(
      departamento,
      levels = dept_order,
      ordered = TRUE
    )
  ) |>
  arrange(departamento)

# Clean up intermediate objects
rm(df_baseline_farmers_chr, df_baseline_land_chr, df_baseline_consumption_chr,
   df_baseline_current_prod_chr, df_baseline_bio_prod_chr)
```

## B7. Baseline Summary Table

Display consolidated baseline metrics for validation. This table shows key departmental metrics that will be used in scenario calculations.

```{r}
#| label: tbl-b7-baseline-summary
#| tbl-cap: "Departmental baseline metrics for scenario modeling"

# =============================================================================
# B7. Generate Baseline Summary Table
# =============================================================================

# Prepare table data with formatted columns
tbl_baseline_summary <- df_market_baseline |>
  transmute(
    departamento,
    n_farmers_total,
    n_farmers_hybrid,
    n_farmers_opv,
    land_ha_total,
    consumption_human_mt,
    consumption_fsi_mt,
    current_production_mt,
    max_bio_production_mt,
    max_coverage_pct
  )

# Export for web resources
export(tbl_baseline_summary, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_b7_baseline_summary.rds")))

# Generate GT table
tbl_baseline_summary |>
  gt() |>
  tab_header(
    title = "Departmental Market Baseline Metrics",
    subtitle = "Farmers, land, consumption, and production capacity by department"
  ) |>
  cols_label(
    departamento = "Department",
    n_farmers_total = "Total Farmers",
    n_farmers_hybrid = "Hybrid Farmers",
    n_farmers_opv = "OPV Farmers",
    land_ha_total = "Land (ha)",
    consumption_human_mt = "Human Cons. (MT/yr)",
    consumption_fsi_mt = "FSI Cons. (MT/yr)",
    current_production_mt = "Current Prod. (MT)",
    max_bio_production_mt = "Max Bio Prod. (MT)",
    max_coverage_pct = "Max Coverage (%)"
  ) |>
  cols_align(align = "left", columns = departamento) |>
  cols_align(align = "right", columns = -departamento) |>
  fmt_number(
    columns = c(n_farmers_total, n_farmers_hybrid, n_farmers_opv),
    use_seps = TRUE,
    decimals = 0
  ) |>
  fmt_number(
    columns = c(land_ha_total, consumption_human_mt, consumption_fsi_mt,
                current_production_mt, max_bio_production_mt),
    use_seps = TRUE,
    decimals = 1
  ) |>
  fmt_number(
    columns = max_coverage_pct,
    decimals = 1,
    pattern = "{x}%"
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(
      columns = max_coverage_pct,
      rows = max_coverage_pct >= 100
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(
      columns = max_coverage_pct,
      rows = max_coverage_pct < 50
    )
  ) |>
  grand_summary_rows(
    columns = c(n_farmers_total, n_farmers_hybrid, n_farmers_opv,
                land_ha_total, consumption_human_mt, consumption_fsi_mt,
                current_production_mt, max_bio_production_mt),
    fns = list(
      Total = ~ sum(., na.rm = TRUE)
    ),
    fmt = ~ fmt_number(., use_seps = TRUE, decimals = 0)
  ) |>
  tab_source_note(
    source_note = "Source: ENCOVI 2023 (MAGA-calibrated weights). Human Cons. = direct human consumption at 110 kg/capita/year (FAO/MAGA 2008). FSI Cons. = Food, Seed & Industrial consumption at 174 kg/capita/year (USDA 2019). Coverage calculated against FSI. Green = surplus (\u2265100%). Red = deficit (<50%)."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    column_labels.font.size = px(10)
  )

# Clean up
rm(tbl_baseline_summary)
```

---

# Phase C: Biofortified Land Share Scale

This phase establishes the scale on which the eleven simulation levels are defined. The simulation presents `penetration` as the national share of maize land cultivated with biofortified seed. The continuous adoption model operates on the internal adoption index `p_internal`, the position along the production-weighted adoption ranking defined in Script 02_06. A fine internal grid maps the continuous adoption index to the resulting national biofortified land share, and each simulation level is associated with the value of `p_internal` whose biofortified land share is closest to its target.

The association is specific to each seed × subsidy combination, since the adoption ranking and the per-farmer entry points depend on the economic scenario. The computation relies exclusively on farmer-level adoption parameters and precedes the gravity-based redistribution, which is evaluated only at the eleven associated values of `p_internal`.

## C1. Internal Adoption Grid

The grid evaluates the internal adoption index at 10,001 equally spaced values between 0 and 1 (step 0.0001). At each grid value, farmers with `p_entry <= p_internal` form the adoption pool, and each admitted farmer converts `land_wtcorn_i × r_i(p_internal)` manzanas of maize land to biofortified seed. The value `p_internal = 0` denotes the empty adoption pool. The national biofortified maize land share at each grid value is:

`land_share(p_internal) = Σ(land_wtcorn_i × r_i(p_internal) × weight_calibrated_i) / Σ(land_wtcorn_i × weight_calibrated_i)`

Numerator and denominator are expressed in manzanas, so the share requires no unit conversion. The denominator is the national maize land and is computed once for the whole grid.

The resulting table `df_penetration_grid` contains, for each grid value, the internal adoption index (`p_internal`), the biofortified maize land (`land_bio`), the national maize land (`land_total`), and the biofortified land share as a proportion (`land_share`) and as a percentage (`land_pct`).

```{r}
#| label: c1-land-share-grid

# =============================================================================
# C1. Build Internal Adoption Grid and National Biofortified Land Share
# =============================================================================

# Grid resolution: 10,000 equal intervals on [0, 1]. Integer division keeps
# both endpoints exactly at 0 and 1.
GRID_N_INTERVALS <- 10000

p_internal_grid <- (0:GRID_N_INTERVALS) / GRID_N_INTERVALS

# Farmer-level adoption inputs (vectors reused at every grid value)
farmer_p_entry   <- df_farmers_raw$p_entry
farmer_profile   <- df_farmers_raw$profile
farmer_land_mz   <- df_farmers_raw$land_wtcorn
farmer_weight    <- df_farmers_raw$weight_calibrated

# National maize land in manzanas (denominator, constant across the grid)
land_total_mz <- sum(farmer_land_mz * farmer_weight, na.rm = TRUE)

# Biofortified maize land at each value of the internal adoption index
land_bio_grid <- map_dbl(
  p_internal_grid,
  function(p_int) {

    # Farmers in the adoption pool at this internal adoption index
    # (p_internal = 0 denotes the empty adoption pool)
    is_admitted <- farmer_p_entry <= p_int & p_int > 0

    # Per-farmer conversion ratio r_i(p_internal)
    r_i <- apply_adoption_curve(p_int,
                                farmer_p_entry[is_admitted],
                                farmer_profile[is_admitted],
                                floor_S, floor_M, floor_L,
                                k_S, k_M, k_L)

    # Weighted biofortified maize land (manzanas)
    sum(farmer_land_mz[is_admitted] * r_i * farmer_weight[is_admitted],
        na.rm = TRUE)
  },
  .progress = TRUE
)

# Complete equivalence table: internal adoption index -> biofortified land share
df_penetration_grid <- tibble(
  p_internal = p_internal_grid,
  land_bio   = land_bio_grid
) |>
  mutate(
    land_total = land_total_mz,
    land_share = land_bio / land_total,
    land_pct   = land_share * 100
  ) |>
  arrange(p_internal)

# Clean up
rm(GRID_N_INTERVALS, p_internal_grid, farmer_p_entry, farmer_profile,
   farmer_land_mz, farmer_weight, land_total_mz, land_bio_grid)
```

## C2. Simulation Level Association

The eleven simulation levels correspond to national biofortified maize land targets from 0% to 100% in 10-percentage-point increments. For each target, the association selects the grid value of `p_internal` whose biofortified land share is closest to the target; among grid values at equal distance, the lowest internal adoption index is selected. The endpoints are fixed explicitly: the 0% level corresponds to `p_internal = 0` (empty adoption pool) and the 100% level corresponds to `p_internal = 1` (full adoption, with `r_i(1) = 1` for every farmer).

The resulting table `df_penetration_lookup` separates the two values used by the simulation: `penetration` is the external level that labels every output, and `p_internal` is the internal value at which all farmer-level adoption calculations are evaluated. The table also records the biofortified land achieved at `p_internal` and the association error, both as a proportion (`error`) and in percentage points (`error_pp`).

```{r}
#| label: c2-penetration-lookup

# =============================================================================
# C2. Associate Simulation Levels with the Internal Adoption Index
# =============================================================================

# External simulation levels: national biofortified maize land share targets
penetration_targets <- seq(0, 1, by = 0.10)

# Nearest grid value for each target (lowest index among equal distances)
nearest_grid_index <- map_int(
  penetration_targets,
  ~ which.min(abs(df_penetration_grid$land_share - .x))
)

# Lookup table: external simulation level -> internal adoption index
df_penetration_lookup <- tibble(
  penetration        = penetration_targets,
  nearest_grid_index = nearest_grid_index
) |>
  mutate(
    # Endpoints fixed explicitly: first grid value (empty adoption pool) and
    # last grid value (full adoption)
    grid_index = case_when(
      near(penetration, 0) ~ 1L,
      near(penetration, 1) ~ nrow(df_penetration_grid),
      TRUE                 ~ nearest_grid_index
    ),
    p_internal          = df_penetration_grid$p_internal[grid_index],
    land_bio            = df_penetration_grid$land_bio[grid_index],
    land_total          = df_penetration_grid$land_total[grid_index],
    land_share_achieved = df_penetration_grid$land_share[grid_index],
    land_pct_achieved   = land_share_achieved * 100,
    error               = land_share_achieved - penetration,
    error_pp            = error * 100
  ) |>
  select(penetration, p_internal, land_bio, land_total,
         land_share_achieved, land_pct_achieved, error, error_pp)

# Clean up
rm(nearest_grid_index)
```

## C3. Land Share Scale Validation

The checks below confirm the integrity of the grid and of the simulation level association before the simulation proceeds. The grid covers [0, 1] in strictly increasing order without missing values, and its biofortified land share lies within [0, 1] and is non-decreasing along the internal adoption index; plateaus are admissible. The lookup contains exactly the eleven simulation levels, with a non-decreasing internal adoption index and achieved land share, the 0% level anchored at the empty adoption pool, the 100% level anchored at full adoption with a land share equal to 1 within the association tolerance, and an association error within 0.25 percentage points. A farmer whose entry point equals 1, at the very end of the ranking, stays at the profile floor at `p_internal = 1` through the numerical guard of `sigmoid_curve()`, so the full land share holds up to a deviation of the order of 10⁻⁷. The script halts if any check fails.

```{r}
#| label: tbl-c3-scale-validation
#| tbl-cap: "Biofortified land share scale validation (grid and simulation level association)"

# =============================================================================
# C3. Validate Biofortified Land Share Scale
# =============================================================================

# Tolerances
NUMERIC_TOLERANCE   <- 1e-9   # Floating-point tolerance for bounds and order
LOOKUP_TOLERANCE_PP <- 0.25   # Maximum association error (percentage points)

# Grid variables required by the association
grid_required_vars <- df_penetration_grid |>
  select(p_internal, land_bio, land_total, land_share, land_pct)

# Steps between consecutive values
grid_land_share_steps   <- diff(df_penetration_grid$land_share)
lookup_p_internal_steps <- diff(df_penetration_lookup$p_internal)
lookup_land_share_steps <- diff(df_penetration_lookup$land_share_achieved)

# Endpoint rows of the lookup
lookup_first <- slice_head(df_penetration_lookup, n = 1)
lookup_last  <- slice_tail(df_penetration_lookup, n = 1)

# Maximum absolute association error
max_abs_error_pp <- max(abs(df_penetration_lookup$error_pp))

# Validation table
tbl_scale_validation <- tibble(
  scope = c(rep("Grid", 5), rep("Lookup", 7)),
  check = c(
    "Internal adoption index within [0, 1]",
    "Internal adoption index in strictly increasing order",
    "No missing values in grid variables",
    "Biofortified land share within [0, 1]",
    "Biofortified land share non-decreasing (plateaus admissible)",
    "Exactly 11 simulation levels",
    "Simulation levels equal to 0, 0.1, ..., 1",
    "Internal adoption index non-decreasing across levels",
    "Achieved land share non-decreasing across levels",
    "0% level at the empty adoption pool (p_internal = 0, land share = 0)",
    "100% level at full adoption (p_internal = 1, land share = 1 within association tolerance)",
    paste0("Maximum association error within ", LOOKUP_TOLERANCE_PP, " pp")
  ),
  value = c(
    sprintf("[%.4f, %.4f]",
            min(df_penetration_grid$p_internal), max(df_penetration_grid$p_internal)),
    paste0(format(nrow(df_penetration_grid), big.mark = ","), " grid values"),
    paste0(sum(is.na(grid_required_vars)), " missing values"),
    sprintf("[%.6f, %.6f]",
            min(df_penetration_grid$land_share), max(df_penetration_grid$land_share)),
    sprintf("Minimum step: %.2e", min(grid_land_share_steps)),
    paste0(nrow(df_penetration_lookup), " levels"),
    paste(df_penetration_lookup$penetration, collapse = ", "),
    sprintf("Minimum step: %.4f", min(lookup_p_internal_steps)),
    sprintf("Minimum step: %.4f", min(lookup_land_share_steps)),
    sprintf("p_internal = %.4f; land share = %.6f",
            lookup_first$p_internal, lookup_first$land_share_achieved),
    sprintf("p_internal = %.4f; deviation from full land share = %.2e",
            lookup_last$p_internal, 1 - lookup_last$land_share_achieved),
    sprintf("%.3f pp", max_abs_error_pp)
  ),
  passed = c(
    all(df_penetration_grid$p_internal >= 0 & df_penetration_grid$p_internal <= 1),
    !is.unsorted(df_penetration_grid$p_internal, strictly = TRUE),
    !anyNA(grid_required_vars),
    all(df_penetration_grid$land_share >= -NUMERIC_TOLERANCE &
          df_penetration_grid$land_share <= 1 + NUMERIC_TOLERANCE),
    all(grid_land_share_steps >= -NUMERIC_TOLERANCE),
    nrow(df_penetration_lookup) == 11,
    isTRUE(all.equal(df_penetration_lookup$penetration, (0:10) / 10)),
    all(lookup_p_internal_steps >= 0),
    all(lookup_land_share_steps >= -NUMERIC_TOLERANCE),
    lookup_first$p_internal == 0 &&
      abs(lookup_first$land_share_achieved) <= NUMERIC_TOLERANCE,
    lookup_last$p_internal == 1 &&
      abs(lookup_last$land_share_achieved - 1) <= LOOKUP_TOLERANCE_PP / 100,
    max_abs_error_pp <= LOOKUP_TOLERANCE_PP
  )
) |>
  mutate(status = if_else(passed %in% TRUE, "Pass", "Fail")) |>
  select(scope, check, value, status)

# Export for web resources
export(tbl_scale_validation, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_c3_scale_validation.rds")))

# Display as GT table
tbl_scale_validation |>
  gt() |>
  tab_header(
    title = "Biofortified Land Share Scale Validation",
    subtitle = "Grid integrity, monotonicity, and simulation level association"
  ) |>
  cols_label(
    scope = "Scope",
    check = "Check",
    value = "Value",
    status = "Status"
  ) |>
  cols_align(align = "left", columns = c(scope, check, value)) |>
  cols_align(align = "center", columns = status) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = status, rows = status == "Pass")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = status, rows = status == "Fail")
  ) |>
  tab_source_note(
    source_note = paste0(
      "Grid: ", format(nrow(df_penetration_grid), big.mark = ","),
      " values of the internal adoption index on [0, 1]. Numeric tolerance: ",
      format(NUMERIC_TOLERANCE, scientific = TRUE),
      ". Association tolerance: ", LOOKUP_TOLERANCE_PP,
      " pp. The script halts if any check fails."
    )
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Halt the pipeline if any check fails
n_failed_checks <- sum(tbl_scale_validation$status == "Fail")

if (n_failed_checks > 0) {
  stop(sprintf("Phase C halted: %d land share scale check(s) failed.",
               n_failed_checks))
}

# Clean up
rm(NUMERIC_TOLERANCE, LOOKUP_TOLERANCE_PP, grid_required_vars,
   grid_land_share_steps, lookup_p_internal_steps, lookup_land_share_steps,
   lookup_first, lookup_last, max_abs_error_pp, n_failed_checks,
   tbl_scale_validation, penetration_targets)
```

## C4. Land Share Verification Table

The table reports the eleven simulation levels with the internal adoption index at which each level is evaluated and the national biofortified maize land share achieved at that index. The difference column expresses the association error in percentage points. Biofortified and total maize land are reported in hectares (1 mz = 0.7 ha).

```{r}
#| label: tbl-c4-land-share-lookup
#| tbl-cap: "National biofortified maize land share by simulation level"

# =============================================================================
# C4. Land Share Verification Table
# =============================================================================

# Prepare table data
tbl_land_share_lookup <- df_penetration_lookup |>
  transmute(
    target_pct    = penetration * 100,
    p_internal,
    land_pct_achieved,
    error_pp,
    # Convert to hectares (1 mz = 0.7 ha)
    land_bio_ha   = land_bio * 0.7,
    land_total_ha = land_total * 0.7
  )

# Export for web resources
export(tbl_land_share_lookup, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_c4_land_share_lookup.rds")))

# Generate GT table
tbl_land_share_lookup |>
  gt() |>
  tab_header(
    title = "Biofortified Maize Land Share by Simulation Level",
    subtitle = paste0("Internal adoption index associated with each national land share target (",
                      seed_code, ", ", subsidy_label, ")")
  ) |>
  cols_label(
    target_pct = "% Biofortified Maize Land Target",
    p_internal = "Internal Adoption Index",
    land_pct_achieved = "% Biofortified Maize Land Achieved",
    error_pp = "Difference (pp)",
    land_bio_ha = "Biofortified Land (ha)",
    land_total_ha = "Total Maize Land (ha)"
  ) |>
  cols_align(align = "right", columns = everything()) |>
  fmt_number(
    columns = target_pct,
    decimals = 0,
    pattern = "{x}%"
  ) |>
  fmt_number(
    columns = p_internal,
    decimals = 4
  ) |>
  fmt_number(
    columns = land_pct_achieved,
    decimals = 2,
    pattern = "{x}%"
  ) |>
  fmt_number(
    columns = error_pp,
    decimals = 3,
    force_sign = TRUE
  ) |>
  fmt_number(
    columns = c(land_bio_ha, land_total_ha),
    use_seps = TRUE,
    decimals = 0
  ) |>
  tab_source_note(
    source_note = "Source: ENCOVI 2023 (MAGA-calibrated weights) with continuous adoption parameters from Script 02_05. Internal adoption index = position along the production-weighted adoption ranking (Script 02_06). Land share = Σ(land_wtcorn × r_i × weight_calibrated) / Σ(land_wtcorn × weight_calibrated). Difference = achieved − target, in percentage points."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    column_labels.font.size = px(10)
  )

# Clean up
rm(tbl_land_share_lookup)
```

## C5. Land Share Curve

The figure displays the national biofortified maize land share along the full internal adoption grid, with the eleven simulation levels overlaid at their associated internal adoption index. Guide lines connect each simulation level to both axes, so the correspondence between the internal adoption index and the land share it represents can be read directly.

```{r}
#| label: fig-c5-land-share-curve
#| fig-cap: "National biofortified maize land share by internal adoption index, with the eleven simulation levels"
#| fig-width: 10
#| fig-height: 8

# =============================================================================
# C5. Visualization - Biofortified Land Share Curve
# =============================================================================

# Curve data: full internal adoption grid
df_land_share_curve <- df_penetration_grid |>
  select(p_internal, land_share, land_pct)

# Simulation levels: selected points on the curve
df_land_share_points <- df_penetration_lookup |>
  transmute(
    penetration,
    p_internal,
    land_pct_achieved,
    point_label = paste0(round(penetration * 100), "%"),
    # Label placement: right of the point at the origin, left elsewhere
    label_hjust = if_else(near(penetration, 0), -0.4, 1.3)
  )

# Export for web resources
export(df_land_share_curve, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_fig_c5_land_share_curve.rds")))
export(df_land_share_points, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_fig_c5_land_share_points.rds")))

# Main visualization
ggplot() +
  # Vertical guide lines (internal adoption index of each level)
  geom_segment(
    data = df_land_share_points,
    aes(x = p_internal * 100, xend = p_internal * 100,
        y = 0, yend = land_pct_achieved),
    linetype = "dotted",
    color = "gray60",
    linewidth = 0.4
  ) +
  # Horizontal guide lines (land share achieved at each level)
  geom_segment(
    data = df_land_share_points,
    aes(x = 0, xend = p_internal * 100,
        y = land_pct_achieved, yend = land_pct_achieved),
    linetype = "dotted",
    color = "gray60",
    linewidth = 0.4
  ) +
  # Land share curve along the internal adoption grid
  geom_line(
    data = df_land_share_curve,
    aes(x = p_internal * 100, y = land_pct),
    color = "#2E86AB",
    linewidth = 1
  ) +
  # Simulation levels
  geom_point(
    data = df_land_share_points,
    aes(x = p_internal * 100, y = land_pct_achieved),
    color = "#08306B",
    size = 3,
    shape = 16
  ) +
  # Labels for simulation levels
  geom_text(
    data = df_land_share_points,
    aes(x = p_internal * 100, y = land_pct_achieved,
        label = point_label, hjust = label_hjust),
    vjust = -0.6,
    size = 3.3,
    color = "#08306B",
    fontface = "bold"
  ) +
  # Scales
  scale_x_continuous(
    breaks = seq(0, 100, 10),
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  scale_y_continuous(
    breaks = seq(0, 100, 10),
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  # Labels
  labs(
    title = "National Biofortified Maize Land Share by Internal Adoption Index",
    subtitle = paste0("Continuous adoption model — Points mark the eleven simulation levels (",
                      seed_code, ", ", subsidy_label, ")"),
    x = "Internal Adoption Index",
    y = "National Biofortified Maize Land Share"
  ) +
  # Theme
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "gray40", size = 12),
    axis.title = element_text(color = "gray40")
  )

# Clean up
rm(df_land_share_curve, df_land_share_points)
```

---

# Phase D: Redistribution Infrastructure & Validation

## D1. Saturation Analysis Before Redistribution

This section generates departmental coverage curves across the eleven simulation levels of national biofortified maize land share (0-100%) using the continuous adoption model.

At each simulation level, farmer-level adoption is evaluated at the associated internal adoption index `p_internal` from Phase C. Every farmer with `p_entry <= p_internal` is in the adoption pool, and the effective biofortified production of farmer `i` is `production_biofortified_i × weight_calibrated_i × r_i(p_internal)`, where `r_i(p_internal) = apply_adoption_curve(p_internal, p_entry_i, profile_i, ...)` is the profile-specific sigmoid conversion ratio. Departmental biofortified production is the sum of these effective contributions across all admitted farmers in that department. Each result row is labelled with the external `penetration` value of its simulation level.

This approach reflects realistic adoption patterns where early adopters are not uniformly distributed across departments, and where each farmer's conversion intensity varies by their behavioural profile and position along the adoption ranking.

```{r}
#| label: d1-saturation-data

# =============================================================================
# D1. Generate Saturation Data - Before Redistribution (Continuous Adoption)
# =============================================================================

# Prepare farmer data with pre-computed base biofortified production
# (the constant factor that is then scaled by r_i(p_internal) at each level)
df_farmers_prep <- df_farmers_raw |>
  mutate(
    bio_base = production_biofortified * weight_calibrated
  )

# Simulation levels: external label (penetration) and internal adoption
# index (p_internal) associated in Phase C
level_penetration <- df_penetration_lookup$penetration
level_p_internal  <- df_penetration_lookup$p_internal

# Calculate departmental coverage at each simulation level
df_saturation_before <- map2_dfr(level_penetration, level_p_internal,
                                 function(pen, p_int) {

  # Farmers in the adoption pool at this internal adoption index
  # (p_internal = 0 denotes the empty adoption pool)
  df_admitted <- df_farmers_prep |>
    filter(p_entry <= p_int, p_int > 0) |>
    mutate(
      r_i = apply_adoption_curve(p_int, p_entry, profile,
                                 floor_S, floor_M, floor_L,
                                 k_S, k_M, k_L),
      bio_weighted_p = bio_base * r_i
    )

  # Biofortified production by department
  df_bio_by_dept <- df_admitted |>
    summarise(
      bio_production_qq = sum(bio_weighted_p, na.rm = TRUE),
      n_adopters = n(),
      .by = departamento
    )

  # Coverage calculation
  df_market_baseline |>
    select(departamento, consumption_fsi_qq) |>
    mutate(departamento = as.character(departamento)) |>
    left_join(df_bio_by_dept, by = "departamento") |>
    replace_na(list(bio_production_qq = 0, n_adopters = 0)) |>
    mutate(
      penetration = pen,
      coverage_pct = bio_production_qq / consumption_fsi_qq * 100,
      is_saturated = coverage_pct >= 100
    ) |>
    select(departamento, penetration, bio_production_qq,
       consumption_fsi_qq, coverage_pct, n_adopters, is_saturated)
})
```

```{r}
#| label: fig-d1-saturation-curves
#| fig-cap: "Departmental coverage evolution by national biofortified maize land share (before redistribution)"
#| fig-width: 10
#| fig-height: 8
#| warning: false

# =============================================================================
# D1. Visualization - Saturation Curves Before Redistribution
# =============================================================================

# Saturation points for annotation
df_saturation_points <- df_saturation_before |>
  filter(is_saturated) |>
  slice_min(penetration, by = departamento)

# Color palette: highlight saturating departments
dept_colors <- df_saturation_before |>
  filter(penetration == 1) |>
  mutate(
    saturates = coverage_pct >= 100,
    color_group = if_else(saturates, departamento, "Other")
  ) |>
  select(departamento, color_group)

df_saturation_plot <- df_saturation_before |>
  left_join(dept_colors, by = "departamento") |>
  mutate(
    alpha_value = if_else(color_group == "Other", 0.4, 1),
    size_value = if_else(color_group == "Other", 0.5, 1)
  )

# Export for web resources
export(df_saturation_plot, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_fig_d1_saturation_curves.rds")))

# Main visualization
ggplot() +
  # Non-saturating departments (background)
  geom_line(
    data = filter(df_saturation_plot, color_group == "Other"),
    aes(x = penetration * 100, y = coverage_pct, group = departamento),
    color = "gray60",
    linewidth = 0.4,
    alpha = 0.5
  ) +
  # Saturating departments (highlighted)
  geom_line(
    data = filter(df_saturation_plot, color_group != "Other"),
    aes(x = penetration * 100, y = coverage_pct, color = departamento),
    linewidth = 1
  ) +
  # Saturation threshold
  geom_hline(
    yintercept = 100,
    linetype = "dashed",
    color = "darkred",
    linewidth = 0.6
  ) +
  # Saturation points
  geom_point(
    data = df_saturation_points,
    aes(x = penetration * 100, y = coverage_pct, color = departamento),
    size = 3,
    shape = 16
  ) +
  # Labels for saturation points (custom positioning)
  geom_text(
    data = filter(df_saturation_points, departamento == "Pet\u00e9n"),
    aes(x = penetration * 100, y = coverage_pct, label = departamento),
    hjust = 1.1,
    vjust = -0.5,
    size = 4.5,
    fontface = "bold"
  ) +
  # Annotation for threshold (moved to the left)
  annotate(
    "text",
    x = 1, y = 105,
    label = "Saturation threshold (100%)",
    hjust = 0,
    size = 3,
    color = "darkred",
    fontface = "italic"
  ) +
  # Scales
  scale_x_continuous(
    breaks = seq(0, 100, 10),
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  scale_y_continuous(
    breaks = seq(0, 300, 50),
    labels = function(x) paste0(x, "%"),
    limits = c(0, 320)
  ) +
  scale_color_manual(
    values = c("Pet\u00e9n" = "#2E86AB")
  ) +
  # Labels
  labs(
    title = "Departmental Coverage by National Biofortified Maize Land Share",
    subtitle = "Before excess redistribution \u2014 Gray lines show non-saturating departments",
    x = "National Biofortified Maize Land Share",
    y = "Departmental Coverage",
    color = "Saturating\nDepartments"
  ) +
  # Theme
  theme_minimal(base_size = 12) +
  theme(
    legend.position = c(0.12, 0.85),
    legend.background = element_rect(fill = "white", color = "gray80"),
    legend.title = element_text(size = 10, face = "bold"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "gray40", size = 12),
    axis.title = element_text(color = "gray40")
  )

# Clean up
rm(df_saturation_points, dept_colors, df_saturation_plot)
```

## D2. Apply Gravity-Based Redistribution

This section applies the gravity-based redistribution model to calculate departmental coverage at each simulation level. For each level, the biofortified production is calculated using the continuous adoption model evaluated at the associated internal adoption index (`r_i(p_internal)` per farmer), then surplus is redistributed through the gravity-weighted network to all deficit departments.

The gravity model distributes surplus from each donor department to all receivers, weighted by `w(i,j) × deficit_j`, where `w(i,j) = 1 / travel_time(i→j)²`.

```{r}
#| label: d2-gravity-redistribution

# =============================================================================
# D2. Apply Gravity-Based Redistribution Across Simulation Levels
#     (Continuous Adoption Model)
# =============================================================================

# Calculate departmental coverage WITH gravity redistribution at each level
df_saturation_redistributed <- map2_dfr(level_penetration, level_p_internal,
                                        function(pen, p_int) {

  # Farmers in the adoption pool at this internal adoption index
  # (p_internal = 0 denotes the empty adoption pool)
  df_admitted <- df_farmers_prep |>
    filter(p_entry <= p_int, p_int > 0) |>
    mutate(
      r_i = apply_adoption_curve(p_int, p_entry, profile,
                                 floor_S, floor_M, floor_L,
                                 k_S, k_M, k_L),
      bio_weighted_p = bio_base * r_i
    )

  # Biofortified production by department
  df_bio_by_dept <- df_admitted |>
    summarise(
      bio_production = sum(bio_weighted_p, na.rm = TRUE),
      .by = departamento
    )

  # Build coverage data frame
  df_coverage <- df_market_baseline |>
    select(departamento, annual_consumption = consumption_fsi_qq) |>
    mutate(departamento = as.character(departamento)) |>
    left_join(df_bio_by_dept, by = "departamento") |>
    replace_na(list(bio_production = 0))

  # Apply gravity-based redistribution
  df_redistributed <- redistribute_gravity(
    df = df_coverage,
    travel_time_df = df_travel_time
  )

  # Add external simulation level label and format output
  df_redistributed |>
    mutate(
      penetration = pen,
      coverage_pct = coverage * 100,
      is_saturated = coverage >= 1
    ) |>
    select(departamento, penetration, bio_production,
           annual_consumption, coverage_pct, excess, is_saturated)
})

# Cap coverage at 100% (excess beyond local demand = implicit export)
df_saturation_after <- df_saturation_redistributed |>
  mutate(
    bio_production = pmin(bio_production, annual_consumption),
    coverage_pct = pmin(coverage_pct, 100),
    is_saturated = coverage_pct >= 100
  )

# Clean up
rm(level_penetration, level_p_internal, df_saturation_redistributed)
```

## D3. Production Conservation Validation

Total biofortified production must be conserved through the redistribution process. The check compares total production before and after redistribution at 100% national biofortified maize land.

```{r}
#| label: tbl-d3-conservation-check
#| tbl-cap: "Production conservation validation at 100% biofortified maize land"

# =============================================================================
# D3. Validate Production Conservation
# =============================================================================

# Total production before redistribution (at 100% penetration)
total_before <- df_saturation_before |>
  filter(penetration == 1) |>
  summarise(total_qq = sum(bio_production_qq, na.rm = TRUE)) |>
  pull(total_qq)

# Total production after redistribution (at 100% penetration)
# Note: after capping, bio_production is capped at annual_consumption
# So we need uncapped data for this check. Recalculate at p=1 without cap.
df_check_uncapped <- df_saturation_after |>
  filter(penetration == 1)

# After redistribution + capping: production received by departments
total_after_capped <- sum(df_check_uncapped$bio_production, na.rm = TRUE)

# Remaining excess (production that exceeded local demand)
total_excess_remaining <- df_saturation_after |>
  filter(penetration == 1) |>
  summarise(total_excess = sum(excess, na.rm = TRUE)) |>
  pull(total_excess)

# Total accounted for = capped production + remaining excess
total_accounted <- total_after_capped + total_excess_remaining

# Difference
production_diff <- total_before - total_accounted
production_diff_pct <- production_diff / total_before * 100

# Summary table
tbl_conservation_check <- tibble(
  metric = c(
    "Total bio production before redistribution (qq)",
    "Total bio production after redistribution \u2014 capped (qq)",
    "Remaining excess after redistribution (qq)",
    "Total accounted (capped + excess) (qq)",
    "Difference (loss/gain) (qq)",
    "Difference as % of total production"
  ),
  value = c(
    format(round(total_before, 0), big.mark = ","),
    format(round(total_after_capped, 0), big.mark = ","),
    format(round(total_excess_remaining, 0), big.mark = ","),
    format(round(total_accounted, 0), big.mark = ","),
    format(round(production_diff, 2), big.mark = ","),
    paste0(round(production_diff_pct, 4), "%")
  )
)

# Export for web resources
export(tbl_conservation_check, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_d3_conservation_check.rds")))

tbl_conservation_check |>
  gt() |>
  tab_header(
    title = "Production Conservation Check",
    subtitle = "Validation at 100% national biofortified maize land \u2014 gravity redistribution"
  ) |>
  cols_label(metric = "Metric", value = "Value") |>
  cols_align(align = "left", columns = metric) |>
  cols_align(align = "right", columns = value) |>
  tab_source_note(
    source_note = "A non-zero difference indicates production loss or creation through redistribution. Values near zero (< 1 qq) are acceptable due to floating-point arithmetic."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Clean up
rm(df_check_uncapped, total_before, total_after_capped,
   total_excess_remaining, total_accounted, production_diff, production_diff_pct,
   tbl_conservation_check)
```

## D4. Saturation Order Summary

This table summarizes the national biofortified maize land share at which each department first reaches 100% coverage after applying the gravity-based redistribution. Departments are ordered by saturation timing.

Departments that do not reach saturation even at 100% national biofortified maize land are marked with "\u2014", indicating insufficient production flow through the redistribution network.

```{r}
#| label: tbl-d4-saturation-order
#| tbl-cap: "Departmental saturation order by national biofortified maize land share (after gravity redistribution)"

# =============================================================================
# D4. Saturation Order Table (After Redistribution)
# =============================================================================

# Identify first saturation point per department
df_saturation_order <- df_saturation_after |>
  filter(is_saturated) |>
  slice_min(penetration, by = departamento) |>
  select(departamento, saturation_penetration = penetration)

# Join with all departments to include non-saturating ones
tbl_saturation_order <- df_market_baseline |>
  select(departamento) |>
  mutate(departamento = as.character(departamento)) |>
  left_join(df_saturation_order, by = "departamento") |>
  arrange(saturation_penetration, departamento) |>
  mutate(
    rank = ifelse(!is.na(saturation_penetration), row_number(), NA_integer_),
    saturation_display = ifelse(
      !is.na(saturation_penetration),
      paste0(saturation_penetration * 100, "%"),
      "\u2014"
    )
  ) |>
  select(rank, departamento, saturation_display)

# Export for web resources
export(tbl_saturation_order, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_d4_saturation_order.rds")))

# Generate GT table
tbl_saturation_order |>
  gt() |>
  tab_header(
    title = "Departmental Saturation Order",
    subtitle = "National biofortified maize land share at which each department reaches 100% coverage (after gravity redistribution)"
  ) |>
  cols_label(
    rank = "Order",
    departamento = "Department",
    saturation_display = "Saturation Point"
  ) |>
  cols_align(align = "center", columns = c(rank, saturation_display)) |>
  cols_align(align = "left", columns = departamento) |>
  sub_missing(columns = rank, missing_text = "\u2014") |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(
      rows = saturation_display != "\u2014"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(
      rows = saturation_display == "\u2014"
    )
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = departamento)
  ) |>
  tab_source_note(
    source_note = "Source: ENCOVI 2023 farmer adoption model with gravity-based redistribution. w(i,j) = 1/travel_time(i\u2192j)\u00b2, \u03b2 = 2."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(11)
  )

# Clean up
rm(df_saturation_order, tbl_saturation_order)
```


## D5. Saturation Curves After Redistribution

This section visualizes departmental coverage evolution after applying the gravity-based redistribution. Compared to the pre-redistribution curves (D1), departments that previously never reached 100% coverage should now show improved trajectories as they receive surplus from producing departments, weighted by geographic proximity.

```{r}
#| label: fig-d5-saturation-after
#| fig-cap: "Departmental coverage evolution by national biofortified maize land share (after gravity redistribution)"
#| fig-width: 10
#| fig-height: 8

# =============================================================================
# D5. Visualization - Saturation Curves After Redistribution
# =============================================================================

# Identify saturation points (first penetration where coverage >= 100%)
df_saturation_points_after <- df_saturation_after |>
  filter(is_saturated) |>
  slice_min(penetration, by = departamento)

# Count saturating departments for subtitle
n_saturating <- n_distinct(df_saturation_points_after$departamento)

# Order saturating departments by saturation point
saturating_depts_ordered <- df_saturation_points_after |>
  arrange(penetration) |>
  pull(departamento)

# Order non-saturating departments by max coverage (highest first = closer to 100%)
non_saturating_depts_ordered <- df_saturation_after |>
  filter(penetration == 1, !departamento %in% saturating_depts_ordered) |>
  arrange(desc(coverage_pct)) |>
  pull(departamento)

# Create blue gradient palette for saturating (dark = earlier saturation)
blue_palette <- colorRampPalette(c("#08306B", "#2171B5", "#6BAED6", "#C6DBEF"))(n_saturating)
names(blue_palette) <- saturating_depts_ordered

# Create gray gradient palette for non-saturating (dark = closer to 100%)
n_non_saturating <- length(non_saturating_depts_ordered)
gray_palette <- colorRampPalette(c("gray40", "gray75"))(max(n_non_saturating, 1))
names(gray_palette) <- non_saturating_depts_ordered

# Combined palette
full_palette <- c(blue_palette, gray_palette)

# Create plot data
df_plot_after <- df_saturation_after |>
  mutate(
    is_saturating = departamento %in% saturating_depts_ordered
  )

# Export for web resources
export(df_plot_after, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_fig_d5_saturation_after.rds")))

# Main visualization
ggplot() +
  # Non-saturating departments (gray gradient)
  geom_line(
    data = filter(df_plot_after, !is_saturating),
    aes(x = penetration * 100, y = coverage_pct, color = departamento),
    linewidth = 0.5,
    alpha = 0.6
  ) +
  # Saturating departments (blue gradient)
  geom_line(
    data = filter(df_plot_after, is_saturating),
    aes(x = penetration * 100, y = coverage_pct, color = departamento),
    linewidth = 0.9
  ) +
  # Saturation threshold
  geom_hline(
    yintercept = 100,
    linetype = "dashed",
    color = "darkred",
    linewidth = 0.6
  ) +
  # Saturation points
  geom_point(
    data = df_saturation_points_after,
    aes(x = penetration * 100, y = coverage_pct, color = departamento),
    size = 2.5,
    shape = 16
  ) +
  # Annotation for threshold
  annotate(
    "text",
    x = 5, y = 105,
    label = "Saturation threshold (100%)",
    hjust = 0,
    size = 3.3,
    color = "darkred",
    fontface = "italic"
  ) +
  # Scales
  scale_x_continuous(
    breaks = seq(0, 100, 10),
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, 25),
    labels = function(x) paste0(x, "%")
  ) +
  scale_color_manual(
    values = full_palette,
    breaks = saturating_depts_ordered  # Only show saturating in legend
  ) +
  # Labels
  labs(
    title = "Departmental Coverage by National Biofortified Maize Land Share",
    subtitle = paste0("After gravity-based redistribution \u2014 ", n_saturating,
                      " departments reach 100% coverage"),
    x = "National Biofortified Maize Land Share",
    y = "Departmental Coverage",
    color = "Saturating Departments\n(dark = earlier saturation)"
  ) +
  # Legend in multiple rows at bottom
  guides(
    color = guide_legend(nrow = ceiling(n_saturating / 4), byrow = TRUE)
  ) +
  # Theme
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 10, color = "gray40"),
    legend.text = element_text(size = 9),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "gray40", size = 12),
    axis.title = element_text(color = "gray40")
  )

# Clean up
rm(df_saturation_points_after, saturating_depts_ordered, non_saturating_depts_ordered,
   blue_palette, gray_palette, full_palette, df_plot_after, n_saturating, n_non_saturating)
```

## D6. Market Metrics After Redistribution

This table presents the final departmental metrics at 100% national biofortified maize land after applying the gravity-based redistribution. It validates that total production is conserved through the redistribution process and shows the final coverage achieved by each department.

```{r}
#| label: tbl-d6-market-metrics-after
#| tbl-cap: "Departmental market metrics at 100% biofortified maize land (after gravity redistribution)"

# =============================================================================
# D6. Market Metrics After Redistribution
# =============================================================================

# Extract data at 100% penetration
df_metrics_after <- df_saturation_after |>
  filter(penetration == 1) |>
  select(departamento, bio_production, annual_consumption, coverage_pct) |>
  mutate(
    bio_production_mt = bio_production * 0.04536,
    annual_consumption_mt = annual_consumption * 0.04536
  ) |>
  arrange(factor(departamento, levels = levels(df_market_baseline$departamento)))

# Join with baseline for comparison
tbl_metrics_comparison <- df_market_baseline |>
  select(
    departamento,
    n_farmers_total,
    land_ha_total,
    max_bio_production_mt,
    max_coverage_before = max_coverage_pct
  ) |>
  mutate(departamento = as.character(departamento)) |>
  left_join(
    df_metrics_after |>
      select(departamento,
             bio_production_after_mt = bio_production_mt,
             coverage_after = coverage_pct),
    by = "departamento"
  ) |>
  mutate(
    coverage_change = coverage_after - max_coverage_before
  )

# Export for web resources
export(tbl_metrics_comparison, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_d6_market_metrics_after.rds")))

# Generate GT table
tbl_metrics_comparison |>
  gt() |>
  tab_header(
    title = "Departmental Market Metrics Comparison",
    subtitle = "Before vs. after gravity redistribution at 100% national biofortified maize land"
  ) |>
  cols_label(
    departamento = "Department",
    n_farmers_total = "Total Farmers",
    land_ha_total = "Land (ha)",
    max_bio_production_mt = "Local Prod. (MT)",
    max_coverage_before = "Coverage Before",
    bio_production_after_mt = "Final Prod. (MT)",
    coverage_after = "Coverage After",
    coverage_change = "Change (pp)"
  ) |>
  cols_align(align = "left", columns = departamento) |>
  cols_align(align = "right", columns = -departamento) |>
  fmt_number(
    columns = c(n_farmers_total),
    use_seps = TRUE,
    decimals = 0
  ) |>
  fmt_number(
    columns = c(land_ha_total, max_bio_production_mt, bio_production_after_mt),
    use_seps = TRUE,
    decimals = 1
  ) |>
  fmt_number(
    columns = c(max_coverage_before, coverage_after),
    decimals = 1,
    pattern = "{x}%"
  ) |>
  fmt_number(
    columns = coverage_change,
    decimals = 1,
    pattern = "{x}pp",
    force_sign = TRUE
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(
      columns = coverage_after,
      rows = coverage_after >= 100
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(
      columns = coverage_after,
      rows = coverage_after < 50
    )
  ) |>
  tab_style(
    style = cell_text(color = "#2E7D32", weight = "bold"),
    locations = cells_body(
      columns = coverage_change,
      rows = coverage_change > 0
    )
  ) |>
  grand_summary_rows(
    columns = c(n_farmers_total, land_ha_total, max_bio_production_mt,
                bio_production_after_mt),
    fns = list(Total = ~ sum(., na.rm = TRUE)),
    fmt = ~ fmt_number(., use_seps = TRUE, decimals = 0)
  ) |>
  tab_source_note(
  source_note = "Source: ENCOVI 2023 with gravity-based redistribution. Coverage calculated against FSI consumption (174 kg/capita/year, USDA 2019). Green = saturated (\u2265100%). Red = severe deficit (<50%)."
) |>
  tab_footnote(
    footnote = paste0("Local Prod. = production from local farmers only. Final Prod. = after gravity redistribution. ", seed_label, "."),
    locations = cells_column_labels(columns = c(max_bio_production_mt, bio_production_after_mt))
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(13),
    heading.subtitle.font.size = px(12),
    column_labels.font.size = px(9)
  )

# Clean up
rm(df_metrics_after, tbl_metrics_comparison)
```

---

# Phase E: Output Validation and Export

## E1. Output Compatibility Validation

The market baseline and the saturation scenarios form the data contract consumed by the downstream scenario scripts (06_02 and 06_03). The checks below confirm, before any file is written, that the saturation scenarios keep the expected column names, order, and types, that `penetration` contains exclusively the eleven simulation levels, and that all 22 departments are present in the market baseline and at every simulation level. Production conservation through redistribution is validated in D3. The script halts if any check fails.

```{r}
#| label: tbl-e1-output-compatibility
#| tbl-cap: "Output compatibility validation for downstream scenario scripts"

# =============================================================================
# E1. Validate Output Compatibility
# =============================================================================

# Expected contract of the saturation scenarios parquet
expected_columns <- c("departamento", "penetration", "bio_production_qq",
                      "annual_consumption_qq", "coverage_pct", "is_saturated")
expected_types   <- c("character", "numeric", "numeric",
                      "numeric", "numeric", "logical")
expected_levels  <- seq(0, 1, by = 0.10)
expected_n_depts <- length(dept_order)

# Market baseline in export format
df_market_baseline_export <- df_market_baseline |>
  mutate(departamento = as.character(departamento))

# Saturation scenarios in export format
df_saturation_export <- df_saturation_after |>
  mutate(departamento = as.character(departamento)) |>
  select(departamento, penetration, bio_production_qq = bio_production,
         annual_consumption_qq = annual_consumption, coverage_pct, is_saturated)

# Column types of the saturation scenarios
export_types <- map_chr(df_saturation_export, ~ class(.x)[1])

# Departments present at each simulation level
df_depts_per_level <- df_saturation_export |>
  summarise(
    n_depts = n_distinct(departamento),
    .by = penetration
  )

# Validation table
tbl_output_compatibility <- tibble(
  output = c(rep("Saturation scenarios", 5), "Market baseline"),
  check = c(
    "Column names and order match the downstream contract",
    "Column types match the downstream contract",
    "Simulation levels equal to 0, 0.1, ..., 1",
    "All departments present at every simulation level",
    "One row per department and simulation level",
    "All departments present, one row per department"
  ),
  value = c(
    paste(names(df_saturation_export), collapse = ", "),
    paste(export_types, collapse = ", "),
    paste(sort(unique(df_saturation_export$penetration)), collapse = ", "),
    paste0(min(df_depts_per_level$n_depts), "–",
           max(df_depts_per_level$n_depts), " departments per level"),
    paste0(format(nrow(df_saturation_export), big.mark = ","), " rows"),
    paste0(n_distinct(df_market_baseline_export$departamento), " departments, ",
           nrow(df_market_baseline_export), " rows")
  ),
  passed = c(
    identical(names(df_saturation_export), expected_columns),
    identical(unname(export_types), expected_types),
    identical(sort(unique(df_saturation_export$penetration)), expected_levels),
    nrow(df_depts_per_level) == length(expected_levels) &&
      all(df_depts_per_level$n_depts == expected_n_depts) &&
      setequal(df_saturation_export$departamento,
               df_market_baseline_export$departamento),
    nrow(df_saturation_export) == expected_n_depts * length(expected_levels) &&
      nrow(distinct(df_saturation_export, departamento, penetration)) ==
        nrow(df_saturation_export),
    n_distinct(df_market_baseline_export$departamento) == expected_n_depts &&
      nrow(df_market_baseline_export) == expected_n_depts
  )
) |>
  mutate(status = if_else(passed %in% TRUE, "Pass", "Fail")) |>
  select(output, check, value, status)

# Export for web resources
export(tbl_output_compatibility, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_e1_output_compatibility.rds")))

# Display as GT table
tbl_output_compatibility |>
  gt() |>
  tab_header(
    title = "Output Compatibility Check",
    subtitle = "Data contract of the market baseline and saturation scenarios"
  ) |>
  cols_label(
    output = "Output",
    check = "Check",
    value = "Value",
    status = "Status"
  ) |>
  cols_align(align = "left", columns = c(output, check, value)) |>
  cols_align(align = "center", columns = status) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = status, rows = status == "Pass")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = status, rows = status == "Fail")
  ) |>
  tab_source_note(
    source_note = "Contract consumed by Scripts 06_02 and 06_03. Production conservation through redistribution is validated in D3. The script halts if any check fails."
  ) |>
  tab_options(
    table.font.size = px(11),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Halt the pipeline if any check fails
n_failed_checks <- sum(tbl_output_compatibility$status == "Fail")

if (n_failed_checks > 0) {
  stop(sprintf("Phase E halted: %d output compatibility check(s) failed.",
               n_failed_checks))
}

# Clean up
rm(expected_columns, expected_types, expected_levels, expected_n_depts,
   export_types, df_depts_per_level, n_failed_checks, tbl_output_compatibility)
```

## E2. Export Market Baseline

This section exports the departmental market baseline metrics for use in downstream scenario analysis. The baseline contains static departmental characteristics: farmer counts, land area, consumption demand, and maximum biofortified production capacity.

```{r}
#| label: e2-export-baseline

# =============================================================================
# E2. Export Market Baseline
# =============================================================================

write_parquet(
  df_market_baseline_export,
  here("01_data", "02_processed",
       "scenarios", paste0(file_prefix, "_market_baseline.parquet"))
)
```

## E3. Export Saturation Scenarios

This section exports the pre-calculated departmental coverage at the eleven simulation levels of national biofortified maize land share (0-100%) with gravity-based redistribution already applied, so the redistribution is not recalculated during scenario simulation.

Each row represents a department × simulation level combination, labelled with the external `penetration` value, with the final biofortified production available after network redistribution.

```{r}
#| label: e3-export-scenarios

# =============================================================================
# E3. Export Saturation Scenarios
# =============================================================================

write_parquet(
  df_saturation_export,
  here("01_data", "02_processed",
       "scenarios", paste0(file_prefix, "_saturation_scenarios.parquet"))
)
```

## E4. Export Land Share Lookup

This section exports the association between each simulation level and its internal adoption index. Each row contains the external `penetration` value, the internal adoption index `p_internal` at which farmer-level adoption is evaluated, the biofortified and total maize land (manzanas), and the land share achieved together with its association error. Downstream scenario scripts read `p_internal` from this table to evaluate farmer-level adoption consistently with the departmental coverage stored in the saturation scenarios.

```{r}
#| label: e4-export-lookup

# =============================================================================
# E4. Export Land Share Lookup
# =============================================================================

write_parquet(
  df_penetration_lookup,
  here("01_data", "02_processed",
       "scenarios", paste0(file_prefix, "_penetration_lookup.parquet"))
)
```

# Notes

This script corresponds to the **`r paste0("Seed ", seed_year, " (", seed_code, ")")`** seed variant under the **`r subsidy_label`** subsidy regime. The yield change factors applied to farmers in the upstream economic impact calculation (Script `r paste0("02_04_", seed_suffix, "_", subsidy_suffix)`) determine the `production_biofortified` values used here, and the Likely Adopters score (Script `r paste0("02_06_", seed_suffix, "_", subsidy_suffix)`) determines the order in which farmers enter the adoption pool along the internal adoption index. The association between the eleven simulation levels of national biofortified maize land share and the internal adoption index is specific to this seed-by-subsidy combination.

The continuous adoption model parameters (per-profile floors and steepness coefficients) are loaded from the calibrated model configuration produced by Script 02_05. These parameters are invariant across the nine seed-by-subsidy combinations: the calibration is performed once on the s22 + sub00 baseline configuration and applied identically to all combinations.

The outputs differ in their sensitivity to the subsidy regime: the structural baseline (`market_baseline.parquet`) is invariant across regimes for a given seed — consumption demand, farmer counts, land area, and maximum biofortified production capacity depend only on physical and demographic quantities. The saturation scenarios (`saturation_scenarios.parquet`) and the land share lookup (`penetration_lookup.parquet`) are subsidy-sensitive at intermediate simulation levels because the adoption-order ranking is driven by an economic score that responds to the subsidy regime; both endpoints (0% and 100% biofortified maize land) converge across regimes by construction, since at 0% the adoption pool is empty and at 100% every farmer has `r_i(1) = 1`.

## Methodological Assumptions

The gravity-based redistribution model incorporates the following assumptions:

1. **Calibrated consumption**: Departmental maize consumption includes two estimates calculated from survey-weighted population:
   - **Direct human consumption**: 110 kg/capita/year (FAO/MAGA 2008)
   - **FSI consumption** (Food, Seed, Industrial): 174 kg/capita/year (USDA 2019)

   All coverage calculations use the **FSI estimate** as the denominator, since biofortified production must satisfy total domestic demand, not just direct human consumption.

2. **Continuous adoption model**: Each farmer's effective biofortified production at internal adoption index `p_internal` is `production_biofortified × r_i(p_internal)`, where `r_i(p_internal)` is a profile-specific sigmoid conversion ratio. The conversion ratio equals the profile floor at entry (`p_internal = p_entry`) and converges to 1 at full adoption (`p_internal = 1`) by construction. The three behavioural profiles (S, M, L) capture distinct adoption intensities: near-binary for small farmers, gradual for large farmers.

3. **Biofortified land share scale**: Simulation levels are defined as national shares of maize land cultivated with biofortified seed, from 0% to 100% in 10-percentage-point increments. Each level is evaluated at the internal adoption index whose land share is nearest to the target on a grid of 10,001 values, with an association error within 0.25 percentage points. The 0% level corresponds to the empty adoption pool and the 100% level to full adoption.

4. **Gravity-weighted redistribution**: Surplus flows from each donor to all deficit departments, weighted by `w(i,j) × deficit_j / Σ_k(w(i,k) × deficit_k)`, where `w(i,j) = 1 / travel_time(i→j)²`. The travel time matrix is asymmetric by design, reflecting real directional differences in Guatemala's road infrastructure.

5. **Iterative convergence**: Redistribution proceeds in passes until total redistributable excess falls below 1 quintal. Departments that receive more than their deficit in one pass become donors in the next.

6. **No transport costs**: The gravity model determines flow patterns but does not deduct transport costs from production volumes.

7. **No intermediaries**: Direct department-to-department flow without wholesale market hubs.

8. **Capped coverage**: Departmental coverage is capped at 100%; excess beyond local demand represents implicit export outside the modeled system.

## References

Anderson, J. E., & van Wincoop, E. (2003). Gravity with Gravitas: A Solution to the Border Puzzle. *American Economic Review*, 93(1), 170–192. https://doi.org/10.1257/000282803321455214
