---
title: "Precompute National Aggregates (`r params$seed_label`, `r params$subsidy_label`, `r params$pathway_label` Pathway)"
subtitle: "`r if (params$pathway == 'pco') 'National aggregation of biofortification impact across penetration levels using pre-computed nutritional profiles (GAM-based stunting model)' else 'National aggregation of biofortification impact across penetration levels using Gunaratna et al. (2010) meta-analysis effect size'`"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
params:
  # Seed identification
  seed_year: "2022"
  seed_label: "2022 Seed"
  seed_suffix: "s22"
  # Subsidy identification
  subsidy_label: "Current subsidy"
  subsidy_suffix: "sub00"
  # Pathway identification
  pathway: "pco"
  pathway_label: "Covariate"
---

```{r}
#| label: setup

# =============================================================================
# Setup: Load packages, configure environment, resolve parameters
# =============================================================================

# Clean environment (preserve params injected by Quarto)
rm(list = setdiff(ls(), "params"))
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  survey,             # Complex survey methodology
  srvyr,              # Tidy wrapper for survey package
  gt,                 # Professional table formatting
  glue,               # String formatting
  tidyverse           # General purpose data manipulation
)

# Packages used via :: (not loaded to keep namespace clean):
#   skimr   - skim_without_charts() in Phase F data quality validation

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter
drop   <- base::drop

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions from centralized file
source(here("02_code", "00_functions", "00_functions.R"))

# --- Resolve parameters into local variables -------------------------------

# Seed
seed_year     <- params$seed_year
seed_label    <- params$seed_label
seed_suffix   <- params$seed_suffix

# Subsidy
subsidy_label  <- params$subsidy_label
subsidy_suffix <- params$subsidy_suffix

# Pathway
pathway       <- params$pathway
pathway_label <- params$pathway_label

# Seed code for plot/table titles (e.g., "S22"). Derived from seed_suffix
# to keep seed_variants.yml unchanged: no additional field required.
seed_code <- toupper(seed_suffix)

# Composite file prefix for all output artefacts (parquet + web_resources rds)
file_prefix <- paste0("06_03_", seed_suffix, "_", subsidy_suffix, "_", pathway)

# Pathway-specific settings
age_var    <- if (pathway == "pco") "edad_mes_nino" else "edad_meses"
is_pco     <- pathway == "pco"

# scenarios/
ensure_output_dir(here("01_data", "02_processed", "scenarios"))

# web_resources/
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

# Overview

> **Seed `r seed_code` · `r subsidy_label` · `r pathway_label` Pathway** — This script corresponds to the Seed `r params$seed_year` (`r seed_code`) scenario under the **`r subsidy_label`** subsidy regime and the `r tolower(pathway_label)` pathway. The methodology is identical across all 18 combinations (3 seeds x 3 subsidy regimes x 2 pathways); differences arise from the upstream yield factors (seed) and the upstream seed-price structure (subsidy regime) that together shape the farmer adoption ranking and economic metrics.

This script calculates national-level impact metrics across 11 biofortification penetration scenarios (0% to 100% in 10% increments). It uses the same methodology as Script `r paste0("06_02_", seed_suffix, "_", subsidy_suffix, "_", pathway)` and aggregates directly to national level from individual-level data. It combines farmer adoption data with the `r if (is_pco) "synthetic child population" else "ENCOVI child population"` to estimate nutrient and stunting outcomes under different market penetration levels.

Penetration is implemented through the continuous adoption model: each farmer has a per-farmer entry point (`p_entry`) and a behavioural profile (`S`, `M`, `L`), which together determine the conversion ratio $r_i(p)$ at each national penetration level $p$. At the farmer's entry point, the conversion ratio equals the profile-specific floor; it then rises along a sigmoid curve towards 1 as $p$ increases. At full penetration ($p = 1$), all farmers have $r_i(1) = 1$ by construction. Economically scalable quantities (production, land, income, seed costs, subsidy flows) are weighted by $r_i(p)$, while farmer counts reflect pool membership (a farmer either has entered the pool or has not).

## Main Operations

1. Load calibrated model configuration for continuous adoption curves (02_05 output)
2. Define helper functions for national scenario simulation pipeline
3. Load `r if (is_pco) "synthetic population with pre-calculated baseline and QPM nutritional/stunting profiles (from 04_08)" else "ENCOVI children with pre-calculated baseline and QPM nutritional/stunting profiles (from 05_06)"`, farmer data, market baseline, and saturation scenarios
4. Calculate baseline metrics at 0% penetration (no biofortification)
5. For each penetration level (10% to 100%):
   - Filter adopting farmers using `p_entry <= p` and compute per-farmer conversion ratio $r_i(p)$
   - Calculate economic metrics at national level, weighting scalable quantities by $r_i(p)$
   - Calculate national coverage from departmental saturation scenarios
   - Assign biofortification flag to children using hierarchical priority
   - Select pre-calculated baseline or QPM nutritional values based on flag
   - Aggregate nutritional and stunting metrics at national level
6. Combine all scenarios and calculate change metrics relative to baseline
7. Export national scenario data for Shiny dashboard

## Methodological Notes

**Direct National Aggregation**: National metrics are calculated directly from individual-level data using survey weights, so the weighted means are computed over the full individual distribution.

**Pre-Calculated Nutrient and Stunting Profiles**: Baseline and QPM nutrient profiles (intake, bioavailability, inadequacy, nutritional index) and GAM-based stunting outcomes are pre-computed in `r if (is_pco) "Script 04_08" else "Script 05_06"`. The simulation loop selects between `*_baseline` and `*_qpm` columns based on biofortification assignment, enabling efficient scenario computation without on-the-fly recalculation.

```{r}
#| echo: false
#| output: asis

if (is_pco) {
  cat("**Pre-computability of GAM Impact**: The GAM-based stunting impact is pre-computed because every quantity the delta ZLEN requires is fixed per-child:

- **Baseline and QPM indices** — `nutritional_index_bio_baseline` and `nutritional_index_bio_qpm` depend only on the child's maize intake and the biofortification factors, which are constant.
- **Fitted response** — the GAM is evaluated at each index, holding the child's age and sex at their observed values, all of which are constant.

Therefore, `delta_zlen_qpm`, the difference between the two fitted values, is a fixed per-child parameter.")
} else {
  cat("**Stunting Impact from Meta-Analysis Effect Size**: Stunting outcomes (`hfa`, `hfa_qpm`, `haz_calibrated`, `haz_qpm`) are derived from Script 05_05, which applies the Gunaratna et al. (2010) systematic review effect size (9% increase in height growth velocity for QPM consumers).")
}
```

**Continuous Adoption Model**: At each national penetration level $p$, every farmer with $p_{\text{entry}} \leq p$ is in the adoption pool. The effective contribution of farmer $i$ to any scalable economic quantity is `quantity_i * weight_calibrated_i * r_i(p)`, where $r_i(p)$ is the profile-specific sigmoid conversion ratio from `apply_adoption_curve()`. Farmer counts (number of adopters) are not scaled by $r_i$: a farmer either has entered the pool or has not.

**Subsidy Variables**: The seven monetary subsidy columns (`cost_seeds_bio_no_subsidy`, `cost_seeds_bio_current_regime`, `subsidy_amount_absolute`, `newseed_subsidy`, `producer_absorption`, `subsidy_reduction_cost`, `newseed_savings_vs_current`) are computed at farmer level in Script 02_04 and propagated through Script 02_06. This script aggregates them at national level using the same `r_i(p)` weighting applied to all scalable economic quantities.

**Hierarchical Biofortification Assignment**: Children are assigned biofortified maize consumption following a priority hierarchy that reflects realistic distribution patterns. The assignment follows a hierarchy based on each household's likelihood of accessing locally-produced biofortified maize, prioritizing in descending order: cereal-producing farm households, other farming households, households of agricultural workers, rural non-agricultural households, and urban households. Within each tier, assignment is randomized until the departmental coverage target is reached.

**Data-linkage limitation:** ENCOVI (farmer survey) and SIVESNU (child anthropometry) cannot be linked at the household level as they share no common identifier. Children of adopting farmers cannot therefore be identified deterministically. The hierarchy uses observable proxies (head-of-household occupation, rural/urban residence) to approximate the probability of biofortified maize access under realistic distribution patterns.

**Pre-calculated Coverage**: Departmental coverage at each penetration level is retrieved from Script `r paste0("06_01_", seed_suffix, "_", subsidy_suffix)`, which already applied the gravity-based redistribution model. National coverage is calculated as a consumption-weighted average of departmental coverages.

# Phase A: Function Definitions

This phase defines the helper functions required for national scenario simulation. Functions A2, A4, and A5 are shared with Script `r paste0("06_02_", seed_suffix, "_", subsidy_suffix, "_", pathway)`. Functions A3 and A6 aggregate at national level.

## A1. Load Model Configuration

Load the calibrated continuous-adoption model configuration from Script 02_05. Extract the six per-profile parameters (floors and steepness coefficients) consumed by `apply_adoption_curve()` in the simulation loop.

```{r}
#| label: a1-load-model-config

# =============================================================================
# A1. Load Model Configuration for Continuous Adoption Curves
# =============================================================================

model_config <- readRDS(
  here("01_data", "02_processed", "models",
       "02_05_likely_adopters_model_config.rds")
)

# Extract continuous adoption parameters
floor_S <- model_config$continuous$floor_S
floor_M <- model_config$continuous$floor_M
floor_L <- model_config$continuous$floor_L
k_S     <- model_config$continuous$k_S
k_M     <- model_config$continuous$k_M
k_L     <- model_config$continuous$k_L
```

```{r}
#| label: tbl-a1-model-config
#| tbl-cap: "Continuous adoption model parameters"

# =============================================================================
# Table: Model Configuration Summary
# =============================================================================

tibble(
  Parameter = c("floor_S", "floor_M", "floor_L", "k_S", "k_M", "k_L"),
  Value     = c(floor_S, floor_M, floor_L, k_S, k_M, k_L),
  Profile   = c("Small", "Medium", "Large", "Small", "Medium", "Large"),
  Role      = c(rep("Floor (conversion ratio at entry)", 3),
                rep("Steepness (sigmoid shape)", 3))
) |>
  gt() |>
  tab_header(
    title = "Continuous Adoption Model Parameters",
    subtitle = glue("Source: 02_05_likely_adopters_model_config.rds")
  ) |>
  fmt_number(columns = Value, decimals = 2) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )
```

## A2. Filter Adopters by Penetration Level

At each national penetration level $p$, farmers with $p_{\text{entry}} \leq p$ are admitted to the adoption pool. For each admitted farmer, the per-farmer conversion ratio $r_i(p)$ is computed via `apply_adoption_curve()`.

**Parameters:**

- `df_farmers`: Data frame of ranked farmers with `p_entry` and `profile`
- `penetration`: National penetration level (0-1)

**Returns:** Data frame of admitted farmers with `r_i` column appended.

```{r}
#| label: a2-filter-adopters

# =============================================================================
# A2. Filter Adopters by Penetration Level (Continuous Adoption Model)
# =============================================================================

filter_adopters_by_penetration <- function(df_farmers, penetration) {

  df_adopters <- df_farmers |>
    filter(p_entry <= penetration) |>
    mutate(
      r_i = apply_adoption_curve(penetration, p_entry, profile,
                                 floor_S, floor_M, floor_L,
                                 k_S, k_M, k_L)
    )

  return(df_adopters)
}
```

## A3. Calculate Economic Metrics at National Level

Aggregate economic impact metrics for adopting farmers at national level. Scalable quantities (land, production, income, seed costs, subsidy flows) are weighted by `r_i(p)` and `weight_calibrated`. Farmer counts reflect pool membership and are weighted by `weight_calibrated` only.

**Parameters:**

- `df_adopters`: Data frame of farmers adopting at current penetration level, with `r_i` column

**Returns:** Data frame with one row containing national economic metrics:

- `n_farmers_bio`: Total adopting farmers (weighted)
- `n_hybrid_bio`: Hybrid segment adopters (weighted)
- `n_opv_bio`: OPV/Criollo segment adopters (weighted)
- `land_bio_mz`: Total biofortified land in manzanas (weighted, scaled by `r_i`)
- `income_increment_total`: Total additional income in GTQ (weighted, scaled by `r_i`)
- `income_increment_median`: Median effective income increment per farmer
- `production_increment_total`: Total additional production in qq (weighted, scaled by `r_i`)
- `cost_seeds_increment_total`: Total seed cost difference in GTQ (weighted, scaled by `r_i`)
- `cost_seeds_bio_no_subsidy_total`: Total unsubsidized biofortified seed cost (GTQ, weighted, scaled by `r_i`)
- `cost_seeds_bio_current_regime_total`: Total farmer-paid biofortified seed cost under active regime (GTQ, weighted, scaled by `r_i`)
- `subsidy_amount_absolute_total`: Total monetary value of the subsidy (GTQ, weighted, scaled by `r_i`)
- `newseed_subsidy_total`: Portion of the total subsidy paid by NewSeed (GTQ, weighted, scaled by `r_i`)
- `producer_absorption_total`: Portion of the total subsidy absorbed by the seed producer (GTQ, weighted, scaled by `r_i`)
- `subsidy_reduction_cost_total`: Extra cost transferred to farmers under subsidy reduction (GTQ, weighted, scaled by `r_i`)
- `newseed_savings_vs_current_total`: NewSeed's savings versus the current subsidy regime (GTQ, weighted, scaled by `r_i`)

```{r}
#| label: a3-calculate-economic-national

# =============================================================================
# A3. Calculate Economic Metrics at National Level (Continuous Adoption Model)
# =============================================================================

calculate_economic_national <- function(df_adopters) {

  # Conversion factor: manzanas to hectares
  FACTOR_MZ_TO_HA <- 0.7

  df_economic <- df_adopters |>
    summarise(

      # -----------------------------------------------------------------------
      # Farmer counts by segment (survey-weighted, NOT scaled by r_i)
      # A farmer either has entered the adoption pool or has not.
      # -----------------------------------------------------------------------

      n_farmers_bio = sum(weight_calibrated, na.rm = TRUE),

      n_hybrid_bio = sum(
        if_else(segment_final != "OPV/Criollo", weight_calibrated, 0),
        na.rm = TRUE
      ),

      n_opv_bio = sum(
        if_else(segment_final == "OPV/Criollo", weight_calibrated, 0),
        na.rm = TRUE
      ),

      # -----------------------------------------------------------------------
      # Land area in manzanas (survey-weighted, scaled by r_i)
      # At penetration p, farmer i converts land_wtcorn * r_i(p) manzanas.
      # -----------------------------------------------------------------------

      land_bio_mz = sum(land_wtcorn * r_i * weight_calibrated, na.rm = TRUE),
      land_bio_ha = sum(land_wtcorn * r_i * weight_calibrated, na.rm = TRUE) * FACTOR_MZ_TO_HA,

      # -----------------------------------------------------------------------
      # Income metrics (scaled by r_i)
      # Total is weighted (population estimate), median is the typical
      # effective per-farmer increment at this penetration level.
      # -----------------------------------------------------------------------

      income_increment_total = sum(
        income_increment * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      income_increment_median = median(income_increment * r_i, na.rm = TRUE),

      # -----------------------------------------------------------------------
      # Production increment in quintales (survey-weighted, scaled by r_i)
      # -----------------------------------------------------------------------

      production_increment_total = sum(
        production_increment * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      production_biofortified_total = sum(
        production_biofortified * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      # -----------------------------------------------------------------------
      # Seed cost increment in GTQ (survey-weighted, scaled by r_i)
      # Represents additional cost farmers pay for biofortified seed,
      # proportional to the fraction of land converted.
      # -----------------------------------------------------------------------

      cost_seeds_increment_total = sum(
        cost_seeds_increment * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      # -----------------------------------------------------------------------
      # Subsidy block (survey-weighted, GTQ, scaled by r_i)
      #
      # All subsidy variables are computed at farmer level in Script 02_04
      # and propagated through Script 02_06. At penetration p, the
      # effective monetary flow for farmer i is quantity_i * r_i(p),
      # reflecting the fraction of land under biofortified seed.
      #
      # Linear identities preserved at aggregate level:
      #   cost_seeds_bio_no_subsidy_total =
      #     cost_seeds_bio_current_regime_total + subsidy_amount_absolute_total
      #     + subsidy_reduction_cost_total
      #   subsidy_amount_absolute_total =
      #     newseed_subsidy_total + producer_absorption_total
      #   newseed_savings_vs_current_total =
      #     producer_absorption_total + subsidy_reduction_cost_total
      # -----------------------------------------------------------------------

      cost_seeds_bio_no_subsidy_total = sum(
        cost_seeds_bio_no_subsidy * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      cost_seeds_bio_current_regime_total = sum(
        cost_seeds_bio_current_regime * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      subsidy_amount_absolute_total = sum(
        subsidy_amount_absolute * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      newseed_subsidy_total = sum(
        newseed_subsidy * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      producer_absorption_total = sum(
        producer_absorption * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      subsidy_reduction_cost_total = sum(
        subsidy_reduction_cost * r_i * weight_calibrated,
        na.rm = TRUE
      ),

      newseed_savings_vs_current_total = sum(
        newseed_savings_vs_current * r_i * weight_calibrated,
        na.rm = TRUE
      )
    )

  return(df_economic)
}
```

## A4. Assign Biofortification Using Hierarchical Priority

Assign biofortified maize consumption to children following a hierarchical priority system that reflects realistic distribution patterns. Within each department, children are prioritized according to their household's likelihood of accessing locally-produced biofortified maize, and assignment within each tier is randomized at the household level until the departmental coverage target is reached.

The function respects survey weights (`weight_calibrated`), ensuring that the weighted proportion of children receiving biofortified maize matches the departmental coverage target.

**Parameters:**

- `df_children`: Children population data frame with pre-calculated baseline and QPM profiles
- `df_coverage`: Data frame with `departamento` and `coverage_pct` (0-100 scale)
- `seed`: Random seed for reproducibility

**Returns:** Data frame with added `flag_biofortified` column (1 = receives biofortified, 0 = does not).

```{r}
#| label: a4-assign-biofortification

# =============================================================================
# A4. Assign Biofortification Using Hierarchical Priority
# =============================================================================

if (is_pco) {
  assign_biofortification_hierarchical <- function(df_children,
                                                    df_coverage,
                                                    seed = 789) {

    set.seed(seed)

    # ===========================================================================
    # STEP 1: Assign priority groups and random tie-breaker
    # ===========================================================================

    df_with_priority <- df_children |>
      mutate(
        priority_group = case_when(
          ocupacion_hogar == "Granjero cereales"    ~ 1L,
          ocupacion_hogar == "Granjero otros"       ~ 2L,
          ocupacion_hogar == "Trabajador del campo" ~ 3L,
          ruralurbano == "Rural"                    ~ 4L,
          ruralurbano == "Urbano"                   ~ 5L,
          TRUE                                      ~ 5L
        ),
        random_tie = runif(n())
      )

    # ===========================================================================
    # STEP 2: Join coverage data
    # ===========================================================================

    df_with_coverage <- df_with_priority |>
      left_join(
        df_coverage |>
          select(departamento, coverage_pct) |>
          mutate(coverage_frac = coverage_pct / 100),
        by = "departamento"
      )

    # ===========================================================================
    # STEP 3: Sort by department, priority, and random tie-breaker
    # ===========================================================================

    df_sorted <- df_with_coverage |>
      arrange(departamento, priority_group, random_tie)

    # ===========================================================================
    # STEP 4: Calculate cumulative weights and assign biofortification flag
    # Children are assigned until departmental coverage target is reached
    # ===========================================================================

    df_assigned <- df_sorted |>
      mutate(
        weight_cumsum = cumsum(weight_calibrated),
        weight_total = sum(weight_calibrated),
        weight_target = coverage_frac * weight_total,
        .by = departamento
      ) |>
      mutate(
        flag_biofortified = if_else(weight_cumsum <= weight_target, 1L, 0L)
      )

    # ===========================================================================
    # STEP 5: Clean up auxiliary columns
    # ===========================================================================

    df_result <- df_assigned |>
      select(
        -priority_group, -random_tie, -weight_cumsum, -weight_total,
        -weight_target, -coverage_pct, -coverage_frac
      )

    return(df_result)
  }
} else {
  assign_biofortification_hierarchical <- function(df_children,
                                                    df_coverage,
                                                    seed = 789) {

    # ===========================================================================
    # STEP 1: Assign household-level random tie-breaker
    # All children in the same household receive the same random value,
    # ensuring consistent within-household assignment.
    # ===========================================================================

    set.seed(seed)

    df_with_tie <- df_children |>
      mutate(random_tie = runif(1), .by = no_hogar)

    # ===========================================================================
    # STEP 2: Join coverage data
    # ===========================================================================

    df_with_coverage <- df_with_tie |>
      left_join(
        df_coverage |>
          select(departamento, coverage_pct) |>
          mutate(coverage_frac = coverage_pct / 100),
        by = "departamento"
      )

    # ===========================================================================
    # STEP 3: Sort by department, priority level, and household random tie
    # ===========================================================================

    df_sorted <- df_with_coverage |>
      arrange(departamento, priority_level, random_tie)

    # ===========================================================================
    # STEP 4: Calculate cumulative weights and assign biofortification flag
    # Children are assigned until departmental coverage target is reached
    # ===========================================================================

    df_assigned <- df_sorted |>
      mutate(
        weight_cumsum = cumsum(weight_calibrated),
        weight_total = sum(weight_calibrated),
        weight_target = coverage_frac * weight_total,
        .by = departamento
      ) |>
      mutate(
        flag_biofortified = if_else(weight_cumsum <= weight_target, 1L, 0L)
      )

    # ===========================================================================
    # STEP 5: Clean up auxiliary columns
    # ===========================================================================

    df_result <- df_assigned |>
      select(
        -random_tie, -weight_cumsum, -weight_total,
        -weight_target, -coverage_pct, -coverage_frac
      )

    return(df_result)
  }
}
```

## A5. Select Pre-Calculated Nutritional Values by Biofortification Status

For each child, select the appropriate pre-calculated nutrient profile based on biofortification assignment. Children with `flag_biofortified == 1` receive QPM values; children with `flag_biofortified == 0` retain baseline values.

All nutrient intake, bioavailability, inadequacy, and stunting outcomes are pre-computed in `r if (is_pco) "Script 04_08" else "Script 05_06"`. This function selects the appropriate profile for each child based on biofortification assignment.

**Mapped Variables (baseline -> QPM):**

| Output Variable | Baseline Source | QPM Source |
|----------------|-----------------|------------|
| `fe_mg_total` | `fe_mg_total_baseline` | `fe_mg_total_qpm` |
| `zn_mg_total` | `zn_mg_total_baseline` | `zn_mg_total_qpm` |
| `prot_g_total` | `prot_g_total_baseline` | `prot_g_total_qpm` |
| `fe_absorbed_mg` | `fe_absorbed_mg_baseline` | `fe_absorbed_mg_qpm` |
| `taz_mg` | `taz_mg_baseline` | `taz_mg_qpm` |
| `hfa_calculated` | `r if (is_pco) "hfa_calculated" else "hfa"` | `r if (is_pco) "hfa_calculated_qpm" else "hfa_qpm"` |
| `zlen_predicted` | `r if (is_pco) "zlen_predicted" else "haz_calibrated"` | `r if (is_pco) "zlen_predicted_qpm" else "haz_qpm"` |

**Parameters:**

- `df_children`: Data frame with `flag_biofortified` and pre-calculated `*_baseline`/`*_qpm` columns

**Returns:** Data frame with scenario-specific nutrient variables selected.

```{r}
#| label: a5-select-nutritional-values

# =============================================================================
# A5. Select Pre-Calculated Nutritional Values by Biofortification Status
# =============================================================================

if (is_pco) {
  select_nutritional_by_flag <- function(df_children) {

    df_selected <- df_children |>
      mutate(
        # =======================================================================
        # TOTAL INTAKE (before absorption)
        # =======================================================================

        fe_mg_total = if_else(
          flag_biofortified == 1L, fe_mg_total_qpm, fe_mg_total_baseline
        ),
        zn_mg_total = if_else(
          flag_biofortified == 1L, zn_mg_total_qpm, zn_mg_total_baseline
        ),
        prot_g_total = if_else(
          flag_biofortified == 1L, prot_g_total_qpm, prot_g_total_baseline
        ),

        # =======================================================================
        # ABSORBED INTAKE
        # =======================================================================

        fe_absorbed_mg = if_else(
          flag_biofortified == 1L, fe_absorbed_mg_qpm, fe_absorbed_mg_baseline
        ),
        taz_mg = if_else(
          flag_biofortified == 1L, taz_mg_qpm, taz_mg_baseline
        ),

        # =======================================================================
        # MAIZE-SOURCE ENERGY
        # Energy content of maize is unchanged by biofortification (factor = 1.0),
        # so ene_kcal_maiz remains the same regardless of assignment.
        # =======================================================================

        ene_kcal_maiz = ene_kcal_maiz,

        # =======================================================================
        # TOTAL ENERGY
        # Energy is unchanged by biofortification (factor = 1.0 for energy),
        # so total energy intake is the same as baseline for all children.
        # =======================================================================

        ene_kcal_total = ene_kcal_maiz + ene_kcal_nomaiz,

        # =======================================================================
        # STUNTING OUTCOMES (pre-computed in 04_08 via GAM)
        # =======================================================================

        hfa_calculated = if_else(
          flag_biofortified == 1L, hfa_calculated_qpm, hfa_calculated
        ),
        zlen_predicted = if_else(
          flag_biofortified == 1L, zlen_predicted_qpm, zlen_predicted
        )
      )

    return(df_selected)
  }
} else {
  select_nutritional_by_flag <- function(df_children) {

    df_selected <- df_children |>
      mutate(
        # =======================================================================
        # TOTAL INTAKE (before absorption)
        # =======================================================================

        fe_mg_total = if_else(
          flag_biofortified == 1L, fe_mg_total_qpm, fe_mg_total_baseline
        ),
        zn_mg_total = if_else(
          flag_biofortified == 1L, zn_mg_total_qpm, zn_mg_total_baseline
        ),
        prot_g_total = if_else(
          flag_biofortified == 1L, prot_g_total_qpm, prot_g_total_baseline
        ),

        # =======================================================================
        # ABSORBED INTAKE
        # =======================================================================

        fe_absorbed_mg = if_else(
          flag_biofortified == 1L, fe_absorbed_mg_qpm, fe_absorbed_mg_baseline
        ),
        taz_mg = if_else(
          flag_biofortified == 1L, taz_mg_qpm, taz_mg_baseline
        ),

        # =======================================================================
        # TOTAL ENERGY
        # =======================================================================

        ene_kcal_total = if_else(
          flag_biofortified == 1L, ene_kcal_total_qpm, ene_kcal_total_baseline
        ),

        # =======================================================================
        # MAIZE-SOURCE ENERGY
        # =======================================================================

        ene_kcal_maiz = kcal_individuo_dia_maiz,

        # =======================================================================
        # STUNTING OUTCOMES (pre-computed in 05_06 via GAM)
        # =======================================================================

        hfa_calculated = if_else(
          flag_biofortified == 1L, hfa_qpm, hfa
        ),
        zlen_predicted = if_else(
          flag_biofortified == 1L, haz_qpm, haz_calibrated
        )
      )

    return(df_selected)
  }
}
```

## A6. Aggregate Nutritional Metrics at National Level

Aggregate nutritional and stunting outcomes at national level using survey-weighted calculations. This function produces the national summary statistics for the Guatemala Total row.

**Metrics calculated:**

| Category | Metrics |
|----------|---------|
| Stunting | `stunting_pct`, `mean_zlen`, `n_children` |
| Zinc | `mean_zinc`, `mean_zinc_total`, `zinc_inad_pct`, `zinc_inad_under12_pct`, `zinc_inad_12to35_pct`, `zinc_inad_36to59_pct` |
| Iron | `mean_iron`, `mean_iron_total`, `iron_inad_pct`, `iron_inad_under36_pct`, `iron_inad_36to59_pct` |
| Protein | `mean_protein`, `protein_inad_pct`, `protein_inad_under36_pct`, `protein_inad_36to59_pct` |

**Inadequacy Thresholds (absorbed values):**

| Nutrient | Age Group | Threshold | Source |
|----------|-----------|-----------|--------|
| Zinc | < 12 months | 0.732 mg/day | EFSA, 2014 |
| Zinc | 12-35 months | 1.074 mg/day | EFSA, 2014 |
| Zinc | 36-59 months | 1.390 mg/day | EFSA, 2014 |
| Iron | All ages | 0.5 mg/day | Allen et al., 2020 / EFSA |
| Protein | < 36 months | 13 g/day | WHO/FAO |
| Protein | 36-59 months | 19 g/day | WHO/FAO |

**Sources:**

- **Zinc**: EFSA, 2014. Scientific Opinion on Dietary Reference Values for zinc. EFSA Journal 2014;12(10):3844. https://efsa.onlinelibrary.wiley.com/doi/epdf/10.2903/j.efsa.2014.3844
- **Iron**: Allen et al., 2020. https://pubmed.ncbi.nlm.nih.gov/31701998/ and WHO/FAO, 2006. https://www.who.int/publications/i/item/9241594012
- **Protein**: WHO/FAO protein requirements for young children.

**Survey design note:**

```{r}
#| echo: false
#| output: asis

if (is_pco) {
  cat("Bootstrap expansion in Script 03_08 broke the original cluster structure. Survey design uses `ids = 1` (simple random sampling) with `weight_calibrated` as weights.")
} else {
  cat("Survey design uses `ids = ~upm` (primary sampling units) with `weight_calibrated` as weights.")
}
```

**Parameters:**

- `df_children`: Data frame with scenario-specific nutritional values (from A5)

**Returns:** Data frame with one row containing all national aggregated metrics.

```{r}
#| label: a6-aggregate-nutritional-national

# =============================================================================
# A6. Aggregate Nutritional Metrics at National Level
# =============================================================================

aggregate_nutritional_national <- function(df_children) {

  # Zinc thresholds by age group (3 levels)
  ZINC_THRESHOLD_UNDER12 <- 0.732
  ZINC_THRESHOLD_12TO35  <- 1.074
  ZINC_THRESHOLD_36TO59  <- 1.390

  # Iron threshold (uniform across ages)
  IRON_THRESHOLD <- 0.5

  # Protein thresholds by age group (2 levels)
  PROT_THRESHOLD_UNDER36 <- 13
  PROT_THRESHOLD_36TO59  <- 19

  # Maize energy-to-weight conversion factor (kcal per gram)
  # Source: INCAP Food Composition Table — nixtamalized maize (tortilla)
  MAIZE_KCAL_PER_G <- 3.6

  # ===========================================================================
  # STEP 1: Create survey design object
  # ===========================================================================

  svy_design <- if (is_pco) {
    df_children |>
      as_survey_design(
        ids = 1,
        weights = weight_calibrated,
        strata = NULL
      )
  } else {
    df_children |>
      as_survey_design(
        ids = upm,
        weights = weight_calibrated,
        strata = NULL
      )
  }

  # ===========================================================================
  # STEP 2: Calculate stunting metrics (national)
  # ===========================================================================

  df_stunting <- svy_design |>
    summarise(
      stunting_pct = survey_mean(
        hfa_calculated == 1L,
        na.rm = TRUE,
        vartype = NULL
      ),
      mean_zlen = survey_mean(
        zlen_predicted,
        na.rm = TRUE,
        vartype = NULL
      ),
      n_children = survey_total(1, vartype = NULL),
      n_children_biofortified = survey_total(
        flag_biofortified,
        vartype = NULL
      )
    ) |>
    mutate(stunting_pct = stunting_pct * 100)

  # ===========================================================================
  # STEP 3: Calculate mean nutritional intake (national)
  # ===========================================================================

  df_nutrition_means <- svy_design |>
    summarise(
      # Absorbed intake (bioavailable)
      mean_zinc = survey_mean(taz_mg, na.rm = TRUE, vartype = NULL),
      mean_iron = survey_mean(fe_absorbed_mg, na.rm = TRUE, vartype = NULL),
      mean_protein = survey_mean(prot_g_total, na.rm = TRUE, vartype = NULL),
      # Total intake (before absorption) — for reference
      mean_zinc_total = survey_mean(zn_mg_total, na.rm = TRUE, vartype = NULL),
      mean_iron_total = survey_mean(fe_mg_total, na.rm = TRUE, vartype = NULL),
      # Consumption tab: maize intake in grams (from energy, factor = 1.0)
      maize_g_day_mean = survey_mean(
        ene_kcal_maiz / MAIZE_KCAL_PER_G,
        na.rm = TRUE, vartype = NULL
      ),
      # Energy tab: total energy intake (baseline, factor = 1.0)
      mean_energy = survey_mean(ene_kcal_total, na.rm = TRUE, vartype = NULL)
    )

  # ---------------------------------------------------------------------------
  # STEP 3b: Calculate mean energy intake by age group (national)
  # ---------------------------------------------------------------------------

  df_energy_under36 <- svy_design |>
    filter(.data[[age_var]] < 36) |>
    summarise(
      mean_energy_under36 = survey_mean(
        ene_kcal_total, na.rm = TRUE, vartype = NULL
      )
    )

  df_energy_36to59 <- svy_design |>
    filter(.data[[age_var]] >= 36) |>
    summarise(
      mean_energy_36to59 = survey_mean(
        ene_kcal_total, na.rm = TRUE, vartype = NULL
      )
    )

  # ===========================================================================
  # STEP 4: Calculate overall inadequacy rates (national)
  # ===========================================================================

  df_inad_overall <- svy_design |>
    mutate(
      zinc_inadequate = case_when(
        .data[[age_var]] < 12 & taz_mg < ZINC_THRESHOLD_UNDER12 ~ 1L,
        .data[[age_var]] >= 12 & .data[[age_var]] < 36 & taz_mg < ZINC_THRESHOLD_12TO35 ~ 1L,
        .data[[age_var]] >= 36 & taz_mg < ZINC_THRESHOLD_36TO59 ~ 1L,
        TRUE ~ 0L
      ),
      iron_inadequate = if_else(fe_absorbed_mg < IRON_THRESHOLD, 1L, 0L),
      protein_inadequate = case_when(
        .data[[age_var]] < 36 & prot_g_total < PROT_THRESHOLD_UNDER36 ~ 1L,
        .data[[age_var]] >= 36 & prot_g_total < PROT_THRESHOLD_36TO59 ~ 1L,
        TRUE ~ 0L
      )
    ) |>
    summarise(
      zinc_inad_pct = survey_mean(zinc_inadequate, na.rm = TRUE, vartype = NULL) * 100,
      iron_inad_pct = survey_mean(iron_inadequate, na.rm = TRUE, vartype = NULL) * 100,
      protein_inad_pct = survey_mean(protein_inadequate, na.rm = TRUE, vartype = NULL) * 100
    )

  # ===========================================================================
  # STEP 5: Calculate inadequacy rates by age group - ZINC (3 groups)
  # ===========================================================================

  df_zinc_under12 <- svy_design |>
    filter(.data[[age_var]] < 12) |>
    mutate(zinc_inadequate = if_else(taz_mg < ZINC_THRESHOLD_UNDER12, 1L, 0L)) |>
    summarise(
      zinc_inad_under12_pct = survey_mean(zinc_inadequate, na.rm = TRUE, vartype = NULL) * 100
    )

  df_zinc_12to35 <- svy_design |>
    filter(.data[[age_var]] >= 12 & .data[[age_var]] < 36) |>
    mutate(zinc_inadequate = if_else(taz_mg < ZINC_THRESHOLD_12TO35, 1L, 0L)) |>
    summarise(
      zinc_inad_12to35_pct = survey_mean(zinc_inadequate, na.rm = TRUE, vartype = NULL) * 100
    )

  df_zinc_36to59 <- svy_design |>
    filter(.data[[age_var]] >= 36) |>
    mutate(zinc_inadequate = if_else(taz_mg < ZINC_THRESHOLD_36TO59, 1L, 0L)) |>
    summarise(
      zinc_inad_36to59_pct = survey_mean(zinc_inadequate, na.rm = TRUE, vartype = NULL) * 100
    )

  # ===========================================================================
  # STEP 6: Calculate inadequacy rates by age group - IRON (2 groups)
  # ===========================================================================

  df_iron_under36 <- svy_design |>
    filter(.data[[age_var]] < 36) |>
    mutate(iron_inadequate = if_else(fe_absorbed_mg < IRON_THRESHOLD, 1L, 0L)) |>
    summarise(
      iron_inad_under36_pct = survey_mean(iron_inadequate, na.rm = TRUE, vartype = NULL) * 100
    )

  df_iron_36to59 <- svy_design |>
    filter(.data[[age_var]] >= 36) |>
    mutate(iron_inadequate = if_else(fe_absorbed_mg < IRON_THRESHOLD, 1L, 0L)) |>
    summarise(
      iron_inad_36to59_pct = survey_mean(iron_inadequate, na.rm = TRUE, vartype = NULL) * 100
    )

  # ===========================================================================
  # STEP 7: Calculate inadequacy rates by age group - PROTEIN (2 groups)
  # ===========================================================================

  df_prot_under36 <- svy_design |>
    filter(.data[[age_var]] < 36) |>
    mutate(protein_inadequate = if_else(prot_g_total < PROT_THRESHOLD_UNDER36, 1L, 0L)) |>
    summarise(
      protein_inad_under36_pct = survey_mean(protein_inadequate, na.rm = TRUE, vartype = NULL) * 100
    )

  df_prot_36to59 <- svy_design |>
    filter(.data[[age_var]] >= 36) |>
    mutate(protein_inadequate = if_else(prot_g_total < PROT_THRESHOLD_36TO59, 1L, 0L)) |>
    summarise(
      protein_inad_36to59_pct = survey_mean(protein_inadequate, na.rm = TRUE, vartype = NULL) * 100
    )

  # ===========================================================================
  # STEP 8: Combine all metrics into single row (bind_cols, not joins)
  # ===========================================================================

  df_aggregated <- df_stunting |>
    bind_cols(df_nutrition_means) |>
    bind_cols(df_inad_overall) |>
    bind_cols(df_zinc_under12) |>
    bind_cols(df_zinc_12to35) |>
    bind_cols(df_zinc_36to59) |>
    bind_cols(df_iron_under36) |>
    bind_cols(df_iron_36to59) |>
    bind_cols(df_prot_under36) |>
    bind_cols(df_prot_36to59) |>
    bind_cols(df_energy_under36) |>
    bind_cols(df_energy_36to59)

  return(df_aggregated)
}
```

---

# Phase B: Data Loading

This phase loads all required datasets for scenario simulation:

1. **`r if (is_pco) "Synthetic population" else "ENCOVI children"`**: Children 6-59 months with pre-calculated baseline and QPM nutritional/stunting profiles (from `r if (is_pco) "04_08" else "05_06"`)
2. **Farmer adoption data**: Ranked farmers with economic metrics, adoption scores, per-farmer entry penetration `p_entry`, and behavioural profile (from 02_06)
3. **Market baseline**: Departmental baseline metrics (farmers, land, consumption) (from 06_01)
4. **Saturation scenarios**: Pre-calculated coverage by department x penetration level (from 06_01)

## B1. Load `r if (is_pco) "Synthetic Population with Pre-Computed Profiles" else "ENCOVI Children with Priority Assignment"`

```{r}
#| echo: false
#| output: asis

if (is_pco) {
  cat("Load the synthetic population from Script 04_08, which contains pre-calculated baseline and QPM nutritional profiles (intake, bioavailability, inadequacy, nutritional index) and GAM-based stunting outcomes. This dataset enables efficient scenario simulation by selecting between `*_baseline` and `*_qpm` columns rather than recalculating on-the-fly.")
} else {
  cat("Load the children aged 6-59 months from Script 05_06, which enriches the 05_05 dataset with `priority_level` and `priority_label`. This dataset contains pre-calculated baseline and QPM nutritional profiles, stunting outcomes, and the household-level priority hierarchy.")
}
```

```{r}
#| label: b1-load-children-population

# =============================================================================
# B1. Load Children Data
# =============================================================================

if (is_pco) {
  df_children_raw <- read_parquet(
    here("01_data", "02_processed",
         "scenarios", "04_08_baseline_qpm_demographics.parquet")
  ) |>
    left_join(
      read_parquet(
        here("01_data", "02_processed",
             "scenarios", "04_08_baseline_qpm_profiles.parquet")
      ),
      by = c("synthetic_id", "departamento", "weight_calibrated")
    )
} else {
  df_children_raw <- read_parquet(
    here("01_data", "02_processed",
         "transfer", "05_06_encovi_children_priority.parquet")
  )
}
```

## B2. Load Farmer Adoption Data

Load farmer-level data with adoption scores, economic metrics, per-farmer entry penetration point (`p_entry`), and behavioural profile (`profile`) from Script `r paste0("02_06_", seed_suffix, "_", subsidy_suffix)`. The parquet integrates the seed-specific yield assumptions and the subsidy-specific seed prices applied in Script `r paste0("02_04_", seed_suffix, "_", subsidy_suffix)`, including the seven monetary subsidy decomposition columns computed in Phase E of Script 02_04.

Farmers are sorted by `adoption_score` (descending). The per-farmer entry penetration point `p_entry` is already present in the parquet; filtering adopters at each penetration level uses `p_entry <= p` and then computes the conversion ratio $r_i(p)$ via `apply_adoption_curve()`.

```{r}
#| label: b2-load-farmer-data

# =============================================================================
# B2. Load Farmer Adoption Data
# =============================================================================

df_farmers_raw <- read_parquet(
  here("01_data", "02_processed",
       "scenarios", paste0("02_06_", seed_suffix, "_", subsidy_suffix,
                           "_farmers_adoption_scores.parquet"))
)

# ===========================================================================
# Sort by adoption score (best candidates first). The entry penetration
# point p_entry and behavioural profile are already present in the parquet.
# ===========================================================================

df_farmers_ranked <- df_farmers_raw |>
  arrange(desc(adoption_score))
```

## B3. Load Market Baseline

Load departmental baseline metrics from Script `r paste0("06_01_", seed_suffix, "_", subsidy_suffix)`. This dataset contains static departmental characteristics that serve as denominators for penetration percentage calculations.

```{r}
#| label: b3-load-market-baseline

# =============================================================================
# B3. Load Market Baseline
# =============================================================================

df_market_baseline <- read_parquet(
  here("01_data", "02_processed",
       "scenarios", paste0("06_01_", seed_suffix, "_", subsidy_suffix,
                           "_market_baseline.parquet"))
)
```

## B4. Load Saturation Scenarios

Load pre-calculated departmental coverage across penetration levels from Script `r paste0("06_01_", seed_suffix, "_", subsidy_suffix)`. The original dataset contains 101 levels (0-100% in 1% increments); we filter to the 11 levels needed for scenario simulation (0%, 10%, 20%... 100%).

**Key insight:** The gravity-based redistribution is already applied in this dataset. Each row represents the final coverage a department achieves after surplus production flows through the gravity model network.

```{r}
#| label: b4-load-saturation

# =============================================================================
# B4. Load Saturation Scenarios
# =============================================================================

df_saturation_raw <- read_parquet(
  here("01_data", "02_processed",
       "scenarios", paste0("06_01_", seed_suffix, "_", subsidy_suffix,
                           "_saturation_scenarios.parquet"))
)

# ===========================================================================
# FILTER TO 11 PENETRATION LEVELS
# Levels: 0%, 10%, 20%, 30%, 40%, 50%, 60%, 70%, 80%, 90%, 100%
# ===========================================================================

df_saturation <- df_saturation_raw |>
  filter(
    near(penetration, 0.0) | near(penetration, 0.1) | near(penetration, 0.2) |
    near(penetration, 0.3) | near(penetration, 0.4) | near(penetration, 0.5) |
    near(penetration, 0.6) | near(penetration, 0.7) | near(penetration, 0.8) |
    near(penetration, 0.9) | near(penetration, 1.0)
  )

# Clean up raw data
rm(df_saturation_raw)
```

---

# Phase C: Baseline Calculations

Calculate baseline metrics at 0% penetration (no biofortification). The `r if (is_pco) "synthetic" else "ENCOVI"` population already contains pre-calculated baseline nutritional and stunting profiles, so we select baseline values and aggregate directly.

## C1. Prepare Baseline Nutrient Variables

Select baseline nutritional values for all children (no biofortification). This maps the pre-calculated baseline columns to the variable names expected by the aggregation function.

```{r}
#| label: c1-prepare-baseline

# =============================================================================
# C1. Prepare Baseline Nutritional Variables
# =============================================================================

if (is_pco) {
  df_baseline <- df_children_raw |>
    mutate(
      # Map baseline columns to aggregation-expected names
      fe_mg_total        = fe_mg_total_baseline,
      zn_mg_total        = zn_mg_total_baseline,
      prot_g_total       = prot_g_total_baseline,
      fe_absorbed_mg     = fe_absorbed_mg_baseline,
      taz_mg             = taz_mg_baseline,
      ene_kcal_maiz      = ene_kcal_maiz,
      ene_kcal_total     = ene_kcal_maiz + ene_kcal_nomaiz,
      # hfa_calculated and zlen_predicted already represent baseline values
      hfa_calculated     = hfa_calculated,
      zlen_predicted     = zlen_predicted,
      # No biofortified assignment exists at 0% penetration
      flag_biofortified  = 0L
    )
} else {
  df_baseline <- df_children_raw |>
    mutate(
      # Map baseline columns to aggregation-expected names
      fe_mg_total    = fe_mg_total_baseline,
      zn_mg_total    = zn_mg_total_baseline,
      prot_g_total   = prot_g_total_baseline,
      ene_kcal_total = ene_kcal_total_baseline,
      fe_absorbed_mg = fe_absorbed_mg_baseline,
      taz_mg         = taz_mg_baseline,
      ene_kcal_maiz  = kcal_individuo_dia_maiz,
      hfa_calculated = hfa,
      zlen_predicted = haz_calibrated,
      # No biofortified assignment exists at 0% penetration
      flag_biofortified = 0L
    )
}
```

## C2. Aggregate Baseline at National Level

Aggregate baseline metrics using the `aggregate_nutritional_national()` function. This produces the reference values for calculating improvements at each penetration level.

```{r}
#| label: c2-aggregate-baseline-national

# =============================================================================
# C2. Aggregate Baseline at National Level
# =============================================================================

df_baseline_national <- aggregate_nutritional_national(df_baseline) |>
  mutate(penetration = 0, .before = everything())
```

## C3. Calculate Baseline Economic Metrics

At 0% penetration, no farmers adopt biofortified seed. All economic metrics are zero. Market baseline totals are included for reference.

```{r}
#| label: c3-baseline-economic

# =============================================================================
# C3. Calculate Baseline Economic Metrics at National Level
# =============================================================================

if (is_pco) {
  df_baseline_economic <- df_market_baseline |>
    summarise(
      population = sum(population, na.rm = TRUE),
      n_farmers_total = sum(n_farmers_total, na.rm = TRUE),
      n_farmers_hybrid = sum(n_farmers_hybrid, na.rm = TRUE),
      n_farmers_opv = sum(n_farmers_opv, na.rm = TRUE),
      land_ha_total = sum(land_ha_total, na.rm = TRUE),
      consumption_fsi_mt = sum(consumption_fsi_mt, na.rm = TRUE)
    ) |>
    mutate(
      penetration = 0,
      n_farmers_bio = 0,
      n_hybrid_bio = 0,
      n_opv_bio = 0,
      land_bio_mz = 0,
      land_bio_ha = 0,
      income_increment_total = 0,
      income_increment_median = 0,
      production_increment_total = 0,
      production_biofortified_total = 0,
      cost_seeds_increment_total = 0,
      # Subsidy block (all zero at baseline: no adopters, no monetary flows)
      cost_seeds_bio_no_subsidy_total = 0,
      cost_seeds_bio_current_regime_total = 0,
      subsidy_amount_absolute_total = 0,
      newseed_subsidy_total = 0,
      producer_absorption_total = 0,
      subsidy_reduction_cost_total = 0,
      newseed_savings_vs_current_total = 0,
      coverage_pct = 0,
      bio_consumed_mt = 0,
      .before = n_farmers_total
    )
} else {
  df_baseline_economic <- df_market_baseline |>
    summarise(
      population = sum(population, na.rm = TRUE),
      n_farmers_total = sum(n_farmers_total, na.rm = TRUE),
      n_farmers_hybrid = sum(n_farmers_hybrid, na.rm = TRUE),
      n_farmers_opv = sum(n_farmers_opv, na.rm = TRUE),
      land_ha_total = sum(land_ha_total, na.rm = TRUE),
      consumption_fsi_mt = sum(consumption_fsi_mt, na.rm = TRUE)
    ) |>
    mutate(
      penetration = 0,
      n_farmers_bio = 0,
      n_hybrid_bio = 0,
      n_opv_bio = 0,
      land_bio_mz = 0,
      land_bio_ha = 0,
      income_increment_total = 0,
      income_increment_median = 0,
      production_increment_total = 0,
      production_biofortified_total = 0,
      cost_seeds_increment_total = 0,
      # Subsidy block (all zero at baseline: no adopters, no monetary flows)
      cost_seeds_bio_no_subsidy_total = 0,
      cost_seeds_bio_current_regime_total = 0,
      subsidy_amount_absolute_total = 0,
      newseed_subsidy_total = 0,
      producer_absorption_total = 0,
      subsidy_reduction_cost_total = 0,
      newseed_savings_vs_current_total = 0,
      coverage_pct = 0,
      bio_consumed_mt = 0,
      .before = n_farmers_total
    )
}
```

## C4. Combine Baseline Datasets

Join economic and nutritional baseline metrics into a single baseline row.

```{r}
#| label: c4-combine-baseline

# =============================================================================
# C4. Combine Baseline Datasets
# =============================================================================

df_scenario_baseline <- df_baseline_economic |>
  bind_cols(df_baseline_national |> select(-penetration))
```

```{r}
#| label: tbl-c4-baseline-summary
#| tbl-cap: "Baseline National Metrics (0% Penetration)"

# =============================================================================
# Baseline Summary Table
# =============================================================================

df_tbl_c4 <- df_scenario_baseline |>
  select(
    n_children, stunting_pct, mean_zlen,
    mean_zinc, mean_iron, mean_protein,
    zinc_inad_pct, iron_inad_pct, protein_inad_pct
  ) |>
  pivot_longer(everything(), names_to = "Metric", values_to = "Value") |>
  mutate(
    Metric = case_match(
      Metric,
      "n_children" ~ "Total Children (N)",
      "stunting_pct" ~ "Stunting Prevalence (%)",
      "mean_zlen" ~ if (is_pco) "Mean ZLEN (SD units)" else "Mean HAZ (SD units)",
      "mean_zinc" ~ "Mean Absorbed Zinc (mg/day)",
      "mean_iron" ~ "Mean Absorbed Iron (mg/day)",
      "mean_protein" ~ "Mean Protein Intake (g/day)",
      "zinc_inad_pct" ~ "Zinc Inadequacy (%)",
      "iron_inad_pct" ~ "Iron Inadequacy (%)",
      "protein_inad_pct" ~ "Protein Inadequacy (%)"
    )
  )

export(df_tbl_c4, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_c4_baseline_summary.rds")))

df_tbl_c4 |>
  gt() |>
  tab_header(
    title = "National Baseline Metrics",
    subtitle = glue("Guatemala Total at 0% biofortification penetration — Seed: {seed_code} | Pathway: {pathway_label}")
  ) |>
  fmt_number(columns = Value, decimals = 2) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(df_tbl_c4)
```

---

# Phase D: Scenario Simulation

Simulate biofortification impact across 10 penetration levels (10% to 100%). For each level:

1. Filter adopting farmers using `p_entry <= p` and compute per-farmer conversion ratio $r_i(p)$
2. Calculate economic impact at national level, weighting scalable quantities by $r_i(p)$
3. Get pre-calculated coverage from saturation scenarios (consumption-weighted national average)
4. Assign biofortification to children `r if (is_pco) "hierarchically" else "using household-level priority"`
5. Select pre-calculated baseline or QPM values per child
6. Aggregate nutritional metrics at national level
7. Combine economic and nutritional results

The loop produces one row per penetration level, which is then combined with baseline for the final output.

## D1. Define Scenario Simulation Wrapper

This function orchestrates the complete simulation pipeline for a single penetration level at national level by chaining all Phase A functions in sequence:

1. **A2**: Filter adopting farmers by `p_entry <= p` and compute $r_i(p)$
2. **A3**: Calculate economic metrics at national level (scalable quantities weighted by $r_i$)
3. **Coverage**: Calculate consumption-weighted national coverage from departmental saturation
4. **A4**: Assign biofortification flag to children hierarchically
5. **A5**: Select pre-calculated nutritional values by flag
6. **A6**: Aggregate nutritional metrics at national level
7. **Combine**: Join economic and nutritional results

**Parameters:**

- `pen`: Penetration level (0-1)
- `df_farmers`: Ranked farmer data with `p_entry` and `profile`
- `df_saturation`: Pre-calculated saturation scenarios from 06_01
- `df_children`: `r if (is_pco) "Synthetic population with pre-calculated profiles (from 04_08)" else "ENCOVI children with pre-calculated profiles and priority (from 05_06)"`
- `df_market_baseline`: Market baseline for reference variables
- `seed`: Random seed for hierarchical assignment reproducibility

**Returns:** Data frame with one row containing economic and nutritional metrics at the specified penetration level.

```{r}
#| label: d1-simulation-wrapper

# =============================================================================
# D1. Define Scenario Simulation Wrapper (National)
# =============================================================================

simulate_scenario_national <- function(pen,
                                       df_farmers,
                                       df_saturation,
                                       df_children,
                                       df_market_baseline,
                                       seed = 789) {

  # ===========================================================================
  # STEP 1: Filter adopting farmers and compute r_i(p) (A2)
  # ===========================================================================

  df_adopters <- filter_adopters_by_penetration(
    df_farmers = df_farmers,
    penetration = pen
  )

  # ===========================================================================
  # STEP 2: Calculate economic metrics at national level (A3)
  # ===========================================================================

  df_economic <- calculate_economic_national(df_adopters)

  # ===========================================================================
  # STEP 3: Get coverage from saturation scenarios
  # Departmental coverage for child assignment
  # National coverage as consumption-weighted average
  # ===========================================================================

  df_coverage <- df_saturation |>
    filter(near(penetration, pen)) |>
    select(departamento, coverage_pct)

  # National coverage (weighted by annual consumption in qq)
  national_coverage <- df_saturation |>
    filter(near(penetration, pen)) |>
    summarise(
      coverage_pct = sum(coverage_pct * annual_consumption_qq, na.rm = TRUE) /
                     sum(annual_consumption_qq, na.rm = TRUE)
    ) |>
    pull(coverage_pct)

  # ===========================================================================
  # STEP 4: Assign biofortification flag to children (A4)
  # ===========================================================================

  df_assigned <- assign_biofortification_hierarchical(
    df_children = df_children,
    df_coverage = df_coverage,
    seed = seed
  )

  # ===========================================================================
  # STEP 5: Select pre-calculated nutritional values by flag (A5)
  # ===========================================================================

  df_selected <- select_nutritional_by_flag(df_assigned)

  # ===========================================================================
  # STEP 6: Aggregate at national level (A6)
  # ===========================================================================

  df_nutritional <- aggregate_nutritional_national(df_selected)

  # ===========================================================================
  # STEP 7: Prepare market baseline reference variables and combine
  # ===========================================================================

  df_market_ref <- df_market_baseline |>
    summarise(
      n_farmers_total = sum(n_farmers_total, na.rm = TRUE),
      n_farmers_hybrid = sum(n_farmers_hybrid, na.rm = TRUE),
      n_farmers_opv = sum(n_farmers_opv, na.rm = TRUE),
      land_ha_total = sum(land_ha_total, na.rm = TRUE),
      consumption_fsi_mt = sum(consumption_fsi_mt, na.rm = TRUE),
      population = sum(population, na.rm = TRUE)
    )

  df_result <- df_economic |>
    bind_cols(df_market_ref) |>
    mutate(
      penetration = pen,
      coverage_pct = national_coverage,
      bio_consumed_mt = consumption_fsi_mt * (national_coverage / 100),
      .before = n_farmers_bio
    ) |>
    bind_cols(df_nutritional)

  return(df_result)
}
```

## D2. Execute Scenario Simulation

Execute the simulation wrapper across 10 penetration levels (10% to 100%). The baseline (0%) was already calculated in Phase C and will be combined in the next step.

```{r}
#| label: d2-execute-scenarios

# =============================================================================
# D2. Execute Scenario Simulation (National)
# =============================================================================

# ---------------------------------------------------------------------------
# Define penetration levels to simulate (0% already calculated as baseline)
# ---------------------------------------------------------------------------

penetration_levels_to_simulate <- seq(0.1, 1.0, by = 0.1)

# ---------------------------------------------------------------------------
# Execute simulation across all penetration levels
# ---------------------------------------------------------------------------

df_scenarios_simulated <- map_dfr(
  penetration_levels_to_simulate,
  ~ simulate_scenario_national(
    pen = .x,
    df_farmers = df_farmers_ranked,
    df_saturation = df_saturation,
    df_children = df_children_raw,
    df_market_baseline = df_market_baseline,
    seed = 789
  ),
  .progress = TRUE
)

# Clean up
rm(penetration_levels_to_simulate)
```

## D3. Combine All Scenarios

Combine baseline (0% penetration) with simulated scenarios (10%-100%) into a single dataset.

The final dataset contains 11 rows (one per penetration level) with complete economic and nutritional metrics at the Guatemala national level.

```{r}
#| label: d3-combine-scenarios

# =============================================================================
# D3. Combine All Scenarios
# =============================================================================

df_scenarios_all <- bind_rows(
  df_scenario_baseline,
  df_scenarios_simulated
) |>
  arrange(penetration)
```

## D4. Validation: Scenario Progression

Validate that scenario simulation produces expected patterns:

1. Stunting rate should decrease as penetration increases
2. Economic metrics (farmers, land, production) should increase monotonically; net income may plateau or decrease at high penetration as less suitable adopters (High segment with yield losses) enter the pool
3. Coverage values should increase with penetration
4. All 11 rows present (one per penetration level)

```{r}
#| label: tbl-d4-scenario-progression
#| tbl-cap: "National scenario progression across penetration levels"

# =============================================================================
# D4. Validation: Scenario Progression Summary
# =============================================================================

df_tbl_d4 <- df_scenarios_all |>
  select(
    penetration, n_farmers_bio, coverage_pct,
    stunting_pct, mean_zlen, zinc_inad_pct, iron_inad_pct
  ) |>
  arrange(penetration)

export(df_tbl_d4, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_d4_scenario_progression.rds")))

df_tbl_d4 |>
  gt() |>
  tab_header(
    title = "National Metrics by Penetration Level",
    subtitle = glue("Validation of national scenario progression — Seed: {seed_code} | Pathway: {pathway_label}")
  ) |>
  cols_label(
    penetration = "Penetration",
    n_farmers_bio = "Farmers Bio",
    coverage_pct = "Coverage",
    stunting_pct = "Stunting (%)",
    mean_zlen = if (is_pco) "Mean ZLEN" else "Mean HAZ",
    zinc_inad_pct = "Zinc Inad. (%)",
    iron_inad_pct = "Iron Inad. (%)"
  ) |>
  fmt_percent(columns = penetration, decimals = 0) |>
  fmt_number(columns = n_farmers_bio, decimals = 0, use_seps = TRUE) |>
  fmt_number(columns = coverage_pct, decimals = 1, pattern = "{x}%") |>
  fmt_number(columns = c(stunting_pct, zinc_inad_pct, iron_inad_pct), decimals = 1) |>
  fmt_number(columns = mean_zlen, decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = stunting_pct, rows = penetration == 1)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = stunting_pct, rows = penetration == 0)
  ) |>
  tab_source_note(
    source_note = glue("Seed: {seed_code} | Pathway: {pathway_label}. Expected: Stunting decreases monotonically with penetration.")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(df_tbl_d4)
```

```{r}
#| label: tbl-d5-nutritional-progression
#| tbl-cap: "National nutritional intake progression across penetration levels"

# =============================================================================
# D5. Validation: Nutritional Intake Progression
# =============================================================================

df_tbl_d5 <- df_scenarios_all |>
  select(penetration, coverage_pct, mean_zinc, mean_iron, mean_protein) |>
  arrange(penetration) |>
  mutate(
    zinc_change_pct = (mean_zinc / first(mean_zinc) - 1) * 100,
    iron_change_pct = (mean_iron / first(mean_iron) - 1) * 100,
    protein_change_pct = (mean_protein / first(mean_protein) - 1) * 100
  )

export(df_tbl_d5, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_d5_nutritional_progression.rds")))

df_tbl_d5 |>
  gt() |>
  tab_header(
    title = "Nutritional Intake by Penetration Level",
    subtitle = glue("Weighted mean bioavailable nutrients: absorbed (Zn/Fe, mg/day), digestible (protein, g/day) — Seed: {seed_code} | Pathway: {pathway_label}")
  ) |>
  cols_label(
    penetration = "Penetration",
    coverage_pct = "Coverage",
    mean_zinc = "Zinc (mg)",
    mean_iron = "Iron (mg)",
    mean_protein = "Protein (g)",
    zinc_change_pct = "Zn %",
    iron_change_pct = "Fe %",
    protein_change_pct = "Prot %"
  ) |>
  fmt_percent(columns = penetration, decimals = 0) |>
  fmt_number(columns = coverage_pct, decimals = 1, pattern = "{x}%") |>
  fmt_number(columns = c(mean_zinc, mean_iron), decimals = 3) |>
  fmt_number(columns = mean_protein, decimals = 2) |>
  fmt_number(columns = c(zinc_change_pct, iron_change_pct, protein_change_pct),
             decimals = 1, pattern = "{x}%", force_sign = TRUE) |>
  tab_spanner(label = "Absorbed Intake", columns = c(mean_zinc, mean_iron, mean_protein)) |>
  tab_spanner(label = "Change vs Baseline", columns = c(zinc_change_pct, iron_change_pct, protein_change_pct)) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = c(zinc_change_pct, iron_change_pct, protein_change_pct),
                           rows = penetration == 1)
  ) |>
  tab_style(
    style = cell_text(color = "#2E7D32", weight = "bold"),
    locations = cells_body(columns = c(zinc_change_pct, iron_change_pct, protein_change_pct),
                           rows = penetration > 0)
  ) |>
  tab_source_note(
    source_note = "Zinc and iron values are absorbed (TAZ via Miller equation, Fe at 5% absorption). Protein is PDCAAS-adjusted."
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(df_tbl_d5)
```

```{r}
#| label: tbl-d6-economic-progression
#| tbl-cap: "National economic and agricultural progression across penetration levels"

# =============================================================================
# D6. Validation: Economic & Agricultural Progression
# =============================================================================

FACTOR_MZ_TO_HA <- 0.7

df_tbl_d6 <- df_scenarios_all |>
  select(
    penetration,
    n_farmers_bio, n_hybrid_bio, n_opv_bio,
    land_bio_mz,
    production_biofortified_total,
    income_increment_total
  ) |>
  arrange(penetration) |>
  mutate(
    land_bio_ha = land_bio_mz * FACTOR_MZ_TO_HA,
    income_increment_mgtq = income_increment_total / 1e6
  ) |>
  select(
    penetration,
    n_farmers_bio, n_hybrid_bio, n_opv_bio,
    land_bio_ha,
    production_biofortified_total,
    income_increment_mgtq
  )

export(df_tbl_d6, here("01_data", "02_processed", "web_resources", paste0(file_prefix, "_tbl_d6_economic_progression.rds")))

df_tbl_d6 |>
  gt() |>
  tab_header(
    title = "Economic Impact by Penetration Level",
    subtitle = glue("Farmer adoption, land use, and economic outcomes — Seed: {seed_code} | Pathway: {pathway_label}")
  ) |>
  cols_label(
    penetration = "Penetration",
    n_farmers_bio = "Total",
    n_hybrid_bio = "Hybrid",
    n_opv_bio = "OPV",
    land_bio_ha = "Ha",
    production_biofortified_total = "Prod Bio (qq)",
    income_increment_mgtq = "Net Income Change (M GTQ)"
  ) |>
  tab_spanner(label = "Adopting Farmers", columns = c(n_farmers_bio, n_hybrid_bio, n_opv_bio)) |>
  fmt_percent(columns = penetration, decimals = 0) |>
  fmt_number(columns = c(n_farmers_bio, n_hybrid_bio, n_opv_bio), decimals = 0, use_seps = TRUE) |>
  fmt_number(columns = land_bio_ha, decimals = 0, use_seps = TRUE) |>
  fmt_number(columns = production_biofortified_total, decimals = 0, use_seps = TRUE) |>
  fmt_number(columns = income_increment_mgtq, decimals = 2) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = income_increment_mgtq, rows = penetration == 1)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = income_increment_mgtq, rows = penetration == 0)
  ) |>
  tab_source_note(
    source_note = "Net Income Change is the change in net income, with the seed cost difference included. Land converted at 0.7 ha/mz."
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Cleanup auxiliary objects
rm(FACTOR_MZ_TO_HA, df_tbl_d6)
```


---

# Phase E: Final Processing

This phase prepares the final dataset for export by calculating change metrics relative to baseline, adding derived variables required for the Overview table, and validating schema completeness.

## E1. Calculate Change Metrics

Calculate improvement metrics by comparing each scenario to the baseline (0% penetration). Change metrics follow the convention: positive values indicate improvement (e.g., stunting reduction, inadequacy reduction).

```{r}
#| label: e1-calculate-change-metrics

# =============================================================================
# E1. Calculate Change Metrics Relative to Baseline
# =============================================================================

# ---------------------------------------------------------------------------
# Extract baseline values (penetration = 0) - single row for national
# ---------------------------------------------------------------------------

baseline_values <- df_scenarios_all |>
  filter(penetration == 0) |>
  summarise(
    stunting_pct_baseline = first(stunting_pct),
    mean_zlen_baseline = first(mean_zlen),
    mean_zinc_total_baseline = first(mean_zinc_total),
    mean_iron_total_baseline = first(mean_iron_total),
    mean_zinc_baseline = first(mean_zinc),
    mean_iron_baseline = first(mean_iron),
    mean_protein_baseline = first(mean_protein),
    # Consumption and energy baselines
    maize_g_day_mean_baseline = first(maize_g_day_mean),
    mean_energy_baseline = first(mean_energy),
    mean_energy_under36_baseline = first(mean_energy_under36),
    mean_energy_36to59_baseline = first(mean_energy_36to59),
    zinc_inad_pct_baseline = first(zinc_inad_pct),
    iron_inad_pct_baseline = first(iron_inad_pct),
    protein_inad_pct_baseline = first(protein_inad_pct)
  )

# ---------------------------------------------------------------------------
# Broadcast baseline to all scenarios and calculate change metrics
# ---------------------------------------------------------------------------

df_scenarios_with_changes <- df_scenarios_all |>
  mutate(
    # Broadcast baseline values to all rows
    stunting_pct_baseline = baseline_values$stunting_pct_baseline,
    mean_zlen_baseline = baseline_values$mean_zlen_baseline,
    mean_zinc_total_baseline = baseline_values$mean_zinc_total_baseline,
    mean_iron_total_baseline = baseline_values$mean_iron_total_baseline,
    mean_zinc_baseline = baseline_values$mean_zinc_baseline,
    mean_iron_baseline = baseline_values$mean_iron_baseline,
    mean_protein_baseline = baseline_values$mean_protein_baseline,
    maize_g_day_mean_baseline = baseline_values$maize_g_day_mean_baseline,
    mean_energy_baseline = baseline_values$mean_energy_baseline,
    mean_energy_under36_baseline = baseline_values$mean_energy_under36_baseline,
    mean_energy_36to59_baseline = baseline_values$mean_energy_36to59_baseline,
    zinc_inad_pct_baseline = baseline_values$zinc_inad_pct_baseline,
    iron_inad_pct_baseline = baseline_values$iron_inad_pct_baseline,
    protein_inad_pct_baseline = baseline_values$protein_inad_pct_baseline,

    # === STUNTING CHANGE ===
    # Positive = improvement (fewer stunted children)
    stunting_change_pp = stunting_pct_baseline - stunting_pct,
    stunting_change_pct = if_else(
      stunting_pct_baseline > 0,
      (stunting_pct_baseline - stunting_pct) / stunting_pct_baseline * 100,
      0
    ),
    zlen_change = mean_zlen - mean_zlen_baseline,

    # === NUTRITIONAL INTAKE CHANGE ===
    zinc_change_pct = if_else(
      mean_zinc_baseline > 0,
      (mean_zinc - mean_zinc_baseline) / mean_zinc_baseline * 100,
      0
    ),
    iron_change_pct = if_else(
      mean_iron_baseline > 0,
      (mean_iron - mean_iron_baseline) / mean_iron_baseline * 100,
      0
    ),
    protein_change_pct = if_else(
      mean_protein_baseline > 0,
      (mean_protein - mean_protein_baseline) / mean_protein_baseline * 100,
      0
    ),

    # === INADEQUACY CHANGE ===
    # Positive = improvement (fewer inadequate children)
    zinc_inad_change_pp = zinc_inad_pct_baseline - zinc_inad_pct,
    iron_inad_change_pp = iron_inad_pct_baseline - iron_inad_pct,
    protein_inad_change_pp = protein_inad_pct_baseline - protein_inad_pct
  )

# Clean up
rm(baseline_values)
```

## E2. Add Derived Variables for Overview Table

Add unit conversions and calculated fields required for the Overview table. This includes conversion from local units (manzanas, quintales, GTQ) to international units (hectares, metric tons), and calculation of subsidy costs.

**Unit Conversions:**

| From | To | Factor |
|------|----|--------|
| Manzanas (mz) | Hectares (ha) | x 0.7 |
| Quintales (qq) | Metric Tons (MT) | x 0.04536 |

**Food System Nutrient Additions:**

The additional quantity of iron and zinc incorporated into the food system through biofortification is calculated from the total volume of QPM maize consumed (all ages, not limited to children 6-59 months) and the incremental nutrient content of biofortified varieties relative to conventional maize.

| Parameter | Iron | Zinc | Source |
|-----------|------|------|--------|
| Content in conventional maize | 2.71 mg/100g | 2.21 mg/100g | INCAP Food Composition Table |
| Biofortification factor | 1.19 (+19%) | 1.38 (+38%) | Semilla Nueva Standard Report, July-December 2020 |
| Incremental content | 0.51 mg/100g | 0.84 mg/100g | = content x (factor - 1) |

```{r}
#| label: e2-add-derived-variables

# =============================================================================
# E2. Add Derived Variables for Overview Table
# =============================================================================

# ---------------------------------------------------------------------------
# Unit conversion factors
# ---------------------------------------------------------------------------

FACTOR_MZ_TO_HA <- 0.7
FACTOR_QQ_TO_MT <- 0.04536

# Food System Nutrient Additions
# Nutrient content of raw white maize (Source: INCAP Food Composition Table)
# Biofortification factors (Source: Semilla Nueva Standard Report 2020)
FE_MG_PER_G_MAIZE <- 2.71 / 100   # 0.0271 mg iron per gram of maize
ZN_MG_PER_G_MAIZE <- 2.21 / 100   # 0.0221 mg zinc per gram of maize
FACTOR_FE <- 1.19                   # QPM iron factor (+19%)
FACTOR_ZN <- 1.38                   # QPM zinc factor (+38%)

# ---------------------------------------------------------------------------
# Calculate derived variables
# Note: At national level, baseline totals already in df from wrapper
# ---------------------------------------------------------------------------

df_scenarios_enriched <- df_scenarios_with_changes |>
  mutate(
    # === REGION IDENTIFIER ===
    region = "Guatemala Total",

    # === LAND CONVERSIONS ===
    # land_bio_ha already calculated in wrapper, but ensure consistency
    land_bio_ha = if ("land_bio_ha" %in% names(df_scenarios_with_changes)) {
      land_bio_ha
    } else {
      land_bio_mz * FACTOR_MZ_TO_HA
    },
    land_bio_pct = if_else(
      land_ha_total > 0,
      land_bio_ha / land_ha_total * 100,
      0
    ),

    # === FARMER PERCENTAGES ===
    farmers_bio_pct = if_else(
      n_farmers_total > 0,
      n_farmers_bio / n_farmers_total * 100,
      0
    ),
    hybrid_bio_pct = if_else(
      n_farmers_hybrid > 0,
      n_hybrid_bio / n_farmers_hybrid * 100,
      0
    ),
    opv_bio_pct = if_else(
      n_farmers_opv > 0,
      n_opv_bio / n_farmers_opv * 100,
      0
    ),

    # === PRODUCTION CONVERSIONS ===
    production_increment_mt = production_increment_total * FACTOR_QQ_TO_MT,
    production_biofortified_mt = production_biofortified_total * FACTOR_QQ_TO_MT,

    # === INCOME PER FARMER ===
    income_per_farmer_total = if_else(
      n_farmers_total > 0,
      income_increment_total / n_farmers_total,
      0
    ),
    income_per_farmer_adopter = if_else(
      n_farmers_bio > 0,
      income_increment_total / n_farmers_bio,
      0
    ),

    # === FOOD SYSTEM NUTRIENT ADDITIONS ===
    iron_added_kg_year = bio_consumed_mt * FE_MG_PER_G_MAIZE * (FACTOR_FE - 1),
    zinc_added_kg_year = bio_consumed_mt * ZN_MG_PER_G_MAIZE * (FACTOR_ZN - 1),

    # === TOTAL POPULATION CONSUMING BIOFORTIFIED MAIZE ===
    n_people_biofortified = population * (coverage_pct / 100),

    # === CHILDREN 6-59 MONTHS CONSUMING BIOFORTIFIED MAIZE ===
    children_coverage_pct = if_else(
      n_children > 0,
      n_children_biofortified / n_children * 100,
      0
    ),

    # === STUNTING ABSOLUTE NUMBERS ===
    n_stunted_baseline = n_children * (stunting_pct_baseline / 100),
    n_stunted_scenario = n_children * (stunting_pct / 100),
    n_stunting_averted = n_stunted_baseline - n_stunted_scenario,

    # === INADEQUACY ABSOLUTE NUMBERS ===
    n_zinc_inad_baseline = n_children * (zinc_inad_pct_baseline / 100),
    n_zinc_inad_scenario = n_children * (zinc_inad_pct / 100),
    n_zinc_inad_averted = n_zinc_inad_baseline - n_zinc_inad_scenario,

    n_iron_inad_baseline = n_children * (iron_inad_pct_baseline / 100),
    n_iron_inad_scenario = n_children * (iron_inad_pct / 100),
    n_iron_inad_averted = n_iron_inad_baseline - n_iron_inad_scenario,

    n_protein_inad_baseline = n_children * (protein_inad_pct_baseline / 100),
    n_protein_inad_scenario = n_children * (protein_inad_pct / 100),
    n_protein_inad_averted = n_protein_inad_baseline - n_protein_inad_scenario
  )

# Clean up
rm(FACTOR_MZ_TO_HA, FACTOR_QQ_TO_MT,
   FE_MG_PER_G_MAIZE, ZN_MG_PER_G_MAIZE, FACTOR_FE, FACTOR_ZN)
```

## E3. Select and Order Final Variables

Select and reorder variables for the final export schema. Variables are organized by category to facilitate downstream use in Shiny dashboard and Overview table generation. Quantiles are not included at national level.

```{r}
#| label: e3-select-final-variables

# =============================================================================
# E3. Select and Order Final Variables
# =============================================================================

df_scenarios_final <- df_scenarios_enriched |>
  mutate(
    # Subsidy identifiers (constant across all rows for this combo,
    # added explicitly so the parquet is self-describing for the dashboard)
    subsidy_scenario      = !!subsidy_label,
    subsidy_suffix        = !!subsidy_suffix,
    subsidy_reduction_pct = as.integer(
      sub("^sub", "", !!subsidy_suffix)
    )
  ) |>
  select(
    # =========================================================================
    # IDENTIFIERS
    # =========================================================================
    penetration,
    region,
    subsidy_scenario,
    subsidy_suffix,
    subsidy_reduction_pct,

    # =========================================================================
    # FARMER METRICS
    # =========================================================================
    n_farmers_total,
    n_farmers_bio,
    farmers_bio_pct,
    n_farmers_hybrid,
    n_hybrid_bio,
    hybrid_bio_pct,
    n_farmers_opv,
    n_opv_bio,
    opv_bio_pct,

    # =========================================================================
    # LAND METRICS
    # =========================================================================
    land_ha_total,
    land_bio_mz,
    land_bio_ha,
    land_bio_pct,

    # =========================================================================
    # ECONOMIC METRICS
    # =========================================================================
    income_increment_total,
    income_increment_median,
    income_per_farmer_total,
    income_per_farmer_adopter,
    cost_seeds_increment_total,
    production_increment_total,
    production_increment_mt,
    production_biofortified_total,
    production_biofortified_mt,

    # =========================================================================
    # SUBSIDY METRICS (GTQ, survey-weighted, adopter aggregates)
    # =========================================================================
    cost_seeds_bio_no_subsidy_total,
    cost_seeds_bio_current_regime_total,
    subsidy_amount_absolute_total,
    newseed_subsidy_total,
    producer_absorption_total,
    subsidy_reduction_cost_total,
    newseed_savings_vs_current_total,

    # =========================================================================
    # COVERAGE METRICS
    # =========================================================================
    coverage_pct,
    consumption_fsi_mt,
    bio_consumed_mt,
    # Food System Nutrient Additions
    iron_added_kg_year,
    zinc_added_kg_year,

    # =========================================================================
    # TOTAL POPULATION (Consumption tab — Total Population view)
    # =========================================================================
    population,
    n_people_biofortified,

    # =========================================================================
    # CONSUMPTION METRICS (Consumption tab)
    # =========================================================================
    maize_g_day_mean_baseline,
    maize_g_day_mean,

    # =========================================================================
    # ENERGY METRICS (Energy tab)
    # =========================================================================
    mean_energy_baseline,
    mean_energy,
    # Age-group means and baselines
    mean_energy_under36_baseline,
    mean_energy_under36,
    mean_energy_36to59_baseline,
    mean_energy_36to59,

    # =========================================================================
    # CHILD POPULATION
    # =========================================================================
    n_children,
    n_children_biofortified,
    children_coverage_pct,

    # =========================================================================
    # STUNTING METRICS
    # =========================================================================
    # Baseline
    stunting_pct_baseline,
    n_stunted_baseline,
    mean_zlen_baseline,
    # Scenario
    stunting_pct,
    n_stunted_scenario,
    mean_zlen,
    # Change
    stunting_change_pp,
    stunting_change_pct,
    n_stunting_averted,
    zlen_change,

    # =========================================================================
    # ZINC METRICS
    # =========================================================================
    # Mean intake
    mean_zinc_total_baseline,
    mean_zinc_total,
    mean_zinc_baseline,
    mean_zinc,
    zinc_change_pct,
    # Overall inadequacy
    zinc_inad_pct_baseline,
    zinc_inad_pct,
    zinc_inad_change_pp,
    # Absolute counts
    n_zinc_inad_baseline,
    n_zinc_inad_scenario,
    n_zinc_inad_averted,
    # Age-specific inadequacy rates
    zinc_inad_under12_pct,
    zinc_inad_12to35_pct,
    zinc_inad_36to59_pct,

    # =========================================================================
    # IRON METRICS
    # =========================================================================
    # Mean intake
    mean_iron_total_baseline,
    mean_iron_total,
    mean_iron_baseline,
    mean_iron,
    iron_change_pct,
    # Overall inadequacy
    iron_inad_pct_baseline,
    iron_inad_pct,
    iron_inad_change_pp,
    # Absolute counts
    n_iron_inad_baseline,
    n_iron_inad_scenario,
    n_iron_inad_averted,
    # Age-specific inadequacy rates
    iron_inad_under36_pct,
    iron_inad_36to59_pct,

    # =========================================================================
    # PROTEIN METRICS
    # =========================================================================
    # Mean intake
    mean_protein_baseline,
    mean_protein,
    protein_change_pct,
    # Overall inadequacy
    protein_inad_pct_baseline,
    protein_inad_pct,
    protein_inad_change_pp,
    # Absolute counts
    n_protein_inad_baseline,
    n_protein_inad_scenario,
    n_protein_inad_averted,
    # Age-specific inadequacy rates
    protein_inad_under36_pct,
    protein_inad_36to59_pct

  ) |>
  arrange(penetration)
```

## E4. Validate Schema Completeness

Verify that all required variables for the Overview table and Shiny dashboard are present in the final dataset.

```{r}
#| label: e4-schema-validation
#| tbl-cap: "Schema validation summary"

# =============================================================================
# E4. Schema Validation
# =============================================================================

# Define required variables by category (no departamento, no quantiles)
required_vars <- list(
  "Identifiers" = c("penetration", "region",
                    "subsidy_scenario", "subsidy_suffix", "subsidy_reduction_pct"),

  "Farmers" = c("n_farmers_total", "n_farmers_bio", "farmers_bio_pct",
                "n_farmers_hybrid", "n_hybrid_bio", "hybrid_bio_pct",
                "n_farmers_opv", "n_opv_bio", "opv_bio_pct"),

  "Land" = c("land_ha_total", "land_bio_mz", "land_bio_ha", "land_bio_pct"),

  "Economic" = c("income_increment_total", "income_increment_median",
                 "income_per_farmer_total", "income_per_farmer_adopter",
                 "cost_seeds_increment_total",
                 "production_increment_total", "production_increment_mt"),

  "Subsidy" = c("cost_seeds_bio_no_subsidy_total",
                "cost_seeds_bio_current_regime_total",
                "subsidy_amount_absolute_total",
                "newseed_subsidy_total",
                "producer_absorption_total",
                "subsidy_reduction_cost_total",
                "newseed_savings_vs_current_total"),

  "Coverage" = c("coverage_pct", "consumption_fsi_mt", "bio_consumed_mt",
                 "iron_added_kg_year", "zinc_added_kg_year"),

  "Population" = c("population", "n_people_biofortified"),

  "Consumption" = c("maize_g_day_mean_baseline", "maize_g_day_mean"),

  "Energy" = c("mean_energy_baseline", "mean_energy",
               "mean_energy_under36_baseline", "mean_energy_under36",
               "mean_energy_36to59_baseline", "mean_energy_36to59"),

  "Children" = c("n_children", "n_children_biofortified",
                 "children_coverage_pct"),

  "Stunting" = c("stunting_pct_baseline", "stunting_pct", "stunting_change_pp",
                 "stunting_change_pct", "n_stunted_baseline", "n_stunted_scenario",
                 "n_stunting_averted", "mean_zlen_baseline", "mean_zlen", "zlen_change"),

  "Zinc" = c("mean_zinc_baseline", "mean_zinc", "zinc_change_pct",
             "zinc_inad_pct_baseline", "zinc_inad_pct", "zinc_inad_change_pp",
             "n_zinc_inad_baseline", "n_zinc_inad_scenario", "n_zinc_inad_averted",
             "zinc_inad_under12_pct", "zinc_inad_12to35_pct", "zinc_inad_36to59_pct"),

  "Iron" = c("mean_iron_baseline", "mean_iron", "iron_change_pct",
             "iron_inad_pct_baseline", "iron_inad_pct", "iron_inad_change_pp",
             "n_iron_inad_baseline", "n_iron_inad_scenario", "n_iron_inad_averted",
             "iron_inad_under36_pct", "iron_inad_36to59_pct"),

  "Protein" = c("mean_protein_baseline", "mean_protein", "protein_change_pct",
                "protein_inad_pct_baseline", "protein_inad_pct", "protein_inad_change_pp",
                "n_protein_inad_baseline", "n_protein_inad_scenario", "n_protein_inad_averted",
                "protein_inad_under36_pct", "protein_inad_36to59_pct")
)

# Check presence of each variable
df_validation <- map_dfr(names(required_vars), function(group) {
  vars <- required_vars[[group]]
  tibble(
    Category = group,
    Required = length(vars),
    Present = sum(vars %in% names(df_scenarios_final)),
    Missing = paste(setdiff(vars, names(df_scenarios_final)), collapse = ", ")
  )
}) |>
  mutate(
    Status = if_else(Required == Present, "Complete", "Incomplete"),
    Missing = if_else(Missing == "", "---", Missing)
  )

# Display validation table
df_validation |>
  gt() |>
  tab_header(
    title = "Schema Validation Summary",
    subtitle = glue("Required variables for national scenarios — Seed: {seed_code} | Pathway: {pathway_label}")
  ) |>
  cols_label(
    Category = "Category",
    Required = "Required",
    Present = "Present",
    Missing = "Missing Variables",
    Status = "Status"
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = Status, rows = Status == "Complete")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = Status, rows = Status == "Incomplete")
  ) |>
  grand_summary_rows(
    columns = c(Required, Present),
    fns = list(Total = ~ sum(.)),
    fmt = ~ fmt_number(., decimals = 0)
  ) |>
  tab_source_note(
    source_note = glue("Seed: {seed_code} | Pathway: {pathway_label}. Final dataset: {nrow(df_scenarios_final)} rows x {ncol(df_scenarios_final)} columns. Quantiles not included at national level.")
  ) |>
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12)
  )

# Clean up
rm(required_vars, df_validation)
```

---

# Phase F: Export

Export the final national scenarios dataset for use in Shiny dashboard and downstream aggregation scripts.

## F1. Data Quality Validation

Validate data integrity before export using comprehensive summary statistics. This step provides a complete overview of the final scenario dataset, ensuring no missing values exist and all metrics fall within expected ranges.

```{r}
#| label: f1-data-quality-validation

# =============================================================================
# F1. Data Quality Validation
# =============================================================================

# ---------------------------------------------------------------------------
# Generate skim summary
# ---------------------------------------------------------------------------

df_skim_numeric <- skimr::skim_without_charts(df_scenarios_final) |>
 filter(skim_type == "numeric") |>
 select(
    Variable = skim_variable,
    Missing = n_missing,
    Complete = complete_rate,
    Mean = numeric.mean,
    SD = numeric.sd,
    Min = numeric.p0,
    Q25 = numeric.p25,
    Median = numeric.p50,
    Q75 = numeric.p75,
    Max = numeric.p100
 )

# ---------------------------------------------------------------------------
# Table: Numeric Variables
# ---------------------------------------------------------------------------

df_skim_numeric |>
 gt() |>
 tab_header(
    title = "Numeric Variables Summary",
    subtitle = glue("Seed: {seed_code} | Pathway: {pathway_label} — {nrow(df_skim_numeric)} numeric variables in dataset")
 ) |>
 fmt_percent(columns = Complete, decimals = 1) |>
 fmt_number(columns = c(Mean, SD, Min, Q25, Median, Q75, Max), decimals = 2) |>
 tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(columns = Missing, rows = Missing == 0)
 ) |>
 tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = Missing, rows = Missing > 0)
 ) |>
 tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
 )

# Clean up
rm(df_skim_numeric)
```

## F2. Export National Scenarios

Export the complete scenario dataset containing 11 rows (one per penetration level) with all metrics required for the Shiny dashboard and Overview table.

```{r}
#| label: f2-export-national-scenarios

# =============================================================================
# F2. Export National Scenarios
# =============================================================================

write_parquet(
  df_scenarios_final,
  here("01_data", "02_processed",
       "scenarios", paste0(file_prefix, "_national_scenarios.parquet"))
)
```

---

# Notes

## Processing Summary

This script generates pre-computed national scenarios for the Shiny dashboard, simulating biofortification impact across 11 penetration levels (0% to 100% in 10% increments). The output dataset contains 11 rows (one per penetration level) with comprehensive metrics for economic impact, nutrient intake, and stunting outcomes at the Guatemala national level.

**Pipeline stages:**

1. **Farmer filtering**: Farmers with `p_entry <= p` are admitted to the adoption pool; each receives a per-farmer conversion ratio $r_i(p)$ via `apply_adoption_curve()`
2. **Economic aggregation**: Survey-weighted totals for adopting farmers, with scalable quantities (land, production, income, seed costs, subsidy flows) weighted by $r_i(p)$
3. **Coverage calculation**: Consumption-weighted national coverage from departmental saturation scenarios (Script `r paste0("06_01_", seed_suffix, "_", subsidy_suffix)`, includes gravity-based redistribution)
4. **Child assignment**: Hierarchical biofortification assignment based on `r if (is_pco) "household occupation" else "pre-existing priority level"` (departmental coverage targets maintained)
5. **Profile selection**: Pre-calculated baseline or QPM nutritional values selected per child based on biofortification flag (from Script `r if (is_pco) "04_08" else "05_06"`)
6. **Aggregation**: Survey-weighted national summaries for all outcome metrics

## Pre-Computed Profiles

Scenario simulation uses pre-computed baseline and QPM profiles from Script `r if (is_pco) "04_08" else "05_06"`. This design provides:

- **Computational efficiency**: Miller equation, iron absorption, z-score normalization, and GAM derivative calculations are resolved once upstream rather than repeated across the 10 scenario iterations
- **Consistency guarantee**: Baseline and QPM profiles are computed once and remain fixed across all scenarios
- **Focused simulation loop**: The `simulate_scenario_national()` wrapper operates directly on the pre-computed profiles, without the GAM model, normalization parameters, or biofortification factor constants
- **Lean dependencies**: The simulation runs without mgcv or gratia

The output schema matches the standard Shiny dashboard schema.

## Continuous Adoption Model

The per-farmer conversion ratio $r_i(p) = \text{sigmoid\_curve}(p, p_{\text{entry}}, \text{floor}_{\text{profile}}, k_{\text{profile}})$ determines the fraction of each farmer's economic activity that is realized at penetration $p$. The model parameters (per-profile floors and steepness coefficients) are loaded from the calibrated model configuration produced by Script 02_05. At the farmer's entry point, the conversion ratio equals the profile-specific floor; at full penetration ($p = 1$), all farmers have $r_i(1) = 1$ by construction.

Scalable economic quantities (land, production, income, seed costs, subsidy flows) are aggregated as $\sum_i \text{quantity}_i \times r_i(p) \times w_i$, where $w_i$ is the survey-calibrated weight. Farmer counts reflect pool membership: a farmer either has entered the pool ($p_{\text{entry}} \leq p$) or has not.

## Subsidy Variables

The seven monetary subsidy columns consumed in this script are computed at farmer level in Script 02_04 (Phase E) and propagated through the 02_06 parquet. No local subsidy calculations are performed in this script. The national aggregates preserve the linear identities established upstream: `cost_seeds_bio_no_subsidy_total = cost_seeds_bio_current_regime_total + subsidy_amount_absolute_total + subsidy_reduction_cost_total`, `subsidy_amount_absolute_total = newseed_subsidy_total + producer_absorption_total`, and `newseed_savings_vs_current_total = producer_absorption_total + subsidy_reduction_cost_total`.

## Seed Variant and Subsidy Regime

This script processes the **Seed `r params$seed_year` (`r seed_code`)** scenario under the **`r subsidy_label`** subsidy regime and the **`r tolower(pathway_label)` pathway**. The scenario simulation methodology is identical across all 18 combinations (3 seeds x 3 subsidy regimes x 2 pathways). Differences in economic metrics originate from the upstream yield factors and seed price structure applied in Script `r paste0("02_04_", seed_suffix, "_", subsidy_suffix)`, which propagate through the adoption model (`r paste0("02_06_", seed_suffix, "_", subsidy_suffix)`) and market baseline (`r paste0("06_01_", seed_suffix, "_", subsidy_suffix)`). Nutritional biofortification factors (iron x1.19, zinc x1.38, protein x1.94) are properties of QPM maize and do not vary between combinations.

## Output Schema

The exported dataset contains 11 rows with variables organized identically to the departmental dataset (`r paste0("06_02_", seed_suffix, "_", subsidy_suffix, "_", pathway)`) for compatibility with downstream processing, with `region = "Guatemala Total"` instead of `departamento`. Quantiles are not included at national level -- the Shiny dashboard uses departmental quantiles from `r paste0("06_02_", seed_suffix, "_", subsidy_suffix, "_", pathway)` for distribution visualizations.

## Hierarchical Assignment Priority

Children are assigned biofortified maize consumption following a household-level hierarchy that reflects realistic distribution patterns. Priority is granted in descending order to: cereal-producing farm households, other farming households, households of agricultural workers, rural non-agricultural households, and urban households. Within each tier, assignment is randomized until the departmental coverage target is reached.

## Key Variables Created

### Economic Variables

| Variable | Description | Unit |
|----------|-------------|------|
| `n_farmers_bio` | Adopting farmers (weighted) | count |
| `land_bio_ha` | Biofortified land area | hectares |
| `income_increment_total` | Total additional income | GTQ |
| `income_per_farmer_total` | Avg income benefit per national farmer | GTQ |
| `income_per_farmer_adopter` | Avg income benefit per adopting farmer | GTQ |

### Nutritional Variables

| Variable | Description | Unit |
|----------|-------------|------|
| `mean_zinc` | Mean absorbed zinc (TAZ) | mg/day |
| `mean_iron` | Mean absorbed iron | mg/day |
| `mean_protein` | Mean PDCAAS-adjusted protein | g/day |
| `*_inad_pct` | Inadequacy prevalence | % |
| `*_inad_averted` | Cases of inadequacy averted | count |

### Stunting Variables

| Variable | Description | Unit |
|----------|-------------|------|
| `stunting_pct` | Stunting prevalence | % |
| `stunting_change_pp` | Reduction vs baseline | percentage points |
| `n_stunting_averted` | Cases of stunting averted | count |
| `mean_zlen` | Mean height-for-age z-score | SD units |

## Limitations

1. **Static adoption hierarchy**: Assignment priority is fixed and does not model dynamic market penetration patterns

2. **No within-household variation**: All children in a household receive the same biofortification status

3. **Uniform biofortification factors**: All biofortified maize assumed to have identical nutrient improvements regardless of variety or growing conditions

4. **Cross-sectional stunting model**: `r if (is_pco) "Predictions represent point-in-time associations, not longitudinal growth trajectories" else "The meta-analysis effect size (9% height growth velocity increase) is applied as a cross-sectional adjustment, not a longitudinal growth trajectory"`

5. **No uncertainty quantification**: Point estimates only; confidence intervals would require bootstrap or Monte Carlo methods (planned for Phase 2.0)

6. **National quantiles excluded**: Distribution visualizations rely on departmental-level quantiles from Script `r paste0("06_02_", seed_suffix, "_", subsidy_suffix, "_", pathway)`
