---
title: "Transfer Model: `r params$nutrient_label`"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
params:
  nutrient_id: "iron"
  nutrient_label: "Iron Intake"
  var_prefix: "fe"
  var_maize: "fe_individuo_dia_maiz"
  var_non_maize: "fe_individuo_dia_no_maiz"
  units: "mg/day"
  script_number: "03_02"
  model_filename: "03_02_iron_models.rds"
  outlier_quantile: 0.95
  outlier_label: "P95 + 1.5×IQR"
  distribution_fn: "determine_distribution"
  overview_text: "This script develops predictive models for iron intake (maize vs non-maize sources) using ENCOVI 2023 data."
---

# Overview

`r params$overview_text`

These models will be applied to SIVESNU 2018 children to estimate their nutrient profiles based on demographic and socioeconomic characteristics.

The workflow:

1.  Load modeling-ready data from the preparation script (cleaned predictors + outcomes)
2.  Create survey design with outlier removal
3.  Evaluate distribution of `r tolower(params$nutrient_label)` variables (maize and non-maize)
4.  Train full gamma regression models with polynomial terms where applicable
5.  Perform variable selection using Wald tests
6.  Train refined models with significant predictors only
7.  Evaluate model diagnostics (residuals, leverage, collinearity)
8.  Save final models for SIVESNU transfer

**Modeling approach**: Survey-weighted gamma regression with log link, suitable for positive continuous outcomes with right-skewed distributions.

# Setup

```{r}
#| label: setup

# Clean environment (preserve params injected by Quarto)
rm(list = setdiff(ls(), "params"))
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export (export() for RDS model output)
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  glue,               # String interpolation in table headers
  broom,              # augment() for model diagnostics
  survey,             # Complex survey methodology
  srvyr,              # Tidy wrapper for survey package
  see,                # Visual backend for performance diagnostics
  tidyverse           # General purpose data manipulation
)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter
drop   <- base::drop

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions, including the distribution helpers resolved at
# runtime through match.fun()
source(here("02_code", "00_functions", "00_functions.R"))

# --- Resolve parameters into local variables -------------------------------

nutrient_id      <- params$nutrient_id
nutrient_label   <- params$nutrient_label
var_prefix       <- params$var_prefix
var_maize        <- params$var_maize
var_non_maize    <- params$var_non_maize
units            <- params$units
model_filename   <- params$model_filename
# Reconstruct script_number from model_filename, whose non-numeric suffix
# keeps it a string through the Quarto execute_params round trip
script_number    <- sub("^([0-9]{2}_[0-9]{2})_.*$", "\\1", model_filename)
outlier_quantile <- params$outlier_quantile
outlier_label    <- params$outlier_label
distribution_fn  <- match.fun(params$distribution_fn)

# Ensure output directories exist
ensure_output_dir(here("01_data", "02_processed", "models"))
ensure_output_dir(here("01_data", "02_processed", "web_resources"))

```

# Phase A: Data Loading and Preparation

## A1. Load Modeling Data

```{r}
#| label: a1-load-data

# Load modeling-ready data
df_raw <- read_parquet(
  here("01_data", "02_processed",
       "transfer", "03_01_modeling_ready_data.parquet")
)

# Define predictor variables for full model
# Polynomial terms are applied to continuous predictors; nominal predictors
# enter as factors, contributing one coefficient per level.
predictor_variables <- c(
  "parentesco", "sexo", "poly(edad, 2)",
  "departamento", "area", "poly(miembros_hogar, 3)",
  "propiedad", "poly(grado_estudios_hogar, 3)", "tipo_vivienda",
  "material_paredes", "material_techo", "material_piso",
  "poly(n_cuartos, 3)", "tipo_sanitario", "fuente_agua",
  "recoleccion_basura", "electricidad", "televisor",
  "telefonia_celular", "computadora"
)
```

## A2. Handle Outliers

Remove extreme values that could distort model training. Zero values cannot be modeled with gamma distribution (log link requires positive values).

**Outlier threshold:** Upper outliers are identified as values exceeding `r outlier_label`. `r if (outlier_quantile == 0.95) "The threshold suits micronutrients, whose dietary intake distributions are markedly right-skewed and where biologically plausible high values arise from consumption of micronutrient-dense or fortified foods. Retaining the upper tail supports model generalizability to populations with varied dietary patterns." else "The threshold suits macronutrients and amino acids, whose distributions are more stable and where extreme values are consistent with measurement error."`

```{r}
#| label: a2-remove-outliers

# Remove outliers and invalid values for modeling
df_clean <- df_raw |>
  # Remove records with missing outcomes
  drop_na(all_of(c(var_maize, var_non_maize))) |>
  # Remove zero values (incompatible with gamma distribution)
  filter(
    .data[[var_maize]] > 0,
    .data[[var_non_maize]] > 0
  ) |>
  # Remove upper outliers
  filter(
    .data[[var_maize]] <= quantile(df_raw[[var_maize]], outlier_quantile, na.rm = TRUE) +
      1.5 * IQR(df_raw[[var_maize]], na.rm = TRUE),
    .data[[var_non_maize]] <= quantile(df_raw[[var_non_maize]], outlier_quantile, na.rm = TRUE) +
      1.5 * IQR(df_raw[[var_non_maize]], na.rm = TRUE)
  )
```


## A3. Create Survey Design

Establish the survey design incorporating sampling weights and clustering structure for population-level inference.

```{r}
#| label: a3-create-survey-design
 
# Create the survey design
survey_design <- df_clean |>
  # Consolidate nominal levels before modeling;
  # "Otro" is retired from tipo_vivienda because the level carries no observations
  mutate(
    material_piso = fct_recode(material_piso, "Madera" = "Parqué"),
    tipo_vivienda = fct_drop(tipo_vivienda, only = "Otro")
  ) |>
  # Select variables for modeling
  transmute(
    # Survey design variables
    upm, factor,
    # Personal characteristics
    parentesco, sexo, edad,
    # Geographic characteristics
    departamento, area,
    # Household characteristics
    miembros_hogar, propiedad, grado_estudios_hogar,
    # Housing characteristics
    tipo_vivienda, material_paredes, material_techo, material_piso,
    n_cuartos, tipo_sanitario,
    # Basic services
    fuente_agua, recoleccion_basura,
    # Household assets
    electricidad, televisor, telefonia_celular, computadora,
    # Outcome variables
    !!sym(var_maize) := .data[[var_maize]],
    !!sym(var_non_maize) := .data[[var_non_maize]]
  ) |>
  # Create the survey design
  as_survey_design(
    ids = upm,              # Primary sampling units (clusters)
    weights = factor,       # Expansion factors
    strata = NULL           # No stratification in ENCOVI 2023
  )
```

---

# Phase B: Model `r nutrient_label` from Maize

## B1. Evaluate Distribution

Assess the distribution of `r tolower(nutrient_label)` from maize to select appropriate regression family.

```{r}
#| label: fig-b1-evaluate-distribution-maize
#| fig-cap: !expr paste0("Distribution assessment for ", tolower(params$nutrient_label), " from maize")
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

# Evaluate distribution fit for nutrient from maize
df_b1_dist_maize <- survey_design$variables[[var_maize]]
export(df_b1_dist_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_fig_b1_evaluate_distribution_maize.rds")))
distribution_fn(df_b1_dist_maize)

# Cleanup auxiliary objects
rm(df_b1_dist_maize)
```

**Decision**: The gamma distribution is selected for `r tolower(nutrient_label)` data (positive, right-skewed, continuous). Modeling proceeds with gamma regression and log link.

## B2. Train Full Model

Train initial model with all predictor variables using survey-weighted gamma regression.

```{r}
#| label: b2-train-full-model-maize
#| warning: FALSE

# Train full gamma regression model for nutrient from maize
results_maize_full <- train_gamma_model(
  formula = as.formula(paste(var_maize, "~", paste(predictor_variables, collapse = " + "))),
  model_label = paste0(nutrient_label, " from Maize (Full Model)"),
  x_label = paste0("Observed ", nutrient_label, " from Maize (", units, ")"),
  y_label = paste0("Predicted ", nutrient_label, " from Maize (", units, ")"),
  design = survey_design
)
```

```{r}
#| label: tbl-b2-coefficients-full-model-maize
#| tbl-cap: !expr paste0("Model coefficients for ", tolower(params$nutrient_label), " from maize (full model)")

# Display model coefficients
df_b2_coef_maize <- results_maize_full$coefficients
export(df_b2_coef_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_tbl_b2_coefficients_full_model_maize.rds")))
df_b2_coef_maize

# Cleanup auxiliary objects
rm(df_b2_coef_maize)
```

```{r}
#| label: fig-b2-diagnostic-plot-full-model-maize
#| fig-cap: !expr paste0("Observed vs predicted ", tolower(params$nutrient_label), " from maize (full model)")

# Display diagnostic scatter plot
df_b2_diag_maize <- augment(results_maize_full$model)
export(df_b2_diag_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_fig_b2_diagnostic_plot_full_model_maize.rds")))
results_maize_full$plot

# Cleanup auxiliary objects
rm(df_b2_diag_maize)
```

## B3. Variable Selection (Wald Tests)

Perform Wald tests to identify statistically significant predictors and remove non-significant variables to improve model parsimony.

```{r}
#| label: tbl-b3-wald-test-maize
#| tbl-cap: !expr paste0("Wald test results for predictor significance (", tolower(params$nutrient_label), " from maize)")

# Perform Wald test for each predictor
df_b3_wald_maize <- map_dfr(
  predictor_variables,
  apply_wald_test,
  model = results_maize_full$model
)
export(df_b3_wald_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_tbl_b3_wald_test_maize.rds")))
df_b3_wald_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = paste("Wald F-test results for", tolower(nutrient_label), "from maize model")
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

**Decision**: Variables with p-value < 0.05 retained for refined model.

## B4. Train Refined Model

```{r}
#| label: b4-train-refined-model-maize

# Automatically select significant predictors from Wald tests
predictors_maize <- df_b3_wald_maize |>
  filter(significant == "*") |>
  pull(variable)

# Train refined gamma regression model
results_maize_refined <- train_gamma_model(
  formula = as.formula(paste(var_maize, "~", paste(predictors_maize, collapse = " + "))),
  model_label = paste0(nutrient_label, " from Maize (Refined Model)"),
  x_label = paste0("Observed ", nutrient_label, " from Maize (", units, ")"),
  y_label = paste0("Predicted ", nutrient_label, " from Maize (", units, ")"),
  design = survey_design
)
```

```{r}
#| label: tbl-b4-coefficients-refined-model-maize
#| tbl-cap: !expr paste0("Model coefficients for ", tolower(params$nutrient_label), " from maize (refined model)")

# Display refined model coefficients
df_b4_coef_maize <- results_maize_refined$coefficients
export(df_b4_coef_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_tbl_b4_coefficients_refined_model_maize.rds")))
df_b4_coef_maize

# Cleanup auxiliary objects
rm(df_b4_coef_maize)
```

```{r}
#| label: fig-b4-diagnostic-plot-refined-model-maize
#| fig-cap: !expr paste0("Observed vs predicted ", tolower(params$nutrient_label), " from maize (refined model)")

# Display diagnostic scatter plot
df_b4_diag_maize <- augment(results_maize_refined$model)
export(df_b4_diag_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_fig_b4_diagnostic_plot_refined_model_maize.rds")))
results_maize_refined$plot

# Cleanup auxiliary objects
rm(df_b4_diag_maize)
```

## B5. Model Diagnostics

Evaluate model assumptions: linearity of residuals, homoscedasticity, influential observations, and multicollinearity.

```{r}
#| label: fig-b5-evaluate-diagnostics-maize
#| fig-cap: !expr paste0("Diagnostic plots for ", tolower(params$nutrient_label), " from maize model")
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

# Generate diagnostic plots
df_b5_diag_maize <- augment(results_maize_refined$model)
export(df_b5_diag_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_fig_b5_evaluate_diagnostics_maize.rds")))
evaluate_gamma_model(results_maize_refined$model)

# Cleanup auxiliary objects
rm(df_b5_diag_maize)
```

---

# Phase C: Model `r nutrient_label` from Non-Maize Sources


Model `r tolower(nutrient_label)` from non-maize foods (vegetables, fruits, animal products, legumes, etc.).

## C1. Evaluate Distribution

```{r}
#| label: fig-c1-evaluate-distribution-non-maize
#| fig-cap: !expr paste0("Distribution assessment for ", tolower(params$nutrient_label), " from non-maize sources")
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

# Evaluate distribution fit for nutrient from non-maize sources
df_c1_dist_non_maize <- survey_design$variables[[var_non_maize]]
export(df_c1_dist_non_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_fig_c1_evaluate_distribution_non_maize.rds")))
distribution_fn(df_c1_dist_non_maize)

# Cleanup auxiliary objects
rm(df_c1_dist_non_maize)
```

**Decision**: The gamma distribution is selected for `r tolower(nutrient_label)` from non-maize sources. Modeling proceeds with gamma regression and log link.

## C2. Train Full Model

```{r}
#| label: c2-train-full-model-non-maize

# Train full gamma regression model for nutrient from non-maize sources
results_non_maize_full <- train_gamma_model(
  formula = as.formula(paste(var_non_maize, "~", paste(predictor_variables, collapse = " + "))),
  model_label = paste0(nutrient_label, " from Non-Maize Sources (Full Model)"),
  x_label = paste0("Observed ", nutrient_label, " from Non-Maize (", units, ")"),
  y_label = paste0("Predicted ", nutrient_label, " from Non-Maize (", units, ")"),
  design = survey_design
)
```

```{r}
#| label: tbl-c2-coefficients-full-model-non-maize
#| tbl-cap: !expr paste0("Model coefficients for ", tolower(params$nutrient_label), " from non-maize sources (full model)")

# Display model coefficients
df_c2_coef_non_maize <- results_non_maize_full$coefficients
export(df_c2_coef_non_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_tbl_c2_coefficients_full_model_non_maize.rds")))
df_c2_coef_non_maize

# Cleanup auxiliary objects
rm(df_c2_coef_non_maize)
```

```{r}
#| label: fig-c2-diagnostic-plot-full-model-non-maize
#| fig-cap: !expr paste0("Observed vs predicted ", tolower(params$nutrient_label), " from non-maize sources (full model)")

# Display diagnostic scatter plot
df_c2_diag_non_maize <- augment(results_non_maize_full$model)
export(df_c2_diag_non_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_fig_c2_diagnostic_plot_full_model_non_maize.rds")))
results_non_maize_full$plot

# Cleanup auxiliary objects
rm(df_c2_diag_non_maize)
```

## C3. Variable Selection (Wald Tests)

```{r}
#| label: tbl-c3-wald-test-non-maize
#| tbl-cap: !expr paste0("Wald test results for predictor significance (", tolower(params$nutrient_label), " from non-maize)")

# Perform Wald test for each predictor
df_c3_wald_non_maize <- map_dfr(
  predictor_variables,
  apply_wald_test,
  model = results_non_maize_full$model
)
export(df_c3_wald_non_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_tbl_c3_wald_test_non_maize.rds")))
df_c3_wald_non_maize |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = paste("Wald F-test results for", tolower(nutrient_label), "from non-maize model")
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

**Decision**: Retain only statistically significant predictors (p < 0.05) for refined model.

## C4. Train Refined Model

```{r}
#| label: c4-train-refined-model-non-maize

# Automatically select significant predictors from Wald tests
predictors_non_maize <- df_c3_wald_non_maize |>
  filter(significant == "*") |>
  pull(variable)

# Train refined gamma regression model
results_non_maize_refined <- train_gamma_model(
  formula = as.formula(paste(var_non_maize, "~", paste(predictors_non_maize, collapse = " + "))),
  model_label = paste0(nutrient_label, " from Non-Maize Sources (Refined Model)"),
  x_label = paste0("Observed ", nutrient_label, " from Non-Maize (", units, ")"),
  y_label = paste0("Predicted ", nutrient_label, " from Non-Maize (", units, ")"),
  design = survey_design
)
```

```{r}
#| label: tbl-c4-coefficients-refined-model-non-maize
#| tbl-cap: !expr paste0("Model coefficients for ", tolower(params$nutrient_label), " from non-maize sources (refined model)")

# Display refined model coefficients
df_c4_coef_non_maize <- results_non_maize_refined$coefficients
export(df_c4_coef_non_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_tbl_c4_coefficients_refined_model_non_maize.rds")))
df_c4_coef_non_maize

# Cleanup auxiliary objects
rm(df_c4_coef_non_maize)
```

```{r}
#| label: fig-c4-diagnostic-plot-refined-model-non-maize
#| fig-cap: !expr paste0("Observed vs predicted ", tolower(params$nutrient_label), " from non-maize sources (refined model)")

# Display diagnostic scatter plot
df_c4_diag_non_maize <- augment(results_non_maize_refined$model)
export(df_c4_diag_non_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_fig_c4_diagnostic_plot_refined_model_non_maize.rds")))
results_non_maize_refined$plot

# Cleanup auxiliary objects
rm(df_c4_diag_non_maize)
```

## C5. Model Diagnostics

```{r}
#| label: fig-c5-evaluate-diagnostics-non-maize
#| fig-cap: !expr paste0("Diagnostic plots for ", tolower(params$nutrient_label), " from non-maize sources model")
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

# Generate diagnostic plots
df_c5_diag_non_maize <- augment(results_non_maize_refined$model)
export(df_c5_diag_non_maize, here("01_data", "02_processed", "web_resources", paste0(script_number, "_fig_c5_evaluate_diagnostics_non_maize.rds")))
evaluate_gamma_model(results_non_maize_refined$model)

# Cleanup auxiliary objects
rm(df_c5_diag_non_maize)
```

---

# Phase D: Export

Save final refined models for application to SIVESNU data in subsequent transfer scripts.

```{r}
#| label: d-save-models

# Extract final models
model_maize     <- results_maize_refined$model
model_non_maize <- results_non_maize_refined$model

# Build named list preserving original naming convention
model_list <- setNames(
  list(model_maize, model_non_maize),
  c(paste0("model_", var_prefix, "_maize"),
    paste0("model_", var_prefix, "_non_maize"))
)

# Save models as RDS
export(
  model_list,
  here("01_data", "02_processed",
       "models", model_filename)
)

# Clean auxiliary objects
rm(results_maize_full, results_maize_refined,
   results_non_maize_full, results_non_maize_refined,
   model_maize, model_non_maize)
```

# Notes

## Model Performance

**`r nutrient_label` from Maize:**

- Final predictors: `r length(predictors_maize)` variables
- Significant predictors: `r paste(predictors_maize, collapse = ", ")`
- Model diagnostics: Assessed via residual plots, leverage analysis, and collinearity checks above

**`r nutrient_label` from Non-Maize:**

- Final predictors: `r length(predictors_non_maize)` variables
- Significant predictors: `r paste(predictors_non_maize, collapse = ", ")`
- Model diagnostics: Assessed via residual plots, leverage analysis, and collinearity checks above
