---
title: "Prepare Nutrient Transfer Modeling Data"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script prepares the ENCOVI 2023 data for building nutrient transfer models that will predict nutrient intake profiles in SIVESNU 2018 children. The workflow includes:

1. Load individual-level nutrient intake data from Module 1
2. Select relevant predictor variables (demographic, socioeconomic, geographic, household)
3. Evaluate predictor correlations to identify multicollinearity issues
4. Identify and remove near-zero variance variables
5. Apply Yeo-Johnson transformations to normalize continuous predictors
6. Save prepared dataset and transformation objects for modeling scripts

**Purpose**: Transfer models predict nutrient intake (iron, zinc, protein, amino acids, plant percentage and energy) in SIVESNU children based on demographic/socioeconomic profiles, enabling integration of nutrient data with anthropometric measurements.

# Setup

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  car,                # Box-Cox/Yeo-Johnson analysis
  gt,                 # Table formatting and presentation
  tidyverse           # General purpose data manipulation
)

# Packages used via :: (not loaded to keep namespace clean):
#   missRanger    - missRanger() via impute_rf_parallel() for RF-based imputation
#   ggcorrplot    - ggcorrplot() for correlation matrix visualization
#   bestNormalize - yeojohnson() for variable normalization

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions
source(here("02_code", "00_functions", "00_functions.R"))

# Ensure output directories exist
ensure_output_dir(here("01_data", "02_processed", "transfer"))
ensure_output_dir(here("01_data", "02_processed", "models"))
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

# Phase A: Data Loading and Variable Selection

## A1. Load Nutrient Intake Data

```{r}
#| label: a1-load-data

# Load individual nutrient intake data from Module 1
df_raw <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_04_individual_nutritional_intake_complete.parquet")
)
```

## A2. Select Predictor Variables

Select relevant predictor variables and impute missing values for modeling.

```{r}
#| label: a2-select-variables

# Select predictor variables and outcome nutrients
df_selected <- df_raw |>
  select(
    # Survey design
    upm, factor,
    # Personal characteristics
    parentesco, sexo, edad,
    # Geographic characteristics
    departamento, area,
    # Household characteristics
    miembros_hogar, grado_estudios_hogar, propiedad,
    # Housing characteristics
    tipo_vivienda, material_paredes, material_techo, material_piso, 
    n_cuartos, n_dormitorios, cocina, tipo_sanitario,
    # Basic services
    fuente_agua, recoleccion_basura,
    # Household assets
    electricidad, televisor, telefonia_fija, telefonia_celular, computadora,
    # Outcome nutrients (maize vs non-maize)
    fe_individuo_dia_maiz, fe_individuo_dia_no_maiz,
    zn_individuo_dia_maiz, zn_individuo_dia_no_maiz,
    prot_individuo_dia_maiz, prot_individuo_dia_no_maiz,
    prot_apro_individuo_dia_maiz, prot_apro_individuo_dia_no_maiz,
    lys_individuo_dia_maiz, lys_individuo_dia_no_maiz,
    trp_individuo_dia_maiz, trp_individuo_dia_no_maiz,
    kcal_individuo_dia_maiz, kcal_individuo_dia_no_maiz,
    porcentaje_vegetales
  )
```

---

# Phase B: Missing Value Treatment

## B1. Diagnose Missing Values

Model training requires a complete predictor matrix. The pipeline applies a Random Forest-based multivariate imputation procedure, which:

- Preserves correlations between variables
- Handles mixed data types (categorical + continuous)
- Draws each imputed value from the predictive relationships in the data
- Provides out-of-bag error estimates for imputation quality

Implementation uses the `impute_rf_parallel()` utility function, which wraps `missRanger` with a multi-threaded `ranger` engine and returns imputed values alongside out-of-bag error estimates (OOBE for continuous variables, PFC for categorical).

```{r}
#| label: tbl-b1-diagnose-missing
#| tbl-cap: "Missing value summary for predictor variables"

# Calculate missing value statistics
missing_summary <- df_selected |>
  select(-c(upm, factor, starts_with("fe_"), starts_with("zn_"), 
            starts_with("prot_"), starts_with("lys_"), starts_with("trp_"),
            starts_with("kcal_"), porcentaje_vegetales)) |>
  summarise(across(everything(), 
                   list(n_missing = ~sum(is.na(.)),
                        pct_missing = ~mean(is.na(.)) * 100))) |>
  pivot_longer(everything(), 
               names_to = c("variable", ".value"),
               names_pattern = "(.+)_(n_missing|pct_missing)") |>
  filter(n_missing > 0) |>
  arrange(desc(pct_missing))

# Export data for web resources
export(
  missing_summary,
  here("01_data", "02_processed", "web_resources", "03_01_tbl_b1_diagnose_missing.rds")
)

# Format as GT table
missing_summary |>
  gt() |>
  tab_header(
    title = "Missing Values in Predictor Variables",
    subtitle = "Variables requiring imputation before modeling"
  ) |>
  cols_label(
    variable = "Variable",
    n_missing = "N Missing",
    pct_missing = "% Missing"
  ) |>
  fmt_number(
    columns = n_missing,
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    columns = pct_missing,
    decimals = 2,
    pattern = "{x}%"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Clean auxiliary objects
rm(missing_summary)
```

## B2. Random Forest Imputation

```{r}
#| label: tbl-b2-impute-missing
#| tbl-cap: "Random Forest imputation quality metrics"

# Separate predictors from outcomes for imputation
predictors_to_impute <- df_selected |>
  select(
    parentesco, sexo, edad, departamento, area,
    miembros_hogar, grado_estudios_hogar, propiedad,
    tipo_vivienda, material_paredes, material_techo, material_piso,
    n_cuartos, n_dormitorios, cocina, tipo_sanitario,
    fuente_agua, recoleccion_basura, electricidad,
    televisor, telefonia_fija, telefonia_celular, computadora
  ) |>
  # Convert ordered factors to regular factors for ranger compatibility
  mutate(departamento = factor(departamento, ordered = FALSE)) |>
  # Convert to data frame
  as.data.frame()

# --- Run Random Forest imputation ---
set.seed(789)  # Azahar standard seed for reproducibility

imputed_result <- impute_rf_parallel(
  data = predictors_to_impute,
  maxiter = 10,
  ntree = 100,
  verbose = FALSE
)

# Extract imputed data
predictors_imputed <- imputed_result$ximp

# Create quality metrics summary table
# Report each metric only when a variable of that type required imputation
n_continuous <- imputed_result$n_continuous
n_categorical <- imputed_result$n_categorical

imputation_quality <- tibble(
  metric = c(
    "OOBE (Continuous Variables)",
    "PFC (Categorical Variables)",
    "Number of Iterations",
    "Trees per Iteration"
  ),
  value = c(
    if (n_continuous > 0) imputed_result$OOBerror["OOBE"] else NA_real_,
    if (n_categorical > 0) imputed_result$OOBerror["PFC"] else NA_real_,
    10,
    100
  ),
  interpretation = c(
    if (n_continuous > 0) {
      if_else(imputed_result$OOBerror["OOBE"] < 0.5, "Acceptable", "Review required")
    } else {
      "No continuous variables required imputation"
    },
    if (n_categorical > 0) {
      if_else(imputed_result$OOBerror["PFC"] < 0.2, "Acceptable", "Review required")
    } else {
      "No categorical variables required imputation"
    },
    "Fixed parameter",
    "Fixed parameter"
  )
)

# Export data for web resources
export(
  imputation_quality,
  here("01_data", "02_processed", "web_resources", "03_01_tbl_b2_impute_missing.rds")
)

imputation_quality |>
  gt() |>
  tab_header(
    title = "Missing Value Imputation Quality Assessment",
    subtitle = "Out-of-bag error estimates from Random Forest imputation"
  ) |>
  cols_label(
    metric = "Metric",
    value = "Value",
    interpretation = "Quality Assessment"
  ) |>
  fmt_number(
    columns = value,
    decimals = 4
  ) |>
  sub_missing(
    columns = value,
    missing_text = "N/A"
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "Acceptable"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "Review required"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#D1ECF1"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "No continuous variables required imputation"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#D1ECF1"),
    locations = cells_body(
      columns = interpretation,
      rows = interpretation == "No categorical variables required imputation"
    )
  ) |>
  tab_footnote(
    footnote = "OOBE: out-of-bag error across continuous variables, the square root of the variance-weighted mean of per-variable unexplained variance. Values < 0.5 indicate acceptable imputation quality. N/A when no continuous variables had missing values.",
    locations = cells_body(columns = metric, rows = 1)
  ) |>
  tab_footnote(
    footnote = "PFC: Proportion of Falsely Classified entries. Values < 0.2 indicate acceptable imputation quality for categorical variables. N/A when no categorical variables had missing values.",
    locations = cells_body(columns = metric, rows = 2)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Clean auxiliary objects from quality summary (keep predictors_imputed
# and imputed_result alive; they are consumed by B2b below).
rm(imputation_quality, n_continuous, n_categorical)
```

## B2b. Completion of Residual Missing Values

The Random Forest imputation engine can leave a residual `NaN` for a small subset of rows when the out-of-bag prediction collapses inside a terminal node with zero within-leaf variance. This affects continuous discrete-scale predictors.

Downstream analyses (correlation, near-zero variance, modeling) operate on a complete predictor matrix, so residual missing values are completed with a variable-specific rule:

- **Numeric / integer variables**: median of the observed values, which preserves the support of discrete-scale ordinal variables (e.g. years of schooling take values such as 1, 1.5, 2, 2.5, ...).
- **Factor variables**: mode of the observed values.

The count of values completed by each rule is reported in the table below.

```{r}
#| label: tbl-b2b-completion-summary
#| tbl-cap: "Residual missing value completion after Random Forest imputation"

# --- Identify residual missing values ---
residual_summary <- predictors_imputed |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n_completed") |>
  filter(n_completed > 0) |>
  mutate(
    rule = vapply(
      variable,
      function(v) {
        x <- predictors_imputed[[v]]
        if (is.numeric(x)) "Median" else "Mode"
      },
      character(1)
    )
  ) |>
  arrange(desc(n_completed))

# --- Apply completion rules ---
mode_value <- function(x) {
  observed <- x[!is.na(x)]
  tab <- table(observed)
  names(tab)[which.max(tab)]
}

predictors_complete <- predictors_imputed |>
  mutate(across(
    where(is.numeric),
    ~ if_else(is.na(.), median(., na.rm = TRUE), .)
  )) |>
  mutate(across(
    where(is.factor),
    ~ {
      mv <- mode_value(.)
      replace(., is.na(.), mv)
    }
  ))

# --- Build summary with post-completion integrity row -----------------------
n_residual_after <- sum(is.na(predictors_complete))

completion_report <- residual_summary |>
  bind_rows(
    tibble(
      variable    = "— Residual NAs after completion —",
      n_completed = n_residual_after,
      rule        = if_else(n_residual_after == 0L,
                            "Integrity check: passed",
                            "Integrity check: review required")
    )
  )

# --- Export summary for web resources ---
export(
  completion_report,
  here("01_data", "02_processed", "web_resources", "03_01_tbl_b2b_completion_summary.rds")
)

# --- Present summary as GT table ---
completion_report |>
  gt() |>
  tab_header(
    title = "Residual Missing Value Completion",
    subtitle = "Values completed by variable-specific rule after RF imputation"
  ) |>
  cols_label(
    variable = "Variable",
    n_completed = "N Completed",
    rule = "Completion Rule"
  ) |>
  fmt_number(
    columns = n_completed,
    decimals = 0,
    use_seps = TRUE
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      rows = rule == "Integrity check: passed"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#F8D7DA"),
    locations = cells_body(
      rows = rule == "Integrity check: review required"
    )
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Combine complete predictors with outcomes and survey design
df <- df_selected |>
  select(upm, factor, starts_with("fe_"), starts_with("zn_"), 
         starts_with("prot_"), starts_with("lys_"), starts_with("trp_"),
         starts_with("kcal_"), porcentaje_vegetales) |>
  bind_cols(predictors_complete)

# Clean auxiliary objects
rm(
  predictors_to_impute, imputed_result,
  predictors_imputed, predictors_complete,
  residual_summary, completion_report, n_residual_after
)
```

---

# Phase C: Variable Quality Assessment

## C1. Evaluate Predictor Correlation

Assess correlation between predictor variables to identify potential multicollinearity issues.

```{r}
#| label: fig-c1-correlation-analysis
#| fig-cap: "Correlation matrix of predictor variables"
#| fig-width: 10
#| fig-height: 10
#| warning: false
  
# Calculate correlation matrix for all predictors
cor_matrix <- df |>
  mutate(factor = as.integer(factor)) |>
  uncount(factor) |>
  transmute(across(
    c(
      parentesco, sexo, edad, departamento, area, miembros_hogar,
      grado_estudios_hogar, propiedad, tipo_vivienda, material_paredes,
      material_techo, material_piso, n_cuartos, n_dormitorios, cocina,
      tipo_sanitario, fuente_agua, recoleccion_basura, electricidad,
      televisor, telefonia_fija, telefonia_celular, computadora
    ),
    as.numeric
  )) |>
  cor() |>
  round(1)

# Export data for web resources
export(
  cor_matrix,
  here("01_data", "02_processed", "web_resources", "03_01_fig_c1_correlation_analysis.rds")
)

cor_matrix |>
  ggcorrplot::ggcorrplot(
    hc.order = TRUE,
    colors = c("#E46726", "white", "#6D9EC1"),
    type = "lower",
    lab = TRUE
  )

# Cleanup auxiliary objects
rm(cor_matrix)
```

**Decision**: Remove `n_dormitorios` (number of bedrooms) due to high correlation with `n_cuartos` (number of rooms). The two variables carry redundant information (r > 0.8), and the models retain `n_cuartos`.

## C2. Identify Near-Zero Variance Predictors

Evaluate variables with near-zero variance that provide minimal information for modeling. Variables with extremely unbalanced distributions or very low variability should be removed to improve model stability.

```{r}
#| label: tbl-c2-near-zero-variance
#| tbl-cap: "Near-zero variance analysis for predictor variables"

# Analyze variance for all predictors (exclude survey design and outcomes)
nzv_data <- df |>
  select(-c(upm, factor, starts_with("fe_"), starts_with("zn_"),
            starts_with("prot_"), starts_with("lys_"), starts_with("trp_"),
            starts_with("kcal_"), porcentaje_vegetales))

# Export data for web resources
export(
  nzv_data,
  here("01_data", "02_processed", "web_resources", "03_01_tbl_c2_near_zero_variance.rds")
)

nzv_data |>
  analyze_variance()

# Cleanup auxiliary objects
rm(nzv_data)
```

**Decision**: Remove `cocina` (kitchen type) and `telefonia_fija` (landline telephone) due to near-zero variance. Both variables show extremely unbalanced distributions (one category dominates >95% of observations), providing minimal predictive information while potentially causing model instability.

---

# Phase D: Variable Transformations

Evaluate normality of continuous predictors and apply Yeo-Johnson transformation where needed. The transformation handles both positive and negative values.

**Variables to evaluate:**
- `miembros_hogar` (household members)
- `n_cuartos` (number of rooms)

## D1. Evaluate Yeo-Johnson

```{r}
#| label: fig-d1-evaluate-yj-miembros-hogar
#| fig-cap: "Yeo-Johnson transformation profile for household members"

# Export data for web resources
export(
  tibble(miembros_hogar = subset(df$miembros_hogar, !is.na(df$miembros_hogar))),
  here("01_data", "02_processed", "web_resources", "03_01_fig_d1_evaluate_yj_miembros_hogar.rds")
)

# Evaluate optimal lambda for Yeo-Johnson transformation
boxCox(
  lm(subset(df$miembros_hogar, !is.na(df$miembros_hogar)) ~ 1),
  main = "Household Members - Transformation Profile",
  family = "yjPower"
)
```

```{r}
#| label: fig-d1-evaluate-yj-n-cuartos
#| fig-cap: "Yeo-Johnson transformation profile for number of rooms"

# Export data for web resources
export(
  tibble(n_cuartos = subset(df$n_cuartos, !is.na(df$n_cuartos))),
  here("01_data", "02_processed", "web_resources", "03_01_fig_d1_evaluate_yj_n_cuartos.rds")
)

# Evaluate optimal lambda for Yeo-Johnson transformation
boxCox(
  lm(subset(df$n_cuartos, !is.na(df$n_cuartos)) ~ 1),
  main = "Number of Rooms - Transformation Profile",
  family = "yjPower"
)
```

**Decision**: Both variables show optimal lambda values near 0 (indicating logarithmic-type transformation needed). Yeo-Johnson transformation automatically handles this, improving normality for modeling. 

## D2. Apply Yeo-Johnson Transformations

```{r}
#| label: d2-apply-yj-transformation

# Create Yeo-Johnson transformation objects
yj_miembros_hogar <- bestNormalize::yeojohnson(df$miembros_hogar)
yj_n_cuartos <- bestNormalize::yeojohnson(df$n_cuartos)
```

Verify transformation effectiveness by re-evaluating normality of transformed variables.

```{r}
#| label: fig-d2-verify-yj-miembros-hogar
#| fig-cap: "Transformation profile for household members (post-transformation)"

# Export data for web resources
export(
  tibble(miembros_hogar_yj = predict(yj_miembros_hogar)),
  here("01_data", "02_processed", "web_resources", "03_01_fig_d2_verify_yj_miembros_hogar.rds")
)

# Verify transformation improved normality
boxCox(
  lm(predict(yj_miembros_hogar) ~ 1),
  main = "Household Members - Post-Transformation",
  family = "yjPower"
)
```

```{r}
#| label: fig-d2-verify-yj-n-cuartos
#| fig-cap: "Transformation profile for number of rooms (post-transformation)"

# Export data for web resources
export(
  tibble(n_cuartos_yj = predict(yj_n_cuartos)),
  here("01_data", "02_processed", "web_resources", "03_01_fig_d2_verify_yj_n_cuartos.rds")
)

# Verify transformation improved normality
boxCox(
  lm(predict(yj_n_cuartos) ~ 1),
  main = "Number of Rooms - Post-Transformation",
  family = "yjPower"
)
```

**Conclusion**: Post-transformation profiles show lambda ≈ 1.

---

# Phase E: Final Dataset and Export

## E1. Apply All Transformations

Apply all identified transformations (remove high-correlation variables, remove near-zero variance variables, apply Yeo-Johnson normalization) to create the final modeling-ready dataset.

```{r}
#| label: e1-create-final-dataset

# Apply all transformations to create final modeling dataset
df_modeling_ready <- df |>
  # Remove high-correlation variable
  select(-n_dormitorios) |>
  # Remove near-zero variance variables
  select(-c(cocina, telefonia_fija)) |>
  # Apply Yeo-Johnson transformations to continuous variables
  mutate(
    miembros_hogar = predict(yj_miembros_hogar),
    n_cuartos = predict(yj_n_cuartos)
  )
```

## E2. Save Outputs

Save the prepared dataset and transformation objects for use in subsequent modeling scripts.

```{r}
#| label: e2-save-outputs

# Save Yeo-Johnson transformation objects for SIVESNU application
export(
  list(
    yj_miembros_hogar = yj_miembros_hogar,
    yj_n_cuartos = yj_n_cuartos
  ),
  here("01_data", "02_processed", 
       "models", "03_01_yj_transforms.rds")
)

# Save modeling-ready dataset
write_parquet(
  df_modeling_ready,
  here("01_data", "02_processed", 
       "transfer", "03_01_modeling_ready_data.parquet")
)

# Clean auxiliary objects
rm(yj_miembros_hogar, yj_n_cuartos)
```

# Notes

## Variable Selection Decisions

**Removed due to high correlation:**
- `n_dormitorios` (bedrooms): Correlation r > 0.8 with `n_cuartos` (rooms), indicating redundancy

**Removed due to near-zero variance:**
- `cocina` (kitchen type): frequency ratio and percent unique reported in the C2 table
- `telefonia_fija` (landline phone): frequency ratio and percent unique reported in the C2 table

## Missing Value Imputation

Random Forest imputation preserves variable correlations and handles mixed
data types. Out-of-bag quality metrics are reported per run in the B2
diagnostic table, with acceptability thresholds of 0.5 for OOBE (continuous
variables) and 0.2 for PFC (categorical variables).

## Yeo-Johnson Transformations

Applied to normalize continuous predictors:
- `miembros_hogar`: Pre-transformation λ ≈ 0 (log-type), post-transformation λ ≈ 1 (normalized)
- `n_cuartos`: Pre-transformation λ ≈ 0 (log-type), post-transformation λ ≈ 1 (normalized)

Transformation objects saved for consistent application to SIVESNU data during transfer modeling.
