---
title: "Transfer Model: Plant-Based Food Percentage"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script develops a predictive model for the percentage of plant-based foods in individual diets using ENCOVI 2023 data. This model will be applied to SIVESNU 2018 children to estimate dietary composition based on demographic and socioeconomic characteristics.

The workflow:

1. Load modeling-ready data from the preparation script (cleaned predictors + outcomes)
2. Create survey design with boundary adjustment for percentage values
3. Evaluate distribution of plant food percentage variable
4. Train full gamma regression model with identity link
5. Perform variable selection using Wald tests
6. Train refined model with significant predictors only
7. Evaluate model diagnostics (residuals, leverage, collinearity)
8. Save final model for SIVESNU transfer

**Modeling approach**: Survey-weighted gamma regression with identity link, suitable for proportion data bounded between 0 and 1. Boundary values (0 and 1) are adjusted to 0.0001 and 0.9999 to satisfy gamma distribution requirements.

# Setup

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,              # Parquet file handling
  rio,                # Robust import/export for various formats
  here,               # Robust file path management
  gt,                 # Table formatting and presentation
  survey,             # Complex survey methodology
  srvyr,              # Tidy wrapper for survey package
  see,                # Visual backend for performance diagnostics
  tidyverse           # General purpose data manipulation
)

# Packages used via :: (not loaded to keep namespace clean):
#   broom - augment() for model diagnostic data frames

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter
drop   <- base::drop

# Set display options
options(scipen = 999, digits = 4)

# Load utility functions
source(here("02_code", "00_functions", "00_functions.R"))

# Ensure output directories exist
ensure_output_dir(here("01_data", "02_processed", "models"))
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

# Phase A: Data Loading and Preparation

## A1. Load Modeling Data

```{r}
#| label: a1-load-data

# Load modeling-ready data
df_raw <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "03_01_modeling_ready_data.parquet")
)

# Define predictor variables for full model
# Polynomial terms are applied to continuous predictors; nominal predictors
# enter as factors, contributing one coefficient per level.
predictor_variables <- c(
  "parentesco", "sexo", "poly(edad, 2)", 
  "departamento", "area", "poly(miembros_hogar, 3)", 
  "propiedad", "poly(grado_estudios_hogar, 3)", "tipo_vivienda",
  "material_paredes", "material_techo", "material_piso",
  "poly(n_cuartos, 3)", "tipo_sanitario", "fuente_agua",
  "recoleccion_basura", "electricidad", "televisor",
  "telefonia_celular", "computadora"
)
```

## A2. Adjust Boundary Values and Create Survey Design

Gamma distribution requires positive values strictly greater than 0 and less than 1 for percentage data. Boundary values (0 and 1) are adjusted to 0.0001 and 0.9999 to satisfy distribution requirements while maintaining interpretability.

```{r}
#| label: a2-create-survey-design
 
# Create the survey design with boundary adjustments
survey_design <- df_raw |>
  # Remove records with missing outcomes
  drop_na(porcentaje_vegetales) |>
  # Consolidate nominal levels before modeling;
  # "Otro" is retired from tipo_vivienda because the level carries no observations
  mutate(
    material_piso = fct_recode(material_piso, "Madera" = "Parqué"),
    tipo_vivienda = fct_drop(tipo_vivienda, only = "Otro")
  ) |>
  # Select variables for modeling
  transmute(
    # Survey design variables
    upm, factor,
    # Personal characteristics
    parentesco, sexo, edad,
    # Geographic characteristics
    departamento, area,
    # Household characteristics
    miembros_hogar, propiedad, grado_estudios_hogar,
    # Housing characteristics
    tipo_vivienda, material_paredes, material_techo, material_piso,
    n_cuartos, tipo_sanitario,
    # Basic services
    fuente_agua, recoleccion_basura,
    # Household assets
    electricidad, televisor, telefonia_celular, computadora,
    # Outcome variable with boundary adjustment
    porcentaje_vegetales = ifelse(porcentaje_vegetales == 1, 0.9999,
                                   ifelse(porcentaje_vegetales == 0, 0.0001,
                                          porcentaje_vegetales))
  ) |>
  # Create the survey design
  as_survey_design(
    ids = upm,              # Primary sampling units (clusters)
    weights = factor,       # Expansion factors
    strata = NULL           # No stratification in ENCOVI 2023
  )
```

---

# Phase B: Model Plant Food Percentage

## B1. Evaluate Distribution

Assess the distribution of plant food percentage to confirm appropriate regression family selection.

```{r}
#| label: fig-b1-evaluate-distribution-plant-percentage
#| fig-cap: "Distribution assessment for plant food percentage"
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

# Export data for web resources
export(
  tibble(porcentaje_vegetales = survey_design$variables$porcentaje_vegetales),
  here("01_data", "02_processed", "web_resources", "03_08_fig_b1_evaluate_distribution_plant_percentage.rds")
)

# Evaluate distribution fit for plant food percentage
determine_distribution(survey_design$variables$porcentaje_vegetales)
```

**Decision**: The gamma distribution is selected for plant food percentage data (bounded between 0 and 1, right-skewed). Modeling proceeds with gamma regression and identity link, which suits proportion data.

## B2. Train Full Model

Train initial model with all predictor variables using survey-weighted gamma regression with identity link.

```{r}
#| label: b2-train-full-model
#| warning: FALSE

# Train full gamma regression model with identity link
results_plant_full <- train_gamma_model_plant_foods(
  formula = as.formula(paste("porcentaje_vegetales ~", 
                             paste(predictor_variables, collapse = " + "))),
  model_label = "Plant Food Percentage (Full Model)",
  x_label = "Observed Plant Food Percentage",
  y_label = "Predicted Plant Food Percentage",
  design = survey_design
)
```

```{r}
#| label: tbl-b2-coefficients-full-model
#| tbl-cap: "Model coefficients for plant food percentage (full model)"

# Export data for web resources
export(
  broom::tidy(results_plant_full$model),
  here("01_data", "02_processed", "web_resources", "03_08_tbl_b2_coefficients_full_model.rds")
)

# Display full model coefficients
results_plant_full$coefficients
```

```{r}
#| label: fig-b2-diagnostic-plot-full-model
#| fig-cap: "Observed vs predicted plant food percentage (full model)"

# Export data for web resources
export(
  results_plant_full$results,
  here("01_data", "02_processed", "web_resources", "03_08_fig_b2_diagnostic_plot_full_model.rds")
)

# Display diagnostic scatter plot
results_plant_full$plot
```

## B3. Variable Selection (Wald Tests)

Perform Wald tests to identify significant predictors and reduce model complexity.

```{r}
#| label: tbl-b3-wald-tests
#| tbl-cap: "Wald test results for predictor significance"

# Perform Wald tests for all predictor variables
wald_results <- map_dfr(
  predictor_variables,
  apply_wald_test,
  model = results_plant_full$model
)

# Export data for web resources
export(
  wald_results,
  here("01_data", "02_processed", "web_resources", "03_08_tbl_b3_wald_tests.rds")
)

wald_results |>
  gt() |>
  tab_header(
    title = "Variable Significance Testing",
    subtitle = "Wald F-test results for Plant Food Percentage"
  ) |>
  fmt_number(
    columns = f_statistic,
    decimals = 2
  ) |>
  tab_style(
    style = cell_fill(color = "#D4EDDA"),
    locations = cells_body(
      columns = everything(),
      rows = significant == "*"
    )
  ) |>
  tab_options(
    table.font.size = 11,
    heading.title.font.size = 13,
    table.width = pct(100)
  )
```

Non-significant variables (p ≥ 0.05) are removed to create a more parsimonious model.

## B4. Train Refined Model

Train final model with only statistically significant predictors.

```{r}
#| label: b4-train-refined-model

# Automatically select significant predictors from Wald tests
predictors_plant_refined <- wald_results |>
  filter(significant == "*") |>
  pull(variable)

# Train refined gamma regression model with identity link
results_plant_refined <- train_gamma_model_plant_foods(
  formula = as.formula(paste("porcentaje_vegetales ~", 
                             paste(predictors_plant_refined, collapse = " + "))),
  model_label = "Plant Food Percentage (Refined Model)",
  x_label = "Observed Plant Food Percentage",
  y_label = "Predicted Plant Food Percentage",
  design = survey_design
)
```

```{r}
#| label: tbl-b4-coefficients-refined-model
#| tbl-cap: "Model coefficients for plant food percentage (refined model)"

# Export data for web resources
export(
  broom::tidy(results_plant_refined$model),
  here("01_data", "02_processed", "web_resources", "03_08_tbl_b4_coefficients_refined_model.rds")
)

# Display refined model coefficients
results_plant_refined$coefficients
```

```{r}
#| label: fig-b4-diagnostic-plot-refined-model
#| fig-cap: "Observed vs predicted plant food percentage (refined model)"

# Export data for web resources
export(
  results_plant_refined$results,
  here("01_data", "02_processed", "web_resources", "03_08_fig_b4_diagnostic_plot_refined_model.rds")
)

# Display diagnostic scatter plot
results_plant_refined$plot
```

## B5. Model Diagnostics

Evaluate model assumptions and identify potential issues with residuals, leverage, and multicollinearity.

```{r}
#| label: fig-b5-evaluate-diagnostics
#| fig-cap: "Diagnostic plots for plant food percentage model"
#| fig-width: 10
#| fig-height: 8
#| warning: FALSE

# Export data for web resources
export(
  broom::augment(results_plant_refined$model),
  here("01_data", "02_processed", "web_resources", "03_08_fig_b5_evaluate_diagnostics.rds")
)

# Generate diagnostic plots
evaluate_gamma_model(results_plant_refined$model)
```

---

# Phase C: Export

Save final refined model for application to SIVESNU data in subsequent transfer scripts.

```{r}
#| label: c-save-models

# Extract final model
model_plant_percentage <- results_plant_refined$model

# Save model as RDS
export(
  list(
    model_plant_percentage = model_plant_percentage
  ),
  here("01_data", "02_processed", 
       "models", "03_08_plant_percentage_model.rds")
)

# Clean auxiliary objects
rm(results_plant_refined, model_plant_percentage)
```

# Notes

## Model Performance

**Plant Food Percentage:**
- Final predictors: variables retained by the Wald significance criterion (p < 0.05), reported in the refined model coefficient table above
- Model fit: Pseudo-R² (Cragg-Uhler and McFadden) reported in the model summary above
- Model diagnostics: residual patterns, leverage and collinearity assessed in the diagnostic plots above
- Identity link function appropriate for bounded proportion data

## Methodological Rationale

The model uses an **identity link**:

1. **Bounded outcome**: Plant food percentage is naturally bounded [0,1]
2. **Linear interpretation**: Identity link preserves direct linear relationships with predictors
3. **Boundary adjustment**: Values of exactly 0 and 1 adjusted to 0.0001 and 0.9999 to satisfy gamma distribution requirements
4. **Convergence**: Gaussian model starting values improve convergence with identity link




