---
title: "Calculate Individual Nutrient Intake"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script completes the nutrient intake calculations by:

1. Loading individual demographic data with synthetic age in months (from script 01_03)
2. Calculating Adult Male Equivalent (AME) using age in months for improved granularity
3. Distributing household intake to individuals via AME proportions
4. Calculating individual nutrient intake for all nutrients
5. Separating maize vs. non-maize contributions at individual level
6. Creating the survey design for statistical analysis
7. Saving final integrated dataset

# Setup

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,          # Parquet file handling
  rio,            # Robust import/export for various formats
  here,           # Robust file path management
  gt,             # Table formatting and presentation
  survey,         # Complex survey methodology
  srvyr,          # Tidy wrapper for survey package
  tidyverse       # General purpose data manipulation
)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter
drop   <- base::drop

# Set display options
options(scipen = 999, digits = 4)

source(here("02_code", "00_functions", "00_functions.R"))

# Ensure output directories exist
ensure_output_dir(here("01_data", "02_processed", "transfer"))
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

---

# Phase A: Data Loading

## A1. Load Household Intake Data

Load all processed datasets from the previous script.

```{r}
#| label: a1-load-data

# Load household socioeconomic data
df_hogar <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_01_household_socioeconomic.parquet")
)

# Load individuals with synthetic age in months
df_individuos <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_03_individuals_with_age_months.parquet")
)

# Load household food intake datasets
df_alimentos <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_02_household_intake_total.parquet")
)

df_maiz <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_02_household_intake_maize.parquet")
)

df_vegetales <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_02_household_intake_plantbased.parquet")
)

df_no_vegetales <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_02_household_intake_animalbased.parquet")
)
```

## A2. Merge Nutrient Data

Combine individual, household, and food intake data.

```{r}
#| label: a2-merge-datasets

df_merged <- df_individuos |>
  # Join household socioeconomic data
  left_join(df_hogar |> select(-departamento), by = join_by(no_hogar, upm, factor)) |>
  # Join total food intake
  left_join(df_alimentos, by = join_by(no_hogar)) |>
  # Join maize-specific intake
  left_join(df_maiz, by = join_by(no_hogar)) |>
  # Join plant-based food intake
  left_join(df_vegetales, by = join_by(no_hogar)) |>
  # Join animal-based food intake
  left_join(df_no_vegetales, by = join_by(no_hogar))
```

## A3. Calculate Adult Male Equivalent (AME)

Assign AME coefficients using age in months and sex, following INCAP energy requirement guidelines for Guatemala (Weisell & Dop, 2012; INCAP, 2012). Children aged 0-5 months receive AME = 0 under the exclusive breastfeeding assumption. Children aged 6-11 months, identified through the synthetic month estimation in script 01_03, receive AME = 0.21 and enter the study population.

```{r}
#| label: a3-calculate-ame

df_with_ame <- df_merged |>
  mutate(
    ame = case_when(
      # Exclusive breastfeeding: 0-5 months
      edad_meses >= 0  & edad_meses <= 5   ~ 0,
      # Complementary feeding: 6-11 months
      edad_meses >= 6  & edad_meses <= 11  ~ 0.21,
      # Children 1-2 years
      edad_meses >= 12 & edad_meses <= 23  ~ 0.27,
      # Children 2-3 years
      edad_meses >= 24 & edad_meses <= 47  ~ 0.37,
      # Children 4-6 years
      edad_meses >= 48 & edad_meses <= 83  ~ 0.44,
      # Males 7+
      edad >= 7  & edad < 10  & sexo == "Hombre" ~ 0.56,
      edad >= 10 & edad < 14  & sexo == "Hombre" ~ 0.73,
      edad >= 14 & edad < 18  & sexo == "Hombre" ~ 0.96,
      edad >= 18 & edad < 30  & sexo == "Hombre" ~ 1.00,
      edad >= 30 & edad < 60  & sexo == "Hombre" ~ 0.95,
      edad >= 60              & sexo == "Hombre" ~ 0.76,
      # Females 7+
      edad >= 7  & edad < 10  & sexo == "Mujer" ~ 0.51,
      edad >= 10 & edad < 14  & sexo == "Mujer" ~ 0.65,
      edad >= 14 & edad < 18  & sexo == "Mujer" ~ 0.73,
      edad >= 18 & edad < 60  & sexo == "Mujer" ~ 0.74,
      edad >= 60              & sexo == "Mujer" ~ 0.65,
      TRUE ~ NA_real_
    )
  ) |>
  # Calculate AME percentage within household
  mutate(
    ame_hogar = sum(ame, na.rm = TRUE),
    .by = no_hogar
  ) |>
  mutate(
    ame_porcentaje = if_else(ame_hogar > 0, ame / ame_hogar, 0)
  ) |>
  select(-ame_hogar)
```

---

# Phase B: Individual Intake Distribution

## B1. Distribute Household Intake by AME

Apply AME percentage to distribute household intake to each individual.

```{r}
#| label: b1-calculate-individual-intake

df_intake <- df_with_ame  |>
  mutate(
    # === TOTAL INTAKE (all foods) ===
    gr_individuo_dia_totales = gr_hogar_dia_totales * ame_porcentaje,
    gr_individuo_dia_vegetales = gr_hogar_dia_vegetales * ame_porcentaje,
    gr_individuo_dia_no_vegetales = gr_hogar_dia_no_vegetales * ame_porcentaje,
    kcal_individuo_dia_totales = kcal_hogar_dia_totales * ame_porcentaje,
    prot_individuo_dia_totales = prot_hogar_dia_totales * ame_porcentaje,
    prot_apro_individuo_dia_totales = prot_apro_hogar_dia_totales * ame_porcentaje,
    fe_individuo_dia_totales = fe_hogar_dia_totales * ame_porcentaje,
    zn_individuo_dia_totales = zn_hogar_dia_totales * ame_porcentaje,
    lys_individuo_dia_totales = lys_hogar_dia_totales * ame_porcentaje,
    met_individuo_dia_totales = met_hogar_dia_totales * ame_porcentaje,
    cys_individuo_dia_totales = cys_hogar_dia_totales * ame_porcentaje,
    thr_individuo_dia_totales = thr_hogar_dia_totales * ame_porcentaje,
    trp_individuo_dia_totales = trp_hogar_dia_totales * ame_porcentaje,
    
    # === MAIZE INTAKE ===
    gr_individuo_dia_maiz = gr_hogar_dia_maiz * ame_porcentaje,
    kcal_individuo_dia_maiz = kcal_hogar_dia_maiz * ame_porcentaje,
    prot_individuo_dia_maiz = prot_hogar_dia_maiz * ame_porcentaje,
    prot_apro_individuo_dia_maiz = prot_apro_hogar_dia_maiz * ame_porcentaje,
    fe_individuo_dia_maiz = fe_hogar_dia_maiz * ame_porcentaje,
    zn_individuo_dia_maiz = zn_hogar_dia_maiz * ame_porcentaje,
    lys_individuo_dia_maiz = lys_hogar_dia_maiz * ame_porcentaje,
    met_individuo_dia_maiz = met_hogar_dia_maiz * ame_porcentaje,
    cys_individuo_dia_maiz = cys_hogar_dia_maiz * ame_porcentaje,
    thr_individuo_dia_maiz = thr_hogar_dia_maiz * ame_porcentaje,
    trp_individuo_dia_maiz = trp_hogar_dia_maiz * ame_porcentaje,
    
    # === NON-MAIZE INTAKE (by subtraction) ===
    gr_individuo_dia_no_maiz = gr_individuo_dia_totales - gr_individuo_dia_maiz,
    kcal_individuo_dia_no_maiz = kcal_individuo_dia_totales - kcal_individuo_dia_maiz,
    prot_individuo_dia_no_maiz = prot_individuo_dia_totales - prot_individuo_dia_maiz,
    prot_apro_individuo_dia_no_maiz = prot_apro_individuo_dia_totales - prot_apro_individuo_dia_maiz,
    fe_individuo_dia_no_maiz = fe_individuo_dia_totales - fe_individuo_dia_maiz,
    zn_individuo_dia_no_maiz = zn_individuo_dia_totales - zn_individuo_dia_maiz,
    lys_individuo_dia_no_maiz = lys_individuo_dia_totales - lys_individuo_dia_maiz,
    met_individuo_dia_no_maiz = met_individuo_dia_totales - met_individuo_dia_maiz,
    cys_individuo_dia_no_maiz = cys_individuo_dia_totales - cys_individuo_dia_maiz,
    thr_individuo_dia_no_maiz = thr_individuo_dia_totales - thr_individuo_dia_maiz,
    trp_individuo_dia_no_maiz = trp_individuo_dia_totales - trp_individuo_dia_maiz
  )
```

---

# Phase C: Finalize Dataset

## C1. Calculate Plant-Based Diet Percentage

Calculate the proportion of plant-based foods in each individual's diet. This variable is used in transfer models for predicting nutrient profiles.

```{r}
#| label: c1-calculate-plant-percentage

df_complete <- df_intake |>
  mutate(
    # Calculate percentage of plant-based foods in diet
    porcentaje_vegetales = gr_individuo_dia_vegetales / gr_individuo_dia_totales,
    # Stratify into ranges for downstream analysis
    per_vegetales_estratificado = factor(
      cut(
        porcentaje_vegetales, 
        breaks = c(-Inf, 0.5, 0.7, 0.9, Inf), 
        labels = c("0-50%", "50-70%", "70-90%", "90-100%")
      ),
      ordered = TRUE
    )
  )
```

## C2. Create Survey Design Object

Create the survey design accounting for ENCOVI's complex sampling structure (clusters and expansion weights). This enables population-representative statistical estimates.

```{r}
#| label: c2-create-survey-design

survey_design <- df_complete |>
  as_survey_design(
    ids = upm,
    weights = factor,
    # No stratification specified in original design
    strata = NULL
  )
```

---

# Phase D: Quality Checks and Export

## D1. Summary Statistics

```{r}
#| label: tbl-d1-summary-sample-characteristics

# Calculate sample characteristics
sample_summary <- survey_design |>
  summarise(
    individuals = n(),
    households = n_distinct(no_hogar),
    departments = n_distinct(departamento),
    pop_estimate = survey_total(vartype = "se")
  )

# Create summary table
tbl_sample_characteristics <- tibble(
  Characteristic = c(
    "Sample size (individuals)",
    "Unique households",
    "Departments covered",
    "Population estimate (weighted)"
  ),
  Value = c(
    format(sample_summary$individuals, big.mark = ","),
    format(sample_summary$households, big.mark = ","),
    sample_summary$departments,
    paste0(
      format(round(sample_summary$pop_estimate / 1e6, 2), nsmall = 2),
      " million (SE: ",
      format(round(sample_summary$pop_estimate_se / 1e6, 2), nsmall = 2),
      ")"
    )
  )
)

# Save data for web resources
export(
  tbl_sample_characteristics,
  here("01_data", "02_processed",
       "web_resources", "01_04_tbl_d1_summary_sample_characteristics.rds")
)

tbl_sample_characteristics |>
  gt() |>
  tab_header(
    title = "Sample Characteristics",
    subtitle = "Survey-weighted estimates for Guatemala"
  ) |>
  cols_label(
    Characteristic = "Characteristic",
    Value = "Value"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(sample_summary, tbl_sample_characteristics)
```

## D2. Distribution Diagnostics

### D2.a Total and Maize Intake

```{r}
#| label: tbl-d2a-summary-total-maize-intake

# Calculate total and maize intake statistics
intake_total_maize <- survey_design |>
  summarise(
    mean_gr_totales = survey_mean(gr_individuo_dia_totales, na.rm = TRUE, vartype = "ci"),
    mean_kcal = survey_mean(kcal_individuo_dia_totales, na.rm = TRUE, vartype = "ci"),
    mean_prot = survey_mean(prot_individuo_dia_totales, na.rm = TRUE, vartype = "ci"),
    mean_fe = survey_mean(fe_individuo_dia_totales, na.rm = TRUE, vartype = "ci"),
    mean_zn = survey_mean(zn_individuo_dia_totales, na.rm = TRUE, vartype = "ci"),
    mean_gr_maiz = survey_mean(gr_individuo_dia_maiz, na.rm = TRUE, vartype = "ci")
  )

# Create table
tbl_total_maize_intake <- tibble(
  Nutrient = c(
    "Food weight (g/day)",
    "Energy (kcal/day)",
    "Protein (g/day)",
    "Iron (mg/day)",
    "Zinc (mg/day)",
    "Maize weight (g/day)"
  ),
  `Mean (95% CI)` = c(
    paste0(round(intake_total_maize$mean_gr_totales, 1), " (",
           round(intake_total_maize$mean_gr_totales_low, 1), " - ",
           round(intake_total_maize$mean_gr_totales_upp, 1), ")"),
    paste0(round(intake_total_maize$mean_kcal, 0), " (",
           round(intake_total_maize$mean_kcal_low, 0), " - ",
           round(intake_total_maize$mean_kcal_upp, 0), ")"),
    paste0(round(intake_total_maize$mean_prot, 1), " (",
           round(intake_total_maize$mean_prot_low, 1), " - ",
           round(intake_total_maize$mean_prot_upp, 1), ")"),
    paste0(round(intake_total_maize$mean_fe, 2), " (",
           round(intake_total_maize$mean_fe_low, 2), " - ",
           round(intake_total_maize$mean_fe_upp, 2), ")"),
    paste0(round(intake_total_maize$mean_zn, 2), " (",
           round(intake_total_maize$mean_zn_low, 2), " - ",
           round(intake_total_maize$mean_zn_upp, 2), ")"),
    paste0(round(intake_total_maize$mean_gr_maiz, 1), " (",
           round(intake_total_maize$mean_gr_maiz_low, 1), " - ",
           round(intake_total_maize$mean_gr_maiz_upp, 1), ")")
  )
)

# Save data for web resources
export(
  tbl_total_maize_intake,
  here("01_data", "02_processed",
       "web_resources", "01_04_tbl_d2a_summary_total_maize_intake.rds")
)

tbl_total_maize_intake |>
  gt() |>
  tab_header(
    title = "Total and Maize Daily Intake",
    subtitle = "Survey-weighted population estimates with 95% confidence intervals"
  ) |>
  cols_label(
    Nutrient = "Nutrient/Metric",
    `Mean (95% CI)` = "Mean (95% CI)"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  ) |>
  tab_footnote(
    footnote = "Estimates account for complex survey design (clustering, weights).",
    locations = cells_title(groups = "subtitle")
  )

# Cleanup auxiliary objects
rm(intake_total_maize, tbl_total_maize_intake)
```

### D2.b Maize Contribution to Total Intake

```{r}
#| label: tbl-d2b-summary-maize-contribution

# Calculate maize contribution percentages
maize_contribution <- survey_design |>
  summarise(
    pct_kcal_maiz = survey_mean(
      kcal_individuo_dia_maiz / kcal_individuo_dia_totales * 100, 
      na.rm = TRUE, 
      vartype = "ci"
    ),
    pct_prot_maiz = survey_mean(
      prot_individuo_dia_maiz / prot_individuo_dia_totales * 100, 
      na.rm = TRUE, 
      vartype = "ci"
    ),
    pct_fe_maiz = survey_mean(
      fe_individuo_dia_maiz / fe_individuo_dia_totales * 100, 
      na.rm = TRUE, 
      vartype = "ci"
    ),
    pct_zn_maiz = survey_mean(
      zn_individuo_dia_maiz / zn_individuo_dia_totales * 100, 
      na.rm = TRUE, 
      vartype = "ci"
    )
  )

# Create table
tbl_maize_contribution <- tibble(
  Nutrient = c(
    "Energy contribution",
    "Protein contribution",
    "Iron contribution",
    "Zinc contribution"
  ),
  `% from Maize (95% CI)` = c(
    paste0(round(maize_contribution$pct_kcal_maiz, 1), "% (",
           round(maize_contribution$pct_kcal_maiz_low, 1), " - ",
           round(maize_contribution$pct_kcal_maiz_upp, 1), ")"),
    paste0(round(maize_contribution$pct_prot_maiz, 1), "% (",
           round(maize_contribution$pct_prot_maiz_low, 1), " - ",
           round(maize_contribution$pct_prot_maiz_upp, 1), ")"),
    paste0(round(maize_contribution$pct_fe_maiz, 1), "% (",
           round(maize_contribution$pct_fe_maiz_low, 1), " - ",
           round(maize_contribution$pct_fe_maiz_upp, 1), ")"),
    paste0(round(maize_contribution$pct_zn_maiz, 1), "% (",
           round(maize_contribution$pct_zn_maiz_low, 1), " - ",
           round(maize_contribution$pct_zn_maiz_upp, 1), ")")
  )
)

# Save data for web resources
export(
  tbl_maize_contribution,
  here("01_data", "02_processed",
       "web_resources", "01_04_tbl_d2b_summary_maize_contribution.rds")
)

tbl_maize_contribution |>
  gt() |>
  tab_header(
    title = "Maize Contribution to Total Nutrient Intake",
    subtitle = "Percentage of daily nutrients from maize products"
  ) |>
  cols_label(
    Nutrient = "Nutrient",
    `% from Maize (95% CI)` = "% from Maize (95% CI)"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(maize_contribution, tbl_maize_contribution)
```

### D2.c Intake by Age Group

```{r}
#| label: tbl-d2c-summary-intake-by-age

# Age bands built on edad_meses so they align with the AME month brackets:
# right = FALSE places the [0, 6) month band (AME = 0) in the first group
intake_by_age <- survey_design |>
  mutate(
    edad_grupos = factor(cut(
      edad_meses,
      breaks = c(-Inf, 6, 12, 24, 48, 84, 120, 168, 216, 360, 720, Inf),
      labels = c("de 0 a 5 meses", "de 6 meses a 1 año", "de 1 a 2 años",
                 "de 2 a 4 años", "de 4 a 7 años", "de 7 a 10 años",
                 "de 10 a 14 años", "de 14 a 18 años", "de 18 a 30 años",
                 "de 30 a 60 años", "más de 60 años"),
      right = FALSE
    ))
  ) |>
  group_by(edad_grupos) |>
  summarise(
    n = n(),
    mean_kcal = survey_mean(kcal_individuo_dia_totales, na.rm = TRUE),
    mean_prot = survey_mean(prot_individuo_dia_totales, na.rm = TRUE),
    mean_fe = survey_mean(fe_individuo_dia_totales, na.rm = TRUE),
    mean_zn = survey_mean(zn_individuo_dia_totales, na.rm = TRUE),
    pct_maiz_kcal = survey_mean(
      kcal_individuo_dia_maiz / kcal_individuo_dia_totales * 100, 
      na.rm = TRUE
    )
  )

# Save data for web resources
export(
  intake_by_age,
  here("01_data", "02_processed",
       "web_resources", "01_04_tbl_d2c_summary_intake_by_age.rds")
)

# Create table
intake_by_age |>
  gt() |>
  tab_header(
    title = "Daily Nutrient Intake by Age Group",
    subtitle = "Survey-weighted population estimates"
  ) |>
  fmt_number(
    columns = c(mean_kcal, mean_kcal_se, mean_prot, mean_prot_se,
                mean_fe, mean_fe_se, mean_zn, mean_zn_se,
                pct_maiz_kcal, pct_maiz_kcal_se),
    decimals = 1
  ) |>
  fmt_number(
    columns = n,
    decimals = 0,
    use_seps = TRUE
  ) |>
  cols_label(
    edad_grupos = "Age Group",
    n = "N",
    mean_kcal = "Energy (kcal)",
    mean_kcal_se = "Energy (SE)",
    mean_prot = "Protein (g)",
    mean_prot_se = "Protein (SE)",
    mean_fe = "Iron (mg)",
    mean_fe_se = "Iron (SE)",
    mean_zn = "Zinc (mg)",
    mean_zn_se = "Zinc (SE)",
    pct_maiz_kcal = "Maize Energy (%)",
    pct_maiz_kcal_se = "Maize Energy (SE)"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(intake_by_age)
```

### D2.d Diet Composition

```{r}
#| label: tbl-d2d-summary-diet-composition

# Calculate distribution of plant-based diet categories
diet_composition <- survey_design |>
  group_by(per_vegetales_estratificado) |>
  summarise(
    n = n(),
    proportion = survey_mean(vartype = "ci")
  )

# Save data for web resources
export(
  diet_composition,
  here("01_data", "02_processed",
       "web_resources", "01_04_tbl_d2d_summary_diet_composition.rds")
)

# Create table
diet_composition |>
  gt() |>
  tab_header(
    title = "Distribution of Plant-Based Diet Categories",
    subtitle = "Survey-weighted population proportions"
  ) |>
  fmt_number(
    columns = n,
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_percent(
    columns = c(proportion, proportion_low, proportion_upp),
    decimals = 1
  ) |>
  cols_label(
    per_vegetales_estratificado = "Plant-based Foods (%)",
    n = "Sample N",
    proportion = "Population %",
    proportion_low = "95% CI Lower",
    proportion_upp = "95% CI Upper"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  ) |>
  tab_footnote(
    footnote = "Categories based on percentage of grams from plant-based foods.",
    locations = cells_column_labels(columns = per_vegetales_estratificado)
  )

# Cleanup auxiliary objects
rm(diet_composition)
```

## D3. Data Quality Checks

### D3.a Missing Values

```{r}
#| label: tbl-d3a-quality-check-missing

# Calculate missing values
missing_summary <- df_complete |>
  summarise(
    missing_ame = sum(is.na(ame_porcentaje)),
    missing_intake = sum(is.na(gr_individuo_dia_totales)),
    missing_maiz = sum(is.na(gr_individuo_dia_maiz)),
    missing_plant_pct = sum(is.na(porcentaje_vegetales))
  )

# Create table
tbl_quality_missing <- tibble(
  Variable = c(
    "AME percentage",
    "Total intake",
    "Maize intake",
    "Plant-based percentage"
  ),
  `Missing N` = c(
    missing_summary$missing_ame,
    missing_summary$missing_intake,
    missing_summary$missing_maiz,
    missing_summary$missing_plant_pct
  ),
  `Missing %` = c(
    missing_summary$missing_ame / nrow(df_complete) * 100,
    missing_summary$missing_intake / nrow(df_complete) * 100,
    missing_summary$missing_maiz / nrow(df_complete) * 100,
    missing_summary$missing_plant_pct / nrow(df_complete) * 100
  )
)

# Save data for web resources
export(
  tbl_quality_missing,
  here("01_data", "02_processed",
       "web_resources", "01_04_tbl_d3a_quality_check_missing.rds")
)

tbl_quality_missing |>
  gt() |>
  tab_header(
    title = "Missing Values by Key Variable",
    subtitle = "Data completeness assessment"
  ) |>
  fmt_number(
    columns = `Missing N`,
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    columns = `Missing %`,
    decimals = 2,
    pattern = "{x}%"
  ) |>
  cols_label(
    Variable = "Variable",
    `Missing N` = "N",
    `Missing %` = "%"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  )

# Cleanup auxiliary objects
rm(missing_summary, tbl_quality_missing)
```

### D3.b Extreme Values

```{r}
#| label: tbl-d3b-quality-check-extremes

# Calculate extreme values
extreme_summary <- df_complete |>
  summarise(
    high_kcal = sum(kcal_individuo_dia_totales > 5000, na.rm = TRUE),
    low_kcal = sum(kcal_individuo_dia_totales < 500, na.rm = TRUE),
    high_prot = sum(prot_individuo_dia_totales > 200, na.rm = TRUE),
    zero_intake = sum(gr_individuo_dia_totales == 0, na.rm = TRUE)
  )

# Create table
tbl_quality_extremes <- tibble(
  Check = c(
    "Very high energy (>5000 kcal)",
    "Very low energy (<500 kcal)",
    "Very high protein (>200g)",
    "Zero food intake"
  ),
  Count = c(
    extreme_summary$high_kcal,
    extreme_summary$low_kcal,
    extreme_summary$high_prot,
    extreme_summary$zero_intake
  ),
  Percentage = c(
    extreme_summary$high_kcal / nrow(df_complete) * 100,
    extreme_summary$low_kcal / nrow(df_complete) * 100,
    extreme_summary$high_prot / nrow(df_complete) * 100,
    extreme_summary$zero_intake / nrow(df_complete) * 100
  )
)

# Save data for web resources
export(
  tbl_quality_extremes,
  here("01_data", "02_processed",
       "web_resources", "01_04_tbl_d3b_quality_check_extremes.rds")
)

tbl_quality_extremes |>
  gt() |>
  tab_header(
    title = "Extreme Value Detection",
    subtitle = "Potential data quality issues or outliers"
  ) |>
  fmt_number(
    columns = Count,
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_number(
    columns = Percentage,
    decimals = 2,
    pattern = "{x}%"
  ) |>
  cols_label(
    Check = "Quality Check",
    Count = "N",
    Percentage = "%"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    table.width = pct(100)
  ) |>
  tab_footnote(
    footnote = "Extreme values may indicate data entry errors, unusual household compositions, or recall bias.",
    locations = cells_title(groups = "subtitle")
  )

# Cleanup auxiliary objects
rm(extreme_summary, tbl_quality_extremes)
```

## D4. Save Final Dataset

```{r}
#| label: d4-save-outputs

# Save complete individual-level dataset
write_parquet(
  df_complete,
  here("01_data", "02_processed", 
       "transfer", "01_04_individual_nutritional_intake_complete.parquet")
)
```

---

# Notes

## Methodology

### Adult Male Equivalent (AME)
AME coefficients follow INCAP energy requirement guidelines for Guatemala, accounting for age and sex differences in energy needs. This script calculates AME using age in months (from script 01_03) for children 0-4 years.

### Maize/Non-Maize Separation
- Supports biofortification impact modeling at the individual level
- Maize-derived nutrients are the fraction modified in intervention scenarios
- Non-maize intake is held constant across scenarios

### Survey Design
- ENCOVI uses complex sampling with Primary Sampling Units (UPM) and expansion weights
- All population-level statistics require survey-weighted estimation
- Standard errors account for clustering within UPMs

## Data Quality

### Infants Under 6 Months
Infants under 6 months have `ame_porcentaje = 0`, resulting in zero calculated intake from solid foods, following the exclusive breastfeeding assumption.

### Missing Values
Missing values in intake variables occur for individuals in households with no food consumption data reported.

### Extreme Values
Extreme values (very high or very low intake) may reflect:
- Data entry errors in original survey
- Unusual household compositions (e.g., large families, guests)
- Reporting period issues (recall bias)
