---
title: "Calculate Household Nutrient Intake"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script processes household food consumption data and calculates nutrient intake at the household level. It performs:

1. Loads cleaned base datasets from previous script
2. Processes food consumption with unit conversions
3. Calculates separate intake for: total foods, maize only, plant-based foods, animal-based foods
4. Applies special adjustments for maize products (tortillas, flour)
5. Aggregates nutrient values by household

All calculations are performed at the **household level** before distributing to individuals in the next script.


# Setup

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,          # Parquet file handling
  rio,            # Robust import/export for various formats
  here,           # Robust file path management
  gt,             # Table formatting and presentation
  tidyverse       # General purpose data manipulation
)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

# Set display options
options(scipen = 999, digits = 4)

source(here("02_code", "00_functions", "00_functions.R"))

# Ensure output directories exist
ensure_output_dir(here("01_data", "02_processed", "transfer"))
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

# Phase A: Data Loading

Load the cleaned datasets created in the previous script.

```{r}
#| label: a-load-data

# Load integrated INCAP food composition data
df_incap <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_01_incap_integrated.parquet")
)

# Load ENCOVI food consumption data
encovi_alimentos <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_01_encovi_alimentos.parquet")
)

# Load food equivalences
equivalencias <- read_parquet(
  here("01_data", "02_processed", 
       "transfer", "01_01_food_equivalences.parquet")
)

# Load reference lists
ref_lists <- import(
  here("01_data", "02_processed", 
       "transfer", "01_01_reference_lists.rds"),
  trust = TRUE
)
nombres_alimentos <- ref_lists$nombres_alimentos
unidades_medida <- ref_lists$unidades_medida
```

---

# Phase B: Food Consumption Processing

Calculate total household nutrient intake from all foods.

## B1. Process Total Food Consumption

**ENCOVI food consumption variables:**
- `p13a07a`: Quantity purchased in last 15 days
- `p13a07c`: Unit code for purchased quantity
- `p13a11a`: Quantity obtained without purchase in last 15 days  
- `p13a11c`: Unit code for obtained quantity

**Maize Products Standardization (INCAP Methodology)**

Following INCAP guidelines for dietary intake assessment in Guatemala, we standardize maize-derived products to maize grain equivalent. Households report maize in three presentations — raw grain (codine 21), tortillas (codine 7) and flour (codine 17) — and all reported records are retained.

**Conversion to maize grain equivalent:**
Weight conversion factors isolate the maize component for biofortification modeling:

- **Tortillas (×0.52)**: Approximately 52% of tortilla weight corresponds to maize grain. The remaining mass consists of water absorbed during nixtamalization, calcium hydroxide (cal), and processing losses.
- **Maize flour (×0.98)**: Flour retains ~98% of original grain weight after dry milling.

**Nutrient value reassignment:**
After weight conversion, nutrient values are reassigned to raw white maize composition (per 100g): 365 kcal, 9.42g protein, 2.71mg iron, 2.21mg zinc. Nutrient calculations then reflect the maize fraction that biofortification modifies. Total dietary energy intake is unchanged by this step: the conversion identifies the maize-derived nutrient fraction within the household diet.

**Source:** Conversion factors derived from INCAP dietary assessment protocols.

```{r}
#| label: b1-process-total-foods

df_alimentos <- encovi_alimentos |>
  # Create unified measurement unit variable
  mutate(medida_encovi = case_when(
    !is.na(p13a07c) ~ p13a07c,
    is.na(p13a07c) & !is.na(p13a11c) ~ p13a11c,
    is.na(p13a07c) & is.na(p13a11c) ~ NA_real_
  )) |>
  # Join with food equivalences for unit conversion
  left_join(equivalencias, by = join_by(codine, medida_encovi == medida)) |>
  # Join with INCAP food composition data
  left_join(df_incap, by = join_by(codine)) |>
  # Recode food and measurement variables as factors
  mutate(
    codine = factor(
      codine, 
      levels = 1:length(nombres_alimentos),
      labels = nombres_alimentos,
      ordered = TRUE
    ),
    medida_encovi_factor = factor(
      medida_encovi,
      levels = c(1:67, 98, 101:125, 128, 129),
      labels = unidades_medida
    )
  ) |>
  # Convert quantities to grams using equivalence table
  mutate(
    gr_comida_comprada = gramos * p13a07a,
    gr_comida_obtenida = gramos * p13a11a
  ) |>
  # Sum purchased and obtained quantities
  rowwise() |>
  mutate(gr_comida_totales = sum(c(gr_comida_comprada, gr_comida_obtenida), na.rm = TRUE)) |>
  ungroup() |>
  # Calculate net grams per day (15-day recall / 15, adjusted for edible portion)
  mutate(gr_netos = (gr_comida_totales / 15) * pct_aprov) |>
  # Standardize maize products to grain equivalent (INCAP methodology)
  mutate(
    # Weight conversion: tortilla to grain (x0.52), flour to grain (x0.98)
    gr_netos = case_when(
      codine == "7 Tortillas de MAIZ" ~ gr_netos * 0.52,
      codine == "17 Harina de MAIZ" ~ gr_netos * 0.98,
      TRUE ~ gr_netos
    ),
    # Reassign nutrient composition to raw white maize values per 100 g
    # Source: INCAP Food Composition Table 2014
    across(
      c(energ_kcal, proteina_g, hierro_mg, zinc_mg, lys_g, met_g, cys_g, thr_g, trp_g),
      ~ if_else(
        codine %in% c("7 Tortillas de MAIZ", "17 Harina de MAIZ"), 
        case_when(
          cur_column() == "energ_kcal" ~ 365 / 100,
          cur_column() == "proteina_g" ~ 9.42 / 100,
          cur_column() == "hierro_mg" ~ 2.71 / 100,
          cur_column() == "zinc_mg" ~ 2.21 / 100,
          cur_column() == "lys_g" ~ 0.27 / 100,
          cur_column() == "met_g" ~ 0.20 / 100,
          cur_column() == "cys_g" ~ 0.17 / 100,
          cur_column() == "thr_g" ~ 0.35 / 100,
          cur_column() == "trp_g" ~ 0.07 / 100
        ),
        .x
      )
    )
  ) |>
  # Calculate nutrient content per food item
  mutate(
    across(
      c(energ_kcal, proteina_g, hierro_mg, zinc_mg, lys_g, met_g, cys_g, thr_g, trp_g),
      ~ if_else(!is.na(gr_netos), gr_netos * .x, NA_real_),
      .names = "{col}_comida"
    ),
    .by = codine
  ) |>
  # Calculate digestible protein (adjusted by PDCAAS)
  mutate(
    proteina_aprovechada_comida = if_else(
      !is.na(gr_netos), 
      gr_netos * proteina_g * pdcaas, 
      NA_real_
    )
  ) |>
  # Aggregate to household level
  summarise(
    gr_hogar_dia_totales = sum(gr_netos, na.rm = TRUE),
    across(
      c(energ_kcal_comida, proteina_g_comida, proteina_aprovechada_comida, 
        hierro_mg_comida, zinc_mg_comida, lys_g_comida,
        met_g_comida, cys_g_comida, thr_g_comida, trp_g_comida),
      ~ sum(.x, na.rm = TRUE),
      .names = "{.col}"
    ),
    .by = c(no_hogar, depto)
  ) |>
  # Rename columns for clarity
  rename(
    kcal_hogar_dia_totales = energ_kcal_comida,
    prot_hogar_dia_totales = proteina_g_comida,
    prot_apro_hogar_dia_totales = proteina_aprovechada_comida,
    fe_hogar_dia_totales = hierro_mg_comida,
    zn_hogar_dia_totales = zinc_mg_comida,
    lys_hogar_dia_totales = lys_g_comida,
    met_hogar_dia_totales = met_g_comida,
    cys_hogar_dia_totales = cys_g_comida,
    thr_hogar_dia_totales = thr_g_comida,
    trp_hogar_dia_totales = trp_g_comida
  ) |>
  select(-depto)
```

## B2. Process Maize-Only Consumption

Calculate household nutrient intake from maize products only (grain, tortillas, flour).

```{r}
#| label: b2-process-maize

df_maiz <- encovi_alimentos |>
  mutate(medida_encovi = case_when(
    !is.na(p13a07c) ~ p13a07c,
    is.na(p13a07c) & !is.na(p13a11c) ~ p13a11c,
    is.na(p13a07c) & is.na(p13a11c) ~ NA_real_
  )) |>
  left_join(equivalencias, by = join_by(codine, medida_encovi == medida)) |>
  left_join(df_incap, by = join_by(codine)) |>
  mutate(
    codine = factor(
      codine, 
      levels = 1:length(nombres_alimentos),
      labels = nombres_alimentos,
      ordered = TRUE
    ),
    medida_encovi_factor = factor(
      medida_encovi,
      levels = c(1:67, 98, 101:125, 128, 129),
      labels = unidades_medida
    )
  ) |>
  mutate(
    gr_comida_comprada = gramos * p13a07a,
    gr_comida_obtenida = gramos * p13a11a
  ) |>
  rowwise() |>
  mutate(gr_comida_totales = sum(c(gr_comida_comprada, gr_comida_obtenida), na.rm = TRUE)) |>
  ungroup() |>
  # Filter for maize products only
  filter(codine %in% c("7 Tortillas de MAIZ", "17 Harina de MAIZ", "21 MAIZ")) |>
  mutate(gr_netos = (gr_comida_totales / 15) * pct_aprov) |>
  # Standardize maize products to grain equivalent
  mutate(
    gr_netos = case_when(
      codine == "7 Tortillas de MAIZ" ~ gr_netos * 0.52,
      codine == "17 Harina de MAIZ" ~ gr_netos * 0.98,
      TRUE ~ gr_netos
    ),
    across(
      c(energ_kcal, proteina_g, hierro_mg, zinc_mg, lys_g, met_g, cys_g, thr_g, trp_g),
      ~ if_else(
        codine %in% c("7 Tortillas de MAIZ", "17 Harina de MAIZ"), 
        case_when(
          cur_column() == "energ_kcal" ~ 365 / 100,
          cur_column() == "proteina_g" ~ 9.42 / 100,
          cur_column() == "hierro_mg" ~ 2.71 / 100,
          cur_column() == "zinc_mg" ~ 2.21 / 100,
          cur_column() == "lys_g" ~ 0.27 / 100,
          cur_column() == "met_g" ~ 0.20 / 100,
          cur_column() == "cys_g" ~ 0.17 / 100,
          cur_column() == "thr_g" ~ 0.35 / 100,
          cur_column() == "trp_g" ~ 0.07 / 100
        ),
        .x
      )
    )
  ) |>
  mutate(
    across(
      c(energ_kcal, proteina_g, hierro_mg, zinc_mg, lys_g, met_g, cys_g, thr_g, trp_g),
      ~ if_else(!is.na(gr_netos), gr_netos * .x, NA_real_),
      .names = "{col}_comida"
    ),
    .by = codine
  ) |>
  mutate(
    proteina_aprovechada_comida = if_else(
      !is.na(gr_netos), 
      gr_netos * proteina_g * pdcaas, 
      NA_real_
    )
  ) |>
  summarise(
    gr_hogar_dia_maiz = sum(gr_netos, na.rm = TRUE),
    across(
      c(energ_kcal_comida, proteina_g_comida, proteina_aprovechada_comida,
        hierro_mg_comida, zinc_mg_comida, lys_g_comida,
        met_g_comida, cys_g_comida, thr_g_comida, trp_g_comida),
      ~ sum(.x, na.rm = TRUE),
      .names = "{.col}"
    ),
    .by = c(no_hogar, depto)
  ) |>
  rename(
    kcal_hogar_dia_maiz = energ_kcal_comida,
    prot_hogar_dia_maiz = proteina_g_comida,
    prot_apro_hogar_dia_maiz = proteina_aprovechada_comida,
    fe_hogar_dia_maiz = hierro_mg_comida,
    zn_hogar_dia_maiz = zinc_mg_comida,
    lys_hogar_dia_maiz = lys_g_comida,
    met_hogar_dia_maiz = met_g_comida,
    cys_hogar_dia_maiz = cys_g_comida,
    thr_hogar_dia_maiz = thr_g_comida,
    trp_hogar_dia_maiz = trp_g_comida
  ) |>
  select(-depto)
```

## B3. Process Plant-Based Foods

Calculate household intake from plant-based foods (for phytate estimation).

```{r}
#| label: b3-process-vegetables

df_vegetales <- encovi_alimentos |>
  mutate(medida_encovi = case_when(
    !is.na(p13a07c) ~ p13a07c,
    is.na(p13a07c) & !is.na(p13a11c) ~ p13a11c,
    is.na(p13a07c) & is.na(p13a11c) ~ NA_real_
  )) |>
  left_join(equivalencias, by = join_by(codine, medida_encovi == medida)) |>
  left_join(df_incap, by = join_by(codine)) |>
  # Filter for plant-based foods only
  filter(vegetal_cereal == "si") |>
  mutate(
    codine = factor(
      codine, 
      levels = 1:length(nombres_alimentos),
      labels = nombres_alimentos,
      ordered = TRUE
    ),
    medida_encovi_factor = factor(
      medida_encovi,
      levels = c(1:67, 98, 101:125, 128, 129),
      labels = unidades_medida
    )
  ) |>
  mutate(
    gr_comida_comprada = gramos * p13a07a,
    gr_comida_obtenida = gramos * p13a11a
  ) |>
  rowwise() |>
  mutate(gr_comida_totales = sum(c(gr_comida_comprada, gr_comida_obtenida), na.rm = TRUE)) |>
  ungroup() |>
  mutate(gr_netos = (gr_comida_totales / 15) * pct_aprov) |>
  # Standardize maize products to grain equivalent
  mutate(
    gr_netos = case_when(
      codine == "7 Tortillas de MAIZ" ~ gr_netos * 0.52,
      codine == "17 Harina de MAIZ" ~ gr_netos * 0.98,
      TRUE ~ gr_netos
    ),
    across(
      c(energ_kcal, proteina_g, hierro_mg, zinc_mg, lys_g, met_g, cys_g, thr_g, trp_g),
      ~ if_else(
        codine %in% c("7 Tortillas de MAIZ", "17 Harina de MAIZ"), 
        case_when(
          cur_column() == "energ_kcal" ~ 365 / 100,
          cur_column() == "proteina_g" ~ 9.42 / 100,
          cur_column() == "hierro_mg" ~ 2.71 / 100,
          cur_column() == "zinc_mg" ~ 2.21 / 100,
          cur_column() == "lys_g" ~ 0.27 / 100,
          cur_column() == "met_g" ~ 0.20 / 100,
          cur_column() == "cys_g" ~ 0.17 / 100,
          cur_column() == "thr_g" ~ 0.35 / 100,
          cur_column() == "trp_g" ~ 0.07 / 100
        ),
        .x
      )
    )
  ) |>
  summarise(
    gr_hogar_dia_vegetales = sum(gr_netos, na.rm = TRUE),
    .by = c(no_hogar, depto)
  ) |>
  select(-depto)
```

## B4. Process Animal-Based Foods

Calculate household intake from animal-based foods.

```{r}
#| label: b4-process-non-vegetables

df_no_vegetales <- encovi_alimentos |>
  mutate(medida_encovi = case_when(
    !is.na(p13a07c) ~ p13a07c,
    is.na(p13a07c) & !is.na(p13a11c) ~ p13a11c,
    is.na(p13a07c) & is.na(p13a11c) ~ NA_real_
  )) |>
  left_join(equivalencias, by = join_by(codine, medida_encovi == medida)) |>
  left_join(df_incap, by = join_by(codine)) |>
  # Filter for animal-based foods only
  filter(vegetal_cereal == "no") |>
  mutate(
    codine = factor(
      codine, 
      levels = 1:length(nombres_alimentos),
      labels = nombres_alimentos,
      ordered = TRUE
    ),
    medida_encovi_factor = factor(
      medida_encovi,
      levels = c(1:67, 98, 101:125, 128, 129),
      labels = unidades_medida
    )
  ) |>
  mutate(
    gr_comida_comprada = gramos * p13a07a,
    gr_comida_obtenida = gramos * p13a11a
  ) |>
  rowwise() |>
  mutate(gr_comida_totales = sum(c(gr_comida_comprada, gr_comida_obtenida), na.rm = TRUE)) |>
  ungroup() |>
  mutate(gr_netos = (gr_comida_totales / 15) * pct_aprov) |>
  # Standardize maize products to grain equivalent
  mutate(
    gr_netos = case_when(
      codine == "7 Tortillas de MAIZ" ~ gr_netos * 0.52,
      codine == "17 Harina de MAIZ" ~ gr_netos * 0.98,
      TRUE ~ gr_netos
    ),
    across(
      c(energ_kcal, proteina_g, hierro_mg, zinc_mg, lys_g, met_g, cys_g, thr_g, trp_g),
      ~ if_else(
        codine %in% c("7 Tortillas de MAIZ", "17 Harina de MAIZ"), 
        case_when(
          cur_column() == "energ_kcal" ~ 365 / 100,
          cur_column() == "proteina_g" ~ 9.42 / 100,
          cur_column() == "hierro_mg" ~ 2.71 / 100,
          cur_column() == "zinc_mg" ~ 2.21 / 100,
          cur_column() == "lys_g" ~ 0.27 / 100,
          cur_column() == "met_g" ~ 0.20 / 100,
          cur_column() == "cys_g" ~ 0.17 / 100,
          cur_column() == "thr_g" ~ 0.35 / 100,
          cur_column() == "trp_g" ~ 0.07 / 100
        ),
        .x
      )
    )
  ) |>
  summarise(
    gr_hogar_dia_no_vegetales = sum(gr_netos, na.rm = TRUE),
    .by = c(no_hogar, depto)
  ) |>
  select(-depto)
```

---

# Phase C: Household Aggregation

```{r}
#| label: tbl-c-summary

# Calculate summary statistics for each dataset
summary_stats <- tibble(
  Metric = c(
    "Households (N)",
    "Grams (median)",
    "Grams (IQR)",
    "Energy - kcal (median)",
    "Energy - kcal (IQR)",
    "Protein - g (median)",
    "Protein - g (IQR)",
    "Iron - mg (median)",
    "Iron - mg (IQR)",
    "Zinc - mg (median)",
    "Zinc - mg (IQR)"
  ),
  Total = c(
    nrow(df_alimentos),
    round(median(df_alimentos$gr_hogar_dia_totales, na.rm = TRUE), 1),
    paste0(
      round(quantile(df_alimentos$gr_hogar_dia_totales, 0.25, na.rm = TRUE), 1),
      " - ",
      round(quantile(df_alimentos$gr_hogar_dia_totales, 0.75, na.rm = TRUE), 1)
    ),
    round(median(df_alimentos$kcal_hogar_dia_totales, na.rm = TRUE), 0),
    paste0(
      round(quantile(df_alimentos$kcal_hogar_dia_totales, 0.25, na.rm = TRUE), 0),
      " - ",
      round(quantile(df_alimentos$kcal_hogar_dia_totales, 0.75, na.rm = TRUE), 0)
    ),
    round(median(df_alimentos$prot_hogar_dia_totales, na.rm = TRUE), 1),
    paste0(
      round(quantile(df_alimentos$prot_hogar_dia_totales, 0.25, na.rm = TRUE), 1),
      " - ",
      round(quantile(df_alimentos$prot_hogar_dia_totales, 0.75, na.rm = TRUE), 1)
    ),
    round(median(df_alimentos$fe_hogar_dia_totales, na.rm = TRUE), 2),
    paste0(
      round(quantile(df_alimentos$fe_hogar_dia_totales, 0.25, na.rm = TRUE), 2),
      " - ",
      round(quantile(df_alimentos$fe_hogar_dia_totales, 0.75, na.rm = TRUE), 2)
    ),
    round(median(df_alimentos$zn_hogar_dia_totales, na.rm = TRUE), 2),
    paste0(
      round(quantile(df_alimentos$zn_hogar_dia_totales, 0.25, na.rm = TRUE), 2),
      " - ",
      round(quantile(df_alimentos$zn_hogar_dia_totales, 0.75, na.rm = TRUE), 2)
    )
  ),
  Maize = c(
    nrow(df_maiz),
    round(median(df_maiz$gr_hogar_dia_maiz, na.rm = TRUE), 1),
    paste0(
      round(quantile(df_maiz$gr_hogar_dia_maiz, 0.25, na.rm = TRUE), 1),
      " - ",
      round(quantile(df_maiz$gr_hogar_dia_maiz, 0.75, na.rm = TRUE), 1)
    ),
    round(median(df_maiz$kcal_hogar_dia_maiz, na.rm = TRUE), 0),
    paste0(
      round(quantile(df_maiz$kcal_hogar_dia_maiz, 0.25, na.rm = TRUE), 0),
      " - ",
      round(quantile(df_maiz$kcal_hogar_dia_maiz, 0.75, na.rm = TRUE), 0)
    ),
    round(median(df_maiz$prot_hogar_dia_maiz, na.rm = TRUE), 1),
    paste0(
      round(quantile(df_maiz$prot_hogar_dia_maiz, 0.25, na.rm = TRUE), 1),
      " - ",
      round(quantile(df_maiz$prot_hogar_dia_maiz, 0.75, na.rm = TRUE), 1)
    ),
    round(median(df_maiz$fe_hogar_dia_maiz, na.rm = TRUE), 2),
    paste0(
      round(quantile(df_maiz$fe_hogar_dia_maiz, 0.25, na.rm = TRUE), 2),
      " - ",
      round(quantile(df_maiz$fe_hogar_dia_maiz, 0.75, na.rm = TRUE), 2)
    ),
    round(median(df_maiz$zn_hogar_dia_maiz, na.rm = TRUE), 2),
    paste0(
      round(quantile(df_maiz$zn_hogar_dia_maiz, 0.25, na.rm = TRUE), 2),
      " - ",
      round(quantile(df_maiz$zn_hogar_dia_maiz, 0.75, na.rm = TRUE), 2)
    )
  ),
  `Plant-based` = c(
    nrow(df_vegetales),
    round(median(df_vegetales$gr_hogar_dia_vegetales, na.rm = TRUE), 1),
    paste0(
      round(quantile(df_vegetales$gr_hogar_dia_vegetales, 0.25, na.rm = TRUE), 1),
      " - ",
      round(quantile(df_vegetales$gr_hogar_dia_vegetales, 0.75, na.rm = TRUE), 1)
    ),
    "—", "—", "—", "—", "—", "—", "—", "—"
  ),
  `Animal-based` = c(
    nrow(df_no_vegetales),
    round(median(df_no_vegetales$gr_hogar_dia_no_vegetales, na.rm = TRUE), 1),
    paste0(
      round(quantile(df_no_vegetales$gr_hogar_dia_no_vegetales, 0.25, na.rm = TRUE), 1),
      " - ",
      round(quantile(df_no_vegetales$gr_hogar_dia_no_vegetales, 0.75, na.rm = TRUE), 1)
    ),
    "—", "—", "—", "—", "—", "—", "—", "—"
  )
)

# Save data for web resources
export(
  summary_stats,
  here("01_data", "02_processed",
       "web_resources", "01_02_tbl_c_summary.rds")
)

# Create GT table
summary_stats |>
  gt() |>
  tab_header(
    title = "Household Daily Intake Summary",
    subtitle = "Median and interquartile range (IQR) by food group"
  ) |>
  cols_label(
    Metric = "Metric",
    Total = "Total Foods",
    Maize = "Maize Only",
    `Plant-based` = "Plant-based",
    `Animal-based` = "Animal-based"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Metric,
      rows = Metric %in% c("Households (N)", "Grams (median)", "Energy - kcal (median)", 
                           "Protein - g (median)", "Iron - mg (median)", "Zinc - mg (median)")
    )
  ) |>
  tab_style(
    style = cell_text(style = "italic", color = "gray50"),
    locations = cells_body(
      columns = Metric,
      rows = grepl("IQR", Metric)
    )
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  ) |>
  tab_footnote(
    footnote = "Plant-based and animal-based datasets track weight (grams) only.",
    locations = cells_column_labels(columns = c(`Plant-based`, `Animal-based`))
  )

# Cleanup auxiliary objects
rm(summary_stats)
```

---

# Phase D: Export

```{r}
#| label: d-save-outputs

# Save all household intake datasets
write_parquet(
  df_alimentos,
  here("01_data", "02_processed", 
       "transfer", "01_02_household_intake_total.parquet")
)

write_parquet(
  df_maiz,
  here("01_data", "02_processed", 
       "transfer", "01_02_household_intake_maize.parquet")
)

write_parquet(
  df_vegetales,
  here("01_data", "02_processed", 
       "transfer", "01_02_household_intake_plantbased.parquet")
)

write_parquet(
  df_no_vegetales,
  here("01_data", "02_processed", 
       "transfer", "01_02_household_intake_animalbased.parquet")
)
```

# Notes

## Methodology

### Unit Conversions
- All food quantities converted to grams using expert-provided equivalence table
- Purchased and obtained (non-purchased) foods combined for total household intake
- Daily intake calculated by dividing 15-day recall period by 15

### Maize Product Adjustments
Conversion factors standardize processed maize to grain equivalent:
- **Tortillas**: 0.52 factor (water absorbed during nixtamalization, cal, and processing losses)
- **Flour**: 0.98 factor (dry milling loss)
- **Nutrient values reassigned** to raw white maize composition

### Edible Portion
All calculations include adjustment for edible portion (`pct_aprov`) to account for waste (peels, bones, etc.).

### Food Groupings
- **Total foods**: All consumed foods
- **Maize**: Grain, tortillas, and flour
- **Plant-based**: All cereals, grains, vegetables, fruits (proxy for phytate content)
- **Animal-based**: All meat, dairy, eggs, fish

## Data Quality

### Missing Values
Missing values in measurement units are handled by prioritizing purchased quantity measurements, then obtained quantity measurements.

### Zero Consumption
Households with zero reported consumption for a food grouping will have corresponding zero values in that dataset.


