---
title: "Load and Prepare Base Data"
author: "Azahar Data Insights"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
editor: source
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This script performs initial data loading, cleaning, and preparation of all base datasets required for the nutrient intake analysis:

1. Loads raw parquet files (ENCOVI 2023, INCAP 2014, external expert data)
2. Loads the integrated food composition table (energy, protein, iron, zinc, amino acids, PDCAAS, NPU)
3. Processes household socioeconomic variables
4. Cleans individual demographic data and calculates Adult Male Equivalent (AME) factors
5. Prepares food equivalence tables and reference lists
6. Saves all cleaned and integrated datasets for subsequent analyses


# Setup

```{r}
#| label: setup

# Clear environment and unload packages
rm(list = ls())
tryCatch(
  suppressWarnings(pacman::p_unload(all)),
  error = function(e) invisible(NULL)
)

# Load required packages
pacman::p_load(
  arrow,          # Parquet file handling
  rio,            # Robust import/export for various formats
  here,           # Robust file path management
  janitor,        # Data cleaning utilities
  gt,             # Table formatting and presentation
  tidyverse       # General purpose data manipulation
)

# Resolve namespace conflicts
select <- dplyr::select
filter <- dplyr::filter

# Set display options
options(scipen = 999, digits = 4)

source(here("02_code", "00_functions", "00_functions.R"))

# Ensure output directories exist
ensure_output_dir(here("01_data", "02_processed", "transfer"))
ensure_output_dir(here("01_data", "02_processed", "web_resources"))
```

# Phase A: Data Loading

## A1. Load ENCOVI 2023 Survey Data

```{r}
#| label: a1-load-encovi

# Load food consumption data
encovi_alimentos <- read_parquet(
  here("01_data", "01_raw", "ENCOVI_2023", "encovi_2023_c13sa_alimentos_3.parquet")
) |>
  clean_names() |>
  remove_empty(c("cols", "rows")) |>
  rename(codine = p13a04a)  # Match INCAP food code

# Load household data
encovi_hogar <- read_parquet(
  here("01_data", "01_raw", "ENCOVI_2023", "encovi_2023_hogares.parquet")
) |>
  clean_names() |>
  remove_empty(c("cols", "rows"))

# Load individual/person data
encovi_individuo <- read_parquet(
  here("01_data", "01_raw", "ENCOVI_2023", "encovi_2023_personas.parquet")
) |>
  clean_names() |>
  remove_empty(c("cols", "rows"))
```

## A2. Load Food Composition Data

```{r}
#| label: a2-load-composition

# Composition table built by Script 00_03, one row per food,
# values per 100 g edible portion
incap_composition <- read_parquet(
  here("01_data", "03_external", "incap_food_composition.parquet")
)
```

## A3. Load Food Equivalences

```{r}
#| label: a3-load-equivalences

# Equivalence table built by Script 00_03, one row per food and
# presentation unit, with unit-to-grams conversions
equivalencias <- read_parquet(
  here("01_data", "03_external", "incap_food_equivalences.parquet")
)
```

---

# Phase B: Data Integration and Processing

## B1. Prepare INCAP Food Composition Database

Normalize nutrient values to a per-gram basis and apply the conservative protein-quality assumption for foods without a PDCAAS score.

```{r}
#| label: b1-prepare-incap

df_incap <- incap_composition |>
  # Convert nutrient values to a per-gram basis
  mutate(across(
    c(energ_kcal, proteina_g, hierro_mg, zinc_mg,
      lys_g, met_g, cys_g, thr_g, trp_g),
    ~ .x / 100
  )) |>
  # Replace missing PDCAAS values with 0.1 (conservative estimate for low-quality protein)
  mutate(pdcaas = if_else(is.na(pdcaas), 0.1, pdcaas))
```


## B2. Process Household Socioeconomic Variables

Clean and recode household socioeconomic variables.

```{r}
#| label: b2-process-household

df_hogar <- encovi_hogar |>
  transmute(
    no_hogar, upm, factor,
    departamento = factor(
      depto, 
      levels = 1:22, 
      labels = c("Guatemala", "El Progreso", "Sacatepéquez", "Chimaltenango", 
                 "Escuintla", "Santa Rosa", "Sololá", "Totonicapán", "Quetzaltenango", 
                 "Suchitepéquez", "Retalhuleu", "San Marcos", "Huehuetenango", "Quiché", 
                 "Baja Verapaz", "Alta Verapaz", "Petén", "Izabal", "Zacapa", "Chiquimula", 
                 "Jalapa", "Jutiapa"), 
      ordered = TRUE
    ),
    area = factor(area, levels = 1:2, labels = c("Urbana", "Rural")),
    miembros_hogar = total_pers,
    propiedad = factor(
      p01b01, 
      levels = c(1:4, 98), 
      labels = c("Propia y totalmente pagada", "Propia y pagandola a plazos", 
                 "Alquilada", "Cedida o prestada", "Otro")
    ),
    material_paredes = factor(
      p01a02,
      levels = c(1:8, 98), 
      labels = c("Ladrillo", "Block", "Concreto", "Adobe", "Madera", 
                 "Lamina metalica", "Bajareque", "Lepa, palo o caña", "Otro")
    ),
    material_techo = factor(
      p01a03,
      levels = c(1:5, 98), 
      labels = c("Concreto", "Lamina metalica", "Asbesto cemento", 
                 "Teja", "Paja, palma o similar", "Otro")
    ),
    material_piso = factor(
      p01a04,
      levels = c(1:7, 98), 
      labels = c("Ladrillo ceramico", "Ladrillo de cemento", "Ladrillo de barro", 
                 "Torta de cemento", "Parqué", "Madera", "Tierra", "Tierra")
    ),
    tipo_vivienda = factor(
      p01a01,
      levels = c(1:5, 98), 
      labels = c("Casa formal", "Apartamento", "Cuarto en casa de vecindad", 
                 "Rancho", "Casa improvisada", "Otro")
    ),
    n_cuartos = p01d01,
    n_dormitorios = p01d02,
    cocina = factor(
      p01d05,
      levels = 1:2, 
      labels = c("De uso exclusivo", "De uso compartido")
    ),
    fuente_agua = factor(
      p01d06,
      levels = c(1:7, 98), 
      labels = c("Tubería dentro de la vivienda", "Tubería fuera de la vivienda", 
                 "Chorro publico", "Pozo perforado", "Río, lago, manantial", 
                 "Camion cisterna", "Agua de lluvia", "Otro")
    ),
    tipo_sanitario = factor(
      p01d18,
      levels = 1:2, 
      labels = c("Uso exclusivo", "Uso compartido")
    ),
    recoleccion_basura = factor(case_when(
      p01d22 == 1 ~ "Servicio municipal",
      p01d22 == 2 ~ "Servicio privado",
      p01d22 == 3 ~ "La queman o la entierran",
      p01d22 == 4 ~ "La queman o la entierran",
      p01d22 == 5 ~ "La tiran en cualquier lugar"
    )),
    electricidad = factor(p01a05c, levels = 1:2, labels = c("Si", "No")),
    televisor = factor(p01d20e, levels = 1:2, labels = c("Si", "No")),
    telefonia_fija = factor(p01d20b, levels = 1:2, labels = c("Si", "No")),
    telefonia_celular = factor(p01d20c, levels = 1:2, labels = c("Si", "No")),
    computadora = factor(p01d20d, levels = 1:2, labels = c("Si", "No"))
  )
```

## B3. Process Individual Demographics

Clean individual demographics.

```{r}
#| label: b3-process-individuals

df_individuos <- encovi_individuo |>
  mutate(
    # Recode relationship to household head
    parentesco = factor(case_when(
      ppa05 == 1 ~ "Jefe del hogar",
      ppa05 == 2 ~ "Esposo(a) o compañero(a)",
      ppa05 == 3 ~ "Hijo(a)",
      ppa05 == 4 ~ "Yerno o nuera",
      ppa05 == 5 ~ "Nieto(a)",
      ppa05 == 6 ~ "Padre o madre",
      ppa05 == 7 ~ "Suegro(a)",
      ppa05 == 8 ~ "Hermano(a)",
      ppa05 %in% c(9, 10) ~ "Otro pariente",
      ppa05 == 11 ~ "Empleado de casa particular",
      ppa05 %in% c(12, 13) ~ "Otro no pariente"
    )),
    # Recode sex
    sexo = factor(ppa02, levels = 1:2, labels = c("Hombre", "Mujer")),
    # Age variable
    edad = ppa03,
    # Recode education level
    grado_estudios = factor(
      if_else(!is.na(p04a02), p04a02, p04a05),
      levels = 1:12, 
      labels = c("Ninguno", "Solo sabe leer", "Preprimaria", 
                 "Primaria incompleta", "Primaria completa", 
                 "Educacion media incompleta", "Educacion media completa", 
                 "Carrera universitaria incompleta", "Carrera universitaria completa", 
                 "Maestria o doctorado", "No lo conoce o conocio", "No sabe")
    )
  ) |>
  # Simplify education categories
  mutate(
    grado_estudios = fct_relevel(
      factor(case_when(
        grado_estudios == "Ninguno" ~ "Ninguno",
        grado_estudios %in% c("Solo sabe leer", "Preprimaria", "Primaria incompleta") ~ "Solo sabe leer",
        grado_estudios %in% c("Primaria completa", "Educacion media incompleta") ~ "Primaria",
        grado_estudios %in% c("Educacion media completa", "Carrera universitaria incompleta") ~ "Secundaria",
        grado_estudios %in% c("Carrera universitaria completa", "Maestria o doctorado") ~ "Superior/universitaria",
        grado_estudios %in% c("No lo conoce o conocio", "No sabe") ~ "Desconocido"
      ), ordered = TRUE), 
      c("Desconocido", "Ninguno", "Solo sabe leer", "Primaria", "Secundaria", "Superior/universitaria")
    ),
    # Convert to numeric for household heads and partners
    grado_estudios_num = case_when(
      grado_estudios == "Desconocido" ~ 1,
      grado_estudios == "Ninguno" ~ 1,
      grado_estudios == "Solo sabe leer" ~ 2,
      grado_estudios == "Primaria" ~ 4,
      grado_estudios == "Secundaria" ~ 7,
      grado_estudios == "Superior/universitaria" ~ 10
    ),
    grado_estudios_num = if_else(
      parentesco %in% c("Jefe del hogar", "Esposo(a) o compañero(a)"), 
      grado_estudios_num, 
      NA_real_
    )
  ) |>
  # Calculate mean education level per household
  mutate(
    grado_estudios_hogar = mean(grado_estudios_num, na.rm = TRUE),
    .by = no_hogar
  ) |>
  # Create age groups
  mutate(
    edad_grupos = factor(cut(
      edad, 
      breaks = c(-Inf, 0.5, 1, 2, 4, 7, 10, 14, 18, 30, 60, Inf), 
      labels = c("de 0 a 6 meses", "de 6 meses a 1 año", "de 1 a 2 años",
                 "de 2 a 4 años", "de 4 a 7 años", "de 7 a 10 años",
                 "de 10 a 14 años", "de 14 a 18 años", "de 18 a 30 años",
                 "de 30 a 60 años", "más de 60 años")
    ))
  ) |>
  # Select key variables
  select(no_hogar, upm, factor, parentesco, grado_estudios, grado_estudios_num, 
         grado_estudios_hogar, sexo, edad, edad_grupos)
```

## B4. Create Reference Lists

Create reference vectors for subsequent data processing.

```{r}
#| label: b4-reference-lists

# Extract unique food names in order from equivalence table
nombres_alimentos <- unique(equivalencias$alimento)

# Define measurement units vector (matches ENCOVI survey codes)
unidades_medida <- c(
  "1 ACRE", "2 ALMUD", "3 ARROBA", "4 BARRIL", "5 BOLSA", "6 BOTELLA", 
  "7 BULTO", "8 CABALLER", "9 CAJA", "10 CANASTO", "11 CARGA", "12 CARRETAD", 
  "13 CIENTO", "14 COPA", "15 CUADRA", "16 CUARTERO", "17 CUERDA", "18 DOCENA", 
  "19 FANEGA", "20 GALON", "21 GARRAFON", "22 GRAMOS", "23 HECTAREA", 
  "24 KILOGRAM", "25 KILOMETR", "26 KILOVATI", "27 LATA", "28 LEGUA", 
  "29 LEÑO O R", "30 LIBRA", "31 LITRO", "32 MANO", "33 MANOJO", "34 MANZANA", 
  "35 MEDIA BO", "36 MEDIDA", "37 MEDIO (S", "38 MEDIO CI", "39 MEDIO GA", 
  "40 METRO CU", "41 METRO", "42 MILLA", "43 MILLAR", "44 OCTAVO", "45 ONZA", 
  "46 PAJA", "47 PAR", "48 PIE LINE", "49 PIE TABL", "50 PULGADA", "51 QUINTAL", 
  "52 RACIMO", "53 RED", "54 SACO", "55 TAREA", "56 TERCIO", "57 TIJANA", 
  "58 TONEL", "59 TONELADA", "60 TRENZA", "61 TROZA", "62 UN CUART", "63 UNIDAD", 
  "64 VARA", "65 VASO", "66 YARDA", "67 PILA", "98 OTRO", "101 CABEZA", 
  "102 SOBRE", "103 BARRA", "104 BOTE", "105 BANDEJA", "106 PAQUETE", 
  "107 CAJETILL", "108 CARTON", "109 BOLA", "110 ENVASE", "111 FRASCO", 
  "112 MARQUETA", "113 PEDAZO", "114 PORCIîN", "115 RODAJA", "116 TAPA/CUA", 
  "117 TALLO", "118 RAJA", "119 RAMA", "120 TAZA", "121 TAMBO", "122 ROLLO", 
  "123 MAZO", "124 FARDO", "125 QUETZAL", "128 HUACAL", "129 PACHA"
)
```

---

# Phase C: Quality Checks and Export

## C1. Data Summary Tables

```{r}
#| label: tbl-c1-data-summary

# === ENCOVI 2023 Summary ===
tbl_encovi_summary <- tibble(
  Dataset = c("Food consumption", "Households", "Individuals"),
  Observations = c(
    nrow(encovi_alimentos),
    nrow(encovi_hogar),
    nrow(encovi_individuo)
  ),
  Variables = c(
    ncol(encovi_alimentos),
    ncol(encovi_hogar),
    ncol(encovi_individuo)
  )
)

tbl_encovi_summary |>
  gt() |>
  tab_header(
    title = "ENCOVI 2023 Survey Data",
    subtitle = "Loaded datasets overview"
  ) |>
  fmt_number(
    columns = c(Observations, Variables),
    decimals = 0,
    use_seps = TRUE
  ) |>
  cols_label(
    Dataset = "Dataset",
    Observations = "N",
    Variables = "Variables"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  )

# === INCAP 2014 Summary ===
tbl_incap_summary <- tibble(
  Component = c(
    "Integrated nutrient database",
    "Energy (kcal)",
    "Protein (g)",
    "Iron (mg)",
    "Zinc (mg)",
    "Lysine (g)",
    "Methionine (g)",
    "Cysteine (g)",
    "Threonine (g)",
    "Tryptophan (g)",
    "PDCAAS scores"
  ),
  Status = c(
    paste(nrow(df_incap), "foods"),
    ifelse("energ_kcal" %in% names(df_incap), "✓", "✗"),
    ifelse("proteina_g" %in% names(df_incap), "✓", "✗"),
    ifelse("hierro_mg" %in% names(df_incap), "✓", "✗"),
    ifelse("zinc_mg" %in% names(df_incap), "✓", "✗"),
    ifelse("lys_g" %in% names(df_incap), "✓", "✗"),
    ifelse("met_g" %in% names(df_incap), "✓", "✗"),
    ifelse("cys_g" %in% names(df_incap), "✓", "✗"),
    ifelse("thr_g" %in% names(df_incap), "✓", "✗"),
    ifelse("trp_g" %in% names(df_incap), "✓", "✗"),
    ifelse("pdcaas" %in% names(df_incap), "✓", "✗")
  )
)

tbl_incap_summary |>
  gt() |>
  tab_header(
    title = "INCAP 2014 Food Composition Database",
    subtitle = "Integrated composition tables with amino acids and PDCAAS"
  ) |>
  cols_label(
    Component = "Component",
    Status = "Available"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  )

# === External Data Summary ===
tbl_external_summary <- tibble(
  Dataset = c(
    "Food equivalences",
    "Unique foods",
    "Measurement units",
    "PDCAAS values"
  ),
  Count = c(
    nrow(equivalencias),
    length(nombres_alimentos),
    length(unidades_medida),
    sum(!is.na(incap_composition$pdcaas))
  ),
  Description = c(
    "Expert-provided unit conversions",
    "Foods tracked in survey",
    "Survey measurement codes",
    "Protein quality scores"
  )
)

tbl_external_summary |>
  gt() |>
  tab_header(
    title = "External Expert Data",
    subtitle = "María José Soto (FINUT) contributions"
  ) |>
  fmt_number(
    columns = Count,
    decimals = 0,
    use_seps = TRUE
  ) |>
  cols_label(
    Dataset = "Dataset",
    Count = "N",
    Description = "Description"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  )

# === Processed Data Summary ===
tbl_processed_summary <- tibble(
  Dataset = c(
    "Household socioeconomic",
    "Individuals (demographic)"
  ),
  Count = c(
    nrow(df_hogar),
    nrow(df_individuos)
  ),
  Percentage = c(
    NA_real_,
    NA_real_
  )
)

tbl_processed_summary |>
  gt() |>
  tab_header(
    title = "Processed Base Data",
    subtitle = "Cleaned and prepared datasets for nutrient intake calculations"
  ) |>
  fmt_number(
    columns = Count,
    decimals = 0,
    use_seps = TRUE
  ) |>
  sub_missing(
    columns = Percentage,
    missing_text = "—"
  ) |>
  cols_label(
    Dataset = "Dataset",
    Count = "N",
    Percentage = "%"
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 11,
    table.width = pct(100)
  )

# Save data for web resources
export(
  list(
    encovi_summary = tbl_encovi_summary,
    incap_summary = tbl_incap_summary,
    external_summary = tbl_external_summary,
    processed_summary = tbl_processed_summary
  ),
  here("01_data", "02_processed",
       "web_resources", "01_01_tbl_c1_data_summary.rds")
)

# Cleanup auxiliary objects
rm(tbl_encovi_summary, tbl_incap_summary, tbl_external_summary, tbl_processed_summary)
```

## C2. Save Processed Datasets

```{r}
#| label: c2-save-outputs

# Save integrated INCAP food composition data
write_parquet(
  df_incap,
  here("01_data", "02_processed",  
       "transfer", "01_01_incap_integrated.parquet")
)

# Save cleaned household socioeconomic data
write_parquet(
  df_hogar,
  here("01_data", "02_processed",  
       "transfer", "01_01_household_socioeconomic.parquet")
)

# Save cleaned individual demographic data
write_parquet(
  df_individuos,
  here("01_data", "02_processed",  
       "transfer", "01_01_individuals_demographic.parquet")
)

# Save ENCOVI food consumption data (unprocessed, for next script)
write_parquet(
  encovi_alimentos,
  here("01_data", "02_processed",  
       "transfer", "01_01_encovi_alimentos.parquet")
)

# Save food equivalence table
write_parquet(
  equivalencias,
  here("01_data", "02_processed",  
       "transfer", "01_01_food_equivalences.parquet")
)

# Save reference lists as RDS
export(
  list(
    nombres_alimentos = nombres_alimentos,
    unidades_medida = unidades_medida
  ),
  here("01_data", "02_processed", 
       "transfer", "01_01_reference_lists.rds")
)
```

# Notes

## Methodology

### Data Integration

- **Food composition source**: The integrated composition table is built from `incap_finut_master.xlsx` (Script 00_03). It combines INCAP food composition values (energy, protein, iron, zinc, edible-portion factor) with the FINUT protein-quality layer (amino acids, PDCAAS, NPU) and Azahar Data Insights derived values for foods without literature scores. Per-value provenance is documented in the workbook's `source_notes` sheet and in `01_data/03_external/README.md`.
- **Nutrient values normalized to per-gram basis**: Supports calculations with varying portion sizes.
- **PDCAAS missing values**: Replaced with 0.1 (conservative estimate for incomplete protein sources).

### Household Variables

Socioeconomic variables are recoded to descriptive factor labels; the original survey codes remain as the underlying factor levels.

### Food Equivalence Table

The `equivalencias` table provides unit-to-grams conversions and plant-based classifications compiled by María José Soto (FINUT) through on-site market research in Guatemala:
- Local Guatemalan unit conversions (bags, packages, trays)
- Plant-based food classifications for phytate calculations
- Market-based measurements reflecting actual retail packaging

See `01_data/03_external/README.md` for provenance and lineage details.



